### 2021-01-22 SpatialDesignVModel.Rmd
initial outline commit

---
title: 'Spatial Analysis: Design-Based vs Model-Based'
author: "Michael Dumelle and Matthew Higham"
output: 
  html_document:
    theme: flatly
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

Insert Overview

# Outline

Here is an outline

## Introduction

Few sentences about the goal of the introduction

## Literature Review

Few sentences about the goal of the literature review

## Numerical Analysis

Few sentences about the goal of the data analysis

### Simulation Based

Few sentences about the goal of the simulations

### Data Based

Few sentences about the goal of the data analysis

## Discussion

Few sentences about the most important takeaways

### 2021-01-25 draft.Rnw
initial tex commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document Start
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% this is where we read in all of the preamble material (packages, etc. - stuff that comes before the document begins)
\input{preamble.tex}

% starting the document
\begin{document}

% printing the title (which is in the preamble)
\maketitle

% adding line numbers
\linenumbers

% page break for abstract
\clearpage

% inserting abstract
% \input{abstract.tex}

% page break for start of manuscript
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\centering {Overview}}

\begin{itemize}
  \item Design-Based Overview
    \begin{itemize}
      \item A classic text overviewing design-based approaches \citep{Sarndal2003model}
      \item A classic text overviewing design-based approaches \citep{Lohr2009sampling}
    \end{itemize}
  \item Model-Based Overview
    \begin{itemize}
      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Cressie2015statistics}
      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Schabenberger2017statistical}
    \end{itemize}
  \item Model-based and Design-based Comparisons
      \begin{itemize}
        \item A comparison of design-based and model-based approaches \citep{Hansen1983evaluation}
        \item A comparison of design-based and model-based approaches \citep{Brus1997random}
        \item A comparison of design-based and model-based approaches for spatial data \citep{VerHoef2002sampling}
        \item A comparison of design-based and model-based approaches for spatial data \citep{Cooper2006sampling}
        \item A comparison of design-based and model-based approaches\citep{sterba2009alternative}
        \item A comparison of design-based and model-based approaches for spatial data \citep{Brus2020statistical}
        \item A Bayesian approach to design-based approaches for spatial data \citep{Chan2020bayesian}
      \end{itemize}
  \item Spatially Balanced Design and Analysis
      \begin{itemize}
        \item GRTS sampling \citep{StevensOlsen2004GRTS}
        \item Local neighborhood variance estimation \citep{StevensOlsen2003VarianceEstimation}
      \end{itemize}
  \item Block Kriging 
      \begin{itemize}
        \item Finite Population Block Kriging (with an overview of Block Kriging) \citep{VerHoef2008spatial}
      \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Outline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\centering OUTLINE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript Start
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\centering INTRODUCTION}

%----------------------
% Background
%----------------------
\section{\centering BACKGROUND}

%----------------------
% Numerical Analysis
%----------------------
\section{\centering NUMERICAL ANALYSIS}

\subsection{Simulation-Based}

\subsection{Data-Based}

\subsection{Software}

%----------------------
% Discussion
%----------------------
\section{\centering DISCUSSION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document End
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% page break for bibliography
\clearpage

% bibliography style
\bibliographystyle{apalike}

% bibliography
\bibliography{draft}

% ending the document
\end{document}

### 2021-01-26 draft.Rnw
minor updates

@@
 
+\subsection*{Why This Paper?}
+
+We believe the distinction between these approaches is often misunderstood and there are several ways we could enchance the literature surrounding this topic:
+
+\begin{enumerate}
+  \item There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
+From what we have seen, these comparisons are between spatial model-based approaches and independent random sample designs.
+While important to study this behavior, this comparison is no longer fair nor modern.
+Spatially balanced sampling has exploded in popularity throughout the last decade, and the design-based vs model-based literature needs to reflect this trend.
+  \item A ``fair'' comparison 
+  \item Misspecification
+  \item Finite AND infinite populations
+  \item Centered around the pragmatic questions practitioners will be faced with
+  \item Provide reliable software
+
+\end{enumerate}
+
+\subsection*{Initial Literature}
+
 \begin{itemize}
-  \item Design-Based Overview
-    \begin{itemize}
-      \item A classic text overviewing design-based approaches \citep{Sarndal2003model}
-      \item A classic text overviewing design-based approaches \citep{Lohr2009sampling}
+  \item Design-Based Overview \citep{Sarndal2003model, Lohr2009sampling}
+  \item Model-Based Overview \citep{Cressie2015statistics, Schabenberger2017statistical}
+  \item Design-Based and Model-Based Comparisons \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian}
+  \item Spatially Balanced Design and Analysis \citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS}
+  \item Finite Population Block Kriging \citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting}
 \end{itemize}
-  \item Model-Based Overview
-    \begin{itemize}
-      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Cressie2015statistics}
-      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Schabenberger2017statistical}
-    \end{itemize}
-  \item Model-based and Design-based Comparisons
-      \begin{itemize}
-        \item A comparison of design-based and model-based approaches \citep{Hansen1983evaluation}
-        \item A comparison of design-based and model-based approaches \citep{Brus1997random}
-        \item A comparison of design-based and model-based approaches for spatial data \citep{VerHoef2002sampling}
-        \item A comparison of design-based and model-based approaches for spatial data \citep{Cooper2006sampling}
-        \item A comparison of design-based and model-based approaches\citep{sterba2009alternative}
-        \item A comparison of design-based and model-based approaches for spatial data \citep{Brus2020statistical}
-        \item A Bayesian approach to design-based approaches for spatial data \citep{Chan2020bayesian}
-      \end{itemize}
-  \item Spatially Balanced Design and Analysis
-      \begin{itemize}
-        \item GRTS sampling \citep{StevensOlsen2004GRTS}
-        \item Local neighborhood variance estimation \citep{StevensOlsen2003VarianceEstimation}
-      \end{itemize}
-  \item Block Kriging 
-      \begin{itemize}
-        \item Finite Population Block Kriging (with an overview of Block Kriging) \citep{VerHoef2008spatial}
-      \end{itemize}
-\end{itemize}
+
+\subsection*{Potential Journals}
 

### 2021-01-29 draft.Rnw
added simulation outline

@@
 
+__MH Thoughts:__ I'd like to keep this section manageable. Perhaps we start with the following examples \begin{itemize}
+  \item simulate via correct model (model outperforms sampling)
+   \item simulate via slightly misspecified model (model still outperforms sampling)
+  \item simulate via very misspecified model (sampling outperforms model)
+  \item simulate via extremely misspecified model (e.g. counts with lots of zeroes and a lot of overdispersion) (neither does well)
+\end{itemize}
+
 \subsection{Data-Based}

### 2021-01-29 draft.Rnw
minor updates

@@
 \begin{enumerate}
-  \item There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
+  \item Spatial Design-Based vs Spatial Model-Based: There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
 From what we have seen, these comparisons are between spatial model-based approaches and independent random sample designs.
@@
 Spatially balanced sampling has exploded in popularity throughout the last decade, and the design-based vs model-based literature needs to reflect this trend.
-  \item A ``fair'' comparison 
-  \item Misspecification
-  \item Finite AND infinite populations
-  \item Centered around the pragmatic questions practitioners will be faced with
+  \item A ``fair'' comparison: We feel that literature in this area has considered scenarios that are more well-suited for the model-based scenario.
+For example, if you simulate a Gaussian erorr with an exponential covariance and then then compare design-based estimates to model-based estimates assuming an exponential covariance, of course the model-based approach will outperform design-based approaches.
+And these are the types of comparisons in the literature, which find model-based approaches generally yield more precise variance estiamtes.
+A challenge lies in creating a comparison scenario that is reasonable and intuitive.
+One this to consider would be exploring the comparison after estimating a misspecified covariance function using model-based approaches.
+  \item Finite AND infinite populations: Literature in the area focuses specifically on finite populations or infinite populations; we want to discuss both in detail.
+  \item Pragmatic Focus: We see papers in this area tend to be fairly technical.
+We want the focus to be less on details, more on discussing the pragmatic questions practitioners will be faced with.
+For example, a thorough discussion of benefits and drawbacks of each method written for practicioners is warranted.
   \item Provide reliable software
@@
 \subsection*{Potential Journals}
+\begin{itemize}
+  \item Ecological Applications
+  \item Methods in Ecology and Evolution
+  \item Journal of Applied Ecology
+  \item Environmetrics
+  \item Environmental and Ecological Statistics
+\end{itemize}
 
@@
 
-__MH Thoughts:__ I'd like to keep this section manageable. Perhaps we start with the following examples \begin{itemize}
+We would like to keep this section manageable. Perhaps we start with the following examples 
+\begin{itemize}
   \item simulate via correct model (model outperforms sampling)

### 2021-01-29 draft.tex
minor updates

@@
 \subsection{Simulation-Based}
+
+We would like to keep this section manageable. Perhaps we start with the following examples 
+\begin{itemize}
+  \item simulate via correct model (model outperforms sampling)
+  \item simulate via slightly misspecified model (model still outperforms sampling)
+  \item simulate via very misspecified model (sampling outperforms model)
+  \item simulate via extremely misspecified model (e.g. counts with lots of zeroes and a lot of overdispersion) (neither does well)
+\end{itemize}
 

### 2021-01-29 draft.Rnw
minor updates

@@
 For example, a thorough discussion of benefits and drawbacks of each method written for practicioners is warranted.
-  \item Provide reliable software
+  \item Provide reliable software and code: Current literature presents the ideas but do not provide adequate access to user-friendly software amd cpde.
+We would clearly explain how to use `spsurvey` to design and analyze spatial probability samples and `sptotal` for FBPK.
+We would make a GitHub repository for the paper walking through code used throughout the manuscript.
 

### 2021-02-02 draft.Rnw
minor updates

--- 2021-01-29 draft.Rnw
@@
 
+This paper would be a comparison of design-based and model-based frameworks for spatial data with an emphasis on pragmatic considerations when choosing among the frameworks and using them to solve applied problems using open-source software.
+
+This paper could be organized as follows.
+In Section \ref{sec:introduction}, we outline the design-based and model-based frameworks and then describe each framework in the context of spatial data.
+In Section \ref{sec:background}, we thoroughly describe and contextualize the assumptions, goals, benefits, and drawbacks of each approach.
+In Section \ref{sec:numerical}, we compare the two methods by estimating means and variances of spatial data in several simulation scenarios as well as real data.
+We also provide reliable software through the use of \texttt{spsurvey} and \texttt{sptotal}.
+In Section \ref{sec:discussion}, we end with a discussion wrapping up loose ends and providing directions for future research.
+
 \subsection*{Why This Paper?}
 
-We believe the distinction between these approaches is often misunderstood and there are several ways we could enchance the literature surrounding this topic:
+We believe the distinction between these approaches is often misunderstood, and there are several ways we could enhance the literature surrounding this topic:
 
 \begin{enumerate}
-  \item Spatial Design-Based vs Spatial Model-Based: There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
+  \item Spatial design-based vs spatial model-based: There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
 From what we have seen, these comparisons are between spatial model-based approaches and independent random sample designs.
@@
   \item A ``fair'' comparison: We feel that literature in this area has considered scenarios that are more well-suited for the model-based scenario.
-For example, if you simulate a Gaussian erorr with an exponential covariance and then then compare design-based estimates to model-based estimates assuming an exponential covariance, of course the model-based approach will outperform design-based approaches.
-And these are the types of comparisons in the literature, which find model-based approaches generally yield more precise variance estiamtes.
+For example, if you simulate a Gaussian error with an exponential covariance and then then compare design-based estimates to model-based estimates assuming an exponential covariance, of course the model-based approach will outperform design-based approaches.
+And these are the types of comparisons in the literature, which find model-based approaches generally yield more precise variance estimates.
 A challenge lies in creating a comparison scenario that is reasonable and intuitive.
@@
 We want the focus to be less on details, more on discussing the pragmatic questions practitioners will be faced with.
-For example, a thorough discussion of benefits and drawbacks of each method written for practicioners is warranted.
-  \item Provide reliable software
+For example, a thorough discussion of benefits and drawbacks of each method written for practitioners is warranted.
+  \item Provide reliable software and code: Current literature presents the ideas but do not provide adequate access to user-friendly software and code.
+We would clearly explain how to use \texttt{spsurvey} to design and analyze spatial probability samples and \texttt{sptotal} for FBPK.
+We would make a GitHub repository / package for the paper walking through code used and containing all manuscript-related documents.
 
@@
   \item Environmental and Ecological Statistics
+  \item Other
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\centering INTRODUCTION}
+\section{\centering INTRODUCTION}\label{sec:introduction}
+
+A brief introduction introducing the design-based and model-based frameworks and outlining the rest of the paper.
 
@@
 %----------------------
-\section{\centering BACKGROUND}
+\section{\centering BACKGROUND}\label{sec:background}
+
+A thorough dive into the assumptions, goals, benefits, and drawbacks of the approaches as they relate to \emph{spatial} data.
+A special focus will be given to pragmatic considerations; we want this paper to be especially for practitioners.
 
@@
 %----------------------
-\section{\centering NUMERICAL ANALYSIS}
+\section{\centering NUMERICAL ANALYSIS}\label{sec:numerical}
+
+We plan to focus on finite populations but also want to accommodate infinite populations
 
@@
 
-We would like to keep this section manageable. Perhaps we start with the following examples 
+We would like to keep this section manageable but useful and realistic. Perhaps we start with the following examples: 
 \begin{itemize}
@@
 
+\begin{itemize}
+  \item Use real data set(s) for which we know population values 
+\end{itemize}
+
 \subsection{Software}
+
+\texttt{spsurvey} ($>$ 5.0; we are wrapping up a major update now to 5.0) and \texttt{sptotal} ($>$ 0.1.0)
 
@@
 %----------------------
-\section{\centering DISCUSSION}
+\section{\centering DISCUSSION}\label{sec:discussion}
+
+We plan to wrap up loose ends, provide directions for future research, and offer any closing thoughts.
 

### 2021-02-04 draft.Rnw
Made minor edits and added the possibility of having an accessible simulation function available to practitioners.

@@
 
-This paper would be a comparison of design-based and model-based frameworks for spatial data with an emphasis on pragmatic considerations when choosing among the frameworks and using them to solve applied problems using open-source software.
+This paper would be a comparison of design-based and model-based frameworks for spatial data with an emphasis on pragmatic applications, like how to choose among the frameworks and how to use them to solve applied problems using open-source software.
 
 This paper could be organized as follows.
-In Section \ref{sec:introduction}, we outline the design-based and model-based frameworks and then describe each framework in the context of spatial data.
+In Section \ref{sec:introduction}, we outline the design-based and model-based frameworks and then we describe each framework in the context of spatial data.
 In Section \ref{sec:background}, we thoroughly describe and contextualize the assumptions, goals, benefits, and drawbacks of each approach.
-In Section \ref{sec:numerical}, we compare the two methods by estimating means and variances of spatial data in several simulation scenarios as well as real data.
+In Section \ref{sec:numerical}, we compare the two methods by comparing means and variances of spatial data in several simulation scenarios as well as in real data.
 We also provide reliable software through the use of \texttt{spsurvey} and \texttt{sptotal}.
-In Section \ref{sec:discussion}, we end with a discussion wrapping up loose ends and providing directions for future research.
+In Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
 
@@
   \item Spatial design-based vs spatial model-based: There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
-From what we have seen, these comparisons are between spatial model-based approaches and independent random sample designs.
+These comparisons are typically between spatial model-based approaches and independent or stratified random sample designs.
 While important to study this behavior, this comparison is no longer fair nor modern.
@@
 For example, if you simulate a Gaussian error with an exponential covariance and then then compare design-based estimates to model-based estimates assuming an exponential covariance, of course the model-based approach will outperform design-based approaches.
-And these are the types of comparisons in the literature, which find model-based approaches generally yield more precise variance estimates.
+These types of comparisons in the literature find model-based approaches generally yield more precise variance estimates.
 A challenge lies in creating a comparison scenario that is reasonable and intuitive.
-One this to consider would be exploring the comparison after estimating a misspecified covariance function using model-based approaches.
-  \item Finite AND infinite populations: Literature in the area focuses specifically on finite populations or infinite populations; we want to discuss both in detail.
+One thing to consider would be exploring the comparison after estimating a misspecified covariance function using model-based approaches.
+  \item Finite AND infinite populations: Literature in the area focuses specifically on finite populations or infinite populations; we want to discuss both, with a focus on finite populations.
   \item Pragmatic Focus: We see papers in this area tend to be fairly technical.
-We want the focus to be less on details, more on discussing the pragmatic questions practitioners will be faced with.
+We want the focus to be less on mathematical details and more on discussing the pragmatic questions practitioners will be faced with.
 For example, a thorough discussion of benefits and drawbacks of each method written for practitioners is warranted.
-  \item Provide reliable software and code: Current literature presents the ideas but do not provide adequate access to user-friendly software and code.
+  \item Provide reliable software and code: Much of the current literature presents the ideas but does not provide adequate access to user-friendly software and code.
 We would clearly explain how to use \texttt{spsurvey} to design and analyze spatial probability samples and \texttt{sptotal} for FBPK.
 We would make a GitHub repository / package for the paper walking through code used and containing all manuscript-related documents.
+Because our code for the simulation study would be easily accessible, practitioners would be able to conduct a small simulation study with parameters that reflect their particular context to compare the two approaches.
 
@@
 A thorough dive into the assumptions, goals, benefits, and drawbacks of the approaches as they relate to \emph{spatial} data.
-A special focus will be given to pragmatic considerations; we want this paper to be especially for practitioners.
+A special focus will be given to pragmatic considerations; we want this paper to be especially useful for practitioners.
 
@@
 
-We plan to focus on finite populations but also want to accommodate infinite populations
+We plan to focus on finite populations but also want to accommodate infinite populations.
 
@@
 \end{itemize}
+
+We may consider providing an `R` script that would let practitioners adjust parameters in simulations freely, requiring only that the practitioner know basic `R` syntax.
 

### 2021-04-15 draft.Rnw
test MRH personal access token.

@@
 \begin{enumerate}
-  \item Spatial design-based vs spatial model-based: There are no comparisons in the literature between spatial model-based approaches and *spatially balanced* design approaches.
+  \item Spatial design-based vs spatial model-based: There are no comparisons in the literature between spatial model-based approaches and \emph{spatially balanced} design approaches.
 These comparisons are typically between spatial model-based approaches and independent or stratified random sample designs.
@@
 \end{itemize}
+
+Test MH personal access token.
 
@@
 
-\texttt{spsurvey} ($>$ 5.0; we are wrapping up a major update now to 5.0) and \texttt{sptotal} ($>$ 0.1.0)
+\texttt{spsurvey} ($>=$ 5.0; we are wrapping up a major update now to 5.0) and \texttt{sptotal} ($>=$ 0.1.0)
 

### 2021-05-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
convert files to package format.

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short Paper},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Short Paper}
    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
   \ead{alice@example.com} 
    \author[Another University]{Bob Security}
   \ead{bob@example.com} 
    \author[Another University]{Cat Memes\corref{2}}
   \ead{cat@example.com} 
    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
   \ead{derek@example.com} 
      \address[Some Institute of Technology]{Department, Street, City,
State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\hypertarget{the-elsevier-article-class}{%
\section{The Elsevier article class}\label{the-elsevier-article-class}}

\hypertarget{installation}{%
\paragraph{Installation}\label{installation}}

If the document class \emph{elsarticle} is not available on your
computer, you can download and install the system package
\emph{texlive-publishers} (Linux) or install the LaTeX package
\emph{elsarticle} using the package manager of your TeX installation,
which is typically TeX Live or MikTeX.

\hypertarget{usage}{%
\paragraph{Usage}\label{usage}}

Once the package is properly installed, you can use the document class
\emph{elsarticle} to create a manuscript. Please make sure that your
manuscript follows the guidelines in the Guide for Authors of the
relevant journal. It is not necessary to typeset your manuscript in
exactly the same way as an article, unless you are submitting to a
camera-ready copy (CRC) journal.

\hypertarget{functionality}{%
\paragraph{Functionality}\label{functionality}}

The Elsevier article class is based on the standard article class and
supports almost all of the functionality of that class. In addition, it
features commands and options to format the

\begin{itemize}
\item
  document style
\item
  baselineskip
\item
  front matter
\item
  keywords and MSC codes
\item
  theorems, definitions and proofs
\item
  lables of enumerations
\item
  citation style and labeling.
\end{itemize}

\hypertarget{front-matter}{%
\section{Front matter}\label{front-matter}}

The author names and affiliations could be formatted in two ways:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  Group the authors per affiliation.
\item
  Use footnotes to indicate the affiliations.
\end{enumerate}

See the front matter of this document for examples. You are recommended
to conform your choice to the journal you are submitting to.

\hypertarget{bibliography-styles}{%
\section{Bibliography styles}\label{bibliography-styles}}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX~to generate your bibliography and include DOIs
whenever available.

Here are two sample references: (Dirac, 1953; Feynman and Vernon Jr.;
1963).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Dirac1953888}{}%
Dirac, P.A.M., 1953. The lorentz transformation and absolute time.
Physica 19, 888--896.
doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}

\leavevmode\hypertarget{ref-Feynman1963118}{}%
Feynman, R.P., Vernon Jr., F.L., 1963. The theory of a general quantum
system interacting with a linear dissipative system. Annals of Physics
24, 118--173.
doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}

\end{CSLReferences}

\end{document}

### 2021-05-07 outline/Introduction_MH.Rmd
convert files to package format.

---
title: "Introduction"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{itemize}
  \item Design-Based Overview \citep{Sarndal2003model, Lohr2009sampling}
  \item Model-Based Overview \citep{Cressie2015statistics, Schabenberger2017statistical}
  \item Design-Based and Model-Based Comparisons \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian}
  \item Spatially Balanced Design and Analysis \citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS}
  \item Finite Population Block Kriging \citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting}
\end{itemize}

References from Brus2020statistical

* de Gruijter, J. J., & ter Braak, C. J. F. (1990). Model-free estimation from spatial samples: A reappraisal of classical sampling theory. Mathematical Geology, 22, 407–415
    * contains info on why classical sampling works with spatially correlated data.

* Brus, D. J., & de Gruijter, J. J. (1997). Random sampling or
geostatistical modelling? Choosing between design-based and
model-based sampling strategies for soil (with discussion). Geoderma, 80, 1–59
    * more comparison between the two appraoches

* I stopped at 2.3

Design-based inference uses characteristics of the sampling design to obtain relevant confidence intervals for any parameters of interest. Typically, there are few assumptions involved because intervals are derived using the sampling design itself. Data is simply "there": we do not assume that it has been generated through some abstract model. We know that if we sampled all sites, we would have an exact value for the true mean, without any uncertainty.

On the other hand, model-based inference imposes additional assumptions on the data with the promise of providing more precise estimates if the additional assumptions hold. Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is typically _prediction_ of an unknown quantity. This is a fundamental philosophical difference between sampling-based and model-based approaches. Instead of _estimating_ an fixed unknown mean, we are _predicting_ the value of the mean, a random variable. We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty. But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not really care much about what it is.

Figure 1a. Data is fixed. In a finite population example, show a 3d surface that can be generated by anything. If we repeatedly sample the surface, then 95% of all 95% CIs will contain the true mean, which never changes.

```{r}
library(plotly)
```

Figure 1b. Spatial process is fixed. In a finite population example, show 10 3d surfaces that are generated from some model. If we repeatedly generate the surface and conduct our sample, then 95% of all 95% PIs will contain the realized means. The mean changes from surface to surface and it's not necessarily the case that 95% of all 95% PIs will contain the true, underlying mean.

### 2021-05-20 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
intro to model based and background of FPBK.

---
title: Short Paper
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
preamble: >
  \usepackage{bm}
  \DeclareMathOperator{\var}{{var}}
  \DeclareMathOperator{\cov}{{cov}}
output: rticles::elsevier_article
---

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

Introduction
==========

## Design-Based Philosphy

Design-Based Overview \citep{Sarndal2003model, Lohr2009sampling}

Design-based inference uses characteristics of the sampling design to to estimate parameters of interest Typically, there are few assumptions involved because intervals are derived using the sampling design itself...........

## Model-Based Philosphy

Model-Based Overview \citep{Cressie2015statistics, Schabenberger2017statistical}

On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimates if the additional assumptions hold. Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often _prediction_ of an unknown quantity. This is a fundamental philosophical difference between sampling-based and model-based approaches. Instead of _estimating_ a fixed unknown mean, we are _predicting_ the value of the mean, a random variable. We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty. But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.

Figure 1a. Data is fixed. In a finite population example, show a 3d surface that can be generated by anything. If we repeatedly sample the surface, then 95% of all 95% CIs will contain the true mean, which never changes.

Figure 1b. Spatial process is fixed. In a finite population example, show 10 3d surfaces that are generated from some model. If we repeatedly generate the surface and obtain a sample, then 95% of all 95% PIs will contain the realized means. The realized mean changes from surface to surface and it's not necessarily the case that 95% of all 95% PIs will contain the true, underlying mean.

## Comparing Design-Based vs. Model-Based

Design-Based and Model-Based Comparisons \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian}

There have been many comparisons between the two paradigms.......

## Spatially Balanced Design and Analysis

\citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS}

## Finite Population Block Kriging

\citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting}

Finite Population Block Kriging (FPBK) is an alternative to samipling-based methods [@VerHoef2008spatial]. FPBK expands the geostatistical kriging framework to the finite population setting. Instead of basing inference off of a specific sampling design, we assume the data were generated by a spatial process with parameters that can be estimated using the framework of a model.

@VerHoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below. For a response variable $\mathbf{z}$ that can be measured on a finite number of $N$ sites, we want to predict some linear function of the response variable, $\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a vector of weights. For example, if we want to predict the total abundance across all sites, then we would use a vector of 1's for the weights.

Typically, however, we only have a sample of the $N$ sites. Denoting quantities that are part of the sampled sites with a subscript \emph{s} and quantities that are part of the unsampled sites with a subscript \emph{u},

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation}
where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled sites, respectively, and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled sites. Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume that $E(\bm{\delta})$ = $\mathbf{0}$.

We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function. Many common choices for this function assume that spatial covariance decreases with increasing Euclidean distance between sites. The primary function used throughout the simulations and applications of this manuscript is the Exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
\mbox{}
\begin{equation}
\cov(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \label{equation:expcov}
\end{equation}
where $h_{i,j}$ is the distance between sites $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$. However, any spatial covariance function could be used in the place of the Exponential, including functions that allow for anisotropy [pg. 80 - 93][@ChilesEtAl1999GeostatisticsModelingSpatial].

With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed. While details of the derivation are in [@VerHoef2008spatial], we note here that the predictor and its variance are both moment-based. Neither require a particular distribution for $\mathbf{z}$.

__Software Implementation__: This probably goes in a different spot (right before simulations?), but putting it here while it's on my mind.

FPBK can be readily performed in `R` with the `sptotal` package [@sptotal2020]. We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).

__Extra Stuff__

References from Brus2020statistical

* de Gruijter, J. J., & ter Braak, C. J. F. (1990). Model-free estimation from spatial samples: A reappraisal of classical sampling theory. Mathematical Geology, 22, 407–415
    * contains info on why classical sampling works with spatially correlated data.

* Brus, D. J., & de Gruijter, J. J. (1997). Random sampling or
geostatistical modelling? Choosing between design-based and
model-based sampling strategies for soil (with discussion). Geoderma, 80, 1–59
    * more comparison between the two appraoches

* I stopped at 2.3

#### Installation

If the document class *elsarticle* is not available on your computer,
you can download and install the system package *texlive-publishers*
(Linux) or install the LaTeX package *elsarticle* using the package
manager of your TeX installation, which is typically TeX Live or MikTeX.

#### Usage

Once the package is properly installed, you can use the document class
*elsarticle* to create a manuscript. Please make sure that your
manuscript follows the guidelines in the Guide for Authors of the
relevant journal. It is not necessary to typeset your manuscript in
exactly the same way as an article, unless you are submitting to a
camera-ready copy (CRC) journal.

#### Functionality

The Elsevier article class is based on the standard article class and
supports almost all of the functionality of that class. In addition, it
features commands and options to format the

-   document style

-   baselineskip

-   front matter

-   keywords and MSC codes

-   theorems, definitions and proofs

-   lables of enumerations

-   citation style and labeling.

Front matter
============

The author names and affiliations could be formatted in two ways:

(1) Group the authors per affiliation.

(2) Use footnotes to indicate the affiliations.

See the front matter of this document for examples. You are recommended
to conform your choice to the journal you are submitting to.

Bibliography styles
===================

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX to generate your bibliography and include DOIs
whenever available.

Here are two sample references: @Feynman1963118 [@Dirac1953888].

References {#references .unnumbered}
==========

### 2021-05-20 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
intro to model based and background of FPBK.

@@
 
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{5}
 % Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
 
@@
 
+\usepackage{bm} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}}
+
 \begin{document}
@@
 
-\hypertarget{the-elsevier-article-class}{%
-\section{The Elsevier article class}\label{the-elsevier-article-class}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+\hypertarget{design-based-philosphy}{%
+\subsection{Design-Based Philosphy}\label{design-based-philosphy}}
+
+\hypertarget{model-based-philosphy}{%
+\subsection{Model-Based Philosphy}\label{model-based-philosphy}}
+
+On the other hand, model-based inference imposes additional assumptions
+on the data with a potential to provide more precise estimates if the
+additional assumptions hold. Instead of estimating true but unknown
+parameters, the goal of model-based inference in the spatial context is
+often \emph{prediction} of an unknown quantity. This is a fundamental
+philosophical difference between sampling-based and model-based
+approaches. Instead of \emph{estimating} an fixed unknown mean, we are
+\emph{predicting} the value of the mean, a random variable. We know that
+if we sampled all sites, we would have an exact prediction for the mean
+of our one realized spatial surface, without any uncertainty. But, the
+true mean of the spatial process that generated our realized data is
+still not known, and, in the prediction context, we typically do not
+care much about what value the mean of the underlying process takes.
+
+Figure 1a. Data is fixed. In a finite population example, show a 3d
+surface that can be generated by anything. If we repeatedly sample the
+surface, then 95\% of all 95\% CIs will contain the true mean, which
+never changes.
+
+Figure 1b. Spatial process is fixed. In a finite population example,
+show 10 3d surfaces that are generated from some model. If we repeatedly
+generate the surface and obtain a sample, then 95\% of all 95\% PIs will
+contain the realized means. The realized mean changes from surface to
+surface and it's not necessarily the case that 95\% of all 95\% PIs will
+contain the true, underlying mean.
+
+\hypertarget{comparing-design-based-vs.-model-based}{%
+\subsection{Comparing Design-Based
+vs.~Model-Based}\label{comparing-design-based-vs.-model-based}}
+
+There have been many comparisons between the two paradigms\ldots\ldots.
+
+\hypertarget{spatially-balanced-design-and-analysis}{%
+\subsection{Spatially Balanced Design and
+Analysis}\label{spatially-balanced-design-and-analysis}}
+
+\hypertarget{finite-population-block-kriging}{%
+\subsection{Finite Population Block
+Kriging}\label{finite-population-block-kriging}}
+
+Finite Population Block Kriging (FPBK) is an alternative to
+samipling-based methods (Ver Hoef, 2008). FPBK expands the
+geostatistical kriging framework to the finite population setting.
+Instead of relying on a specific sampling design, we assume the data
+were produced by a spatial stochastic process with spatial parameters
+that can be estimated.
+
+Ver Hoef (2008) gives details on the theory of FPBK, but some of the
+basic principles are summarized below. For a response variable
+\(\mathbf{z}\) that can be measured on a finite number of \(N\) sites,
+our goal is to predict some linear function of all of the sample units,
+\(\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
+\(\mathbf{b}\) is a vector of weights. One common vector of weights is a
+vector of 1's to predict the total abundance in the region.
+
+Typically, however, we only have a sample of the \(N\) sites. Denoting
+quantities that are part of the sampled sites with a subscript \emph{s}
+and quantities that are part of the unsampled sites with a subscript
+\emph{u},
+
+\begin{equation}
+\begin{pmatrix} \label{equation:Zmarginal}
+    \mathbf{z}_s      \\
+    \mathbf{z}_u
+\end{pmatrix}
+=
+\begin{pmatrix}
+  \mathbf{X}_s    \\
+  \mathbf{X}_u
+\end{pmatrix}
+\bm{\beta} +
+\begin{pmatrix}
+\bm{\delta}_s    \\
+\bm{\delta}_u
+\end{pmatrix},
+\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
+design matrices for the sampled and unsampled sites, respectively, and
+\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are zero-mean random errors for
+the sampled and unsampled sites. Denoting
+\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), then we
+assume that \(E(\bm{\delta})\) = \(\mathbf{0}\).
+
+We also typically assume that there is spatial correlation in
+\(\bm{\delta}\), which can be modeled using a covariance function. Many
+common choices for this function assume that spatial covariance
+decreases with increasing Euclidean distance between sites. The primary
+function used throughout the simulations and applications of this
+manuscript is the Exponential covariance function: the \(i,j^{th}\)
+entry for \(\mathop{\mathrm{{var}}}(\bm{\delta})\) is \mbox{}
+\begin{equation}
+\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \label{equation:expcov}
+\end{equation} where \(h_{i,j}\) is the distance between sites \(i\) and
+\(j\), and \(\bm{\theta}\) is a vector of spatial covariance parameters
+of the partial sill \(\theta_1\), the range \(\theta_2\), and the nugget
+\(\theta_3\). However, any spatial covariance function could be used in
+the place of the Exponential, including functions that allow for
+anisotropy {[}pg. 80 - 93{]}(Chiles and Delfiner, 1999).
+
+With the above model formulation, the Best Linear Unbiased Predictor
+(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\) and its prediction variance
+can be computed. While details of the derivation are in (Ver Hoef,
+2008), we note here that the predictor and its variance are both
+moment-based. Neither require a particular distribution for
+\(\mathbf{z}\).
+
+\textbf{Software Implementation}: This probably goes in a different spot
+(right before simulations?), but putting it here while it's on my mind.
+
+FPBK can be readily performed in \texttt{R} with the \texttt{sptotal}
+package (Matt et al., 2020). We use \texttt{sptotal} for both the
+simulation analysis and the application, estimating parameters with
+Restricted Maximum Likelihood (REML).
 
@@
 
-Here are two sample references: (Dirac, 1953; Feynman and Vernon Jr.;
-1963).
+Here are two sample references: (\textbf{Feynman1963118?};
+\textbf{Dirac1953888?}).
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Dirac1953888}{}%
-Dirac, P.A.M., 1953. The lorentz transformation and absolute time.
-Physica 19, 888--896.
-doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}
+\leavevmode\hypertarget{ref-ChilesEtAl1999GeostatisticsModelingSpatial}{}%
+Chiles, J.-P., Delfiner, P., 1999. Geostatistics: {Modeling Spatial
+Uncertainty}. {John Wiley \& Sons}, New York.
 
-\leavevmode\hypertarget{ref-Feynman1963118}{}%
-Feynman, R.P., Vernon Jr., F.L., 1963. The theory of a general quantum
-system interacting with a linear dissipative system. Annals of Physics
-24, 118--173.
-doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}
+\leavevmode\hypertarget{ref-sptotal2020}{}%
+Matt, H., Jay, V.H., Bryce, F., 2020. Sptotal: Predicting totals and
+weighted sums from spatial data.
+
+\leavevmode\hypertarget{ref-VerHoef2008spatial}{}%
+Ver Hoef, J.M., 2008. Spatial methods for plot-based sampling of
+wildlife populations. Environmental and Ecological Statistics 15, 3--13.
 

### 2021-05-20 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updates to "spatially balanced design and analysis"

@@
 ---
-title: Short Paper
+title: A comparison of model-based and design-based approaches for spatial data.
 author:
-  - name: Alice Anonymous
-    email: alice@example.com
-    affiliation: Some Institute of Technology
+  - name: Michael Dumelle
+    email: Dumelle.Michael@epa.gov
+    affiliation: USEPA
     footnote: 1
-  - name: Bob Security
-    email: bob@example.com
-    affiliation: Another University
-  - name: Cat Memes
-    email: cat@example.com
-    affiliation: Another University
-    footnote: 2
-  - name: Derek Zoolander
-    email: derek@example.com
-    affiliation: Some Institute of Technology
-    footnote: 2
+  - name: Matthew Higham
+    email: mhigham@stlaw.edu
+    affiliation: STLAW
+    footnote: 1
+  - name: Lisa Madsen
+    affiliation: OSU
+  - name: Anthony R. Olsen
+    affiliation: USEPA
+  - name: Jay M. Ver Hoef
+    affiliation: NOAA
 address:
-  - code: Some Institute of Technology
-    address: Department, Street, City, State, Zip
-  - code: Another University
-    address: Department, Street, City, State, Zip
+  - code: USEPA
+    address: United States Environmental Protection Agency, 200 SW 35th St, Corvallis, Oregon, 97333
+  - code: STLAW
+    address: Saint Lawrence University Department of Math, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
+  - code: OSU
+    address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
+  - code: NOAA
+    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
 footnote:
@@
     text: "Corresponding Author"
-  - code: 2
-    text: "Equal contribution"
 abstract: |
@@
   \usepackage{bm}
+  \usepackage{bbm}
+  \usepackage{color}
   \DeclareMathOperator{\var}{{var}}
@@
 
-Introduction
-==========
+# Introduction
+
+Please leave comments in your \textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}: \textcolor{blue}{Michael}, \textcolor{magenta}{Matt}, \textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
+
+# Background
 
@@
 
-Design-Based Overview \citep{Sarndal2003model, Lohr2009sampling}
+Design-Based Overview
+
+<!-- \citep{Sarndal2003model, Lohr2009sampling} -->
 
@@
 
-Model-Based Overview \citep{Cressie2015statistics, Schabenberger2017statistical}
+Model-Based Overview
+
+<!-- \citep{Cressie2015statistics, Schabenberger2017statistical} -->
 
@@
 
-Design-Based and Model-Based Comparisons \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian}
+Design-Based and Model-Based Comparisons
+
+<!-- \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian} -->
 
@@
 
-\citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS}
+Spatially balanced sampling algorithms use spatial information to obtain samples spread out in space.
+Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from an sampling algorithm that is not spatially balanced ([@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design]).
+Many spatially balanced sampling algorithms exist, including the Generalized Random Tessellation Stratified [@stevens2004spatially], the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
+Here we focus on the Generalized Random Tessellation Stratified (GRTS) algorithm, which has several attractive properties that we discuss next.
+
+The GRTS algorithm is used to sample from finite and infinite sample frames.
+A finite sample frames contains a finite number of sampling units and is related to a point geometry.
+An infinite sample frame contains an infinite number of sampling units and is related to linear and polygon geometries.
+Examples of point, linear, and polygon resources include lake centroids, stream networks, and wetland areas, respectively.
+In addition to its applicability for finite and infinite sample frames, the GRTS algorithm naturally accommodates stratified designs and designs with unequal selection probabilities.
+The algorithm has also been used to select replacement sites using reverse hierarchical ordering [@stevens2004spatially].
+Replacement sites are used to replace sites in the original sample that cannot be sampled, often as a result of physical difficulty in reaching the site or landowner denial of access to the sites.
+More recently, the GRTS algorithm also accommodates legacy (historical) sites, minimum distance between sites, and nearest neighbor replacement sites.
+The GRTS algorithm is implemented in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
+
+describe grts
+
+discuss spatial balance
+
+describe estimation and local neighborhood variance
+
+<!-- \citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS} -->
 
@@
 
-\citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting}
+<!-- \citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting} -->
 
-Finite Population Block Kriging (FPBK) is an alternative to samipling-based methods [@VerHoef2008spatial].
+Finite Population Block Kriging (FPBK) is an alternative to samipling-based methods [@verhoef2008spatial].
 FPBK expands the geostatistical kriging framework to the finite population setting.
@@
 
-@VerHoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
+@verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
 For a response variable $\mathbf{z}$ that can be measured on a finite number of $N$ sites, we want to predict some linear function of the response variable, $\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a vector of weights.
@@
 \mbox{}
-\begin{equation}
-\cov(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \label{equation:expcov}
-\end{equation}
+\begin{align}\label{equation:expcov}
+\cov(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \\
+\textcolor{blue}{\sigma^2[(1 - v)\exp(-3\mathbf{h}_{i,j}/\phi) + v\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}]}, 
+\end{align}
 where $h_{i,j}$ is the distance between sites $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$.
 However, any spatial covariance function could be used in the place of the Exponential, including functions that allow for anisotropy [pg.
-80 - 93][@ChilesEtAl1999GeostatisticsModelingSpatial].
+80 - 93][@chiles1999geostatistics].
 
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@VerHoef2008spatial], we note here that the predictor and its variance are both moment-based.
+With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ \textcolor{blue}{Did you mean to give the form of the BLUP here?
+$\tau(\mathbf{b}'\mathbf{z})$ is vague} and its prediction variance can be computed.
+While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
 Neither require a particular distribution for $\mathbf{z}$.
 
-__Software Implementation__: This probably goes in a different spot (right before simulations?), but putting it here while it's on my mind.
+# Numerical Study
 
-FPBK can be readily performed in `R` with the `sptotal` package [@sptotal2020].
+## Software
+
+FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
 We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
 
-__Extra Stuff__
+# Discussion
 
-References from Brus2020statistical
+<!-- References from Brus2020statistical, de1990model, Brus1997random -->
 
-* de Gruijter, J. J., & ter Braak, C. J. F. (1990). Model-free estimation from spatial samples: A reappraisal of classical sampling theory. Mathematical Geology, 22, 407–415
-    * contains info on why classical sampling works with spatially correlated data.
-
-* Brus, D. J., & de Gruijter, J. J. (1997). Random sampling or
-geostatistical modelling? Choosing between design-based and
-model-based sampling strategies for soil (with discussion). Geoderma, 80, 1–59
-    * more comparison between the two appraoches
-
-* I stopped at 2.3
-
-#### Installation
-
-If the document class *elsarticle* is not available on your computer,
-you can download and install the system package *texlive-publishers*
-(Linux) or install the LaTeX package *elsarticle* using the package
-manager of your TeX installation, which is typically TeX Live or MikTeX.
-
-#### Usage
-
-Once the package is properly installed, you can use the document class
-*elsarticle* to create a manuscript. Please make sure that your
-manuscript follows the guidelines in the Guide for Authors of the
-relevant journal. It is not necessary to typeset your manuscript in
-exactly the same way as an article, unless you are submitting to a
-camera-ready copy (CRC) journal.
-
-#### Functionality
-
-The Elsevier article class is based on the standard article class and
-supports almost all of the functionality of that class. In addition, it
-features commands and options to format the
-
--   document style
-
--   baselineskip
-
--   front matter
-
--   keywords and MSC codes
-
--   theorems, definitions and proofs
-
--   lables of enumerations
-
--   citation style and labeling.
-
-Front matter
-============
-
-The author names and affiliations could be formatted in two ways:
-
-(1) Group the authors per affiliation.
-
-(2) Use footnotes to indicate the affiliations.
-
-See the front matter of this document for examples. You are recommended
-to conform your choice to the journal you are submitting to.
-
-Bibliography styles
-===================
-
-There are various bibliography styles available. You can select the
-style of your choice in the preamble of this document. These styles are
-Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use BibTeX to generate your bibliography and include DOIs
-whenever available.
-
-Here are two sample references: @Feynman1963118 [@Dirac1953888].
-
-References {#references .unnumbered}
-==========
+# References {#references .unnumbered}

### 2021-05-20 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updates to "spatially balanced design and analysis"

@@
             pdfauthor={},
-            pdftitle={Short Paper},
+            pdftitle={A comparison of model-based and design-based approaches for spatial data.},
             colorlinks=false,
@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\usepackage{bm} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}}
+\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}}
 
@@
 
-  \title{Short Paper}
-    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
-   \ead{alice@example.com} 
-    \author[Another University]{Bob Security}
-   \ead{bob@example.com} 
-    \author[Another University]{Cat Memes\corref{2}}
-   \ead{cat@example.com} 
-    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
-   \ead{derek@example.com} 
-      \address[Some Institute of Technology]{Department, Street, City,
-State, Zip}
-    \address[Another University]{Department, Street, City, State, Zip}
+  \title{A comparison of model-based and design-based approaches for spatial
+data.}
+    \author[USEPA]{Michael Dumelle\corref{1}}
+   \ead{Dumelle.Michael@epa.gov} 
+    \author[STLAW]{Matthew Higham\corref{1}}
+   \ead{mhigham@stlaw.edu} 
+    \author[OSU]{Lisa Madsen}
+
+    \author[USEPA]{Anthony R. Olsen}
+
+    \author[NOAA]{Jay M. Ver Hoef}
+
+      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
+Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Math, Computer Science, and
+Statistics, 23 Romoda Drive, Canton, New York, 13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
+Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
+Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
       \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
 
@@
 
+Please leave comments in your
+\textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}:
+\textcolor{blue}{Michael}, \textcolor{magenta}{Matt},
+\textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
+
+\hypertarget{background}{%
+\section{Background}\label{background}}
+
 \hypertarget{design-based-philosphy}{%
@@
 
+Design-Based Overview
+
+Design-based inference uses characteristics of the sampling design to to
+estimate parameters of interest Typically, there are few assumptions
+involved because intervals are derived using the sampling design
+itself\ldots\ldots\ldots..
+
 \hypertarget{model-based-philosphy}{%
 \subsection{Model-Based Philosphy}\label{model-based-philosphy}}
+
+Model-Based Overview
 
@@
 philosophical difference between sampling-based and model-based
-approaches. Instead of \emph{estimating} an fixed unknown mean, we are
+approaches. Instead of \emph{estimating} a fixed unknown mean, we are
 \emph{predicting} the value of the mean, a random variable. We know that
@@
 
+Design-Based and Model-Based Comparisons
+
 There have been many comparisons between the two paradigms\ldots\ldots.
@@
 Analysis}\label{spatially-balanced-design-and-analysis}}
+
+Spatially balanced sampling algorithms use spatial information to obtain
+samples spread out in space. Spatially balanced samples are useful
+because they tend to yield estimators that are more precise than
+estimators constructed from an sampling algorithm that is not spatially
+balanced ((Barabesi and Franceschi, 2011; Benedetti et al., 2017;
+Grafström and Lundström, 2013; Robertson et al., 2013; Stevens Jr and
+Olsen, 2004; Wang et al., 2013)). Many spatially balanced sampling
+algorithms exist, including the Generalized Random Tessellation
+Stratified (Stevens Jr and Olsen, 2004), the Local Pivotal Method
+(Grafström et al., 2012; Grafström and Matei, 2018), Spatially
+Correlated Poisson Sampling (Grafström, 2012), Balanced Acceptance
+Sampling (Robertson et al., 2013), Within-Sample-Distance (Benedetti and
+Piersimoni, 2017), and Halton Iterative Partitioning (Robertson et al.,
+2018) algorithms. Here we focus on the Generalized Random Tessellation
+Stratified (GRTS) algorithm, which has several attractive properties
+that we discuss next.
+
+The GRTS algorithm is used to sample from finite and infinite sample
+frames. A finite sample frames contains a finite number of sampling
+units and is related to a point geometry. An infinite sample frame
+contains an infinite number of sampling units and is related to linear
+and polygon geometries. Examples of point, linear, and polygon resources
+include lake centroids, stream networks, and wetland areas,
+respectively. In addition to its applicability for finite and infinite
+sample frames, the GRTS algorithm naturally accommodates stratified
+designs and designs with unequal selection probabilities. The algorithm
+has also been used to select replacement sites using reverse
+hierarchical ordering (Stevens Jr and Olsen, 2004). Replacement sites
+are used to replace sites in the original sample that cannot be sampled,
+often as a result of physical difficulty in reaching the site or
+landowner denial of access to the sites. More recently, the GRTS
+algorithm also accommodates legacy (historical) sites, minimum distance
+between sites, and nearest neighbor replacement sites. The GRTS
+algorithm is implemented in the \textbf{\textsf{R}} package
+\texttt{spsurvey} (Dumelle et al., 2021).
+
+accomodates stratification, unequal selection probabilities, oversample
+sites It has also recently been updated to accommodate legacy
+(historical) sites, minimum distance between sites, and a nearest
+neighbor replacement sites comp efficient software
 
@@
 geostatistical kriging framework to the finite population setting.
-Instead of relying on a specific sampling design, we assume the data
-were produced by a spatial stochastic process with spatial parameters
-that can be estimated.
+Instead of basing inference off of a specific sampling design, we assume
+the data were generated by a spatial process with parameters that can be
+estimated using the framework of a model.
 
@@
 \(\mathbf{z}\) that can be measured on a finite number of \(N\) sites,
-our goal is to predict some linear function of all of the sample units,
+we want to predict some linear function of the response variable,
 \(\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
-\(\mathbf{b}\) is a vector of weights. One common vector of weights is a
-vector of 1's to predict the total abundance in the region.
+\(\mathbf{b}\) is a vector of weights. For example, if we want to
+predict the total abundance across all sites, then we would use a vector
+of 1's for the weights.
 
@@
 design matrices for the sampled and unsampled sites, respectively, and
-\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are zero-mean random errors for
-the sampled and unsampled sites. Denoting
-\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), then we
-assume that \(E(\bm{\delta})\) = \(\mathbf{0}\).
+\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
+sampled and unsampled sites. Denoting
+\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
+that \(E(\bm{\delta})\) = \(\mathbf{0}\).
 
@@
 entry for \(\mathop{\mathrm{{var}}}(\bm{\delta})\) is \mbox{}
-\begin{equation}
-\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \label{equation:expcov}
-\end{equation} where \(h_{i,j}\) is the distance between sites \(i\) and
+\begin{align}\label{equation:expcov}
+\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \\
+\textcolor{blue}{\sigma^2[(1 - v)\exp(-3\mathbf{h}_{i,j}/\phi) + v\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}]}, 
+\end{align} where \(h_{i,j}\) is the distance between sites \(i\) and
 \(j\), and \(\bm{\theta}\) is a vector of spatial covariance parameters
@@
 With the above model formulation, the Best Linear Unbiased Predictor
-(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\) and its prediction variance
-can be computed. While details of the derivation are in (Ver Hoef,
-2008), we note here that the predictor and its variance are both
-moment-based. Neither require a particular distribution for
-\(\mathbf{z}\).
+(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\)
+\textcolor{blue}{Did you mean to give the form of the BLUP here? $\tau(\mathbf{b}'\mathbf{z})$ is vague}
+and its prediction variance can be computed. While details of the
+derivation are in (Ver Hoef, 2008), we note here that the predictor and
+its variance are both moment-based. Neither require a particular
+distribution for \(\mathbf{z}\).
 
-\textbf{Software Implementation}: This probably goes in a different spot
-(right before simulations?), but putting it here while it's on my mind.
+\hypertarget{numerical-study}{%
+\section{Numerical Study}\label{numerical-study}}
+
+\hypertarget{software}{%
+\subsection{Software}\label{software}}
 
 FPBK can be readily performed in \texttt{R} with the \texttt{sptotal}
-package (Matt et al., 2020). We use \texttt{sptotal} for both the
+package (Higham et al., 2020). We use \texttt{sptotal} for both the
 simulation analysis and the application, estimating parameters with
@@
 
-\hypertarget{installation}{%
-\paragraph{Installation}\label{installation}}
-
-If the document class \emph{elsarticle} is not available on your
-computer, you can download and install the system package
-\emph{texlive-publishers} (Linux) or install the LaTeX package
-\emph{elsarticle} using the package manager of your TeX installation,
-which is typically TeX Live or MikTeX.
-
-\hypertarget{usage}{%
-\paragraph{Usage}\label{usage}}
-
-Once the package is properly installed, you can use the document class
-\emph{elsarticle} to create a manuscript. Please make sure that your
-manuscript follows the guidelines in the Guide for Authors of the
-relevant journal. It is not necessary to typeset your manuscript in
-exactly the same way as an article, unless you are submitting to a
-camera-ready copy (CRC) journal.
-
-\hypertarget{functionality}{%
-\paragraph{Functionality}\label{functionality}}
-
-The Elsevier article class is based on the standard article class and
-supports almost all of the functionality of that class. In addition, it
-features commands and options to format the
-
-\begin{itemize}
-\item
-  document style
-\item
-  baselineskip
-\item
-  front matter
-\item
-  keywords and MSC codes
-\item
-  theorems, definitions and proofs
-\item
-  lables of enumerations
-\item
-  citation style and labeling.
-\end{itemize}
-
-\hypertarget{front-matter}{%
-\section{Front matter}\label{front-matter}}
-
-The author names and affiliations could be formatted in two ways:
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\item
-  Group the authors per affiliation.
-\item
-  Use footnotes to indicate the affiliations.
-\end{enumerate}
-
-See the front matter of this document for examples. You are recommended
-to conform your choice to the journal you are submitting to.
-
-\hypertarget{bibliography-styles}{%
-\section{Bibliography styles}\label{bibliography-styles}}
-
-There are various bibliography styles available. You can select the
-style of your choice in the preamble of this document. These styles are
-Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use BibTeX~to generate your bibliography and include DOIs
-whenever available.
-
-Here are two sample references: (\textbf{Feynman1963118?};
-\textbf{Dirac1953888?}).
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-ChilesEtAl1999GeostatisticsModelingSpatial}{}%
-Chiles, J.-P., Delfiner, P., 1999. Geostatistics: {Modeling Spatial
-Uncertainty}. {John Wiley \& Sons}, New York.
+\leavevmode\hypertarget{ref-barabesi2011sampling}{}%
+Barabesi, L., Franceschi, S., 2011. Sampling properties of spatial total
+estimators under tessellation stratified designs. Environmetrics 22,
+271--278.
 
-\leavevmode\hypertarget{ref-sptotal2020}{}%
-Matt, H., Jay, V.H., Bryce, F., 2020. Sptotal: Predicting totals and
-weighted sums from spatial data.
+\leavevmode\hypertarget{ref-benedetti2017spatially}{}%
+Benedetti, R., Piersimoni, F., 2017. A spatially balanced design with
+probability function proportional to the within sample distance.
+Biometrical Journal 59, 1067--1084.
 
-\leavevmode\hypertarget{ref-VerHoef2008spatial}{}%
+\leavevmode\hypertarget{ref-benedetti2017spatiallyreview}{}%
+Benedetti, R., Piersimoni, F., Postiglione, P., 2017. Spatially balanced
+sampling: A review and a reappraisal. International Statistical Review
+85, 439--454.
+
+\leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
+Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
+Uncertainty. John Wiley \& Sons, New York.
+
+\leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
+Dumelle, M., Olsen, A.R., Kincaid, T., Weber, M., 2021. Selecting and
+analyzing spatial probability samples in r using spsurvey. Manuscript
+Submitted for Publication.
+
+\leavevmode\hypertarget{ref-grafstrom2012spatiallypoisson}{}%
+Grafström, A., 2012. Spatially correlated poisson sampling. Journal of
+Statistical Planning and Inference 142, 139--147.
+
+\leavevmode\hypertarget{ref-grafstrom2013well}{}%
+Grafström, A., Lundström, N.L., 2013. Why well spread probability
+samples are balanced. Open Journal of Statistics 3, 36--41.
+
+\leavevmode\hypertarget{ref-grafstrom2012spatially}{}%
+Grafström, A., Lundström, N.L., Schelin, L., 2012. Spatially balanced
+sampling through the pivotal method. Biometrics 68, 514--520.
+
+\leavevmode\hypertarget{ref-grafstrom2018spatially}{}%
+Grafström, A., Matei, A., 2018. Spatially balanced sampling of
+continuous populations. Scandinavian Journal of Statistics 45, 792--805.
+
+\leavevmode\hypertarget{ref-higham2020sptotal}{}%
+Higham, M., Ver Hoef, J., Bryce, F., 2020. Sptotal: Predicting totals
+and weighted sums from spatial data.
+
+\leavevmode\hypertarget{ref-robertson2013bas}{}%
+Robertson, B., Brown, J., McDonald, T., Jaksons, P., 2013. BAS: Balanced
+acceptance sampling of natural resources. Biometrics 69, 776--784.
+
+\leavevmode\hypertarget{ref-robertson2018halton}{}%
+Robertson, B., McDonald, T., Price, C., Brown, J., 2018. Halton
+iterative partitioning: Spatially balanced sampling via partitioning.
+Environmental and Ecological Statistics 25, 305--323.
+
+\leavevmode\hypertarget{ref-stevens2004spatially}{}%
+Stevens Jr, D.L., Olsen, A.R., 2004. Spatially balanced sampling of
+natural resources. Journal of the american Statistical association 99,
+262--278.
+
+\leavevmode\hypertarget{ref-verhoef2008spatial}{}%
 Ver Hoef, J.M., 2008. Spatial methods for plot-based sampling of
@@
 
-\end{CSLReferences}
+\leavevmode\hypertarget{ref-wang2013design}{}%
+Wang, J.-F., Jiang, C.-S., Hu, M.-G., Cao, Z.-D., Guo, Y.-S., Li, L.-F.,
+Liu, T.-J., Meng, B., 2013. Design-based spatial sampling: Theory and
+implementation. Environmental modelling \& software 40, 280--288.
 

### 2021-05-20 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor update

@@
 Spatially balanced sampling algorithms use spatial information to obtain samples spread out in space.
-Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from an sampling algorithm that is not spatially balanced ([@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design]).
+Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from an sampling algorithm that is not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
 Many spatially balanced sampling algorithms exist, including the Generalized Random Tessellation Stratified [@stevens2004spatially], the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
grts background updates

@@
 
+Several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS or SIRS (stratified IRS).
+No comparison has been made between spatial models and spatially balanced design-based approaches.
+
 # Background
@@
 Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from an sampling algorithm that is not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
+To quantify spatial balance, @stevens2004spatially proposed statistics based on Voroni polygons.
 Many spatially balanced sampling algorithms exist, including the Generalized Random Tessellation Stratified [@stevens2004spatially], the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
@@
 
-describe grts
+The GRTS algorithm works by mapping two-dimensional space into one-dimensional space while incorporating some elements of randomness.
+First a square bounding box is placed over the sample frame and divided into four equally sized squares randomly given an level-one address of 0, 1, 2, or 3.
+Then the inclusion probabilities of the sampling units in each square are summed.
+If this sum is at least one in any of the cells, a set of four sub-squares are placed within each square.
+Each set of four sub-squares are randomly given a second address (a level-two address) of 0, 1, 2, 3.
+Then the inclusion probabilities in each sub-square are summed, and if any of these sums are at least one, a third level of squares are added in the same manner.
+This hierarchical addressing continues until the inclusion probabilities in the smallest squares are all less than 1.
+The squares are then mapped to the one-dimensional line in order of their addresses.
+The length of this line equals $n$, the desired sample size (the sum of the inclusion probabilities).
+A uniform random variable, denoted by $u_1$ is simulated in $[0, 1]$ and placed on this line.
+The location of $u_1$ on the line can be linked to a sampling unit, denoted by $s_1$.
+Then $u_2 \equiv u_1 + 1$, which can be linked to another sampling unit, denoted by $s_2$.
+Because the GRTS algorithm requires the inclusion probabilities of the smallest sub-cells to be less than one, $s_1$ and $s_2$ must be distinct.
+This process continues until the $n$ sampling units have been selected as part of the sample.
+Further details are provided by [@stevens2004spatially] and [@dumelle2021spsurvey].
 
-discuss spatial balance
-
-describe estimation and local neighborhood variance
-
-<!-- \citep{StevensOlsen2003VarianceEstimation, StevensOlsen2004GRTS} -->
+The Horvitz-Thompson estimator [@horvitz1952generalization] and its continuous analog [@cordy1993extension] yield unbiased estimates of population totals (and means).
+If $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$, denoted by $\hat{\tau}_{HT}$, is given by
+\begin{align}\label{eq:htest}
+  \hat{\tau}_{HT} = \sum_{i = 1}^n Y_i \pi_i^{-1},
+\end{align}
+where $Y_i$ is the observed value of the $i$th sampling unit, and $\pi_i$ is the inclusion probability of the $i$th sampling unit.
+Estimating the variance of $\hat{\tau}_{HT}$ directly relies on knowing the probability that $s_i$ and $s_j$ are both included in the sample (second-order inclusion probability), which can be unfeasible to compute.
+Furthermore, this variance estimator does incorporate spatial locations.
+To address these challenges, [@stevens2003variance] proposed an alternative variance estimator: the local neighborhood variance estimator.
+The local neighborhood variance estimator does not require second-order inclusion probabilities.
+It also incorporates spatial locations, which tends to reduce the estimated variance of compared $\hat{\tau}_{HT}$ to a variance estimator ignoring spatial location [@stevens2003variance].
 

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
grts background updates

@@
 
+Several comparisons between model-based and design-based for spatial
+data have been considered, but they tend to compare spatial models to
+IRS or SIRS (stratified IRS). No comparison has been made between
+spatial models and spatially balanced design-based approaches.
+
 \hypertarget{background}{%
@@
 estimators constructed from an sampling algorithm that is not spatially
-balanced ((Barabesi and Franceschi, 2011; Benedetti et al., 2017;
+balanced (Barabesi and Franceschi, 2011; Benedetti et al., 2017;
 Grafström and Lundström, 2013; Robertson et al., 2013; Stevens Jr and
-Olsen, 2004; Wang et al., 2013)). Many spatially balanced sampling
-algorithms exist, including the Generalized Random Tessellation
-Stratified (Stevens Jr and Olsen, 2004), the Local Pivotal Method
-(Grafström et al., 2012; Grafström and Matei, 2018), Spatially
-Correlated Poisson Sampling (Grafström, 2012), Balanced Acceptance
-Sampling (Robertson et al., 2013), Within-Sample-Distance (Benedetti and
-Piersimoni, 2017), and Halton Iterative Partitioning (Robertson et al.,
-2018) algorithms. Here we focus on the Generalized Random Tessellation
-Stratified (GRTS) algorithm, which has several attractive properties
-that we discuss next.
+Olsen, 2004; Wang et al., 2013). To quantify spatial balance, Stevens Jr
+and Olsen (2004) proposed statistics based on Voroni polygons. Many
+spatially balanced sampling algorithms exist, including the Generalized
+Random Tessellation Stratified (Stevens Jr and Olsen, 2004), the Local
+Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
+Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
+Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
+(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
+(Robertson et al., 2018) algorithms. Here we focus on the Generalized
+Random Tessellation Stratified (GRTS) algorithm, which has several
+attractive properties that we discuss next.
 
@@
 
-accomodates stratification, unequal selection probabilities, oversample
-sites It has also recently been updated to accommodate legacy
-(historical) sites, minimum distance between sites, and a nearest
-neighbor replacement sites comp efficient software
+The GRTS algorithm works by mapping two-dimensional space into
+one-dimensional space while incorporating some elements of randomness.
+First a square bounding box is placed over the sample frame and divided
+into four equally sized squares randomly given an level-one address of
+0, 1, 2, or 3. Then the inclusion probabilities of the sampling units in
+each square are summed. If this sum is at least one in any of the cells,
+a set of four sub-squares are placed within each square. Each set of
+four sub-squares are randomly given a second address (a level-two
+address) of 0, 1, 2, 3. Then the inclusion probabilities in each
+sub-square are summed, and if any of these sums are at least one, a
+third level of squares are added in the same manner. This hierarchical
+addressing continues until the inclusion probabilities in the smallest
+squares are all less than 1. The squares are then mapped to the
+one-dimensional line in order of their addresses. The length of this
+line equals \(n\), the desired sample size (the sum of the inclusion
+probabilities). A uniform random variable, denoted by \(u_1\) is
+simulated in \([0, 1]\) and placed on this line. The location of \(u_1\)
+on the line can be linked to a sampling unit, denoted by \(s_1\). Then
+\(u_2 \equiv u_1 + 1\), which can be linked to another sampling unit,
+denoted by \(s_2\). Because the GRTS algorithm requires the inclusion
+probabilities of the smallest sub-cells to be less than one, \(s_1\) and
+\(s_2\) must be distinct. This process continues until the \(n\)
+sampling units have been selected as part of the sample. Further details
+are provided by (Stevens Jr and Olsen, 2004) and (Dumelle et al., 2021).
+
+The Horvitz-Thompson estimator (Horvitz and Thompson, 1952) and its
+continuous analog (Cordy, 1993) yield unbiased estimates of population
+totals (and means). If \(\tau\) is a population total, then the
+Horvitz-Thompson estimator of \(\tau\), denoted by \(\hat{\tau}_{HT}\),
+is given by \begin{align}\label{eq:htest}
+  \hat{\tau}_{HT} = \sum_{i = 1}^n Y_i \pi_i^{-1},
+\end{align} where \(Y_i\) is the observed value of the \(i\)th sampling
+unit, and \(\pi_i\) is the inclusion probability of the \(i\)th sampling
+unit. Estimating the variance of \(\hat{\tau}_{HT}\) directly relies on
+knowing the probability that \(s_i\) and \(s_j\) are both included in
+the sample (second-order inclusion probability), which can be unfeasible
+to compute. Furthermore, this variance estimator does incorporate
+spatial locations. To address these challenges, (Stevens Jr and Olsen,
+2003) proposed an alternative variance estimator: the local neighborhood
+variance estimator. The local neighborhood variance estimator does not
+require second-order inclusion probabilities. It also incorporates
+spatial locations, which tends to reduce the estimated variance of
+compared \(\hat{\tau}_{HT}\) to a variance estimator ignoring spatial
+location (Stevens Jr and Olsen, 2003).
 
@@
 
+\leavevmode\hypertarget{ref-cordy1993extension}{}%
+Cordy, C.B., 1993. An extension of the horvitz---thompson theorem to
+point sampling from a continuous universe. Statistics \& Probability
+Letters 18, 353--362.
+
 \leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
@@
 
+\leavevmode\hypertarget{ref-horvitz1952generalization}{}%
+Horvitz, D.G., Thompson, D.J., 1952. A generalization of sampling
+without replacement from a finite universe. Journal of the American
+statistical Association 47, 663--685.
+
 \leavevmode\hypertarget{ref-robertson2013bas}{}%
@@
 Environmental and Ecological Statistics 25, 305--323.
+
+\leavevmode\hypertarget{ref-stevens2003variance}{}%
+Stevens Jr, D.L., Olsen, A.R., 2003. Variance estimation for spatially
+balanced samples of environmental resources. Environmetrics 14,
+593--610.
 

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
design-based overview updates

@@
 
-Design-based inference uses characteristics of the sampling design to to estimate parameters of interest Typically, there are few assumptions involved because intervals are derived using the sampling design itself...........
+The design-based approach to inference uses characteristics of the sampling design to estimate parameters of interest.
+The design-based approach assumes the data, $\mathbf{Y}$, are fixed.
+Randomness is incorporated through the random selection of sampling units based on probabilities derived using the sampling design itself.
+Some examples of commonly used sampling designs include independent random sampling (IRS), stratified random sampling, and cluster sampling.
+Viewing the data as fixed and incorporating randomness through the sampling design yields estimators having very few assumptions.
+Asymptotic properties of these types of estimators are derived using limiting arguments.
+Means and totals, for example, are asymptotically Gaussian.
+@sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
 
@@
 On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimates if the additional assumptions hold.
-Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often _prediction_ of an unknown quantity.
+Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
 This is a fundamental philosophical difference between sampling-based and model-based approaches.
-Instead of _estimating_ a fixed unknown mean, we are _predicting_ the value of the mean, a random variable.
+Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable.
 We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty.

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
design-based overview updates

@@
 
-Design-based inference uses characteristics of the sampling design to to
-estimate parameters of interest Typically, there are few assumptions
-involved because intervals are derived using the sampling design
-itself\ldots\ldots\ldots..
+The design-based approach to inference uses characteristics of the
+sampling design to estimate parameters of interest. The design-based
+approach assumes the data, \(\mathbf{Y}\), are fixed. Randomness is
+incorporated through the random selection of sampling units based on
+probabilities derived using the sampling design itself. Some examples of
+commonly used sampling designs include independent random sampling
+(IRS), stratified random sampling, and cluster sampling. Viewing the
+data as fixed and incorporating randomness through the sampling design
+yields estimators having very few assumptions. Asymptotic properties of
+these types of estimators are derived using limiting arguments. Means
+and totals, for example, are asymptotically Gaussian. Särndal et al.
+(2003) and Lohr (2009) provide thorough reviews of the design-based
+approach.
 
@@
 
+\leavevmode\hypertarget{ref-lohr2009sampling}{}%
+Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.
+
 \leavevmode\hypertarget{ref-robertson2013bas}{}%
@@
 Environmental and Ecological Statistics 25, 305--323.
+
+\leavevmode\hypertarget{ref-sarndal2003model}{}%
+Särndal, C.-E., Swensson, B., Wretman, J., 2003. Model assisted survey
+sampling. Springer Science \& Business Media.
 

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor typo fixes in initial draft.

@@
 
-Several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS or SIRS (stratified IRS).
+Several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS).
 No comparison has been made between spatial models and spatially balanced design-based approaches.
@@
 
-On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimates if the additional assumptions hold.
+On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimators if the additional assumptions hold.
 Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
@@
 Spatially balanced sampling algorithms use spatial information to obtain samples spread out in space.
-Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from an sampling algorithm that is not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
+Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from a sampling algorithm that is not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
 To quantify spatial balance, @stevens2004spatially proposed statistics based on Voroni polygons.
@@
 The GRTS algorithm is used to sample from finite and infinite sample frames.
-A finite sample frames contains a finite number of sampling units and is related to a point geometry.
+A finite sample frame contains a finite number of sampling units and is related to a point geometry.
 An infinite sample frame contains an infinite number of sampling units and is related to linear and polygon geometries.
@@
 Estimating the variance of $\hat{\tau}_{HT}$ directly relies on knowing the probability that $s_i$ and $s_j$ are both included in the sample (second-order inclusion probability), which can be unfeasible to compute.
-Furthermore, this variance estimator does incorporate spatial locations.
+Furthermore, this variance estimator does not incorporate spatial locations.
 To address these challenges, [@stevens2003variance] proposed an alternative variance estimator: the local neighborhood variance estimator.

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
outline for introduction.
added some lit overview

@@
 
+define sampling design, define parameters, define model-based, etc.
+
+introduction of spatial data / definition of spatial and maybe general background?
+
+2-3 sentences about each design
+
+goals and outline of paper
+
 Several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS).
 No comparison has been made between spatial models and spatially balanced design-based approaches.
+
+* hansen1983evaluation - general overview (in non spatial context), talk about population vs superpopulation give a set of guiding principles in the discussion
+
+* brus1997random - design and model based explanation and simulation study for spatial means, make a decision tree for choosing among approaches
+
+* verhoef2002sampling - strictly defines classical sampling, compares FPBK to classical, BK to classical
 

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added to outline from model comparison papers.

@@
 
-define sampling design, define parameters, define model-based, etc.
+define sampling design, define parameters, define model-based, sample, population, super-population?, etc.
 
 introduction of spatial data / definition of spatial and maybe general background?
+
+give some examples of possible spatial data
 
@@
 goals and outline of paper
+
+brus2020statistical: most similar to what this paper will be, directly comparing the two approaches.
+This has a lot of other potentially useful references to go back to.
+(published in 2021 so likely doesn't have many that have cited it yet).
+
+de1990model: early comparison of design v model in spatial context
 
@@
 
+Other model based stuff:
+chan2020bayesian: Bayesian model-based approach.
+Based on splines: 
+Cicchitelli, G. and Montanari, G. E. (2012). “Model-Assisted Estimation of a Spatial Population
+Mean.” International Statistical Review, 80(1): 111–126.
 <!-- \citep{Cressie2015statistics, Schabenberger2017statistical} -->
@@
 <!-- \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian} -->
+
+sterba2009alternative: psychological context comparison, has a lot of the older background (e.g. Fisher, Neyman, etc.)
+
+Cooper2006sampling: ecological context, also has a nice very general background comparing two approaches, introduces a "model-assisted" variance estimator.
 

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor updates

@@
 \begin{align}\label{equation:expcov}
-\cov(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \\
-\textcolor{blue}{\sigma^2[(1 - v)\exp(-3\mathbf{h}_{i,j}/\phi) + v\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}]}, 
+\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
 \end{align}
@@
 
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ \textcolor{blue}{Did you mean to give the form of the BLUP here?
-$\tau(\mathbf{b}'\mathbf{z})$ is vague} and its prediction variance can be computed.
+With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ $\tau(\mathbf{b}'\mathbf{z})$ is vague} and its prediction variance can be computed.
 While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor updates

@@
 data have been considered, but they tend to compare spatial models to
-IRS or SIRS (stratified IRS). No comparison has been made between
-spatial models and spatially balanced design-based approaches.
+IRS (Independent Random Sampling) or SIRS (stratified IRS). No
+comparison has been made between spatial models and spatially balanced
+design-based approaches.
 
@@
 On the other hand, model-based inference imposes additional assumptions
-on the data with a potential to provide more precise estimates if the
+on the data with a potential to provide more precise estimators if the
 additional assumptions hold. Instead of estimating true but unknown
@@
 because they tend to yield estimators that are more precise than
-estimators constructed from an sampling algorithm that is not spatially
+estimators constructed from a sampling algorithm that is not spatially
 balanced (Barabesi and Franceschi, 2011; Benedetti et al., 2017;
@@
 The GRTS algorithm is used to sample from finite and infinite sample
-frames. A finite sample frames contains a finite number of sampling
-units and is related to a point geometry. An infinite sample frame
-contains an infinite number of sampling units and is related to linear
-and polygon geometries. Examples of point, linear, and polygon resources
-include lake centroids, stream networks, and wetland areas,
-respectively. In addition to its applicability for finite and infinite
-sample frames, the GRTS algorithm naturally accommodates stratified
-designs and designs with unequal selection probabilities. The algorithm
-has also been used to select replacement sites using reverse
-hierarchical ordering (Stevens Jr and Olsen, 2004). Replacement sites
-are used to replace sites in the original sample that cannot be sampled,
-often as a result of physical difficulty in reaching the site or
-landowner denial of access to the sites. More recently, the GRTS
-algorithm also accommodates legacy (historical) sites, minimum distance
-between sites, and nearest neighbor replacement sites. The GRTS
-algorithm is implemented in the \textbf{\textsf{R}} package
-\texttt{spsurvey} (Dumelle et al., 2021).
+frames. A finite sample frame contains a finite number of sampling units
+and is related to a point geometry. An infinite sample frame contains an
+infinite number of sampling units and is related to linear and polygon
+geometries. Examples of point, linear, and polygon resources include
+lake centroids, stream networks, and wetland areas, respectively. In
+addition to its applicability for finite and infinite sample frames, the
+GRTS algorithm naturally accommodates stratified designs and designs
+with unequal selection probabilities. The algorithm has also been used
+to select replacement sites using reverse hierarchical ordering (Stevens
+Jr and Olsen, 2004). Replacement sites are used to replace sites in the
+original sample that cannot be sampled, often as a result of physical
+difficulty in reaching the site or landowner denial of access to the
+sites. More recently, the GRTS algorithm also accommodates legacy
+(historical) sites, minimum distance between sites, and nearest neighbor
+replacement sites. The GRTS algorithm is implemented in the
+\textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al., 2021).
 
@@
 the sample (second-order inclusion probability), which can be unfeasible
-to compute. Furthermore, this variance estimator does incorporate
+to compute. Furthermore, this variance estimator does not incorporate
 spatial locations. To address these challenges, (Stevens Jr and Olsen,
@@
 \begin{align}\label{equation:expcov}
-\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_3 + \theta_1\exp(-h_{i,j}/\theta_2), \\
-\textcolor{blue}{\sigma^2[(1 - v)\exp(-3\mathbf{h}_{i,j}/\phi) + v\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}]}, 
+\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
 \end{align} where \(h_{i,j}\) is the distance between sites \(i\) and
@@
 (BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\)
-\textcolor{blue}{Did you mean to give the form of the BLUP here? $\tau(\mathbf{b}'\mathbf{z})$ is vague}
-and its prediction variance can be computed. While details of the
-derivation are in (Ver Hoef, 2008), we note here that the predictor and
-its variance are both moment-based. Neither require a particular
-distribution for \(\mathbf{z}\).
+\(\tau(\mathbf{b}'\mathbf{z})\) is vague\} and its prediction variance
+can be computed. While details of the derivation are in (Ver Hoef,
+2008), we note here that the predictor and its variance are both
+moment-based. Neither require a particular distribution for
+\(\mathbf{z}\).
 

### 2021-06-07 R/sim_one.R
initial simulation functions.

@@ -1,0 +1,1 @@
+## get one simulation

### 2021-06-07 R/sim_pop.R

@@ -1,0 +1,4 @@
+## simulate population
+sim_pop <- function(N = 100, n = 50, cortype, psill, erange, nugget) {
+
+}

### 2021-06-07 inst/scripts/simulation1_skeleton.R

@@ -5,6 +5,12 @@ library(sptotal)
 # create simulation function 1
 sim1 <- function(seed, ...) { # dots are placeholders for other arguments
 
+  ## sim 1 model specified correctly
+  ##     - grts gets grts sample
+  ##     - model gets srs
+  ##         - one range, one sample size, for now
+  ##         - predict/estimate total
+  ##
   # simulate data
 
   # fit models

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd

@@
 
-Several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS).
+Several comparisons between modegitl-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS).
 No comparison has been made between spatial models and spatially balanced design-based approaches.
@@
 * verhoef2002sampling - strictly defines classical sampling, compares FPBK to classical, BK to classical
+
+* wang2012spatial - sampling and inference steps, thorough overview in terms of spatial context, lots of references
 

### 2021-06-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor edit.

@@
 
+Several comparisons between the two paradigms have been made in many different contexts. @sterba2009alternative give psych context, old history fisher / Neyman
+
+Many new methods have been constructed that combine aspects of both design-based and model-based approaches.
+Our goal is not to expand upon these methods nor is it to develop a novel method that combines the two approaches.
+
 Design-Based and Model-Based Comparisons
 
-<!-- \citep{Hansen1983evaluation, Brus1997random, VerHoef2002sampling, Cooper2006sampling, sterba2009alternative, Brus2020statistical, Chan2020bayesian} -->
-
-sterba2009alternative: psychological context comparison, has a lot of the older background (e.g. Fisher, Neyman, etc.)
+: psychological context comparison, has a lot of the older background (e.g. Fisher, Neyman, etc.)
 
 Cooper2006sampling: ecological context, also has a nice very general background comparing two approaches, introduces a "model-assisted" variance estimator.
-
-There have been many comparisons between the two paradigms.......
 

### 2021-06-07 R/point_sim.R
grid and random point samples.

@@ -1,0 +1,54 @@
+#' Creates a systematic grid of points.
+#'
+#' Creates a systematic grid of points. This function is only used
+#' in simulating data sets.
+#'
+#' @param lower_x_lim the lower limit for x-coordinate, default is 0
+#' @param upper_x_lim the upper limit for x-coordinate, default is 1
+#' @param lower_y_lim the lower limit for y-coordinate, default is 0
+#' @param upper_y_lim the upper limit for y-coordinate, default is 1
+#' @param ncol the number of cols in the systematic grid, default is 10
+#' @param nrow the number of rows in the systematic grid, default is 10
+
+#' @return A data.frame with x- and y-coordinates of simulated locations
+#' @author Jay Ver Hoef
+#' @export
+
+pointSimSyst <- function (nrow = 10, ncol = 10, lower_x_lim = 0, upper_x_lim = 1,
+                          lower_y_lim = 0, upper_y_lim = 1)
+{
+  x_range <- upper_x_lim - lower_x_lim
+  y_range <- upper_y_lim - lower_y_lim
+  y_mat <- lower_y_lim + y_range * (nrow - matrix(rep(1:nrow,
+                                                      times = ncol), nrow = nrow, ncol = ncol))/(nrow) + y_range/(2 *
+                                                                                                                    nrow)
+  x_mat <- lower_x_lim + x_range * (t(matrix(rep(1:ncol, times = nrow),
+                                             nrow = ncol, ncol = nrow)) - 1)/(ncol) + x_range/(2 *
+                                                                                                 ncol)
+  data.frame(x = matrix(x_mat, ncol = 1), y = matrix(y_mat,
+                                                     ncol = 1))
+}
+
+#' simulate completely spatially random point patterns.
+#'
+#' simulates a completely spatially random point patterns. This function is
+#' only used in simulating data sets.
+#'
+#' @param npoints number of points to add that are completely spatially random (CSR), default = 100
+#' @param lower_x_lim left limit of boundary, default = 0
+#' @param upper_x_lim right limit of boundary, default = 1
+#' @param lower_y_lim lower limit of boundary, default = 0
+#' @param upper_y_lim upper limit of boundary, default = 1
+#'
+#' @return data.frame of two columns, x-coordinate in the first, and y-coordinate in the second.
+#'
+#' @author Jay Ver Hoef
+
[…]

### 2021-06-08 R/sim_one.R
added true mean and var.

@@ -59,10 +59,14 @@ sim_one <- function(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", ps
   ## just return the mean info
   design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
 
+  realized_mean <- mean(data$response)
+  realized_var <- var(data$response)
 
   # store output
   output <- data.frame(
     approach = c("Design", "Model"),
+    true_mean = c(realized_mean, realized_mean),
+    true_var = c(realized_var, realized_var),
     estimate = c(design_mean$Estimate, as.vector(model_mean)),
     sd = c(design_mean$StdError, as.vector(model_se)),
     lb = c(design_mean$LCB95Pct, as.vector(model_lb)),
@@ -71,3 +75,12 @@ sim_one <- function(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", ps
 
   output
 }
+
+
+
+## example
+# df <- sim_pop(N = 100, n = 50, gridded = TRUE, cortype = "exponential",
+#               psill = 1.1, erange = 1.01, nugget = 0.3)
+# library(dplyr)
+# sim_one(data = df, xcoord = "x", ycoord = "y", response = "response",
+#         cortype_est = "exponential")

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
update to paradigm comparison section.

@@
 
-Several comparisons between the two paradigms have been made in many different contexts. @sterba2009alternative give psych context, old history fisher / Neyman
-
-Many new methods have been constructed that combine aspects of both design-based and model-based approaches.
-Our goal is not to expand upon these methods nor is it to develop a novel method that combines the two approaches.
-
-Design-Based and Model-Based Comparisons
-
-: psychological context comparison, has a lot of the older background (e.g. Fisher, Neyman, etc.)
-
-Cooper2006sampling: ecological context, also has a nice very general background comparing two approaches, introduces a "model-assisted" variance estimator.
+Several comparisons between the two paradigms have been made in many different contexts.
+@sterba2009alternative give some history of the two paradigms as well as some applications in psychological contexts.
+@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
+We note that while there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
+CITE some more f these), our goal is not to expand upon or improve these methods.
 

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
Added citations and started introductory paragraph.

@@
 
+@de1990model give an early comparison of the paradigms for spatial data, quashing the belief that classical sampling-based methods could not be used for spatially correlated data.
+Thereafter, several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS) [@brus1997random].
+No comparison has been made between spatial models and spatially balanced design-based approaches.
+
 brus2020statistical: most similar to what this paper will be, directly comparing the two approaches.
@@
 
-de1990model: early comparison of design v model in spatial context
-
-Several comparisons between modegitl-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS).
-No comparison has been made between spatial models and spatially balanced design-based approaches.
-
 * hansen1983evaluation - general overview (in non spatial context), talk about population vs superpopulation give a set of guiding principles in the discussion
-
-* brus1997random - design and model based explanation and simulation study for spatial means, make a decision tree for choosing among approaches
 
@@
 The design-based approach to inference uses characteristics of the sampling design to estimate parameters of interest.
-The design-based approach assumes the data, $\mathbf{Y}$, are fixed.
+The design-based approach assumes the data, $\mathbf{Z}$, are fixed.
 Randomness is incorporated through the random selection of sampling units based on probabilities derived using the sampling design itself.
@@
 
-Other model based stuff:
-chan2020bayesian: Bayesian model-based approach.
-Based on splines: 
-Cicchitelli, G. and Montanari, G. E. (2012). “Model-Assisted Estimation of a Spatial Population
-Mean.” International Statistical Review, 80(1): 111–126.
 <!-- \citep{Cressie2015statistics, Schabenberger2017statistical} -->
@@
 But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.
+
+We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
+However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
 
@@
 We note that while there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-CITE some more f these), our goal is not to expand upon or improve these methods.
+@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, CITE some more of these), our goal is not to expand upon or improve these methods.
 
@@
 \begin{align}\label{eq:htest}
-  \hat{\tau}_{HT} = \sum_{i = 1}^n Y_i \pi_i^{-1},
+  \hat{\tau}_{HT} = \sum_{i = 1}^n Z_i \pi_i^{-1},
 \end{align}
-where $Y_i$ is the observed value of the $i$th sampling unit, and $\pi_i$ is the inclusion probability of the $i$th sampling unit.
+where $Z_i$ is the observed value of the $i$th sampling unit, and $\pi_i$ is the inclusion probability of the $i$th sampling unit.
 Estimating the variance of $\hat{\tau}_{HT}$ directly relies on knowing the probability that $s_i$ and $s_j$ are both included in the sample (second-order inclusion probability), which can be unfeasible to compute.

### 2021-06-08 R/sim_one.R
simulation updates

@@ -1,5 +1,6 @@
 ## get one simulation
-sim_one <- function(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill, erange, nugget, cortype_est = "Exponential", ...) {
+sim_one <- function(seed, N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill, erange, nugget, cortype_est = "Exponential", ...) {
+  set.seed(seed)
   data <- sim_pop(N, n, gridded, cortype, psill, erange, nugget, ...)
   irs_samp <- sample_n(data, n)
   irs_unsamp <- anti_join(data, irs_samp)
@@ -65,6 +66,7 @@ sim_one <- function(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", ps
   # store output
   output <- data.frame(
     approach = c("Design", "Model"),
+    seed = c(seed, seed),
     true_mean = c(realized_mean, realized_mean),
     true_var = c(realized_var, realized_var),
     estimate = c(design_mean$Estimate, as.vector(model_mean)),
@@ -79,8 +81,4 @@ sim_one <- function(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", ps
 
 
 ## example
-# df <- sim_pop(N = 100, n = 50, gridded = TRUE, cortype = "exponential",
-#               psill = 1.1, erange = 1.01, nugget = 0.3)
-# library(dplyr)
-# sim_one(data = df, xcoord = "x", ycoord = "y", response = "response",
-#         cortype_est = "exponential")
+## data <- sim_one(psill = 1, erange = 1, nugget = 1)

### 2021-06-08 R/sim_pop.R

@@ -1,5 +1,5 @@
 ## simulate population
-# need spsurvey and sptotal
+# need spsurvey and sptotal and dplyr
 sim_pop <- function(N = 100, n = 50, gridded = TRUE, cortype, psill, erange, nugget, ...) {
 
   # simulating the locations

### 2021-06-08 R/sim_one.R
Updated description as well as documentation for functions in roxygen.

@@ -1,9 +1,48 @@
-## get one simulation
-sim_one <- function(seed, N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill, erange, nugget, cortype_est = "Exponential", ...) {
+#' Simulate spatially correlated data
+#'
+#' Take a sample from and subsequently analyze spatially correlated data
+#' using \enumerate{
+#'          \item GRTS with a spatially balanced sample
+#'          \item FPBK with a simple random sample
+#' }
+#'
+#' @param seed is a random seed, by default a random integer.
+#' @param N is the total number of data points.
+#'  For \code{gridded = TRUE}, this must be a perfect square.
+#' @param n is the number of data points sampled.
+#' @param gridded \code{TRUE} for gridded sites and \code{FALSE} for
+#' points with random locations.
+#' @param cortype is the true correlation function, which is
+#' \code{"Exponential"} by default.
+#' @param psill is the partial sill.
+#' @param erange is the effective range.
+#' @param nugget is the nugget.
+#' @param cortype_est is the (possibly misspecified) correlation
+#' function used to estimate/predict the mean, total, or other quantity.
+#' @param ... further arguments passed to or from other methods.
+#' @return a data frame with \itemize{
+#'   \item \code{approach}, the name of th approach (\code{"Degign"}
+#'   or \code{"Model"}).
+#'   \item \code{seed}, a non-random integer seed.
+#'   \item \code{true_mean}, the realized mean from the entire population.
+#'   \item \code{true_var}, the realized variance from the entire population.
+#'   \item \code{estimate}, the estimated/predicted mean.
+#'   \item \code{sd}, the standard error of the estimated/predicted mean.
+#'   \item \code{lb}, a lower 95% confidence bound
+#'   \item \code{ub}, an upper 95% confidence bound.
+#' }
+#' @examples
+#' sim_one(seed = sample.int(1e7, size = 1), N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1, nugget = 0.2, cortype_est = "Exponential")
+#' @export
+
+sim_one <- function(seed = sample.int(1e7, size = 1),
+                    N = 100, n = 50, gridded = TRUE,
+                    cortype = "Exponential", psill, erange,
+                    nugget, cortype_est = "Exponential", ...) {
[…]

### 2021-06-08 R/sim_pop.R

@@ -6,11 +6,10 @@
 #'
 #' @param N is the total number of data points. For \code{gridded = TRUE},
 #' this must be a perfect square.
-#' @param n is number of data points sampled.
 #' @param gridded \code{TRUE} for gridded sites and \code{FALSE} for
 #' points with random locations.
 #' @param cortype is the correlation function, which is
-#' \code{"Exponential"}
+#' \code{"Exponential"} by default.
 #' @param psill is the partial sill.
 #' @param erange is the effective range.
 #' @param nugget is the nugget.
@@ -21,12 +20,14 @@
 #'   \item \code{response}, a column with the response variable.
 #' }
 #' @examples
-#' sim_pop(N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1, nugget = 0.2)
+#' sim_pop(N = 100, gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1, nugget = 0.2)
 #' @export
 
 ## simulate population
 # need spsurvey and sptotal and dplyr
-sim_pop <- function(N = 100, n = 50, gridded = TRUE, cortype, psill, erange, nugget, ...) {
+sim_pop <- function(N = 100, gridded = TRUE,
+                    cortype = "Exponential", psill, erange,
+                    nugget, ...) {
 
   # simulating the locations
   if (gridded) {

### 2021-06-08 R/sim_one.R
fixes for r package checks notes.

@@ -32,7 +32,10 @@
 #'   \item \code{ub}, an upper 95% confidence bound.
 #' }
 #' @examples
-#' sim_one(seed = sample.int(1e7, size = 1), N = 100, n = 50, gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1, nugget = 0.2, cortype_est = "Exponential")
+#' sim_one(seed = sample.int(1e7, size = 1), N = 100, n = 50,
+#' gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1,
+#' nugget = 0.2, cortype_est = "Exponential")
+#' @import stats
 #' @export
 
 sim_one <- function(seed = sample.int(1e7, size = 1),
@@ -44,12 +47,13 @@ sim_one <- function(seed = sample.int(1e7, size = 1),
   irs_samp <- dplyr::sample_n(data, n)
   irs_unsamp <- dplyr::anti_join(data, irs_samp)
   irs_unsamp$response <- NA
-  full_df <- bind_rows(irs_samp, irs_unsamp)
+  full_df <- dplyr::bind_rows(irs_samp, irs_unsamp)
   full_df$wts <- 1 / nrow(full_df)
-  mod <- slmfit(formula = response ~ 1, data = full_df, xcoordcol = "x",
-                ycoordcol = "y", CorModel = cortype_est)
+  mod <- sptotal::slmfit(formula = response ~ 1,
+                         data = full_df, xcoordcol = "x",
+                         ycoordcol = "y", CorModel = cortype_est)
 
-  pred_mod <- predict(mod, wtscol = "wts")
+  pred_mod <- sptotal::predict(mod, wtscol = "wts")
   model_mean <- pred_mod$FPBK_Prediction
   model_se <- sqrt(pred_mod$PredVar)
   model_lb <- model_mean + -1 * 1.96 * model_se
@@ -58,13 +62,13 @@ sim_one <- function(seed = sample.int(1e7, size = 1),
   # browser()
   # take a sample
   ## convert data to sf object (for spsurvey)
-  data <- st_as_sf(data, coords = c("x", "y"), crs = 5070)
+  data <- sf::st_as_sf(data, coords = c("x", "y"), crs = 5070)
   ## select grts sample
-  grts_samp <- grts(data, n_base = n, ...)
+  grts_samp <- spsurvey::grts(data, n_base = n, ...)
   ## convert to usable
-  grts_bind <- sprbind(grts_samp, ...)
[…]

### 2021-06-08 R/sim_pop.R

@@ -80,7 +80,7 @@ covmx_exp <- function(distmx, psill, erange, nugget) {
 
 #' @describeIn covmx_exp Gaussian Covariance
 covmx_gauss <- function(distmx, psill, erange, nugget) {
-  psill * exp(-sqrt(3) * distance.matrix ^ 2 / erange) +
+  psill * exp(-sqrt(3) * distmx ^ 2 / erange) +
     nugget * (distmx == 0)
 }
 

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
simulation outline discussion points.
some introduction edits

@@
 
-# Introduction
+# Introduction {#sec:introduction}
 
 Please leave comments in your \textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}: \textcolor{blue}{Michael}, \textcolor{magenta}{Matt}, \textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
+
+<!-- Brief introduction to model based and design based -->
+
+There are two approaches to statistical inference: design-based and model-based.
+Both approaches aim to study a population.
+Because data generally cannot be sampled at all units in the population (known as population units), the population units must be sampled.
+In the design-based approach, the samples are often used to estimate parameters of a population.
+Properties of these estimators are derived through random selection of the samples.
+In the model-based approach, the samples are often used to predict random variables of a population.
+Properties of these predictions are derived through a set of distributional assumptions that specify the population's structure.
+Randomness is crucial to both approaches but for different reasons; the design-based approach uses randomness when selecting a sample of population units, and the model-based approach uses randomness when specifying distributional assumptions.
+@sterba2009alternative provides historical context for each approach, and @hansen1983evaluation discusses the advantages of each approach [@hansen1983evaluation].
+
+<!-- Shift focus to spatial data -->
+
+Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus comparing these approaches for spatial data.
+@de1990model gave an early comparison of these approaches, quashing the belief that design-based approaches could not be used for spatially correlated data.
+Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random, @verhoef2002sampling, @verhoef2008spatial].
+No comparison has been made design-based approaches that incorporate spatial location in sampling and model-based approaches.
 
@@
 
-@de1990model give an early comparison of the paradigms for spatial data, quashing the belief that classical sampling-based methods could not be used for spatially correlated data.
-Thereafter, several comparisons between model-based and design-based for spatial data have been considered, but they tend to compare spatial models to IRS (Independent Random Sampling) or SIRS (stratified IRS) [@brus1997random].
-No comparison has been made between spatial models and spatially balanced design-based approaches.
-
 brus2020statistical: most similar to what this paper will be, directly comparing the two approaches.
@@
 
-* hansen1983evaluation - general overview (in non spatial context), talk about population vs superpopulation give a set of guiding principles in the discussion
+wang2012spatial - sampling and inference steps, thorough overview in terms of spatial context, lots of references
 
-* verhoef2002sampling - strictly defines classical sampling, compares FPBK to classical, BK to classical
+<!-- Outline for the rest of the paper -->
 
-* wang2012spatial - sampling and inference steps, thorough overview in terms of spatial context, lots of references
+In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
+In Section \ref{sec:numstudy}, we use simulated and real data to compare the properties of parameter estimates from both aproaches.
+And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
 
-# Background
+# Background {#sec:background}
 
@@
 
-# Numerical Study
+# Numerical Study {#sec:numstudy}
+
+cases
+
+__Base Simulations__
+
+* both good: correctly specified model with high correlation
+* break model: non-gaussian errors
+* break design: small area estimation
+
+* both good?: misspecified covariance model with high correlation
+* break both? non-gaussian areas with smaller sample size
+
+* change n or sampling fraction
+
+* model-based: how should sample be drawn?
 
@@
 
-# Discussion
+# Discussion {#sec:discussion}
 

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
some introduction edits

@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\hypertarget{sec:introduction}{%
+\section{Introduction}\label{sec:introduction}}
 
@@
 
-Several comparisons between model-based and design-based for spatial
-data have been considered, but they tend to compare spatial models to
-IRS (Independent Random Sampling) or SIRS (stratified IRS). No
-comparison has been made between spatial models and spatially balanced
-design-based approaches.
+There are two approaches to statistical inference for spatial data:
+design-based and model-based. In the design-based approach, statistical
+inference is based on random sampling. In the model-based approach,
+statistical inference is based on a set of distributional assumptions
+for the data.
 
-\hypertarget{background}{%
-\section{Background}\label{background}}
+In either approach, statistical inference requires estimating parameters
+of interest. Often, these parameters are means or totals. In the
+design-based approach, parameters correspond to some population. In the
+model-based approach, parameters correspond to some statistical process,
+sometimes called a superpopulation.
+
+Spatial data paragraph
+
+define sampling design, define parameters, define model-based, sample,
+population, super-population?, etc.
+
+introduction of spatial data / definition of spatial and maybe general
+background?
+
+give some examples of possible spatial data
+
+2-3 sentences about each design
+
+goals and outline of paper
+
+De Gruijter and Ter Braak (1990) give an early comparison of the
+paradigms for spatial data, quashing the belief that classical
+sampling-based methods could not be used for spatially correlated data.
+Thereafter, several comparisons between model-based and design-based for
+spatial data have been considered, but they tend to compare spatial
+models to IRS (Independent Random Sampling) or SIRS (stratified IRS)
+(Brus and De Gruijter, 1997). No comparison has been made between
+spatial models and spatially balanced design-based approaches.
+
+brus2020statistical: most similar to what this paper will be, directly
+comparing the two approaches. This has a lot of other potentially useful
+references to go back to. (published in 2021 so likely doesn't have many
+that have cited it yet).
+
+\begin{itemize}
+\item
+  hansen1983evaluation - general overview (in non spatial context), talk
+  about population vs superpopulation give a set of guiding principles
+  in the discussion
+\item
+  verhoef2002sampling - strictly defines classical sampling, compares
+  FPBK to classical, BK to classical
+\item
+  wang2012spatial - sampling and inference steps, thorough overview in
+  terms of spatial context, lots of references
+\end{itemize}
+
+In Section \ref{sec:background}
+
+\hypertarget{sec:background}{%
+\section{Background}\label{sec:background}}
 
@@
 sampling design to estimate parameters of interest. The design-based
-approach assumes the data, \(\mathbf{Y}\), are fixed. Randomness is
+approach assumes the data, \(\mathbf{Z}\), are fixed. Randomness is
 incorporated through the random selection of sampling units based on
@@
 
+We only use FPBK in this paper in order to focus more on comparing the
+design-based and model-based approaches. However, k-nearest-neighbors
+(Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013), random forest
+(Breiman, 2001), Bayesian models (Chan-Golston et al., 2020), among
+others, can also be used to obtain predictions for a mean or total from
+spatially correlated responses in a finite population setting.
+
 Figure 1a. Data is fixed. In a finite population example, show a 3d
@@
 
-Design-Based and Model-Based Comparisons
-
-There have been many comparisons between the two paradigms\ldots\ldots.
+Several comparisons between the two paradigms have been made in many
+different contexts. Sterba (2009) give some history of the two paradigms
+as well as some applications in psychological contexts. Cooper (2006)
+review the two approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. We note that while there has been substantial research and
+development into estimators that use both design and model-based
+principles (see e.g.~Cicchitelli and Montanari (2012), Chan-Golston et
+al. (2020) for a Bayesian approach, CITE some more of these), our goal
+is not to expand upon or improve these methods.
 
@@
 is given by \begin{align}\label{eq:htest}
-  \hat{\tau}_{HT} = \sum_{i = 1}^n Y_i \pi_i^{-1},
-\end{align} where \(Y_i\) is the observed value of the \(i\)th sampling
+  \hat{\tau}_{HT} = \sum_{i = 1}^n Z_i \pi_i^{-1},
+\end{align} where \(Z_i\) is the observed value of the \(i\)th sampling
 unit, and \(\pi_i\) is the inclusion probability of the \(i\)th sampling
@@
 
+cases
+
+\textbf{Base Simulations}
+
+\begin{itemize}
+\item
+  both good: correctly specified model with high correlation
+\item
+  break model: non-gaussian errors
+\item
+  break design: small area estimation
+\item
+  both good?: misspecified covariance model with high correlation
+\item
+  break both? non-gaussian areas with smaller sample size
+\item
+  change n or sampling fraction
+\item
+  model-based: how should sample be drawn?
+\end{itemize}
+
 \hypertarget{software}{%
@@
 
+\leavevmode\hypertarget{ref-breiman2001random}{}%
+Breiman, L., 2001. Random forests. Machine learning 45, 5--32.
+
+\leavevmode\hypertarget{ref-brus1997random}{}%
+Brus, D., De Gruijter, J., 1997. Random sampling or geostatistical
+modelling? Choosing between design-based and model-based sampling
+strategies for soil (with discussion). Geoderma 80, 1--44.
+
+\leavevmode\hypertarget{ref-chan2020bayesian}{}%
+Chan-Golston, A.M., Banerjee, S., Handcock, M.S., 2020. Bayesian
+inference for finite populations under spatial process settings.
+Environmetrics 31, e2606.
+
 \leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
@@
 Uncertainty. John Wiley \& Sons, New York.
+
+\leavevmode\hypertarget{ref-cicchitelli2012model}{}%
+Cicchitelli, G., Montanari, G.E., 2012. Model-assisted estimation of a
+spatial population mean. International Statistical Review 80, 111--126.
+
+\leavevmode\hypertarget{ref-cooper2006sampling}{}%
+Cooper, C., 2006. Sampling and variance estimation on continuous
+domains. Environmetrics: The official journal of the International
+Environmetrics Society 17, 539--553.
 
@@
 
+\leavevmode\hypertarget{ref-de1990model}{}%
+De Gruijter, J., Ter Braak, C., 1990. Model-free estimation from spatial
+samples: A reappraisal of classical sampling theory. Mathematical
+geology 22, 407--415.
+
 \leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
@@
 Submitted for Publication.
+
+\leavevmode\hypertarget{ref-fix1951discriminatory}{}%
+Fix, E., Hodges, J.L., 1951. Discriminatory analysis, nonparametric
+discrimination: Consistency properties. USAF School of Aviation
+Medicine.
 
@@
 
+\leavevmode\hypertarget{ref-sterba2009alternative}{}%
+Sterba, S.K., 2009. Alternative model-based and design-based frameworks
+for inference from samples to populations: From polarization to
+integration. Multivariate behavioral research 44, 711--740.
+
 \leavevmode\hypertarget{ref-stevens2003variance}{}%
@@
 
+\leavevmode\hypertarget{ref-ver2013comparison}{}%
+Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
+model to nearest neighbor (k-nn) methods for forestry applications. PloS
+one 8, e59129.
+
 \leavevmode\hypertarget{ref-wang2013design}{}%

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
introduction revision.

@@
 
+finish spatial introduction
+revise section 2
+reread brus and wang
+write down potential data sets
+
 Please leave comments in your \textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}: \textcolor{blue}{Michael}, \textcolor{magenta}{Matt}, \textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
@@
 
-There are two approaches to statistical inference: design-based and model-based.
-Both approaches aim to study a population.
-Because data generally cannot be sampled at all units in the population (known as population units), the population units must be sampled.
-In the design-based approach, the samples are often used to estimate parameters of a population.
-Properties of these estimators are derived through random selection of the samples.
-In the model-based approach, the samples are often used to predict random variables of a population.
-Properties of these predictions are derived through a set of distributional assumptions that specify the population's structure.
-Randomness is crucial to both approaches but for different reasons; the design-based approach uses randomness when selecting a sample of population units, and the model-based approach uses randomness when specifying distributional assumptions.
-@sterba2009alternative provides historical context for each approach, and @hansen1983evaluation discusses the advantages of each approach [@hansen1983evaluation].
+There are two general approaches for using data to make statistical inferences about a population: design-based approaches and model-based approaches.
+When data cannot be obtained for all units in a population (known as population units), data on a subset of the population units is collected in a  sample.
+In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
+Alternatively, in the model-based approach, inferences are made from specific assumptions made about the underlying process that generated the data.
+Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
 
@@
 
-Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus comparing these approaches for spatial data.
-@de1990model gave an early comparison of these approaches, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random, @verhoef2002sampling, @verhoef2008spatial].
-No comparison has been made design-based approaches that incorporate spatial location in sampling and model-based approaches.
+Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
 
-define sampling design, define parameters, define model-based, sample, population, super-population?, etc.
+<!-- ????????? -->
+We define spatial data as [blah blah blah, any data collected at specific locations blah blah blah blah]. Examples include [do we want to give a list of general examples?]. 
+<!-- ???????????? -->
 
-introduction of spatial data / definition of spatial and maybe general background?
-
-give some examples of possible spatial data
-
-2-3 sentences about each design
-
-goals and outline of paper
-
-brus2020statistical: most similar to what this paper will be, directly comparing the two approaches.
-This has a lot of other potentially useful references to go back to.
-(published in 2021 so likely doesn't have many that have cited it yet).
-
-wang2012spatial - sampling and inference steps, thorough overview in terms of spatial context, lots of references
+@de1990model give an early comparison of these approaches, quashing the belief that design-based approaches could not be used for spatially correlated data.
+Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
+More recent overviews include @brus2020statistical and @wang2012spatial, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.
 
@@
 
+The rest of this paper is organized as follows.
 In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
@@
 
+## Applied Example
+
 # Discussion {#sec:discussion}

### 2021-06-08 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
introduction revision.

@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-  \title{A comparison of model-based and design-based approaches for spatial
-data.}
+  \title{A comparison of model-based and design-based approaches for
+spatial data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Math, Computer Science, and
-Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Math,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
       \cortext[1]{Corresponding Author}
@@
 
-There are two approaches to statistical inference for spatial data:
-design-based and model-based. In the design-based approach, statistical
-inference is based on random sampling. In the model-based approach,
-statistical inference is based on a set of distributional assumptions
-for the data.
+There are two general approaches for using data to make statistical
+inferences about a population: design-based approaches and model-based
+approaches. When data cannot be obtained for all units in a population
+(known as population units), data on a subset of the population units is
+collected in a sample. In the design-based approach, inferences about
+the underlying population are informed from a probabilistic process in
+which population units are selected to be in the sample. Alternatively,
+in the model-based approach, inferences are made from specific
+assumptions made about the underlying process that generated the data.
+Each paradigm has a deep historical context (Sterba, 2009) and its own
+set of general advantages (Hansen et al., 1983).
 
-In either approach, statistical inference requires estimating parameters
-of interest. Often, these parameters are means or totals. In the
-design-based approach, parameters correspond to some population. In the
-model-based approach, parameters correspond to some statistical process,
-sometimes called a superpopulation.
+Though the design-based and model-based approaches apply to statistical
+inference in a broad sense, we focus on comparing these approaches for
+spatial data.
 
-Spatial data paragraph
+We define spatial data as {[}blah blah blah, any data collected at
+specific locations blah blah blah blah{]}. Examples include {[}do we
+want to give a list of general examples?{]}.
 
-define sampling design, define parameters, define model-based, sample,
-population, super-population?, etc.
+De Gruijter and Ter Braak (1990) give an early comparison of these
+approaches, quashing the belief that design-based approaches could not
+be used for spatially correlated data. Thereafter, several comparisons
+between design-based and model-based for spatial data have been
+considered, but they tend to compare design-based approaches that ignore
+spatial location in sampling to model-based approaches (Brus and De
+Gruijter, 1997; Brus, 2020; Ver Hoef, 2002; Ver Hoef, 2008). No
+comparison has been made between design-based approaches that
+incorporate spatial location in sampling and model-based approaches.
 
-introduction of spatial data / definition of spatial and maybe general
-background?
-
-give some examples of possible spatial data
-
-2-3 sentences about each design
-
-goals and outline of paper
-
-De Gruijter and Ter Braak (1990) give an early comparison of the
-paradigms for spatial data, quashing the belief that classical
-sampling-based methods could not be used for spatially correlated data.
-Thereafter, several comparisons between model-based and design-based for
-spatial data have been considered, but they tend to compare spatial
-models to IRS (Independent Random Sampling) or SIRS (stratified IRS)
-(Brus and De Gruijter, 1997). No comparison has been made between
-spatial models and spatially balanced design-based approaches.
-
-brus2020statistical: most similar to what this paper will be, directly
-comparing the two approaches. This has a lot of other potentially useful
-references to go back to. (published in 2021 so likely doesn't have many
-that have cited it yet).
-
-\begin{itemize}
-\item
-  hansen1983evaluation - general overview (in non spatial context), talk
-  about population vs superpopulation give a set of guiding principles
-  in the discussion
-\item
-  verhoef2002sampling - strictly defines classical sampling, compares
-  FPBK to classical, BK to classical
-\item
 wang2012spatial - sampling and inference steps, thorough overview in
 terms of spatial context, lots of references
-\end{itemize}
 
-In Section \ref{sec:background}
+In Section \ref{sec:background}, we compare sampling and estimation
+procedures between the design-based approach and the model-based
+approach. In Section \ref{sec:numstudy}, we use simulated and real data
+to compare the properties of parameter estimates from both aproaches.
+And in Section \ref{sec:discussion}, we end with a discussion and
+provide directions for future research.
 
@@
 development into estimators that use both design and model-based
-principles (see e.g.~Cicchitelli and Montanari (2012), Chan-Golston et
+principles (see e.g. Cicchitelli and Montanari (2012), Chan-Golston et
 al. (2020) for a Bayesian approach, CITE some more of these), our goal
@@
 
-\hypertarget{numerical-study}{%
-\section{Numerical Study}\label{numerical-study}}
+\hypertarget{sec:numstudy}{%
+\section{Numerical Study}\label{sec:numstudy}}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\hypertarget{sec:discussion}{%
+\section{Discussion}\label{sec:discussion}}
 
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-barabesi2011sampling}{}%
@@
 
+\leavevmode\hypertarget{ref-brus2020statistical}{}%
+Brus, D.J., 2020. Statistical approaches for spatial sample survey:
+Persistent misconceptions and new developments. European Journal of Soil
+Science.
+
 \leavevmode\hypertarget{ref-chan2020bayesian}{}%
@@
 \leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
-Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
-Uncertainty. John Wiley \& Sons, New York.
+Chiles, J.-P., Delfiner, P., 1999. Geostatistics: {Modeling Spatial
+Uncertainty}. {John Wiley \& Sons}, New York.
 
@@
 
+\leavevmode\hypertarget{ref-hansen1983evaluation}{}%
+Hansen, M.H., Madow, W.G., Tepping, B.J., 1983. An evaluation of
+model-dependent and probability-sampling inferences in sample surveys.
+Journal of the American Statistical Association 78, 776--793.
+
 \leavevmode\hypertarget{ref-higham2020sptotal}{}%
@@
 
+\leavevmode\hypertarget{ref-verhoef2002sampling}{}%
+Ver Hoef, J., 2002. Sampling and geostatistics for spatial data.
+Ecoscience 9, 152--161.
+
 \leavevmode\hypertarget{ref-verhoef2008spatial}{}%
@@
 Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
-model to nearest neighbor (k-nn) methods for forestry applications. PloS
+model to nearest neighbor (k-NN) methods for forestry applications. PloS
 one 8, e59129.
@@
 
+\end{CSLReferences}
+
 \end{document}

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated introduction.

@@
 Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-
-<!-- ????????? -->
-We define spatial data as [blah blah blah, any data collected at specific locations blah blah blah blah]. Examples include [do we want to give a list of general examples?]. 
-<!-- ???????????? -->
-
-@de1990model give an early comparison of these approaches, quashing the belief that design-based approaches could not be used for spatially correlated data.
+We define spatial data as variables measured at specific geographic locations.
+@de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
 Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated introduction.

@@
 
+finish spatial introduction revise section 2 reread brus and wang write
+down potential data sets
+
 Please leave comments in your
@@
 inference in a broad sense, we focus on comparing these approaches for
-spatial data.
+spatial data. We define spatial data as variables measured at specific
+geographic locations. De Gruijter and Ter Braak (1990) give an early
+comparison of design-based and model-based approaches for spatial data,
+quashing the belief that design-based approaches could not be used for
+spatially correlated data. Thereafter, several comparisons between
+design-based and model-based for spatial data have been considered, but
+they tend to compare design-based approaches that ignore spatial
+location in sampling to model-based approaches (Brus and De Gruijter,
+1997; Ver Hoef, 2002; Ver Hoef, 2008). More recent overviews include
+Brus (2020) and (\textbf{wang2012spatial?}), but no numerical comparison
+has been made between design-based approaches that incorporate spatial
+location in sampling and model-based approaches.
 
-We define spatial data as {[}blah blah blah, any data collected at
-specific locations blah blah blah blah{]}. Examples include {[}do we
-want to give a list of general examples?{]}.
-
-De Gruijter and Ter Braak (1990) give an early comparison of these
-approaches, quashing the belief that design-based approaches could not
-be used for spatially correlated data. Thereafter, several comparisons
-between design-based and model-based for spatial data have been
-considered, but they tend to compare design-based approaches that ignore
-spatial location in sampling to model-based approaches (Brus and De
-Gruijter, 1997; Brus, 2020; Ver Hoef, 2002; Ver Hoef, 2008). No
-comparison has been made between design-based approaches that
-incorporate spatial location in sampling and model-based approaches.
-
-wang2012spatial - sampling and inference steps, thorough overview in
-terms of spatial context, lots of references
-
-In Section \ref{sec:background}, we compare sampling and estimation
-procedures between the design-based approach and the model-based
-approach. In Section \ref{sec:numstudy}, we use simulated and real data
-to compare the properties of parameter estimates from both aproaches.
-And in Section \ref{sec:discussion}, we end with a discussion and
-provide directions for future research.
+The rest of this paper is organized as follows. In Section
+\ref{sec:background}, we compare sampling and estimation procedures
+between the design-based approach and the model-based approach. In
+Section \ref{sec:numstudy}, we use simulated and real data to compare
+the properties of parameter estimates from both aproaches. And in
+Section \ref{sec:discussion}, we end with a discussion and provide
+directions for future research.
 
@@
 
+\hypertarget{applied-example}{%
+\subsection{Applied Example}\label{applied-example}}
+
 \hypertarget{sec:discussion}{%

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
adjusted final intro paragraph
update to title

@@
 ---
-title: A comparison of model-based and design-based approaches for spatial data.
+title: A comparison of design-based and model-based approaches for spatial data.
 author:
@@
 In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
-In Section \ref{sec:numstudy}, we use simulated and real data to compare the properties of parameter estimates from both aproaches.
+In Section \ref{sec:numstudy}, we use simulated and real data to study the properties of parameter estimates from both approaches.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
update to title

@@
 
-  \title{A comparison of model-based and design-based approaches for
-spatial data.}
+  \title{A comparison of model-based and design-based approaches for spatial
+data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200
-SW 35th St, Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Math,
-Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
-13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239
-Weniger Hall, Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
-Center, National Oceanic and Atmospheric Administration, Seattle,
-Washington, 98115}
+      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
+Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Math, Computer Science, and
+Statistics, 23 Romoda Drive, Canton, New York, 13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
+Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
+Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
       \cortext[1]{Corresponding Author}
@@
 location in sampling to model-based approaches (Brus and De Gruijter,
-1997; Ver Hoef, 2002; Ver Hoef, 2008). More recent overviews include
-Brus (2020) and (\textbf{wang2012spatial?}), but no numerical comparison
-has been made between design-based approaches that incorporate spatial
-location in sampling and model-based approaches.
+1997; Ver Hoef, 2002, 2008). More recent overviews include Brus (2020)
+and ({\textbf{???}}), but no numerical comparison has been made between
+design-based approaches that incorporate spatial location in sampling
+and model-based approaches.
 
@@
 between the design-based approach and the model-based approach. In
-Section \ref{sec:numstudy}, we use simulated and real data to compare
-the properties of parameter estimates from both aproaches. And in
-Section \ref{sec:discussion}, we end with a discussion and provide
-directions for future research.
+Section \ref{sec:numstudy}, we use simulated and real data to study the
+properties of parameter estimates from both approaches. And in Section
+\ref{sec:discussion}, we end with a discussion and provide directions
+for future research.
 
@@
 development into estimators that use both design and model-based
-principles (see e.g. Cicchitelli and Montanari (2012), Chan-Golston et
+principles (see e.g.~Cicchitelli and Montanari (2012), Chan-Golston et
 al. (2020) for a Bayesian approach, CITE some more of these), our goal
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-barabesi2011sampling}{}%
@@
 \leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
-Chiles, J.-P., Delfiner, P., 1999. Geostatistics: {Modeling Spatial
-Uncertainty}. {John Wiley \& Sons}, New York.
+Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
+Uncertainty. John Wiley \& Sons, New York.
 
@@
 Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
-model to nearest neighbor (k-NN) methods for forestry applications. PloS
+model to nearest neighbor (k-nn) methods for forestry applications. PloS
 one 8, e59129.
@@
 
-\end{CSLReferences}
-
 \end{document}

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
fixed wang 2012 citation

@@
 Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-More recent overviews include @brus2020statistical and @wang2012spatial, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.
+More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
re-organized section.
first paragraph of background revision.

@@
 
-## Design-Based Philosphy
+The design-based and model-based approaches incorporate randomness in fundamentally different ways.
+In this section, we describe the role of randomness and its effects on subsequent inferences.
+We then discuss specific inference methods for the design-based and model-based approaches for spatial data.
 
-Design-Based Overview
+## Comparing Design-Based vs. Model-Based
 
-<!-- \citep{Sarndal2003model, Lohr2009sampling} -->
+### Design-Based Philosphy
 
-The design-based approach to inference uses characteristics of the sampling design to estimate parameters of interest.
-The design-based approach assumes the data, $\mathbf{Z}$, are fixed.
-Randomness is incorporated through the random selection of sampling units based on probabilities derived using the sampling design itself.
+The design-based approach assumes the data are fixed.
+Randomness is incorporated in the selection of population units according to a sampling design.
+A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit.
 Some examples of commonly used sampling designs include independent random sampling (IRS), stratified random sampling, and cluster sampling.
-Viewing the data as fixed and incorporating randomness through the sampling design yields estimators having very few assumptions.
-Asymptotic properties of these types of estimators are derived using limiting arguments.
-Means and totals, for example, are asymptotically Gaussian.
+The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
+These population parameters are typically assumed to be fixed but unknown.
+
+Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions.
+Confidence intervals for these types of estimators are typically derived using limiting arguments.
+Means and totals, for example, are asymptotically normally distributed by the Central Limit Theorem.
 @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
 
-## Model-Based Philosphy
-
-Model-Based Overview
+### Model-Based Philosphy
 
@@
 On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimators if the additional assumptions hold.
-Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
+Instead of estimating fixed but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
 This is a fundamental philosophical difference between sampling-based and model-based approaches.
@@
 
-We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
+Several comparisons between the two paradigms have been made in many different contexts.
+@sterba2009alternative give some history of the two paradigms as well as some applications in psychological contexts.
+@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
+We note that while there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
+@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, CITE some more of these), our goal is not to expand upon or improve these methods.
 
 Figure 1a.
-Data is fixed.
+Brus (2020): Data is fixed.
 In a finite population example, show a 3d surface that can be generated by anything.
@@
 The realized mean changes from surface to surface and it's not necessarily the case that 95% of all 95% PIs will contain the true, underlying mean.
-
-## Comparing Design-Based vs. Model-Based
-
-Several comparisons between the two paradigms have been made in many different contexts.
-@sterba2009alternative give some history of the two paradigms as well as some applications in psychological contexts.
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-We note that while there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, CITE some more of these), our goal is not to expand upon or improve these methods.
 
@@
 ## Finite Population Block Kriging
+
+We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
+However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
 
@@
 
-* model-based: how should sample be drawn?
+* model-based: how should sample be drawn? should locations be fixed?
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
model-based background revision.

@@
 
-On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimators if the additional assumptions hold.
-Instead of estimating fixed but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
-This is a fundamental philosophical difference between sampling-based and model-based approaches.
+The model-based approach assumes the data are a random realization of a process.
+Randomness is often incorporated through distributional assumptions on the data-generating process.
+Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
+For example, suppose the realized mean of all population units is the quantity of interest.
 Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty.
-But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.
+We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
+But the true mean of the spatial process that generated our realized data is still not known.
+When predicting the realized mean, we typically are not interested in the underlying process' true mean.
+
+Assuming the data is a realization of a specific data-generating process yields predictors that are linked to distributional assumptions.
+These distributional assumptions are used to derive prediction intervals.
+The distributional assumptions allow the prediction intervals can be more precise.
+@Cressie2015statistics and @Schabenberger2017statistical provide reviews of model-based approaches for spatial data.
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
draft of figure outline.

@@
 
-Several comparisons between the two paradigms have been made in many different contexts.
+Several comparisons between the model-based paradigm and the design-based paradigm have been made in many different contexts.
 @sterba2009alternative give some history of the two paradigms as well as some applications in psychological contexts.
@@
 If we repeatedly sample the surface, then 95% of all 95% CIs will contain the true mean, which never changes.
+
+```{r}
+library(tidyverse)
+library(here)
+
+set.seed(06092021)
+
+source(here("R", "sim_pop.R"))
+N <- 25
+pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
+                  psill = 0.9, erange = 1, nugget = 0.1)
+
+n <- 5
+samp_df <- pop_df %>% sample_n(n) 
+samp_df2 <- pop_df %>% sample_n(n)
+
+library(cowplot)
+d1 <- ggplot(data = pop_df, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df, size = 5.5, shape = 1, 
+             show.legend = FALSE, stroke = 1.5) +
+  theme_minimal_grid() +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Design 1")
+d1
+
+d2 <- ggplot(data = pop_df, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df2, size = 5.5, shape = 1, show.legend = FALSE,
+             stroke = 1.5) +
+  theme_minimal_grid() +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Design 2")
+
+pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
+                      psill = 0.9, erange = 1, nugget = 0.1)
+pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
+                       psill = 0.9, erange = 1, nugget = 0.1)
+indeces <- sample(1:N, size = n, replace = FALSE)
+samp_df_mod <- pop_df_mod %>% slice(indeces)
+samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
+
+m1 <- ggplot(data = pop_df_mod, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
+  theme_minimal_grid() +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Model 1")
+
+m2 <- ggplot(data = pop_df_mod2, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
+  theme_minimal_grid() +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  ggtitle("Model 2")
+
+library(gridExtra)
+grid.arrange(d1, d2, m1, m2)
+```
 

### 2021-06-09 R/sim_pop.R

@@ -43,8 +43,8 @@ sim_pop <- function(N = 100, gridded = TRUE,
     data <- expand.grid(x = x, y = y)
   } else {
     ## simulate square
-    x <- rnorm(N)
-    y <- rnorm(N)
+    x <- runif(N)
+    y <- runif(N)
     data <- data.frame(x = x, y = y)
   }
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor figure update and chunk option fixes.

@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
 library(tidyverse)
@@
   ggtitle("Design 1")
-d1
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor figure update and chunk option fixes.

--- 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
@@
             pdfauthor={},
-            pdftitle={A comparison of model-based and design-based approaches for spatial data.},
+            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
             colorlinks=false,
@@
 
-  \title{A comparison of model-based and design-based approaches for
+  \title{A comparison of design-based and model-based approaches for
 spatial data.}
@@
 1997; Ver Hoef, 2002; Ver Hoef, 2008). More recent overviews include
-Brus (2020) and (\textbf{wang2012spatial?}), but no numerical comparison
-has been made between design-based approaches that incorporate spatial
-location in sampling and model-based approaches.
+Brus (2020) and Wang et al. (2012), but no numerical comparison has been
+made between design-based approaches that incorporate spatial location
+in sampling and model-based approaches.
 
@@
 
+The design-based and model-based approaches incorporate randomness in
+fundamentally different ways. In this section, we describe the role of
+randomness and its effects on subsequent inferences. We then discuss
+specific inference methods for the design-based and model-based
+approaches for spatial data.
+
+\hypertarget{comparing-design-based-vs.-model-based}{%
+\subsection{Comparing Design-Based
+vs.~Model-Based}\label{comparing-design-based-vs.-model-based}}
+
 \hypertarget{design-based-philosphy}{%
-\subsection{Design-Based Philosphy}\label{design-based-philosphy}}
+\subsubsection{Design-Based Philosphy}\label{design-based-philosphy}}
 
-Design-Based Overview
+The design-based approach assumes the data are fixed. Randomness is
+incorporated in the selection of population units according to a
+sampling design. A sampling design assigns a positive probability of
+inclusion in the sample (inclusion probability) to each population unit.
+Some examples of commonly used sampling designs include independent
+random sampling (IRS), stratified random sampling, and cluster sampling.
+The goal is to use the sampling design and the sampled data to estimate
+population parameters like means and totals. These population parameters
+are typically assumed to be fixed but unknown.
 
-The design-based approach to inference uses characteristics of the
-sampling design to estimate parameters of interest. The design-based
-approach assumes the data, \(\mathbf{Z}\), are fixed. Randomness is
-incorporated through the random selection of sampling units based on
-probabilities derived using the sampling design itself. Some examples of
-commonly used sampling designs include independent random sampling
-(IRS), stratified random sampling, and cluster sampling. Viewing the
-data as fixed and incorporating randomness through the sampling design
-yields estimators having very few assumptions. Asymptotic properties of
-these types of estimators are derived using limiting arguments. Means
-and totals, for example, are asymptotically Gaussian. Särndal et al.
-(2003) and Lohr (2009) provide thorough reviews of the design-based
-approach.
+Treating the data as fixed and incorporating randomness through the
+sampling design yields estimators having very few other assumptions.
+Confidence intervals for these types of estimators are typically derived
+using limiting arguments. Means and totals, for example, are
+asymptotically normally distributed by the Central Limit Theorem.
+Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
+design-based approach.
 
 \hypertarget{model-based-philosphy}{%
-\subsection{Model-Based Philosphy}\label{model-based-philosphy}}
+\subsubsection{Model-Based Philosphy}\label{model-based-philosphy}}
 
-Model-Based Overview
+The model-based approach assumes the data are a random realization of a
+process. Randomness is often incorporated through distributional
+assumptions on the data-generating process. Instead of estimating fixed
+but unknown parameters (as in the design-based approach), the goal of
+model-based inference in the spatial context is often \emph{prediction}
+of an unknown quantity. For example, suppose the realized mean of all
+population units is the quantity of interest. Instead of
+\emph{estimating} a fixed unknown mean, we are \emph{predicting} the
+value of the mean, a random variable. We know that if we sampled all
+population units, we would have an exact prediction for the mean of our
+one realized process, without any uncertainty. But the true mean of the
+spatial process that generated our realized data is still not known.
+When predicting the realized mean, we typically are not interested in
+the underlying process' true mean.
 
-On the other hand, model-based inference imposes additional assumptions
-on the data with a potential to provide more precise estimators if the
-additional assumptions hold. Instead of estimating true but unknown
-parameters, the goal of model-based inference in the spatial context is
-often \emph{prediction} of an unknown quantity. This is a fundamental
-philosophical difference between sampling-based and model-based
-approaches. Instead of \emph{estimating} a fixed unknown mean, we are
-\emph{predicting} the value of the mean, a random variable. We know that
-if we sampled all sites, we would have an exact prediction for the mean
-of our one realized spatial surface, without any uncertainty. But, the
-true mean of the spatial process that generated our realized data is
-still not known, and, in the prediction context, we typically do not
-care much about what value the mean of the underlying process takes.
+Assuming the data is a realization of a specific data-generating process
+yields predictors that are linked to distributional assumptions. These
+distributional assumptions are used to derive prediction intervals. The
+distributional assumptions allow the prediction intervals can be more
+precise. (\textbf{Cressie2015statistics?}) and
+(\textbf{Schabenberger2017statistical?}) provide reviews of model-based
+approaches for spatial data.
 
-We only use FPBK in this paper in order to focus more on comparing the
-design-based and model-based approaches. However, k-nearest-neighbors
-(Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013), random forest
-(Breiman, 2001), Bayesian models (Chan-Golston et al., 2020), among
-others, can also be used to obtain predictions for a mean or total from
-spatially correlated responses in a finite population setting.
+Several comparisons between the model-based paradigm and the
+design-based paradigm have been made in many different contexts. Sterba
+(2009) give some history of the two paradigms as well as some
+applications in psychological contexts. Cooper (2006) review the two
+approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. We note that while there has been substantial research and
+development into estimators that use both design and model-based
+principles (see e.g. Cicchitelli and Montanari (2012), Chan-Golston et
+al. (2020) for a Bayesian approach, CITE some more of these), our goal
+is not to expand upon or improve these methods.
 
-Figure 1a. Data is fixed. In a finite population example, show a 3d
-surface that can be generated by anything. If we repeatedly sample the
-surface, then 95\% of all 95\% CIs will contain the true mean, which
-never changes.
+Figure 1a. Brus (2020): Data is fixed. In a finite population example,
+show a 3d surface that can be generated by anything. If we repeatedly
+sample the surface, then 95\% of all 95\% CIs will contain the true
+mean, which never changes.
+
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-1-1}
 
@@
 contain the true, underlying mean.
-
-\hypertarget{comparing-design-based-vs.-model-based}{%
-\subsection{Comparing Design-Based
-vs.~Model-Based}\label{comparing-design-based-vs.-model-based}}
-
-Several comparisons between the two paradigms have been made in many
-different contexts. Sterba (2009) give some history of the two paradigms
-as well as some applications in psychological contexts. Cooper (2006)
-review the two approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. We note that while there has been substantial research and
-development into estimators that use both design and model-based
-principles (see e.g. Cicchitelli and Montanari (2012), Chan-Golston et
-al. (2020) for a Bayesian approach, CITE some more of these), our goal
-is not to expand upon or improve these methods.
 
@@
 
+We only use FPBK in this paper in order to focus more on comparing the
+design-based and model-based approaches. However, k-nearest-neighbors
+(Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013), random forest
+(Breiman, 2001), Bayesian models (Chan-Golston et al., 2020), among
+others, can also be used to obtain predictions for a mean or total from
+spatially correlated responses in a finite population setting.
+
 Finite Population Block Kriging (FPBK) is an alternative to
@@
 \item
-  model-based: how should sample be drawn?
+  model-based: how should sample be drawn? should locations be fixed?
 \end{itemize}
@@
 
+\leavevmode\hypertarget{ref-wang2012review}{}%
+Wang, J.-F., Stein, A., Gao, B.-B., Ge, Y., 2012. A review of spatial
+sampling. Spatial Statistics 2, 1--14.
+
 \end{CSLReferences}

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
adjusted grts section and added simulation one table

@@
 # Introduction {#sec:introduction}
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>", 
+  warning = FALSE, 
+  message = FALSE,
+  include = TRUE, 
+  echo = FALSE
+)
+library(tidyverse)
+library(here)
+library(xtable) # latex tables
+```
 
@@
 
-<!-- \citep{Cressie2015statistics, Schabenberger2017statistical} -->
-
 The model-based approach assumes the data are a random realization of a process.
@@
 The distributional assumptions allow the prediction intervals can be more precise.
-@Cressie2015statistics and @Schabenberger2017statistical provide reviews of model-based approaches for spatial data.
+@cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
 
@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%"}
-library(tidyverse)
-library(here)
-
+```{r, out.width = "85%"}
 set.seed(06092021)
@@
 
-Spatially balanced sampling algorithms use spatial information to obtain samples spread out in space.
-Spatially balanced samples are useful because they tend to yield estimators that are more precise than estimators constructed from a sampling algorithm that is not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
-To quantify spatial balance, @stevens2004spatially proposed statistics based on Voroni polygons.
-Many spatially balanced sampling algorithms exist, including the Generalized Random Tessellation Stratified [@stevens2004spatially], the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
-Here we focus on the Generalized Random Tessellation Stratified (GRTS) algorithm, which has several attractive properties that we discuss next.
+Sampling designs can incorporate spatial locations to obtain samples that are spatially balanced with respect to the population [@stevens2004spatially].
+A sample is spatially with respect to the population if the sampled population units are a miniature of the population units.
+A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
+Spatial balance with respect to the population is is different than spatial balance with respect to geography.
+A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and are not meant to be miniatures of the population.
+When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
 
-The GRTS algorithm is used to sample from finite and infinite sample frames.
-A finite sample frame contains a finite number of sampling units and is related to a point geometry.
-An infinite sample frame contains an infinite number of sampling units and is related to linear and polygon geometries.
-Examples of point, linear, and polygon resources include lake centroids, stream networks, and wetland areas, respectively.
-In addition to its applicability for finite and infinite sample frames, the GRTS algorithm naturally accommodates stratified designs and designs with unequal selection probabilities.
-The algorithm has also been used to select replacement sites using reverse hierarchical ordering [@stevens2004spatially].
-Replacement sites are used to replace sites in the original sample that cannot be sampled, often as a result of physical difficulty in reaching the site or landowner denial of access to the sites.
-More recently, the GRTS algorithm also accommodates legacy (historical) sites, minimum distance between sites, and nearest neighbor replacement sites.
-The GRTS algorithm is implemented in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
+Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
+To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
+The first spatiallly balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
+Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
+We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because the algorithm has several attractive properties detailed by @stevens2004spatially and @dumelle2021spsurvey.
 
-The GRTS algorithm works by mapping two-dimensional space into one-dimensional space while incorporating some elements of randomness.
-First a square bounding box is placed over the sample frame and divided into four equally sized squares randomly given an level-one address of 0, 1, 2, or 3.
-Then the inclusion probabilities of the sampling units in each square are summed.
-If this sum is at least one in any of the cells, a set of four sub-squares are placed within each square.
-Each set of four sub-squares are randomly given a second address (a level-two address) of 0, 1, 2, 3.
-Then the inclusion probabilities in each sub-square are summed, and if any of these sums are at least one, a third level of squares are added in the same manner.
-This hierarchical addressing continues until the inclusion probabilities in the smallest squares are all less than 1.
-The squares are then mapped to the one-dimensional line in order of their addresses.
-The length of this line equals $n$, the desired sample size (the sum of the inclusion probabilities).
-A uniform random variable, denoted by $u_1$ is simulated in $[0, 1]$ and placed on this line.
-The location of $u_1$ on the line can be linked to a sampling unit, denoted by $s_1$.
-Then $u_2 \equiv u_1 + 1$, which can be linked to another sampling unit, denoted by $s_2$.
-Because the GRTS algorithm requires the inclusion probabilities of the smallest sub-cells to be less than one, $s_1$ and $s_2$ must be distinct.
-This process continues until the $n$ sampling units have been selected as part of the sample.
-Further details are provided by [@stevens2004spatially] and [@dumelle2021spsurvey].
+The GRTS algorithm is used to sample from finite and infinite populations and works by utilizing a mapping between two-dimensional and one-dimensional space.
+The population units in two-dimensional space are divided into cells using a hierarchical index.
+Population units are then mapped to a one-dimensional line via the hierarchical indexing.
+The line length of each population unit equals its inclusion probability.
+A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample size.
+@stevens2004spatially provide and @dumelle2021spsurvey provide further details.
+The GRTS algorithm is available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
 
-The Horvitz-Thompson estimator [@horvitz1952generalization] and its continuous analog [@cordy1993extension] yield unbiased estimates of population totals (and means).
-If $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$, denoted by $\hat{\tau}_{HT}$, is given by
-\begin{align}\label{eq:htest}
-  \hat{\tau}_{HT} = \sum_{i = 1}^n Z_i \pi_i^{-1},
+After collecting a sample, the data are used to estimate population parameters.
+The Horvitz-Thompson estimator [@horvitz1952generalization] and its continuous analog [@cordy1993extension] yield unbiased estimates of population means and totals.
+If $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
+\begin{align}\label{eq:ht}
+  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
 \end{align}
-where $Z_i$ is the observed value of the $i$th sampling unit, and $\pi_i$ is the inclusion probability of the $i$th sampling unit.
-Estimating the variance of $\hat{\tau}_{HT}$ directly relies on knowing the probability that $s_i$ and $s_j$ are both included in the sample (second-order inclusion probability), which can be unfeasible to compute.
-Furthermore, this variance estimator does not incorporate spatial locations.
-To address these challenges, [@stevens2003variance] proposed an alternative variance estimator: the local neighborhood variance estimator.
-The local neighborhood variance estimator does not require second-order inclusion probabilities.
-It also incorporates spatial locations, which tends to reduce the estimated variance of compared $\hat{\tau}_{HT}$ to a variance estimator ignoring spatial location [@stevens2003variance].
+where $Z_i$ and $pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
+@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
+First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, which can be very difficult to calculate.
+Second, they ignore the spatial locations of the population units.
+To address these drawbacks, @stevens2003variance proposed a local neighborhood variance estimator.
+The local neighborhood variance estimator does not rely on $\pi_{ij}$, and it incorporates spatial locations by assigning higher weights to nearby observations.
+@stevens2003variance show this variance estimators tends to reduce the variability associated with estimating $\tau$.
+This yields confidence intervals for $\tau$ that are narrower than confidence intervals constructed from variance estimators ignoring spatial locations.
 
@@
 
-cases
+```{r, results = "asis"}
+sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
+colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
+sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
+print(sim_one_table, include.rownames = FALSE, comment = FALSE)
+```
 

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
adjusted grts section and added simulation one table

@@
 
-  \title{A comparison of design-based and model-based approaches for
-spatial data.}
+  \title{A comparison of design-based and model-based approaches for spatial
+data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 location in sampling to model-based approaches (Brus and De Gruijter,
-1997; Ver Hoef, 2002; Ver Hoef, 2008). More recent overviews include
-Brus (2020) and Wang et al. (2012), but no numerical comparison has been
-made between design-based approaches that incorporate spatial location
-in sampling and model-based approaches.
+1997; Ver Hoef, 2002, 2008). More recent overviews include Brus (2020)
+and Wang et al. (2012), but no numerical comparison has been made
+between design-based approaches that incorporate spatial location in
+sampling and model-based approaches.
 
@@
 distributional assumptions allow the prediction intervals can be more
-precise. (\textbf{Cressie2015statistics?}) and
-(\textbf{Schabenberger2017statistical?}) provide reviews of model-based
-approaches for spatial data.
+precise. Cressie (1993) and Schabenberger and Gotway (2017) provide
+reviews of model-based approaches for spatial data.
 
@@
 
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-1-1}
+\includegraphics[width=0.85\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-2-1}
 
@@
 
-Spatially balanced sampling algorithms use spatial information to obtain
-samples spread out in space. Spatially balanced samples are useful
-because they tend to yield estimators that are more precise than
-estimators constructed from a sampling algorithm that is not spatially
-balanced (Barabesi and Franceschi, 2011; Benedetti et al., 2017;
-Grafström and Lundström, 2013; Robertson et al., 2013; Stevens Jr and
-Olsen, 2004; Wang et al., 2013). To quantify spatial balance, Stevens Jr
-and Olsen (2004) proposed statistics based on Voroni polygons. Many
-spatially balanced sampling algorithms exist, including the Generalized
-Random Tessellation Stratified (Stevens Jr and Olsen, 2004), the Local
-Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
-Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
-Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
-(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
-(Robertson et al., 2018) algorithms. Here we focus on the Generalized
-Random Tessellation Stratified (GRTS) algorithm, which has several
-attractive properties that we discuss next.
+Sampling designs can incorporate spatial locations to obtain samples
+that are spatially balanced with respect to the population (Stevens Jr
+and Olsen, 2004). A sample is spatially with respect to the population
+if the sampled population units are a miniature of the population units.
+A sample is a miniature of the population if the distribution of the
+sampled population units mirrors the density of all population units.
+Spatial balance with respect to the population is is different than
+spatial balance with respect to geography. A sample that is spatially
+balanced with respect to geography is spread out in some type of
+equidistant manner over geographical space and are not meant to be
+miniatures of the population. When we refer to spatial balance
+henceforth, we mean spatial balance with respect to the population.
 
-The GRTS algorithm is used to sample from finite and infinite sample
-frames. A finite sample frame contains a finite number of sampling units
-and is related to a point geometry. An infinite sample frame contains an
-infinite number of sampling units and is related to linear and polygon
-geometries. Examples of point, linear, and polygon resources include
-lake centroids, stream networks, and wetland areas, respectively. In
-addition to its applicability for finite and infinite sample frames, the
-GRTS algorithm naturally accommodates stratified designs and designs
-with unequal selection probabilities. The algorithm has also been used
-to select replacement sites using reverse hierarchical ordering (Stevens
-Jr and Olsen, 2004). Replacement sites are used to replace sites in the
-original sample that cannot be sampled, often as a result of physical
-difficulty in reaching the site or landowner denial of access to the
-sites. More recently, the GRTS algorithm also accommodates legacy
-(historical) sites, minimum distance between sites, and nearest neighbor
-replacement sites. The GRTS algorithm is implemented in the
-\textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al., 2021).
+Spatially balanced samples are useful because they tend to yield
+estimates that have lower variance than estimates constructed from
+sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
+Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
+2013; Stevens Jr and Olsen, 2004; Wang et al., 2013). To quantify
+spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
+based on Voroni polygons. The first spatiallly balanced sampling
+algorithm that saw widespread use was the Generalized Random
+Tessellation Stratified (Stevens Jr and Olsen, 2004). Since GRTS was
+developed, several other spatially balanced sampling algorithms have
+emerged, including the Local Pivotal Method (Grafström et al., 2012;
+Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
+(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
+2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
+Halton Iterative Partitioning (Robertson et al., 2018) algorithms. We
+focus on the Generalized Random Tessellation Stratified (GRTS) algorithm
+to select spatially balanced sampling because the algorithm has several
+attractive properties detailed by Stevens Jr and Olsen (2004) and
+Dumelle et al. (2021).
 
-The GRTS algorithm works by mapping two-dimensional space into
-one-dimensional space while incorporating some elements of randomness.
-First a square bounding box is placed over the sample frame and divided
-into four equally sized squares randomly given an level-one address of
-0, 1, 2, or 3. Then the inclusion probabilities of the sampling units in
-each square are summed. If this sum is at least one in any of the cells,
-a set of four sub-squares are placed within each square. Each set of
-four sub-squares are randomly given a second address (a level-two
-address) of 0, 1, 2, 3. Then the inclusion probabilities in each
-sub-square are summed, and if any of these sums are at least one, a
-third level of squares are added in the same manner. This hierarchical
-addressing continues until the inclusion probabilities in the smallest
-squares are all less than 1. The squares are then mapped to the
-one-dimensional line in order of their addresses. The length of this
-line equals \(n\), the desired sample size (the sum of the inclusion
-probabilities). A uniform random variable, denoted by \(u_1\) is
-simulated in \([0, 1]\) and placed on this line. The location of \(u_1\)
-on the line can be linked to a sampling unit, denoted by \(s_1\). Then
-\(u_2 \equiv u_1 + 1\), which can be linked to another sampling unit,
-denoted by \(s_2\). Because the GRTS algorithm requires the inclusion
-probabilities of the smallest sub-cells to be less than one, \(s_1\) and
-\(s_2\) must be distinct. This process continues until the \(n\)
-sampling units have been selected as part of the sample. Further details
-are provided by (Stevens Jr and Olsen, 2004) and (Dumelle et al., 2021).
+The GRTS algorithm is used to sample from finite and infinite
+populations and works by utilizing a mapping between two-dimensional and
+one-dimensional space. The population units in two-dimensional space are
+divided into cells using a hierarchical index. Population units are then
+mapped to a one-dimensional line via the hierarchical indexing. The line
+length of each population unit equals its inclusion probability. A
+systematic sample is conducted on the line and these samples are linked
+to a population unit in two-dimensional space, which results in the
+desired sample size. Stevens Jr and Olsen (2004) provide and Dumelle et
+al. (2021) provide further details. The GRTS algorithm is available in
+the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al.,
+2021).
 
-The Horvitz-Thompson estimator (Horvitz and Thompson, 1952) and its
-continuous analog (Cordy, 1993) yield unbiased estimates of population
-totals (and means). If \(\tau\) is a population total, then the
-Horvitz-Thompson estimator of \(\tau\), denoted by \(\hat{\tau}_{HT}\),
-is given by \begin{align}\label{eq:htest}
-  \hat{\tau}_{HT} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align} where \(Z_i\) is the observed value of the \(i\)th sampling
-unit, and \(\pi_i\) is the inclusion probability of the \(i\)th sampling
-unit. Estimating the variance of \(\hat{\tau}_{HT}\) directly relies on
-knowing the probability that \(s_i\) and \(s_j\) are both included in
-the sample (second-order inclusion probability), which can be unfeasible
-to compute. Furthermore, this variance estimator does not incorporate
-spatial locations. To address these challenges, (Stevens Jr and Olsen,
-2003) proposed an alternative variance estimator: the local neighborhood
-variance estimator. The local neighborhood variance estimator does not
-require second-order inclusion probabilities. It also incorporates
-spatial locations, which tends to reduce the estimated variance of
-compared \(\hat{\tau}_{HT}\) to a variance estimator ignoring spatial
-location (Stevens Jr and Olsen, 2003).
+After collecting a sample, the data are used to estimate population
+parameters. The Horvitz-Thompson estimator (Horvitz and Thompson, 1952)
+and its continuous analog (Cordy, 1993) yield unbiased estimates of
+population means and totals. If \(\tau\) is a population total, then the
+Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
+is given by \begin{align}\label{eq:ht}
+  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
+\end{align} where \(Z_i\) and \(pi_i\) are the observed value and
+inclusion probability of the \(i\)th population unit selected in the
+sample. Horvitz and Thompson (1952) and Sen (1953) provide variance
+estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
+they rely on calculating \(\pi_{ij}\), the probability that population
+unit \(i\) and population unit \(j\) are included in the sample, which
+can be very difficult to calculate. Second, they ignore the spatial
+locations of the population units. To address these drawbacks, Stevens
+Jr and Olsen (2003) proposed a local neighborhood variance estimator.
+The local neighborhood variance estimator does not rely on \(\pi_{ij}\),
+and it incorporates spatial locations by assigning higher weights to
+nearby observations. Stevens Jr and Olsen (2003) show this variance
+estimators tends to reduce the variability associated with estimating
+\(\tau\). This yields confidence intervals for \(\tau\) that are
+narrower than confidence intervals constructed from variance estimators
+ignoring spatial locations.
 
@@
 
-cases
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrrrr}
+  \hline
+Approach & Bias & RMSE & MedAE & Coverage & PClose & MedIL \\ 
+  \hline
+Design & 0.0003 & 0.0353 & 0.0251 & 0.9461 & 0.4889 & 0.1362 \\ 
+  Model & -0.0001 & 0.0362 & 0.0253 & 0.9480 & 0.5111 & 0.1430 \\ 
+   \hline
+\end{tabular}
+\caption{Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)}
+\label{tab:sim_one}
+\end{table}
 
@@
 Letters 18, 353--362.
+
+\leavevmode\hypertarget{ref-cressie1993statistics}{}%
+Cressie, N., 1993. Statistics for spatial data. John Wiley \& Sons.
 
@@
 
+\leavevmode\hypertarget{ref-schabenberger2017statistical}{}%
+Schabenberger, O., Gotway, C.A., 2017. Statistical methods for spatial
+data analysis. CRC press.
+
+\leavevmode\hypertarget{ref-sen1953estimate}{}%
+Sen, A.R., 1953. On the estimate of the variance in sampling with
+varying probabilities. Journal of the Indian Society of Agricultural
+Statistics 5, 127.
+
 \leavevmode\hypertarget{ref-sterba2009alternative}{}%

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added description of table.

@@
 
+For the following simulation results, we simulated 1040 different gridded populations, each of size 900 with sample size 150.
+For the model-based approach (FPBK), sites were selected via Independent Random Sample.
+For GRTS, the local mean variance was used.
+
+The response was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1.
+For model-based, we assumed the correct form of the covariance function (Exponential), but estimated the spatial parameters with REML.
+
 ```{r, results = "asis"}

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added notes .Rmds, potential data sets, and potential journals.

@@
 
-  It consists of two paragraphs.
-
 journal: "An awesome journal"
@@
 bibliography: mybibfile.bib
-#linenumbers: true
+linenumbers: true
 numbersections: true
@@
 _Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
+
+Potential Journals:
+
+* Ecological Applications
+* Methods in Ecology and Evolution
+* Journal of Applied Ecology
+* Environmetrics
+* Environmental and Ecological Statistics
 
@@
 
-finish spatial introduction
-revise section 2
-reread brus and wang
-write down potential data sets
-
-Please leave comments in your \textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}: \textcolor{blue}{Michael}, \textcolor{magenta}{Matt}, \textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
-
 <!-- Brief introduction to model based and design based -->
@@
 Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial location in sampling and model-based approaches.
+@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
+In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
+@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
+More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial sampling and model-based approaches.
 
@@
 
-### Design-Based Philosphy
-
 The design-based approach assumes the data are fixed.
@@
 @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
-
-### Model-Based Philosphy
 
@@
 
-Several comparisons between the model-based paradigm and the design-based paradigm have been made in many different contexts.
-@sterba2009alternative give some history of the two paradigms as well as some applications in psychological contexts.
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-We note that while there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, CITE some more of these), our goal is not to expand upon or improve these methods.
-
-Figure 1a.
+<!-- Figure 1a.
 Brus (2020): Data is fixed.
@@
 If we repeatedly sample the surface, then 95% of all 95% CIs will contain the true mean, which never changes.
+-->
 
-```{r, out.width = "85%"}
+```{r, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
+In the top row, we have one fixed population, and two random samples.
+In the bottom row, we have two realizations of the same spatial process sampled at the same locations."}
 set.seed(06092021)
@@
 
-Figure 1b.
+<!-- Figure 1b.
 Spatial process is fixed.
@@
 The realized mean changes from surface to surface and it's not necessarily the case that 95% of all 95% PIs will contain the true, underlying mean.
+-->
 
@@
 
+Potential Data Sets:
+
+* National Lakes Assessment 
+* Moose in Alaska
+* Temperature Data from NOAA
+
 # Discussion {#sec:discussion}

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added notes .Rmds, potential data sets, and potential journals.

@@
 \usepackage{lineno} % add
+  \linenumbers % turns line numbering on
 \providecommand{\tightlist}{%
@@
 
-  \title{A comparison of design-based and model-based approaches for spatial
-data.}
+  \title{A comparison of design-based and model-based approaches for
+spatial data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
   This is the abstract.
-
-  It consists of two paragraphs.
   \end{abstract}
@@
 
+Potential Journals:
+
+\begin{itemize}
+\tightlist
+\item
+  Ecological Applications
+\item
+  Methods in Ecology and Evolution
+\item
+  Journal of Applied Ecology
+\item
+  Environmetrics
+\item
+  Environmental and Ecological Statistics
+\end{itemize}
+
 \hypertarget{sec:introduction}{%
 \section{Introduction}\label{sec:introduction}}
-
-finish spatial introduction revise section 2 reread brus and wang write
-down potential data sets
-
-Please leave comments in your
-\textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}:
-\textcolor{blue}{Michael}, \textcolor{magenta}{Matt},
-\textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.
 
@@
 location in sampling to model-based approaches (Brus and De Gruijter,
-1997; Ver Hoef, 2002, 2008). More recent overviews include Brus (2020)
-and Wang et al. (2012), but no numerical comparison has been made
-between design-based approaches that incorporate spatial location in
-sampling and model-based approaches.
+1997; Ver Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two
+approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. In addition to Cooper (2006), there has been substantial
+research and development into estimators that use both design and
+model-based principles (see e.g. Cicchitelli and Montanari (2012),
+Chan-Golston et al. (2020) for a Bayesian approach, and Sterba (2009)).
+More recent overviews include Brus (2020) and Wang et al. (2012), but no
+numerical comparison has been made between design-based approaches that
+incorporate spatial sampling and model-based approaches.
 
@@
 
-\hypertarget{design-based-philosphy}{%
-\subsubsection{Design-Based Philosphy}\label{design-based-philosphy}}
-
 The design-based approach assumes the data are fixed. Randomness is
@@
 design-based approach.
-
-\hypertarget{model-based-philosphy}{%
-\subsubsection{Model-Based Philosphy}\label{model-based-philosphy}}
 
@@
 
-Several comparisons between the model-based paradigm and the
-design-based paradigm have been made in many different contexts. Sterba
-(2009) give some history of the two paradigms as well as some
-applications in psychological contexts. Cooper (2006) review the two
-approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. We note that while there has been substantial research and
-development into estimators that use both design and model-based
-principles (see e.g.~Cicchitelli and Montanari (2012), Chan-Golston et
-al. (2020) for a Bayesian approach, CITE some more of these), our goal
-is not to expand upon or improve these methods.
-
-Figure 1a. Brus (2020): Data is fixed. In a finite population example,
-show a 3d surface that can be generated by anything. If we repeatedly
-sample the surface, then 95\% of all 95\% CIs will contain the true
-mean, which never changes.
-
-\includegraphics[width=0.85\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-2-1}
-
-Figure 1b. Spatial process is fixed. In a finite population example,
-show 10 3d surfaces that are generated from some model. If we repeatedly
-generate the surface and obtain a sample, then 95\% of all 95\% PIs will
-contain the realized means. The realized mean changes from surface to
-surface and it's not necessarily the case that 95\% of all 95\% PIs will
-contain the true, underlying mean.
+\begin{figure}
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-2-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
+In the top row, we have one fixed population, and two random samples.
+In the bottom row, we have two realizations of the same spatial process sampled at the same locations.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 
+For the following simulation results, we simulated 1040 different
+gridded populations, each of size 900 with sample size 150. For the
+model-based approach (FPBK), sites were selected via Independent Random
+Sample. For GRTS, the local mean variance was used.
+
+The response was normally distributed with an exponential covariance
+function with partial sill of 0.9, effective range of \(\sqrt{2}\), and
+a nugget of 0.1. For model-based, we assumed the correct form of the
+covariance function (Exponential), but estimated the spatial parameters
+with REML.
+
 \begin{table}[ht]
@@
 
+Potential Data Sets:
+
+\begin{itemize}
+\tightlist
+\item
+  National Lakes Assessment
+\item
+  Moose in Alaska
+\item
+  Temperature Data from NOAA
+\end{itemize}
+
 \hypertarget{sec:discussion}{%

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor revisions and comments.

@@
     footnote: 1
-  - name: Matthew Higham
+  - name: Matt Higham
     email: mhigham@stlaw.edu
@@
 In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions made about the underlying process that generated the data.
+Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process that generated the data.
 Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
@@
 But the true mean of the spatial process that generated our realized data is still not known.
-When predicting the realized mean, we typically are not interested in the underlying process' true mean.
+When predicting the realized mean, we typically are not interested in the underlying process's true mean.
+<!-- (and is often not of interest) -->
 
@@
 These distributional assumptions are used to derive prediction intervals.
-The distributional assumptions allow the prediction intervals can be more precise.
+The distributional assumptions allow the prediction intervals to be more precise.
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
@@
 
+<!-- a sample is spatially balanced -->
+
 Sampling designs can incorporate spatial locations to obtain samples that are spatially balanced with respect to the population [@stevens2004spatially].
@@
 A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and are not meant to be miniatures of the population.
+Spatial balance with respect to the population is different than spatial balance with respect to geography.
+A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
 When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
+
+<!-- shorten?
+Spatial balance with respect to the population is different than spatial balance with respect to geography.
+A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
+When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
+-->
 
@@
 To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
-The first spatiallly balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
+The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
 Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
 We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because the algorithm has several attractive properties detailed by @stevens2004spatially and @dumelle2021spsurvey.
+
+<!-- algorithms. -->
 
@@
 
+<!-- This makes it seem like HT is being used to estimate population parameters in GRTS?
+Needs to be in there, but maybe a sentence making it clear that this isn't being used?
+Or just change the order and talk about Local variance first and then mention these for more general settings?
+-->
+
 After collecting a sample, the data are used to estimate population parameters.
 The Horvitz-Thompson estimator [@horvitz1952generalization] and its continuous analog [@cordy1993extension] yield unbiased estimates of population means and totals.
-If $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
+For example, if $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
 \begin{align}\label{eq:ht}
@@
 \end{align}
-where $Z_i$ and $pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
+where $Z_i$ and $\pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
@@
 
-We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
-
 <!-- \citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting} -->
 
-Finite Population Block Kriging (FPBK) is an alternative to samipling-based methods [@verhoef2008spatial].
+Finite Population Block Kriging (FPBK) is an alternative to sampling-based methods [@verhoef2008spatial].
 FPBK expands the geostatistical kriging framework to the finite population setting.
@@
 Neither require a particular distribution for $\mathbf{z}$.
+
+We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
+However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
 

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor updates from unlink pdf issue

@@
 FPBK expands the geostatistical kriging framework to the finite population setting.
-Instead of basing inference off of a specific sampling design, we assume the data were generated by a spatial process with parameters that can be estimated using the framework of a model.
+Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process with parameters that can be estimated using the framework of a model.
+
+<!-- don't like this: the framework of a model.  -->
 
@@
 For a response variable $\mathbf{z}$ that can be measured on a finite number of $N$ sites, we want to predict some linear function of the response variable, $\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a vector of weights.
-For example, if we want to predict the total abundance across all sites, then we would use a vector of 1's for the weights.
+For example, if we want to predict the population total across all sites, then we would use a vector of 1's for the weights.
+
+<!-- how explicitly do we want to define \mathbf{z} -->
+
+<!-- finite number of -->
+
+<!-- confusing having two tau's: tau for total and a tau as a function -->
 
@@
 
+<!-- do we want the general model or do we want a constant mean model with a note that this can be expanded to include covariates? -->
+
 We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
@@
 
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ $\tau(\mathbf{b}'\mathbf{z})$ is vague} and its prediction variance can be computed.
+<!-- need to add stationarity to this definition -->
+
+<!-- need to define indicator variable definition -->
+
+With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
 While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
@@
 
-We only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
+<!-- a little weird to talk about inference being based on distributional assumptions and then throw in this sentence about not assuming a distribution for z. -->
+
+We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
 However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
@@
 # Numerical Study {#sec:numstudy}
+
+__Sample Simulation__
 
@@
 * both good: correctly specified model with high correlation
-* break model: non-gaussian errors
+* break model: highly non-normal errors with small sample size
 * break design: small area estimation
+
+__Simulation Discussion Questions__
+
+* model-based: how should sample be drawn? should locations be fixed?
+* change n or sampling fraction?
+
+__Other Base Settings?__
 
@@
 
-* change n or sampling fraction
+## Software
 
-* model-based: how should sample be drawn? should locations be fixed?
-
-## Software
+<!-- spsurvey is talked about in GRTS. so either move that down here or move this to FPBK. -->
 

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor updates from unlink pdf issue

@@
    \ead{Dumelle.Michael@epa.gov} 
-    \author[STLAW]{Matthew Higham\corref{1}}
+    \author[STLAW]{Matt Higham\corref{1}}
    \ead{mhigham@stlaw.edu} 
@@
 in the model-based approach, inferences are made from specific
-assumptions made about the underlying process that generated the data.
-Each paradigm has a deep historical context (Sterba, 2009) and its own
-set of general advantages (Hansen et al., 1983).
+assumptions about the underlying process that generated the data. Each
+paradigm has a deep historical context (Sterba, 2009) and its own set of
+general advantages (Hansen et al., 1983).
 
@@
 When predicting the realized mean, we typically are not interested in
-the underlying process' true mean.
+the underlying process's true mean.
 
@@
 distributional assumptions are used to derive prediction intervals. The
-distributional assumptions allow the prediction intervals can be more
+distributional assumptions allow the prediction intervals to be more
 precise. Cressie (1993) and Schabenberger and Gotway (2017) provide
@@
 sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is is different than
-spatial balance with respect to geography. A sample that is spatially
-balanced with respect to geography is spread out in some type of
-equidistant manner over geographical space and are not meant to be
-miniatures of the population. When we refer to spatial balance
-henceforth, we mean spatial balance with respect to the population.
+Spatial balance with respect to the population is different than spatial
+balance with respect to geography. A sample that is spatially balanced
+with respect to geography is spread out in some type of equidistant
+manner over geographical space and is not meant to be miniatures of the
+population. When we refer to spatial balance henceforth, we mean spatial
+balance with respect to the population.
 
@@
 spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
-based on Voroni polygons. The first spatiallly balanced sampling
+based on Voroni polygons. The first spatially balanced sampling
 algorithm that saw widespread use was the Generalized Random
@@
 and its continuous analog (Cordy, 1993) yield unbiased estimates of
-population means and totals. If \(\tau\) is a population total, then the
-Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
-is given by \begin{align}\label{eq:ht}
+population means and totals. For example, if \(\tau\) is a population
+total, then the Horvitz-Thompson estimator of \(\tau\) (denoted by
+\(\hat{\tau}_{ht}\)), is given by \begin{align}\label{eq:ht}
   \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align} where \(Z_i\) and \(pi_i\) are the observed value and
+\end{align} where \(Z_i\) and \(\pi_i\) are the observed value and
 inclusion probability of the \(i\)th population unit selected in the
@@
 
-We only use FPBK in this paper in order to focus more on comparing the
-design-based and model-based approaches. However, k-nearest-neighbors
-(Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013), random forest
-(Breiman, 2001), Bayesian models (Chan-Golston et al., 2020), among
-others, can also be used to obtain predictions for a mean or total from
-spatially correlated responses in a finite population setting.
-
 Finite Population Block Kriging (FPBK) is an alternative to
-samipling-based methods (Ver Hoef, 2008). FPBK expands the
-geostatistical kriging framework to the finite population setting.
-Instead of basing inference off of a specific sampling design, we assume
-the data were generated by a spatial process with parameters that can be
-estimated using the framework of a model.
+sampling-based methods (Ver Hoef, 2008). FPBK expands the geostatistical
+kriging framework to the finite population setting. Instead of basing
+inference off of a specific sampling design, we assume the data are
+generated by a spatial process with parameters that can be estimated
+using the framework of a model.
 
@@
 \(\mathbf{b}\) is a vector of weights. For example, if we want to
-predict the total abundance across all sites, then we would use a vector
-of 1's for the weights.
+predict the population total across all sites, then we would use a
+vector of 1's for the weights.
 
@@
 With the above model formulation, the Best Linear Unbiased Predictor
-(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\)
-\(\tau(\mathbf{b}'\mathbf{z})\) is vague\} and its prediction variance
+(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\) and its prediction variance
 can be computed. While details of the derivation are in (Ver Hoef,
@@
 
+We note that we only use FPBK in this paper in order to focus more on
+comparing the design-based and model-based approaches. However,
+k-nearest-neighbors (Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013),
+random forest (Breiman, 2001), Bayesian models (Chan-Golston et al.,
+2020), among others, can also be used to obtain predictions for a mean
+or total from spatially correlated responses in a finite population
+setting.
+
 \hypertarget{sec:numstudy}{%
 \section{Numerical Study}\label{sec:numstudy}}
+
+\textbf{Sample Simulation}
 
@@
 \begin{itemize}
+\tightlist
 \item
@@
 \item
-  break model: non-gaussian errors
+  break model: highly non-normal errors with small sample size
 \item
   break design: small area estimation
+\end{itemize}
+
+\textbf{Simulation Discussion Questions}
+
+\begin{itemize}
+\tightlist
+\item
+  model-based: how should sample be drawn? should locations be fixed?
+\item
+  change n or sampling fraction?
+\end{itemize}
+
+\textbf{Other Base Settings?}
+
+\begin{itemize}
+\tightlist
 \item
@@
   break both? non-gaussian areas with smaller sample size
-\item
-  change n or sampling fraction
-\item
-  model-based: how should sample be drawn? should locations be fixed?
 \end{itemize}

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor updates

@@
 There are two general approaches for using data to make statistical inferences about a population: design-based approaches and model-based approaches.
-When data cannot be obtained for all units in a population (known as population units), data on a subset of the population units is collected in a  sample.
+When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected in a  sample.
 In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
@@
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial location in sampling to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
+Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
 @cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
@@
 @cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial sampling and model-based approaches.
+More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
 
@@
 In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
-In Section \ref{sec:numstudy}, we use simulated and real data to study the properties of parameter estimates from both approaches.
+In Section \ref{sec:numstudy}, we use simulated and real data to study the the behavior of both approaches.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 
-The model-based approach assumes the data are a random realization of a process.
-Randomness is often incorporated through distributional assumptions on the data-generating process.
+The model-based approach assumes the data are a random realization of a data-generating process.
+Randomness is often incorporated through distributional assumptions on this process.
 Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
@@
 We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-But the true mean of the spatial process that generated our realized data is still not known.
-When predicting the realized mean, we typically are not interested in the underlying process's true mean.
-<!-- (and is often not of interest) -->
+But we are typically not interested in the true, unknown mean of the underlying process.
 
@@
 
-```{r, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
+```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
 In the top row, we have one fixed population, and two random samples.
@@
 
+Description of Figure \ref{fig:fig1} goes here.
+
 <!-- Figure 1b.
@@
 
-<!-- a sample is spatially balanced -->
-
-Sampling designs can incorporate spatial locations to obtain samples that are spatially balanced with respect to the population [@stevens2004spatially].
-A sample is spatially with respect to the population if the sampled population units are a miniature of the population units.
+The design-based approach can use spatial locations to obtain spatially balanced samples.
+First we discuss spatial balance with respect to the population [@stevens2004spatially].
+A sample is spatially balanced with respect to the population if the sampled population units are a miniature of the population units.
 A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
@@
 
-<!-- shorten?
-Spatial balance with respect to the population is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
-When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
--->
-
-Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @benedetti2017spatiallyreview; @wang2013design].
+Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
 To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
 The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton] algorithms.
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because the algorithm has several attractive properties detailed by @stevens2004spatially and @dumelle2021spsurvey.
+Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
+We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties detailed by @stevens2004spatially and @dumelle2021spsurvey.
 
@@
 The line length of each population unit equals its inclusion probability.
-A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample size.
+A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample.
 @stevens2004spatially provide and @dumelle2021spsurvey provide further details.
-The GRTS algorithm is available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
 
@@
 
-After collecting a sample, the data are used to estimate population parameters.
-The Horvitz-Thompson estimator [@horvitz1952generalization] and its continuous analog [@cordy1993extension] yield unbiased estimates of population means and totals.
+After collecting a sample using the GRTS algorithm, the data are used to estimate population parameters.
+The Horvitz-Thompson estimator [@horvitz1952generalization] yields unbiased estimates of population means and totals.
 For example, if $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
@@
 where $Z_i$ and $\pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
+A similar formula exists for estimating the mean, $\mu$.
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, which can be very difficult to calculate.
+First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, and this can be very difficult to calculate.
 Second, they ignore the spatial locations of the population units.
@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$, and it incorporates spatial locations by assigning higher weights to nearby observations.
-@stevens2003variance show this variance estimators tends to reduce the variability associated with estimating $\tau$.
-This yields confidence intervals for $\tau$ that are narrower than confidence intervals constructed from variance estimators ignoring spatial locations.
+@stevens2003variance show this variance estimators tends to reduce the estimated standard error of $\hat{\tau}$, yielding narrower confidence confidence intervals for $\tau$.
 
@@
 
-<!-- \citep{VerHoef2002sampling, VerHoef2008spatial, Higham2020adjusting} -->
-
-Finite Population Block Kriging (FPBK) is an alternative to sampling-based methods [@verhoef2008spatial].
-FPBK expands the geostatistical kriging framework to the finite population setting.
-Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process with parameters that can be estimated using the framework of a model.
-
-<!-- don't like this: the framework of a model.  -->
-
+Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
+Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process.
 @verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-For a response variable $\mathbf{z}$ that can be measured on a finite number of $N$ sites, we want to predict some linear function of the response variable, $\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a vector of weights.
-For example, if we want to predict the population total across all sites, then we would use a vector of 1's for the weights.
+Let $\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
+.
+.
+, \text{z}(s_N) \}$ be a response variable that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
+Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a $1 \times N$ vector of weights.
+For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
 
@@
 
-Typically, however, we only have a sample of the $N$ sites.
-Denoting quantities that are part of the sampled sites with a subscript \emph{s} and quantities that are part of the unsampled sites with a subscript \emph{u},
+Typically, however, we only have a sample of the $N$ population units.
+Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
 
@@
 \end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled sites, respectively, and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled sites.
-Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume that $E(\bm{\delta})$ = $\mathbf{0}$.
-
-<!-- do we want the general model or do we want a constant mean model with a note that this can be expanded to include covariates? -->
+where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively; $\beta$ is the parameter vector of fixed effects; and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
+Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume the expectation of $\bm{\delta}$ equals $\mathbf{0}$.
 
 We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
-Many common choices for this function assume that spatial covariance decreases with increasing Euclidean distance between sites.
+Many common choices for this function assume that spatial covariance decreases with increasing Euclidean distance between population units.
 The primary function used throughout the simulations and applications of this manuscript is the Exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
@@
 \end{align}
-where $h_{i,j}$ is the distance between sites $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$.
+where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$.
 However, any spatial covariance function could be used in the place of the Exponential, including functions that allow for anisotropy [pg.
@@
 For the following simulation results, we simulated 1040 different gridded populations, each of size 900 with sample size 150.
-For the model-based approach (FPBK), sites were selected via Independent Random Sample.
+For the model-based approach (FPBK), population units were selected via Independent Random Sample.
 For GRTS, the local mean variance was used.
@@
 
+The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
 FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor updates

@@
 approaches. When data cannot be obtained for all units in a population
-(known as population units), data on a subset of the population units is
+(population units), data on a subset of the population units is
 collected in a sample. In the design-based approach, inferences about
@@
 they tend to compare design-based approaches that ignore spatial
-location in sampling to model-based approaches (Brus and De Gruijter,
-1997; Ver Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two
-approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. In addition to Cooper (2006), there has been substantial
-research and development into estimators that use both design and
-model-based principles (see e.g. Cicchitelli and Montanari (2012),
-Chan-Golston et al. (2020) for a Bayesian approach, and Sterba (2009)).
-More recent overviews include Brus (2020) and Wang et al. (2012), but no
-numerical comparison has been made between design-based approaches that
-incorporate spatial sampling and model-based approaches.
+locations to model-based approaches (Brus and De Gruijter, 1997; Ver
+Hoef, 2002, 2008). Cooper (2006) review the two approaches in an
+ecological context before introducing a ``model-assisted'' variance
+estimator that combines aspects from each approach. In addition to
+Cooper (2006), there has been substantial research and development into
+estimators that use both design and model-based principles (see
+e.g.~Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
+Bayesian approach, and Sterba (2009)). More recent overviews include
+Brus (2020) and Wang et al. (2012), but no numerical comparison has been
+made between design-based approaches that incorporate spatial locations
+and model-based approaches.
 
@@
 Section \ref{sec:numstudy}, we use simulated and real data to study the
-properties of parameter estimates from both approaches. And in Section
-\ref{sec:discussion}, we end with a discussion and provide directions
-for future research.
+the behavior of both approaches. And in Section \ref{sec:discussion}, we
+end with a discussion and provide directions for future research.
 
@@
 The model-based approach assumes the data are a random realization of a
-process. Randomness is often incorporated through distributional
-assumptions on the data-generating process. Instead of estimating fixed
+data-generating process. Randomness is often incorporated through
+distributional assumptions on this process. Instead of estimating fixed
 but unknown parameters (as in the design-based approach), the goal of
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-2-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/fig1-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
 In the top row, we have one fixed population, and two random samples.
-In the bottom row, we have two realizations of the same spatial process sampled at the same locations.}\label{fig:unnamed-chunk-2}
+In the bottom row, we have two realizations of the same spatial process sampled at the same locations.}\label{fig:fig1}
 \end{figure}
+
+Figure \ref{fig:fig1} shows
 
@@
 that are spatially balanced with respect to the population (Stevens Jr
-and Olsen, 2004). A sample is spatially with respect to the population
-if the sampled population units are a miniature of the population units.
-A sample is a miniature of the population if the distribution of the
-sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is different than spatial
-balance with respect to geography. A sample that is spatially balanced
-with respect to geography is spread out in some type of equidistant
-manner over geographical space and is not meant to be miniatures of the
-population. When we refer to spatial balance henceforth, we mean spatial
-balance with respect to the population.
+and Olsen, 2004). A sample is spatially balanced with respect to the
+population if the sampled population units are a miniature of the
+population units. A sample is a miniature of the population if the
+distribution of the sampled population units mirrors the density of all
+population units. Spatial balance with respect to the population is
+different than spatial balance with respect to geography. A sample that
+is spatially balanced with respect to geography is spread out in some
+type of equidistant manner over geographical space and is not meant to
+be miniatures of the population. When we refer to spatial balance
+henceforth, we mean spatial balance with respect to the population.
 

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor edits.

@@
 We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
-Many common choices for this function assume that spatial covariance decreases with increasing Euclidean distance between population units.
-The primary function used throughout the simulations and applications of this manuscript is the Exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
+Many common choices for this function assume stationarity and isotropy (CITE Cressie?) with spatial covariance that decreases with increasing Euclidean distance between population units.
+The primary function used throughout the simulations and applications of this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
 \mbox{}
@@
 \end{align}
-where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$.
-However, any spatial covariance function could be used in the place of the Exponential, including functions that allow for anisotropy [pg.
+where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$, and $\mathbbm{1}$ is an indicator function.
+However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [pg.
 80 - 93][@chiles1999geostatistics].
-
-<!-- need to add stationarity to this definition -->
-
-<!-- need to define indicator variable definition -->
 

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor grammatical updates

@@
 author:
-  - name: Michael Dumelle
+  - name: In alphabetical order Michael Dumelle
     email: Dumelle.Michael@epa.gov
@@
 We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
-Many common choices for this function assume stationarity and isotropy (CITE Cressie?) with spatial covariance that decreases with increasing Euclidean distance between population units.
-The primary function used throughout the simulations and applications of this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\var(\bm{\delta})$ is
+It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
+Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\cov(\bm{\delta})$ is
 \mbox{}
@@
 where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$, and $\mathbbm{1}$ is an indicator function.
-However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [pg.
-80 - 93][@chiles1999geostatistics].
+However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
+80-93].
 
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $\tau(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
+With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
 While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
-Neither require a particular distribution for $\mathbf{z}$.
-
-<!-- a little weird to talk about inference being based on distributional assumptions and then throw in this sentence about not assuming a distribution for z. -->
 
@@
 
-For the following simulation results, we simulated 1040 different gridded populations, each of size 900 with sample size 150.
-For the model-based approach (FPBK), population units were selected via Independent Random Sample.
-For GRTS, the local mean variance was used.
+For the following simulation results, we simulated 1040 different gridded populations, each of size 900 (on the unit square) with sample size 150.
+For the design-based approach, population units were selected via GRTS, the Horvitz-Thompson estimator was used,and the local mean variance was used.
+For the model-based approach (FPBK), population units were selected via Independent Random Sampling (IRS) and the appropriate prediction and prediction variance formulas were used.
 
 The response was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1.
-For model-based, we assumed the correct form of the covariance function (Exponential), but estimated the spatial parameters with REML.
+For model-based, we assumed the correct form of the covariance function (exponential), but estimated the spatial parameters with REML.
 
@@
 
-* both good: correctly specified model with high correlation
+* both good: correctly specified model with high correlation (we did this in Table \ref{tab:sim_one})
 * break model: highly non-normal errors with small sample size
@@
 
-<!-- spsurvey is talked about in GRTS. so either move that down here or move this to FPBK. -->
-
 The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
@@
 
-<!-- References from Brus2020statistical, de1990model, Brus1997random -->
-
 # References {#references .unnumbered}

### 2021-06-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor grammatical updates

@@
 data.}
-    \author[USEPA]{Michael Dumelle\corref{1}}
+    \author[USEPA]{In alphabetical order Michael Dumelle\corref{1}}
    \ead{Dumelle.Michael@epa.gov} 
@@
 population units, we would have an exact prediction for the mean of our
-one realized process, without any uncertainty. But the true mean of the
-spatial process that generated our realized data is still not known.
-When predicting the realized mean, we typically are not interested in
-the underlying process's true mean.
+one realized process, without any uncertainty. But we are typically not
+interested in the true, unknown mean of the underlying process.
 
@@
 
-Figure \ref{fig:fig1} shows
+Description of Figure \ref{fig:fig1} goes here.
 
@@
 
-Sampling designs can incorporate spatial locations to obtain samples
-that are spatially balanced with respect to the population (Stevens Jr
-and Olsen, 2004). A sample is spatially balanced with respect to the
-population if the sampled population units are a miniature of the
-population units. A sample is a miniature of the population if the
-distribution of the sampled population units mirrors the density of all
-population units. Spatial balance with respect to the population is
-different than spatial balance with respect to geography. A sample that
-is spatially balanced with respect to geography is spread out in some
-type of equidistant manner over geographical space and is not meant to
-be miniatures of the population. When we refer to spatial balance
-henceforth, we mean spatial balance with respect to the population.
+The design-based approach can use spatial locations to obtain spatially
+balanced samples. First we discuss spatial balance with respect to the
+population (Stevens Jr and Olsen, 2004). A sample is spatially balanced
+with respect to the population if the sampled population units are a
+miniature of the population units. A sample is a miniature of the
+population if the distribution of the sampled population units mirrors
+the density of all population units. Spatial balance with respect to the
+population is different than spatial balance with respect to geography.
+A sample that is spatially balanced with respect to geography is spread
+out in some type of equidistant manner over geographical space and is
+not meant to be miniatures of the population. When we refer to spatial
+balance henceforth, we mean spatial balance with respect to the
+population.
 
@@
 2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
-Halton Iterative Partitioning (Robertson et al., 2018) algorithms. We
-focus on the Generalized Random Tessellation Stratified (GRTS) algorithm
-to select spatially balanced sampling because the algorithm has several
-attractive properties detailed by Stevens Jr and Olsen (2004) and
-Dumelle et al. (2021).
+Halton Iterative Partitioning (Robertson et al., 2018). We focus on the
+Generalized Random Tessellation Stratified (GRTS) algorithm to select
+spatially balanced sampling because it has several attractive properties
+detailed by Stevens Jr and Olsen (2004) and Dumelle et al. (2021).
 
@@
 to a population unit in two-dimensional space, which results in the
-desired sample size. Stevens Jr and Olsen (2004) provide and Dumelle et
-al. (2021) provide further details. The GRTS algorithm is available in
-the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al.,
-2021).
+desired sample. Stevens Jr and Olsen (2004) provide and Dumelle et al.
+(2021) provide further details.
 
-After collecting a sample, the data are used to estimate population
-parameters. The Horvitz-Thompson estimator (Horvitz and Thompson, 1952)
-and its continuous analog (Cordy, 1993) yield unbiased estimates of
-population means and totals. For example, if \(\tau\) is a population
-total, then the Horvitz-Thompson estimator of \(\tau\) (denoted by
-\(\hat{\tau}_{ht}\)), is given by \begin{align}\label{eq:ht}
+After collecting a sample using the GRTS algorithm, the data are used to
+estimate population parameters. The Horvitz-Thompson estimator (Horvitz
+and Thompson, 1952) yields unbiased estimates of population means and
+totals. For example, if \(\tau\) is a population total, then the
+Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
+is given by \begin{align}\label{eq:ht}
   \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
@@
 inclusion probability of the \(i\)th population unit selected in the
-sample. Horvitz and Thompson (1952) and Sen (1953) provide variance
-estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
-they rely on calculating \(\pi_{ij}\), the probability that population
-unit \(i\) and population unit \(j\) are included in the sample, which
-can be very difficult to calculate. Second, they ignore the spatial
-locations of the population units. To address these drawbacks, Stevens
-Jr and Olsen (2003) proposed a local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on \(\pi_{ij}\),
-and it incorporates spatial locations by assigning higher weights to
-nearby observations. Stevens Jr and Olsen (2003) show this variance
-estimators tends to reduce the variability associated with estimating
-\(\tau\). This yields confidence intervals for \(\tau\) that are
-narrower than confidence intervals constructed from variance estimators
-ignoring spatial locations.
+sample. A similar formula exists for estimating the mean, \(\mu\).
+Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
+for \(\hat{\tau}_{ht}\), but they have two drawbacks. First, they rely
+on calculating \(\pi_{ij}\), the probability that population unit \(i\)
+and population unit \(j\) are included in the sample, and this can be
+very difficult to calculate. Second, they ignore the spatial locations
+of the population units. To address these drawbacks, Stevens Jr and
+Olsen (2003) proposed a local neighborhood variance estimator. The local
+neighborhood variance estimator does not rely on \(\pi_{ij}\), and it
+incorporates spatial locations by assigning higher weights to nearby
+observations. Stevens Jr and Olsen (2003) show this variance estimators
+tends to reduce the estimated standard error of \(\hat{\tau}\), yielding
+narrower confidence confidence intervals for \(\tau\).
 
@@
 
-Finite Population Block Kriging (FPBK) is an alternative to
-sampling-based methods (Ver Hoef, 2008). FPBK expands the geostatistical
-kriging framework to the finite population setting. Instead of basing
-inference off of a specific sampling design, we assume the data are
-generated by a spatial process with parameters that can be estimated
-using the framework of a model.
+Finite Population Block Kriging (FPBK) is a model-based approach that
+expands the geostatistical Kriging framework to the finite population
+setting (Ver Hoef, 2008). Instead of basing inference off of a specific
+sampling design, we assume the data are generated by a spatial process.
+Ver Hoef (2008) gives details on the theory of FPBK, but some of the
+basic principles are summarized below. Let
+\(\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}\)
+be a response variable that can be measured at the \(N\) population
+units and is represented as an \(N \times 1\) vector. Suppose we want to
+predict some linear function of the response variable,
+\(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where \(\mathbf{b}\)
+is a \(1 \times N\) vector of weights. For example, if we want to
+predict the population total across all population units, then we would
+use a vector of 1's for the weights.
 
-Ver Hoef (2008) gives details on the theory of FPBK, but some of the
-basic principles are summarized below. For a response variable
-\(\mathbf{z}\) that can be measured on a finite number of \(N\) sites,
-we want to predict some linear function of the response variable,
-\(\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
-\(\mathbf{b}\) is a vector of weights. For example, if we want to
-predict the population total across all sites, then we would use a
-vector of 1's for the weights.
-
-Typically, however, we only have a sample of the \(N\) sites. Denoting
-quantities that are part of the sampled sites with a subscript \emph{s}
-and quantities that are part of the unsampled sites with a subscript
-\emph{u},
+Typically, however, we only have a sample of the \(N\) population units.
+Denoting quantities that are part of the sampled population units with a
+subscript \emph{s} and quantities that are part of the unsampled
+population units with a subscript \emph{u},
 
@@
 \end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
-design matrices for the sampled and unsampled sites, respectively, and
+design matrices for the sampled and unsampled population units,
+respectively; \(\beta\) is the parameter vector of fixed effects; and
 \(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
-sampled and unsampled sites. Denoting
+sampled and unsampled population units, respectively. Denoting
 \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
-that \(E(\bm{\delta})\) = \(\mathbf{0}\).
+the expectation of \(\bm{\delta}\) equals \(\mathbf{0}\).
 
 We also typically assume that there is spatial correlation in
-\(\bm{\delta}\), which can be modeled using a covariance function. Many
-common choices for this function assume that spatial covariance
-decreases with increasing Euclidean distance between sites. The primary
-function used throughout the simulations and applications of this
-manuscript is the Exponential covariance function: the \(i,j^{th}\)
-entry for \(\mathop{\mathrm{{var}}}(\bm{\delta})\) is \mbox{}
+\(\bm{\delta}\), which can be modeled using a covariance function. It is
+common to assume the covariance function is second-order stationary and
+isotropic (Cressie, 1993), and that the spatial covariance decreases as
+the separation between population units increases. Many spatial
+covariance functions exist, but the primary function we use throughout
+the simulations and applications in this manuscript is the exponential
+covariance function: the \(i,j^{th}\) entry for
+\(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
 \begin{align}\label{equation:expcov}
 \mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
-\end{align} where \(h_{i,j}\) is the distance between sites \(i\) and
-\(j\), and \(\bm{\theta}\) is a vector of spatial covariance parameters
-of the partial sill \(\theta_1\), the range \(\theta_2\), and the nugget
-\(\theta_3\). However, any spatial covariance function could be used in
-the place of the Exponential, including functions that allow for
-anisotropy {[}pg. 80 - 93{]}(Chiles and Delfiner, 1999).
+\end{align} where \(h_{i,j}\) is the distance between population units
+\(i\) and \(j\), and \(\bm{\theta}\) is a vector of spatial covariance
+parameters of the partial sill \(\theta_1\), the range \(\theta_2\), and
+the nugget \(\theta_3\), and \(\mathbbm{1}\) is an indicator function.
+However, any spatial covariance function could be used in the place of
+the exponential, including functions that allow for non-stationarity or
+anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
 
 With the above model formulation, the Best Linear Unbiased Predictor
-(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\) and its prediction variance
-can be computed. While details of the derivation are in (Ver Hoef,
-2008), we note here that the predictor and its variance are both
-moment-based. Neither require a particular distribution for
-\(\mathbf{z}\).
+(BLUP) for \(f(\mathbf{b}'\mathbf{z})\) and its prediction variance can
+be computed. While details of the derivation are in (Ver Hoef, 2008), we
+note here that the predictor and its variance are both moment-based.
 
@@
 For the following simulation results, we simulated 1040 different
-gridded populations, each of size 900 with sample size 150. For the
-model-based approach (FPBK), sites were selected via Independent Random
-Sample. For GRTS, the local mean variance was used.
+gridded populations, each of size 900 (on the unit square) with sample
+size 150. For the design-based approach, population units were selected
+via GRTS, the Horvitz-Thompson estimator was used,and the local mean
+variance was used. For the model-based approach (FPBK), population units
+were selected via Independent Random Sampling (IRS) and the appropriate
+prediction and prediction variance formulas were used.
 
@@
 a nugget of 0.1. For model-based, we assumed the correct form of the
-covariance function (Exponential), but estimated the spatial parameters
+covariance function (exponential), but estimated the spatial parameters
 with REML.
@@
 \item
-  both good: correctly specified model with high correlation
+  both good: correctly specified model with high correlation (we did
+  this in Table \ref{tab:sim_one})
 \item
@@
 
-FPBK can be readily performed in \texttt{R} with the \texttt{sptotal}
-package (Higham et al., 2020). We use \texttt{sptotal} for both the
-simulation analysis and the application, estimating parameters with
-Restricted Maximum Likelihood (REML).
+The GRTS algorithm and the local neighborhood variance estimator are
+available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
+et al., 2021). FPBK can be readily performed in \texttt{R} with the
+\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
+for both the simulation analysis and the application, estimating
+parameters with Restricted Maximum Likelihood (REML).
 
@@
 Environmetrics Society 17, 539--553.
-
-\leavevmode\hypertarget{ref-cordy1993extension}{}%
-Cordy, C.B., 1993. An extension of the horvitz---thompson theorem to
-point sampling from a continuous universe. Statistics \& Probability
-Letters 18, 353--362.
 

### 2021-06-14 inst/SpatialDVM_Manuscript/notes_Olsen.Rmd
uploaded tony notes

# Please Leave Your Notes Here

Rmd 2: Is the paper restricted to finite populations? if so then change title to
"a comparison of design-based and model-based approaches for finite population 
spatial data"

Rmd 76: finite population or not? if so, then "inferences about a finite population" 
and "all units in a finite population".  Much but not all of the literature assumes
finite populations for comparisons. My understanding is that restricting to finite
populations is what is intended for this paper.

Rmd 80: design-based vs model-based in spatial context. Incorporating spatial location
can be done at the design stage, i.e., selection of the sample units from the finite 
set of population units, or at the analysis stage, incorporating spatial information
for estimation.  Should this paragraph address this?

"We define spatial data...": Virtually all data is spatial. Is just a question for us
whether the spatial information is either incorporated into the selection of units or
into the statistical estimation.  Could redo sentence to be: "We define spatial data as
data that incorporates the specific locations of the population units into either the 
design or estimation process."

In general, not clear throughout the paragraph whether refer to process of site selection 
or estimation.

Rmd 90: Like the section as it focuses on whether statistical inference comes from
process of selecting the sites (design-based design and analysis) or realization of 
statistical process (model-based analysis of realizations). While not explicitly stated,
either analysis could be based on explicit use of spatial location or not at all. That is, 
while the examples are spatial in nature neither their site selection nor analysis needs
to be spatial.

Rmd 92: "independent random sampling (IRS), stratified random sampling, and cluster sampling." Replace by "simple random sampling, stratified random sample, and cluster
sampling, which we refer to as Independent Random Sampling (IRS) survey designs". My 
assumption is that all of these are interpreted as using srs. GRTS-based survey designs applies to equal probability designs, stratified designs and cluster designs as well. 
Just as any spatially-balanced-based design does as well.

"typically assumed" - always? drop typically or what am I missing?

Rmd 170: Before this section is it useful to have a section that lays out the general
site selection and general analysis options. Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based. Then general analysis options as design-based no spatial, design-based spatial, model-based. This four by three table would show that model-based analyses are possible for all selection options. Design-based options with no spatial info possible for IRS-based and GRTS-based. Design-based options with spatial info possible for GRTS-based.

Then the focus of the paper is to compare WHAT? design-based GRTS site selection/design-based spatial analysis to IRS-based site selection/model-based FPBK?

### 2021-06-25 inst/SpatialDVM_Manuscript/notes_to_do.Rmd
added to-do list with notes from June 24 meeting.

---
title: "To Do List"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction and Background:

* consider re-wording the model-based use of the word "parameter" to make the parameter generating the process distinct from using "parameter" as, for example, the realized mean (which we treat as fixed for one particular realization in the finite population setting).

Simulations:

* put in SRS as base-line sampling setting.
* add FPBK as traditional SRS analysis as methods to analyze SRS sample.
* add FPBK and traditional SRS analysis as additional methods to analyze GRTS sample.
* add GRTS with 10% nugget sites as additional sampling setting? If so, add FPBK to analyze this sample.

### 2021-07-12 inst/scripts/application.Rmd
initial application analysis with DOC and lakes data.

---
title: "Application with Lakes Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

__Notes__:

* Major drawback of only sampling once for each design (IRS and GRTS) for skewed data: the results will be very dependent on how many of the skewed outleirs are selected in the sample. See this analysis for an example of IRS being artificially more "precise" because GRTS happens to select a large outlier to be in its sample.
* for both IRS and GRTS, model-based analyses have a lower standard error (and therefore a more narrow prediction interval). In terms of prediction error, design IRS < model IRS < model GRTS < design GRTS (but, again the fact that GRTS by chance selected an outlier to be in the sample made its prediction error worse). 
* these analyses are finding the mean DOC across all lakes, ignoring the lake area/volume.

__Things to Think About__:

* based on these results, I'm inclined toward simulating this application many times: otherwise, results change depending on what the seed is set to. We could also then (hopefully) report that the methods cover fine for this real data example.
* are we happy with using `DOC` for extremely skewed data or should we consider a different response? Would we want to drop the two __extremely large__ values from the analysis? It doesn't seem like they are errors in data entry so I'm not sure that this would be justified.
* one option for symmetric data is to pretend that we are interested in predicting the mean of log(DOC). The downside is that this isn't really a realistic goal that people would have in practice but the upside is that it's the same response variable. If not log(DOC), is there another variable you had in mind?
* are we happy with a sample size of 200 or do we want it to be smaller to more closely match simulations?

## Data Exploration

Copied from `sptotal` vignette:

As another example, we took data from the [National Aquatic Resource Surveys](https://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys). With concerns about global warming, the earth's capacity to store carbon is of great interest, and dissolved organic carbon (DOC) is an estimate of a lake's ability to store carbon.  We will estimate the mean mg/L for DOC from a sample of lakes.  If the total lake volume could be calculated (we will not attempt that), then the total dissolved carbon in a population of lakes could be estimated. We will examine DOC in lakes from the 2012 surveys.  We combined [site data](https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_siteinfo_08232016.csv), [DOC data](https://www.epa.gov/sites/production/files/2016-12/nla2012_waterchem_wide.csv), and [habitat metrics](https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_phabmet_10202016.csv) to create a data set of 1206 lakes in the conterminous United States.

The following code explores the variables in the data set, focusing on dissolved organic carbon (DOC). We can see tht `DOC_RESULT` is highly skewed with at least one extreme outlier and that the log of `DOC_RESULT` is approximately symmetric with slightly larger tails than a normal distribution would have.

```{r, results = "hide", fig.keep = "none"}
library(sptotal)
data(USlakes)
library(skimr)
skim(USlakes)

library(tidyverse)
USlakes <- USlakes %>% mutate(logdoc = log(DOC_RESULT))

ggplot(data = USlakes, aes(x = log(DOC_RESULT))) +
  geom_histogram(bins = 20)

ggplot(data = USlakes, aes(x = DOC_RESULT)) +
  geom_histogram(bins = 20)

summary(USlakes$DOC_RESULT)
```

We can keep log DOC in mind, but focus for now on the untransformed variable. The following plot shows that there are two observations with extremely large `DOC_RESULT` values that are clustered together.

```{r}
ggplot(data = USlakes, aes(x = XCOORD, y = YCOORD)) +
  geom_point(aes(colour = DOC_RESULT)) +
  scale_colour_viridis_c()
```

Filtering these two values out, we see from the next plot that there is evidence of spatial correlation in DOC, as many of the observations with higher DOC's are clustered together in the northern plains of the United States and in the southeastern state of Florida.

```{r}
ggplot(data = USlakes %>% filter(DOC_RESULT < 300),
       aes(x = XCOORD, y = YCOORD)) +
  geom_point(aes(colour = DOC_RESULT)) +
  scale_colour_viridis_c()
```

In the following analyses, we will keep all observations (and not remove any extreme outliers).

## IRS Sample, Model Analysis

Copied from sptotal vignette: We have the whole population of lakes, but, with budget cuts, it is likely that this whole population will not always be surveyed in its entirety.  So, we will ask the question, "If we sample from this population, can we still get a fairly precise prediction of the mean DOC?"

Copied from sptotal vignette: We will do the same thing that we did with the simulated data, and take a random sample of 200 lakes.  Also, because we want the mean, and not a total, we will create a weights column for the `lakeobs` data set, with each element $1/N$, where, here, $N = 1204$.

```{r}
set.seed(2)
lakesobs <- USlakes %>% sample_n(200) ## sample 200 lakes
lakesunobs <- anti_join(USlakes, lakesobs)
lakesunobs$DOC_RESULT <- NA

lakes_test <- bind_rows(lakesobs, lakesunobs) 
lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
```

We first analyze the IRS sample with a model-based analysis.

```{r}
slmfitout_exp_lakes <- slmfit(formula = DOC_RESULT ~ 1,
                              data = lakes_test, 
                              xcoordcol = 'XCOORD',
                              ycoordcol = 'YCOORD',
                              CorModel = "Exponential")
summary(slmfitout_exp_lakes)
plot(slmfitout_exp_lakes)
```

```{r}
pred_exp_lakes <- predict(slmfitout_exp_lakes,  wtscol = "wts",
                          conf_level = 0.95)
print(pred_exp_lakes)
mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
ub_irs_mod <- pred_exp_lakes$conf_bounds[2]

realized_mean <- mean(USlakes$DOC_RESULT) ## compute realized mean
```

## IRS Sample, Design Analysis

```{r}
irs_info <- lakesobs %>% summarise(meandoc = mean(DOC_RESULT),
                                   vardoc = var(DOC_RESULT))
N <- nrow(USlakes); n <- nrow(lakesobs)
irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
irs_info$meandoc; sqrt(irs_info$vardoc)
irs_lb_irs_samp; irs_ub_irs_samp
```

## GRTS Sample, Design Analysis

Next, we perform design-based and model-based analyses using a GRTS sample of 200 sites. The following code obtains the GRTS sample.

```{r, results = "hide"}
set.seed(320215)
data_sf <- sf::st_as_sf(USlakes, coords = c("XCOORD", "YCOORD"),
                        crs = 5070)
library(spsurvey)
grts_samp <- grts(data_sf, n_base = n)
grts_bind <- sprbind(grts_samp)

## get coordinates
grts_coords <- sf::st_coordinates(grts_bind)
## make data frame
grts_df <- data.frame(
  response = grts_bind$DOC_RESULT,
  x = grts_coords[, "X"],
  y = grts_coords[, "Y"],
  siteID = grts_bind$siteID,
  wgt = grts_bind$wgt
)
## head(grts_df)
summary(grts_df$response)
```

We then perform a design-based analysis on the GRTS sample.

```{r}
design_analysis <- cont_analysis(
  grts_df,
  siteID = "siteID",
  vars = "response",
  weight = "wgt",
  xcoord = "x",
  ycoord = "y"
)
## just return the mean info
## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
design_mean_grts <- design_analysis$Mean
design_mean_grts$Estimate
design_mean_grts$StdError
design_mean_grts$LCB95Pct
design_mean_grts$UCB95Pct
```

## GRTS Sample, Model Analysis

Finally, we use a model-based analysis on the GRTS sample.

```{r}
grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
  rename(DOC_RESULT = "response", XCOORD = "x", YCOORD = "y")
grts_unsamp <- anti_join(USlakes, grts_coords_resp)
grts_unsamp$DOC_RESULT <- NA
grts_unsamp <- grts_unsamp %>% select(DOC_RESULT, XCOORD, YCOORD)
grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
```

```{r}
grts_full$wts <- 1 / nrow(grts_full)
mod_grts <- slmfit(DOC_RESULT ~ 1, data = grts_full,
                   xcoordcol = "XCOORD",
                   ycoordcol = "YCOORD")
pred_mod_grts <- predict(mod_grts, wtscol = "wts")
model_mean_grts <- pred_mod_grts$FPBK_Prediction
model_se_grts <- sqrt(pred_mod_grts$PredVar)
model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
model_mean_grts
model_se_grts
model_lb_grts
model_ub_grts
```

Finally, we combine all results in a data frame for easy comparison:

```{r}
tibble(approach = c("Design IRS", "Design GRTS", "Model IRS", "Model GRTS"),
       realized_mean = c(realized_mean, realized_mean, realized_mean,
                     realized_mean),
       estimate = c(irs_info$meandoc,
                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
                    as.vector(model_mean_grts)),
       sd = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
              as.vector(model_se_grts)),
       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
              as.vector(model_lb_grts)),
       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
              as.vector(model_ub_grts)))
```

All 95% intervals contain the realized mean of 8.32. We also see that model-based GRTS is more precise than design-based GRTS, and that model-based IRS is more precise than design-based IRS for this sample. The fact that IRS is more precise than GRTS is likely a consequence of GRTS (by chance) obtaining a very large value for `DOC_RESULT` in its sample.

```{r}
## grts obtains a very large response value in its sample
ggplot(grts_df, aes(x = response)) +
         geom_histogram(colour  = "black", fill = "white", bins = 20)

## srs does not
ggplot(lakesobs, aes(x = DOC_RESULT)) +
         geom_histogram(colour  = "black", fill = "white", bins = 20)
```

### 2021-07-26 inst/scripts/application2.Rmd
new application with different lakes data set.

---
title: "Application with Different Lakes Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(sptotal)
library(skimr)
nla_df <- read_csv(here("data", "nla12_keyvariables_data.csv"))
skim(nla_df)
```

```{r}
View(nla_df)
factor(nla_df$TOTALHG_UNITS) ## all same units
```

```{r}
nla_df %>% group_by(SITE_ID) %>%
  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
  ungroup() %>% summarise(maxnonmiss = max(nonmiss))

nla_df %>% group_by(SITE_ID) %>%
  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
  summarise(sumnomiss = sum(nonmiss == 0))
```

HG never measured twice at the same site. 35 sites never have HG measured. These will be dropped as well as the duplicate sites.

```{r}
nla_nomiss <- nla_df %>% filter(!is.na(TOTALHG_RESULT)) %>%
  rename(total_hg = "TOTALHG_RESULT")

## check
nla_nomiss %>% group_by(SITE_ID) %>% count() %>%
  ungroup() %>%
  summarise(not1 = sum(n != 1))
```

Exploratory graph

```{r}
ggplot(data = nla_nomiss, aes(x = total_hg)) +
  geom_histogram(bins = 20)
```

Exploratory summary stats

```{r}
summary(nla_nomiss$total_hg)
```

```{r}
ggplot(data = nla_nomiss, aes(x = INDEX_LON_DD, y = INDEX_LAT_DD)) +
  geom_point(aes(colour = total_hg)) +
  scale_colour_viridis_c()
```

We see some spatial correlation (higher values in northeast, north, south, and somewhat high on the western coast. lower values in middle america, where it tends to be more rural).

transform coordinates

```{r}
coord_list <- LLtoTM(cm = mean(nla_nomiss$INDEX_LON_DD), lat = nla_nomiss$INDEX_LAT_DD,
       lon = nla_nomiss$INDEX_LON_DD)
nla_nomiss$xcoords <- coord_list$xy[ ,1]
nla_nomiss$ycoords <- coord_list$xy[ ,2]
ggplot(data = nla_nomiss, aes(x = xcoords, y = ycoords)) +
  geom_point(aes(colour = total_hg)) +
  scale_colour_viridis_c()
```

## IRS Sample, Model Analysis

```{r}
set.seed(07262021)
lakesobs <- nla_nomiss %>% sample_n(200) ## sample 200 lakes
lakesunobs <- anti_join(nla_nomiss, lakesobs)
lakesunobs$total_hg <- NA

lakes_test <- bind_rows(lakesobs, lakesunobs) 
lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
```

```{r}
slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
                              data = lakes_test, 
                              xcoordcol = 'xcoords',
                              ycoordcol = 'ycoords',
                              CorModel = "Exponential")
summary(slmfitout_exp_lakes)
plot(slmfitout_exp_lakes)
```

```{r}
pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
                          conf_level = 0.95)
print(pred_exp_lakes)
mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
ub_irs_mod <- pred_exp_lakes$conf_bounds[2]

realized_mean <- mean(nla_nomiss$total_hg) ## compute realized mean
```

## IRS Sample, Design Analysis

```{r}
irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
                                   vardoc = var(total_hg))
N <- nrow(nla_nomiss); n <- nrow(lakesobs)
irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
irs_info$meandoc; sqrt(irs_info$vardoc)
irs_lb_irs_samp; irs_ub_irs_samp
```

## GRTS Sample, Design Analysis

Next, we perform design-based and model-based analyses using a GRTS sample of 200 sites. The following code obtains the GRTS sample.

```{r, results = "hide"}
set.seed(07262021)
data_sf <- sf::st_as_sf(nla_nomiss, coords = c("xcoords", "ycoords"),
                        crs = 5070)
library(spsurvey)
grts_samp <- grts(data_sf, n_base = n)
grts_bind <- sprbind(grts_samp)

## get coordinates
grts_coords <- sf::st_coordinates(grts_bind)
## make data frame
grts_df <- data.frame(
  response = grts_bind$total_hg,
  x = grts_coords[, "X"],
  y = grts_coords[, "Y"],
  siteID = grts_bind$siteID,
  wgt = grts_bind$wgt
)
## head(grts_df)
summary(grts_df$response)
```

We then perform a design-based analysis on the GRTS sample.

```{r}
design_analysis <- cont_analysis(
  grts_df,
  siteID = "siteID",
  vars = "response",
  weight = "wgt",
  xcoord = "x",
  ycoord = "y"
)
## just return the mean info
## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
design_mean_grts <- design_analysis$Mean
design_mean_grts$Estimate
design_mean_grts$StdError
design_mean_grts$LCB95Pct
design_mean_grts$UCB95Pct
```

## GRTS Sample, Model Analysis

Finally, we use a model-based analysis on the GRTS sample.

```{r}
grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
  rename(total_hg = "response", xcoords = "x", ycoords = "y")
grts_unsamp <- anti_join(nla_nomiss, grts_coords_resp)
grts_unsamp$total_hg <- NA
grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
```

```{r}
grts_full$wts <- 1 / nrow(grts_full)
mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
                   xcoordcol = "xcoords",
                   ycoordcol = "ycoords")
pred_mod_grts <- predict(mod_grts, wtscol = "wts")
model_mean_grts <- pred_mod_grts$FPBK_Prediction
model_se_grts <- sqrt(pred_mod_grts$PredVar)
model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
model_mean_grts
model_se_grts
model_lb_grts
model_ub_grts
```

Combine all results:

```{r}
res_df <- tibble(approach = c("Design IRS", "Design GRTS", "Model IRS", "Model GRTS"),
       realized_mean = c(realized_mean, realized_mean, realized_mean,
                     realized_mean),
       estimate = c(irs_info$meandoc,
                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
                    as.vector(model_mean_grts)),
       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
              as.vector(model_se_grts)),
       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
              as.vector(model_lb_grts)),
       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
              as.vector(model_ub_grts)))
res_df$estimate
```

All 95% CI's contain the true mean for this one simulation.

For this simulation, in terms of standard error:

Model GRTS < Design GRTS < Model IRS < Design IRS

In terms of how close the prediction is to the true mean (prediction - true_mean), from closest to furthest:

Model IRS < Design IRS < Model GRTS < Design GRTS

```{r}
## irs sample
ggplot(lakesobs, aes(x = total_hg)) +
         geom_histogram(colour  = "black", fill = "white", bins = 20)

## grts sample
ggplot(grts_df, aes(x = response)) +
         geom_histogram(colour  = "black", fill = "white", bins = 20)
```

Candidates for symmetric variable.

```{r}
## disturbance for benthic microinvertebrates
ggplot(data = nla_df, aes(x = MMI_BENT_NLA12)) +
  geom_histogram(colour = "black", fill = "white", bins = 20)

## disturbance for zooplankton
ggplot(data = nla_df, aes(x = MMI_ZOOP_NLA6)) +
  geom_histogram(colour = "black", fill = "white", bins = 20)

## lakeshore disturbance indicator
ggplot(data = nla_df, aes(x = RDis_IX)) +
  geom_histogram(colour = "black", fill = "white", bins = 20)
```

But, does it make sense to want to find an average for any of these?

### 2021-07-28 inst/scripts/application2.Rmd
sloppy html file of application.
modified documentation and namespace so that package loads without errors, warnings, or notes.

@@
 library(skimr)
-nla_df <- read_csv(here("data", "nla12_keyvariables_data.csv"))
-skim(nla_df)
+load(here("data", "nla.rda"))
 ```
 
-```{r}
+```{r, results = "hide", eval = FALSE}
 View(nla_df)

### 2021-07-28 R/sim_pop.R

@@ -2,7 +2,7 @@
 #'
 #' simulates a spatially correlated response on either a gridded or
 #' random unit square. The resulting data can then be used in
-#' \code{\link{sim_one}()}.
+#' \code{\link{sim_trial}()}.
 #'
 #' @param N is the total number of data points. For \code{gridded = TRUE},
 #' this must be a perfect square.

### 2021-07-28 R/sim_trial.R

@@ -19,6 +19,8 @@
 #' @param nugget is the nugget.
 #' @param cortype_est is the (possibly misspecified) correlation
 #' function used to estimate/predict the mean, total, or other quantity.
+#' @param resptype is the response variable type, either \code{
+#' "normal"} or \code{"lognormal"}.
 #' @param ... further arguments passed to or from other methods.
 #' @return a data frame with \itemize{
 #'   \item \code{approach}, the name of th approach (\code{"Degign"}
@@ -36,6 +38,9 @@
 #' gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1,
 #' nugget = 0.2, cortype_est = "Exponential")
 #' @import stats
+#' @import sptotal
+#' @import spsurvey
+#' @importFrom rlang .data
 #' @export
 
 ## library(devtools)
@@ -86,17 +91,17 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
 
   # make data frame removing id and wgt
   irs_coords_resp <- irs_df %>%
-    select(-siteID, -wgt)
+    dplyr::select(-.data$siteID, -.data$wgt)
 
   # find unsampled sites
-  irs_unsamp <- anti_join(data, irs_coords_resp)
+  irs_unsamp <- dplyr::anti_join(data, irs_coords_resp)
 
   # make response NA for unsampled
   irs_unsamp$response <- NA
 
   # only keep necessary columns
   irs_unsamp <- irs_unsamp %>%
-    select(response, x, y)
+    dplyr::select(.data$response, .data$x, .data$y)
 
   # put them together
   irs_full <- dplyr::bind_rows(irs_coords_resp, irs_unsamp)
@@ -128,17 +133,17 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
 
   # make data frame removing id and wgt
   grts_coords_resp <- grts_df %>%
-    select(-siteID, -wgt)
+    dplyr::select(-.data$siteID, -.data$wgt)
 
   # find unsampled sites
-  grts_unsamp <- anti_join(data, grts_coords_resp)
+  grts_unsamp <- dplyr::anti_join(data, grts_coords_resp)
 
   # make response NA for unsampled
   grts_unsamp$response <- NA
 
   # only keep necessary columns
   grts_unsamp <- grts_unsamp %>%
-    select(response, x, y)
+    dplyr::select(.data$response, .data$x, .data$y)
 
   # put them together
   grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)

### 2021-08-03 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added application.

@@
 In the bottom row, we have two realizations of the same spatial process sampled at the same locations."}
-set.seed(06092021)
+set.seed(08032021)
 
@@
 
-```{r, results = "asis"}
+```{r, results = "asis", eval = FALSE}
 sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
@@
 
-## Applied Example
+# Application
 
-Potential Data Sets:
+The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
+We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
+Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
+Therefore, we will explore whether or not we can still obtain a relatively precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
 
-* National Lakes Assessment 
-* Moose in Alaska
-* Temperature Data from NOAA
+```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
+library(tidyverse)
+library(here)
+library(SpatialDesignVModel)
+data(nla)
+nla_df
+factor(nla_df$TOTALHG_UNITS) ## all same units
+
+nla_df %>% group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
+## HG never measured twice at the same site.
+
+nla_df %>% group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  summarise(sumnomiss = sum(nonmiss == 0))
+## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
+
+## in duplicate sites, fill in the value for RDis_IX
+nla_nomiss <- nla_df %>%
+  group_by(SITE_ID) %>%
+  fill(RDis_IX, .direction = "downup") %>%
+  ungroup() %>%
+  filter(!is.na(TOTALHG_RESULT)) %>%
+  rename(total_hg = "TOTALHG_RESULT")
+
+nla_nomiss_both <- nla_nomiss
+## check
+nla_nomiss_both %>% group_by(SITE_ID) %>% count() %>%
+  ungroup() %>%
+  summarise(not1 = sum(n != 1))
+
+## transform coordinates
+library(sptotal)
+coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
+                     lat = nla_nomiss_both$INDEX_LAT_DD, 
+                     lon = nla_nomiss_both$INDEX_LON_DD)
+nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
+nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
+```
+
+```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.
+Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.", message = FALSE, warning = FALSE}
+merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
+  geom_histogram(colour = "black", fill = "white", bins = 15) +
+  labs(x = "HG (ng / g)")
+# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
+#   geom_histogram(colour = "black", fill = "white", bins = 15) +
+#   labs(x = "Lakeshore Disturbance")
+merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
+  geom_point(aes(colour = total_hg)) +
+  scale_colour_viridis_c() +
+  labs(x = "", y = "", colour = "hg") +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks = element_blank()) +
+  coord_quickmap() 
+# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
+#   geom_point(aes(colour = RDis_IX)) +
+#   scale_colour_viridis_c() +
+#   labs(x = "", y = "", colour = "Dist") +
+#   theme(axis.text.x = element_blank(),
+#         axis.text.y = element_blank(),
+#         axis.ticks = element_blank())
+library(gridExtra)
+# grid.arrange(merc, merc_map, nrow = 2)
+merc_map
+```
+
+```{r, echo = FALSE, results = "hide"}
+mean(nla_nomiss_both$total_hg)
+# mean(nla_nomiss_both$RDis_IX)
+```
+
+Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
+Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
+Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.03 ng / g.
+
+```{r, results = "hide", fig.keep = "none"}
+## IRS Sample, Model Analysis
+
+set.seed(080421)
+n <- 100
+lakesobs <- nla_nomiss_both %>% sample_n(n) 
+lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
+lakesunobs$total_hg <- NA
+
+lakes_test <- bind_rows(lakesobs, lakesunobs) 
+lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
+
+slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
+                              data = lakes_test, 
+                              xcoordcol = 'xcoords',
+                              ycoordcol = 'ycoords',
+                              CorModel = "Exponential")
+summary(slmfitout_exp_lakes)
+plot(slmfitout_exp_lakes)
+
+pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
+                          conf_level = 0.95)
+print(pred_exp_lakes)
+mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
+se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
+lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
+ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
+
+realized_mean <- mean(nla_nomiss_both$total_hg) ## compute realized mean
+```
+
+```{r, results = "hide"}
+## IRS Sample, Design Analysis
+## 
+irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
+                                   vardoc = var(total_hg))
+N <- nrow(nla_nomiss_both); n <- nrow(lakesobs)
+irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
+irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
+irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
+irs_info$meandoc; sqrt(irs_info$vardoc)
+irs_lb_irs_samp; irs_ub_irs_samp
+```
+
+```{r, results = "hide"}
+## GRTS Sample, Design Analysis
+data_sf <- sf::st_as_sf(nla_nomiss_both, coords = c("xcoords", "ycoords"),
+                        crs = 5070)
+library(spsurvey)
+grts_samp <- grts(data_sf, n_base = n)
+grts_bind <- sprbind(grts_samp)
+
+## get coordinates
+grts_coords <- sf::st_coordinates(grts_bind)
+## make data frame
+grts_df <- data.frame(
+  response = grts_bind$total_hg,
+  x = grts_coords[, "X"],
+  y = grts_coords[, "Y"],
+  siteID = grts_bind$siteID,
+  wgt = grts_bind$wgt
+)
+## head(grts_df)
+summary(grts_df$response)
+
+design_analysis <- cont_analysis(
+  grts_df,
+  siteID = "siteID",
+  vars = "response",
+  weight = "wgt",
+  xcoord = "x",
+  ycoord = "y"
+)
+## just return the mean info
+## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
+design_mean_grts <- design_analysis$Mean
+design_mean_grts$Estimate
+design_mean_grts$StdError
+design_mean_grts$LCB95Pct
+design_mean_grts$UCB95Pct
+```
+
+```{r, results = "hide"}
+## GRTS Sample, Model Analysis
+
+grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
+  rename(total_hg = "response", xcoords = "x", ycoords = "y")
+grts_unsamp <- anti_join(nla_nomiss_both, grts_coords_resp)
+grts_unsamp$total_hg <- NA
+grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
+grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
+
+grts_full$wts <- 1 / nrow(grts_full)
+mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
+                   xcoordcol = "xcoords",
+                   ycoordcol = "ycoords")
+pred_mod_grts <- predict(mod_grts, wtscol = "wts")
+model_mean_grts <- pred_mod_grts$FPBK_Prediction
+model_se_grts <- sqrt(pred_mod_grts$PredVar)
+model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
+model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
+model_mean_grts
+model_se_grts
+model_lb_grts
+model_ub_grts
+```
+
+```{r, results = "hide"}
+## combine results
+res_df <- tibble(approach = c("Design IRS", "Design GRTS", "Model IRS", "Model GRTS"),
+       realized_mean = c(realized_mean, realized_mean, realized_mean,
+                     realized_mean),
+       estimate = c(irs_info$meandoc,
+                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
+                    as.vector(model_mean_grts)),
+       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
+              as.vector(model_se_grts)),
+       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
+              as.vector(model_lb_grts)),
+       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
+              as.vector(model_ub_grts)))
+res_df <- res_df %>% slice(1, 3, 2, 4)
+```
+
+```{r apptab}
+library(knitr)
+names(res_df) <- c("Approach", "Realized Mean", "Estimate", "SE", "95% LB", "95% UB")
+kable(res_df, digits = 1, caption = "\\label{tab:appliedtab} Table XXX. Application of design-based and model-based approaches to the NLA data set on mercury concentration.")
+```
+
+Table \ref{tab:appliedtab} shows the application of a design-based analysis on an IRS, a model-based analysis on an IRS, a design-based analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
+We see that, for all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
+However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
+
+But, we do note a couple of patterns.
+The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not use the spatial correlation in mercury concentration across the contiguous United States.
+We also see that, for the samples drawn, the both analyses with the GRTS sampling design have a lower standard error than the analyses with the IRS sampling design.
+We would expect this to be the case for most samples because mercury concentration exhibits spatial correlation so a spatially balanced sample should usually yield a lower standard error.
+If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling just 50 lakes to save time and money.
 

### 2021-08-03 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added application.

@@
 \bibliographystyle{elsarticle-harv}
+\usepackage{longtable}
 \ifxetex
@@
 % Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-  \title{A comparison of design-based and model-based approaches for spatial
-data.}
+  \title{A comparison of design-based and model-based approaches for
+spatial data.}
     \author[USEPA]{In alphabetical order Michael Dumelle\corref{1}}
@@
 locations to model-based approaches (Brus and De Gruijter, 1997; Ver
-Hoef, 2002, 2008). Cooper (2006) review the two approaches in an
-ecological context before introducing a ``model-assisted'' variance
+Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two approaches in
+an ecological context before introducing a ``model-assisted'' variance
 estimator that combines aspects from each approach. In addition to
 Cooper (2006), there has been substantial research and development into
-estimators that use both design and model-based principles (see
-e.g.~Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
+estimators that use both design and model-based principles (see e.g.
+Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
 Bayesian approach, and Sterba (2009)). More recent overviews include
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrrrr}
-  \hline
-Approach & Bias & RMSE & MedAE & Coverage & PClose & MedIL \\ 
-  \hline
-Design & 0.0003 & 0.0353 & 0.0251 & 0.9461 & 0.4889 & 0.1362 \\ 
-  Model & -0.0001 & 0.0362 & 0.0253 & 0.9480 & 0.5111 & 0.1430 \\ 
-   \hline
-\end{tabular}
-\caption{Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)}
-\label{tab:sim_one}
-\end{table}
-
 \textbf{Base Simulations}
@@
 
-\hypertarget{applied-example}{%
-\subsection{Applied Example}\label{applied-example}}
+\hypertarget{application}{%
+\section{Application}\label{application}}
 
-Potential Data Sets:
+The Environmental Protection Agency (EPA), states, and tribes
+periodically conduct National Aquatic Research Surveys (NARS) in the
+United States to assess the water quality of various bodies of water. We
+will use the 2012 National Lakes Assessment (NLA), which measures
+various aspects of lake health and quality in lakes in the contiguous
+United States, to obtain an interval for mean mercury concentration.
+Although all lakes in the survey were measured in 2012, there may not
+always be enough time or money to do so. Therefore, we will explore
+whether or not we can still obtain a relatively precise estimate for the
+realized mean mercury concentration if we only take a sample of 100 of
+the 986 lakes.
 
-\begin{itemize}
-\tightlist
-\item
-  National Lakes Assessment
-\item
-  Moose in Alaska
-\item
-  Temperature Data from NOAA
-\end{itemize}
+\begin{figure}
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.
+Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.}\label{fig:figdata}
+\end{figure}
+
+Figure \ref{fig:figdata} shows that mercury concentration is
+right-skewed, with most lakes having a low value of mercury
+concentration but a few having a much higher concentration. Mercury
+concentration exhibits some spatial correlation, with high mercury
+concentrations in lakes in the northeast and north central United
+States. Because we are considering these lakes to be our entire
+population, we know that the realized mean mercury concentration is
+103.03 ng / g.
+
+\begin{longtable}[]{@{}lrrrrr@{}}
+\caption{\label{tab:appliedtab} Table XXX. Application of design-based
+and model-based approaches to the NLA data set on mercury
+concentration.}\tabularnewline
+\toprule
+Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+\midrule
+\endfirsthead
+\toprule
+Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+\midrule
+\endhead
+Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\
+Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\
+Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\
+Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\
+\bottomrule
+\end{longtable}
+
+Table \ref{tab:appliedtab} shows the application of a design-based
+analysis on an IRS, a model-based analysis on an IRS, a design-based
+analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
+We see that, for all four analyses, the true realized mean mercury
+concentration is within the bounds of the 95\% intervals. However, we
+should not generalize the results of this particular realization to any
+other data set or even to other potential samples of this data set.
+
+But, we do note a couple of patterns. The design-based IRS analysis
+shows the largest standard error: a likely reason is that this is the
+only approach that does not use the spatial correlation in mercury
+concentration across the contiguous United States. We also see that, for
+the samples drawn, the both analyses with the GRTS sampling design have
+a lower standard error than the analyses with the IRS sampling design.
+We would expect this to be the case for most samples because mercury
+concentration exhibits spatial correlation so a spatially balanced
+sample should usually yield a lower standard error. If it is acceptable
+to have an interval for mean mercury concentration of about 25 ng / g
+and if we ignore the other variables that the EPA collects information
+on in these NLA surveys, then the EPA could consider sampling just 50
+lakes to save time and money.
 

### 2021-08-03 inst/scripts/application2.Rmd
added application.

@@
 
+IGNORE THIS FILE
+Relevant work was tranferred to the main document.
+
 ```{r}
@@
 load(here("data", "nla.rda"))
-```
-
-```{r, results = "hide", eval = FALSE}
-View(nla_df)
-factor(nla_df$TOTALHG_UNITS) ## all same units
-```
-
-```{r}
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-```
-
-HG never measured twice at the same site. 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-
-```{r}
-nla_nomiss <- nla_df %>% filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-## check
-nla_nomiss %>% group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-```
-
-Exploratory graph
-
-```{r}
-ggplot(data = nla_nomiss, aes(x = total_hg)) +
-  geom_histogram(bins = 20)
 ```
@@
 
-transform coordinates
+```{r}
 
-```{r}
-coord_list <- LLtoTM(cm = mean(nla_nomiss$INDEX_LON_DD), lat = nla_nomiss$INDEX_LAT_DD,
-       lon = nla_nomiss$INDEX_LON_DD)
-nla_nomiss$xcoords <- coord_list$xy[ ,1]
-nla_nomiss$ycoords <- coord_list$xy[ ,2]
-ggplot(data = nla_nomiss, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c()
-```
-
-## IRS Sample, Model Analysis
-
-```{r}
-set.seed(07262021)
-lakesobs <- nla_nomiss %>% sample_n(200) ## sample 200 lakes
-lakesunobs <- anti_join(nla_nomiss, lakesobs)
-lakesunobs$total_hg <- NA
-
-lakes_test <- bind_rows(lakesobs, lakesunobs) 
-lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
 ```
@@
 ```{r}
-slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
-                              data = lakes_test, 
-                              xcoordcol = 'xcoords',
-                              ycoordcol = 'ycoords',
-                              CorModel = "Exponential")
-summary(slmfitout_exp_lakes)
-plot(slmfitout_exp_lakes)
 ```
-
-```{r}
-pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
-                          conf_level = 0.95)
-print(pred_exp_lakes)
-mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
-se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
-lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
-ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
-
-realized_mean <- mean(nla_nomiss$total_hg) ## compute realized mean
-```
-
-## IRS Sample, Design Analysis
-
-```{r}
-irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
-                                   vardoc = var(total_hg))
-N <- nrow(nla_nomiss); n <- nrow(lakesobs)
-irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
-irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
-irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
-irs_info$meandoc; sqrt(irs_info$vardoc)
-irs_lb_irs_samp; irs_ub_irs_samp
-```
-
-## GRTS Sample, Design Analysis
 
@@
 set.seed(07262021)
-data_sf <- sf::st_as_sf(nla_nomiss, coords = c("xcoords", "ycoords"),
-                        crs = 5070)
-library(spsurvey)
-grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sprbind(grts_samp)
 
-## get coordinates
-grts_coords <- sf::st_coordinates(grts_bind)
-## make data frame
-grts_df <- data.frame(
-  response = grts_bind$total_hg,
-  x = grts_coords[, "X"],
-  y = grts_coords[, "Y"],
-  siteID = grts_bind$siteID,
-  wgt = grts_bind$wgt
-)
-## head(grts_df)
-summary(grts_df$response)
 ```
@@
 
-```{r}
-design_analysis <- cont_analysis(
-  grts_df,
-  siteID = "siteID",
-  vars = "response",
-  weight = "wgt",
-  xcoord = "x",
-  ycoord = "y"
-)
-## just return the mean info
-## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
-design_mean_grts <- design_analysis$Mean
-design_mean_grts$Estimate
-design_mean_grts$StdError
-design_mean_grts$LCB95Pct
-design_mean_grts$UCB95Pct
-```
-
-## GRTS Sample, Model Analysis
-
-Finally, we use a model-based analysis on the GRTS sample.
-
-```{r}
-grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
-  rename(total_hg = "response", xcoords = "x", ycoords = "y")
-grts_unsamp <- anti_join(nla_nomiss, grts_coords_resp)
-grts_unsamp$total_hg <- NA
-grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
-grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
-```
-
-```{r}
-grts_full$wts <- 1 / nrow(grts_full)
-mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
-                   xcoordcol = "xcoords",
-                   ycoordcol = "ycoords")
-pred_mod_grts <- predict(mod_grts, wtscol = "wts")
-model_mean_grts <- pred_mod_grts$FPBK_Prediction
-model_se_grts <- sqrt(pred_mod_grts$PredVar)
-model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
-model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
-model_mean_grts
-model_se_grts
-model_lb_grts
-model_ub_grts
-```
-
 Combine all results:
@@
 ```{r}
-res_df <- tibble(approach = c("Design IRS", "Design GRTS", "Model IRS", "Model GRTS"),
-       realized_mean = c(realized_mean, realized_mean, realized_mean,
-                     realized_mean),
-       estimate = c(irs_info$meandoc,
-                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
-                    as.vector(model_mean_grts)),
-       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
-              as.vector(model_se_grts)),
-       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
-              as.vector(model_lb_grts)),
-       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
-              as.vector(model_ub_grts)))
-res_df$estimate
+
 ```

### 2021-08-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
wrote functions to allow a more easy comparison between grid and random locations, and added simulation thoughts to manuscript.

@@
 
+```{r, results = "hide", message = FALSE, warning = FALSE}
+library(tidyverse)
+files <- list.files(path = here("inst", "output"), pattern = "*.csv",
+                    full.names = TRUE)
+
+combo_data <- purrr::map_df(files,
+                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
+  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
+           sep = "_") %>%
+  select(-junk)
+combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))
+
+combo_data_grts <- combo_data %>% filter(approach %in% c("Design GRTS", "Model GRTS"))
+
+dep_only <- combo_data_grts %>% filter(error_dep == "deperr")
+dep_long <- dep_only %>%
+  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
+  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
+```
+
+Major Points from August 3 Simulations:
+
+1.
+In most of the dependent error simulation settings, either all four approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model) perform equally or GRTS-Design and GRTS-Model outperform IRS-Design and IRS-Model.
+Exceptions to this are a couple of the settings with very small sample sizes (n = 10), in which the IRS does better than GRTS.
+In the independent error settings, it usually doesn't matter much which approach is used, which makes sense.
+
+```{r}
+ggplot(data = combo_data, aes(x = sim, y = rmspe, colour = approach)) +
+  geom_point() +
+  coord_flip()
+```
+
+2.
+We will now focus in a bit more on comparing Design-GRTS to Model-GRTS, the two best approaches for any reasonable sample size.
+In the independent error settings, the two approaches perform very similarly, so those results are omitted in the following graph.
+In the dependent error settings, using rmspe as the performance criterion, Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two approaches perform very similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3 settings.
+
+```{r}
+ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
+  geom_point() +
+  coord_flip() +
+  labs(y = "Percent Change in rmspe (Design - Model) / Model",
+       caption = "Percent > 0 means Model has better rmspe")
+```
+
+3.
+Focusing in on the three settings where Design-GRTS outperforms Model-GRTS, we see that, in two of the settings, the log-normal response has a large variance, corresponding to a large right-skew after exponentiation.
+All three settings have sites in random locations.
+However, in only one of these settings would we recommend actually using Design-GRTS.
+In the other two settings, the data are sufficiently skewed that a practitioner should not use either approach, though it is "safer" to use Design-GRTS.
+
+```{r}
+design_better <- dep_long %>% filter(perc < -5)
+design_better_full <- inner_join(combo_data_grts, design_better, by = c("sim"))
+ggplot(data = design_better_full, aes(x = sim, y = coverage, colour = approach)) +
+  geom_point() +
+  coord_flip() +
+  geom_hline(aes(yintercept = 0.95), alpha = 0.5)
+```
+
+4. Take-home messages
+
+- In terms of rmspe, a model-based analysis on a GRTS design yields an rmspe similar to or lower than a design-based analysis on a GRTS design, as long as the response variable is not "too skewed." 
+- If the response variable is very skewed, then neither analysis is appropriate, but, the design-based analysis is quicker. 
+- a spatially balanced GRTS sample outperforms IRS in nearly all dependent error settings, as expected.
+- methods that use spatial correlation generally perform better on random location points than they do on gridded points.
+This makes some intuitive sense because (1) on average, the minimum distance between an unobserved point and its nearest observed neighbor should be lower for random points and (2) the span of the study area is maximized for a grid based on the way that we set up the simulations (with the random points being drawn as uniform random variables within the boundary of the grid).
+- comparison of Design-GRTS and Model-GRTS between two settings with different locations of points, but otherwise the same simulation parameters, should really be done on the same surface realization.
+One very strange realized response vector could drastically alter the results, especially on the exponentiated log data.
+In the same way that we compare the four approaches on the same realized data, we should also try to do the same with the locations, if they are of interest.
+(The realized mean won't be exactly the same but should be close).
+
 __Sample Simulation__

### 2021-08-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
wrote functions to allow a more easy comparison between grid and random locations, and added simulation thoughts to manuscript.

@@
 \usepackage{longtable}
+\usepackage{graphicx}
 \ifxetex
@@
 
+Major Points from August 3 Simulations:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  In most of the dependent error simulation settings, either all four
+  approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model)
+  perform equally or GRTS-Design and GRTS-Model outperform IRS-Design
+  and IRS-Model. Exceptions to this are a couple of the settings with
+  very small sample sizes (n = 10), in which the IRS does better than
+  GRTS. In the independent error settings, it usually doesn't matter
+  much which approach is used, which makes sense.
+\end{enumerate}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  We will now focus in a bit more on comparing Design-GRTS to
+  Model-GRTS, the two best approaches for any reasonable sample size. In
+  the independent error settings, the two approaches perform very
+  similarly, so those results are omitted in the following graph. In the
+  dependent error settings, using rmspe as the performance criterion,
+  Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two
+  approaches perform very similarly in 3 settings, and Design-GRTS
+  outperforms Model-GRTS in 3 settings.
+\end{enumerate}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Focusing in on the three settings where Design-GRTS outperforms
+  Model-GRTS, we see that, in two of the settings, the log-normal
+  response has a large variance, corresponding to a large right-skew
+  after exponentiation. All three settings have sites in random
+  locations. However, in only one of these settings would we recommend
+  actually using Design-GRTS. In the other two settings, the data are
+  sufficiently skewed that a practitioner should not use either
+  approach, though it is ``safer'' to use Design-GRTS.
+\end{enumerate}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Take-home messages
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  In terms of rmspe, a model-based analysis on a GRTS design yields an
+  rmspe similar to or lower than a design-based analysis on a GRTS
+  design, as long as the response variable is not ``too skewed.''
+\item
+  If the response variable is very skewed, then neither analysis is
+  appropriate, but, the design-based analysis is quicker.
+\item
+  a spatially balanced GRTS sample outperforms IRS in nearly all
+  dependent error settings, as expected.
+\item
+  methods that use spatial correlation generally perform better on
+  random location points than they do on gridded points. This makes some
+  intuitive sense because (1) on average, the minimum distance between
+  an unobserved point and its nearest observed neighbor should be lower
+  for random points and (2) the span of the study area is maximized for
+  a grid based on the way that we set up the simulations (with the
+  random points being drawn as uniform random variables within the
+  boundary of the grid).
+\item
+  comparison of Design-GRTS and Model-GRTS between two settings with
+  different locations of points, but otherwise the same simulation
+  parameters, should really be done on the same surface realization. One
+  very strange realized response vector could drastically alter the
+  results, especially on the exponentiated log data. In the same way
+  that we compare the four approaches on the same realized data, we
+  should also try to do the same with the locations, if they are of
+  interest. (The realized mean won't be exactly the same but should be
+  close).
+\end{itemize}
+
 \textbf{Sample Simulation}

### 2021-08-04 R/sim_pop.R

@@ -83,6 +83,7 @@ sim_pop <- function(N = 100, gridded = TRUE,
 #' @param erange is the effective range.
 #' @param nugget is the nugget effect.
 #' @return Covariance Matrix
+#' @export
 
 #' @describeIn covmx_exp Exponential Covariance
 covmx_exp <- function(distmx, psill, erange, nugget) {

### 2021-08-04 inst/scripts/simulation_exploration.R

@@ -1,3 +1,5 @@
+library(tidyverse)
+library(here)
 read_csv(here("inst", "output", "sim_deperr_lognormal_n50_gridloc.csv"))
 
 files <- list.files(path = here("inst", "output"), pattern = "*.csv",
@@ -44,10 +46,20 @@ ggplot(data = combo_data_grts, aes(x = sim, y = rmspe, colour = approach)) +
   geom_point() +
   coord_flip()
 
+dep_only <- combo_data_grts %>% filter(error_dep == "deperr")
+dep_long <- dep_only %>%
+  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
+  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
+ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
+  geom_point() +
+  coord_flip() +
+  labs(x = "Percent Change in rmspe (Design - Model) / Model",
+       caption = "Percent > 0 means Model has better rmspe")
+
 combo_perc <- combo_data_grts %>%
   pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
   mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
-ggplot(data = combo_perc, aes(x = sim, y = perc)) +
+ggplot(data = combo_perc, aes(x = fct_reorder(sim, perc), y = perc)) +
   geom_point() +
   coord_flip()
 ## investigate discrepancy in grid vs random
@@ -64,5 +76,21 @@ ggplot(data = combo_perc, aes(x = sim, y = perc)) +
 ggplot(data = combo_data_grts, aes(x = sim, y = coverage, colour = approach)) +
   geom_point() +
   coord_flip()
+
+
+
+## all
+
+grid_df <- read_csv(here("inst", "all_output", "sim_deperr_lognormalbig_n150_gridloc.csv"))
+grid_df$location <- "grid"
+rand_df <- read_csv(here("inst", "all_output", "sim_deperr_lognormalbig_n150_randloc.csv"))
+rand_df$location <- "rand"
+
+both_df <- bind_rows(grid_df, rand_df)
+both_df %>% group_by(location) %>%
+  summarise(max(true_mean))
+ggplot(data = both_df, aes(x = true_mean)) +
+  geom_histogram(bins = 15) +
+  facet_wrap(~ location)
 
 

### 2021-08-06 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added mae.

@@
   mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
+
+dep_long_abs <- dep_only %>%
+  pivot_wider(names_from = approach, values_from = medae, sim) %>%
+  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
 ```
@@
 ```{r}
+ggplot(data = dep_long_abs, aes(x = fct_reorder(sim, perc), y = perc)) +
+  geom_point() +
+  coord_flip() +
+  labs(y = "Percent Change in medae (Design - Model) / Model",
+       caption = "Percent > 0 means Model has better rmspe")
+
 ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
@@
 ```{r}
-design_better <- dep_long %>% filter(perc < -5)
+design_better <- dep_long %>% dplyr::filter(perc < -5)
 design_better_full <- inner_join(combo_data_grts, design_better, by = c("sim"))

### 2021-08-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
coverage and pros/cons updates

@@
 
+# mike chunk
+# combo_data <- purrr::map_df(files,
+#                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
+#   separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
+#            sep = "_") %>%
+#   select(-junk, -junk2)
+
+# matt chunk
 combo_data <- purrr::map_df(files,
@@
   select(-junk)
+
 combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))
@@
 
-4. Take-home messages
+4. Coverage
+
+For Gaussian errors, coverage for all approaches tended to be near 0.95.
+There was less between-approach deviation in coverages for random locations compared to grid locations.
+Generally, the larger the skew, the worse the coverage,and the larger the sample size, the better the coverage.
+Design GRTS (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
+
+```{r}
+ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
+  geom_hline(yintercept = 0.95) +
+  geom_point() +
+  coord_flip()
+```
+
+5. Take-home messages
 
@@
 ```
-
-__Base Simulations__
-
-* both good: correctly specified model with high correlation (we did this in Table \ref{tab:sim_one})
-* break model: highly non-normal errors with small sample size
-* break design: small area estimation
-
-__Simulation Discussion Questions__
-
-* model-based: how should sample be drawn? should locations be fixed?
-* change n or sampling fraction?
-
-__Other Base Settings?__
-
-* both good?: misspecified covariance model with high correlation
-* break both? non-gaussian areas with smaller sample size
 
@@
 
+* Pros of Design-Based (items we are not exploring): computationally efficient, few assumptions, more naturally handles binary data,
+
+* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?
+
 # References {#references .unnumbered}

### 2021-08-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
knitted writeup
coverage and pros/cons updates

@@
 % Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-  \title{A comparison of design-based and model-based approaches for
-spatial data.}
+  \title{A comparison of design-based and model-based approaches for spatial
+data.}
     \author[USEPA]{In alphabetical order Michael Dumelle\corref{1}}
@@
 locations to model-based approaches (Brus and De Gruijter, 1997; Ver
-Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two approaches in
-an ecological context before introducing a ``model-assisted'' variance
+Hoef, 2002, 2008). Cooper (2006) review the two approaches in an
+ecological context before introducing a ``model-assisted'' variance
 estimator that combines aspects from each approach. In addition to
 Cooper (2006), there has been substantial research and development into
-estimators that use both design and model-based principles (see e.g.
-Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
+estimators that use both design and model-based principles (see
+e.g.~Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
 Bayesian approach, and Sterba (2009)). More recent overviews include
@@
 \includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-2.pdf}
 
@@
 \setcounter{enumi}{3}
+\tightlist
+\item
+  Coverage
+\end{enumerate}
+
+For Gaussian errors, coverage for all approaches tended to be near 0.95.
+There was less between-approach deviation in coverages for random
+locations compared to grid locations. Generally, the larger the skew,
+the worse the coverage,and the larger the sample size, the better the
+coverage. Design GRTS (local neighborhood variance) tended to slightly
+undercover, a result Tony was familiar with.
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
 \tightlist
@@
 
-\textbf{Base Simulations}
-
-\begin{itemize}
-\tightlist
-\item
-  both good: correctly specified model with high correlation (we did
-  this in Table \ref{tab:sim_one})
-\item
-  break model: highly non-normal errors with small sample size
-\item
-  break design: small area estimation
-\end{itemize}
-
-\textbf{Simulation Discussion Questions}
-
-\begin{itemize}
-\tightlist
-\item
-  model-based: how should sample be drawn? should locations be fixed?
-\item
-  change n or sampling fraction?
-\end{itemize}
-
-\textbf{Other Base Settings?}
-
-\begin{itemize}
-\tightlist
-\item
-  both good?: misspecified covariance model with high correlation
-\item
-  break both? non-gaussian areas with smaller sample size
-\end{itemize}
-
 \hypertarget{software}{%
@@
 \toprule
-Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Realized Mean & Estimate & SE & 95\% LB & 95\%
+UB\tabularnewline
 \midrule
@@
 \toprule
-Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Realized Mean & Estimate & SE & 95\% LB & 95\%
+UB\tabularnewline
 \midrule
 \endhead
-Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\
-Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\
-Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\
-Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\
+Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9\tabularnewline
+Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9\tabularnewline
+Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7\tabularnewline
+Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9\tabularnewline
 \bottomrule
@@
 
+\begin{itemize}
+\item
+  Pros of Design-Based (items we are not exploring): computationally
+  efficient, few assumptions, more naturally handles binary data,
+\item
+  Pros of Model-Based (items we are not exploring): covariate inference,
+  more efficient small-area estimation, model selection?
+\end{itemize}
+
 \hypertarget{references}{%

### 2021-08-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor updates

@@
 
+There were several variables to alter in the simulations, and the names of the scenarios in future plots mirror these variables
+
+* correlation type: dependent errors or independent errors
+* error type:
+    * normal: mean 0, variance 2
+    * lognormal: log scale mean 0, log scale variance 2
+    * lognormalbig: log scale mean 0, log scale variance 4
+* sample sizes: $n = 10, 50, 150; N = 900$
+* layout: gridded vs random population locations
+
+So for example, the `inderror.normal.n50.randloc` is the simulation having independent random errors that are normal, a sample size of 50, and random population locations.
+
+In each simulation, a GRTS sample and an IRS sample were selected.
+Then for the GRTS sample, the design-based approach using the local neighborhood variance (Design GRTS) and a model-based approach were applied (Model GRTS).
+Then for the IRS sample, the design-based approach using the simple random sample variance (Design IRS) and a model-based approach were applied (Model IRS).
+
 ```{r, results = "hide", message = FALSE, warning = FALSE}
@@
 # mike chunk
+combo_data <- purrr::map_df(files,
+                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
+  separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
+           sep = "_") %>%
+  select(-junk, -junk2)
+
+# matt chunk
 # combo_data <- purrr::map_df(files,
 #                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
-#   separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
+#   separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
 #            sep = "_") %>%
-#   select(-junk, -junk2)
-
-# matt chunk
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
-  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
-           sep = "_") %>%
-  select(-junk)
+#   select(-junk)
 

### 2021-08-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor updates

@@
 \section{Numerical Study}\label{sec:numstudy}}
+
+There were several variables to alter in the simulations, and the names
+of the scenarios in future plots mirror these variables
+
+\begin{itemize}
+\tightlist
+\item
+  correlation type: dependent errors or independent errors
+\item
+  error type:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    normal: mean 0, variance 2
+  \item
+    lognormal: log scale mean 0, log scale variance 2
+  \item
+    lognormalbig: log scale mean 0, log scale variance 4
+  \end{itemize}
+\item
+  sample sizes: \(n = 10, 50, 150; N = 900\)
+\item
+  layout: gridded vs random population locations
+\end{itemize}
+
+So for example, the \texttt{inderror.normal.n50.randloc} is the
+simulation having independent random errors that are normal, a sample
+size of 50, and random population locations.
+
+In each simulation, a GRTS sample and an IRS sample were selected. Then
+for the GRTS sample, the design-based approach using the local
+neighborhood variance (Design GRTS) and a model-based approach were
+applied (Model GRTS). Then for the IRS sample, the design-based approach
+using the simple random sample variance (Design IRS) and a model-based
+approach were applied (Model IRS).
 

### 2021-08-12 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
figure pdfs.

@@
 # mike chunk
+# Mike: see if this works. If so, delete one of the duplicates
 combo_data <- purrr::map_df(files,
                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
-  separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
+  mutate(filetemp = str_extract(filename, pattern = "sim_\\w+")) %>% 
+  separate(filetemp, into = c("junk", "error_dep", "dist", "n", "location"),
            sep = "_") %>%
-  select(-junk, -junk2)
+  select(-junk)
+
+# combo_data <- purrr::map_df(files,
+#                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
+#   separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
+#            sep = "_") %>%
+#   select(-junk, -junk2)
 
 # matt chunk
-# combo_data <- purrr::map_df(files,
-#                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
-#   separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
-#            sep = "_") %>%
-#   select(-junk)
+combo_data <- purrr::map_df(files,
+                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
+  mutate(filetemp = str_extract(filename, pattern = "sim_\\w+")) %>% 
+  separate(filetemp, into = c("junk", "error_dep", "dist", "n", "location"),
+           sep = "_") %>%
+  select(-junk)
 
@@
 
-combo_data_grts <- combo_data %>% filter(approach %in% c("Design GRTS", "Model GRTS"))
+combo_data_grts <- combo_data %>% dplyr::filter(approach %in% c("Design GRTS", "Model GRTS"))
 
-dep_only <- combo_data_grts %>% filter(error_dep == "deperr")
+dep_only <- combo_data_grts %>% dplyr::filter(error_dep == "deperr")
 dep_long <- dep_only %>%

### 2021-08-12 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
figure pdfs.

--- 2021-08-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
@@
 
+There were several variables to alter in the simulations, and the names
+of the scenarios in future plots mirror these variables
+
+\begin{itemize}
+\tightlist
+\item
+  correlation type: dependent errors or independent errors
+\item
+  error type:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    normal: mean 0, variance 2
+  \item
+    lognormal: log scale mean 0, log scale variance 2
+  \item
+    lognormalbig: log scale mean 0, log scale variance 4
+  \end{itemize}
+\item
+  sample sizes: \(n = 10, 50, 150; N = 900\)
+\item
+  layout: gridded vs random population locations
+\end{itemize}
+
+So for example, the \texttt{inderror.normal.n50.randloc} is the
+simulation having independent random errors that are normal, a sample
+size of 50, and random population locations.
+
+In each simulation, a GRTS sample and an IRS sample were selected. Then
+for the GRTS sample, the design-based approach using the local
+neighborhood variance (Design GRTS) and a model-based approach were
+applied (Model GRTS). Then for the IRS sample, the design-based approach
+using the simple random sample variance (Design IRS) and a model-based
+approach were applied (Model IRS).
+
 Major Points from August 3 Simulations:
@@
 with REML.
-
-\textbf{Base Simulations}
-
-\begin{itemize}
-\tightlist
-\item
-  both good: correctly specified model with high correlation (we did
-  this in Table \ref{tab:sim_one})
-\item
-  break model: highly non-normal errors with small sample size
-\item
-  break design: small area estimation
-\end{itemize}
-
-\textbf{Simulation Discussion Questions}
-
-\begin{itemize}
-\tightlist
-\item
-  model-based: how should sample be drawn? should locations be fixed?
-\item
-  change n or sampling fraction?
-\end{itemize}
-
-\textbf{Other Base Settings?}
-
-\begin{itemize}
-\tightlist
-\item
-  both good?: misspecified covariance model with high correlation
-\item
-  break both? non-gaussian areas with smaller sample size
-\end{itemize}
 
@@
 \section{Discussion}\label{sec:discussion}}
+
+\begin{itemize}
+\item
+  Pros of Design-Based (items we are not exploring): computationally
+  efficient, few assumptions, more naturally handles binary data,
+\item
+  Pros of Model-Based (items we are not exploring): covariate inference,
+  more efficient small-area estimation, model selection?
+\end{itemize}
 

### 2021-08-12 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added a couple more discussion points.

@@
 
-* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?
+* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?, estimated spatial parameters to better understand spatial structure, site-by-site predictions/prediction map
 

### 2021-08-12 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added a couple more discussion points.

@@
   Pros of Model-Based (items we are not exploring): covariate inference,
-  more efficient small-area estimation, model selection?
+  more efficient small-area estimation, model selection?, estimated
+  spatial parameters to better understand spatial structure,
+  site-by-site predictions/prediction map
 \end{itemize}

### 2021-08-13 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
minor updates

--- 2021-08-11 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
@@
   \item
-    lognormal: log scale mean 0, log scale variance 2
+    lognormal: log scale mean 0, log scale variance 2 (total variance
+    47)
   \item
-    lognormalbig: log scale mean 0, log scale variance 4
+    lognormalbig: log scale mean 0, log scale variance 4 (total variance
+    2,926)
   \end{itemize}
@@
 
+There were 2000 trials for each simulation. The original response
+(before exponentiating if applicable) for the dependent error cases was
+normally distributed with an exponential covariance function with
+partial sill of 0.9, effective range of \(\sqrt{2}\), and a nugget of
+0.1. For the independent error cases, the partial sill was 0 and the
+nugget was 1.
+
 In each simulation, a GRTS sample and an IRS sample were selected. Then
@@
 approach were applied (Model IRS).
+
+The GRTS algorithm and the local neighborhood variance estimator are
+available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
+et al., 2021). FPBK can be readily performed in \texttt{R} with the
+\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
+for both the simulation analysis and the application, estimating
+parameters with Restricted Maximum Likelihood (REML).
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-2.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
   If the response variable is very skewed, then neither analysis is
-  appropriate, but, the design-based analysis is quicker.
+  appropriate, but, the design-based analysis is better.
 \item
@@
 \end{itemize}
-
-\textbf{Sample Simulation}
-
-For the following simulation results, we simulated 1040 different
-gridded populations, each of size 900 (on the unit square) with sample
-size 150. For the design-based approach, population units were selected
-via GRTS, the Horvitz-Thompson estimator was used,and the local mean
-variance was used. For the model-based approach (FPBK), population units
-were selected via Independent Random Sampling (IRS) and the appropriate
-prediction and prediction variance formulas were used.
-
-The response was normally distributed with an exponential covariance
-function with partial sill of 0.9, effective range of \(\sqrt{2}\), and
-a nugget of 0.1. For model-based, we assumed the correct form of the
-covariance function (exponential), but estimated the spatial parameters
-with REML.
-
-\hypertarget{software}{%
-\subsection{Software}\label{software}}
-
-The GRTS algorithm and the local neighborhood variance estimator are
-available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
-et al., 2021). FPBK can be readily performed in \texttt{R} with the
-\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
-for both the simulation analysis and the application, estimating
-parameters with Restricted Maximum Likelihood (REML).
 

### 2021-08-16 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd

@@
     * normal: mean 0, variance 2
-    * lognormal: log scale mean 0, log scale variance 2
-    * lognormalbig: log scale mean 0, log scale variance 4
+    * lognormal: log scale mean 0, log scale variance 2 (total variance 47)
+    * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926)
 * sample sizes: $n = 10, 50, 150; N = 900$
@@
 
+There were 2000 trials for each simulation.
+The original response (before exponentiating if applicable) for the dependent error cases was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1.
+For the independent error cases, the partial sill was 0 and the nugget was 1.
+
 In each simulation, a GRTS sample and an IRS sample were selected.
@@
 Then for the IRS sample, the design-based approach using the simple random sample variance (Design IRS) and a model-based approach were applied (Model IRS).
+
+```{r, results = "asis", eval = FALSE}
+sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
+colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
+sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
+print(sim_one_table, include.rownames = FALSE, comment = FALSE)
+```
+
+The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
+FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
+We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
 
@@
 
-# mike chunk
-# Mike: see if this works. If so, delete one of the duplicates
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
-  mutate(filetemp = str_extract(filename, pattern = "sim_\\w+")) %>% 
-  separate(filetemp, into = c("junk", "error_dep", "dist", "n", "location"),
-           sep = "_") %>%
-  select(-junk)
-
-# combo_data <- purrr::map_df(files,
-#                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
-#   separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
-#            sep = "_") %>%
-#   select(-junk, -junk2)
-
-# matt chunk
 combo_data <- purrr::map_df(files,
@@
 - In terms of rmspe, a model-based analysis on a GRTS design yields an rmspe similar to or lower than a design-based analysis on a GRTS design, as long as the response variable is not "too skewed." 
-- If the response variable is very skewed, then neither analysis is appropriate, but, the design-based analysis is quicker. 
+- If the response variable is very skewed, then neither analysis is appropriate, but, the design-based analysis is better. 
 - a spatially balanced GRTS sample outperforms IRS in nearly all dependent error settings, as expected.
@@
 (The realized mean won't be exactly the same but should be close).
-
-__Sample Simulation__
-
-For the following simulation results, we simulated 1040 different gridded populations, each of size 900 (on the unit square) with sample size 150.
-For the design-based approach, population units were selected via GRTS, the Horvitz-Thompson estimator was used,and the local mean variance was used.
-For the model-based approach (FPBK), population units were selected via Independent Random Sampling (IRS) and the appropriate prediction and prediction variance formulas were used.
-
-The response was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1.
-For model-based, we assumed the correct form of the covariance function (exponential), but estimated the spatial parameters with REML.
-
-```{r, results = "asis", eval = FALSE}
-sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
-colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
-sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
-print(sim_one_table, include.rownames = FALSE, comment = FALSE)
-```
-
-## Software
-
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
-We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
 

### 2021-09-03 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
incorporated revisions from Lisa.

@@
 
+Lisa M.: Add paragraph describing contribution of manuscript.
+
 The rest of this paper is organized as follows.
@@
 
-The design-based approach assumes the data are fixed.
+The design-based approach assumes the population is fixed.
 Randomness is incorporated in the selection of population units according to a sampling design.
@@
 The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
-These population parameters are typically assumed to be fixed but unknown.
+These population parameters are traditionally assumed to be fixed but unknown.
 
@@
 We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-But we are typically not interested in the true, unknown mean of the underlying process.
+But we are often not interested in the true, unknown mean of the underlying process.
 
@@
 ```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
+Points circled are those that are sampled.
 In the top row, we have one fixed population, and two random samples.
@@
 
+Lisa M.: Need a more precise definition of "miniature" in this context, and need an example.
+
 The design-based approach can use spatial locations to obtain spatially balanced samples.
@@
 Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties detailed by @stevens2004spatially and @dumelle2021spsurvey.
+We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including Lisa M.: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
 
@@
 A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially provide and @dumelle2021spsurvey provide further details.
+@stevens2004spatially and @dumelle2021spsurvey provide further details.
 
@@
 , \text{z}(s_N) \}$ be a response variable that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
-Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}$ is a $1 \times N$ vector of weights.
+Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
 For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
@@
 
-Typically, however, we only have a sample of the $N$ population units.
+However, we often only have a sample of the $N$ population units.
 Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
@@
 
-We also typically assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
+We also assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
 It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
@@
 \end{align}
-where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$, and $\mathbbm{1}$ is an indicator function.
+where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$; and, $\mathbbm{1}$ is equal to 1 when distance $h_i,j$ is equal to 0, and equal to 0 otherwise.
 However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
 80-93].
+
+Lisa M. : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
 
@@
 However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
+We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
 
@@
 * sample sizes: $n = 10, 50, 150; N = 900$
-* layout: gridded vs random population locations
+* layout: gridded vs random uniform population locations confined to a 1 x 1 unit square
 
@@
 
-In each simulation, a GRTS sample and an IRS sample were selected.
+Lisa M. Notes: adding an intermediate level of spatial dependency?
+Transfer simulation scenarios to a table
+Explain what the effective range is
+Fill in the details of what exactly each approach means (perhaps a table would be a good way to do this)
+Reorder the sims in the first figure by some criterion.
+Think about what would be a "reasonable sample size" instead of 10.
+Define medae
+If the data has a large right-skew, wouldn't one consider a transformation before the analysis?
+We should address this by stating that the BLUP for the log response does not mean that e^logBLUP is the BLUP for the response on the original scale.
+
+In each simulated data set, a GRTS sample and an IRS sample were selected.
 Then for the GRTS sample, the design-based approach using the local neighborhood variance (Design GRTS) and a model-based approach were applied (Model GRTS).
@@
 2.
-We will now focus in a bit more on comparing Design-GRTS to Model-GRTS, the two best approaches for any reasonable sample size.
+We will now focus on comparing Design-GRTS to Model-GRTS, the two best approaches for any reasonable sample size.
 In the independent error settings, the two approaches perform very similarly, so those results are omitted in the following graph.
@@
 Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
-Therefore, we will explore whether or not we can still obtain a relatively precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
+Therefore, we will explore whether or not we can still obtain an adequately precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
 
@@
 Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.03 ng / g.
+Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.2 ng / g.
 
@@
 library(knitr)
-names(res_df) <- c("Approach", "Realized Mean", "Estimate", "SE", "95% LB", "95% UB")
-kable(res_df, digits = 1, caption = "\\label{tab:appliedtab} Table XXX. Application of design-based and model-based approaches to the NLA data set on mercury concentration.")
+res_df_drop <- res_df[ ,-2]
+names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
+kable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
+The true mean concentration is 103.2 103.2 ng / g")
 ```

### 2021-09-03 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
incorporated revisions from Lisa.

--- 2021-08-12 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
@@
 
+Lisa M.: Add paragraph describing contribution of manuscript.
+
 The rest of this paper is organized as follows. In Section
@@
 
-The design-based approach assumes the data are fixed. Randomness is
+The design-based approach assumes the population is fixed. Randomness is
 incorporated in the selection of population units according to a
@@
 population parameters like means and totals. These population parameters
-are typically assumed to be fixed but unknown.
+are traditionally assumed to be fixed but unknown.
 
@@
 population units, we would have an exact prediction for the mean of our
-one realized process, without any uncertainty. But we are typically not
+one realized process, without any uncertainty. But we are often not
 interested in the true, unknown mean of the underlying process.
@@
 \includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/fig1-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
+Points circled are those that are sampled.
 In the top row, we have one fixed population, and two random samples.
@@
 
+Lisa M.: Need a more precise definition of ``miniature'' in this
+context, and need an example.
+
 The design-based approach can use spatial locations to obtain spatially
 balanced samples. First we discuss spatial balance with respect to the
-population (Stevens Jr and Olsen, 2004). A sample is spatially balanced
+population (Stevens and Olsen, 2004). A sample is spatially balanced
 with respect to the population if the sampled population units are a
@@
 Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
-2013; Stevens Jr and Olsen, 2004; Wang et al., 2013). To quantify
-spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
-based on Voroni polygons. The first spatially balanced sampling
-algorithm that saw widespread use was the Generalized Random
-Tessellation Stratified (Stevens Jr and Olsen, 2004). Since GRTS was
-developed, several other spatially balanced sampling algorithms have
-emerged, including the Local Pivotal Method (Grafström et al., 2012;
-Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
-(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
-2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
-Halton Iterative Partitioning (Robertson et al., 2018). We focus on the
-Generalized Random Tessellation Stratified (GRTS) algorithm to select
-spatially balanced sampling because it has several attractive properties
-detailed by Stevens Jr and Olsen (2004) and Dumelle et al. (2021).
+2013; Stevens and Olsen, 2004; Wang et al., 2013). To quantify spatial
+balance, Stevens and Olsen (2004) proposed loss functions based on
+Voroni polygons. The first spatially balanced sampling algorithm that
+saw widespread use was the Generalized Random Tessellation Stratified
+(Stevens and Olsen, 2004). Since GRTS was developed, several other
+spatially balanced sampling algorithms have emerged, including the Local
+Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
+Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
+Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
+(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
+(Robertson et al., 2018). We focus on the Generalized Random
+Tessellation Stratified (GRTS) algorithm to select spatially balanced
+sampling because it has several attractive properties, including Lisa
+M.: List major attractive properties, and detailed by Stevens and Olsen
+(2004) and Dumelle et al. (2021).
 
@@
 to a population unit in two-dimensional space, which results in the
-desired sample. Stevens Jr and Olsen (2004) provide and Dumelle et al.
-(2021) provide further details.
+desired sample. Stevens and Olsen (2004) and Dumelle et al. (2021)
+provide further details.
 
@@
 very difficult to calculate. Second, they ignore the spatial locations
-of the population units. To address these drawbacks, Stevens Jr and
-Olsen (2003) proposed a local neighborhood variance estimator. The local
+of the population units. To address these drawbacks, Stevens and Olsen
+(2003) proposed a local neighborhood variance estimator. The local
 neighborhood variance estimator does not rely on \(\pi_{ij}\), and it
 incorporates spatial locations by assigning higher weights to nearby
-observations. Stevens Jr and Olsen (2003) show this variance estimators
+observations. Stevens and Olsen (2003) show this variance estimators
 tends to reduce the estimated standard error of \(\hat{\tau}\), yielding
@@
 predict some linear function of the response variable,
-\(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where \(\mathbf{b}\)
-is a \(1 \times N\) vector of weights. For example, if we want to
-predict the population total across all population units, then we would
-use a vector of 1's for the weights.
+\(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
+\(\mathbf{b}^\prime\) is a \(1 \times N\) vector of weights. For
+example, if we want to predict the population total across all
+population units, then we would use a vector of 1's for the weights.
 
-Typically, however, we only have a sample of the \(N\) population units.
+However, we often only have a sample of the \(N\) population units.
 Denoting quantities that are part of the sampled population units with a
@@
 
-We also typically assume that there is spatial correlation in
-\(\bm{\delta}\), which can be modeled using a covariance function. It is
-common to assume the covariance function is second-order stationary and
-isotropic (Cressie, 1993), and that the spatial covariance decreases as
-the separation between population units increases. Many spatial
-covariance functions exist, but the primary function we use throughout
-the simulations and applications in this manuscript is the exponential
+We also assume that there is spatial correlation in \(\bm{\delta}\),
+which can be modeled using a covariance function. It is common to assume
+the covariance function is second-order stationary and isotropic
+(Cressie, 1993), and that the spatial covariance decreases as the
+separation between population units increases. Many spatial covariance
+functions exist, but the primary function we use throughout the
+simulations and applications in this manuscript is the exponential
 covariance function: the \(i,j^{th}\) entry for
@@
 parameters of the partial sill \(\theta_1\), the range \(\theta_2\), and
-the nugget \(\theta_3\), and \(\mathbbm{1}\) is an indicator function.
-However, any spatial covariance function could be used in the place of
-the exponential, including functions that allow for non-stationarity or
+the nugget \(\theta_3\); and, \(\mathbbm{1}\) is equal to 1 when
+distance \(h_i,j\) is equal to 0, and equal to 0 otherwise. However, any
+spatial covariance function could be used in the place of the
+exponential, including functions that allow for non-stationarity or
 anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
+
+Lisa M. : Include formulas. Perhaps, but, these are very heavy in
+notation and matrix algebra. We might consider, however, adding the
+formulas to an Appendix.
 
@@
 or total from spatially correlated responses in a finite population
-setting.
+setting. We choose to use FPBK because it is faster than a Bayesian
+approach and random forest and because Ver Hoef and Temesgen (2013)
+showed that the method outperforms k-nearest-neighbors in many
+scenarios.
 
@@
 \item
-  layout: gridded vs random population locations
+  layout: gridded vs random uniform population locations confined to a 1
+  x 1 unit square
 \end{itemize}
@@
 
-In each simulation, a GRTS sample and an IRS sample were selected. Then
-for the GRTS sample, the design-based approach using the local
-neighborhood variance (Design GRTS) and a model-based approach were
-applied (Model GRTS). Then for the IRS sample, the design-based approach
-using the simple random sample variance (Design IRS) and a model-based
-approach were applied (Model IRS).
+There were 2000 trials for each simulation. The original response
+(before exponentiating if applicable) for the dependent error cases was
+normally distributed with an exponential covariance function with
+partial sill of 0.9, effective range of \(\sqrt{2}\), and a nugget of
+0.1. For the independent error cases, the partial sill was 0 and the
+nugget was 1.
+
+Lisa M. Notes: adding an intermediate level of spatial dependency?
+Transfer simulation scenarios to a table Explain what the effective
+range is Fill in the details of what exactly each approach means
+(perhaps a table would be a good way to do this) Reorder the sims in the
+first figure by some criterion. Think about what would be a ``reasonable
+sample size'' instead of 10. Define medae If the data has a large
+right-skew, wouldn't one consider a transformation before the analysis?
+We should address this by stating that the BLUP for the log response
+does not mean that e\^{}logBLUP is the BLUP for the response on the
+original scale.
+
+In each simulated data set, a GRTS sample and an IRS sample were
+selected. Then for the GRTS sample, the design-based approach using the
+local neighborhood variance (Design GRTS) and a model-based approach
+were applied (Model GRTS). Then for the IRS sample, the design-based
+approach using the simple random sample variance (Design IRS) and a
+model-based approach were applied (Model IRS).
+
+The GRTS algorithm and the local neighborhood variance estimator are
+available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
+et al., 2021). FPBK can be readily performed in \texttt{R} with the
+\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
+for both the simulation analysis and the application, estimating
+parameters with Restricted Maximum Likelihood (REML).
 
@@
 \item
-  We will now focus in a bit more on comparing Design-GRTS to
-  Model-GRTS, the two best approaches for any reasonable sample size. In
-  the independent error settings, the two approaches perform very
-  similarly, so those results are omitted in the following graph. In the
-  dependent error settings, using rmspe as the performance criterion,
-  Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two
-  approaches perform very similarly in 3 settings, and Design-GRTS
-  outperforms Model-GRTS in 3 settings.
+  We will now focus on comparing Design-GRTS to Model-GRTS, the two best
+  approaches for any reasonable sample size. In the independent error
+  settings, the two approaches perform very similarly, so those results
+  are omitted in the following graph. In the dependent error settings,
+  using rmspe as the performance criterion, Model-GRTS outperforms
+  Design-GRTS in 12 of the 18 settings, the two approaches perform very
+  similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3
+  settings.
 \end{enumerate}
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-2.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
 
@@
 
-\textbf{Sample Simulation}
-
-For the following simulation results, we simulated 1040 different
-gridded populations, each of size 900 (on the unit square) with sample
-size 150. For the design-based approach, population units were selected
-via GRTS, the Horvitz-Thompson estimator was used,and the local mean
-variance was used. For the model-based approach (FPBK), population units
-were selected via Independent Random Sampling (IRS) and the appropriate
-prediction and prediction variance formulas were used.
-
-The response was normally distributed with an exponential covariance
-function with partial sill of 0.9, effective range of \(\sqrt{2}\), and
-a nugget of 0.1. For model-based, we assumed the correct form of the
-covariance function (exponential), but estimated the spatial parameters
-with REML.
-
-\hypertarget{software}{%
-\subsection{Software}\label{software}}
-
-The GRTS algorithm and the local neighborhood variance estimator are
-available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
-et al., 2021). FPBK can be readily performed in \texttt{R} with the
-\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
-for both the simulation analysis and the application, estimating
-parameters with Restricted Maximum Likelihood (REML).
-
 \hypertarget{application}{%
@@
 always be enough time or money to do so. Therefore, we will explore
-whether or not we can still obtain a relatively precise estimate for the
-realized mean mercury concentration if we only take a sample of 100 of
-the 986 lakes.
+whether or not we can still obtain an adequately precise estimate for
+the realized mean mercury concentration if we only take a sample of 100
+of the 986 lakes.
 
@@
 population, we know that the realized mean mercury concentration is
-103.03 ng / g.
+103.2 ng / g.
 
-\begin{longtable}[]{@{}lrrrrr@{}}
-\caption{\label{tab:appliedtab} Table XXX. Application of design-based
-and model-based approaches to the NLA data set on mercury
-concentration.}\tabularnewline
+\begin{longtable}[]{@{}lrrrr@{}}
+\caption{\label{tab:appliedtab} Application of design-based and
+model-based approaches to the NLA data set on mercury concentration. The
+true mean concentration is 103.2 103.2 ng / g}\tabularnewline
 \toprule
-Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Estimate & SE & 95\% LB & 95\% UB \\
 \midrule
@@
 \toprule
-Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Estimate & SE & 95\% LB & 95\% UB \\
 \midrule
 \endhead
-Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\
-Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\
-Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\
-Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\
+Design IRS & 112.7 & 8.8 & 95.4 & 129.9 \\
+Model IRS & 110.5 & 7.9 & 95.0 & 125.9 \\
+Design GRTS & 101.8 & 6.1 & 89.8 & 113.7 \\
+Model GRTS & 102.3 & 5.9 & 90.8 & 113.9 \\
 \bottomrule
@@
 \leavevmode\hypertarget{ref-stevens2003variance}{}%
-Stevens Jr, D.L., Olsen, A.R., 2003. Variance estimation for spatially
+Stevens, D.L., Olsen, A.R., 2003. Variance estimation for spatially
 balanced samples of environmental resources. Environmetrics 14,
@@
 \leavevmode\hypertarget{ref-stevens2004spatially}{}%
-Stevens Jr, D.L., Olsen, A.R., 2004. Spatially balanced sampling of
-natural resources. Journal of the american Statistical association 99,
-262--278.
+Stevens, D.L., Olsen, A.R., 2004. Spatially balanced sampling of natural
+resources. Journal of the american Statistical association 99, 262--278.
 

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added some of Tony's comments to the manuscript.

@@
 
+__Tony O.__: Should this paragraph address that spatial information can be incorporated in the design stage or in the analysis stage (or both).
+In general, it's not clear whether we are referring to site selection process or the estimation process
+
 Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as variables measured at specific geographic locations.
+We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process.
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
@@
 
-Lisa M.: Add paragraph describing contribution of manuscript.
+__Lisa M.__: Add paragraph describing contribution of manuscript.
 
@@
 A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit.
-Some examples of commonly used sampling designs include independent random sampling (IRS), stratified random sampling, and cluster sampling.
+Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
+sampling, which we refer to as Independent Random Sampling (IRS) survey designs.
 The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
@@
 
+__Tony O.__: Before this section is it useful to have a section that lays out the general site selection and general analysis options.
+Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based.
+Then general analysis options as design-based no spatial, design-based spatial, model-based.
+This four by three table would show that model-based analyses are possible for all selection options.
+Design-based options with no spatial info possible for IRS-based and GRTS-based.
+Design-based options with spatial info possible for GRTS-based.
+
 ## Spatially Balanced Design and Analysis
 
-Lisa M.: Need a more precise definition of "miniature" in this context, and need an example.
+__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
 
@@
 Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including Lisa M.: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
+We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including __Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
 
@@
 
-Lisa M. : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
+__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
 
@@
 
-Lisa M. Notes: adding an intermediate level of spatial dependency?
+__Lisa M.__ Notes: adding an intermediate level of spatial dependency?
 Transfer simulation scenarios to a table
@@
                     full.names = TRUE)
+
+## read_csv("inst/output/sim_deperr_normal_n10_gridloc.csv")
+## read_csv("inst/output/sim_deperr_normal_n50_gridloc.csv")
 
@@
 ```{r}
+## View(combo_data %>% filter(approach == "Design IRS"))
 ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added some of Tony's comments to the manuscript.

@@
 
+\textbf{Tony O.}: Should this paragraph address that spatial information
+can be incorporated in the design stage or in the analysis stage (or
+both). In general, it's not clear whether we are referring to site
+selection process or the estimation process
+
 Though the design-based and model-based approaches apply to statistical
 inference in a broad sense, we focus on comparing these approaches for
-spatial data. We define spatial data as variables measured at specific
-geographic locations. De Gruijter and Ter Braak (1990) give an early
+spatial data. We define spatial data as data that incorporates the
+specific locations of the population units into either the design or
+estimation process. De Gruijter and Ter Braak (1990) give an early
 comparison of design-based and model-based approaches for spatial data,
@@
 
-Lisa M.: Add paragraph describing contribution of manuscript.
+\textbf{Lisa M.}: Add paragraph describing contribution of manuscript.
 
@@
 inclusion in the sample (inclusion probability) to each population unit.
-Some examples of commonly used sampling designs include independent
-random sampling (IRS), stratified random sampling, and cluster sampling.
-The goal is to use the sampling design and the sampled data to estimate
-population parameters like means and totals. These population parameters
-are traditionally assumed to be fixed but unknown.
+Some examples of commonly used sampling designs include simple random
+sampling, stratified random sample, and cluster sampling, which we refer
+to as Independent Random Sampling (IRS) survey designs. The goal is to
+use the sampling design and the sampled data to estimate population
+parameters like means and totals. These population parameters are
+traditionally assumed to be fixed but unknown.
 
@@
 
+\textbf{Tony O.}: Before this section is it useful to have a section
+that lays out the general site selection and general analysis options.
+Thinking about site selection as design-based IRS, design-based GRTS,
+Arbitrary set of sites, selection for model-based. Then general analysis
+options as design-based no spatial, design-based spatial, model-based.
+This four by three table would show that model-based analyses are
+possible for all selection options. Design-based options with no spatial
+info possible for IRS-based and GRTS-based. Design-based options with
+spatial info possible for GRTS-based.
+
 \hypertarget{spatially-balanced-design-and-analysis}{%
@@
 
-Lisa M.: Need a more precise definition of ``miniature'' in this
-context, and need an example.
+\textbf{Lisa M.}: Need a more precise definition of ``miniature'' in
+this context, and need an example.
 
@@
 Tessellation Stratified (GRTS) algorithm to select spatially balanced
-sampling because it has several attractive properties, including Lisa
-M.: List major attractive properties, and detailed by Stevens and Olsen
-(2004) and Dumelle et al. (2021).
+sampling because it has several attractive properties, including
+\textbf{Lisa M.}: List major attractive properties, and detailed by
+Stevens and Olsen (2004) and Dumelle et al. (2021).
 
@@
 
-Lisa M. : Include formulas. Perhaps, but, these are very heavy in
-notation and matrix algebra. We might consider, however, adding the
+\textbf{Lisa M.} : Include formulas. Perhaps, but, these are very heavy
+in notation and matrix algebra. We might consider, however, adding the
 formulas to an Appendix.
@@
 
-Lisa M. Notes: adding an intermediate level of spatial dependency?
-Transfer simulation scenarios to a table Explain what the effective
-range is Fill in the details of what exactly each approach means
-(perhaps a table would be a good way to do this) Reorder the sims in the
-first figure by some criterion. Think about what would be a ``reasonable
-sample size'' instead of 10. Define medae If the data has a large
-right-skew, wouldn't one consider a transformation before the analysis?
-We should address this by stating that the BLUP for the log response
-does not mean that e\^{}logBLUP is the BLUP for the response on the
-original scale.
+\textbf{Lisa M.} Notes: adding an intermediate level of spatial
+dependency? Transfer simulation scenarios to a table Explain what the
+effective range is Fill in the details of what exactly each approach
+means (perhaps a table would be a good way to do this) Reorder the sims
+in the first figure by some criterion. Think about what would be a
+``reasonable sample size'' instead of 10. Define medae If the data has a
+large right-skew, wouldn't one consider a transformation before the
+analysis? We should address this by stating that the BLUP for the log
+response does not mean that e\^{}logBLUP is the BLUP for the response on
+the original scale.
 

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
Added Jay's comments to the manuscript.

@@
 
+__Jay VH__: I think it is important to stress that the limiting distribution is over all possible randomizations, constrained by whatever design is used.
+
+__Jay VH__: quantity is vague.
+We should stick with variables, or realized variables (we might also call these values, but we should define and establish a consistent terminology early on.) __Matt H__: I think, though this comment is for this paragraph, we should establish the terminology earlier.
+
 The model-based approach assumes the data are a random realization of a data-generating process.
@@
 
+__Jay VH__: What about the design for model-based inference?
+Strictly speaking, it is fixed – there is no probabilistic use of a randomized design.
+However, we are going to have to deal with Diggle et al.
+(2010).
+
 ## Spatially Balanced Design and Analysis
@@
 __Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
+
+__Jay VH__: Saying "the distribution of the sampled population units mirrors the density of..." is confusing to me.
+Are these formal statistical definitions of distribution (cumulative distribution function) and density (probability density function)?
+Wouldn’t IRS sample be a miniature, as it should, on average, mirror a population?
 
@@
 .
-, \text{z}(s_N) \}$ be a response variable that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
+, \text{z}(s_N) \}$ be a response vector at locations $s_1$, $s_2$, .
+.
+.
+, $s_N$ that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
 Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
@@
 
+__Jay VH__: Use relative efficiency for figure, using design IRS as the base method.
+
+__Jay VH__: There is not a complete crossing of all simulation factors. We need to add text description about why some independent error settings were excluded.
+
 ```{r}
@@
 Design GRTS (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
+
+__Jay VH__: Need to figure out the weird results for n = 50.
+
+__Jay VH__: For the design estimator, verify that the simulation MSPE is equal to the estimated variance for the design estimator, on average, to make sure things are correct for this estimator.
+
+__Jay VH__: Not a ton of difference in grid vs. random so we can consider removing that factor to simplify the presentation, or, put it in an appendix.
 

### 2021-09-07 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
Added Jay's comments to the manuscript.

@@
 
+\textbf{Jay VH}: I think it is important to stress that the limiting
+distribution is over all possible randomizations, constrained by
+whatever design is used.
+
+\textbf{Jay VH}: quantity is vague. We should stick with variables, or
+realized variables (we might also call these values, but we should
+define and establish a consistent terminology early on.) \textbf{Matt
+H}: I think, though this comment is for this paragraph, we should
+establish the terminology earlier.
+
 The model-based approach assumes the data are a random realization of a
@@
 
+\textbf{Jay VH}: What about the design for model-based inference?
+Strictly speaking, it is fixed -- there is no probabilistic use of a
+randomized design. However, we are going to have to deal with Diggle et
+al.~(2010).
+
 \hypertarget{spatially-balanced-design-and-analysis}{%
@@
 this context, and need an example.
+
+\textbf{Jay VH}: Saying ``the distribution of the sampled population
+units mirrors the density of\ldots{}'' is confusing to me. Are these
+formal statistical definitions of distribution (cumulative distribution
+function) and density (probability density function)? Wouldn't IRS
+sample be a miniature, as it should, on average, mirror a population?
 
@@
 \(\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}\)
-be a response variable that can be measured at the \(N\) population
-units and is represented as an \(N \times 1\) vector. Suppose we want to
-predict some linear function of the response variable,
+be a response vector at locations \(s_1\), \(s_2\), . . . , \(s_N\) that
+can be measured at the \(N\) population units and is represented as an
+\(N \times 1\) vector. Suppose we want to predict some linear function
+of the response variable,
 \(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
@@
 
+\textbf{Jay VH}: Use relative efficiency for figure, using design IRS as
+the base method.
+
+\textbf{Jay VH}: There is not a complete crossing of all simulation
+factors. We need to add text description about why some independent
+error settings were excluded.
+
 \includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
@@
 undercover, a result Tony was familiar with.
+
+\textbf{Jay VH}: Need to figure out the weird results for n = 50.
+
+\textbf{Jay VH}: For the design estimator, verify that the simulation
+MSPE is equal to the estimated variance for the design estimator, on
+average, to make sure things are correct for this estimator.
+
+\textbf{Jay VH}: Not a ton of difference in grid vs.~random so we can
+consider removing that factor to simplify the presentation, or, put it
+in an appendix.
 

### 2021-09-14 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
Updated simulations to include 48 settings and be run in a single .r file.

@@
 * Environmental and Ecological Statistics
+
+<!-- Notes from today -->
+<!-- •	Finite populations require a smaller scope but we need to discuss the impact of infinite populations as well – put finite vs infinite population context in the introduction / similarities and differences -->
+<!-- •	What should a lower sample size be? 50 seems reasonable 30 may be too small – maybe 50, 100, 200 -->
+<!-- •	IRS results seem off for dependent error sample size 50 -->
+<!-- •	Take home points – don’t have to do GRTS design with GRTS sample -->
+<!-- •	Take home points – GRTS is better for both model and design, while spatial approach for design is close to spatial model -->
+<!-- •	Add medium dependence -->
+<!-- •	Remind readers what choices of the four approaches in the application mean – what do they mean to practitioners -->
 

### 2021-09-14 inst/scripts/all_sims.R

@@ -1,0 +1,99 @@
+library(tidyverse)
+library(parallel)
+
+# write results?
+write_out <- TRUE
+
+## set up grid for simulation settings
+N <- 30 ^ 2
+n <- c(10, 50, 100, 200)
+total_var <- 2
+psill_ratio <- c(0, 0.5, 0.9)
+erange <- sqrt(4)
+gridded <- c(TRUE, FALSE)
+resptype <- c("normal", "lognormal")
+
+parm_df <- expand_grid(N, n, total_var, psill_ratio,
+                       erange, gridded, resptype) %>%
+  mutate(nugget_ratio = 1 - psill_ratio,
+         psill = total_var * psill_ratio,
+         nugget = total_var * nugget_ratio)
+
+# source functions for now
+source("R/sim_pop.R")
+source("R/sim_trial.R")
+
+## loop through each row of parm_df
+for (i in 1:nrow(parm_df)) {
+  n_trials <- 5
+  seed <- runif(n_trials, 1, 1e7)
+  n_cluster <- detectCores() # find cores (48 on mine)
+  cluster <- makeCluster(n_cluster) # make cluster
+  clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
+  clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
+  clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
+  clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
+
+  parm_df_sim <- parm_df %>% slice(i) ## grab row for ith parameter combo
+  sim_output <- parLapply(
+    cluster, # the cluster
+    seed, # some seeds
+    safely(sim_trial), # see if any errors
+    ## pull each parameter from the single row data frame
+    N = parm_df_sim %>% pull(N), # pop size
+    n = parm_df_sim %>% pull(n), # sample size
+    psill = parm_df_sim %>% pull(psill), # partial sill
+    nugget = parm_df_sim %>% pull(nugget), # nugget
+    erange = parm_df_sim %>% pull(erange), # effective range
+    gridded = parm_df_sim %>% pull(gridded), # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
+    resptype = parm_df_sim %>% pull(resptype) # normal response
+  )
+
+  stopCluster(cluster) # stop cluster
+
+  # check errors
+  map(sim_output, "error") %>%
+    bind_rows() %>%
[…]

### 2021-09-14 inst/scripts/sim_deperr_lognormal_n10_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-09-14 inst/scripts/simulation_exploration.R

@@ -1,16 +1,17 @@
 library(tidyverse)
 library(here)
-read_csv(here("inst", "output", "sim_deperr_lognormal_n50_gridloc.csv"))
 
-files <- list.files(path = here("inst", "output"), pattern = "*.csv",
-                    full.names = TRUE)
+files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
 
 combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
-  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
-           sep = "_") %>%
-  select(-junk)
-combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))
+                            ~read_csv(.x) %>% mutate(filename = .x))
+
+# %>%
+#   separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
+#            sep = "_") %>%
+#   select(-junk)
+
+combo_data <- combo_data %>% mutate(sim = interaction(n, psill, resptype, gridded))
 
 ##
 ggplot(data = combo_data, aes(x = sim, y = rmspe, colour = approach)) +
@@ -27,17 +28,8 @@ summary(lm(rmspe ~ approach, data = combo_data_grts))
 ## evidence that normal distribution gives better estimators (makes sense)
 ## evidence that n of 50 gives worse estimators (makes sense)
 
-summary(lm(rmspe ~ approach + error_dep + dist + n + location,
+summary(lm(rmspe ~ approach + resptype + psill + n + gridded,
            data = combo_data_grts))
-
-summary(lm(rmspe ~ approach + error_dep*approach + dist* approach + n * approach + location * approach, data = combo_data_grts))
-
-
-## checking other things
-summary(lm(med_ci_len ~ approach, data = combo_data_grts))
-summary(lm(med_ci_len ~ approach + error_dep + dist + n + location,
-           data = combo_data_grts))
-
 
 
 ## visually explore the two approaches
@@ -46,7 +38,7 @@ ggplot(data = combo_data_grts, aes(x = sim, y = rmspe, colour = approach)) +
   geom_point() +
   coord_flip()
 
[…]

### 2021-09-16 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated simulations and a start to an updated write-up.

@@
 
+MAJOR POINTS for the following Figure, which has relative efficiency (rmspe / Design IRS rmspe): We see that
+
+- When there is no spatial correlation (top row), we aren't losing that much by using a spatial method over Design IRS, even if assumptions are violated.
+- When there is a lot of spatial correlation Model GRTS tends to perform best, but difference in relative efficiency between Model GRTS and Design GRTS is not very big.
+In many settings Design GRTS outperforms Model IRS by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is much more important than the analysis decision (whether to analyze using model assumptions or not).
+- If there is a large amount of spatial correlation, we should __not__ use IRS.
+Even though it's assumptions are satisfied, the resulting estimator is much worse than an estimator using a spatially balanced sample.
+- If we are comparing design-based and model-based methods, we should not use a poor design to compare with (give examples?).
+
 ```{r, results = "hide", message = FALSE, warning = FALSE}
 library(tidyverse)
-files <- list.files(path = here("inst", "output"), pattern = "*.csv",
-                    full.names = TRUE)
+library(here)
 
-## read_csv("inst/output/sim_deperr_normal_n10_gridloc.csv")
-## read_csv("inst/output/sim_deperr_normal_n50_gridloc.csv")
+files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
 
 combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
-  mutate(filetemp = str_extract(filename, pattern = "sim_\\w+")) %>% 
-  separate(filetemp, into = c("junk", "error_dep", "dist", "n", "location"),
-           sep = "_") %>%
-  select(-junk)
+                            ~read_csv(.x) %>% mutate(filename = .x))
 
-combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))
+combo_data <- combo_data %>%
+  dplyr::filter(n != 10 & gridded == FALSE) %>%
+  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
+  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
+                                            "Design GRTS", "Model GRTS"))) %>%
+  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal")))
+combo_data <- combo_data %>% group_by(sim) %>% mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
+  fill(designirsrmspe, .direction = "downup") %>%
+  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
+  ungroup()
 
-combo_data_grts <- combo_data %>% dplyr::filter(approach %in% c("Design GRTS", "Model GRTS"))
+ggplot(data = combo_data, aes(x = n, y = rel_efficiency, colour = approach)) +
+  facet_grid(psill ~ resptype) +
+  geom_jitter(width = 7) +
+  scale_colour_manual(values = c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4"))
+```
 
-dep_only <- combo_data_grts %>% dplyr::filter(error_dep == "deperr")
-dep_long <- dep_only %>%
-  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
-  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
+MAJOR POINTS for the following Figure:
 
-dep_long_abs <- dep_only %>%
-  pivot_wider(names_from = approach, values_from = medae, sim) %>%
-  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
+- when we drop the IRS samples to compare the GRTS samples more closely (looking at the relative efficiency of model rmspe / design rmspe), we see that the model-based approach is usually better than the design-based approach (but, keep in mind that the improvement was small compared to the improvement of both methods over Design IRS).
+- when the model used is the same model that generated the data, the Model-based approach far outperforms the Design-based approach, especially when there is a lot of spatial correlation (bottom-left facet).
+The methods perform similarly when there is no spatial correlation.
+- even when the model that generates the data is different than the model used to fit the data (lognormal), the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
+
+```{r}
+combo_data_grts <- combo_data %>%
+  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
+  mutate(designgrts = if_else(approach == "Design GRTS", 
+                              true = rmspe,
+                              false = NA_real_)) %>%
+  group_by(sim) %>%
+  fill(designgrts, .direction = "updown") %>%
+  ungroup() %>%
+  mutate(releffgrts = rmspe / designgrts)
+combo_data_grts_model <- combo_data_grts %>%
+  filter(approach == "Model GRTS")
+
+ggplot(data = combo_data_grts_model, aes(x = n, y = rel_efficiency)) +
+  facet_grid(psill ~ resptype) +
+  geom_point()
+```
+
+MAJOR POINTS for the following Figure:
+
+- coverage for the model-based estimator is slightly higher than coverage for the design-based estimator in the normal settings.
+Coverages are about equal in the lognormal settings with a slight edge to model-based (this is the point that will be tougher to explain: I would have expected the design-based estimator to have better coverage in the lognormal settings because it has fewer assumptions).
+- coverage is at or near the nominal 95% in all of the normal settings, where assumptions for the model approach and the design approach are satisfied.
+- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean.
+The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
+- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval.
+Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.
+
+```{r}
+ggplot(data = combo_data_grts, aes(x = n, y = coverage, colour = approach)) +
+  geom_hline(yintercept = 0.95) +
+  geom_jitter(width = 7) + 
+  facet_grid(psill ~ resptype) +
+  scale_colour_manual(values = c("mediumpurple1", "mediumpurple4"))
 ```
@@
 
-__Jay VH__: Use relative efficiency for figure, using design IRS as the base method.
-
-__Jay VH__: There is not a complete crossing of all simulation factors. We need to add text description about why some independent error settings were excluded.
-
-```{r}
-ggplot(data = combo_data, aes(x = sim, y = rmspe, colour = approach)) +
-  geom_point() +
-  coord_flip()
-```
-
 2.
@@
 In the dependent error settings, using rmspe as the performance criterion, Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two approaches perform very similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3 settings.
-
-```{r}
-ggplot(data = dep_long_abs, aes(x = fct_reorder(sim, perc), y = perc)) +
-  geom_point() +
-  coord_flip() +
-  labs(y = "Percent Change in medae (Design - Model) / Model",
-       caption = "Percent > 0 means Model has better rmspe")
-
-ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
-  geom_point() +
-  coord_flip() +
-  labs(y = "Percent Change in rmspe (Design - Model) / Model",
-       caption = "Percent > 0 means Model has better rmspe")
-```
 
@@
 
-```{r}
-design_better <- dep_long %>% dplyr::filter(perc < -5)
-design_better_full <- inner_join(combo_data_grts, design_better, by = c("sim"))
-ggplot(data = design_better_full, aes(x = sim, y = coverage, colour = approach)) +
-  geom_point() +
-  coord_flip() +
-  geom_hline(aes(yintercept = 0.95), alpha = 0.5)
-```
-
 4. Coverage
@@
 Design GRTS (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
-
-__Jay VH__: Need to figure out the weird results for n = 50.
-
-__Jay VH__: For the design estimator, verify that the simulation MSPE is equal to the estimated variance for the design estimator, on average, to make sure things are correct for this estimator.
-
-__Jay VH__: Not a ton of difference in grid vs. random so we can consider removing that factor to simplify the presentation, or, put it in an appendix.
 

### 2021-09-16 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated simulations and a start to an updated write-up.

@@
 
+MAJOR POINTS for the following Figure, which has relative efficiency
+(rmspe / Design IRS rmspe): We see that
+
+\begin{itemize}
+\tightlist
+\item
+  When there is no spatial correlation (top row), we aren't losing that
+  much by using a spatial method over Design IRS, even if assumptions
+  are violated.
+\item
+  When there is a lot of spatial correlation Model GRTS tends to perform
+  best, but difference in relative efficiency between Model GRTS and
+  Design GRTS is not very big. In many settings Design GRTS outperforms
+  Model IRS by a large margin, suggesting that the design decision
+  (whether to use IRS or GRTS) is much more important than the analysis
+  decision (whether to analyze using model assumptions or not).
+\item
+  If there is a large amount of spatial correlation, we should
+  \textbf{not} use IRS. Even though it's assumptions are satisfied, the
+  resulting estimator is much worse than an estimator using a spatially
+  balanced sample.
+\item
+  If we are comparing design-based and model-based methods, we should
+  not use a poor design to compare with (give examples?).
+\end{itemize}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+MAJOR POINTS for the following Figure:
+
+\begin{itemize}
+\tightlist
+\item
+  when we drop the IRS samples to compare the GRTS samples more closely
+  (looking at the relative efficiency of model rmspe / design rmspe), we
+  see that the model-based approach is usually better than the
+  design-based approach (but, keep in mind that the improvement was
+  small compared to the improvement of both methods over Design IRS).
+\item
+  when the model used is the same model that generated the data, the
+  Model-based approach far outperforms the Design-based approach,
+  especially when there is a lot of spatial correlation (bottom-left
+  facet). The methods perform similarly when there is no spatial
+  correlation.
+\item
+  even when the model that generates the data is different than the
+  model used to fit the data (lognormal), the model-based approach still
+  outperforms the design-based approach when there is a high amount of
+  spatial correlation.
+\end{itemize}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+MAJOR POINTS for the following Figure:
+
+\begin{itemize}
+\tightlist
+\item
+  coverage for the model-based estimator is slightly higher than
+  coverage for the design-based estimator in the normal settings.
+  Coverages are about equal in the lognormal settings with a slight edge
+  to model-based (this is the point that will be tougher to explain: I
+  would have expected the design-based estimator to have better coverage
+  in the lognormal settings because it has fewer assumptions).
+\item
+  coverage is at or near the nominal 95\% in all of the normal settings,
+  where assumptions for the model approach and the design approach are
+  satisfied.
+\item
+  for the model-based approach, the more skewed the population is, the
+  higher the sample size needed to satisfy CLT for predicting a mean.
+  The derivation of the BLUP is entirely moment-based (no distribution
+  assumed) but we still need to assume a distribution to estimate
+  spatial parameters and to generate bounds of a prediction interval.
+\item
+  many confidence intervals generated for design-based approaches also
+  rely on the CLT and the normal distribution to generate the interval.
+  Again, for highly skewed data with a small sample size, this
+  assumption is violated even though all of the assumptions for
+  generating the estimator are valid.
+\end{itemize}
+
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+
 Major Points from August 3 Simulations:
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
   much which approach is used, which makes sense.
-\end{enumerate}
-
-\textbf{Jay VH}: Use relative efficiency for figure, using design IRS as
-the base method.
-
-\textbf{Jay VH}: There is not a complete crossing of all simulation
-factors. We need to add text description about why some independent
-error settings were excluded.
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
 \item
@@
   settings.
-\end{enumerate}
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
 \item
@@
   approach, though it is ``safer'' to use Design-GRTS.
-\end{enumerate}
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{3}
-\tightlist
 \item
@@
 
-\textbf{Jay VH}: Need to figure out the weird results for n = 50.
-
-\textbf{Jay VH}: For the design estimator, verify that the simulation
-MSPE is equal to the estimated variance for the design estimator, on
-average, to make sure things are correct for this estimator.
-
-\textbf{Jay VH}: Not a ton of difference in grid vs.~random so we can
-consider removing that factor to simplify the presentation, or, put it
-in an appendix.
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 

### 2021-09-16 inst/scripts/all_sims.R

@@ -26,7 +26,7 @@ source("R/sim_trial.R")
 ## loop through each row of parm_df
 for (i in 1:nrow(parm_df)) {
   n_trials <- 5
-  seed <- runif(n_trials, 1, 1e7)
+  seed <- sample.int(1e7, size = n_trials)
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
   clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
@@ -52,9 +52,9 @@ for (i in 1:nrow(parm_df)) {
   stopCluster(cluster) # stop cluster
 
   # check errors
-  map(sim_output, "error") %>%
-    bind_rows() %>%
-    nrow()
+  # map(sim_output, "error") %>%
+  #   bind_rows() %>%
+  #   nrow()
 
   # do summaries
   safe_output <- map(sim_output, "result") %>%

### 2021-09-16 inst/scripts/simulation_exploration.R

@@ -6,11 +6,6 @@ files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv",
 combo_data <- purrr::map_df(files,
                             ~read_csv(.x) %>% mutate(filename = .x))
 
-# %>%
-#   separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
-#            sep = "_") %>%
-#   select(-junk)
-
 combo_data <- combo_data %>% mutate(sim = interaction(n, psill, resptype, gridded))
 
 ##
@@ -86,4 +81,4 @@ ggplot(data = combo_data_grts, aes(x = sim, y = coverage, colour = approach)) +
 #   geom_histogram(bins = 15) +
 #   facet_wrap(~ location)
 #
-#
+

### 2021-09-17 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added simulation parameter tables and graph of lognormal distribution.

@@
 
-There were several variables to alter in the simulations, and the names of the scenarios in future plots mirror these variables
+We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.
+
+```{r, echo = FALSE}
+tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
+       nrow = 2) %>%
+  as.table()
+row.names(tab) <- c("Design", "Model")
+library(knitr)
+kable(tab, col.names = c("IRS", "GRTS"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
+The columns give the two types of sampling designs while the rows give the two types of analyses.")
+## I really want "Analysis" to appear as a row header and "Sampling Design" 
+## to appear as a column header, but this isn't easy to figure out.
+## I couldn't figure this out with xtable either so it might need to be
+## made "by hand."
+```
+
+We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
+All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded).
+For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
+A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
+Therefore, when the model-based methods were used, the correlation is mis-specified.
+We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness (Figure \ref{fig:figlog}).
+
+```{r}
+n <- c("50", "100", "200")
+site_locations <- c("Random", "Gridded", "")
+psill_ratio <- c("0", "0.5", "0.9")
+resp_type <- c("Normal", "Lognormal", "")
+parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
+row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
+                         "Partial Sill Variance Ratio", "Response Type")
+kable(parm_tab, col.names = c("", "", ""),
+      caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+```
+
+```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
+The distribution exhibits moderate to strong right-skewness."}
+mu <- 0; var <- 2
+xvar <- seq(0.001, 10, by = 0.01)
+yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
+df <- tibble(xvar = xvar, yvar = yvar)
+ggplot(df, aes(x = xvar, y = yvar)) +
+  geom_line()
+```
 
@@
     * lognormal: log scale mean 0, log scale variance 2 (total variance 47)
-    * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926)
+    <!-- * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926) -->
 * sample sizes: $n = 10, 50, 150; N = 900$

### 2021-09-17 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added simulation parameter tables and graph of lognormal distribution.

@@
 
-There were several variables to alter in the simulations, and the names
-of the scenarios in future plots mirror these variables
+We used a numerical simulation study to investigate performance of four
+design-analysis combinations, summarized in Table
+\ref{tab:designanalysis}.
+
+\begin{longtable}[]{@{}lll@{}}
+\caption{\label{tab:designanalysis} Types of Sampling Design and
+Analysis combinations considered in the simulation study. The columns
+give the two types of sampling designs while the rows give the two types
+of analyses.}\tabularnewline
+\toprule
+& IRS & GRTS \\
+\midrule
+\endfirsthead
+\toprule
+& IRS & GRTS \\
+\midrule
+\endhead
+Design & IRS-Design & GRTS-Design \\
+Model & IRS-Model & GRTS-Model \\
+\bottomrule
+\end{longtable}
+
+We used a crossed design with the simulation parameters given in Table
+\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
+exponential correlation with an effective range of \(\sqrt{2}\) for
+\(N = 900\) response values simulated on the unit square in either
+random locations (Site Locations = Random) or gridded locations (Site
+Locations = Gridded). For the lognormal scenarios, the response values
+were simulated using the specified correlation parameters using a normal
+distribution and were subsequently exponentiated. A total variance of 2
+and a mean of 0 on the normal scale is equivalent to a total variance of
+47 and a mean of 2.72 after exponentiation. Therefore, when the
+model-based methods were used, the correlation is mis-specified. We
+chose to simulate values with a lognormal distribution so that we could
+test the model-based analysis approach with a mis-specified model and so
+that we could test both analysis approaches on data that exhibits a
+large amount of skewness (Figure \ref{fig:figlog}).
+
+\begin{longtable}[]{@{}llll@{}}
+\caption{\label{tab:parmtab} Simulation parameters. Total variability
+for all scenarios was 2 so that the partial sill was 0, 1, or
+1.8.}\tabularnewline
+\toprule
+& & & \\
+\midrule
+\endfirsthead
+\toprule
+& & & \\
+\midrule
+\endhead
+Sample Size (n) & 50 & 100 & 200 \\
+Site Locations & Random & Gridded & \\
+Partial Sill Variance Ratio & 0 & 0.5 & 0.9 \\
+Response Type & Normal & Lognormal & \\
+\bottomrule
+\end{longtable}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figlog-1} \caption{Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
+The distribution exhibits moderate to strong right-skewness.}\label{fig:figlog}
+\end{figure}
 
@@
     47)
-  \item
-    lognormalbig: log scale mean 0, log scale variance 4 (total variance
-    2,926)
   \end{itemize}
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
 

### 2021-10-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
minor graph update.

@@
   facet_grid(psill ~ resptype) +
-  scale_colour_manual(values = c("mediumpurple1", "mediumpurple4"))
+  scale_colour_manual(values = c("mediumpurple1", "mediumpurple4")) +
+  theme_bw()
 ```

### 2021-10-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated figure and added plot notes.

@@
 -->
+
+Plot Note: change to 3 x 3, add third column, investigate colour vs. size.
 
@@
 
-ggplot(data = combo_data, aes(x = n, y = rel_efficiency, colour = approach)) +
+ggplot(data = combo_data, aes(x = factor(n), y = rel_efficiency, colour = approach)) +
   facet_grid(psill ~ resptype) +
@@
 ```
+
+Plot Note: change colours and think about shape
+change partial sill variance to proportion of partial sill
+think about legend going on graph
+
+Plot Note: Think about raw data.
 
@@
 
-ggplot(data = combo_data_grts_model, aes(x = n, y = rel_efficiency)) +
+ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
   facet_grid(psill ~ resptype) +
@@
 ```
+
+Plot Note: Do we want this figure?
 
@@
 Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.
+
+Plot Note: Change to coverage plot to include all types
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
edited conceptual Design v Model Figure 1.

@@
   - code: STLAW
-    address: Saint Lawrence University Department of Math, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
+    address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
   - code: OSU
@@
 
-Plot Note: change to 3 x 3, add third column, investigate colour vs. size.
-
 ```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
 Points circled are those that are sampled.
-In the top row, we have one fixed population, and two random samples.
-In the bottom row, we have two realizations of the same spatial process sampled at the same locations."}
-set.seed(08032021)
+In the top row, we have one fixed population, and three random samples.
+In the bottom row, we have three realizations of the same spatial process sampled at the same locations."}
+set.seed(10102021)
 
 source(here("R", "sim_pop.R"))
-N <- 25
+N <- 9
 pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
@@
 
-n <- 5
+n <- 4
 samp_df <- pop_df %>% sample_n(n) 
 samp_df2 <- pop_df %>% sample_n(n) 
+samp_df3 <- pop_df %>% sample_n(n)
 
@@
   theme_minimal_grid() +
-  theme(axis.title.x = element_blank(),
-        axis.title.y = element_blank()) +
-  ggtitle("Design 1")
+  labs(title = "Design 1",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
 
@@
   theme_minimal_grid() +
-  theme(axis.title.x = element_blank(),
-        axis.title.y = element_blank()) +
-  ggtitle("Design 2")
+  labs(title = "Design 2",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
+
+d3 <- ggplot(data = pop_df, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df3, size = 5.5, shape = 1, show.legend = FALSE,
+             stroke = 1.5) +
+  theme_minimal_grid() +
+  labs(title = "Design 3",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(), 
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
 
@@
                        psill = 0.9, erange = 1, nugget = 0.1)
+pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
+                      psill = 0.9, erange = 1, nugget = 0.1)
+
 indeces <- sample(1:N, size = n, replace = FALSE)
@@
 samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
+samp_df_mod3 <- pop_df_mod3 %>% slice(indeces)
 
@@
   theme_minimal_grid() +
-  theme(axis.title.x = element_blank(),
-        axis.title.y = element_blank()) +
-  ggtitle("Model 1")
+  labs(title = "Model 1",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(), 
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
 
@@
   theme_minimal_grid() +
-  theme(axis.title.x = element_blank(),
-        axis.title.y = element_blank()) +
-  ggtitle("Model 2")
+  labs(title = "Model 2",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(), 
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
 
-library(gridExtra)
-grid.arrange(d1, d2, m1, m2)
+m3 <- ggplot(data = pop_df_mod3, aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
+  theme_minimal_grid() +
+  labs(title = "Model 3",
+       colour = "Response") +
+  theme(axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank(), 
+        legend.position = "none",
+        panel.grid.major = element_blank()) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
+
+library(ggpubr)
+ggarrange(d1, d2, d3, m1, m2, m3, nrow = 2, ncol = 3,
+          common.legend = TRUE, legend = "right")
 ```

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
edited conceptual Design v Model Figure 1.

@@
 SW 35th St, Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Math,
+    \address[STLAW]{Saint Lawrence University Department of Mathematics,
 Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
@@
 Points circled are those that are sampled.
-In the top row, we have one fixed population, and two random samples.
-In the bottom row, we have two realizations of the same spatial process sampled at the same locations.}\label{fig:fig1}
+In the top row, we have one fixed population, and three random samples.
+In the bottom row, we have three realizations of the same spatial process sampled at the same locations.}\label{fig:fig1}
 \end{figure}
@@
 
+Plot Note: change colours and think about shape change partial sill
+variance to proportion of partial sill think about legend going on graph
+
+Plot Note: Think about raw data.
+
 MAJOR POINTS for the following Figure:
@@
 \includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+Plot Note: Do we want this figure?
 
@@
 \end{itemize}
+
+Plot Note: Change to coverage plot to include all types
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
revised text for first conceptual figure.

@@
 
-Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions.
+Treating the data as fixed and incorporating randomness through the sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here since our figure is similar?))) yields estimators having very few other assumptions.
 Confidence intervals for these types of estimators are typically derived using limiting arguments.
 Means and totals, for example, are asymptotically normally distributed by the Central Limit Theorem.
+If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
 @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
@@
 The model-based approach assumes the data are a random realization of a data-generating process.
-Randomness is often incorporated through distributional assumptions on this process.
+Randomness is often incorporated through distributional assumptions on this process and need not be incorporated through random sampling (bottom row of Figure \ref{fig:fig1}).
 Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
@@
 We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-But we are often not interested in the true, unknown mean of the underlying process.
 
@@
 The distributional assumptions allow the prediction intervals to be more precise.
+If we repeatedly generate the response values from a fixed spatial process and obtain a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-
-<!-- Figure 1a.
-Brus (2020): Data is fixed.
-In a finite population example, show a 3d surface that can be generated by anything.
-If we repeatedly sample the surface, then 95% of all 95% CIs will contain the true mean, which never changes.
--->
 
@@
 Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations."}
+In the top row, we have one fixed population, and three random samples of n = 4.
+The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
+In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
+The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels."}
 set.seed(10102021)
@@
 ```
-
-Description of Figure \ref{fig:fig1} goes here.
-
-<!-- Figure 1b.
-Spatial process is fixed.
-In a finite population example, show 10 3d surfaces that are generated from some model.
-If we repeatedly generate the surface and obtain a sample, then 95% of all 95% PIs will contain the realized means.
-The realized mean changes from surface to surface and it's not necessarily the case that 95% of all 95% PIs will contain the true, underlying mean.
--->
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
revised text for first conceptual figure.

@@
 Treating the data as fixed and incorporating randomness through the
-sampling design yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived
-using limiting arguments. Means and totals, for example, are
-asymptotically normally distributed by the Central Limit Theorem.
-Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
-design-based approach.
+sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here
+since our figure is similar?))) yields estimators having very few other
+assumptions. Confidence intervals for these types of estimators are
+typically derived using limiting arguments. Means and totals, for
+example, are asymptotically normally distributed by the Central Limit
+Theorem. If we repeatedly sample the surface, then 95\% of all 95\%
+confidence intervals constructed from a procedure with appropriate
+coverage will contain the true, fixed mean. Särndal et al. (2003) and
+Lohr (2009) provide thorough reviews of the design-based approach.
 
@@
 data-generating process. Randomness is often incorporated through
-distributional assumptions on this process. Instead of estimating fixed
-but unknown parameters (as in the design-based approach), the goal of
-model-based inference in the spatial context is often \emph{prediction}
-of an unknown quantity. For example, suppose the realized mean of all
-population units is the quantity of interest. Instead of
-\emph{estimating} a fixed unknown mean, we are \emph{predicting} the
-value of the mean, a random variable. We know that if we sampled all
-population units, we would have an exact prediction for the mean of our
-one realized process, without any uncertainty. But we are often not
-interested in the true, unknown mean of the underlying process.
+distributional assumptions on this process and need not be incorporated
+through random sampling (bottom row of Figure \ref{fig:fig1}). Instead
+of estimating fixed but unknown parameters (as in the design-based
+approach), the goal of model-based inference in the spatial context is
+often \emph{prediction} of an unknown quantity. For example, suppose the
+realized mean of all population units is the quantity of interest.
+Instead of \emph{estimating} a fixed unknown mean, we are
+\emph{predicting} the value of the mean, a random variable. We know that
+if we sampled all population units, we would have an exact prediction
+for the mean of our one realized process, without any uncertainty.
 
@@
 distributional assumptions allow the prediction intervals to be more
-precise. Cressie (1993) and Schabenberger and Gotway (2017) provide
-reviews of model-based approaches for spatial data.
+precise. If we repeatedly generate the response values from a fixed
+spatial process and obtain a sample, then 95\% of all 95\% prediction
+intervals constructed from a procedure with appropriate coverage will
+contain their respective realized means. Cressie (1993) and
+Schabenberger and Gotway (2017) provide reviews of model-based
+approaches for spatial data.
 
@@
 Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.}\label{fig:fig1}
+In the top row, we have one fixed population, and three random samples of n = 4.
+The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
+In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
+The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels.}\label{fig:fig1}
 \end{figure}
-
-Description of Figure \ref{fig:fig1} goes here.
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
edited text around Figure on relative efficiency.

@@
              show.legend = FALSE, stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Design 1",
@@
              stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Design 2",
@@
              stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Design 3",
@@
   geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Model 1",
@@
   geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Model 2",
@@
   geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_minimal_grid() +
+  theme_bw() +
   labs(title = "Model 3",
@@
 All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded).
+The mean for the spatial process generating the response was set to zero.
+
 For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
 A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
-Therefore, when the model-based methods were used, the correlation is mis-specified.
-We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness (Figure \ref{fig:figlog}).
+Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
+We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.
 
@@
 row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
-                         "Partial Sill Variance Ratio", "Response Type")
+                         "Partial Sill / Total Variance", "Response Type")
 kable(parm_tab, col.names = c("", "", ""),
@@
 ```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness."}
+The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
 mu <- 0; var <- 2
@@
 
-* correlation type: dependent errors or independent errors
-* error type:
-    * normal: mean 0, variance 2
-    * lognormal: log scale mean 0, log scale variance 2 (total variance 47)
-    <!-- * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926) -->
-* sample sizes: $n = 10, 50, 150; N = 900$
-* layout: gridded vs random uniform population locations confined to a 1 x 1 unit square
-
-So for example, the `inderror.normal.n50.randloc` is the simulation having independent random errors that are normal, a sample size of 50, and random population locations.
-
-There were 2000 trials for each simulation.
-The original response (before exponentiating if applicable) for the dependent error cases was normally distributed with an exponential covariance function with partial sill of 0.9, effective range of $\sqrt{2}$, and a nugget of 0.1.
-For the independent error cases, the partial sill was 0 and the nugget was 1.
-
-__Lisa M.__ Notes: adding an intermediate level of spatial dependency?
-Transfer simulation scenarios to a table
-Explain what the effective range is
-Fill in the details of what exactly each approach means (perhaps a table would be a good way to do this)
-Reorder the sims in the first figure by some criterion.
-Think about what would be a "reasonable sample size" instead of 10.
-Define medae
-If the data has a large right-skew, wouldn't one consider a transformation before the analysis?
-We should address this by stating that the BLUP for the log response does not mean that e^logBLUP is the BLUP for the response on the original scale.
-
-In each simulated data set, a GRTS sample and an IRS sample were selected.
-Then for the GRTS sample, the design-based approach using the local neighborhood variance (Design GRTS) and a model-based approach were applied (Model GRTS).
-Then for the IRS sample, the design-based approach using the simple random sample variance (Design IRS) and a model-based approach were applied (Model IRS).
+There were 2000 simulation trials for each of the 36 parameter combinations.
+In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected.
+For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model).
+For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).
 
@@
 
-MAJOR POINTS for the following Figure, which has relative efficiency (rmspe / Design IRS rmspe): We see that
+Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
+\mbox{}
+\begin{equation*}
+\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+\end{equation*}
 
-- When there is no spatial correlation (top row), we aren't losing that much by using a spatial method over Design IRS, even if assumptions are violated.
-- When there is a lot of spatial correlation Model GRTS tends to perform best, but difference in relative efficiency between Model GRTS and Design GRTS is not very big.
-In many settings Design GRTS outperforms Model IRS by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is much more important than the analysis decision (whether to analyze using model assumptions or not).
-- If there is a large amount of spatial correlation, we should __not__ use IRS.
-Even though it's assumptions are satisfied, the resulting estimator is much worse than an estimator using a spatially balanced sample.
-- If we are comparing design-based and model-based methods, we should not use a poor design to compare with (give examples?).
+where rMSPE is the root-Mean-Squared-Prediction-Error.
+When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
+So, using GRTS or using a spatial model does not result in much loss in efficiency even if the response variable is not spatially correlated.
+When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is not very big.
+In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
 
-```{r, results = "hide", message = FALSE, warning = FALSE}
+Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design.
+However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
+
+```{r figeff, out.width = "100%", fig.cap = "Relative Efficiency of the four design-analysis approaches.
+The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
 library(tidyverse)
@@
   mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  ungroup()
+  ungroup() %>%
+  mutate(psill_ratio = 1 - nugget_ratio) %>%
+  mutate(n_factor = factor(n))
 
-ggplot(data = combo_data, aes(x = factor(n), y = rel_efficiency, colour = approach)) +
-  facet_grid(psill ~ resptype) +
-  geom_jitter(width = 7) +
-  scale_colour_manual(values = c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4"))
+effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
+                              colour = approach)) +
+  facet_grid(psill_ratio ~ resptype) +
+  geom_jitter(width = 0.19, size = 2.5) +
+  scale_colour_manual(values = c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")) +
+  theme_bw(base_size = 14) +
+  labs(x = "Sample Size", colour = "Approach",
+       y = "Relative Efficiency")
+effplot
 ```
@@
 Plot Note: change colours and think about shape
-change partial sill variance to proportion of partial sill
 think about legend going on graph
 
-Plot Note: Think about raw data.
-
-MAJOR POINTS for the following Figure:
-
-- when we drop the IRS samples to compare the GRTS samples more closely (looking at the relative efficiency of model rmspe / design rmspe), we see that the model-based approach is usually better than the design-based approach (but, keep in mind that the improvement was small compared to the improvement of both methods over Design IRS).
-- when the model used is the same model that generated the data, the Model-based approach far outperforms the Design-based approach, especially when there is a lot of spatial correlation (bottom-left facet).
-The methods perform similarly when there is no spatial correlation.
-- even when the model that generates the data is different than the model used to fit the data (lognormal), the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
-
-```{r}
+```{r, fig.keep = "none"}
 combo_data_grts <- combo_data %>%
@@
 ```
-
-Plot Note: Do we want this figure?
 
@@
 Generally, the larger the skew, the worse the coverage,and the larger the sample size, the better the coverage.
-Design GRTS (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
+GRTS-Design (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
 
 ```{r}
-## View(combo_data %>% filter(approach == "Design IRS"))
+## View(combo_data %>% filter(approach == "IRS-Design"))
 ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
@@
 ## combine results
-res_df <- tibble(approach = c("Design IRS", "Design GRTS", "Model IRS", "Model GRTS"),
+res_df <- tibble(approach = c("IRS-Design", "GRTS-Design", "IRS-Model", "GRTS-Model"),
        realized_mean = c(realized_mean, realized_mean, realized_mean,

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
edited text around Figure on relative efficiency.

@@
 random locations (Site Locations = Random) or gridded locations (Site
-Locations = Gridded). For the lognormal scenarios, the response values
-were simulated using the specified correlation parameters using a normal
-distribution and were subsequently exponentiated. A total variance of 2
-and a mean of 0 on the normal scale is equivalent to a total variance of
-47 and a mean of 2.72 after exponentiation. Therefore, when the
-model-based methods were used, the correlation is mis-specified. We
+Locations = Gridded). The mean for the spatial process generating the
+response was set to zero.
+
+For the lognormal scenarios, the response values were simulated using
+the specified correlation parameters using a normal distribution and
+were subsequently exponentiated. A total variance of 2 and a mean of 0
+on the normal scale is equivalent to a total variance of 47 and a mean
+of 2.72 after exponentiation. Therefore, when the model-based methods
+were used for lognormal response, the correlation was mis-specified. We
 chose to simulate values with a lognormal distribution so that we could
@@
 that we could test both analysis approaches on data that exhibits a
-large amount of skewness (Figure \ref{fig:figlog}).
+large amount of skewness.
 
@@
 Site Locations & Random & Gridded & \\
-Partial Sill Variance Ratio & 0 & 0.5 & 0.9 \\
+Partial Sill / Total Variance & 0 & 0.5 & 0.9 \\
 Response Type & Normal & Lognormal & \\
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figlog-1} \caption{Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness.}\label{fig:figlog}
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  correlation type: dependent errors or independent errors
-\item
-  error type:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    normal: mean 0, variance 2
-  \item
-    lognormal: log scale mean 0, log scale variance 2 (total variance
-    47)
-  \end{itemize}
-\item
-  sample sizes: \(n = 10, 50, 150; N = 900\)
-\item
-  layout: gridded vs random uniform population locations confined to a 1
-  x 1 unit square
-\end{itemize}
-
-So for example, the \texttt{inderror.normal.n50.randloc} is the
-simulation having independent random errors that are normal, a sample
-size of 50, and random population locations.
-
-There were 2000 trials for each simulation. The original response
-(before exponentiating if applicable) for the dependent error cases was
-normally distributed with an exponential covariance function with
-partial sill of 0.9, effective range of \(\sqrt{2}\), and a nugget of
-0.1. For the independent error cases, the partial sill was 0 and the
-nugget was 1.
-
-\textbf{Lisa M.} Notes: adding an intermediate level of spatial
-dependency? Transfer simulation scenarios to a table Explain what the
-effective range is Fill in the details of what exactly each approach
-means (perhaps a table would be a good way to do this) Reorder the sims
-in the first figure by some criterion. Think about what would be a
-``reasonable sample size'' instead of 10. Define medae If the data has a
-large right-skew, wouldn't one consider a transformation before the
-analysis? We should address this by stating that the BLUP for the log
-response does not mean that e\^{}logBLUP is the BLUP for the response on
-the original scale.
-
-In each simulated data set, a GRTS sample and an IRS sample were
-selected. Then for the GRTS sample, the design-based approach using the
-local neighborhood variance (Design GRTS) and a model-based approach
-were applied (Model GRTS). Then for the IRS sample, the design-based
-approach using the simple random sample variance (Design IRS) and a
-model-based approach were applied (Model IRS).
+There were 2000 simulation trials for each of the 36 parameter
+combinations. In each trial, response values were generated from a
+spatial process with the specified parameters, and a GRTS sample and an
+IRS sample were selected. For the GRTS sample, the design-based approach
+using the local neighborhood variance (GRTS-Design) and a model-based
+approach were applied (GRTS-Model). For the IRS sample, the design-based
+approach using the simple random sample variance (IRS-Design) and a
+model-based approach were applied (IRS-Model).
 
@@
 
-MAJOR POINTS for the following Figure, which has relative efficiency
-(rmspe / Design IRS rmspe): We see that
+Figure \ref{fig:figeff} shows the relative efficiency of the four
+approaches from Table \ref{tab:designanalysis} with ``IRS-Design'' as
+the baseline: \mbox{} \begin{equation*}
+\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+\end{equation*}
 
-\begin{itemize}
-\tightlist
-\item
-  When there is no spatial correlation (top row), we aren't losing that
-  much by using a spatial method over Design IRS, even if assumptions
-  are violated.
-\item
-  When there is a lot of spatial correlation Model GRTS tends to perform
-  best, but difference in relative efficiency between Model GRTS and
-  Design GRTS is not very big. In many settings Design GRTS outperforms
-  Model IRS by a large margin, suggesting that the design decision
-  (whether to use IRS or GRTS) is much more important than the analysis
-  decision (whether to analyze using model assumptions or not).
-\item
-  If there is a large amount of spatial correlation, we should
-  \textbf{not} use IRS. Even though it's assumptions are satisfied, the
-  resulting estimator is much worse than an estimator using a spatially
-  balanced sample.
-\item
-  If we are comparing design-based and model-based methods, we should
-  not use a poor design to compare with (give examples?).
-\end{itemize}
+where rMSPE is the root-Mean-Squared-Prediction-Error. When there is no
+spatial correlation (top row), the four approaches have approximately
+equal rMSPE, even when the assumptions of the model-based approaches are
+violated. So, using GRTS or using a spatial model does not result in
+much loss in efficiency even if the response variable is not spatially
+correlated. When there is high spatial correlation (bottom row), the
+GRTS-Model approach tends to perform best, but difference in relative
+efficiency between GRTS-Model and GRTS-Design is not very big. In the
+lognormal, high partial sill settings (bottom-right facet), GRTS-Design
+outperforms IRS-Model by a large margin, suggesting that the design
+decision (whether to use IRS or GRTS) is more important than the
+analysis decision (whether to analyze using model assumptions or not).
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
+Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the
+assumptions for GRTS-Model are satisfied, the approach outperforms
+GRTS-Design. However, even when the model that generates the data is
+different than the model used to fit the data, as in the lognormal
+response, the model-based approach still outperforms the design-based
+approach when there is a high amount of spatial correlation.
 
-Plot Note: change colours and think about shape change partial sill
-variance to proportion of partial sill think about legend going on graph
+\begin{figure}
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figeff-1} \caption{Relative Efficiency of the four design-analysis approaches.
+The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figeff}
+\end{figure}
 
-Plot Note: Think about raw data.
-
-MAJOR POINTS for the following Figure:
-
-\begin{itemize}
-\tightlist
-\item
-  when we drop the IRS samples to compare the GRTS samples more closely
-  (looking at the relative efficiency of model rmspe / design rmspe), we
-  see that the model-based approach is usually better than the
-  design-based approach (but, keep in mind that the improvement was
-  small compared to the improvement of both methods over Design IRS).
-\item
-  when the model used is the same model that generated the data, the
-  Model-based approach far outperforms the Design-based approach,
-  especially when there is a lot of spatial correlation (bottom-left
-  facet). The methods perform similarly when there is no spatial
-  correlation.
-\item
-  even when the model that generates the data is different than the
-  model used to fit the data (lognormal), the model-based approach still
-  outperforms the design-based approach when there is a high amount of
-  spatial correlation.
-\end{itemize}
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-Plot Note: Do we want this figure?
+Plot Note: change colours and think about shape think about legend going
+on graph
 
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 the worse the coverage,and the larger the sample size, the better the
-coverage. Design GRTS (local neighborhood variance) tended to slightly
+coverage. GRTS-Design (local neighborhood variance) tended to slightly
 undercover, a result Tony was familiar with.
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
 \endhead
-Design IRS & 112.7 & 8.8 & 95.4 & 129.9 \\
-Model IRS & 110.5 & 7.9 & 95.0 & 125.9 \\
-Design GRTS & 101.8 & 6.1 & 89.8 & 113.7 \\
-Model GRTS & 102.3 & 5.9 & 90.8 & 113.9 \\
+IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\
+IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\
+GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\
+GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\
 \bottomrule

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated the coverage plot and cut the text from the old simulations.

@@
 
+colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
@@
   geom_jitter(width = 0.19, size = 2.5) +
-  scale_colour_manual(values = c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")) +
+  scale_colour_manual(values = colour_scale) +
   theme_bw(base_size = 14) +
@@
 
-Plot Note: Change to coverage plot to include all types
-
 ```{r}
-ggplot(data = combo_data_grts, aes(x = n, y = coverage, colour = approach)) +
+ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 7) + 
-  facet_grid(psill ~ resptype) +
-  scale_colour_manual(values = c("mediumpurple1", "mediumpurple4")) +
-  theme_bw()
+  geom_jitter(width = 0.19, size = 2.5) + 
+  facet_grid(psill_ratio ~ resptype) +
+  scale_colour_manual(values = colour_scale) +
+  ylim(0.80, 1) +
+  theme_bw(base_size = 14) +
+  labs(x = "Sample Size", colour = "Approach",
+       y = "Interval Coverage")
 ```
-
-Major Points from August 3 Simulations:
-
-1.
-In most of the dependent error simulation settings, either all four approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model) perform equally or GRTS-Design and GRTS-Model outperform IRS-Design and IRS-Model.
-Exceptions to this are a couple of the settings with very small sample sizes (n = 10), in which the IRS does better than GRTS.
-In the independent error settings, it usually doesn't matter much which approach is used, which makes sense.
-
-2.
-We will now focus on comparing Design-GRTS to Model-GRTS, the two best approaches for any reasonable sample size.
-In the independent error settings, the two approaches perform very similarly, so those results are omitted in the following graph.
-In the dependent error settings, using rmspe as the performance criterion, Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two approaches perform very similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3 settings.
-
-3.
-Focusing in on the three settings where Design-GRTS outperforms Model-GRTS, we see that, in two of the settings, the log-normal response has a large variance, corresponding to a large right-skew after exponentiation.
-All three settings have sites in random locations.
-However, in only one of these settings would we recommend actually using Design-GRTS.
-In the other two settings, the data are sufficiently skewed that a practitioner should not use either approach, though it is "safer" to use Design-GRTS.
-
-4. Coverage
-
-For Gaussian errors, coverage for all approaches tended to be near 0.95.
-There was less between-approach deviation in coverages for random locations compared to grid locations.
-Generally, the larger the skew, the worse the coverage,and the larger the sample size, the better the coverage.
-GRTS-Design (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.
-
-```{r}
-## View(combo_data %>% filter(approach == "IRS-Design"))
-ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
-  geom_hline(yintercept = 0.95) +
-  geom_point() +
-  coord_flip()
-```
-
-5. Take-home messages
-
-- In terms of rmspe, a model-based analysis on a GRTS design yields an rmspe similar to or lower than a design-based analysis on a GRTS design, as long as the response variable is not "too skewed." 
-- If the response variable is very skewed, then neither analysis is appropriate, but, the design-based analysis is better. 
-- a spatially balanced GRTS sample outperforms IRS in nearly all dependent error settings, as expected.
-- methods that use spatial correlation generally perform better on random location points than they do on gridded points.
-This makes some intuitive sense because (1) on average, the minimum distance between an unobserved point and its nearest observed neighbor should be lower for random points and (2) the span of the study area is maximized for a grid based on the way that we set up the simulations (with the random points being drawn as uniform random variables within the boundary of the grid).
-- comparison of Design-GRTS and Model-GRTS between two settings with different locations of points, but otherwise the same simulation parameters, should really be done on the same surface realization.
-One very strange realized response vector could drastically alter the results, especially on the exponentiated log data.
-In the same way that we compare the four approaches on the same realized data, we should also try to do the same with the locations, if they are of interest.
-(The realized mean won't be exactly the same but should be close).
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated the coverage plot and cut the text from the old simulations.

@@
 
-Plot Note: Change to coverage plot to include all types
-
 \includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-Major Points from August 3 Simulations:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  In most of the dependent error simulation settings, either all four
-  approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model)
-  perform equally or GRTS-Design and GRTS-Model outperform IRS-Design
-  and IRS-Model. Exceptions to this are a couple of the settings with
-  very small sample sizes (n = 10), in which the IRS does better than
-  GRTS. In the independent error settings, it usually doesn't matter
-  much which approach is used, which makes sense.
-\item
-  We will now focus on comparing Design-GRTS to Model-GRTS, the two best
-  approaches for any reasonable sample size. In the independent error
-  settings, the two approaches perform very similarly, so those results
-  are omitted in the following graph. In the dependent error settings,
-  using rmspe as the performance criterion, Model-GRTS outperforms
-  Design-GRTS in 12 of the 18 settings, the two approaches perform very
-  similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3
-  settings.
-\item
-  Focusing in on the three settings where Design-GRTS outperforms
-  Model-GRTS, we see that, in two of the settings, the log-normal
-  response has a large variance, corresponding to a large right-skew
-  after exponentiation. All three settings have sites in random
-  locations. However, in only one of these settings would we recommend
-  actually using Design-GRTS. In the other two settings, the data are
-  sufficiently skewed that a practitioner should not use either
-  approach, though it is ``safer'' to use Design-GRTS.
-\item
-  Coverage
-\end{enumerate}
-
-For Gaussian errors, coverage for all approaches tended to be near 0.95.
-There was less between-approach deviation in coverages for random
-locations compared to grid locations. Generally, the larger the skew,
-the worse the coverage,and the larger the sample size, the better the
-coverage. GRTS-Design (local neighborhood variance) tended to slightly
-undercover, a result Tony was familiar with.
-
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{4}
-\tightlist
-\item
-  Take-home messages
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  In terms of rmspe, a model-based analysis on a GRTS design yields an
-  rmspe similar to or lower than a design-based analysis on a GRTS
-  design, as long as the response variable is not ``too skewed.''
-\item
-  If the response variable is very skewed, then neither analysis is
-  appropriate, but, the design-based analysis is better.
-\item
-  a spatially balanced GRTS sample outperforms IRS in nearly all
-  dependent error settings, as expected.
-\item
-  methods that use spatial correlation generally perform better on
-  random location points than they do on gridded points. This makes some
-  intuitive sense because (1) on average, the minimum distance between
-  an unobserved point and its nearest observed neighbor should be lower
-  for random points and (2) the span of the study area is maximized for
-  a grid based on the way that we set up the simulations (with the
-  random points being drawn as uniform random variables within the
-  boundary of the grid).
-\item
-  comparison of Design-GRTS and Model-GRTS between two settings with
-  different locations of points, but otherwise the same simulation
-  parameters, should really be done on the same surface realization. One
-  very strange realized response vector could drastically alter the
-  results, especially on the exponentiated log data. In the same way
-  that we compare the four approaches on the same realized data, we
-  should also try to do the same with the locations, if they are of
-  interest. (The realized mean won't be exactly the same but should be
-  close).
-\end{itemize}
 

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated text surrounding the confidence interval figure.

@@
   facet_grid(psill_ratio ~ resptype) +
-  geom_jitter(width = 0.19, size = 2.5) +
+  geom_jitter(width = 0.24, size = 2.5) +
   scale_colour_manual(values = colour_scale) +
@@
 
-MAJOR POINTS for the following Figure:
+Figure \ref{fig:figconf} shows the coverage for each of the four approaches.
+We see that the four approaches have somewhat similar coverages in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
 
-- coverage for the model-based estimator is slightly higher than coverage for the design-based estimator in the normal settings.
-Coverages are about equal in the lognormal settings with a slight edge to model-based (this is the point that will be tougher to explain: I would have expected the design-based estimator to have better coverage in the lognormal settings because it has fewer assumptions).
-- coverage is at or near the nominal 95% in all of the normal settings, where assumptions for the model approach and the design approach are satisfied.
+In the normal response settings, where assumptions for the model-based approaches and the design-based approaches are satisfied, all approaches have coverage around the nominal 0.95.
+Because the intervals are symmetric, normal-based intervals, the coverages are generally lower than the nominal 95%.
+We see that, because the sampling distribution of the mean is asymptotically normal (reference for this?
+a spatial version of CLT), the coverages when the sample size is 200 are much closer to the nominal 95% than the coverages when the sample size is 50.
+In general, the more skewed the distribution of the response, the larger the sample size needed to ensure proper coverage of these normal-based intervals.
+
+Do we want to mention stuff like what's in these next bullet points or no?
+
 - for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean.
@@
 
-```{r}
+```{r figconf, out.width = "100%", fig.cap = "Coverage of the four design-analysis approaches.
+All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
+The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
 ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 0.19, size = 2.5) + 
+  geom_jitter(width = 0.24, size = 2.5) + 
   facet_grid(psill_ratio ~ resptype) +
@@
 
+<!-- not sure why this is getting cutoff. temporary fix: -->
+
+\pagebreak
+
 But, we do note a couple of patterns.

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated text surrounding the confidence interval figure.

@@
 \usepackage{longtable}
-\usepackage{graphicx}
 \ifxetex
@@
 
-MAJOR POINTS for the following Figure:
+Figure \ref{fig:figconf} shows the coverage for each of the four
+approaches. We see that the four approaches have somewhat similar
+coverages in all settings, with GRTS-Design having slightly lower
+coverage when the response is normal.
+
+In the normal response settings, where assumptions for the model-based
+approaches and the design-based approaches are satisfied, all approaches
+have coverage around the nominal 0.95. Because the intervals are
+symmetric, normal-based intervals, the coverages are generally lower
+than the nominal 95\%. We see that, because the sampling distribution of
+the mean is asymptotically normal (reference for this? a spatial version
+of CLT), the coverages when the sample size is 200 are much closer to
+the nominal 95\% than the coverages when the sample size is 50. In
+general, the more skewed the distribution of the response, the larger
+the sample size needed to ensure proper coverage of these normal-based
+intervals.
+
+Do we want to mention stuff like what's in these next bullet points or
+no?
 
@@
 \tightlist
-\item
-  coverage for the model-based estimator is slightly higher than
-  coverage for the design-based estimator in the normal settings.
-  Coverages are about equal in the lognormal settings with a slight edge
-  to model-based (this is the point that will be tougher to explain: I
-  would have expected the design-based estimator to have better coverage
-  in the lognormal settings because it has fewer assumptions).
-\item
-  coverage is at or near the nominal 95\% in all of the normal settings,
-  where assumptions for the model approach and the design approach are
-  satisfied.
 \item
@@
 
-\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+\begin{figure}
+\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figconf-1} \caption{Coverage of the four design-analysis approaches.
+All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
+The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figconf}
+\end{figure}
 
@@
 other data set or even to other potential samples of this data set.
+
+\pagebreak
 

### 2021-10-12 inst/manuscript/manuscript.Rmd
reorganized file structure

--- 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
@@
 grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sprbind(grts_samp)
+grts_bind <- sp_rbind(grts_samp) # Matt you need to install newest version
+# devtools::install_github("USEPA/spsurvey", ref = "main")
 

### 2021-10-12 inst/manuscript/manuscript.tex
reorganized file structure

--- 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200
-SW 35th St, Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics,
-Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
-13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239
-Weniger Hall, Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
-Center, National Oceanic and Atmospheric Administration, Seattle,
-Washington, 98115}
+      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
+Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
+and Statistics, 23 Romoda Drive, Canton, New York, 13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
+Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
+Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
       \cortext[1]{Corresponding Author}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/fig1-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/fig1-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
 Points circled are those that are sampled.
@@
 \toprule
-& IRS & GRTS \\
+& IRS & GRTS\tabularnewline
 \midrule
@@
 \toprule
-& IRS & GRTS \\
+& IRS & GRTS\tabularnewline
 \midrule
 \endhead
-Design & IRS-Design & GRTS-Design \\
-Model & IRS-Model & GRTS-Model \\
+Design & IRS-Design & GRTS-Design\tabularnewline
+Model & IRS-Model & GRTS-Model\tabularnewline
 \bottomrule
@@
 \toprule
-& & & \\
-\midrule
-\endfirsthead
-\toprule
-& & & \\
-\midrule
 \endhead
-Sample Size (n) & 50 & 100 & 200 \\
-Site Locations & Random & Gridded & \\
-Partial Sill / Total Variance & 0 & 0.5 & 0.9 \\
-Response Type & Normal & Lognormal & \\
+Sample Size (n) & 50 & 100 & 200\tabularnewline
+Site Locations & Random & Gridded &\tabularnewline
+Partial Sill / Total Variance & 0 & 0.5 & 0.9\tabularnewline
+Response Type & Normal & Lognormal &\tabularnewline
 \bottomrule
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figeff-1} \caption{Relative Efficiency of the four design-analysis approaches.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figeff-1} \caption{Relative Efficiency of the four design-analysis approaches.
 The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figeff}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figconf-1} \caption{Coverage of the four design-analysis approaches.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1} \caption{Coverage of the four design-analysis approaches.
 All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.
 Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.}\label{fig:figdata}
@@
 \toprule
-Approach & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
 \midrule
@@
 \toprule
-Approach & Estimate & SE & 95\% LB & 95\% UB \\
+Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
 \midrule
 \endhead
-IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\
-IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\
-GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\
-GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\
+IRS-Design & 112.7 & 8.8 & 95.4 & 129.9\tabularnewline
+IRS-Model & 110.5 & 7.9 & 95.0 & 125.9\tabularnewline
+GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7\tabularnewline
+GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9\tabularnewline
 \bottomrule

### 2021-10-12 inst/manuscript/notes_to_do.Rmd
reorganized file structure

--- 2021-06-25 inst/SpatialDVM_Manuscript/notes_to_do.Rmd
@@
 
-* put in SRS as base-line sampling setting.
-* add FPBK as traditional SRS analysis as methods to analyze SRS sample.
-* add FPBK and traditional SRS analysis as additional methods to analyze GRTS sample.
+* put in IRS as base-line sampling setting.
+* add FPBK as traditional IRS analysis as methods to analyze SRS sample.
+* add FPBK and traditional IRS analysis as additional methods to analyze GRTS sample.
 * add GRTS with 10% nugget sites as additional sampling setting? If so, add FPBK to analyze this sample.
+
+* no correlation - normal errors
+* strong correlation - normal errors
+* no correlation - non-normal errors
+* strong correlation - log normal

### 2021-10-12 inst/scripts/sim_deperr_lognormal_n10_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormal_n150_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormal_n150_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormal_n50_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormal_n50_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n10_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n10_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n150_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n150_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n50_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_lognormalbig_n50_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 4
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n10_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n10_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 10, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n150_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n150_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n50_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_deperr_normal_n50_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0.9, # partial sill
-  nugget = total_var * 0.1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_lognormal_n150_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_lognormal_n150_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_lognormal_n50_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_lognormal_n50_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "lognormal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_normal_n150_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_normal_n150_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 150, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_normal_n50_gridloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = TRUE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/scripts/sim_inderr_normal_n50_randloc.R

@@ -1,76 +1,0 @@
-# write results?
-write_out <- TRUE
-
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-
-# parallel
-n_trials <- 2000
-total_var <- 2
-seed <- seq_len(n_trials)
-library(parallel) # load parallel
-library(purrr) # load purrr
-n_cluster <- detectCores() # find cores (48 on mine)
-cluster <- makeCluster(n_cluster) # make cluster
-clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
-clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
-sim_output <- parLapply(
-  cluster, # the cluster
-  seed, # some seeds
-  safely(sim_trial), # see if any errors
-  N = 30^2, # pop size
-  n = 50, # sample size
-  psill = total_var * 0, # partial sill
-  nugget = total_var * 1, # nugget
-  erange = sqrt(4), # effective range
-  gridded = FALSE, # grid (TRUE) or random (FALSE) both in [0, 1] x [0,1]
-  resptype = "normal" # normal response
-)
-stopCluster(cluster) # stop cluster
-
-# model summaries
-library(dplyr)
-
-# check errors
-map(sim_output, "error") %>%
-  bind_rows() %>%
-  nrow()
-
-# do summaries
-safe_output <- map(sim_output, "result") %>%
-  bind_rows()
-
-# write output
-all_output <- safe_output %>%
-  mutate(
-    resid = true_mean - estimate,
-    cover = 1 * (lb <= true_mean & true_mean <= ub)
-  ) %>%
-  group_by(seed) %>%
-  mutate(closer = case_when(
-    resid == min(resid) ~ 1,
-    resid != min(resid) ~ 0
-  ))
-
-# write summaries
-summ_output <- all_output %>%
-  group_by(approach) %>%
-  summarize(
-    mpbias = mean(resid),
-    rmspe = sqrt(mean(resid^2)),
-    medae = median(abs(resid)),
-    coverage = mean(cover),
-    prcloser = mean(closer),
-    med_ci_len = median(ub - lb)
-  )
-
-if (write_out) {
-  library(readr)
[…]

### 2021-10-12 inst/manuscript/manuscript.Rmd
updates to section 1 and 2

@@
 
-There are two general approaches for using data to make statistical inferences about a population: design-based approaches and model-based approaches.
-When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected in a  sample.
+There are two general approaches for using data to make statistical inferences about a population: design-based and model-based.
+When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected and called a sample.
 In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
@@
 
-__Tony O.__: Should this paragraph address that spatial information can be incorporated in the design stage or in the analysis stage (or both).
-In general, it's not clear whether we are referring to site selection process or the estimation process
-
 Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
@@
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
+Thereafter, several comparisons between design-based and model-based for spatial data have been considered [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
 @cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
@@
 @cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
+More recent overviews include @brus2020statistical and @wang2012review.
 
-<!-- Outline for the rest of the paper -->
-
-__Lisa M.__: Add paragraph describing contribution of manuscript.
-
-The rest of this paper is organized as follows.
-In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
-In Section \ref{sec:numstudy}, we use simulated and real data to study the the behavior of both approaches.
+<!-- Paragraph about the contribution of the manuscript and outlining the rest -->
+Though comparisons between design-based and model-based approaches to spatial data have been studied, no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
+In this manuscript, we compare design-based approaches that incorporate spatial locations to model-based approaches for spatial data.
+The rest of the manuscript is organized as follows.
+In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach for spatial data.
+In Section \ref{sec:numstudy}, we use a simulation approach to study the behavior and performance of both approaches.
+In Section \ref{application}, we use both approaches to analyze real data.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 
-## Comparing Design-Based vs. Model-Based
+## Comparing Design-Based and Model-Based Approaches
 
 The design-based approach assumes the population is fixed.
-Randomness is incorporated in the selection of population units according to a sampling design.
-A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit.
+Randomness is incorporated via the selection of units in a sampling frame according to a sampling design.
+A sampling frame is the set of all units available to be sampled.
+A sampling design assigns a positive probability of inclusion (inclusion probability) to each unit in the sampling frame.
 Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
-sampling, which we refer to as Independent Random Sampling (IRS) survey designs.
-The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
-These population parameters are traditionally assumed to be fixed but unknown.
+sampling.
+These sampling designs tend to select units from the sampling frame independently of other units, so we call them "Independent Random Sampling" (IRS) designs.
+The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate parameters (e.g.
+means and totals) of a population.
+Generally, these population parameters are assumed to be fixed, unknown constants.
 
-Treating the data as fixed and incorporating randomness through the sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here since our figure is similar?))) yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived using limiting arguments.
-Means and totals, for example, are asymptotically normally distributed by the Central Limit Theorem.
+Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions.
+Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design.
+Means and totals, for example, are asymptotically Gaussian by the Central Limit Theorem (under some assumptions).
 If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
@@
 
-__Jay VH__: I think it is important to stress that the limiting distribution is over all possible randomizations, constrained by whatever design is used.
-
-__Jay VH__: quantity is vague.
-We should stick with variables, or realized variables (we might also call these values, but we should define and establish a consistent terminology early on.) __Matt H__: I think, though this comment is for this paragraph, we should establish the terminology earlier.
-
 The model-based approach assumes the data are a random realization of a data-generating process.
 Randomness is often incorporated through distributional assumptions on this process and need not be incorporated through random sampling (bottom row of Figure \ref{fig:fig1}).
-Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
-For example, suppose the realized mean of all population units is the quantity of interest.
-Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-
-Assuming the data is a realization of a specific data-generating process yields predictors that are linked to distributional assumptions.
-These distributional assumptions are used to derive prediction intervals.
-The distributional assumptions allow the prediction intervals to be more precise.
-If we repeatedly generate the response values from a fixed spatial process and obtain a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
+Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
+For example, suppose the realized mean of all population units is the value of interest.
+Instead of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the value of the mean, a random variable.
+Prediction intervals are then derived leveraging assumptions of the data generating process.
+If we repeatedly generate the response values from a fixed spatial process and obtained a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
+A visual comparison of the design-based and model-based assumptinos is provided in Figure \ref{fig:fig1} (@brus2020statistical provides a similar figure).
 
@@
 
-# Application
+# Application {#application}
 

### 2021-10-12 inst/manuscript/manuscript.tex
updates to section 1 and 2

@@
 There are two general approaches for using data to make statistical
-inferences about a population: design-based approaches and model-based
-approaches. When data cannot be obtained for all units in a population
-(population units), data on a subset of the population units is
-collected in a sample. In the design-based approach, inferences about
-the underlying population are informed from a probabilistic process in
-which population units are selected to be in the sample. Alternatively,
-in the model-based approach, inferences are made from specific
-assumptions about the underlying process that generated the data. Each
-paradigm has a deep historical context (Sterba, 2009) and its own set of
-general advantages (Hansen et al., 1983).
-
-\textbf{Tony O.}: Should this paragraph address that spatial information
-can be incorporated in the design stage or in the analysis stage (or
-both). In general, it's not clear whether we are referring to site
-selection process or the estimation process
+inferences about a population: design-based and model-based. When data
+cannot be obtained for all units in a population (population units),
+data on a subset of the population units is collected and called a
+sample. In the design-based approach, inferences about the underlying
+population are informed from a probabilistic process in which population
+units are selected to be in the sample. Alternatively, in the
+model-based approach, inferences are made from specific assumptions
+about the underlying process that generated the data. Each paradigm has
+a deep historical context (Sterba, 2009) and its own set of general
+advantages (Hansen et al., 1983).
 
@@
 spatially correlated data. Thereafter, several comparisons between
-design-based and model-based for spatial data have been considered, but
-they tend to compare design-based approaches that ignore spatial
-locations to model-based approaches (Brus and De Gruijter, 1997; Ver
-Hoef, 2002, 2008). Cooper (2006) review the two approaches in an
-ecological context before introducing a ``model-assisted'' variance
-estimator that combines aspects from each approach. In addition to
-Cooper (2006), there has been substantial research and development into
-estimators that use both design and model-based principles (see
-e.g.~Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
-Bayesian approach, and Sterba (2009)). More recent overviews include
-Brus (2020) and Wang et al. (2012), but no numerical comparison has been
-made between design-based approaches that incorporate spatial locations
-and model-based approaches.
+design-based and model-based for spatial data have been considered (Brus
+and De Gruijter, 1997; Ver Hoef, 2002, 2008). Cooper (2006) review the
+two approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. In addition to Cooper (2006), there has been substantial
+research and development into estimators that use both design and
+model-based principles (see e.g.~Cicchitelli and Montanari (2012),
+Chan-Golston et al. (2020) for a Bayesian approach, and Sterba (2009)).
+More recent overviews include Brus (2020) and Wang et al. (2012).
 
-\textbf{Lisa M.}: Add paragraph describing contribution of manuscript.
-
-The rest of this paper is organized as follows. In Section
-\ref{sec:background}, we compare sampling and estimation procedures
-between the design-based approach and the model-based approach. In
-Section \ref{sec:numstudy}, we use simulated and real data to study the
-the behavior of both approaches. And in Section \ref{sec:discussion}, we
-end with a discussion and provide directions for future research.
+Though comparisons between design-based and model-based approaches to
+spatial data have been studied, no numerical comparison has been made
+between design-based approaches that incorporate spatial locations and
+model-based approaches. In this manuscript, we compare design-based
+approaches that incorporate spatial locations to model-based approaches
+for spatial data. The rest of the manuscript is organized as follows. In
+Section \ref{sec:background}, we compare sampling and estimation
+procedures between the design-based approach and the model-based
+approach for spatial data. In Section \ref{sec:numstudy}, we use a
+simulation approach to study the behavior and performance of both
+approaches. In Section \ref{application}, we use both approaches to
+analyze real data. And in Section \ref{sec:discussion}, we end with a
+discussion and provide directions for future research.
 
@@
 
-\hypertarget{comparing-design-based-vs.-model-based}{%
-\subsection{Comparing Design-Based
-vs.~Model-Based}\label{comparing-design-based-vs.-model-based}}
+\hypertarget{comparing-design-based-and-model-based-approaches}{%
+\subsection{Comparing Design-Based and Model-Based
+Approaches}\label{comparing-design-based-and-model-based-approaches}}
 
 The design-based approach assumes the population is fixed. Randomness is
-incorporated in the selection of population units according to a
-sampling design. A sampling design assigns a positive probability of
-inclusion in the sample (inclusion probability) to each population unit.
+incorporated via the selection of units in a sampling frame according to
+a sampling design. A sampling frame is the set of all units available to
+be sampled. A sampling design assigns a positive probability of
+inclusion (inclusion probability) to each unit in the sampling frame.
 Some examples of commonly used sampling designs include simple random
-sampling, stratified random sample, and cluster sampling, which we refer
-to as Independent Random Sampling (IRS) survey designs. The goal is to
-use the sampling design and the sampled data to estimate population
-parameters like means and totals. These population parameters are
-traditionally assumed to be fixed but unknown.
+sampling, stratified random sample, and cluster sampling. These sampling
+designs tend to select units from the sampling frame independently of
+other units, so we call them ``Independent Random Sampling'' (IRS)
+designs. The design-based approach combines the randomness of the
+sampling design and the data collected via the sample to estimate
+parameters (e.g.~means and totals) of a population. Generally, these
+population parameters are assumed to be fixed, unknown constants.
 
 Treating the data as fixed and incorporating randomness through the
-sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here
-since our figure is similar?))) yields estimators having very few other
-assumptions. Confidence intervals for these types of estimators are
-typically derived using limiting arguments. Means and totals, for
-example, are asymptotically normally distributed by the Central Limit
-Theorem. If we repeatedly sample the surface, then 95\% of all 95\%
-confidence intervals constructed from a procedure with appropriate
+sampling design yields estimators having very few other assumptions.
+Confidence intervals for these types of estimators are typically derived
+using limiting arguments that incorporate all possible randomizations of
+sampling units selected via the sampling design. Means and totals, for
+example, are asymptotically Gaussian by the Central Limit Theorem (under
+some assumptions). If we repeatedly sample the surface, then 95\% of all
+95\% confidence intervals constructed from a procedure with appropriate
 coverage will contain the true, fixed mean. Särndal et al. (2003) and
 Lohr (2009) provide thorough reviews of the design-based approach.
-
-\textbf{Jay VH}: I think it is important to stress that the limiting
-distribution is over all possible randomizations, constrained by
-whatever design is used.
-
-\textbf{Jay VH}: quantity is vague. We should stick with variables, or
-realized variables (we might also call these values, but we should
-define and establish a consistent terminology early on.) \textbf{Matt
-H}: I think, though this comment is for this paragraph, we should
-establish the terminology earlier.
 
@@
 approach), the goal of model-based inference in the spatial context is
-often \emph{prediction} of an unknown quantity. For example, suppose the
-realized mean of all population units is the quantity of interest.
-Instead of \emph{estimating} a fixed unknown mean, we are
-\emph{predicting} the value of the mean, a random variable. We know that
-if we sampled all population units, we would have an exact prediction
-for the mean of our one realized process, without any uncertainty.
-
-Assuming the data is a realization of a specific data-generating process
-yields predictors that are linked to distributional assumptions. These
-distributional assumptions are used to derive prediction intervals. The
-distributional assumptions allow the prediction intervals to be more
-precise. If we repeatedly generate the response values from a fixed
-spatial process and obtain a sample, then 95\% of all 95\% prediction
-intervals constructed from a procedure with appropriate coverage will
-contain their respective realized means. Cressie (1993) and
-Schabenberger and Gotway (2017) provide reviews of model-based
-approaches for spatial data.
+often to predict a realized variable, or value. For example, suppose the
+realized mean of all population units is the value of interest. Instead
+of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the
+value of the mean, a random variable. We know that if we sampled all
+population units (and not any units outside the population), we would be
+able to exactly predict (without any uncertainty) the mean of our one
+realized process. Assuming the data are realized from a specific
+data-generating process yields predictions that are linked to
+distributional assumptions. These distributional assumptions are used to
+derive prediction intervals. If we repeatedly generate the response
+values from a fixed spatial process and obtained a sample, then 95\% of
+all 95\% prediction intervals constructed from a procedure with
+appropriate coverage will contain their respective realized means.
+Cressie (1993) and Schabenberger and Gotway (2017) provide reviews of
+model-based approaches for spatial data. A visual comparison of the
+design-based and model-based assumptinos is provided in Figure
+\ref{fig:fig1} (Brus (2020) provides a similar figure).
 

### 2021-10-12 inst/manuscript/notes_Madsen.Rmd
updates to section 1 and 2

# Please Leave Your Notes Here

__Lisa M__: Add a paragraph describing the contribution of the manuscript (before the "the rest of the manuscript is organize as follows etc." in the introduction)

### 2021-10-12 inst/manuscript/notes_VerHoef.Rmd
updates to section 1 and 2

--- 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
@@
----
-title: A comparison of design-based and model-based approaches for spatial data.
-author:
-  - name: In alphabetical order Michael Dumelle
-    email: Dumelle.Michael@epa.gov
-    affiliation: USEPA
-    footnote: 1
-  - name: Matt Higham
-    email: mhigham@stlaw.edu
-    affiliation: STLAW
-    footnote: 1
-  - name: Lisa Madsen
-    affiliation: OSU
-  - name: Anthony R. Olsen
-    affiliation: USEPA
-  - name: Jay M. Ver Hoef
-    affiliation: NOAA
-address:
-  - code: USEPA
-    address: United States Environmental Protection Agency, 200 SW 35th St, Corvallis, Oregon, 97333
-  - code: STLAW
-    address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
-  - code: OSU
-    address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
-  - code: NOAA
-    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
-  This is the abstract.
+# Please Leave Your Notes Here
 
-journal: "An awesome journal"
-date: "`r Sys.Date()`"
-bibliography: mybibfile.bib
-linenumbers: true
-numbersections: true
-csl: elsevier-harvard.csl
-preamble: >
-  \usepackage{bm}
-  \usepackage{bbm}
-  \usepackage{color}
-  \DeclareMathOperator{\var}{{var}}
-  \DeclareMathOperator{\cov}{{cov}}
-output: rticles::elsevier_article
----
-
-_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
-
-Potential Journals:
-
-* Ecological Applications
-* Methods in Ecology and Evolution
-* Journal of Applied Ecology
-* Environmetrics
-* Environmental and Ecological Statistics
-
-<!-- Notes from today -->
-<!-- •	Finite populations require a smaller scope but we need to discuss the impact of infinite populations as well – put finite vs infinite population context in the introduction / similarities and differences -->
-<!-- •	What should a lower sample size be? 50 seems reasonable 30 may be too small – maybe 50, 100, 200 -->
-<!-- •	IRS results seem off for dependent error sample size 50 -->
-<!-- •	Take home points – don’t have to do GRTS design with GRTS sample -->
-<!-- •	Take home points – GRTS is better for both model and design, while spatial approach for design is close to spatial model -->
-<!-- •	Add medium dependence -->
-<!-- •	Remind readers what choices of the four approaches in the application mean – what do they mean to practitioners -->
-
-# Introduction {#sec:introduction}
-
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>", 
-  warning = FALSE, 
-  message = FALSE,
-  include = TRUE, 
-  echo = FALSE
-)
-library(tidyverse)
-library(here)
-library(xtable) # latex tables
-```
-
-<!-- Brief introduction to model based and design based -->
-
-There are two general approaches for using data to make statistical inferences about a population: design-based approaches and model-based approaches.
-When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected in a  sample.
-In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process that generated the data.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
-
-<!-- Shift focus to spatial data -->
-
-__Tony O.__: Should this paragraph address that spatial information can be incorporated in the design stage or in the analysis stage (or both).
-In general, it's not clear whether we are referring to site selection process or the estimation process
-
-Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process.
-@de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
-
-<!-- Outline for the rest of the paper -->
-
-__Lisa M.__: Add paragraph describing contribution of manuscript.
-
-The rest of this paper is organized as follows.
-In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
-In Section \ref{sec:numstudy}, we use simulated and real data to study the the behavior of both approaches.
-And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
-
-# Background {#sec:background}
-
-The design-based and model-based approaches incorporate randomness in fundamentally different ways.
-In this section, we describe the role of randomness and its effects on subsequent inferences.
-We then discuss specific inference methods for the design-based and model-based approaches for spatial data.
-
-## Comparing Design-Based vs. Model-Based
-
-The design-based approach assumes the population is fixed.
-Randomness is incorporated in the selection of population units according to a sampling design.
-A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit.
-Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
-sampling, which we refer to as Independent Random Sampling (IRS) survey designs.
-The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
-These population parameters are traditionally assumed to be fixed but unknown.
-
-Treating the data as fixed and incorporating randomness through the sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here since our figure is similar?))) yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived using limiting arguments.
-Means and totals, for example, are asymptotically normally distributed by the Central Limit Theorem.
-If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
-@sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
-
-__Jay VH__: I think it is important to stress that the limiting distribution is over all possible randomizations, constrained by whatever design is used.
+__Jay VH__: I think it is important to stress that the limiting distribution (related to confidence intervals) is over all possible randomizations, constrained by whatever design is used.
 
@@
 We should stick with variables, or realized variables (we might also call these values, but we should define and establish a consistent terminology early on.) __Matt H__: I think, though this comment is for this paragraph, we should establish the terminology earlier.
-
-The model-based approach assumes the data are a random realization of a data-generating process.
-Randomness is often incorporated through distributional assumptions on this process and need not be incorporated through random sampling (bottom row of Figure \ref{fig:fig1}).
-Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
-For example, suppose the realized mean of all population units is the quantity of interest.
-Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-
-Assuming the data is a realization of a specific data-generating process yields predictors that are linked to distributional assumptions.
-These distributional assumptions are used to derive prediction intervals.
-The distributional assumptions allow the prediction intervals to be more precise.
-If we repeatedly generate the response values from a fixed spatial process and obtain a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
-@cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-
-```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
-Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of n = 4.
-The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
-The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels."}
-set.seed(10102021)
-
-source(here("R", "sim_pop.R"))
-N <- 9
-pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                  psill = 0.9, erange = 1, nugget = 0.1)
-
-n <- 4
-samp_df <- pop_df %>% sample_n(n) 
-samp_df2 <- pop_df %>% sample_n(n) 
-samp_df3 <- pop_df %>% sample_n(n)
-
-library(cowplot)
-d1 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df, size = 5.5, shape = 1, 
-             show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d2 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df2, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d3 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df3, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                       psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-
-indeces <- sample(1:N, size = n, replace = FALSE)
-samp_df_mod <- pop_df_mod %>% slice(indeces)
-samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
-samp_df_mod3 <- pop_df_mod3 %>% slice(indeces)
-
-m1 <- ggplot(data = pop_df_mod, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m2 <- ggplot(data = pop_df_mod2, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m3 <- ggplot(data = pop_df_mod3, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-library(ggpubr)
-ggarrange(d1, d2, d3, m1, m2, m3, nrow = 2, ncol = 3,
-          common.legend = TRUE, legend = "right")
-```
-
-__Tony O.__: Before this section is it useful to have a section that lays out the general site selection and general analysis options.
-Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based.
-Then general analysis options as design-based no spatial, design-based spatial, model-based.
-This four by three table would show that model-based analyses are possible for all selection options.
-Design-based options with no spatial info possible for IRS-based and GRTS-based.
-Design-based options with spatial info possible for GRTS-based.
-
-__Jay VH__: What about the design for model-based inference?
-Strictly speaking, it is fixed – there is no probabilistic use of a randomized design.
-However, we are going to have to deal with Diggle et al.
-(2010).
-
-## Spatially Balanced Design and Analysis
-
-__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
-
-__Jay VH__: Saying "the distribution of the sampled population units mirrors the density of..." is confusing to me.
-Are these formal statistical definitions of distribution (cumulative distribution function) and density (probability density function)?
-Wouldn’t IRS sample be a miniature, as it should, on average, mirror a population?
-
-The design-based approach can use spatial locations to obtain spatially balanced samples.
-First we discuss spatial balance with respect to the population [@stevens2004spatially].
-A sample is spatially balanced with respect to the population if the sampled population units are a miniature of the population units.
-A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
-When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
-
-Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
-The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including __Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
-
-<!-- algorithms. -->
-
-The GRTS algorithm is used to sample from finite and infinite populations and works by utilizing a mapping between two-dimensional and one-dimensional space.
-The population units in two-dimensional space are divided into cells using a hierarchical index.
-Population units are then mapped to a one-dimensional line via the hierarchical indexing.
-The line length of each population unit equals its inclusion probability.
-A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially and @dumelle2021spsurvey provide further details.
-
-<!-- This makes it seem like HT is being used to estimate population parameters in GRTS?
-Needs to be in there, but maybe a sentence making it clear that this isn't being used?
-Or just change the order and talk about Local variance first and then mention these for more general settings?
--->
-
-After collecting a sample using the GRTS algorithm, the data are used to estimate population parameters.
-The Horvitz-Thompson estimator [@horvitz1952generalization] yields unbiased estimates of population means and totals.
-For example, if $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
-\begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align}
-where $Z_i$ and $\pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
-A similar formula exists for estimating the mean, $\mu$.
-@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, and this can be very difficult to calculate.
-Second, they ignore the spatial locations of the population units.
-To address these drawbacks, @stevens2003variance proposed a local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$, and it incorporates spatial locations by assigning higher weights to nearby observations.
-@stevens2003variance show this variance estimators tends to reduce the estimated standard error of $\hat{\tau}$, yielding narrower confidence confidence intervals for $\tau$.
-
-## Finite Population Block Kriging
-
-Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
-Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process.
-@verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-Let $\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
-.
-.
-, \text{z}(s_N) \}$ be a response vector at locations $s_1$, $s_2$, .
-.
-.
-, $s_N$ that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
-Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
-For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
-
-<!-- how explicitly do we want to define \mathbf{z} -->
-
-<!-- finite number of -->
-
-<!-- confusing having two tau's: tau for total and a tau as a function -->
-
-However, we often only have a sample of the $N$ population units.
-Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
-
-\begin{equation}
-\begin{pmatrix} \label{equation:Zmarginal}
-    \mathbf{z}_s      \\
-    \mathbf{z}_u
-\end{pmatrix}
-=
-\begin{pmatrix}
-  \mathbf{X}_s    \\
-  \mathbf{X}_u
-\end{pmatrix}
-\bm{\beta} +
-\begin{pmatrix}
-\bm{\delta}_s    \\
-\bm{\delta}_u
-\end{pmatrix},
-\end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively; $\beta$ is the parameter vector of fixed effects; and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
-Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume the expectation of $\bm{\delta}$ equals $\mathbf{0}$.
-
-We also assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
-It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\cov(\bm{\delta})$ is
-\mbox{}
-\begin{align}\label{equation:expcov}
-\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
-\end{align}
-where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$; and, $\mathbbm{1}$ is equal to 1 when distance $h_i,j$ is equal to 0, and equal to 0 otherwise.
-However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
-80-93].
-
-__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
-
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
-
-We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
-We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
-
-# Numerical Study {#sec:numstudy}
-
-We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.
-
-```{r, echo = FALSE}
-tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-       nrow = 2) %>%
-  as.table()
-row.names(tab) <- c("Design", "Model")
-library(knitr)
-kable(tab, col.names = c("IRS", "GRTS"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The columns give the two types of sampling designs while the rows give the two types of analyses.")
-## I really want "Analysis" to appear as a row header and "Sampling Design" 
-## to appear as a column header, but this isn't easy to figure out.
-## I couldn't figure this out with xtable either so it might need to be
-## made "by hand."
-```
-
-We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded).
-The mean for the spatial process generating the response was set to zero.
-
-For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
-A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
-Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
-We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.
-
-```{r}
-n <- c("50", "100", "200")
-site_locations <- c("Random", "Gridded", "")
-psill_ratio <- c("0", "0.5", "0.9")
-resp_type <- c("Normal", "Lognormal", "")
-parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
-                         "Partial Sill / Total Variance", "Response Type")
-kable(parm_tab, col.names = c("", "", ""),
-      caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-```
-
-```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
-mu <- 0; var <- 2
-xvar <- seq(0.001, 10, by = 0.01)
-yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
-df <- tibble(xvar = xvar, yvar = yvar)
-ggplot(df, aes(x = xvar, y = yvar)) +
-  geom_line()
-```
-
-There were 2000 simulation trials for each of the 36 parameter combinations.
-In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected.
-For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model).
-For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).
-
-```{r, results = "asis", eval = FALSE}
-sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
-colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
-sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
-print(sim_one_table, include.rownames = FALSE, comment = FALSE)
-```
-
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
-We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
-
-Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
-\mbox{}
-\begin{equation*}
-\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
-\end{equation*}
-
-where rMSPE is the root-Mean-Squared-Prediction-Error.
-When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
-So, using GRTS or using a spatial model does not result in much loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is not very big.
-In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
-
-Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design.
-However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
-
-```{r figeff, out.width = "100%", fig.cap = "Relative Efficiency of the four design-analysis approaches.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
-library(tidyverse)
-library(here)
-
-files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
-
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x))
-
-combo_data <- combo_data %>%
-  dplyr::filter(n != 10 & gridded == FALSE) %>%
-  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
-  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
-                                            "Design GRTS", "Model GRTS"))) %>%
-  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal")))
-combo_data <- combo_data %>% group_by(sim) %>% mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
-  fill(designirsrmspe, .direction = "downup") %>%
-  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  ungroup() %>%
-  mutate(psill_ratio = 1 - nugget_ratio) %>%
-  mutate(n_factor = factor(n))
-
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
-                              colour = approach)) +
-  facet_grid(psill_ratio ~ resptype) +
-  geom_jitter(width = 0.19, size = 2.5) +
-  scale_colour_manual(values = colour_scale) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Relative Efficiency")
-effplot
-```
-
-Plot Note: change colours and think about shape
-think about legend going on graph
-
-```{r, fig.keep = "none"}
-combo_data_grts <- combo_data %>%
-  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
-  mutate(designgrts = if_else(approach == "Design GRTS", 
-                              true = rmspe,
-                              false = NA_real_)) %>%
-  group_by(sim) %>%
-  fill(designgrts, .direction = "updown") %>%
-  ungroup() %>%
-  mutate(releffgrts = rmspe / designgrts)
-combo_data_grts_model <- combo_data_grts %>%
-  filter(approach == "Model GRTS")
-
-ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
-  facet_grid(psill ~ resptype) +
-  geom_point()
-```
-
-MAJOR POINTS for the following Figure:
-
-- coverage for the model-based estimator is slightly higher than coverage for the design-based estimator in the normal settings.
-Coverages are about equal in the lognormal settings with a slight edge to model-based (this is the point that will be tougher to explain: I would have expected the design-based estimator to have better coverage in the lognormal settings because it has fewer assumptions).
-- coverage is at or near the nominal 95% in all of the normal settings, where assumptions for the model approach and the design approach are satisfied.
-- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean.
-The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
-- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval.
-Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.
-
-```{r}
-ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
-  geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 0.19, size = 2.5) + 
-  facet_grid(psill_ratio ~ resptype) +
-  scale_colour_manual(values = colour_scale) +
-  ylim(0.80, 1) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Interval Coverage")
-```
-
-# Application
-
-The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
-Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
-Therefore, we will explore whether or not we can still obtain an adequately precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
-
-```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
-library(tidyverse)
-library(here)
-library(SpatialDesignVModel)
-data(nla)
-nla_df
-factor(nla_df$TOTALHG_UNITS) ## all same units
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
-## HG never measured twice at the same site.
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-
-## in duplicate sites, fill in the value for RDis_IX
-nla_nomiss <- nla_df %>%
-  group_by(SITE_ID) %>%
-  fill(RDis_IX, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-nla_nomiss_both <- nla_nomiss
-## check
-nla_nomiss_both %>% group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-
-## transform coordinates
-library(sptotal)
-coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
-                     lat = nla_nomiss_both$INDEX_LAT_DD, 
-                     lon = nla_nomiss_both$INDEX_LON_DD)
-nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
-nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-```
-
-```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.
-Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.", message = FALSE, warning = FALSE}
-merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(colour = "black", fill = "white", bins = 15) +
-  labs(x = "HG (ng / g)")
-# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
-#   geom_histogram(colour = "black", fill = "white", bins = 15) +
-#   labs(x = "Lakeshore Disturbance")
-merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "hg") +
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank()) +
-  coord_quickmap() 
-# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-#   geom_point(aes(colour = RDis_IX)) +
-#   scale_colour_viridis_c() +
-#   labs(x = "", y = "", colour = "Dist") +
-#   theme(axis.text.x = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.ticks = element_blank())
-library(gridExtra)
-# grid.arrange(merc, merc_map, nrow = 2)
-merc_map
-```
-
-```{r, echo = FALSE, results = "hide"}
-mean(nla_nomiss_both$total_hg)
-# mean(nla_nomiss_both$RDis_IX)
-```
-
-Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.2 ng / g.
-
-```{r, results = "hide", fig.keep = "none"}
-## IRS Sample, Model Analysis
-
-set.seed(080421)
-n <- 100
-lakesobs <- nla_nomiss_both %>% sample_n(n) 
-lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
-lakesunobs$total_hg <- NA
-
-lakes_test <- bind_rows(lakesobs, lakesunobs) 
-lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
-
-slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
-                              data = lakes_test, 
-                              xcoordcol = 'xcoords',
-                              ycoordcol = 'ycoords',
-                              CorModel = "Exponential")
-summary(slmfitout_exp_lakes)
-plot(slmfitout_exp_lakes)
-
-pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
-                          conf_level = 0.95)
-print(pred_exp_lakes)
-mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
-se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
-lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
-ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
-
-realized_mean <- mean(nla_nomiss_both$total_hg) ## compute realized mean
-```
-
-```{r, results = "hide"}
-## IRS Sample, Design Analysis
-## 
-irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
-                                   vardoc = var(total_hg))
-N <- nrow(nla_nomiss_both); n <- nrow(lakesobs)
-irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
-irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
-irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
-irs_info$meandoc; sqrt(irs_info$vardoc)
-irs_lb_irs_samp; irs_ub_irs_samp
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Design Analysis
-data_sf <- sf::st_as_sf(nla_nomiss_both, coords = c("xcoords", "ycoords"),
-                        crs = 5070)
-library(spsurvey)
-grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sprbind(grts_samp)
-
-## get coordinates
-grts_coords <- sf::st_coordinates(grts_bind)
-## make data frame
-grts_df <- data.frame(
-  response = grts_bind$total_hg,
-  x = grts_coords[, "X"],
-  y = grts_coords[, "Y"],
-  siteID = grts_bind$siteID,
-  wgt = grts_bind$wgt
-)
-## head(grts_df)
-summary(grts_df$response)
-
-design_analysis <- cont_analysis(
-  grts_df,
-  siteID = "siteID",
-  vars = "response",
-  weight = "wgt",
-  xcoord = "x",
-  ycoord = "y"
-)
-## just return the mean info
-## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
-design_mean_grts <- design_analysis$Mean
-design_mean_grts$Estimate
-design_mean_grts$StdError
-design_mean_grts$LCB95Pct
-design_mean_grts$UCB95Pct
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Model Analysis
-
-grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
-  rename(total_hg = "response", xcoords = "x", ycoords = "y")
-grts_unsamp <- anti_join(nla_nomiss_both, grts_coords_resp)
-grts_unsamp$total_hg <- NA
-grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
-grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
-
-grts_full$wts <- 1 / nrow(grts_full)
-mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
-                   xcoordcol = "xcoords",
-                   ycoordcol = "ycoords")
-pred_mod_grts <- predict(mod_grts, wtscol = "wts")
-model_mean_grts <- pred_mod_grts$FPBK_Prediction
-model_se_grts <- sqrt(pred_mod_grts$PredVar)
-model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
-model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
-model_mean_grts
-model_se_grts
-model_lb_grts
-model_ub_grts
-```
-
-```{r, results = "hide"}
-## combine results
-res_df <- tibble(approach = c("IRS-Design", "GRTS-Design", "IRS-Model", "GRTS-Model"),
-       realized_mean = c(realized_mean, realized_mean, realized_mean,
-                     realized_mean),
-       estimate = c(irs_info$meandoc,
-                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
-                    as.vector(model_mean_grts)),
-       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
-              as.vector(model_se_grts)),
-       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
-              as.vector(model_lb_grts)),
-       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
-              as.vector(model_ub_grts)))
-res_df <- res_df %>% slice(1, 3, 2, 4)
-```
-
-```{r apptab}
-library(knitr)
-res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-kable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 103.2 ng / g")
-```
-
-Table \ref{tab:appliedtab} shows the application of a design-based analysis on an IRS, a model-based analysis on an IRS, a design-based analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
-We see that, for all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
-However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
-
-But, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not use the spatial correlation in mercury concentration across the contiguous United States.
-We also see that, for the samples drawn, the both analyses with the GRTS sampling design have a lower standard error than the analyses with the IRS sampling design.
-We would expect this to be the case for most samples because mercury concentration exhibits spatial correlation so a spatially balanced sample should usually yield a lower standard error.
-If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling just 50 lakes to save time and money.
-
-# Discussion {#sec:discussion}
-
-* Pros of Design-Based (items we are not exploring): computationally efficient, few assumptions, more naturally handles binary data,
-
-* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?, estimated spatial parameters to better understand spatial structure, site-by-site predictions/prediction map
-
-# References {#references .unnumbered}

### 2021-10-12 inst/scripts/all_sims.R
added some comments to simulation script

@@ -23,9 +23,12 @@ parm_df <- expand_grid(N, n, total_var, psill_ratio,
 source("R/sim_pop.R")
 source("R/sim_trial.R")
 
-## loop through each row of parm_df
+# library for later
+# library(SpatialDesignVModel)
+
+## loop through each row of parm_df (could be done with purrr::pmap() instead)
 for (i in 1:nrow(parm_df)) {
-  n_trials <- 5
+  n_trials <- 5 # 2000 for the simulation section in the manuscript
   seed <- sample.int(1e7, size = n_trials)
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
@@ -51,7 +54,7 @@ for (i in 1:nrow(parm_df)) {
 
   stopCluster(cluster) # stop cluster
 
-  # check errors
+  # check errors (deprecated)
   # map(sim_output, "error") %>%
   #   bind_rows() %>%
   #   nrow()

### 2021-10-12 inst/manuscript/manuscript.Rmd
section 2 updates

@@
 ---
-title: A comparison of design-based and model-based approaches for spatial data.
+title: A comparison of design-based and model-based approaches for finite population spatial data.
 author:
-  - name: In alphabetical order Michael Dumelle
+  - name: Michael Dumelle
     email: Dumelle.Michael@epa.gov
@@
     affiliation: STLAW
-    footnote: 1
   - name: Lisa Madsen
@@
 In this manuscript, we compare design-based approaches that incorporate spatial locations to model-based approaches for spatial data.
+Though these comparisons generalize to both finite populations (e.g.
+point resources) and infinite populations (e.g.
+linear and areal resources), we focus on applications to finite populations.
 The rest of the manuscript is organized as follows.
@@
 
-## Comparing Design-Based and Model-Based Approaches
+## Comparing Design-Based and Model-Based Approaches {#subsec:dvm_compare}
 
@@
 A sampling design assigns a positive probability of inclusion (inclusion probability) to each unit in the sampling frame.
-Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
-sampling.
+Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster sampling.
 These sampling designs tend to select units from the sampling frame independently of other units, so we call them "Independent Random Sampling" (IRS) designs.
+Sampling designs incorporating the spatial locations of units in the sample frame are called spatially balanced designs.
+Spatially balanced designs can be obtained using the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
 The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate parameters (e.g.
@@
 The model-based approach assumes the data are a random realization of a data-generating process.
-Randomness is often incorporated through distributional assumptions on this process and need not be incorporated through random sampling (bottom row of Figure \ref{fig:fig1}).
+Randomness is incorporated through distributional assumptions on this process.
+Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
+To guard against preferential sampling, model-based approaches often still implement random sampling.
+
 Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
@@
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-A visual comparison of the design-based and model-based assumptinos is provided in Figure \ref{fig:fig1} (@brus2020statistical provides a similar figure).
+A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@brus2020statistical provides a similar figure).
 
@@
 
-__Tony O.__: Before this section is it useful to have a section that lays out the general site selection and general analysis options.
-Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based.
-Then general analysis options as design-based no spatial, design-based spatial, model-based.
-This four by three table would show that model-based analyses are possible for all selection options.
-Design-based options with no spatial info possible for IRS-based and GRTS-based.
-Design-based options with spatial info possible for GRTS-based.
+## Spatially Balanced Design and Analysis {#subsec:spb_design}
 
-__Jay VH__: What about the design for model-based inference?
-Strictly speaking, it is fixed – there is no probabilistic use of a randomized design.
-However, we are going to have to deal with Diggle et al.
-(2010).
+Spatially balanced samples can be obtained using the design-based approach.
+Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
+The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
+To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voroni polygons.
+Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
+In this manuscript, we use Generalized Random Tessellation Stratified (GRTS) sampling because it has several attractive properties: GRTS sampling accommodates finite and infinite sampling frames; accommodates equal, unequal, and proportional (to) size inclusion probabilities; accommodates legacy (historical) sampling; accommodates a minimum distance between units in a sample; accommodates reverse hierarchically ordered replacement units in a sample (replacement units are units available to be sampled if an original unit cannot be sampled); and is available in the spsurvey R package @dumelle2021spsurvey.
 
-## Spatially Balanced Design and Analysis
-
-__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
-
-__Jay VH__: Saying "the distribution of the sampled population units mirrors the density of..." is confusing to me.
-Are these formal statistical definitions of distribution (cumulative distribution function) and density (probability density function)?
-Wouldn’t IRS sample be a miniature, as it should, on average, mirror a population?
-
-The design-based approach can use spatial locations to obtain spatially balanced samples.
-First we discuss spatial balance with respect to the population [@stevens2004spatially].
-A sample is spatially balanced with respect to the population if the sampled population units are a miniature of the population units.
-A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
-When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
-
-Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
-The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including __Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
-
-<!-- algorithms. -->
-
-The GRTS algorithm is used to sample from finite and infinite populations and works by utilizing a mapping between two-dimensional and one-dimensional space.
-The population units in two-dimensional space are divided into cells using a hierarchical index.
-Population units are then mapped to a one-dimensional line via the hierarchical indexing.
-The line length of each population unit equals its inclusion probability.
-A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample.
+The GRTS algorithm samples from finite and infinite populations by utilizing a mapping between two-dimensional and one-dimensional space.
+The units in the two-dimensional sampling frame are divided into cells using a hierarchical address.
+This hierarchical address is then used to map the units from two-dimensional space to a one-dimensional line where each unit's line length equals its inclusion probability.
+A systematic sample is conducted on the line and linked back to a unit in two-dimensional space, which results in the desired sample.
 @stevens2004spatially and @dumelle2021spsurvey provide further details.
@@
 
-After collecting a sample using the GRTS algorithm, the data are used to estimate population parameters.
-The Horvitz-Thompson estimator [@horvitz1952generalization] yields unbiased estimates of population means and totals.
-For example, if $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
+After selecting a GRTS sample, data are collected and used to estimate population parameters.
+To unbiasedly estimate population means and totals from sample data, one can use the Horvitz-Thompson estimator [@horvitz1952generalization].
+If $\tau$ is a population total, the Horvitz-Thompson estimate of $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
 \begin{align}\label{eq:ht}
@@
 \end{align}
-where $Z_i$ and $\pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
-A similar formula exists for estimating the mean, $\mu$.
+where $Z_i$ is the value of the $i$th unit in the sample and $\pi_i$ is the inclusion probability of the $i$th unit in the sample.
+An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by the population size.
+
+While the Horvitz-Thompson estimator is unbiased for population means and totals, it is also important to quantify the uncertainty in these estimates.
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, and this can be very difficult to calculate.
-Second, they ignore the spatial locations of the population units.
-To address these drawbacks, @stevens2003variance proposed a local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$, and it incorporates spatial locations by assigning higher weights to nearby observations.
-@stevens2003variance show this variance estimators tends to reduce the estimated standard error of $\hat{\tau}$, yielding narrower confidence confidence intervals for $\tau$.
+First, these estimators rely on calculating $\pi_{ij}$, the probability that unit $i$ and unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
+Second, these estimators ignore the spatial locations of the units in the sampling frame.
+To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
+The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
+@stevens2003variance show the local neigborhood variance estimator tends reduce $\text{Var}(\hat{\tau})$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
 
@@
 @verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-Let $\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
+Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
 .
 .
-, \text{z}(s_N) \}$ be a response vector at locations $s_1$, $s_2$, .
+, \text{z}(s_N) \}}$ be a response vector at locations $s_1$, $s_2$, .
 .
@@
 
-However, we often only have a sample of the $N$ population units.
+We often only have a sample of the $N$ population units.
 Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
@@
 
-We also assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
+In addition to assuming the expectation of $\bm{\delta}$ equals $\mathbf{0}$, we also assume that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
 It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\cov(\bm{\delta})$ is
+Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th entry for $\cov(\bm{\delta})$ is
 \mbox{}
 \begin{align}\label{equation:expcov}
-\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
+\cov(\delta_i, \delta_j) = 
+\begin{cases} 
+\sigma^2_{ps}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
+\sigma^2_{ps} + \sigma^2_n & h_{i,j} = 0
+\end{cases}
+,
 \end{align}
-where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$; and, $\mathbbm{1}$ is equal to 1 when distance $h_i,j$ is equal to 0, and equal to 0 otherwise.
-However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
+where $\sigma^2_{ps}$ is the partial sill measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the nugget measuring fine-scale (indepenent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+Any spatial covariance function could be used in the place of the exponential, however, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
 80-93].
 
-__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
-
 With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
+While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based, meaning they don't rely on any distributional assumptions.
 
 We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
+However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
 We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, results = "asis"}
 tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
@@
   as.table()
-row.names(tab) <- c("Design", "Model")
-library(knitr)
-kable(tab, col.names = c("IRS", "GRTS"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The columns give the two types of sampling designs while the rows give the two types of analyses.")
+row.names(tab) <- c("IRS", "GRTS")
+xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
+The rows give the two types of sampling designs while the columns give the two types of analyses.")
+align(xtab1) <- "r|ll"
+names(xtab1) <- c("IRS", "GRTS")
+print(xtab1, comment = FALSE)
 ## I really want "Analysis" to appear as a row header and "Sampling Design" 
@@
 
-```{r}
+```{r, echo = FALSE, results = "asis"}
 n <- c("50", "100", "200")
@@
 parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
-                         "Partial Sill / Total Variance", "Response Type")
-kable(parm_tab, col.names = c("", "", ""),
-      caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+parm_tab[c(2, 4), 3] <- "-"
+row.names(parm_tab) <- c("Sample Size (n)", "Layout", 
+                         "Proportion of Dependent Error", "Response Type")
+# library(knitr) kable with htlm has trouble with page breaks -- kable with latex harder for me than xtable
+# kable(parm_tab, col.names = c("", "", ""),
+#       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+align(parm_xtab) <- "r|lll"
+print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
 ```

### 2021-10-12 inst/manuscript/manuscript.tex
section 2 updates

@@
             pdfauthor={},
-            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
+            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data.},
             colorlinks=false,
@@
 
-  \title{A comparison of design-based and model-based approaches for spatial
-data.}
-    \author[USEPA]{In alphabetical order Michael Dumelle\corref{1}}
+  \title{A comparison of design-based and model-based approaches for finite
+population spatial data.}
+    \author[USEPA]{Michael Dumelle\corref{1}}
    \ead{Dumelle.Michael@epa.gov} 
-    \author[STLAW]{Matt Higham\corref{1}}
+    \author[STLAW]{Matt Higham}
    \ead{mhigham@stlaw.edu} 
@@
 approaches that incorporate spatial locations to model-based approaches
-for spatial data. The rest of the manuscript is organized as follows. In
-Section \ref{sec:background}, we compare sampling and estimation
-procedures between the design-based approach and the model-based
-approach for spatial data. In Section \ref{sec:numstudy}, we use a
-simulation approach to study the behavior and performance of both
-approaches. In Section \ref{application}, we use both approaches to
-analyze real data. And in Section \ref{sec:discussion}, we end with a
-discussion and provide directions for future research.
+for spatial data. Though these comparisons generalize to both finite
+populations (e.g.~point resources) and infinite populations (e.g.~linear
+and areal resources), we focus on applications to finite populations.
+The rest of the manuscript is organized as follows. In Section
+\ref{sec:background}, we compare sampling and estimation procedures
+between the design-based approach and the model-based approach for
+spatial data. In Section \ref{sec:numstudy}, we use a simulation
+approach to study the behavior and performance of both approaches. In
+Section \ref{application}, we use both approaches to analyze real data.
+And in Section \ref{sec:discussion}, we end with a discussion and
+provide directions for future research.
 
@@
 
-\hypertarget{comparing-design-based-and-model-based-approaches}{%
+\hypertarget{subsec:dvm_compare}{%
 \subsection{Comparing Design-Based and Model-Based
-Approaches}\label{comparing-design-based-and-model-based-approaches}}
+Approaches}\label{subsec:dvm_compare}}
 
@@
 other units, so we call them ``Independent Random Sampling'' (IRS)
-designs. The design-based approach combines the randomness of the
-sampling design and the data collected via the sample to estimate
-parameters (e.g.~means and totals) of a population. Generally, these
-population parameters are assumed to be fixed, unknown constants.
+designs. Sampling designs incorporating the spatial locations of units
+in the sample frame are called spatially balanced designs. Spatially
+balanced designs can be obtained using the Generalized Random
+Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004),
+which we discuss in more detail in Section \ref{subsec:spb_design}. The
+design-based approach combines the randomness of the sampling design and
+the data collected via the sample to estimate parameters (e.g.~means and
+totals) of a population. Generally, these population parameters are
+assumed to be fixed, unknown constants.
 
@@
 The model-based approach assumes the data are a random realization of a
-data-generating process. Randomness is often incorporated through
-distributional assumptions on this process and need not be incorporated
-through random sampling (bottom row of Figure \ref{fig:fig1}). Instead
-of estimating fixed but unknown parameters (as in the design-based
-approach), the goal of model-based inference in the spatial context is
-often to predict a realized variable, or value. For example, suppose the
-realized mean of all population units is the value of interest. Instead
-of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the
-value of the mean, a random variable. We know that if we sampled all
-population units (and not any units outside the population), we would be
-able to exactly predict (without any uncertainty) the mean of our one
-realized process. Assuming the data are realized from a specific
-data-generating process yields predictions that are linked to
-distributional assumptions. These distributional assumptions are used to
-derive prediction intervals. If we repeatedly generate the response
-values from a fixed spatial process and obtained a sample, then 95\% of
-all 95\% prediction intervals constructed from a procedure with
-appropriate coverage will contain their respective realized means.
-Cressie (1993) and Schabenberger and Gotway (2017) provide reviews of
-model-based approaches for spatial data. A visual comparison of the
-design-based and model-based assumptinos is provided in Figure
-\ref{fig:fig1} (Brus (2020) provides a similar figure).
+data-generating process. Randomness is incorporated through
+distributional assumptions on this process. Strictly speaking,
+randomness need not be incorporated through random sampling, though
+Diggle et al. (2010) warn against preferential sampling Preferential
+sampling occurs when the process generating the data locations and the
+process being modeled are not independent of one another. To guard
+against preferential sampling, model-based approaches often still
+implement random sampling.
+
+Instead of estimating fixed but unknown parameters (as in the
+design-based approach), the goal of model-based inference in the spatial
+context is often to predict a realized variable, or value. For example,
+suppose the realized mean of all population units is the value of
+interest. Instead of \emph{estimating} a fixed, unknown mean, we are
+\emph{predicting} the value of the mean, a random variable. Prediction
+intervals are then derived leveraging assumptions of the data generating
+process. If we repeatedly generate the response values from a fixed
+spatial process and obtained a sample, then 95\% of all 95\% prediction
+intervals constructed from a procedure with appropriate coverage will
+contain their respective realized means. Cressie (1993) and
+Schabenberger and Gotway (2017) provide reviews of model-based
+approaches for spatial data. A visual comparison of the design-based and
+model-based assumptions is provided in Figure \ref{fig:fig1} (Brus
+(2020) provides a similar figure).
 
@@
 
-\textbf{Tony O.}: Before this section is it useful to have a section
-that lays out the general site selection and general analysis options.
-Thinking about site selection as design-based IRS, design-based GRTS,
-Arbitrary set of sites, selection for model-based. Then general analysis
-options as design-based no spatial, design-based spatial, model-based.
-This four by three table would show that model-based analyses are
-possible for all selection options. Design-based options with no spatial
-info possible for IRS-based and GRTS-based. Design-based options with
-spatial info possible for GRTS-based.
+\hypertarget{subsec:spb_design}{%
+\subsection{Spatially Balanced Design and
+Analysis}\label{subsec:spb_design}}
 
-\textbf{Jay VH}: What about the design for model-based inference?
-Strictly speaking, it is fixed -- there is no probabilistic use of a
-randomized design. However, we are going to have to deal with Diggle et
-al.~(2010).
+Spatially balanced samples can be obtained using the design-based
+approach. Spatially balanced samples are useful because parameter
+estimates from these samples tend to vary less than parameter estimates
+from samples that are not spatially balanced (Barabesi and Franceschi,
+2011; Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson
+et al., 2013; Stevens and Olsen, 2004; Wang et al., 2013). The first
+spatially balanced sampling algorithm that saw widespread use was the
+Generalized Random Tessellation Stratified (GRTS) algorithm (Stevens and
+Olsen, 2004). To quantify the spatial balance of a sample, Stevens and
+Olsen (2004) proposed loss metrics based on Voroni polygons. Since GRTS
+was developed, several other spatially balanced sampling algorithms have
+emerged, including the Local Pivotal Method (Grafström et al., 2012;
+Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
+(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
+2013), Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017),
+and Halton Iterative Partitioning Sampling (Robertson et al., 2018). In
+this manuscript, we use Generalized Random Tessellation Stratified
+(GRTS) sampling because it has several attractive properties: GRTS
+sampling accommodates finite and infinite sampling frames; accommodates
+equal, unequal, and proportional (to) size inclusion probabilities;
+accommodates legacy (historical) sampling; accommodates a minimum
+distance between units in a sample; accommodates reverse hierarchically
+ordered replacement units in a sample (replacement units are units
+available to be sampled if an original unit cannot be sampled); and is
+available in the spsurvey R package Dumelle et al. (2021).
 
-\hypertarget{spatially-balanced-design-and-analysis}{%
-\subsection{Spatially Balanced Design and
-Analysis}\label{spatially-balanced-design-and-analysis}}
+The GRTS algorithm samples from finite and infinite populations by
+utilizing a mapping between two-dimensional and one-dimensional space.
+The units in the two-dimensional sampling frame are divided into cells
+using a hierarchical address. This hierarchical address is then used to
+map the units from two-dimensional space to a one-dimensional line where
+each unit's line length equals its inclusion probability. A systematic
+sample is conducted on the line and linked back to a unit in
+two-dimensional space, which results in the desired sample. Stevens and
+Olsen (2004) and Dumelle et al. (2021) provide further details.
 
-\textbf{Lisa M.}: Need a more precise definition of ``miniature'' in
-this context, and need an example.
+After selecting a GRTS sample, data are collected and used to estimate
+population parameters. To unbiasedly estimate population means and
+totals from sample data, one can use the Horvitz-Thompson estimator
+(Horvitz and Thompson, 1952). If \(\tau\) is a population total, the
+Horvitz-Thompson estimate of \(\tau\), denoted by \(\hat{\tau}_{ht}\),
+is is given by \begin{align}\label{eq:ht}
+  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
+\end{align} where \(Z_i\) is the value of the \(i\)th unit in the sample
+and \(\pi_i\) is the inclusion probability of the \(i\)th unit in the
+sample. An estimate of the population mean is obtained by dividing
+\(\hat{\tau}_{ht}\) by the population size.
 
-\textbf{Jay VH}: Saying ``the distribution of the sampled population
-units mirrors the density of\ldots{}'' is confusing to me. Are these
-formal statistical definitions of distribution (cumulative distribution
-function) and density (probability density function)? Wouldn't IRS
-sample be a miniature, as it should, on average, mirror a population?
-
-The design-based approach can use spatial locations to obtain spatially
-balanced samples. First we discuss spatial balance with respect to the
-population (Stevens and Olsen, 2004). A sample is spatially balanced
-with respect to the population if the sampled population units are a
-miniature of the population units. A sample is a miniature of the
-population if the distribution of the sampled population units mirrors
-the density of all population units. Spatial balance with respect to the
-population is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread
-out in some type of equidistant manner over geographical space and is
-not meant to be miniatures of the population. When we refer to spatial
-balance henceforth, we mean spatial balance with respect to the
-population.
-
-Spatially balanced samples are useful because they tend to yield
-estimates that have lower variance than estimates constructed from
-sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
-Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
-2013; Stevens and Olsen, 2004; Wang et al., 2013). To quantify spatial
-balance, Stevens and Olsen (2004) proposed loss functions based on
-Voroni polygons. The first spatially balanced sampling algorithm that
-saw widespread use was the Generalized Random Tessellation Stratified
-(Stevens and Olsen, 2004). Since GRTS was developed, several other
-spatially balanced sampling algorithms have emerged, including the Local
-Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
-Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
-Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
-(Benedetti and Piersimoni, 2017), and Halton Iterative Partitioning
-(Robertson et al., 2018). We focus on the Generalized Random
-Tessellation Stratified (GRTS) algorithm to select spatially balanced
-sampling because it has several attractive properties, including
-\textbf{Lisa M.}: List major attractive properties, and detailed by
-Stevens and Olsen (2004) and Dumelle et al. (2021).
-
-The GRTS algorithm is used to sample from finite and infinite
-populations and works by utilizing a mapping between two-dimensional and
-one-dimensional space. The population units in two-dimensional space are
-divided into cells using a hierarchical index. Population units are then
-mapped to a one-dimensional line via the hierarchical indexing. The line
-length of each population unit equals its inclusion probability. A
-systematic sample is conducted on the line and these samples are linked
-to a population unit in two-dimensional space, which results in the
-desired sample. Stevens and Olsen (2004) and Dumelle et al. (2021)
-provide further details.
-
-After collecting a sample using the GRTS algorithm, the data are used to
-estimate population parameters. The Horvitz-Thompson estimator (Horvitz
-and Thompson, 1952) yields unbiased estimates of population means and
-totals. For example, if \(\tau\) is a population total, then the
-Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
-is given by \begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align} where \(Z_i\) and \(\pi_i\) are the observed value and
-inclusion probability of the \(i\)th population unit selected in the
-sample. A similar formula exists for estimating the mean, \(\mu\).
-Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
-for \(\hat{\tau}_{ht}\), but they have two drawbacks. First, they rely
-on calculating \(\pi_{ij}\), the probability that population unit \(i\)
-and population unit \(j\) are included in the sample, and this can be
-very difficult to calculate. Second, they ignore the spatial locations
-of the population units. To address these drawbacks, Stevens and Olsen
-(2003) proposed a local neighborhood variance estimator. The local
-neighborhood variance estimator does not rely on \(\pi_{ij}\), and it
-incorporates spatial locations by assigning higher weights to nearby
-observations. Stevens and Olsen (2003) show this variance estimators
-tends to reduce the estimated standard error of \(\hat{\tau}\), yielding
-narrower confidence confidence intervals for \(\tau\).
+While the Horvitz-Thompson estimator is unbiased for population means
+and totals, it is also important to quantify the uncertainty in these
+estimates. Horvitz and Thompson (1952) and Sen (1953) provide variance
+estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
+these estimators rely on calculating \(\pi_{ij}\), the probability that
+unit \(i\) and unit \(j\) are both in the sample -- this quantity can be
+challenging if not impossible to calculate analytically. Second, these
+estimators ignore the spatial locations of the units in the sampling
+frame. To address these two drawbacks simultaneously, Stevens and Olsen
+(2003) proposed the local neighborhood variance estimator. The local
+neighborhood variance estimator does not rely on \(\pi_{ij}\) and
+incorporates spatial locations -- for technical details see Stevens and
+Olsen (2003). Stevens and Olsen (2003) show the local neigborhood
+variance estimator tends reduce \(\text{Var}(\hat{\tau})\) compared to
+variance estimators ignoring spatial locations, yielding narrower
+confidence intervals for \(\tau\).
 
@@
 basic principles are summarized below. Let
-\(\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}\)
+\({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
 be a response vector at locations \(s_1\), \(s_2\), . . . , \(s_N\) that
@@
 
-However, we often only have a sample of the \(N\) population units.
-Denoting quantities that are part of the sampled population units with a
+We often only have a sample of the \(N\) population units. Denoting
+quantities that are part of the sampled population units with a
 subscript \emph{s} and quantities that are part of the unsampled
@@
 
-We also assume that there is spatial correlation in \(\bm{\delta}\),
-which can be modeled using a covariance function. It is common to assume
-the covariance function is second-order stationary and isotropic
-(Cressie, 1993), and that the spatial covariance decreases as the
-separation between population units increases. Many spatial covariance
-functions exist, but the primary function we use throughout the
-simulations and applications in this manuscript is the exponential
-covariance function: the \(i,j^{th}\) entry for
+In addition to assuming the expectation of \(\bm{\delta}\) equals
+\(\mathbf{0}\), we also assume that there is spatial correlation in
+\(\bm{\delta}\) that can be modeled using a covariance function. It is
+common to assume the covariance function is second-order stationary and
+isotropic (Cressie, 1993), and that the spatial covariance decreases as
+the separation between population units increases. Many spatial
+covariance functions exist, but the primary function we use throughout
+the simulations and applications in this manuscript is the exponential
+covariance function: the \(i,j\)th entry for
 \(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
 \begin{align}\label{equation:expcov}
-\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
-\end{align} where \(h_{i,j}\) is the distance between population units
-\(i\) and \(j\), and \(\bm{\theta}\) is a vector of spatial covariance
-parameters of the partial sill \(\theta_1\), the range \(\theta_2\), and
-the nugget \(\theta_3\); and, \(\mathbbm{1}\) is equal to 1 when
-distance \(h_i,j\) is equal to 0, and equal to 0 otherwise. However, any
-spatial covariance function could be used in the place of the
-exponential, including functions that allow for non-stationarity or
-anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
-
-\textbf{Lisa M.} : Include formulas. Perhaps, but, these are very heavy
-in notation and matrix algebra. We might consider, however, adding the
-formulas to an Appendix.
+\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = 
+\begin{cases} 
+\sigma^2_{ps}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
+\sigma^2_{ps} + \sigma^2_n & h_{i,j} = 0
+\end{cases}
+,
+\end{align} where \(\sigma^2_{ps}\) is the partial sill measuring
+coarse-scale (correlated) variability, \(\sigma^2_{n}\) is the nugget
+measuring fine-scale (indepenent) variability, \(\phi\) is the range
+parameter measuring the distance-decay rate of the covariance, and
+\(h_{i,j}\) is the Euclidean distance between population units \(i\) and
+\(j\). Any spatial covariance function could be used in the place of the
+exponential, however, including functions that allow for
+non-stationarity or anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
 
@@
 be computed. While details of the derivation are in (Ver Hoef, 2008), we
-note here that the predictor and its variance are both moment-based.
+note here that the predictor and its variance are both moment-based,
+meaning they don't rely on any distributional assumptions.
 
@@
 2020), among others, can also be used to obtain predictions for a mean
-or total from spatially correlated responses in a finite population
-setting. We choose to use FPBK because it is faster than a Bayesian
-approach and random forest and because Ver Hoef and Temesgen (2013)
-showed that the method outperforms k-nearest-neighbors in many
-scenarios.
+or total from spatially correlated responses of a finite population. We
+choose to use FPBK because it is faster than a Bayesian approach and
+random forest and because Ver Hoef and Temesgen (2013) showed that the
+method outperforms k-nearest-neighbors in many scenarios.
 
@@
 
-\begin{longtable}[]{@{}lll@{}}
-\caption{\label{tab:designanalysis} Types of Sampling Design and
-Analysis combinations considered in the simulation study. The columns
-give the two types of sampling designs while the rows give the two types
-of analyses.}\tabularnewline
-\toprule
-& IRS & GRTS\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-& IRS & GRTS\tabularnewline
-\midrule
-\endhead
-Design & IRS-Design & GRTS-Design\tabularnewline
-Model & IRS-Model & GRTS-Model\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{table}[ht]
+\centering
+\begin{tabular}{r|ll}
+  \hline
+ & IRS & GRTS \\ 
+  \hline
+IRS & IRS-Design & GRTS-Design \\ 
+  GRTS & IRS-Model & GRTS-Model \\ 
+   \hline
+\end{tabular}
+\caption{\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
+The rows give the two types of sampling designs while the columns give the two types of analyses.}
+\end{table}
 
@@
 
-\begin{longtable}[]{@{}llll@{}}
-\caption{\label{tab:parmtab} Simulation parameters. Total variability
-for all scenarios was 2 so that the partial sill was 0, 1, or
-1.8.}\tabularnewline
-\toprule
-\endhead
-Sample Size (n) & 50 & 100 & 200\tabularnewline
-Site Locations & Random & Gridded &\tabularnewline
-Partial Sill / Total Variance & 0 & 0.5 & 0.9\tabularnewline
-Response Type & Normal & Lognormal &\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{table}[ht]
+\centering
+\begin{tabular}{r|lll}
+   \hline
+Sample Size (n) & 50 & 100 & 200 \\ 
+  Layout & Random & Gridded & - \\ 
+  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
+  Response Type & Normal & Lognormal & - \\ 
+   \hline
+\end{tabular}
+\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.} 
+\end{table}
 
@@
 
+\leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
+Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
+under preferential sampling. Journal of the Royal Statistical Society:
+Series C (Applied Statistics) 59, 191--232.
+
 \leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%

### 2021-10-12 inst/manuscript/notes_Madsen.Rmd
section 2 updates

@@
 __Lisa M__: Add a paragraph describing the contribution of the manuscript (before the "the rest of the manuscript is organize as follows etc." in the introduction)
+
+__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
+
+__Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
+
+__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.

### 2021-10-12 inst/manuscript/notes_VerHoef.Rmd
section 2 updates

@@
 We should stick with variables, or realized variables (we might also call these values, but we should define and establish a consistent terminology early on.) __Matt H__: I think, though this comment is for this paragraph, we should establish the terminology earlier.
+
+__Jay VH__: What about the design for model-based inference?
+Strictly speaking, it is fixed – there is no probabilistic use of a randomized design.
+However, we are going to have to deal with Diggle et al.
+(2010).
+
+__Jay VH__: Saying "the distribution of the sampled population units mirrors the density of..." is confusing to me.
+Are these formal statistical definitions of distribution (cumulative distribution function) and density (probability density function)?
+Wouldn’t IRS sample be a miniature, as it should, on average, mirror a population?

### 2021-10-13 inst/manuscript/manuscript.Rmd
section 3 and 4 edits

@@
 Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design.
-Means and totals, for example, are asymptotically Gaussian by the Central Limit Theorem (under some assumptions).
+Means and totals, for example, are asymptotically normally distributed (normal) by the Central Limit Theorem (under some assumptions).
 If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
@@
 We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded).
+All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
 The mean for the spatial process generating the response was set to zero.
@@
 The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
+FPBK can be readily performed in `R` with the `sptotal` package [@higham2021sptotal].
 We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
 
+**Mike** For design-based, it's really RMSE -- how should we address this?
 Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
@@
 \begin{equation*}
-\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
 \end{equation*}
 
-where rMSPE is the root-Mean-Squared-Prediction-Error.
+where rMSPE is the root-mean-squared-prediction error.
 When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
-So, using GRTS or using a spatial model does not result in much loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is not very big.
+So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated.
+When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small.
 In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
@@
 
-Plot Note: change colours and think about shape
-think about legend going on graph
-
 ```{r, fig.keep = "none"}
@@
 
-Figure \ref{fig:figconf} shows the coverage for each of the four approaches.
-We see that the four approaches have somewhat similar coverages in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
+We also studied 95% interval coverage among the approaches.
+The design-based 95% confidence intervals and model-based 95% prediction intervals are constructed using the normal distribution.
+Justification for the design-based intervals lies in the asymptotic normality of totals via the Central Limit Theorem, and justification for the model-based intervals lies in the normality assumption of the errors.
+Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches.
+All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
 
-In the normal response settings, where assumptions for the model-based approaches and the design-based approaches are satisfied, all approaches have coverage around the nominal 0.95.
-Because the intervals are symmetric, normal-based intervals, the coverages are generally lower than the nominal 95%.
-We see that, because the sampling distribution of the mean is asymptotically normal (reference for this?
-a spatial version of CLT), the coverages when the sample size is 200 are much closer to the nominal 95% than the coverages when the sample size is 50.
-In general, the more skewed the distribution of the response, the larger the sample size needed to ensure proper coverage of these normal-based intervals.
-
-Do we want to mention stuff like what's in these next bullet points or no?
-
-- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean.
-The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
-- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval.
-Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.
+In the normal response settings, all approaches have coverage around 95%.
+This is expected, as the intervals are also based on the normal distribution In the lognormal response settings, however, all approaches have coverage below 95%.
+This is also expected, as the intervals are still based on the normal distribution.
+In the lognormal response settings, interval coverage increases both as the sample size increases and as the strength of spatial dependence increases.
+This suggests that the larger the sample size and the stronger the spatial dependence, the more resistant these intervals are to departures from normality of the data.
 
@@
 We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
-Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
-Therefore, we will explore whether or not we can still obtain an adequately precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
+Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
 
@@
 library(SpatialDesignVModel)
-data(nla)
+data(nla_df) # may need to reinstall package
 nla_df
@@
 
-```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.
-Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.", message = FALSE, warning = FALSE}
+```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.", message = FALSE, warning = FALSE}
 merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
@@
         axis.ticks = element_blank()) +
-  coord_quickmap() 
+  coord_quickmap() +
+  theme_bw(base_size = 14)
 # dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
@@
 
-```{r, echo = FALSE, results = "hide"}
-mean(nla_nomiss_both$total_hg)
+```{r, echo = FALSE}
+mean_hg <- mean(nla_nomiss_both$total_hg)
 # mean(nla_nomiss_both$RDis_IX)
@@
 Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.2 ng / g.
+The realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
 
@@
 
-```{r apptab}
-library(knitr)
+```{r apptab, results = "asis"}
 res_df_drop <- res_df[ ,-2]
 names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-kable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 103.2 ng / g")
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
+The true mean concentration is 103.2 ng / g.")
+print(xtab_app, include.rownames = FALSE, comment = FALSE)
 ```
 
-Table \ref{tab:appliedtab} shows the application of a design-based analysis on an IRS, a model-based analysis on an IRS, a design-based analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
-We see that, for all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
+Table \ref{tab:appliedtab} shows the application of a design-based analysis of an IRS sample, a model-based analysis of an IRS sample, a design-based analysis of a GRTS sample, and a model-based analysis of a GRTS sample.
+For all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
 However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
 
-<!-- not sure why this is getting cutoff. temporary fix: -->
-
-\pagebreak
+<!-- not sure why this is getting cutoff. temporary fix: \pagebreak 
+probably because it is an html table -- I used xtable to fix -->
 
 But, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not use the spatial correlation in mercury concentration across the contiguous United States.
-We also see that, for the samples drawn, the both analyses with the GRTS sampling design have a lower standard error than the analyses with the IRS sampling design.
-We would expect this to be the case for most samples because mercury concentration exhibits spatial correlation so a spatially balanced sample should usually yield a lower standard error.
-If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling just 50 lakes to save time and money.
+The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration across the contiguous United States.
+We also see that both approaches using the GRTS sample have a lower standard error than the both approaches using the IRS sample.
+We would expect this to be the case for most samples because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.
 

### 2021-10-13 inst/manuscript/manuscript.tex
section 3 and 4 edits

@@
 \bibliographystyle{elsarticle-harv}
-\usepackage{longtable}
 \ifxetex
@@
 sampling units selected via the sampling design. Means and totals, for
-example, are asymptotically Gaussian by the Central Limit Theorem (under
-some assumptions). If we repeatedly sample the surface, then 95\% of all
-95\% confidence intervals constructed from a procedure with appropriate
-coverage will contain the true, fixed mean. Särndal et al. (2003) and
-Lohr (2009) provide thorough reviews of the design-based approach.
+example, are asymptotically normally distributed (normal) by the Central
+Limit Theorem (under some assumptions). If we repeatedly sample the
+surface, then 95\% of all 95\% confidence intervals constructed from a
+procedure with appropriate coverage will contain the true, fixed mean.
+Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
+design-based approach.
 
@@
 \(N = 900\) response values simulated on the unit square in either
-random locations (Site Locations = Random) or gridded locations (Site
-Locations = Gridded). The mean for the spatial process generating the
-response was set to zero.
+random locations (Layout = Random) or gridded locations (Layout =
+Gridded). The mean for the spatial process generating the response was
+set to zero.
 
@@
 et al., 2021). FPBK can be readily performed in \texttt{R} with the
-\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
+\texttt{sptotal} package (Higham et al., 2021). We use \texttt{sptotal}
 for both the simulation analysis and the application, estimating
@@
 
-Figure \ref{fig:figeff} shows the relative efficiency of the four
-approaches from Table \ref{tab:designanalysis} with ``IRS-Design'' as
-the baseline: \mbox{} \begin{equation*}
-\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+\textbf{Mike} For design-based, it's really RMSE -- how should we
+address this? Figure \ref{fig:figeff} shows the relative efficiency of
+the four approaches from Table \ref{tab:designanalysis} with
+``IRS-Design'' as the baseline: \mbox{} \begin{equation*}
+\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
 \end{equation*}
 
-where rMSPE is the root-Mean-Squared-Prediction-Error. When there is no
+where rMSPE is the root-mean-squared-prediction error. When there is no
 spatial correlation (top row), the four approaches have approximately
@@
 violated. So, using GRTS or using a spatial model does not result in
-much loss in efficiency even if the response variable is not spatially
-correlated. When there is high spatial correlation (bottom row), the
-GRTS-Model approach tends to perform best, but difference in relative
-efficiency between GRTS-Model and GRTS-Design is not very big. In the
+much, if any, loss in efficiency even if the response variable is not
+spatially correlated. When there is high spatial correlation (bottom
+row), the GRTS-Model approach tends to perform best, but difference in
+relative efficiency between GRTS-Model and GRTS-Design is small. In the
 lognormal, high partial sill settings (bottom-right facet), GRTS-Design
@@
 
-Plot Note: change colours and think about shape think about legend going
-on graph
-
-Figure \ref{fig:figconf} shows the coverage for each of the four
-approaches. We see that the four approaches have somewhat similar
-coverages in all settings, with GRTS-Design having slightly lower
+We also studied 95\% interval coverage among the approaches. The
+design-based 95\% confidence intervals and model-based 95\% prediction
+intervals are constructed using the normal distribution. Justification
+for the design-based intervals lies in the asymptotic normality of
+totals via the Central Limit Theorem, and justification for the
+model-based intervals lies in the normality assumption of the errors.
+Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
+the four approaches. All four approaches have somewhat similar interval
+coverage in all settings, with GRTS-Design having slightly lower
 coverage when the response is normal.
 
-In the normal response settings, where assumptions for the model-based
-approaches and the design-based approaches are satisfied, all approaches
-have coverage around the nominal 0.95. Because the intervals are
-symmetric, normal-based intervals, the coverages are generally lower
-than the nominal 95\%. We see that, because the sampling distribution of
-the mean is asymptotically normal (reference for this? a spatial version
-of CLT), the coverages when the sample size is 200 are much closer to
-the nominal 95\% than the coverages when the sample size is 50. In
-general, the more skewed the distribution of the response, the larger
-the sample size needed to ensure proper coverage of these normal-based
-intervals.
-
-Do we want to mention stuff like what's in these next bullet points or
-no?
-
-\begin{itemize}
-\tightlist
-\item
-  for the model-based approach, the more skewed the population is, the
-  higher the sample size needed to satisfy CLT for predicting a mean.
-  The derivation of the BLUP is entirely moment-based (no distribution
-  assumed) but we still need to assume a distribution to estimate
-  spatial parameters and to generate bounds of a prediction interval.
-\item
-  many confidence intervals generated for design-based approaches also
-  rely on the CLT and the normal distribution to generate the interval.
-  Again, for highly skewed data with a small sample size, this
-  assumption is violated even though all of the assumptions for
-  generating the estimator are valid.
-\end{itemize}
+In the normal response settings, all approaches have coverage around
+95\%. This is expected, as the intervals are also based on the normal
+distribution In the lognormal response settings, however, all approaches
+have coverage below 95\%. This is also expected, as the intervals are
+still based on the normal distribution. In the lognormal response
+settings, interval coverage increases both as the sample size increases
+and as the strength of spatial dependence increases. This suggests that
+the larger the sample size and the stronger the spatial dependence, the
+more resistant these intervals are to departures from normality of the
+data.
 
@@
 United States, to obtain an interval for mean mercury concentration.
-Although all lakes in the survey were measured in 2012, there may not
-always be enough time or money to do so. Therefore, we will explore
-whether or not we can still obtain an adequately precise estimate for
-the realized mean mercury concentration if we only take a sample of 100
-of the 986 lakes.
+Although we know the true mean mercury concentration values for the 986
+lakes from the 2012 NLA, we will explore whether or not we obtain an
+adequately precise estimate for the realized mean mercury concentration
+if we sample only 100 of the 986 lakes.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.
-Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.}\label{fig:figdata}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.}\label{fig:figdata}
 \end{figure}
@@
 concentrations in lakes in the northeast and north central United
-States. Because we are considering these lakes to be our entire
-population, we know that the realized mean mercury concentration is
+States. The realized mean mercury concentration in the 986 lakes is
 103.2 ng / g.
 
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tab:appliedtab} Application of design-based and
-model-based approaches to the NLA data set on mercury concentration. The
-true mean concentration is 103.2 103.2 ng / g}\tabularnewline
-\toprule
-Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
-\midrule
-\endhead
-IRS-Design & 112.7 & 8.8 & 95.4 & 129.9\tabularnewline
-IRS-Model & 110.5 & 7.9 & 95.0 & 125.9\tabularnewline
-GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7\tabularnewline
-GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9\tabularnewline
-\bottomrule
-\end{longtable}
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Approach & Estimate & SE & 95\% LB & 95\% UB \\ 
+  \hline
+IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\ 
+  IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\ 
+  GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\ 
+  GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\ 
+   \hline
+\end{tabular}
+\caption{\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.} 
+\end{table}
 
 Table \ref{tab:appliedtab} shows the application of a design-based
-analysis on an IRS, a model-based analysis on an IRS, a design-based
-analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
-We see that, for all four analyses, the true realized mean mercury
+analysis of an IRS sample, a model-based analysis of an IRS sample, a
+design-based analysis of a GRTS sample, and a model-based analysis of a
+GRTS sample. For all four analyses, the true realized mean mercury
 concentration is within the bounds of the 95\% intervals. However, we
@@
 
-\pagebreak
-
 But, we do note a couple of patterns. The design-based IRS analysis
 shows the largest standard error: a likely reason is that this is the
-only approach that does not use the spatial correlation in mercury
-concentration across the contiguous United States. We also see that, for
-the samples drawn, the both analyses with the GRTS sampling design have
-a lower standard error than the analyses with the IRS sampling design.
-We would expect this to be the case for most samples because mercury
-concentration exhibits spatial correlation so a spatially balanced
-sample should usually yield a lower standard error. If it is acceptable
-to have an interval for mean mercury concentration of about 25 ng / g
-and if we ignore the other variables that the EPA collects information
-on in these NLA surveys, then the EPA could consider sampling just 50
-lakes to save time and money.
+only approach that does not incorporate any spatial information
+regarding mercury concentration across the contiguous United States. We
+also see that both approaches using the GRTS sample have a lower
+standard error than the both approaches using the IRS sample. We would
+expect this to be the case for most samples because mercury
+concentration exhibits spatial patterning, so a spatially balanced
+sample should usually yield a lower standard error.
 
@@
 
-\leavevmode\hypertarget{ref-higham2020sptotal}{}%
-Higham, M., Ver Hoef, J., Bryce, F., 2020. Sptotal: Predicting totals
-and weighted sums from spatial data.
+\leavevmode\hypertarget{ref-higham2021sptotal}{}%
+Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
+Predicting totals and weighted sums from spatial data.
 

### 2021-10-13 inst/manuscript/notes_Higham.Rmd
section 3 and 4 edits

--- 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
@@
----
-title: A comparison of design-based and model-based approaches for spatial data.
-author:
-  - name: In alphabetical order Michael Dumelle
-    email: Dumelle.Michael@epa.gov
-    affiliation: USEPA
-    footnote: 1
-  - name: Matt Higham
-    email: mhigham@stlaw.edu
-    affiliation: STLAW
-    footnote: 1
-  - name: Lisa Madsen
-    affiliation: OSU
-  - name: Anthony R. Olsen
-    affiliation: USEPA
-  - name: Jay M. Ver Hoef
-    affiliation: NOAA
-address:
-  - code: USEPA
-    address: United States Environmental Protection Agency, 200 SW 35th St, Corvallis, Oregon, 97333
-  - code: STLAW
-    address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
-  - code: OSU
-    address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
-  - code: NOAA
-    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-abstract: |
-  This is the abstract.
+Notes for Matthew Higham
 
-journal: "An awesome journal"
-date: "`r Sys.Date()`"
-bibliography: mybibfile.bib
-linenumbers: true
-numbersections: true
-csl: elsevier-harvard.csl
-preamble: >
-  \usepackage{bm}
-  \usepackage{bbm}
-  \usepackage{color}
-  \DeclareMathOperator{\var}{{var}}
-  \DeclareMathOperator{\cov}{{cov}}
-output: rticles::elsevier_article
----
-
-_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
-
-Potential Journals:
-
-* Ecological Applications
-* Methods in Ecology and Evolution
-* Journal of Applied Ecology
-* Environmetrics
-* Environmental and Ecological Statistics
-
-<!-- Notes from today -->
-<!-- •	Finite populations require a smaller scope but we need to discuss the impact of infinite populations as well – put finite vs infinite population context in the introduction / similarities and differences -->
-<!-- •	What should a lower sample size be? 50 seems reasonable 30 may be too small – maybe 50, 100, 200 -->
-<!-- •	IRS results seem off for dependent error sample size 50 -->
-<!-- •	Take home points – don’t have to do GRTS design with GRTS sample -->
-<!-- •	Take home points – GRTS is better for both model and design, while spatial approach for design is close to spatial model -->
-<!-- •	Add medium dependence -->
-<!-- •	Remind readers what choices of the four approaches in the application mean – what do they mean to practitioners -->
-
-# Introduction {#sec:introduction}
-
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>", 
-  warning = FALSE, 
-  message = FALSE,
-  include = TRUE, 
-  echo = FALSE
-)
-library(tidyverse)
-library(here)
-library(xtable) # latex tables
-```
-
-<!-- Brief introduction to model based and design based -->
-
-There are two general approaches for using data to make statistical inferences about a population: design-based approaches and model-based approaches.
-When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected in a  sample.
-In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process that generated the data.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
-
-<!-- Shift focus to spatial data -->
-
-__Tony O.__: Should this paragraph address that spatial information can be incorporated in the design stage or in the analysis stage (or both).
-In general, it's not clear whether we are referring to site selection process or the estimation process
-
-Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process.
-@de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered, but they tend to compare design-based approaches that ignore spatial locations to model-based approaches [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review, but no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
-
-<!-- Outline for the rest of the paper -->
-
-__Lisa M.__: Add paragraph describing contribution of manuscript.
-
-The rest of this paper is organized as follows.
-In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach.
-In Section \ref{sec:numstudy}, we use simulated and real data to study the the behavior of both approaches.
-And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
-
-# Background {#sec:background}
-
-The design-based and model-based approaches incorporate randomness in fundamentally different ways.
-In this section, we describe the role of randomness and its effects on subsequent inferences.
-We then discuss specific inference methods for the design-based and model-based approaches for spatial data.
-
-## Comparing Design-Based vs. Model-Based
-
-The design-based approach assumes the population is fixed.
-Randomness is incorporated in the selection of population units according to a sampling design.
-A sampling design assigns a positive probability of inclusion in the sample (inclusion probability) to each population unit.
-Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster
-sampling, which we refer to as Independent Random Sampling (IRS) survey designs.
-The goal is to use the sampling design and the sampled data to estimate population parameters like means and totals.
-These population parameters are traditionally assumed to be fixed but unknown.
-
-Treating the data as fixed and incorporating randomness through the sampling design (top row of Figure \ref{fig:fig1} ((cite Brus 2021 here since our figure is similar?))) yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived using limiting arguments.
-Means and totals, for example, are asymptotically normally distributed by the Central Limit Theorem.
-If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
-@sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
-
-__Jay VH__: I think it is important to stress that the limiting distribution is over all possible randomizations, constrained by whatever design is used.
-
-__Jay VH__: quantity is vague.
-We should stick with variables, or realized variables (we might also call these values, but we should define and establish a consistent terminology early on.) __Matt H__: I think, though this comment is for this paragraph, we should establish the terminology earlier.
-
-The model-based approach assumes the data are a random realization of a data-generating process.
-Randomness is often incorporated through distributional assumptions on this process and need not be incorporated through random sampling (bottom row of Figure \ref{fig:fig1}).
-Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity.
-For example, suppose the realized mean of all population units is the quantity of interest.
-Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty.
-
-Assuming the data is a realization of a specific data-generating process yields predictors that are linked to distributional assumptions.
-These distributional assumptions are used to derive prediction intervals.
-The distributional assumptions allow the prediction intervals to be more precise.
-If we repeatedly generate the response values from a fixed spatial process and obtain a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
-@cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-
-```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
-Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of n = 4.
-The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
-The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels."}
-set.seed(10102021)
-
-source(here("R", "sim_pop.R"))
-N <- 9
-pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                  psill = 0.9, erange = 1, nugget = 0.1)
-
-n <- 4
-samp_df <- pop_df %>% sample_n(n) 
-samp_df2 <- pop_df %>% sample_n(n) 
-samp_df3 <- pop_df %>% sample_n(n)
-
-library(cowplot)
-d1 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df, size = 5.5, shape = 1, 
-             show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d2 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df2, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d3 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df3, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                       psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-
-indeces <- sample(1:N, size = n, replace = FALSE)
-samp_df_mod <- pop_df_mod %>% slice(indeces)
-samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
-samp_df_mod3 <- pop_df_mod3 %>% slice(indeces)
-
-m1 <- ggplot(data = pop_df_mod, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m2 <- ggplot(data = pop_df_mod2, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m3 <- ggplot(data = pop_df_mod3, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-library(ggpubr)
-ggarrange(d1, d2, d3, m1, m2, m3, nrow = 2, ncol = 3,
-          common.legend = TRUE, legend = "right")
-```
-
-__Tony O.__: Before this section is it useful to have a section that lays out the general site selection and general analysis options.
-Thinking about site selection as design-based IRS, design-based GRTS, Arbitrary set of sites, selection for model-based.
-Then general analysis options as design-based no spatial, design-based spatial, model-based.
-This four by three table would show that model-based analyses are possible for all selection options.
-Design-based options with no spatial info possible for IRS-based and GRTS-based.
-Design-based options with spatial info possible for GRTS-based.
-
-__Jay VH__: What about the design for model-based inference?
-Strictly speaking, it is fixed – there is no probabilistic use of a randomized design.
-However, we are going to have to deal with Diggle et al.
-(2010).
-
-## Spatially Balanced Design and Analysis
-
-__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.
-
-__Jay VH__: Saying "the distribution of the sampled population units mirrors the density of..." is confusing to me.
-Are these formal statistical definitions of distribution (cumulative distribution function) and density (probability density function)?
-Wouldn’t IRS sample be a miniature, as it should, on average, mirror a population?
-
-The design-based approach can use spatial locations to obtain spatially balanced samples.
-First we discuss spatial balance with respect to the population [@stevens2004spatially].
-A sample is spatially balanced with respect to the population if the sampled population units are a miniature of the population units.
-A sample is a miniature of the population if the distribution of the sampled population units mirrors the density of all population units.
-Spatial balance with respect to the population is different than spatial balance with respect to geography.
-A sample that is spatially balanced with respect to geography is spread out in some type of equidistant manner over geographical space and is not meant to be miniatures of the population.
-When we refer to spatial balance henceforth, we mean spatial balance with respect to the population.
-
-Spatially balanced samples are useful because they tend to yield estimates that have lower variance than estimates constructed from sampling designs lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-To quantify spatial balance, @stevens2004spatially proposed loss functions based on Voroni polygons.
-The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified [@stevens2004spatially].
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance [@benedetti2017spatially], and Halton Iterative Partitioning [@robertson2018halton].
-We focus on the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced sampling because it has several attractive properties, including __Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.
-
-<!-- algorithms. -->
-
-The GRTS algorithm is used to sample from finite and infinite populations and works by utilizing a mapping between two-dimensional and one-dimensional space.
-The population units in two-dimensional space are divided into cells using a hierarchical index.
-Population units are then mapped to a one-dimensional line via the hierarchical indexing.
-The line length of each population unit equals its inclusion probability.
-A systematic sample is conducted on the line and these samples are linked to a population unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially and @dumelle2021spsurvey provide further details.
-
-<!-- This makes it seem like HT is being used to estimate population parameters in GRTS?
-Needs to be in there, but maybe a sentence making it clear that this isn't being used?
-Or just change the order and talk about Local variance first and then mention these for more general settings?
--->
-
-After collecting a sample using the GRTS algorithm, the data are used to estimate population parameters.
-The Horvitz-Thompson estimator [@horvitz1952generalization] yields unbiased estimates of population means and totals.
-For example, if $\tau$ is a population total, then the Horvitz-Thompson estimator of $\tau$ (denoted by $\hat{\tau}_{ht}$), is given by
-\begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align}
-where $Z_i$ and $\pi_i$ are the observed value and inclusion probability of the $i$th population unit selected in the sample.
-A similar formula exists for estimating the mean, $\mu$.
-@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are included in the sample, and this can be very difficult to calculate.
-Second, they ignore the spatial locations of the population units.
-To address these drawbacks, @stevens2003variance proposed a local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$, and it incorporates spatial locations by assigning higher weights to nearby observations.
-@stevens2003variance show this variance estimators tends to reduce the estimated standard error of $\hat{\tau}$, yielding narrower confidence confidence intervals for $\tau$.
-
-## Finite Population Block Kriging
-
-Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
-Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process.
-@verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-Let $\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
-.
-.
-, \text{z}(s_N) \}$ be a response vector at locations $s_1$, $s_2$, .
-.
-.
-, $s_N$ that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
-Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
-For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
-
-<!-- how explicitly do we want to define \mathbf{z} -->
-
-<!-- finite number of -->
-
-<!-- confusing having two tau's: tau for total and a tau as a function -->
-
-However, we often only have a sample of the $N$ population units.
-Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
-
-\begin{equation}
-\begin{pmatrix} \label{equation:Zmarginal}
-    \mathbf{z}_s      \\
-    \mathbf{z}_u
-\end{pmatrix}
-=
-\begin{pmatrix}
-  \mathbf{X}_s    \\
-  \mathbf{X}_u
-\end{pmatrix}
-\bm{\beta} +
-\begin{pmatrix}
-\bm{\delta}_s    \\
-\bm{\delta}_u
-\end{pmatrix},
-\end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively; $\beta$ is the parameter vector of fixed effects; and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
-Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume the expectation of $\bm{\delta}$ equals $\mathbf{0}$.
-
-We also assume that there is spatial correlation in $\bm{\delta}$, which can be modeled using a covariance function.
-It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j^{th}$ entry for $\cov(\bm{\delta})$ is
-\mbox{}
-\begin{align}\label{equation:expcov}
-\cov(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
-\end{align}
-where $h_{i,j}$ is the distance between population units $i$ and $j$, and $\bm{\theta}$ is a vector of spatial covariance parameters of the partial sill $\theta_1$, the range $\theta_2$, and the nugget $\theta_3$; and, $\mathbbm{1}$ is equal to 1 when distance $h_i,j$ is equal to 0, and equal to 0 otherwise.
-However, any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
-80-93].
-
-__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.
-
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based.
-
-We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses in a finite population setting.
-We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
-
-# Numerical Study {#sec:numstudy}
-
-We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.
-
-```{r, echo = FALSE}
-tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-       nrow = 2) %>%
-  as.table()
-row.names(tab) <- c("Design", "Model")
-library(knitr)
-kable(tab, col.names = c("IRS", "GRTS"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The columns give the two types of sampling designs while the rows give the two types of analyses.")
-## I really want "Analysis" to appear as a row header and "Sampling Design" 
-## to appear as a column header, but this isn't easy to figure out.
-## I couldn't figure this out with xtable either so it might need to be
-## made "by hand."
-```
-
-We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded).
-The mean for the spatial process generating the response was set to zero.
-
-For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
-A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
-Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
-We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.
-
-```{r}
-n <- c("50", "100", "200")
-site_locations <- c("Random", "Gridded", "")
-psill_ratio <- c("0", "0.5", "0.9")
-resp_type <- c("Normal", "Lognormal", "")
-parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
-                         "Partial Sill / Total Variance", "Response Type")
-kable(parm_tab, col.names = c("", "", ""),
-      caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-```
-
-```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
-mu <- 0; var <- 2
-xvar <- seq(0.001, 10, by = 0.01)
-yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
-df <- tibble(xvar = xvar, yvar = yvar)
-ggplot(df, aes(x = xvar, y = yvar)) +
-  geom_line()
-```
-
-There were 2000 simulation trials for each of the 36 parameter combinations.
-In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected.
-For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model).
-For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).
-
-```{r, results = "asis", eval = FALSE}
-sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
-colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
-sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
-print(sim_one_table, include.rownames = FALSE, comment = FALSE)
-```
-
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal].
-We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
-
-Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
-\mbox{}
-\begin{equation*}
-\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
-\end{equation*}
-
-where rMSPE is the root-Mean-Squared-Prediction-Error.
-When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
-So, using GRTS or using a spatial model does not result in much loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is not very big.
-In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
-
-Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design.
-However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
-
-```{r figeff, out.width = "100%", fig.cap = "Relative Efficiency of the four design-analysis approaches.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
-library(tidyverse)
-library(here)
-
-files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
-
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x))
-
-combo_data <- combo_data %>%
-  dplyr::filter(n != 10 & gridded == FALSE) %>%
-  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
-  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
-                                            "Design GRTS", "Model GRTS"))) %>%
-  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal")))
-combo_data <- combo_data %>% group_by(sim) %>% mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
-  fill(designirsrmspe, .direction = "downup") %>%
-  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  ungroup() %>%
-  mutate(psill_ratio = 1 - nugget_ratio) %>%
-  mutate(n_factor = factor(n))
-
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
-                              colour = approach)) +
-  facet_grid(psill_ratio ~ resptype) +
-  geom_jitter(width = 0.19, size = 2.5) +
-  scale_colour_manual(values = colour_scale) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Relative Efficiency")
-effplot
-```
-
-Plot Note: change colours and think about shape
-think about legend going on graph
-
-```{r, fig.keep = "none"}
-combo_data_grts <- combo_data %>%
-  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
-  mutate(designgrts = if_else(approach == "Design GRTS", 
-                              true = rmspe,
-                              false = NA_real_)) %>%
-  group_by(sim) %>%
-  fill(designgrts, .direction = "updown") %>%
-  ungroup() %>%
-  mutate(releffgrts = rmspe / designgrts)
-combo_data_grts_model <- combo_data_grts %>%
-  filter(approach == "Model GRTS")
-
-ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
-  facet_grid(psill ~ resptype) +
-  geom_point()
-```
-
-MAJOR POINTS for the following Figure:
-
-- coverage for the model-based estimator is slightly higher than coverage for the design-based estimator in the normal settings.
-Coverages are about equal in the lognormal settings with a slight edge to model-based (this is the point that will be tougher to explain: I would have expected the design-based estimator to have better coverage in the lognormal settings because it has fewer assumptions).
-- coverage is at or near the nominal 95% in all of the normal settings, where assumptions for the model approach and the design approach are satisfied.
 - for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean.
@@
 
-```{r}
-ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
-  geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 0.19, size = 2.5) + 
-  facet_grid(psill_ratio ~ resptype) +
-  scale_colour_manual(values = colour_scale) +
-  ylim(0.80, 1) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Interval Coverage")
-```
+-Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
 
-# Application
-
-The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
-Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.
-Therefore, we will explore whether or not we can still obtain an adequately precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.
-
-```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
-library(tidyverse)
-library(here)
-library(SpatialDesignVModel)
-data(nla)
-nla_df
-factor(nla_df$TOTALHG_UNITS) ## all same units
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
-## HG never measured twice at the same site.
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-
-## in duplicate sites, fill in the value for RDis_IX
-nla_nomiss <- nla_df %>%
-  group_by(SITE_ID) %>%
-  fill(RDis_IX, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-nla_nomiss_both <- nla_nomiss
-## check
-nla_nomiss_both %>% group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-
-## transform coordinates
-library(sptotal)
-coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
-                     lat = nla_nomiss_both$INDEX_LAT_DD, 
-                     lon = nla_nomiss_both$INDEX_LON_DD)
-nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
-nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-```
-
-```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.
-Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.", message = FALSE, warning = FALSE}
-merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(colour = "black", fill = "white", bins = 15) +
-  labs(x = "HG (ng / g)")
-# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
-#   geom_histogram(colour = "black", fill = "white", bins = 15) +
-#   labs(x = "Lakeshore Disturbance")
-merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "hg") +
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank()) +
-  coord_quickmap() 
-# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-#   geom_point(aes(colour = RDis_IX)) +
-#   scale_colour_viridis_c() +
-#   labs(x = "", y = "", colour = "Dist") +
-#   theme(axis.text.x = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.ticks = element_blank())
-library(gridExtra)
-# grid.arrange(merc, merc_map, nrow = 2)
-merc_map
-```
-
-```{r, echo = FALSE, results = "hide"}
-mean(nla_nomiss_both$total_hg)
-# mean(nla_nomiss_both$RDis_IX)
-```
-
-Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.2 ng / g.
-
-```{r, results = "hide", fig.keep = "none"}
-## IRS Sample, Model Analysis
-
-set.seed(080421)
-n <- 100
-lakesobs <- nla_nomiss_both %>% sample_n(n) 
-lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
-lakesunobs$total_hg <- NA
-
-lakes_test <- bind_rows(lakesobs, lakesunobs) 
-lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
-
-slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
-                              data = lakes_test, 
-                              xcoordcol = 'xcoords',
-                              ycoordcol = 'ycoords',
-                              CorModel = "Exponential")
-summary(slmfitout_exp_lakes)
-plot(slmfitout_exp_lakes)
-
-pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
-                          conf_level = 0.95)
-print(pred_exp_lakes)
-mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
-se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
-lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
-ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
-
-realized_mean <- mean(nla_nomiss_both$total_hg) ## compute realized mean
-```
-
-```{r, results = "hide"}
-## IRS Sample, Design Analysis
-## 
-irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
-                                   vardoc = var(total_hg))
-N <- nrow(nla_nomiss_both); n <- nrow(lakesobs)
-irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
-irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
-irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
-irs_info$meandoc; sqrt(irs_info$vardoc)
-irs_lb_irs_samp; irs_ub_irs_samp
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Design Analysis
-data_sf <- sf::st_as_sf(nla_nomiss_both, coords = c("xcoords", "ycoords"),
-                        crs = 5070)
-library(spsurvey)
-grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sprbind(grts_samp)
-
-## get coordinates
-grts_coords <- sf::st_coordinates(grts_bind)
-## make data frame
-grts_df <- data.frame(
-  response = grts_bind$total_hg,
-  x = grts_coords[, "X"],
-  y = grts_coords[, "Y"],
-  siteID = grts_bind$siteID,
-  wgt = grts_bind$wgt
-)
-## head(grts_df)
-summary(grts_df$response)
-
-design_analysis <- cont_analysis(
-  grts_df,
-  siteID = "siteID",
-  vars = "response",
-  weight = "wgt",
-  xcoord = "x",
-  ycoord = "y"
-)
-## just return the mean info
-## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
-design_mean_grts <- design_analysis$Mean
-design_mean_grts$Estimate
-design_mean_grts$StdError
-design_mean_grts$LCB95Pct
-design_mean_grts$UCB95Pct
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Model Analysis
-
-grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
-  rename(total_hg = "response", xcoords = "x", ycoords = "y")
-grts_unsamp <- anti_join(nla_nomiss_both, grts_coords_resp)
-grts_unsamp$total_hg <- NA
-grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
-grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
-
-grts_full$wts <- 1 / nrow(grts_full)
-mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
-                   xcoordcol = "xcoords",
-                   ycoordcol = "ycoords")
-pred_mod_grts <- predict(mod_grts, wtscol = "wts")
-model_mean_grts <- pred_mod_grts$FPBK_Prediction
-model_se_grts <- sqrt(pred_mod_grts$PredVar)
-model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
-model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
-model_mean_grts
-model_se_grts
-model_lb_grts
-model_ub_grts
-```
-
-```{r, results = "hide"}
-## combine results
-res_df <- tibble(approach = c("IRS-Design", "GRTS-Design", "IRS-Model", "GRTS-Model"),
-       realized_mean = c(realized_mean, realized_mean, realized_mean,
-                     realized_mean),
-       estimate = c(irs_info$meandoc,
-                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
-                    as.vector(model_mean_grts)),
-       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
-              as.vector(model_se_grts)),
-       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
-              as.vector(model_lb_grts)),
-       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
-              as.vector(model_ub_grts)))
-res_df <- res_df %>% slice(1, 3, 2, 4)
-```
-
-```{r apptab}
-library(knitr)
-res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-kable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 103.2 ng / g")
-```
-
-Table \ref{tab:appliedtab} shows the application of a design-based analysis on an IRS, a model-based analysis on an IRS, a design-based analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
-We see that, for all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
-However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
-
-But, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not use the spatial correlation in mercury concentration across the contiguous United States.
-We also see that, for the samples drawn, the both analyses with the GRTS sampling design have a lower standard error than the analyses with the IRS sampling design.
-We would expect this to be the case for most samples because mercury concentration exhibits spatial correlation so a spatially balanced sample should usually yield a lower standard error.
-If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling just 50 lakes to save time and money.
-
-# Discussion {#sec:discussion}
-
-* Pros of Design-Based (items we are not exploring): computationally efficient, few assumptions, more naturally handles binary data,
-
-* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?, estimated spatial parameters to better understand spatial structure, site-by-site predictions/prediction map
-
-# References {#references .unnumbered}
+- (mike: may get in trouble for this one) If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling only 100 lakes in the future to save time and money.

### 2021-10-13 inst/scripts/application.Rmd
section 3 and 4 edits

@@
 grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sprbind(grts_samp)
+grts_bind <- sp_rbind(grts_samp)
 

### 2021-10-13 R/data.R

@@ -56,7 +56,7 @@
 #' }
 #'
 #' @examples
-#' data(nla)
+#' data(nla_df)
 #' names(nla_df)
 #' summary(nla_df)
 "nla_df"

### 2021-10-13 R/sim_pop_both.R

@@ -69,8 +69,8 @@ sim_pop_both <- function(N = 100,
   }
 
   ## return data
-  data_grid <- data %>% dplyr::filter(method == "grid")
-  data_rand <- data %>% dplyr::filter(method == "rand")
+  data_grid <- data[data$method == "grid", ] # using dplyr::filter gives note about NSE
+  data_rand <- data[data$method == "rand", ]
   data <- list(data_grid, data_rand)
   return(data)
 }

### 2021-10-13 R/sim_trial.R

@@ -45,7 +45,7 @@
 
 ## library(devtools)
 ## install_github("https://github.com/highamm/sptotal")
-## install_github('USEPA/spsurvey', ref = "develop_5.0")
+## install_github('USEPA/spsurvey', ref = "main")
 ## library(sptotal)
 ## library(spsurvey)
 ## library(dplyr)
@@ -75,7 +75,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   irs_samp <- irs(data_sf, n_base = n, ...)
 
   # put it together
-  irs_bind <- sprbind(irs_samp, ...)
+  irs_bind <- sp_rbind(irs_samp, ...)
 
   # get coordinates
   irs_coords <- sf::st_coordinates(irs_bind)
@@ -117,7 +117,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   grts_samp <- grts(data_sf, n_base = n, ...)
 
   # put it together
-  grts_bind <- sprbind(grts_samp, ...)
+  grts_bind <- sp_rbind(grts_samp, ...)
 
   # get coordinates
   grts_coords <- sf::st_coordinates(grts_bind)
@@ -170,7 +170,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   ## IRS Sample IRS Analysis
   ###############################
 
-  irs_analysis <- cont_analysis(irs_df, vars = "response", siteID = "siteID", weight = "wgt", vartype = "SRS", fpc = N, statistics = "mean")$Mean
+  irs_analysis <- cont_analysis(irs_df, vars = "response", siteID = "siteID", weight = "wgt", vartype = "SRS", fpc = N, statistics = "Mean")$Mean
   irs_analysis_mean <- irs_analysis$Estimate
   irs_analysis_se <- irs_analysis$StdError
   irs_analysis_lb <- irs_analysis$LCB95Pct
@@ -195,7 +195,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   ###############################
 
   grts_analysis <- cont_analysis(grts_df, vars = "response", siteID = "siteID",
-                                 weight = "wgt", xcoord = "x", ycoord = "y", statistics = "mean")$Mean
+                                 weight = "wgt", xcoord = "x", ycoord = "y", statistics = "Mean")$Mean
   grts_analysis_mean <- grts_analysis$Estimate
   grts_analysis_se <- grts_analysis$StdError
   grts_analysis_lb <- grts_analysis$LCB95Pct

### 2021-10-13 R/sim_trial_both.R

@@ -75,7 +75,7 @@ sim_trial_both <- function(seed = sample.int(1e7, size = 1),
     irs_samp <- irs(data_sf, n_base = n)
 
     # put it together
-    irs_bind <- sprbind(irs_samp)
+    irs_bind <- sp_rbind(irs_samp)
 
     # get coordinates
     irs_coords <- sf::st_coordinates(irs_bind)
@@ -117,7 +117,7 @@ sim_trial_both <- function(seed = sample.int(1e7, size = 1),
     grts_samp <- grts(data_sf, n_base = n)
 
     # put it together
-    grts_bind <- sprbind(grts_samp)
+    grts_bind <- sp_rbind(grts_samp)
 
     # get coordinates
     grts_coords <- sf::st_coordinates(grts_bind)
@@ -170,7 +170,7 @@ sim_trial_both <- function(seed = sample.int(1e7, size = 1),
     ## IRS Sample IRS Analysis
     ###############################
 
-    irs_analysis <- cont_analysis(irs_df, vars = "response", siteID = "siteID", weight = "wgt", vartype = "SRS", fpc = N, statistics = "mean")$Mean
+    irs_analysis <- cont_analysis(irs_df, vars = "response", siteID = "siteID", weight = "wgt", vartype = "SRS", fpc = N, statistics = "Mean")$Mean
     irs_analysis_mean <- irs_analysis$Estimate
     irs_analysis_se <- irs_analysis$StdError
     irs_analysis_lb <- irs_analysis$LCB95Pct
@@ -195,7 +195,7 @@ sim_trial_both <- function(seed = sample.int(1e7, size = 1),
     ###############################
 
     grts_analysis <- cont_analysis(grts_df, vars = "response", siteID = "siteID",
-                                   weight = "wgt", xcoord = "x", ycoord = "y", statistics = "mean")$Mean
+                                   weight = "wgt", xcoord = "x", ycoord = "y", statistics = "Mean")$Mean
     grts_analysis_mean <- grts_analysis$Estimate
     grts_analysis_se <- grts_analysis$StdError
     grts_analysis_lb <- grts_analysis$LCB95Pct

### 2021-10-13 inst/manuscript/manuscript.Rmd
section 5 edits

@@
 
-* Pros of Design-Based (items we are not exploring): computationally efficient, few assumptions, more naturally handles binary data,
+The design-based and model-based approaches to inference are fundamentally different paradigms by which to select samples and analyze data.
+The design-based approach incorporates randomness through sampling to estimate a population parameter.
+The model-based approach incorporates randomness through distributional assumptions to predict the realized value of a random variable.
+Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
+Aside from the theoretical differences described, a few analytical findings were particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); Independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation increasingly than IRS as the strength of spatial correlation increases; the gap in relative efficiency between design-GRTS and model-GRTS widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
 
-* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?, estimated spatial parameters to better understand spatial structure, site-by-site predictions/prediction map
+There are several benefits and drawbacks of the design-based and model-based approaches for spatial data, some of which we have not yet discussed but are worthy of consideration in future research.
+The design-based approach relies on few assumptions, while the model-based approach relies on rigid distributional ones.
+The Horvitz-Thompson estimator of means and totals is unbiased and computationally efficient, while model-based estimation of covariance parameters can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
+The design-based approach also more naturally handles binary data, free from the more complicated logistic regression formulation commonly used to handle binary data in a model-based approach.
+The model-based approach, however, can quantify the relationship between covariates (predictor variables) and the response variable, something the design-based approach cannot do naturally.
+The model-based approach also yields estimated spatial covariance parameters, which help better understand the process of study.
+Model selection is also possible using model-based approaches and criteria such as likelihood ratio tests or AIC [@akaike1974new].
+Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
+Mdoel-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
+The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be heavily considered when choosing among them.
+This is especially true from an analysis perspective, as we found that using a spatially balanced sampling algorithm benefits both design-based and model-based analyses.
 

### 2021-10-13 inst/manuscript/manuscript.tex
section 5 edits

@@
 
-\begin{itemize}
-\item
-  Pros of Design-Based (items we are not exploring): computationally
-  efficient, few assumptions, more naturally handles binary data,
-\item
-  Pros of Model-Based (items we are not exploring): covariate inference,
-  more efficient small-area estimation, model selection?, estimated
-  spatial parameters to better understand spatial structure,
-  site-by-site predictions/prediction map
-\end{itemize}
+The design-based and model-based approaches to inference are
+fundamentally different paradigms by which to select samples and analyze
+data. The design-based approach incorporates randomness through sampling
+to estimate a population parameter. The model-based approach
+incorporates randomness through distributional assumptions to predict
+the realized value of a random variable. Though these approaches have
+often been compared in the literature both from theoretical and
+analytical perspectives, our contribution lies in studying them in a
+spatial context while implementing spatially balanced sampling. Aside
+from the theoretical differences described, a few analytical findings
+were particularly notable: the design decision (GRTS vs IRS) seems much
+more important than the analysis decision (design-based vs model-based);
+Independent of the analysis approach, there is no reason to prefer IRS
+over GRTS for spatial data -- GRTS tends to perform at least as well as
+IRS when there is no spatial correlation increasingly than IRS as the
+strength of spatial correlation increases; the gap in relative
+efficiency between design-GRTS and model-GRTS widens as the strength of
+spatial correlation increases; and when the data are skewed, interval
+coverage for all approaches improves both as the sample size increases
+and as the strength of correlation increases.
+
+There are several benefits and drawbacks of the design-based and
+model-based approaches for spatial data, some of which we have not yet
+discussed but are worthy of consideration in future research. The
+design-based approach relies on few assumptions, while the model-based
+approach relies on rigid distributional ones. The Horvitz-Thompson
+estimator of means and totals is unbiased and computationally efficient,
+while model-based estimation of covariance parameters can be
+computationally burdensome, especially for likelihood-based methods such
+as REML that rely on inverting a covariance matrix. The design-based
+approach also more naturally handles binary data, free from the more
+complicated logistic regression formulation commonly used to handle
+binary data in a model-based approach. The model-based approach,
+however, can quantify the relationship between covariates (predictor
+variables) and the response variable, something the design-based
+approach cannot do naturally. The model-based approach also yields
+estimated spatial covariance parameters, which help better understand
+the process of study. Model selection is also possible using model-based
+approaches and criteria such as likelihood ratio tests or AIC (Akaike,
+1974). Model-based approaches are capable of more efficient small-area
+estimation than design-based approaches by leveraging distributional
+assumptions in areas with few observed sites. Mdoel-based approaches can
+also compute site-by-site predictions at unobserved locations and use
+them to construct informative visualizations. The benefits and drawbacks
+of both approaches, alongside our theoretical and analytical
+comparisons, should be heavily considered when choosing among them. This
+is especially true from an analysis perspective, as we found that using
+a spatially balanced sampling algorithm benefits both design-based and
+model-based analyses.
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-akaike1974new}{}%
+Akaike, H., 1974. A new look at the statistical model identification.
+IEEE transactions on automatic control 19, 716--723.
+
 \leavevmode\hypertarget{ref-barabesi2011sampling}{}%

### 2021-10-13 inst/manuscript/manuscript.Rmd
spell check and minor updates

@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neigborhood variance estimator tends reduce $\text{Var}(\hat{\tau})$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
+@stevens2003variance show the local neighborhood variance estimator tends reduce $\text{Var}(\hat{\tau})$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
 
@@
 \end{align}
-where $\sigma^2_{ps}$ is the partial sill measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the nugget measuring fine-scale (indepenent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+where $\sigma^2_{ps}$ is the partial sill measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the nugget measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
 Any spatial covariance function could be used in the place of the exponential, however, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
@@
 tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-       nrow = 2) %>%
+       nrow = 2, byrow = TRUE) %>%
   as.table()
@@
 align(xtab1) <- "r|ll"
-names(xtab1) <- c("IRS", "GRTS")
+names(xtab1) <- c("Design", "Model")
 print(xtab1, comment = FALSE)
@@
 Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
-Aside from the theoretical differences described, a few analytical findings were particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); Independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation increasingly than IRS as the strength of spatial correlation increases; the gap in relative efficiency between design-GRTS and model-GRTS widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
+Aside from the theoretical differences described, a few analytical findings were particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); Independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation increasingly than IRS as the strength of spatial correlation increases; the gap in relative efficiency between GRTS-design and GRTS-model widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
 
@@
 Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
-Mdoel-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
+Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
 The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be heavily considered when choosing among them.

### 2021-10-13 inst/manuscript/manuscript.tex
spell check and minor updates

@@
 incorporates spatial locations -- for technical details see Stevens and
-Olsen (2003). Stevens and Olsen (2003) show the local neigborhood
+Olsen (2003). Stevens and Olsen (2003) show the local neighborhood
 variance estimator tends reduce \(\text{Var}(\hat{\tau})\) compared to
@@
 coarse-scale (correlated) variability, \(\sigma^2_{n}\) is the nugget
-measuring fine-scale (indepenent) variability, \(\phi\) is the range
+measuring fine-scale (independent) variability, \(\phi\) is the range
 parameter measuring the distance-decay rate of the covariance, and
@@
   \hline
- & IRS & GRTS \\ 
+ & Design & Model \\ 
   \hline
-IRS & IRS-Design & GRTS-Design \\ 
-  GRTS & IRS-Model & GRTS-Model \\ 
+IRS & IRS-Design & IRS-Model \\ 
+  GRTS & GRTS-Design & GRTS-Model \\ 
    \hline
@@
 strength of spatial correlation increases; the gap in relative
-efficiency between design-GRTS and model-GRTS widens as the strength of
+efficiency between GRTS-design and GRTS-model widens as the strength of
 spatial correlation increases; and when the data are skewed, interval
@@
 estimation than design-based approaches by leveraging distributional
-assumptions in areas with few observed sites. Mdoel-based approaches can
+assumptions in areas with few observed sites. Model-based approaches can
 also compute site-by-site predictions at unobserved locations and use

### 2021-10-19 R/data.R
file cleanup

@@ -1,6 +1,6 @@
 #' National Lakes Assessment Data
 #'
-#' A data set that can be used with the \code{SpatialDesignVModel} package.
+#' A data set that can be used with the \code{DvMsp} package.
 #'
 #' @format A data frame object including:
 #' \describe{

### 2021-10-19 R/utils-pipe.R

@@ -10,5 +10,5 @@
 #' @usage lhs \%>\% rhs
 #' @param lhs A value or the magrittr placeholder.
 #' @param rhs A function call using the magrittr semantics.
-#' @return The result of calling `rhs(lhs)`.
+#' @return The result of calling \code{rhs(lhs)}.
 NULL

### 2021-10-19 inst/scripts/sim_runs.R

@@ -92,8 +92,8 @@ for (i in 1:nrow(parm_df)) {
 
   if (write_out == TRUE) {
     library(readr)
-    write_csv(all_output, glue::glue("inst/output/sims914raw/allsims_raw", i, ".csv"))
-    write_csv(summ_output, glue::glue("inst/output/sims914/allsims", i, ".csv"))
+    write_csv(all_output, glue::glue("inst/output/simulation_raw/allsims_raw", i, ".csv"))
+    write_csv(summ_output, glue::glue("inst/output/simulation_summary/allsims", i, ".csv"))
   }
 
   print(i)

### 2021-10-19 inst/scripts/simulation_exploration.R

@@ -1,84 +1,0 @@
-library(tidyverse)
-library(here)
-
-files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
-
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x))
-
-combo_data <- combo_data %>% mutate(sim = interaction(n, psill, resptype, gridded))
-
-##
-ggplot(data = combo_data, aes(x = sim, y = rmspe, colour = approach)) +
-  geom_point() +
-  coord_flip()
-
-combo_data_grts <- combo_data %>% filter(approach %in% c("Design GRTS", "Model GRTS"))
-
-## the significance (p-values) of these are fairly meaningless, but still
-## might help as a starting point
-summary(lm(rmspe ~ approach, data = combo_data_grts))
-
-## evidence that independent error gives worse estimators (makes sense)
-## evidence that normal distribution gives better estimators (makes sense)
-## evidence that n of 50 gives worse estimators (makes sense)
-
-summary(lm(rmspe ~ approach + resptype + psill + n + gridded,
-           data = combo_data_grts))
-
-
-## visually explore the two approaches
-
-ggplot(data = combo_data_grts, aes(x = sim, y = rmspe, colour = approach)) +
-  geom_point() +
-  coord_flip()
-
-dep_only <- combo_data_grts %>% filter(psill == max(psill))
-dep_long <- dep_only %>%
-  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
-  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
-ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
-  geom_point() +
-  coord_flip() +
-  labs(x = "Percent Change in rmspe (Design - Model) / Model",
-       caption = "Percent > 0 means Model has better rmspe")
-
-combo_perc <- combo_data_grts %>%
-  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
-  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
-ggplot(data = combo_perc, aes(x = fct_reorder(sim, perc), y = perc)) +
-  geom_point() +
-  coord_flip()
-
[…]

### 2021-10-19 inst/manuscript/manuscript.tex

@@
 
-  \title{A comparison of design-based and model-based approaches for finite
-population spatial data.}
+  \title{A comparison of design-based and model-based approaches for
+finite population spatial data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
-and Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Mathematics,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
       \cortext[1]{Corresponding Author}
@@
 design-based and model-based for spatial data have been considered (Brus
-and De Gruijter, 1997; Ver Hoef, 2002, 2008). Cooper (2006) review the
-two approaches in an ecological context before introducing a
+and De Gruijter, 1997; Ver Hoef, 2002; Ver Hoef, 2008). Cooper (2006)
+review the two approaches in an ecological context before introducing a
 ``model-assisted'' variance estimator that combines aspects from each
@@
 research and development into estimators that use both design and
-model-based principles (see e.g.~Cicchitelli and Montanari (2012),
+model-based principles (see e.g. Cicchitelli and Montanari (2012),
 Chan-Golston et al. (2020) for a Bayesian approach, and Sterba (2009)).
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-akaike1974new}{}%
@@
 \leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
-Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
-under preferential sampling. Journal of the Royal Statistical Society:
-Series C (Applied Statistics) 59, 191--232.
+Diggle, P.J., Menezes, R., Su, T., 2010. Geostatistical inference under
+preferential sampling. Journal of the Royal Statistical Society: Series
+C (Applied Statistics) 59, 191--232.
 

### 2021-10-19 inst/manuscript/manuscript.Rmd
added supplementary material

@@
 <!-- Brief introduction to model based and design based -->
+
+Add sentence about bias.
+Add sentence about user functions.
 
@@
 We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
+All scenarios used exponential correlation with a  $\sqrt{2} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
+The range was chosen so that the decayed to nearly zero at the largest distance possible in the domain, $\sqrt{2}$, otherwise known as the "effective" range (for the exponential covariance, the effective range is $3\phi$).
 The mean for the spatial process generating the response was set to zero.
@@
 
-files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
+files <- list.files(path = here("inst", "output", "simulation_summary"), pattern = "*.csv", full.names = TRUE)
 

### 2021-10-19 inst/manuscript/manuscript.tex
added supplementary material

--- 2021-10-13 inst/manuscript/manuscript.tex
@@
 \section{Introduction}\label{sec:introduction}}
+
+Add sentence about bias. Add sentence about user functions.
 
@@
 \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
-exponential correlation with an effective range of \(\sqrt{2}\) for
-\(N = 900\) response values simulated on the unit square in either
-random locations (Layout = Random) or gridded locations (Layout =
-Gridded). The mean for the spatial process generating the response was
-set to zero.
+exponential correlation with a \(\sqrt{2} / 3\) for \(N = 900\) response
+values simulated on the unit square in either random locations (Layout =
+Random) or gridded locations (Layout = Gridded). The range was chosen so
+that the decayed to nearly zero at the largest distance possible in the
+domain, \(\sqrt{2}\), otherwise known as the ``effective'' range (for
+the exponential covariance, the effective range is \(3\phi\)). The mean
+for the spatial process generating the response was set to zero.
 

### 2021-10-19 inst/manuscript/supplementary.Rmd
added supplementary material

--- 2021-10-13 inst/manuscript/manuscript.Rmd
@@
 ---
-title: A comparison of design-based and model-based approaches for finite population spatial data.
+title: A comparison of design-based and model-based approaches for finite population spatial data -- Supplementary Material.
 author:
@@
     text: "Corresponding Author"
-abstract: |
-  This is the abstract.
-
 journal: "An awesome journal"
 date: "`r Sys.Date()`"
-bibliography: mybibfile.bib
 linenumbers: true
@@
   \DeclareMathOperator{\cov}{{cov}}
+  \renewenvironment{abstract}{}{}
 output: rticles::elsevier_article
@@
 
-_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
-
-Potential Journals:
-
-* Ecological Applications
-* Methods in Ecology and Evolution
-* Journal of Applied Ecology
-* Environmetrics
-* Environmental and Ecological Statistics
-
-<!-- Notes from today -->
-<!-- •	Finite populations require a smaller scope but we need to discuss the impact of infinite populations as well – put finite vs infinite population context in the introduction / similarities and differences -->
-<!-- •	What should a lower sample size be? 50 seems reasonable 30 may be too small – maybe 50, 100, 200 -->
-<!-- •	IRS results seem off for dependent error sample size 50 -->
-<!-- •	Take home points – don’t have to do GRTS design with GRTS sample -->
-<!-- •	Take home points – GRTS is better for both model and design, while spatial approach for design is close to spatial model -->
-<!-- •	Add medium dependence -->
-<!-- •	Remind readers what choices of the four approaches in the application mean – what do they mean to practitioners -->
-
-# Introduction {#sec:introduction}
+# Numerical Study Tables {#sec:numstudy}
 
@@
 
-<!-- Brief introduction to model based and design based -->
-
-There are two general approaches for using data to make statistical inferences about a population: design-based and model-based.
-When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected and called a sample.
-In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process that generated the data.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
-
-<!-- Shift focus to spatial data -->
-
-Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process.
-@de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review.
-
-<!-- Paragraph about the contribution of the manuscript and outlining the rest -->
-Though comparisons between design-based and model-based approaches to spatial data have been studied, no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
-In this manuscript, we compare design-based approaches that incorporate spatial locations to model-based approaches for spatial data.
-Though these comparisons generalize to both finite populations (e.g.
-point resources) and infinite populations (e.g.
-linear and areal resources), we focus on applications to finite populations.
-The rest of the manuscript is organized as follows.
-In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach for spatial data.
-In Section \ref{sec:numstudy}, we use a simulation approach to study the behavior and performance of both approaches.
-In Section \ref{application}, we use both approaches to analyze real data.
-And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
-
-# Background {#sec:background}
-
-The design-based and model-based approaches incorporate randomness in fundamentally different ways.
-In this section, we describe the role of randomness and its effects on subsequent inferences.
-We then discuss specific inference methods for the design-based and model-based approaches for spatial data.
-
-## Comparing Design-Based and Model-Based Approaches {#subsec:dvm_compare}
-
-The design-based approach assumes the population is fixed.
-Randomness is incorporated via the selection of units in a sampling frame according to a sampling design.
-A sampling frame is the set of all units available to be sampled.
-A sampling design assigns a positive probability of inclusion (inclusion probability) to each unit in the sampling frame.
-Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster sampling.
-These sampling designs tend to select units from the sampling frame independently of other units, so we call them "Independent Random Sampling" (IRS) designs.
-Sampling designs incorporating the spatial locations of units in the sample frame are called spatially balanced designs.
-Spatially balanced designs can be obtained using the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
-The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate parameters (e.g.
-means and totals) of a population.
-Generally, these population parameters are assumed to be fixed, unknown constants.
-
-Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design.
-Means and totals, for example, are asymptotically normally distributed (normal) by the Central Limit Theorem (under some assumptions).
-If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
-@sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
-
-The model-based approach assumes the data are a random realization of a data-generating process.
-Randomness is incorporated through distributional assumptions on this process.
-Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
-To guard against preferential sampling, model-based approaches often still implement random sampling.
-
-Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
-For example, suppose the realized mean of all population units is the value of interest.
-Instead of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-Prediction intervals are then derived leveraging assumptions of the data generating process.
-If we repeatedly generate the response values from a fixed spatial process and obtained a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
-@cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@brus2020statistical provides a similar figure).
-
-```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
-Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of n = 4.
-The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
-The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels."}
-set.seed(10102021)
-
-source(here("R", "sim_pop.R"))
-N <- 9
-pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                  psill = 0.9, erange = 1, nugget = 0.1)
-
-n <- 4
-samp_df <- pop_df %>% sample_n(n) 
-samp_df2 <- pop_df %>% sample_n(n) 
-samp_df3 <- pop_df %>% sample_n(n)
-
-library(cowplot)
-d1 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df, size = 5.5, shape = 1, 
-             show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d2 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df2, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d3 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df3, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                       psill = 0.9, erange = 1, nugget = 0.1)
-pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
-
-indeces <- sample(1:N, size = n, replace = FALSE)
-samp_df_mod <- pop_df_mod %>% slice(indeces)
-samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
-samp_df_mod3 <- pop_df_mod3 %>% slice(indeces)
-
-m1 <- ggplot(data = pop_df_mod, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m2 <- ggplot(data = pop_df_mod2, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m3 <- ggplot(data = pop_df_mod3, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-library(ggpubr)
-ggarrange(d1, d2, d3, m1, m2, m3, nrow = 2, ncol = 3,
-          common.legend = TRUE, legend = "right")
-```
-
-## Spatially Balanced Design and Analysis {#subsec:spb_design}
-
-Spatially balanced samples can be obtained using the design-based approach.
-Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
-To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voroni polygons.
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we use Generalized Random Tessellation Stratified (GRTS) sampling because it has several attractive properties: GRTS sampling accommodates finite and infinite sampling frames; accommodates equal, unequal, and proportional (to) size inclusion probabilities; accommodates legacy (historical) sampling; accommodates a minimum distance between units in a sample; accommodates reverse hierarchically ordered replacement units in a sample (replacement units are units available to be sampled if an original unit cannot be sampled); and is available in the spsurvey R package @dumelle2021spsurvey.
-
-The GRTS algorithm samples from finite and infinite populations by utilizing a mapping between two-dimensional and one-dimensional space.
-The units in the two-dimensional sampling frame are divided into cells using a hierarchical address.
-This hierarchical address is then used to map the units from two-dimensional space to a one-dimensional line where each unit's line length equals its inclusion probability.
-A systematic sample is conducted on the line and linked back to a unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially and @dumelle2021spsurvey provide further details.
-
-<!-- This makes it seem like HT is being used to estimate population parameters in GRTS?
-Needs to be in there, but maybe a sentence making it clear that this isn't being used?
-Or just change the order and talk about Local variance first and then mention these for more general settings?
--->
-
-After selecting a GRTS sample, data are collected and used to estimate population parameters.
-To unbiasedly estimate population means and totals from sample data, one can use the Horvitz-Thompson estimator [@horvitz1952generalization].
-If $\tau$ is a population total, the Horvitz-Thompson estimate of $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
-\begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align}
-where $Z_i$ is the value of the $i$th unit in the sample and $\pi_i$ is the inclusion probability of the $i$th unit in the sample.
-An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by the population size.
-
-While the Horvitz-Thompson estimator is unbiased for population means and totals, it is also important to quantify the uncertainty in these estimates.
-@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, these estimators rely on calculating $\pi_{ij}$, the probability that unit $i$ and unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
-Second, these estimators ignore the spatial locations of the units in the sampling frame.
-To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neighborhood variance estimator tends reduce $\text{Var}(\hat{\tau})$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
-
-## Finite Population Block Kriging
-
-Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
-Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process.
-@verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
-.
-.
-, \text{z}(s_N) \}}$ be a response vector at locations $s_1$, $s_2$, .
-.
-.
-, $s_N$ that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
-Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
-For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
-
-<!-- how explicitly do we want to define \mathbf{z} -->
-
-<!-- finite number of -->
-
-<!-- confusing having two tau's: tau for total and a tau as a function -->
-
-We often only have a sample of the $N$ population units.
-Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
-
-\begin{equation}
-\begin{pmatrix} \label{equation:Zmarginal}
-    \mathbf{z}_s      \\
-    \mathbf{z}_u
-\end{pmatrix}
-=
-\begin{pmatrix}
-  \mathbf{X}_s    \\
-  \mathbf{X}_u
-\end{pmatrix}
-\bm{\beta} +
-\begin{pmatrix}
-\bm{\delta}_s    \\
-\bm{\delta}_u
-\end{pmatrix},
-\end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively; $\beta$ is the parameter vector of fixed effects; and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
-Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume the expectation of $\bm{\delta}$ equals $\mathbf{0}$.
-
-In addition to assuming the expectation of $\bm{\delta}$ equals $\mathbf{0}$, we also assume that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
-It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th entry for $\cov(\bm{\delta})$ is
-\mbox{}
-\begin{align}\label{equation:expcov}
-\cov(\delta_i, \delta_j) = 
-\begin{cases} 
-\sigma^2_{ps}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
-\sigma^2_{ps} + \sigma^2_n & h_{i,j} = 0
-\end{cases}
-,
-\end{align}
-where $\sigma^2_{ps}$ is the partial sill measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the nugget measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-Any spatial covariance function could be used in the place of the exponential, however, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
-80-93].
-
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based, meaning they don't rely on any distributional assumptions.
-
-We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
-We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
-
-# Numerical Study {#sec:numstudy}
-
-We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.
+Recall the four design-analysis combinations and the parameter configurations, summarized in Table \ref{tab:designanalysis} and \ref{tab:parmtab}, respectively.
 
@@
 ```
-
-We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
-The mean for the spatial process generating the response was set to zero.
-
-For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
-A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
-Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
-We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.
 
@@
 
-```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
-mu <- 0; var <- 2
-xvar <- seq(0.001, 10, by = 0.01)
-yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
-df <- tibble(xvar = xvar, yvar = yvar)
-ggplot(df, aes(x = xvar, y = yvar)) +
-  geom_line()
-```
-
-There were 2000 simulation trials for each of the 36 parameter combinations.
-In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected.
-For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model).
-For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).
-
 ```{r, results = "asis", eval = FALSE}
@@
 
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2021sptotal].
-We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
-
-**Mike** For design-based, it's really RMSE -- how should we address this?
-Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
-\mbox{}
-\begin{equation*}
-\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
-\end{equation*}
-
-where rMSPE is the root-mean-squared-prediction error.
-When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
-So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small.
-In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
-
-Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design.
-However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
-
-```{r figeff, out.width = "100%", fig.cap = "Relative Efficiency of the four design-analysis approaches.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
-library(tidyverse)
-library(here)
-
-files <- list.files(path = here("inst", "output", "sims914"), pattern = "*.csv", full.names = TRUE)
-
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x))
-
-combo_data <- combo_data %>%
-  dplyr::filter(n != 10 & gridded == FALSE) %>%
-  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
-  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
-                                            "Design GRTS", "Model GRTS"))) %>%
-  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal")))
-combo_data <- combo_data %>% group_by(sim) %>% mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
-  fill(designirsrmspe, .direction = "downup") %>%
-  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  ungroup() %>%
-  mutate(psill_ratio = 1 - nugget_ratio) %>%
-  mutate(n_factor = factor(n))
-
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
-                              colour = approach)) +
-  facet_grid(psill_ratio ~ resptype) +
-  geom_jitter(width = 0.24, size = 2.5) +
-  scale_colour_manual(values = colour_scale) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Relative Efficiency")
-effplot
+Next we present a summary table for each of the 36 simulation scenarios.
+```{r}
+make_sim_table <- function(table_value) {
+  sim_table <- read_csv(here("inst", "output", "simulation_summary", paste0("allsims", table_value, ".csv")))
+  sim_table <- sim_table %>%
+    dplyr::select(Approach = approach,
+                  Layout = gridded,
+                  Response = resptype,
+                  `DE Prop` = psill_ratio,
+                  n = n,
+                  Bias = mpbias,
+                  `rMS(P)E` = rmspe,
+                  Coverage = coverage) %>%
+    mutate(
+     Approach = case_when(
+       Approach == "Design GRTS" ~ "GRTS-Design",
+       Approach == "Design IRS" ~ "IRS-Design",
+       Approach == "Model GRTS" ~ "GRTS-Model",
+       Approach == "Model IRS" ~ "IRS-Model"
+     ),
+     Layout = case_when(
+        Layout ~ "Gridded",
+        !Layout ~ "Random"
+      ),
+     Response = case_when(
+       Response == "normal" ~ "Normal",
+       Response == "lognormal" ~ "Lognormal"
+     ),
+     across(c(`DE Prop`, n), as.character)
+    )
+sim_table <- sim_table[c(2, 4, 1, 3), ]
+sim_table_out <- xtable(
+  sim_table,
+  digits = 4,
+  caption = "Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\\% interval coverage (Coverage) for a simulation scenario.",
+   type = "latex",
+  latex.environments = "center"
+)
+print(sim_table_out, include.rownames = FALSE, comment = FALSE)
+}
 ```
 
-```{r, fig.keep = "none"}
-combo_data_grts <- combo_data %>%
-  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
-  mutate(designgrts = if_else(approach == "Design GRTS", 
-                              true = rmspe,
-                              false = NA_real_)) %>%
-  group_by(sim) %>%
-  fill(designgrts, .direction = "updown") %>%
-  ungroup() %>%
-  mutate(releffgrts = rmspe / designgrts)
-combo_data_grts_model <- combo_data_grts %>%
-  filter(approach == "Model GRTS")
-
-ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
-  facet_grid(psill ~ resptype) +
-  geom_point()
+```{r, results = "asis"}
+for (i in 13:48) {
+  make_sim_table(i)
+}
 ```
-
-We also studied 95% interval coverage among the approaches.
-The design-based 95% confidence intervals and model-based 95% prediction intervals are constructed using the normal distribution.
-Justification for the design-based intervals lies in the asymptotic normality of totals via the Central Limit Theorem, and justification for the model-based intervals lies in the normality assumption of the errors.
-Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches.
-All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
-
-In the normal response settings, all approaches have coverage around 95%.
-This is expected, as the intervals are also based on the normal distribution In the lognormal response settings, however, all approaches have coverage below 95%.
-This is also expected, as the intervals are still based on the normal distribution.
-In the lognormal response settings, interval coverage increases both as the sample size increases and as the strength of spatial dependence increases.
-This suggests that the larger the sample size and the stronger the spatial dependence, the more resistant these intervals are to departures from normality of the data.
-
-```{r figconf, out.width = "100%", fig.cap = "Coverage of the four design-analysis approaches.
-All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
-ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
-  geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 0.24, size = 2.5) + 
-  facet_grid(psill_ratio ~ resptype) +
-  scale_colour_manual(values = colour_scale) +
-  ylim(0.80, 1) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Interval Coverage")
-```
-
-# Application {#application}
-
-The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
-Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-
-```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
-library(tidyverse)
-library(here)
-library(SpatialDesignVModel)
-data(nla_df) # may need to reinstall package
-nla_df
-factor(nla_df$TOTALHG_UNITS) ## all same units
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
-## HG never measured twice at the same site.
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-
-## in duplicate sites, fill in the value for RDis_IX
-nla_nomiss <- nla_df %>%
-  group_by(SITE_ID) %>%
-  fill(RDis_IX, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-nla_nomiss_both <- nla_nomiss
-## check
-nla_nomiss_both %>% group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-
-## transform coordinates
-library(sptotal)
-coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
-                     lat = nla_nomiss_both$INDEX_LAT_DD, 
-                     lon = nla_nomiss_both$INDEX_LON_DD)
-nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
-nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-```
-
-```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.", message = FALSE, warning = FALSE}
-merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(colour = "black", fill = "white", bins = 15) +
-  labs(x = "HG (ng / g)")
-# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
-#   geom_histogram(colour = "black", fill = "white", bins = 15) +
-#   labs(x = "Lakeshore Disturbance")
-merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "hg") +
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank()) +
-  coord_quickmap() +
-  theme_bw(base_size = 14)
-# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-#   geom_point(aes(colour = RDis_IX)) +
-#   scale_colour_viridis_c() +
-#   labs(x = "", y = "", colour = "Dist") +
-#   theme(axis.text.x = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.ticks = element_blank())
-library(gridExtra)
-# grid.arrange(merc, merc_map, nrow = 2)
-merc_map
-```
-
-```{r, echo = FALSE}
-mean_hg <- mean(nla_nomiss_both$total_hg)
-# mean(nla_nomiss_both$RDis_IX)
-```
-
-Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-The realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
-
-```{r, results = "hide", fig.keep = "none"}
-## IRS Sample, Model Analysis
-
-set.seed(080421)
-n <- 100
-lakesobs <- nla_nomiss_both %>% sample_n(n) 
-lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
-lakesunobs$total_hg <- NA
-
-lakes_test <- bind_rows(lakesobs, lakesunobs) 
-lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
-
-slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
-                              data = lakes_test, 
-                              xcoordcol = 'xcoords',
-                              ycoordcol = 'ycoords',
-                              CorModel = "Exponential")
-summary(slmfitout_exp_lakes)
-plot(slmfitout_exp_lakes)
-
-pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
-                          conf_level = 0.95)
-print(pred_exp_lakes)
-mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
-se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
-lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
-ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
-
-realized_mean <- mean(nla_nomiss_both$total_hg) ## compute realized mean
-```
-
-```{r, results = "hide"}
-## IRS Sample, Design Analysis
-## 
-irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
-                                   vardoc = var(total_hg))
-N <- nrow(nla_nomiss_both); n <- nrow(lakesobs)
-irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
-irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
-irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
-irs_info$meandoc; sqrt(irs_info$vardoc)
-irs_lb_irs_samp; irs_ub_irs_samp
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Design Analysis
-data_sf <- sf::st_as_sf(nla_nomiss_both, coords = c("xcoords", "ycoords"),
-                        crs = 5070)
-library(spsurvey)
-grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sp_rbind(grts_samp) # Matt you need to install newest version
-# devtools::install_github("USEPA/spsurvey", ref = "main")
-
-## get coordinates
-grts_coords <- sf::st_coordinates(grts_bind)
-## make data frame
-grts_df <- data.frame(
-  response = grts_bind$total_hg,
-  x = grts_coords[, "X"],
-  y = grts_coords[, "Y"],
-  siteID = grts_bind$siteID,
-  wgt = grts_bind$wgt
-)
-## head(grts_df)
-summary(grts_df$response)
-
-design_analysis <- cont_analysis(
-  grts_df,
-  siteID = "siteID",
-  vars = "response",
-  weight = "wgt",
-  xcoord = "x",
-  ycoord = "y"
-)
-## just return the mean info
-## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
-design_mean_grts <- design_analysis$Mean
-design_mean_grts$Estimate
-design_mean_grts$StdError
-design_mean_grts$LCB95Pct
-design_mean_grts$UCB95Pct
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Model Analysis
-
-grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
-  rename(total_hg = "response", xcoords = "x", ycoords = "y")
-grts_unsamp <- anti_join(nla_nomiss_both, grts_coords_resp)
-grts_unsamp$total_hg <- NA
-grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
-grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
-
-grts_full$wts <- 1 / nrow(grts_full)
-mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
-                   xcoordcol = "xcoords",
-                   ycoordcol = "ycoords")
-pred_mod_grts <- predict(mod_grts, wtscol = "wts")
-model_mean_grts <- pred_mod_grts$FPBK_Prediction
-model_se_grts <- sqrt(pred_mod_grts$PredVar)
-model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
-model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
-model_mean_grts
-model_se_grts
-model_lb_grts
-model_ub_grts
-```
-
-```{r, results = "hide"}
-## combine results
-res_df <- tibble(approach = c("IRS-Design", "GRTS-Design", "IRS-Model", "GRTS-Model"),
-       realized_mean = c(realized_mean, realized_mean, realized_mean,
-                     realized_mean),
-       estimate = c(irs_info$meandoc,
-                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
-                    as.vector(model_mean_grts)),
-       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
-              as.vector(model_se_grts)),
-       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
-              as.vector(model_lb_grts)),
-       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
-              as.vector(model_ub_grts)))
-res_df <- res_df %>% slice(1, 3, 2, 4)
-```
-
-```{r apptab, results = "asis"}
-res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 ng / g.")
-print(xtab_app, include.rownames = FALSE, comment = FALSE)
-```
-
-Table \ref{tab:appliedtab} shows the application of a design-based analysis of an IRS sample, a model-based analysis of an IRS sample, a design-based analysis of a GRTS sample, and a model-based analysis of a GRTS sample.
-For all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
-However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
-
-<!-- not sure why this is getting cutoff. temporary fix: \pagebreak 
-probably because it is an html table -- I used xtable to fix -->
-
-But, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration across the contiguous United States.
-We also see that both approaches using the GRTS sample have a lower standard error than the both approaches using the IRS sample.
-We would expect this to be the case for most samples because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.
-
-# Discussion {#sec:discussion}
-
-The design-based and model-based approaches to inference are fundamentally different paradigms by which to select samples and analyze data.
-The design-based approach incorporates randomness through sampling to estimate a population parameter.
-The model-based approach incorporates randomness through distributional assumptions to predict the realized value of a random variable.
-Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
-Aside from the theoretical differences described, a few analytical findings were particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); Independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation increasingly than IRS as the strength of spatial correlation increases; the gap in relative efficiency between GRTS-design and GRTS-model widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
-
-There are several benefits and drawbacks of the design-based and model-based approaches for spatial data, some of which we have not yet discussed but are worthy of consideration in future research.
-The design-based approach relies on few assumptions, while the model-based approach relies on rigid distributional ones.
-The Horvitz-Thompson estimator of means and totals is unbiased and computationally efficient, while model-based estimation of covariance parameters can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
-The design-based approach also more naturally handles binary data, free from the more complicated logistic regression formulation commonly used to handle binary data in a model-based approach.
-The model-based approach, however, can quantify the relationship between covariates (predictor variables) and the response variable, something the design-based approach cannot do naturally.
-The model-based approach also yields estimated spatial covariance parameters, which help better understand the process of study.
-Model selection is also possible using model-based approaches and criteria such as likelihood ratio tests or AIC [@akaike1974new].
-Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
-Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
-The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be heavily considered when choosing among them.
-This is especially true from an analysis perspective, as we found that using a spatially balanced sampling algorithm benefits both design-based and model-based analyses.
-
-# References {#references .unnumbered}

### 2021-10-19 inst/manuscript/supplementary.tex
added supplementary material

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supplementary Material.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing

% Pandoc header

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \renewenvironment{abstract}{}{}

\begin{document}
\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for finite
population spatial data -- Supplementary Material.}
    \author[USEPA]{Michael Dumelle\corref{1}}
   \ead{Dumelle.Michael@epa.gov} 
    \author[STLAW]{Matt Higham}
   \ead{mhigham@stlaw.edu} 
    \author[OSU]{Lisa Madsen}

    \author[USEPA]{Anthony R. Olsen}

    \author[NOAA]{Jay M. Ver Hoef}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
and Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
      \cortext[1]{Corresponding Author}

  \begin{abstract}

  \end{abstract}

 \end{frontmatter}

\hypertarget{sec:numstudy}{%
\section{Numerical Study Tables}\label{sec:numstudy}}

Recall the four design-analysis combinations and the parameter
configurations, summarized in Table \ref{tab:designanalysis} and
\ref{tab:parmtab}, respectively.

\begin{table}[ht]
\centering
\begin{tabular}{r|ll}
  \hline
 & Design & Model \\ 
  \hline
IRS & IRS-Design & IRS-Model \\ 
  GRTS & GRTS-Design & GRTS-Model \\ 
   \hline
\end{tabular}
\caption{\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study. The rows give the two types of sampling designs while the columns give the two types of analyses.} 
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{r|lll}
   \hline
Sample Size (n) & 50 & 100 & 200 \\ 
  Layout & Random & Gridded & - \\ 
  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
  Response Type & Normal & Lognormal & - \\ 
   \hline
\end{tabular}
\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.} 
\end{table}

Next we present a summary table for each of the 36 simulation scenarios.

\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 50 & -0.0001 & 0.1939 & 0.9475 \\ 
  IRS-Model & Gridded & Normal & 0 & 50 & -0.0001 & 0.1960 & 0.9370 \\ 
  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0010 & 0.1930 & 0.9175 \\ 
  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0016 & 0.1936 & 0.9405 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 50 & 0.0077 & 1.0498 & 0.8355 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0274 & 1.2011 & 0.8230 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 50 & 0.0092 & 0.9618 & 0.7910 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0069 & 0.9808 & 0.8100 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 50 & 0.0045 & 0.1948 & 0.9410 \\ 
  IRS-Model & Random & Normal & 0 & 50 & 0.0048 & 0.1966 & 0.9360 \\ 
  GRTS-Design & Random & Normal & 0 & 50 & 0.0056 & 0.1943 & 0.9000 \\ 
  GRTS-Model & Random & Normal & 0 & 50 & 0.0057 & 0.1947 & 0.9365 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 50 & 0.0116 & 0.9684 & 0.8205 \\ 
  IRS-Model & Random & Lognormal & 0 & 50 & -0.0235 & 1.0055 & 0.8150 \\ 
  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0005 & 0.9151 & 0.8075 \\ 
  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0175 & 0.9426 & 0.8240 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 50 & -0.0006 & 0.1571 & 0.9600 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 50 & -0.0005 & 0.1509 & 0.9435 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0006 & 0.1463 & 0.9240 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 50 & 0.0006 & 0.1443 & 0.9445 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0154 & 0.8578 & 0.8700 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0209 & 0.9091 & 0.8490 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & -0.0046 & 0.8302 & 0.8400 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0167 & 0.8301 & 0.8495 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 50 & 0.0089 & 0.1720 & 0.9385 \\ 
  IRS-Model & Random & Normal & 0.5 & 50 & 0.0074 & 0.1580 & 0.9310 \\ 
  GRTS-Design & Random & Normal & 0.5 & 50 & 0.0060 & 0.1460 & 0.9145 \\ 
  GRTS-Model & Random & Normal & 0.5 & 50 & 0.0059 & 0.1439 & 0.9365 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0006 & 0.8479 & 0.8670 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0447 & 0.8845 & 0.8485 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 50 & 0.0022 & 0.8663 & 0.8230 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0044 & 0.8620 & 0.8470 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 50 & -0.0021 & 0.1185 & 0.9790 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 50 & -0.0023 & 0.0925 & 0.9475 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 50 & -0.0004 & 0.0957 & 0.9280 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 50 & 0.0015 & 0.0874 & 0.9300 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0191 & 0.6345 & 0.9310 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 50 & 0.0043 & 0.6031 & 0.8850 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & -0.0072 & 0.6496 & 0.8890 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0144 & 0.5873 & 0.8890 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 50 & -0.0022 & 0.1451 & 0.9445 \\ 
  IRS-Model & Random & Normal & 0.9 & 50 & -0.0016 & 0.1018 & 0.9350 \\ 
  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0022 & 0.0963 & 0.9235 \\ 
  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0016 & 0.0881 & 0.9360 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 50 & -0.0279 & 0.7417 & 0.9080 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0479 & 0.6787 & 0.8720 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0101 & 0.5712 & 0.8995 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0079 & 0.5143 & 0.8935 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 100 & -0.0031 & 0.1343 & 0.9430 \\ 
  IRS-Model & Gridded & Normal & 0 & 100 & -0.0035 & 0.1349 & 0.9405 \\ 
  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0016 & 0.1327 & 0.9200 \\ 
  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0015 & 0.1327 & 0.9395 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 100 & 0.0123 & 0.6517 & 0.8595 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 100 & 0.0037 & 0.6698 & 0.8540 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0175 & 0.6309 & 0.8275 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0108 & 0.6450 & 0.8460 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 100 & 0.0058 & 0.1363 & 0.9350 \\ 
  IRS-Model & Random & Normal & 0 & 100 & 0.0062 & 0.1373 & 0.9315 \\ 
  GRTS-Design & Random & Normal & 0 & 100 & 0.0033 & 0.1351 & 0.9175 \\ 
  GRTS-Model & Random & Normal & 0 & 100 & 0.0033 & 0.1350 & 0.9460 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 100 & -0.0187 & 0.6427 & 0.8595 \\ 
  IRS-Model & Random & Lognormal & 0 & 100 & -0.0394 & 0.6726 & 0.8565 \\ 
  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0049 & 0.6366 & 0.8445 \\ 
  GRTS-Model & Random & Lognormal & 0 & 100 & 0.0011 & 0.6431 & 0.8585 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0005 & 0.1070 & 0.9640 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0006 & 0.1041 & 0.9335 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 100 & -0.0010 & 0.1014 & 0.9145 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 100 & -0.0011 & 0.1006 & 0.9365 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0156 & 0.5285 & 0.9060 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0020 & 0.5346 & 0.8800 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0061 & 0.4945 & 0.8595 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0015 & 0.4965 & 0.8705 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 100 & 0.0002 & 0.1169 & 0.9440 \\ 
  IRS-Model & Random & Normal & 0.5 & 100 & 0.0023 & 0.1042 & 0.9450 \\ 
  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0030 & 0.1032 & 0.9140 \\ 
  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0024 & 0.1022 & 0.9335 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0020 & 0.5609 & 0.8895 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0209 & 0.5570 & 0.8835 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 100 & -0.0106 & 0.5261 & 0.8800 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0199 & 0.5358 & 0.8875 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 100 & 0.0008 & 0.0702 & 0.9915 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 100 & 0.0016 & 0.0573 & 0.9510 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0002 & 0.0600 & 0.9315 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 100 & -0.0001 & 0.0540 & 0.9515 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0165 & 0.4844 & 0.9595 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0062 & 0.4274 & 0.8975 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0056 & 0.3959 & 0.9135 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0015 & 0.3764 & 0.9200 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 100 & -0.0034 & 0.1021 & 0.9425 \\ 
  IRS-Model & Random & Normal & 0.9 & 100 & -0.0015 & 0.0611 & 0.9480 \\ 
  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0614 & 0.9275 \\ 
  GRTS-Model & Random & Normal & 0.9 & 100 & -0.0004 & 0.0548 & 0.9410 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0048 & 0.5829 & 0.9275 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0287 & 0.5423 & 0.9205 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 100 & 0.0010 & 0.4351 & 0.9180 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0025 & 0.3975 & 0.9140 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 200 & 0.0005 & 0.0882 & 0.9455 \\ 
  IRS-Model & Gridded & Normal & 0 & 200 & 0.0005 & 0.0882 & 0.9435 \\ 
  GRTS-Design & Gridded & Normal & 0 & 200 & 0.0034 & 0.0882 & 0.9430 \\ 
  GRTS-Model & Gridded & Normal & 0 & 200 & 0.0034 & 0.0882 & 0.9500 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0109 & 0.4122 & 0.8815 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 200 & 0.0102 & 0.4128 & 0.8790 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0128 & 0.4355 & 0.8825 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0132 & 0.4362 & 0.8865 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 200 & -0.0025 & 0.0869 & 0.9500 \\ 
  IRS-Model & Random & Normal & 0 & 200 & -0.0024 & 0.0872 & 0.9505 \\ 
  GRTS-Design & Random & Normal & 0 & 200 & 0.0027 & 0.0875 & 0.9470 \\ 
  GRTS-Model & Random & Normal & 0 & 200 & 0.0027 & 0.0876 & 0.9505 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 200 & 0.0166 & 0.4146 & 0.8975 \\ 
  IRS-Model & Random & Lognormal & 0 & 200 & 0.0074 & 0.4479 & 0.8930 \\ 
  GRTS-Design & Random & Lognormal & 0 & 200 & -0.0044 & 0.4800 & 0.8870 \\ 
  GRTS-Model & Random & Lognormal & 0 & 200 & -0.0063 & 0.4820 & 0.8875 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 200 & -0.0003 & 0.0682 & 0.9735 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 200 & -0.0005 & 0.0664 & 0.9435 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 200 & 0.0002 & 0.0661 & 0.9415 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 200 & 0.0001 & 0.0656 & 0.9450 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0008 & 0.3556 & 0.9379 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0030 & 0.3540 & 0.9144 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0001 & 0.3305 & 0.9029 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.3287 & 0.9049 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 200 & 0.0039 & 0.0771 & 0.9464 \\ 
  IRS-Model & Random & Normal & 0.5 & 200 & 0.0024 & 0.0684 & 0.9489 \\ 
  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0001 & 0.0658 & 0.9374 \\ 
  GRTS-Model & Random & Normal & 0.5 & 200 & -0.0003 & 0.0652 & 0.9444 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 200 & 0.0091 & 0.3861 & 0.9164 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0048 & 0.3783 & 0.9069 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 200 & 0.0012 & 0.4088 & 0.9029 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0005 & 0.3941 & 0.9029 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0002 & 0.0413 & 0.9965 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 200 & -0.0008 & 0.0356 & 0.9475 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0004 & 0.0381 & 0.9475 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0000 & 0.0349 & 0.9490 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0026 & 0.2868 & 0.9835 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0016 & 0.2560 & 0.9154 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0038 & 0.2341 & 0.9309 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0062 & 0.2290 & 0.9239 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 200 & -0.0037 & 0.0673 & 0.9410 \\ 
  IRS-Model & Random & Normal & 0.9 & 200 & -0.0012 & 0.0385 & 0.9505 \\ 
  GRTS-Design & Random & Normal & 0.9 & 200 & 0.0002 & 0.0389 & 0.9370 \\ 
  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0005 & 0.0349 & 0.9435 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}
\begin{table}[ht]
\centering
\begin{tabular}{lllllrrr}
  \hline
Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 200 & 0.0154 & 0.3554 & 0.9313 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 200 & 0.0008 & 0.2657 & 0.9238 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 200 & 0.0058 & 0.2299 & 0.9218 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 200 & 0.0043 & 0.2206 & 0.9147 \\ 
   \hline
\end{tabular}
\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.} 
\end{table}

\end{document}

### 2021-10-21 inst/manuscript/manuscript.Rmd
minor discussion edits.

@@
 <!-- Brief introduction to model based and design based -->
-
-Add sentence about bias.
-Add sentence about user functions.
 
@@
 grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sp_rbind(grts_samp) # Matt you need to install newest version
-# devtools::install_github("USEPA/spsurvey", ref = "main")
+grts_bind <- sp_rbind(grts_samp)
 
@@
 
-The design-based and model-based approaches to inference are fundamentally different paradigms by which to select samples and analyze data.
+The design-based and model-based approaches to inference are fundamentally different paradigms by which to samples are selected and data are analyzed.
 The design-based approach incorporates randomness through sampling to estimate a population parameter.
@@
 Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
-Aside from the theoretical differences described, a few analytical findings were particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); Independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation increasingly than IRS as the strength of spatial correlation increases; the gap in relative efficiency between GRTS-design and GRTS-model widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
+Aside from the theoretical differences described, a few analytical findings are particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation and increasingly better than IRS as the strength of spatial correlation increases; the gap in relative efficiency between GRTS-design and GRTS-model widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
+
+<!-- The design-based approach relies on few assumptions, while the model-based approach relies on rigid distributional ones. -->
 
 There are several benefits and drawbacks of the design-based and model-based approaches for spatial data, some of which we have not yet discussed but are worthy of consideration in future research.
-The design-based approach relies on few assumptions, while the model-based approach relies on rigid distributional ones.
-The Horvitz-Thompson estimator of means and totals is unbiased and computationally efficient, while model-based estimation of covariance parameters can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
+Design-based approaches are often computationally efficient, while model-based estimation of covariance parameters can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
 The design-based approach also more naturally handles binary data, free from the more complicated logistic regression formulation commonly used to handle binary data in a model-based approach.
-The model-based approach, however, can quantify the relationship between covariates (predictor variables) and the response variable, something the design-based approach cannot do naturally.
+The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
 The model-based approach also yields estimated spatial covariance parameters, which help better understand the process of study.

### 2021-10-21 inst/manuscript/manuscript.Rmd
minor updates

@@
   - name: Matt Higham
-    email: mhigham@stlaw.edu
     affiliation: STLAW
@@
 
-journal: "An awesome journal"
+journal: "TBD"
 date: "`r Sys.Date()`"
@@
 
-```{r, results = "asis", eval = FALSE}
-sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
-colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
-sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
-print(sim_one_table, include.rownames = FALSE, comment = FALSE)
-```
-
 The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
@@
 
-**Mike** For design-based, it's really RMSE -- how should we address this?
 Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
@@
 \begin{equation*}
-\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+\text{EFF} = \frac{\text{rMS(P)E of approach}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*}
 
-where rMSPE is the root-mean-squared-prediction error.
+where rMS(P)E is the root-mean-squared error (design-based) or the root-mean-squared-prediction error (model-based).
 When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
@@
 ```
+
+In addition to rMS(P)E and interval coverage, we also recorded average bias.
+The average bias is nearly zero for all approaches in all scenarios, so we omit a visualization of the results here.
+The supplementary material contains tables with mean bias, rMS(P)E, and interval coverage for all 36 simulation scenarios.
 
@@
 
+# Data and Code Availability {.unnumbered}
+
+This manuscript has a supplementary R package that contains all of the data and code used.
+Instructions for download at available at [https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
+
+# Supplementary Material {.unnumbered}
+
+In the supplementary material, we provide summary statistics for all 36 simulation scenarios.
+
 # References {#references .unnumbered}

### 2021-10-21 inst/manuscript/manuscript.tex
minor updates

@@
 
-  \journal{An awesome journal} % Sets Journal name
+  \journal{TBD} % Sets Journal name
 
@@
     \author[STLAW]{Matt Higham}
-   \ead{mhigham@stlaw.edu} 
+
     \author[OSU]{Lisa Madsen}
@@
 \section{Introduction}\label{sec:introduction}}
-
-Add sentence about bias. Add sentence about user functions.
 
@@
 
-\textbf{Mike} For design-based, it's really RMSE -- how should we
-address this? Figure \ref{fig:figeff} shows the relative efficiency of
-the four approaches from Table \ref{tab:designanalysis} with
-``IRS-Design'' as the baseline: \mbox{} \begin{equation*}
-\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
+Figure \ref{fig:figeff} shows the relative efficiency of the four
+approaches from Table \ref{tab:designanalysis} with ``IRS-Design'' as
+the baseline: \mbox{} \begin{equation*}
+\text{EFF} = \frac{\text{rMS(P)E of approach}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*}
 
-where rMSPE is the root-mean-squared-prediction error. When there is no
+where rMS(P)E is the root-mean-squared error (design-based) or the
+root-mean-squared-prediction error (model-based). When there is no
 spatial correlation (top row), the four approaches have approximately
@@
 \end{figure}
+
+In addition to rMS(P)E and interval coverage, we also recorded average
+bias. The average bias is nearly zero for all approaches in all
+scenarios, so we omit a visualization of the results here. The
+supplementary material contains tables with mean bias, rMS(P)E, and
+interval coverage for all 36 simulation scenarios.
 
@@
 The design-based and model-based approaches to inference are
-fundamentally different paradigms by which to select samples and analyze
-data. The design-based approach incorporates randomness through sampling
-to estimate a population parameter. The model-based approach
-incorporates randomness through distributional assumptions to predict
-the realized value of a random variable. Though these approaches have
-often been compared in the literature both from theoretical and
+fundamentally different paradigms by which to samples are selected and
+data are analyzed. The design-based approach incorporates randomness
+through sampling to estimate a population parameter. The model-based
+approach incorporates randomness through distributional assumptions to
+predict the realized value of a random variable. Though these approaches
+have often been compared in the literature both from theoretical and
 analytical perspectives, our contribution lies in studying them in a
@@
 from the theoretical differences described, a few analytical findings
-were particularly notable: the design decision (GRTS vs IRS) seems much
+are particularly notable: the design decision (GRTS vs IRS) seems much
 more important than the analysis decision (design-based vs model-based);
-Independent of the analysis approach, there is no reason to prefer IRS
+independent of the analysis approach, there is no reason to prefer IRS
 over GRTS for spatial data -- GRTS tends to perform at least as well as
-IRS when there is no spatial correlation increasingly than IRS as the
-strength of spatial correlation increases; the gap in relative
-efficiency between GRTS-design and GRTS-model widens as the strength of
-spatial correlation increases; and when the data are skewed, interval
-coverage for all approaches improves both as the sample size increases
-and as the strength of correlation increases.
+IRS when there is no spatial correlation and increasingly better than
+IRS as the strength of spatial correlation increases; the gap in
+relative efficiency between GRTS-design and GRTS-model widens as the
+strength of spatial correlation increases; and when the data are skewed,
+interval coverage for all approaches improves both as the sample size
+increases and as the strength of correlation increases.
 
@@
 model-based approaches for spatial data, some of which we have not yet
-discussed but are worthy of consideration in future research. The
-design-based approach relies on few assumptions, while the model-based
-approach relies on rigid distributional ones. The Horvitz-Thompson
-estimator of means and totals is unbiased and computationally efficient,
-while model-based estimation of covariance parameters can be
-computationally burdensome, especially for likelihood-based methods such
-as REML that rely on inverting a covariance matrix. The design-based
-approach also more naturally handles binary data, free from the more
-complicated logistic regression formulation commonly used to handle
-binary data in a model-based approach. The model-based approach,
-however, can quantify the relationship between covariates (predictor
-variables) and the response variable, something the design-based
-approach cannot do naturally. The model-based approach also yields
-estimated spatial covariance parameters, which help better understand
-the process of study. Model selection is also possible using model-based
-approaches and criteria such as likelihood ratio tests or AIC (Akaike,
-1974). Model-based approaches are capable of more efficient small-area
-estimation than design-based approaches by leveraging distributional
-assumptions in areas with few observed sites. Model-based approaches can
-also compute site-by-site predictions at unobserved locations and use
-them to construct informative visualizations. The benefits and drawbacks
-of both approaches, alongside our theoretical and analytical
-comparisons, should be heavily considered when choosing among them. This
-is especially true from an analysis perspective, as we found that using
-a spatially balanced sampling algorithm benefits both design-based and
-model-based analyses.
+discussed but are worthy of consideration in future research.
+Design-based approaches are often computationally efficient, while
+model-based estimation of covariance parameters can be computationally
+burdensome, especially for likelihood-based methods such as REML that
+rely on inverting a covariance matrix. The design-based approach also
+more naturally handles binary data, free from the more complicated
+logistic regression formulation commonly used to handle binary data in a
+model-based approach. The model-based approach, however, can more
+naturally quantify the relationship between covariates (predictor
+variables) and the response variable. The model-based approach also
+yields estimated spatial covariance parameters, which help better
+understand the process of study. Model selection is also possible using
+model-based approaches and criteria such as likelihood ratio tests or
+AIC (Akaike, 1974). Model-based approaches are capable of more efficient
+small-area estimation than design-based approaches by leveraging
+distributional assumptions in areas with few observed sites. Model-based
+approaches can also compute site-by-site predictions at unobserved
+locations and use them to construct informative visualizations. The
+benefits and drawbacks of both approaches, alongside our theoretical and
+analytical comparisons, should be heavily considered when choosing among
+them. This is especially true from an analysis perspective, as we found
+that using a spatially balanced sampling algorithm benefits both
+design-based and model-based analyses.
+
+\hypertarget{data-and-code-availability}{%
+\section*{Data and Code Availability}\label{data-and-code-availability}}
+\addcontentsline{toc}{section}{Data and Code Availability}
+
+This manuscript has a supplementary R package that contains all of the
+data and code used. Instructions for download at available at
+\url{https://github.com/michaeldumelle/DvMsp}.
+
+\hypertarget{supplementary-material}{%
+\section*{Supplementary Material}\label{supplementary-material}}
+\addcontentsline{toc}{section}{Supplementary Material}
+
+In the supplementary material, we provide summary statistics for all 36
+simulation scenarios.
 

### 2021-10-21 inst/manuscript/supplementary.Rmd
minor updates

@@
     text: "Corresponding Author"
-journal: "An awesome journal"
+journal: "TBD"
 date: "`r Sys.Date()`"

### 2021-10-22 R/sim_trial.R
added vignettes folder.

@@ -43,17 +43,12 @@
 #' @importFrom rlang .data
 #' @export
 
-## library(devtools)
-## install_github("https://github.com/highamm/sptotal")
-## install_github('USEPA/spsurvey', ref = "main")
-## library(sptotal)
-## library(spsurvey)
-## library(dplyr)
 sim_trial <- function(seed = sample.int(1e7, size = 1),
                     N = 100, n = 50, gridded = TRUE,
                     cortype = "Exponential", psill, erange,
-                    nugget, cortype_est = "Exponential", resptype = "normal", ...) {
-  # browser()
+                    nugget, cortype_est = "Exponential",
+                    resptype = "normal", ...) {
+
   ###############################
   ## Data Prep
   ###############################

### 2021-10-22 R/sim_pop.R
added the mean, mu, as a parameter to sim_pop().

@@ -15,6 +15,7 @@
 #' @param nugget is the nugget.
 #' @param resptype the response type. The default is Gaussian errors.
 #' @param ... further arguments passed to or from other methods.
+#' @param mu is the mean of the response and is \code{0} by default.
 #' @return a data frame with \itemize{
 #'   \item \code{x}, a column with the spatial x-coordinates.
 #'   \item \code{y}, a column with the spatial y-coordinates.
@@ -29,7 +30,7 @@
 # need spsurvey and sptotal and dplyr
 sim_pop <- function(N = 100, gridded = TRUE,
                     cortype = "Exponential", psill, erange,
-                    nugget, resptype = "normal", ...) {
+                    nugget, resptype = "normal", mu = 0, ...) {
 
   # simulating the locations
   if (gridded) {
@@ -59,7 +60,7 @@ sim_pop <- function(N = 100, gridded = TRUE,
                   )
   ## simulate response
   chol_covmx <- chol(covmx)
-  response <- as.vector(t(chol_covmx) %*% rnorm(N))
+  response <- mu + as.vector(t(chol_covmx) %*% rnorm(N))
 
   if (resptype == "normal") {
     data$response <- response

### 2021-10-22 inst/manuscript/manuscript.Rmd
added to vignett and got rid of effective range line.

@@
 We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with a  $\sqrt{2} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
-The range was chosen so that the decayed to nearly zero at the largest distance possible in the domain, $\sqrt{2}$, otherwise known as the "effective" range (for the exponential covariance, the effective range is $3\phi$).
+All scenarios used exponential correlation with a  $\sqrt{4} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
 The mean for the spatial process generating the response was set to zero.
+
+<!-- The range was chosen so that the decayed to nearly zero at the largest distance possible in the domain, $\sqrt{2}$, otherwise known as the "effective" range (for the exponential covariance, the effective range is $3\phi$).
+-->
 

### 2021-10-22 R/sim_trial.R

@@ -14,11 +14,11 @@
 #' points with random locations.
 #' @param cortype is the true correlation function, which is
 #' \code{"Exponential"} by default.
+#' @param cortype_est is the (possibly misspecified) correlation
+#' function used to estimate/predict the mean, total, or other quantity.
 #' @param psill is the partial sill.
 #' @param erange is the effective range.
 #' @param nugget is the nugget.
-#' @param cortype_est is the (possibly misspecified) correlation
-#' function used to estimate/predict the mean, total, or other quantity.
 #' @param resptype is the response variable type, either \code{
 #' "normal"} or \code{"lognormal"}.
 #' @param ... further arguments passed to or from other methods.
@@ -33,6 +33,8 @@
 #'   \item \code{lb}, a lower 95% confidence bound
 #'   \item \code{ub}, an upper 95% confidence bound.
 #' }
+#'
+#'
 #' @examples
 #' sim_trial(seed = sample.int(1e7, size = 1), N = 100, n = 50,
 #' gridded = TRUE, cortype = "Exponential", psill = 1, erange = 1,
@@ -47,7 +49,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
                     N = 100, n = 50, gridded = TRUE,
                     cortype = "Exponential", psill, erange,
                     nugget, cortype_est = "Exponential",
-                    resptype = "normal", ...) {
+                    resptype = "normal", mu = 0, ...) {
 
   ###############################
   ## Data Prep

### 2021-10-27 inst/manuscript/manuscript.Rmd
reanalyzing simulation data

@@
 pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                  psill = 0.9, erange = 1, nugget = 0.1)
+                  psill = 0.9, range = 1, nugget = 0.1)
 
@@
 pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
+                      psill = 0.9, range = 1, nugget = 0.1)
 pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                       psill = 0.9, erange = 1, nugget = 0.1)
+                       psill = 0.9, range = 1, nugget = 0.1)
 pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, erange = 1, nugget = 0.1)
+                      psill = 0.9, range = 1, nugget = 0.1)
 
@@
 
-Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
+Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} using the random location layout, where "IRS-Design" is the baseline:
 \mbox{}
@@
 However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
+Additionally, as the sample size increases, IRS-Design performs relatively worse compared to the other approaches.
+These conclusions were similar to those observed when the data were gridded.
 
@@
 combo_data <- combo_data %>%
-  dplyr::filter(n != 10 & gridded == FALSE) %>%
+  dplyr::filter(gridded == FALSE) %>%
   mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
@@
 Justification for the design-based intervals lies in the asymptotic normality of totals via the Central Limit Theorem, and justification for the model-based intervals lies in the normality assumption of the errors.
-Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches.
+Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches in the random location layout.
 All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
@@
 This suggests that the larger the sample size and the stronger the spatial dependence, the more resistant these intervals are to departures from normality of the data.
+These conclusions were similar to those observed when the data were gridded.
 

### 2021-10-27 inst/manuscript/manuscript.tex
reanalyzing simulation data

@@
 \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
-exponential correlation with a \(\sqrt{2} / 3\) for \(N = 900\) response
+exponential correlation with a \(\sqrt{4} / 3\) for \(N = 900\) response
 values simulated on the unit square in either random locations (Layout =
-Random) or gridded locations (Layout = Gridded). The range was chosen so
-that the decayed to nearly zero at the largest distance possible in the
-domain, \(\sqrt{2}\), otherwise known as the ``effective'' range (for
-the exponential covariance, the effective range is \(3\phi\)). The mean
-for the spatial process generating the response was set to zero.
+Random) or gridded locations (Layout = Gridded). The mean for the
+spatial process generating the response was set to zero.
 
@@
 Figure \ref{fig:figeff} shows the relative efficiency of the four
-approaches from Table \ref{tab:designanalysis} with ``IRS-Design'' as
-the baseline: \mbox{} \begin{equation*}
+approaches from Table \ref{tab:designanalysis} using the random location
+layout, where ``IRS-Design'' is the baseline: \mbox{} \begin{equation*}
 \text{EFF} = \frac{\text{rMS(P)E of approach}}{\text{rMS(P)E of IRS-Design}},
@@
 approach when there is a high amount of spatial correlation.
+Additionally, as the sample size increases, IRS-Design performs
+relatively worse compared to the other approaches. These conclusions
+were similar to those observed when the data were gridded.
 
@@
 Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
-the four approaches. All four approaches have somewhat similar interval
-coverage in all settings, with GRTS-Design having slightly lower
-coverage when the response is normal.
+the four approaches in the random location layout. All four approaches
+have somewhat similar interval coverage in all settings, with
+GRTS-Design having slightly lower coverage when the response is normal.
 
@@
 more resistant these intervals are to departures from normality of the
-data.
+data. These conclusions were similar to those observed when the data
+were gridded.
 

### 2021-10-27 R/sim_trial.R

@@ -58,9 +58,6 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   # make an sf object (for spsurvey)
   data_sf <- sf::st_as_sf(data, coords = c("x", "y"), crs = 5070)
 
-  # reshuffle data (for spsurvey irs() but in 5.0.0 -- will be fixed in 5.1.0)
-  data_sf <- data_sf[sample(1:nrow(data_sf)), ]
-
   ###############################
   ## IRS Sample
   ###############################

### 2021-10-27 inst/scripts/sim_runs.R

@@ -7,14 +7,20 @@ write_out <- TRUE
 ## set up grid for simulation settings
 N <- 30 ^ 2
 n <- c(50, 100, 200)
-total_var <- 2
 psill_ratio <- c(0, 0.5, 0.9)
-range <- 2 / 3
+range <- sqrt(2) / 3
 gridded <- c(TRUE, FALSE)
 resptype <- c("normal", "lognormal")
+total_var_norm <- 2
+# find variance that gives lognormal var of 2
+total_var_lognorm <- uniroot(function(x) (exp(x) - 1) * exp(x) - total_var_norm, interval = c(0, 10))$root
 
-parm_df <- expand_grid(N, n, total_var, psill_ratio,
+parm_df <- expand_grid(N, n, psill_ratio,
                        range, gridded, resptype) %>%
+  mutate(total_var = case_when(
+    resptype == "normal" ~ total_var_norm,
+    resptype == "lognormal" ~ total_var_lognorm
+  )) %>%
   mutate(nugget_ratio = 1 - psill_ratio,
          psill = total_var * psill_ratio,
          nugget = total_var * nugget_ratio)
@@ -29,13 +35,13 @@ source("R/sim_trial.R")
 ## loop through each row of parm_df (could be done with purrr::pmap() instead)
 for (i in 1:nrow(parm_df)) {
   n_trials <- 2000
-  seed <- sample.int(1e7, size = n_trials)
+  seed <- 1:n_trials
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
   clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
-  clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster
+  clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster v >= 5.1.0
   clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
-  clusterEvalQ(cluster, library(dplyr)) # export sptotal to cluster
+  clusterEvalQ(cluster, library(dplyr)) # export dplyr to cluster
 
   parm_df_sim <- parm_df %>% slice(i) ## grab row for ith parameter combo
   sim_output <- parLapply(

### 2021-10-27 inst/manuscript/manuscript.Rmd
minor manuscript updates

@@
 For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
-A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation.
+A total variance of 0.6931 and a mean of 0 on the normal scale yielded a total variance of 2 and a mean of 1.414 after exponentiation.
 Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
@@
 #       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.")
 align(parm_xtab) <- "r|lll"

### 2021-10-27 inst/manuscript/manuscript.tex
minor manuscript updates

@@
 the specified correlation parameters using a normal distribution and
-were subsequently exponentiated. A total variance of 2 and a mean of 0
-on the normal scale is equivalent to a total variance of 47 and a mean
-of 2.72 after exponentiation. Therefore, when the model-based methods
-were used for lognormal response, the correlation was mis-specified. We
-chose to simulate values with a lognormal distribution so that we could
-test the model-based analysis approach with a mis-specified model and so
-that we could test both analysis approaches on data that exhibits a
-large amount of skewness.
+were subsequently exponentiated. A total variance of 0.6931 and a mean
+of 0 on the normal scale yielded a total variance of 2 and a mean of
+1.414 after exponentiation. Therefore, when the model-based methods were
+used for lognormal response, the correlation was mis-specified. We chose
+to simulate values with a lognormal distribution so that we could test
+the model-based analysis approach with a mis-specified model and so that
+we could test both analysis approaches on data that exhibits a large
+amount of skewness.
 
@@
 \end{tabular}
-\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.} 
+\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.} 
 \end{table}

### 2021-10-27 inst/manuscript/supplementary.Rmd
reanalyzing simulation data
minor manuscript updates

@@
 ```{r, results = "asis"}
-for (i in 13:48) {
+for (i in 1:36) {
   make_sim_table(i)

### 2021-11-02 inst/scripts/sim_runs.R
Updated vignette to be functional and generate summary tables.

@@ -65,7 +65,7 @@ for (i in 1:nrow(parm_df)) {
   #   bind_rows() %>%
   #   nrow()
 
-  # do summaries
+  # do summaries, could be done with a for loop instead ;)
   safe_output <- map(sim_output, "result") %>%
     bind_rows()
 

### 2021-11-02 R/sim_pop.R
more minor updates

@@ -85,28 +85,25 @@ sim_pop <- function(N = 100, gridded = TRUE,
 
 #' Spatial Correlation Models
 #'
-#' Note that, currently, only one of these models is implemented
-#' in this package: \code{covmx_exp()}.
-#'
 #' @param distmx is the distance matrix for sampled sites.
 #' @param psill is the partial sill.
 #' @param range is the effective range.
 #' @param nugget is the nugget effect.
 #' @return Covariance Matrix
-#' @export
+#' @noRd
 
-#' @describeIn covmx_exp Exponential Covariance
+# Exponential Covariance
 covmx_exp <- function(distmx, psill, range, nugget) {
   psill * exp(- distmx / range) + nugget * (distmx == 0)
 }
 
-#' @describeIn covmx_exp Gaussian Covariance
+# Gaussian Covariance
 covmx_gauss <- function(distmx, psill, range, nugget) {
   psill * exp(- distmx^2 / range^2) +
     nugget * (distmx == 0)
 }
 
-#' @describeIn covmx_exp Spherical Covariance
+# Spherical Covariance
 covmx_sph <- function(distmx, psill, range, nugget) {
   distratio <- distmx / range
   psill * (1 - (3 / 2) * distratio + (1 / 2) * distratio^3) + nugget * (distmx == 0)

### 2021-11-11 inst/manuscript/manuscript.Rmd
section 1 and 2 jay updates

@@
   - name: Michael Dumelle
-    email: Dumelle.Michael@epa.gov
     affiliation: USEPA
@@
     affiliation: STLAW
+  - name: Jay M. Ver Hoef
+    affiliation: NOAA
+  - name: Anthony R. Olsen
+    affiliation: USEPA
   - name: Lisa Madsen
     affiliation: OSU
-  - name: Anthony R. Olsen
-    affiliation: USEPA
-  - name: Jay M. Ver Hoef
-    affiliation: NOAA
 address:
@@
     address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
+  - code: NOAA
+    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
   - code: OSU
     address: Oregon State University Department of Statistics, 239 Weniger Hall, Corvallis, Oregon, 97331
-  - code: NOAA
-    address: Marine Mammal Laboratory, Alaska Fisheries Science Center, National Oceanic and Atmospheric Administration, Seattle, Washington, 98115
 footnote:
-  - code: 1
-    text: "Corresponding Author"
+    code: 1
+    text: "Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)"
 abstract: |
@@
 
-journal: "TBD"
+journal: "Methods in Ecology and Evolution"
 date: "`r Sys.Date()`"
@@
 ---
-
-_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_
-
-Potential Journals:
-
-* Ecological Applications
-* Methods in Ecology and Evolution
-* Journal of Applied Ecology
-* Environmetrics
-* Environmental and Ecological Statistics
-
-<!-- Notes from today -->
-<!-- •	Finite populations require a smaller scope but we need to discuss the impact of infinite populations as well – put finite vs infinite population context in the introduction / similarities and differences -->
-<!-- •	What should a lower sample size be? 50 seems reasonable 30 may be too small – maybe 50, 100, 200 -->
-<!-- •	IRS results seem off for dependent error sample size 50 -->
-<!-- •	Take home points – don’t have to do GRTS design with GRTS sample -->
-<!-- •	Take home points – GRTS is better for both model and design, while spatial approach for design is close to spatial model -->
-<!-- •	Add medium dependence -->
-<!-- •	Remind readers what choices of the four approaches in the application mean – what do they mean to practitioners -->
 
@@
 
-There are two general approaches for using data to make statistical inferences about a population: design-based and model-based.
-When data cannot be obtained for all units in a population (population units), data on a subset of the population units is collected and called a sample.
-In the design-based approach, inferences about the underlying population are informed from a probabilistic process in which population units are selected to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process that generated the data.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of general advantages [@hansen1983evaluation].
+There are two general approaches for using data to make frequentist statistical inferences about a population: design-based and model-based.
+When data cannot be collected for all units in a population (population units), data are collected on a subset of the population units.
+This subset is called a sample.
+In the design-based approach, inferences about the underlying population are informed via a probabilistic process assigning some population units to the sample.
+Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process generating the data.
+Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation].
 
@@
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Thereafter, several comparisons between design-based and model-based for spatial data have been considered [@brus1997random; @verhoef2002sampling; @verhoef2008spatial].
-@cooper2006sampling review the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g.
-@cicchitelli2012model, @chan2020bayesian for a Bayesian approach, and @sterba2009alternative).
-More recent overviews include @brus2020statistical and @wang2012review.
+Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @wang2012review; @brus2020statistical].
+@cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
+In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g., @sterba2009alternative, @cicchitelli2012model, @chan2020bayesian for a Bayesian approach).
 
@@
 In this manuscript, we compare design-based approaches that incorporate spatial locations to model-based approaches for spatial data.
-Though these comparisons generalize to both finite populations (e.g.
-point resources) and infinite populations (e.g.
-linear and areal resources), we focus on applications to finite populations.
+We focus on finite populations, but these comparisons generalize to infinite populations as well.
+A finite population contains a finite number of population units; an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
+An infinite population contains an infinite number of population units; an example is locations within a single lake.
+
 The rest of the manuscript is organized as follows.
-In Section \ref{sec:background}, we compare sampling and estimation procedures between the design-based approach and the model-based approach for spatial data.
+In Section \ref{sec:background}, we introduce and compare several sampling and estimation procedures of the design-based and model-based approaches for finite population spatial data.
 In Section \ref{sec:numstudy}, we use a simulation approach to study the behavior and performance of both approaches.
-In Section \ref{application}, we use both approaches to analyze real data.
+In Section \ref{application}, we use both approaches to analyze real data consisting of mercury concentration from lakes in the contiguous United States.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 In this section, we describe the role of randomness and its effects on subsequent inferences.
-We then discuss specific inference methods for the design-based and model-based approaches for spatial data.
+We then discuss specific inference methods of the approaches for spatial data.
 
@@
 The design-based approach assumes the population is fixed.
-Randomness is incorporated via the selection of units in a sampling frame according to a sampling design.
+Randomness is incorporated via the selection of units in a smapling frame according to a sampling design.
 A sampling frame is the set of all units available to be sampled.
 A sampling design assigns a positive probability of inclusion (inclusion probability) to each unit in the sampling frame.
-Some examples of commonly used sampling designs include simple random sampling, stratified random sample, and cluster sampling.
-These sampling designs tend to select units from the sampling frame independently of other units, so we call them "Independent Random Sampling" (IRS) designs.
-Sampling designs incorporating the spatial locations of units in the sample frame are called spatially balanced designs.
+Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
+If a sampling design selects units from the sampling frame while ignoring their spatial locations, we call them "Independent Random Sampling" (IRS) designs.
+If a sampling design selects units from the sampling frame while incorporating their spatial locations, we call them spatially balanced designs.
 Spatially balanced designs can be obtained using the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
-The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate parameters (e.g.
-means and totals) of a population.
-Generally, these population parameters are assumed to be fixed, unknown constants.
+The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate fixed, unknown parameters (e.g., means and totals) of a population.
 
@@
 
-The model-based approach assumes the data are a random realization of a data-generating process.
+The model-based approach assumes the data are a random realization of a data-generating stochastic process.
 Randomness is incorporated through distributional assumptions on this process.
-Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
+Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling.
+Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
 To guard against preferential sampling, model-based approaches often still implement random sampling.
 
-Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
+Instead of estimating fixed but unknown parameters like a mean or total (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
 For example, suppose the realized mean of all population units is the value of interest.
 Instead of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-Prediction intervals are then derived leveraging assumptions of the data generating process.
-If we repeatedly generate the response values from a fixed spatial process and obtained a sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
+Prediction intervals are then derived using assumptions of the data generating process.
+If we repeatedly generate the response values from the same spatial process and sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@brus2020statistical provides a similar figure).
+A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@verhoef2002sampling and @brus2020statistical provide similar figures).
 
@@
 Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of n = 4.
+In the top row, we have one fixed population, and three random samples of size four.
 The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
@@
 
-Spatially balanced samples can be obtained using the design-based approach.
+The design-based approach can be used to select samples that are "well-spread" in space, or spatially balanced.
 Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
@@
 To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voroni polygons.
-Since GRTS was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we use Generalized Random Tessellation Stratified (GRTS) sampling because it has several attractive properties: GRTS sampling accommodates finite and infinite sampling frames; accommodates equal, unequal, and proportional (to) size inclusion probabilities; accommodates legacy (historical) sampling; accommodates a minimum distance between units in a sample; accommodates reverse hierarchically ordered replacement units in a sample (replacement units are units available to be sampled if an original unit cannot be sampled); and is available in the spsurvey R package @dumelle2021spsurvey.
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
+In this manuscript, we use the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced samples sampling because the algorithm has several attractive properties.
+It accommodates finite and infinite sampling frames.
+It accommodates equal, unequal, and proportional (to size) inclusion probabilities.
+It accommodates legacy (historical) sampling [@foster2017spatially].
+It accommodates a minimum distance between units in a sample.
+It accommodates replacement units in a sample, which are units that can be sampled in place of an original unit that can no longer be sampled.
+Lastly, it can be implemented using the spsurvey R package @dumelle2021spsurvey.
 
@@
 A systematic sample is conducted on the line and linked back to a unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially and @dumelle2021spsurvey provide further details.
+@stevens2004spatially provides further details.
 
@@
 
-After selecting a GRTS sample, data are collected and used to estimate population parameters.
+After selecting a spatially balanced sample using the GRTS algorithm (i.e., a GRTS sample), data are collected and used to estimate population parameters.
 To unbiasedly estimate population means and totals from sample data, one can use the Horvitz-Thompson estimator [@horvitz1952generalization].
@@
 where $Z_i$ is the value of the $i$th unit in the sample and $\pi_i$ is the inclusion probability of the $i$th unit in the sample.
-An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by the population size.
+An estimate of the population mean can be obtained by dividing $\hat{\tau}_{ht}$ by number of population units.
 
 While the Horvitz-Thompson estimator is unbiased for population means and totals, it is also important to quantify the uncertainty in these estimates.
-@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but they have two drawbacks.
-First, these estimators rely on calculating $\pi_{ij}$, the probability that unit $i$ and unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
+@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
+First, they rely on calculating $\pi_{ij}$, the probability that unit $i$ and unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
 Second, these estimators ignore the spatial locations of the units in the sampling frame.
@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neighborhood variance estimator tends reduce $\text{Var}(\hat{\tau})$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
+@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
 
@@
 Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
-Instead of basing inference off of a specific sampling design, we assume the data are generated by a spatial process.
+Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial process.
 @verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
@@
 .
-, \text{z}(s_N) \}}$ be a response vector at locations $s_1$, $s_2$, .
+, \text{z}(s_N) \}}$ be an $N \times 1$ response vector at locations $s_1$, $s_2$, .
 .
 .
-, $s_N$ that can be measured at the $N$ population units and is represented as an $N \times 1$ vector.
+, $s_N$ that can be measured at the $N$ population units.
 Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
@@
 
-<!-- how explicitly do we want to define \mathbf{z} -->
-
-<!-- finite number of -->
-
-<!-- confusing having two tau's: tau for total and a tau as a function -->
-
 We often only have a sample of the $N$ population units.
-Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u},
+Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with subscript \emph{u}, let
 
@@
 \end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively; $\beta$ is the parameter vector of fixed effects; and $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
-Denoting $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, we assume the expectation of $\bm{\delta}$ equals $\mathbf{0}$.
+where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, and $\bm{\beta}$ is the parameter vector of fixed effects.
 
-In addition to assuming the expectation of $\bm{\delta}$ equals $\mathbf{0}$, we also assume that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
+Let $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for teh sampled and unsampled population units, respectively.
+We assume $E(\bm{\delta} = \mathbf{0}$ and that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
 It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th entry for $\cov(\bm{\delta})$ is
+Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th elment of hte matrix $\cov(\bm{\delta})$ is
 \mbox{}
@@
 \begin{cases} 
-\sigma^2_{ps}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
-\sigma^2_{ps} + \sigma^2_n & h_{i,j} = 0
+\sigma^2_{1}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
+\sigma^2_{1} + \sigma^2_2 & h_{i,j} = 0
 \end{cases}
@@
 \end{align}
-where $\sigma^2_{ps}$ is the partial sill measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the nugget measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-Any spatial covariance function could be used in the place of the exponential, however, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
+where $\sigma^2_{1}$ is dependent random error variance measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the independent random error variance measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the correlation, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+Often $\sigma^2_{1}$ and $\sigma^2_{2}$ are called the partial sill and nugget, respectively.
+Any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
 80-93].
@@
 With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based, meaning they don't rely on any distributional assumptions.
+While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
 
 We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-However, k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, can also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
-We choose to use FPBK because it is faster than a Bayesian approach and random forest and because @ver2013comparison showed that the method outperforms k-nearest-neighbors in many scenarios.
+Other methods, such as k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
+We choose to use FPBK because it is faster than a Bayesian approach and it was developed with theoretically-based variance estimators of means and totals for spatial data, whereas random forests and k-nearest-neighbors use ad-hoc variance estimators in most cases [@ver2013comparison]; additionally, FBPK outperformed the other methods in most scenarios.
 
@@
 library(here)
-library(SpatialDesignVModel)
+library(DvMsp)
 data(nla_df) # may need to reinstall package

### 2021-11-11 inst/manuscript/manuscript.tex
section 1 and 2 jay updates

@@
 
-  \journal{TBD} % Sets Journal name
+  \journal{Methods in Ecology and Evolution} % Sets Journal name
 
@@
 \usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 
-  \title{A comparison of design-based and model-based approaches for finite
-population spatial data.}
+  \title{A comparison of design-based and model-based approaches for
+finite population spatial data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
-   \ead{Dumelle.Michael@epa.gov} 
+
     \author[STLAW]{Matt Higham}
+
+    \author[NOAA]{Jay M. Ver Hoef}
+
+    \author[USEPA]{Anthony R. Olsen}
 
@@
 
-    \author[USEPA]{Anthony R. Olsen}
-
-    \author[NOAA]{Jay M. Ver Hoef}
-
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
-and Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
-      \cortext[1]{Corresponding Author}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Mathematics,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+      \cortext[1]{Corresponding Author: Michael Dumelle
+(Dumelle.Michael@epa.gov)}
 
@@
 
-\emph{Text based on elsarticle sample manuscript, see
-\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}
-
-Potential Journals:
-
-\begin{itemize}
-\tightlist
-\item
-  Ecological Applications
-\item
-  Methods in Ecology and Evolution
-\item
-  Journal of Applied Ecology
-\item
-  Environmetrics
-\item
-  Environmental and Ecological Statistics
-\end{itemize}
-
 \hypertarget{sec:introduction}{%
@@
 
-There are two general approaches for using data to make statistical
-inferences about a population: design-based and model-based. When data
-cannot be obtained for all units in a population (population units),
-data on a subset of the population units is collected and called a
-sample. In the design-based approach, inferences about the underlying
-population are informed from a probabilistic process in which population
-units are selected to be in the sample. Alternatively, in the
+There are two general approaches for using data to make frequentist
+statistical inferences about a population: design-based and model-based.
+When data cannot be collected for all units in a population (population
+units), data are collected on a subset of the population units. This
+subset is called a sample. In the design-based approach, inferences
+about the underlying population are informed via a probabilistic process
+assigning some population units to the sample. Alternatively, in the
 model-based approach, inferences are made from specific assumptions
-about the underlying process that generated the data. Each paradigm has
-a deep historical context (Sterba, 2009) and its own set of general
-advantages (Hansen et al., 1983).
+about the underlying process generating the data. Each paradigm has a
+deep historical context (Sterba, 2009) and its own set of benefits and
+drawbacks (Hansen et al., 1983).
 
@@
 quashing the belief that design-based approaches could not be used for
-spatially correlated data. Thereafter, several comparisons between
-design-based and model-based for spatial data have been considered (Brus
-and De Gruijter, 1997; Ver Hoef, 2002, 2008). Cooper (2006) review the
-two approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. In addition to Cooper (2006), there has been substantial
-research and development into estimators that use both design and
-model-based principles (see e.g.~Cicchitelli and Montanari (2012),
-Chan-Golston et al. (2020) for a Bayesian approach, and Sterba (2009)).
-More recent overviews include Brus (2020) and Wang et al. (2012).
+spatially correlated data. Since then, there have been several general
+comparisons between design-based and model-based approaches for spatial
+data (Brus and De Gruijter, 1997; Brus, 2020; Ver Hoef, 2002; Ver Hoef,
+2008; Wang et al., 2012). Cooper (2006) reviews the two approaches in an
+ecological context before introducing a ``model-assisted'' variance
+estimator that combines aspects from each approach. In addition to
+Cooper (2006), there has been substantial research and development into
+estimators that use both design and model-based principles (see e.g.,
+Sterba (2009), Cicchitelli and Montanari (2012), Chan-Golston et al.
+(2020) for a Bayesian approach).
 
@@
 approaches that incorporate spatial locations to model-based approaches
-for spatial data. Though these comparisons generalize to both finite
-populations (e.g.~point resources) and infinite populations (e.g.~linear
-and areal resources), we focus on applications to finite populations.
+for spatial data. We focus on finite populations, but these comparisons
+generalize to infinite populations as well. A finite population contains
+a finite number of population units; an example is lakes (treated as a
+whole with the lake centroid representing location) in the contiguous
+United States. An infinite population contains an infinite number of
+population units; an example is locations within a single lake.
+
 The rest of the manuscript is organized as follows. In Section
-\ref{sec:background}, we compare sampling and estimation procedures
-between the design-based approach and the model-based approach for
-spatial data. In Section \ref{sec:numstudy}, we use a simulation
-approach to study the behavior and performance of both approaches. In
-Section \ref{application}, we use both approaches to analyze real data.
-And in Section \ref{sec:discussion}, we end with a discussion and
-provide directions for future research.
+\ref{sec:background}, we introduce and compare several sampling and
+estimation procedures of the design-based and model-based approaches for
+finite population spatial data. In Section \ref{sec:numstudy}, we use a
+simulation approach to study the behavior and performance of both
+approaches. In Section \ref{application}, we use both approaches to
+analyze real data consisting of mercury concentration from lakes in the
+contiguous United States. And in Section \ref{sec:discussion}, we end
+with a discussion and provide directions for future research.
 
@@
 randomness and its effects on subsequent inferences. We then discuss
-specific inference methods for the design-based and model-based
-approaches for spatial data.
+specific inference methods of the approaches for spatial data.
 
@@
 The design-based approach assumes the population is fixed. Randomness is
-incorporated via the selection of units in a sampling frame according to
+incorporated via the selection of units in a smapling frame according to
 a sampling design. A sampling frame is the set of all units available to
@@
 Some examples of commonly used sampling designs include simple random
-sampling, stratified random sample, and cluster sampling. These sampling
-designs tend to select units from the sampling frame independently of
-other units, so we call them ``Independent Random Sampling'' (IRS)
-designs. Sampling designs incorporating the spatial locations of units
-in the sample frame are called spatially balanced designs. Spatially
-balanced designs can be obtained using the Generalized Random
-Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004),
-which we discuss in more detail in Section \ref{subsec:spb_design}. The
-design-based approach combines the randomness of the sampling design and
-the data collected via the sample to estimate parameters (e.g.~means and
-totals) of a population. Generally, these population parameters are
-assumed to be fixed, unknown constants.
+sampling, stratified random sampling, and cluster sampling. If a
+sampling design selects units from the sampling frame while ignoring
+their spatial locations, we call them ``Independent Random Sampling''
+(IRS) designs. If a sampling design selects units from the sampling
+frame while incorporating their spatial locations, we call them
+spatially balanced designs. Spatially balanced designs can be obtained
+using the Generalized Random Tessellation Stratified (GRTS) algorithm
+(Stevens and Olsen, 2004), which we discuss in more detail in Section
+\ref{subsec:spb_design}. The design-based approach combines the
+randomness of the sampling design and the data collected via the sample
+to estimate fixed, unknown parameters (e.g., means and totals) of a
+population.
 
@@
 The model-based approach assumes the data are a random realization of a
-data-generating process. Randomness is incorporated through
+data-generating stochastic process. Randomness is incorporated through
 distributional assumptions on this process. Strictly speaking,
 randomness need not be incorporated through random sampling, though
-Diggle et al. (2010) warn against preferential sampling Preferential
+Diggle et al. (2010) warn against preferential sampling. Preferential
 sampling occurs when the process generating the data locations and the
@@
 
-Instead of estimating fixed but unknown parameters (as in the
-design-based approach), the goal of model-based inference in the spatial
-context is often to predict a realized variable, or value. For example,
-suppose the realized mean of all population units is the value of
-interest. Instead of \emph{estimating} a fixed, unknown mean, we are
-\emph{predicting} the value of the mean, a random variable. Prediction
-intervals are then derived leveraging assumptions of the data generating
-process. If we repeatedly generate the response values from a fixed
-spatial process and obtained a sample, then 95\% of all 95\% prediction
+Instead of estimating fixed but unknown parameters like a mean or total
+(as in the design-based approach), the goal of model-based inference in
+the spatial context is often to predict a realized variable, or value.
+For example, suppose the realized mean of all population units is the
+value of interest. Instead of \emph{estimating} a fixed, unknown mean,
+we are \emph{predicting} the value of the mean, a random variable.
+Prediction intervals are then derived using assumptions of the data
+generating process. If we repeatedly generate the response values from
+the same spatial process and sample, then 95\% of all 95\% prediction
 intervals constructed from a procedure with appropriate coverage will
@@
 approaches for spatial data. A visual comparison of the design-based and
-model-based assumptions is provided in Figure \ref{fig:fig1} (Brus
-(2020) provides a similar figure).
+model-based assumptions is provided in Figure \ref{fig:fig1} (Ver Hoef
+(2002) and Brus (2020) provide similar figures).
 
@@
 
-Spatially balanced samples can be obtained using the design-based
-approach. Spatially balanced samples are useful because parameter
-estimates from these samples tend to vary less than parameter estimates
-from samples that are not spatially balanced (Barabesi and Franceschi,
-2011; Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson
-et al., 2013; Stevens and Olsen, 2004; Wang et al., 2013). The first
-spatially balanced sampling algorithm that saw widespread use was the
-Generalized Random Tessellation Stratified (GRTS) algorithm (Stevens and
-Olsen, 2004). To quantify the spatial balance of a sample, Stevens and
-Olsen (2004) proposed loss metrics based on Voroni polygons. Since GRTS
+The design-based approach can be used to select samples that are
+``well-spread'' in space, or spatially balanced. Spatially balanced
+samples are useful because parameter estimates from these samples tend
+to vary less than parameter estimates from samples that are not
+spatially balanced (Barabesi and Franceschi, 2011; Benedetti et al.,
+2017; Grafström and Lundström, 2013; Robertson et al., 2013; Stevens and
+Olsen, 2004; Wang et al., 2013). The first spatially balanced sampling
+algorithm that saw widespread use was the Generalized Random
+Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004). To
+quantify the spatial balance of a sample, Stevens and Olsen (2004)
+proposed loss metrics based on Voroni polygons. After the GRTS algorithm
 was developed, several other spatially balanced sampling algorithms have
@@
 and Halton Iterative Partitioning Sampling (Robertson et al., 2018). In
-this manuscript, we use Generalized Random Tessellation Stratified
-(GRTS) sampling because it has several attractive properties: GRTS
-sampling accommodates finite and infinite sampling frames; accommodates
-equal, unequal, and proportional (to) size inclusion probabilities;
-accommodates legacy (historical) sampling; accommodates a minimum
-distance between units in a sample; accommodates reverse hierarchically
-ordered replacement units in a sample (replacement units are units
-available to be sampled if an original unit cannot be sampled); and is
-available in the spsurvey R package Dumelle et al. (2021).
+this manuscript, we use the Generalized Random Tessellation Stratified
+(GRTS) algorithm to select spatially balanced samples sampling because
+the algorithm has several attractive properties. It accommodates finite
+and infinite sampling frames. It accommodates equal, unequal, and
+proportional (to size) inclusion probabilities. It accommodates legacy
+(historical) sampling (Foster et al., 2017). It accommodates a minimum
+distance between units in a sample. It accommodates replacement units in
+a sample, which are units that can be sampled in place of an original
+unit that can no longer be sampled. Lastly, it can be implemented using
+the spsurvey R package Dumelle et al. (2021).
 
@@
 two-dimensional space, which results in the desired sample. Stevens and
-Olsen (2004) and Dumelle et al. (2021) provide further details.
+Olsen (2004) provides further details.
 
-After selecting a GRTS sample, data are collected and used to estimate
+After selecting a spatially balanced sample using the GRTS algorithm
+(i.e., a GRTS sample), data are collected and used to estimate
 population parameters. To unbiasedly estimate population means and
@@
 and \(\pi_i\) is the inclusion probability of the \(i\)th unit in the
-sample. An estimate of the population mean is obtained by dividing
-\(\hat{\tau}_{ht}\) by the population size.
+sample. An estimate of the population mean can be obtained by dividing
+\(\hat{\tau}_{ht}\) by number of population units.
 
@@
 estimates. Horvitz and Thompson (1952) and Sen (1953) provide variance
-estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
-these estimators rely on calculating \(\pi_{ij}\), the probability that
-unit \(i\) and unit \(j\) are both in the sample -- this quantity can be
-challenging if not impossible to calculate analytically. Second, these
-estimators ignore the spatial locations of the units in the sampling
-frame. To address these two drawbacks simultaneously, Stevens and Olsen
-(2003) proposed the local neighborhood variance estimator. The local
-neighborhood variance estimator does not rely on \(\pi_{ij}\) and
+estimators for \(\hat{\tau}_{ht}\), but these estimators have two
+drawbacks. First, they rely on calculating \(\pi_{ij}\), the probability
+that unit \(i\) and unit \(j\) are both in the sample -- this quantity
+can be challenging if not impossible to calculate analytically. Second,
+these estimators ignore the spatial locations of the units in the
+sampling frame. To address these two drawbacks simultaneously, Stevens
+and Olsen (2003) proposed the local neighborhood variance estimator. The
+local neighborhood variance estimator does not rely on \(\pi_{ij}\) and
 incorporates spatial locations -- for technical details see Stevens and
 Olsen (2003). Stevens and Olsen (2003) show the local neighborhood
-variance estimator tends reduce \(\text{Var}(\hat{\tau})\) compared to
-variance estimators ignoring spatial locations, yielding narrower
-confidence intervals for \(\tau\).
+variance estimator tends to reduce the estimated variance of
+\(\hat{\tau}\) compared to variance estimators ignoring spatial
+locations, yielding narrower confidence intervals for \(\tau\).
 
@@
 expands the geostatistical Kriging framework to the finite population
-setting (Ver Hoef, 2008). Instead of basing inference off of a specific
-sampling design, we assume the data are generated by a spatial process.
-Ver Hoef (2008) gives details on the theory of FPBK, but some of the
-basic principles are summarized below. Let
+setting (Ver Hoef, 2008). Instead of developing inference based on a
+specific sampling design, we assume the data are generated by a spatial
+process. Ver Hoef (2008) gives details on the theory of FPBK, but some
+of the basic principles are summarized below. Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
-be a response vector at locations \(s_1\), \(s_2\), . . . , \(s_N\) that
-can be measured at the \(N\) population units and is represented as an
-\(N \times 1\) vector. Suppose we want to predict some linear function
-of the response variable,
+be an \(N \times 1\) response vector at locations \(s_1\), \(s_2\), . .
+. , \(s_N\) that can be measured at the \(N\) population units. Suppose
+we want to predict some linear function of the response variable,
 \(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
@@
 subscript \emph{s} and quantities that are part of the unsampled
-population units with a subscript \emph{u},
+population units with subscript \emph{u}, let
 
@@
 design matrices for the sampled and unsampled population units,
-respectively; \(\beta\) is the parameter vector of fixed effects; and
-\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
-sampled and unsampled population units, respectively. Denoting
-\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
-the expectation of \(\bm{\delta}\) equals \(\mathbf{0}\).
+respectively, and \(\bm{\beta}\) is the parameter vector of fixed
+effects.
 
-In addition to assuming the expectation of \(\bm{\delta}\) equals
-\(\mathbf{0}\), we also assume that there is spatial correlation in
+Let \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
+\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for teh
+sampled and unsampled population units, respectively. We assume
+\(E(\bm{\delta} = \mathbf{0}\) and that there is spatial correlation in
 \(\bm{\delta}\) that can be modeled using a covariance function. It is
@@
 the simulations and applications in this manuscript is the exponential
-covariance function: the \(i,j\)th entry for
+covariance function: the \(i,j\)th elment of hte matrix
 \(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
@@
 \begin{cases} 
-\sigma^2_{ps}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
-\sigma^2_{ps} + \sigma^2_n & h_{i,j} = 0
+\sigma^2_{1}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
+\sigma^2_{1} + \sigma^2_2 & h_{i,j} = 0
 \end{cases}
 ,
-\end{align} where \(\sigma^2_{ps}\) is the partial sill measuring
-coarse-scale (correlated) variability, \(\sigma^2_{n}\) is the nugget
-measuring fine-scale (independent) variability, \(\phi\) is the range
-parameter measuring the distance-decay rate of the covariance, and
-\(h_{i,j}\) is the Euclidean distance between population units \(i\) and
-\(j\). Any spatial covariance function could be used in the place of the
-exponential, however, including functions that allow for
+\end{align} where \(\sigma^2_{1}\) is dependent random error variance
+measuring coarse-scale (correlated) variability, \(\sigma^2_{n}\) is the
+independent random error variance measuring fine-scale (independent)
+variability, \(\phi\) is the range parameter measuring the
+distance-decay rate of the correlation, and \(h_{i,j}\) is the Euclidean
+distance between population units \(i\) and \(j\). Often
+\(\sigma^2_{1}\) and \(\sigma^2_{2}\) are called the partial sill and
+nugget, respectively. Any spatial covariance function could be used in
+the place of the exponential, including functions that allow for
 non-stationarity or anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
@@
 note here that the predictor and its variance are both moment-based,
-meaning they don't rely on any distributional assumptions.
+meaning that they do not rely on any distributional assumptions.
 
 We note that we only use FPBK in this paper in order to focus more on
-comparing the design-based and model-based approaches. However,
-k-nearest-neighbors (Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013),
-random forest (Breiman, 2001), Bayesian models (Chan-Golston et al.,
-2020), among others, can also be used to obtain predictions for a mean
-or total from spatially correlated responses of a finite population. We
-choose to use FPBK because it is faster than a Bayesian approach and
-random forest and because Ver Hoef and Temesgen (2013) showed that the
-method outperforms k-nearest-neighbors in many scenarios.
+comparing the design-based and model-based approaches. Other methods,
+such as k-nearest-neighbors (Fix and Hodges, 1951; Ver Hoef and
+Temesgen, 2013), random forest (Breiman, 2001), Bayesian models
+(Chan-Golston et al., 2020), among others, could also be used to obtain
+predictions for a mean or total from spatially correlated responses of a
+finite population. We choose to use FPBK because it is faster than a
+Bayesian approach and it was developed with theoretically-based variance
+estimators of means and totals for spatial data, whereas random forests
+and k-nearest-neighbors use ad-hoc variance estimators in most cases
+(Ver Hoef and Temesgen, 2013); additionally, FBPK outperformed the other
+methods in most scenarios.
 
@@
 \leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
-Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
-under preferential sampling. Journal of the Royal Statistical Society:
-Series C (Applied Statistics) 59, 191--232.
+Diggle, P.J., Menezes, R., Su, T., 2010. Geostatistical inference under
+preferential sampling. Journal of the Royal Statistical Society: Series
+C (Applied Statistics) 59, 191--232.
 
 \leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
-Dumelle, M., Olsen, A.R., Kincaid, T., Weber, M., 2021. Selecting and
-analyzing spatial probability samples in r using spsurvey. Manuscript
-Submitted for Publication.
+Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2021. Spsurvey:
+Spatial sampling design and analysis.
 
@@
 Medicine.
+
+\leavevmode\hypertarget{ref-foster2017spatially}{}%
+Foster, S.D., Hosack, G.R., Lawrence, E., Przeslawski, R., Hedge, P.,
+Caley, M.J., Barrett, N.S., Williams, A., Li, J., Lynch, T., others,
+2017. Spatially balanced designs that incorporate legacy sites. Methods
+in Ecology and Evolution 8, 1433--1442.
 

### 2021-11-11 inst/manuscript/manuscript.Rmd
minor updates

@@
 It accommodates a minimum distance between units in a sample.
-It accommodates replacement units in a sample, which are units that can be sampled in place of an original unit that can no longer be sampled.
-Lastly, it can be implemented using the spsurvey R package @dumelle2021spsurvey.
-
+Lastly, it accommodates replacement units in a sample, which are units that can be sampled in place of an original unit that can no longer be sampled.
 The GRTS algorithm samples from finite and infinite populations by utilizing a mapping between two-dimensional and one-dimensional space.
@@
 \end{align}
-where $\sigma^2_{1}$ is dependent random error variance measuring coarse-scale (correlated) variability, $\sigma^2_{n}$ is the independent random error variance measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the correlation, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+where $\sigma^2_{1}$ is dependent random error variance measuring coarse-scale (correlated) variability, $\sigma^2_{2}$ is the independent random error variance measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the correlation, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
 Often $\sigma^2_{1}$ and $\sigma^2_{2}$ are called the partial sill and nugget, respectively.
@@
 
-We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.
+We used a simulation study to investigate performance of four sampling-analysis combinations: IRS-Design, IRS with a design-based analysis; IRS-Model, IRS with a model-based analysis; GRTS-Design, GRTS sampling with a design-based analysis; and GRTS-Model, GRTS sampling with a model-based analysis.
+These combinations are also provided in Table \ref{tab:designanalysis}.
 
@@
 row.names(tab) <- c("IRS", "GRTS")
-xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The rows give the two types of sampling designs while the columns give the two types of analyses.")
+xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-analysis combinations in the simulation study.
+The rows give the two types of sampling designs and the columns give the two types of analyses.")
 align(xtab1) <- "r|ll"
@@
 
-We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
+Performance of the four sampling-analysis combinations were evaluated in 36 different simulation scenarios.
+The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error.
+The three sample sizes ($n$) were $n = 50, n = 100,$ and $n = 200$.
+Samples were always selected from a population size ($N$) of $N = 900$.
+The two location layouts were random and gridded.
+Locations in the random layout were selected randomly from the unit square ($[0, 1] \times [0, 1]$).
+Locations in the gridded layout were selected randomly on a fixed grid from the unit square.
+The two response types were normal and lognormal.
+For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error.
+The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are from Equation$~$\ref{equation:expcov}.
+The total variance ($\sigma^2_1 + \sigma^2_2$) was always 2.
+The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decays to nearly zero at the largest possible distance between two units in the domain.
+For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
+The response was then exponentiated, yielding a total variance of 2 on the exponentiated scale.
+The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed.
+
+a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
 All scenarios used exponential correlation with a  $\sqrt{4} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
@@
 parm_tab[c(2, 4), 3] <- "-"
-row.names(parm_tab) <- c("Sample Size (n)", "Layout", 
+row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
                          "Proportion of Dependent Error", "Response Type")
-# library(knitr) kable with htlm has trouble with page breaks -- kable with latex harder for me than xtable
-# kable(parm_tab, col.names = c("", "", ""),
-#       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.")
+# parm_tab <- parm_tab[c(1, 2, 4, 3), ]
+parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the mean was zero and the variance was two.")
 align(parm_xtab) <- "r|lll"
@@
 
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey].
-FPBK can be readily performed in `R` with the `sptotal` package [@higham2021sptotal].
-We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).
+The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package spsurvey [@dumelle2021spsurvey].
+FPBK can be readily performed using the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
+We use sptotal for both the simulation analysis and the application, estimating covariance parameters with Restricted Maximum Likelihood (REML).
 
-Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} using the random location layout, where "IRS-Design" is the baseline:
-\mbox{}
+We define rMS(P)E as the root-mean-squared error (design-based) or the root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline.
+More formally, the relative rMS(P)E is defined as
 \begin{equation*}
-\text{EFF} = \frac{\text{rMS(P)E of approach}}{\text{rMS(P)E of IRS-Design}},
+\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*}
 
-where rMS(P)E is the root-mean-squared error (design-based) or the root-mean-squared-prediction error (model-based).
-When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
+When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
 So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small.
-In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).
+When there is high spatial correlation (Figure \ref{fig:figeff}, bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small.
+In the lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom row and right column), IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
 
@@
 
-```{r figeff, out.width = "100%", fig.cap = "Relative Efficiency of the four design-analysis approaches.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
+```{r figeff, out.width = "100%", fig.cap = "Relative rMS(P)E for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.", message = FALSE, warning = FALSE}
 library(tidyverse)
@@
 We also studied 95% interval coverage among the approaches.
-The design-based 95% confidence intervals and model-based 95% prediction intervals are constructed using the normal distribution.
-Justification for the design-based intervals lies in the asymptotic normality of totals via the Central Limit Theorem, and justification for the model-based intervals lies in the normality assumption of the errors.
+The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution.
+Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem (under some assumptions).
+
 Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches in the random location layout.
 All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
-
-In the normal response settings, all approaches have coverage around 95%.
-This is expected, as the intervals are also based on the normal distribution In the lognormal response settings, however, all approaches have coverage below 95%.
-This is also expected, as the intervals are still based on the normal distribution.
-In the lognormal response settings, interval coverage increases both as the sample size increases and as the strength of spatial dependence increases.
-This suggests that the larger the sample size and the stronger the spatial dependence, the more resistant these intervals are to departures from normality of the data.
+Coverage in the normal response settings tended to be near 95%, slightly higher than coverage in the lognormal settings, which still generally exceeded 90%.
+Coverage for all approaches increased with the sample size for both the normal and lognormal scenarios.
+At a sample size of 200, all four sampling-design combinations had approximately 95% interval coverage in both response scenarios.
 These conclusions were similar to those observed when the data were gridded.
 
-```{r figconf, out.width = "100%", fig.cap = "Coverage of the four design-analysis approaches.
-All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
+```{r figconf, out.width = "100%", fig.cap = "Interval coverage for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line in each plot represents 95% coverage.", message = FALSE, warning = FALSE}
 ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +

### 2021-11-11 inst/manuscript/manuscript.tex
minor updates

@@
 (historical) sampling (Foster et al., 2017). It accommodates a minimum
-distance between units in a sample. It accommodates replacement units in
-a sample, which are units that can be sampled in place of an original
-unit that can no longer be sampled. Lastly, it can be implemented using
-the spsurvey R package Dumelle et al. (2021).
-
-The GRTS algorithm samples from finite and infinite populations by
-utilizing a mapping between two-dimensional and one-dimensional space.
-The units in the two-dimensional sampling frame are divided into cells
-using a hierarchical address. This hierarchical address is then used to
-map the units from two-dimensional space to a one-dimensional line where
-each unit's line length equals its inclusion probability. A systematic
-sample is conducted on the line and linked back to a unit in
-two-dimensional space, which results in the desired sample. Stevens and
-Olsen (2004) provides further details.
+distance between units in a sample. Lastly, it accommodates replacement
+units in a sample, which are units that can be sampled in place of an
+original unit that can no longer be sampled. The GRTS algorithm samples
+from finite and infinite populations by utilizing a mapping between
+two-dimensional and one-dimensional space. The units in the
+two-dimensional sampling frame are divided into cells using a
+hierarchical address. This hierarchical address is then used to map the
+units from two-dimensional space to a one-dimensional line where each
+unit's line length equals its inclusion probability. A systematic sample
+is conducted on the line and linked back to a unit in two-dimensional
+space, which results in the desired sample. Stevens and Olsen (2004)
+provides further details.
 
@@
 \end{align} where \(\sigma^2_{1}\) is dependent random error variance
-measuring coarse-scale (correlated) variability, \(\sigma^2_{n}\) is the
+measuring coarse-scale (correlated) variability, \(\sigma^2_{2}\) is the
 independent random error variance measuring fine-scale (independent)
@@
 
-We used a numerical simulation study to investigate performance of four
-design-analysis combinations, summarized in Table
-\ref{tab:designanalysis}.
+We used a simulation study to investigate performance of four
+sampling-analysis combinations: IRS-Design, IRS with a design-based
+analysis; IRS-Model, IRS with a model-based analysis; GRTS-Design, GRTS
+sampling with a design-based analysis; and GRTS-Model, GRTS sampling
+with a model-based analysis. These combinations are also provided in
+Table \ref{tab:designanalysis}.
 
@@
 \end{tabular}
-\caption{\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The rows give the two types of sampling designs while the columns give the two types of analyses.}
+\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
 \end{table}
 
-We used a crossed design with the simulation parameters given in Table
+Performance of the four sampling-analysis combinations were evaluated in
+36 different simulation scenarios. The 36 scenarios resulted from the
+crossing of three sample sizes, two location layouts, two response
+types, and three proportions of dependent random error. The three sample
+sizes (\(n\)) were \(n = 50, n = 100,\) and \(n = 200\). Samples were
+always selected from a population size (\(N\)) of \(N = 900\). The two
+location layouts were random and gridded. Locations in the random layout
+were selected randomly from the unit square (\([0, 1] \times [0, 1]\)).
+Locations in the gridded layout were selected randomly on a fixed grid
+from the unit square. The two response types were normal and lognormal.
+For the normal response type, the response was simulated using mean-zero
+random errors with the exponential covariance
+(Equation\(~\)\ref{equation:expcov}) for varying proportions of
+dependent random error. The proportion of dependent random error is
+represented by \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where
+\(\sigma^2_1\) and \(\sigma^2_2\) are from
+Equation\(~\)\ref{equation:expcov}. The total variance
+(\(\sigma^2_1 + \sigma^2_2\)) was always 2. The range was always
+\(\sqrt{2} / 3\), which means that the correlation in the dependent
+random error decays to nearly zero at the largest possible distance
+between two units in the domain. For the lognormal response type, the
+response was first simulated using the same approach as for the normal
+response type, except that the total variance was 0.6931 instead of 2.
+The response was then exponentiated, yielding a total variance of 2 on
+the exponentiated scale. The lognormal responses were used to evaluate
+performance of the sampling-analysis approaches for data that were
+skewed.
+
+a crossed design with the simulation parameters given in Table
 \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
@@
 Sample Size (n) & 50 & 100 & 200 \\ 
-  Layout & Random & Gridded & - \\ 
+  Location Layout & Random & Gridded & - \\ 
   Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
@@
 \end{tabular}
-\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.} 
+\caption{\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the mean was zero and the variance was two.}
 \end{table}
@@
 The GRTS algorithm and the local neighborhood variance estimator are
-available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
-et al., 2021). FPBK can be readily performed in \texttt{R} with the
-\texttt{sptotal} package (Higham et al., 2021). We use \texttt{sptotal}
-for both the simulation analysis and the application, estimating
+available in the \textbf{\textsf{R}} package spsurvey (Dumelle et al.,
+2021). FPBK can be readily performed using the \texttt{sptotal}
+\textbf{\textsf{R}} package (Higham et al., 2021). We use sptotal for
+both the simulation analysis and the application, estimating covariance
 parameters with Restricted Maximum Likelihood (REML).
 
-Figure \ref{fig:figeff} shows the relative efficiency of the four
-approaches from Table \ref{tab:designanalysis} using the random location
-layout, where ``IRS-Design'' is the baseline: \mbox{} \begin{equation*}
-\text{EFF} = \frac{\text{rMS(P)E of approach}}{\text{rMS(P)E of IRS-Design}},
+We define rMS(P)E as the root-mean-squared error (design-based) or the
+root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff}
+shows the relative rMS(P)E of the four approaches from Table
+\ref{tab:designanalysis} using the random location layout with
+``IRS-Design'' is the baseline. More formally, the relative rMS(P)E is
+defined as \begin{equation*}
+\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*}
 
-where rMS(P)E is the root-mean-squared error (design-based) or the
-root-mean-squared-prediction error (model-based). When there is no
-spatial correlation (top row), the four approaches have approximately
-equal rMSPE, even when the assumptions of the model-based approaches are
-violated. So, using GRTS or using a spatial model does not result in
-much, if any, loss in efficiency even if the response variable is not
-spatially correlated. When there is high spatial correlation (bottom
+When there is no spatial correlation (Figure \ref{fig:figeff}, top row),
+the four approaches have approximately equal rMSPE, even when the
+assumptions of the model-based approaches are violated. So, using GRTS
+or using a spatial model does not result in much, if any, loss in
+efficiency even if the response variable is not spatially correlated.
+When there is high spatial correlation (Figure \ref{fig:figeff}, bottom
 row), the GRTS-Model approach tends to perform best, but difference in
 relative efficiency between GRTS-Model and GRTS-Design is small. In the
-lognormal, high partial sill settings (bottom-right facet), GRTS-Design
-outperforms IRS-Model by a large margin, suggesting that the design
-decision (whether to use IRS or GRTS) is more important than the
-analysis decision (whether to analyze using model assumptions or not).
+lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom
+row and right column), IRS-Model outperforms IRS-Design by a large
+margin, suggesting that the poor design properties of IRS are largely
+mitigated by the model-based analysis.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figeff-1} \caption{Relative Efficiency of the four design-analysis approaches.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figeff}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figeff-1} \caption{Relative rMS(P)E for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.}\label{fig:figeff}
 \end{figure}
@@
 design-based 95\% confidence intervals and model-based 95\% prediction
-intervals are constructed using the normal distribution. Justification
-for the design-based intervals lies in the asymptotic normality of
-totals via the Central Limit Theorem, and justification for the
-model-based intervals lies in the normality assumption of the errors.
+intervals were constructed using the normal distribution. Justification
+for the design-based and model-based intervals comes from the asymptotic
+normality of totals via the Central Limit Theorem (under some
+assumptions).
+
 Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
@@
 GRTS-Design having slightly lower coverage when the response is normal.
+Coverage in the normal response settings tended to be near 95\%,
+slightly higher than coverage in the lognormal settings, which still
+generally exceeded 90\%. Coverage for all approaches increased with the
+sample size for both the normal and lognormal scenarios. At a sample
+size of 200, all four sampling-design combinations had approximately
+95\% interval coverage in both response scenarios. These conclusions
+were similar to those observed when the data were gridded.
 
-In the normal response settings, all approaches have coverage around
-95\%. This is expected, as the intervals are also based on the normal
-distribution In the lognormal response settings, however, all approaches
-have coverage below 95\%. This is also expected, as the intervals are
-still based on the normal distribution. In the lognormal response
-settings, interval coverage increases both as the sample size increases
-and as the strength of spatial dependence increases. This suggests that
-the larger the sample size and the stronger the spatial dependence, the
-more resistant these intervals are to departures from normality of the
-data. These conclusions were similar to those observed when the data
-were gridded.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1} \caption{Coverage of the four design-analysis approaches.
-All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line.
-The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figconf}
-\end{figure}
+\textbackslash begin\{figure\}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1}
+\textbackslash caption\{Interval coverage for the four sampling-analysis
+combinations. The rows indicate the proportion of dependent error and
+the columns indicate the response type. The solid, black line in each
+plot represents 95\% coverage.\}\label{fig:figconf}
+\textbackslash end\{figure\}
 

### 2021-11-12 inst/manuscript/manuscript.Rmd
section 3 updates

@@
 
-Performance of the four sampling-analysis combinations were evaluated in 36 different simulation scenarios.
+Performance of the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
 The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error.
@@
 The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are from Equation$~$\ref{equation:expcov}.
-The total variance ($\sigma^2_1 + \sigma^2_2$) was always 2.
-The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decays to nearly zero at the largest possible distance between two units in the domain.
+The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
+The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decayed to nearly zero at the largest possible distance between two units in the domain.
 For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a total variance of 2 on the exponentiated scale.
+The response was then exponentiated, yielding a random variable whose total variance is 2.
 The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed.
-
-a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios.
-All scenarios used exponential correlation with a  $\sqrt{4} / 3$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded).
-The mean for the spatial process generating the response was set to zero.
-
-<!-- The range was chosen so that the decayed to nearly zero at the largest distance possible in the domain, $\sqrt{2}$, otherwise known as the "effective" range (for the exponential covariance, the effective range is $3\phi$).
--->
-
-For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated.
-A total variance of 0.6931 and a mean of 0 on the normal scale yielded a total variance of 2 and a mean of 1.414 after exponentiation.
-Therefore, when the model-based methods were used for lognormal response, the correlation was mis-specified.
-We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness.
 
@@
 All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the mean was zero and the variance was two.")
+In each simualtion scenario, the total variance was two.")
 align(parm_xtab) <- "r|lll"
@@
 
-```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale.
-The distribution exhibits moderate to strong right-skewness.", fig.keep = "none"}
-mu <- 0; var <- 2
-xvar <- seq(0.001, 10, by = 0.01)
-yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
-df <- tibble(xvar = xvar, yvar = yvar)
-ggplot(df, aes(x = xvar, y = yvar)) +
-  geom_line()
-```
+In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
+In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate the mean and construct confidence (design-based) or prediction (model-based) intervals.
+We recorded the bias, squared error, and interval coverage for all sampling-analysis combinations in each trial.
+Then we summarized the performance of the combinations across trials by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the rate at which the true mean is contained in its 95% interval.
+The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package `spsurvey` [@dumelle2021spsurvey].
+FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal] and covariance parameters were estimated using Restricted Maximum Likelihood  [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
 
-There were 2000 simulation trials for each of the 36 parameter combinations.
-In each trial, response values were generated from a spatial process with the specified parameters, and a GRTS sample and an IRS sample were selected.
-For the GRTS sample, the design-based approach using the local neighborhood variance (GRTS-Design) and a model-based approach were applied (GRTS-Model).
-For the IRS sample, the design-based approach using the simple random sample variance (IRS-Design) and a model-based approach were applied (IRS-Model).
-
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package spsurvey [@dumelle2021spsurvey].
-FPBK can be readily performed using the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
-We use sptotal for both the simulation analysis and the application, estimating covariance parameters with Restricted Maximum Likelihood (REML).
-
-We define rMS(P)E as the root-mean-squared error (design-based) or the root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline.
+Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline.
 More formally, the relative rMS(P)E is defined as
@@
 \end{equation*}
-
-When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated.
-So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (Figure \ref{fig:figeff}, bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small.
-In the lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom row and right column), IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
-
-Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the assumptions for GRTS-Model are satisfied, the approach outperforms GRTS-Design.
-However, even when the model that generates the data is different than the model used to fit the data, as in the lognormal response, the model-based approach still outperforms the design-based approach when there is a high amount of spatial correlation.
-Additionally, as the sample size increases, IRS-Design performs relatively worse compared to the other approaches.
-These conclusions were similar to those observed when the data were gridded.
+When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even when the response variable is not spatially correlated.
+When there is spatial correlation (Figure \ref{fig:figeff}, middle and bottom row), the GRTS-Model combination tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively small.
+As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other combinations widens.
+Finally we note that when there is spatial correlation, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
+These conclusions are similar to those observed in the grid location layout.
+Tables for rMS(P)E in all 36 simulation scenarios are provided in the supplementary material.
 
@@
 
-```{r, fig.keep = "none"}
-combo_data_grts <- combo_data %>%
-  dplyr::filter(approach %in% c("Design GRTS", "Model GRTS")) %>%
-  mutate(designgrts = if_else(approach == "Design GRTS", 
-                              true = rmspe,
-                              false = NA_real_)) %>%
-  group_by(sim) %>%
-  fill(designgrts, .direction = "updown") %>%
-  ungroup() %>%
-  mutate(releffgrts = rmspe / designgrts)
-combo_data_grts_model <- combo_data_grts %>%
-  filter(approach == "Model GRTS")
+We also studied 95% interval coverage among the combinations.
+The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution.
+Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem.
 
-ggplot(data = combo_data_grts_model, aes(x = n, y = releffgrts)) +
-  facet_grid(psill ~ resptype) +
-  geom_point()
-```
-
-We also studied 95% interval coverage among the approaches.
-The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution.
-Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem (under some assumptions).
-
-Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches in the random location layout.
-All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.
-Coverage in the normal response settings tended to be near 95%, slightly higher than coverage in the lognormal settings, which still generally exceeded 90%.
-Coverage for all approaches increased with the sample size for both the normal and lognormal scenarios.
-At a sample size of 200, all four sampling-design combinations had approximately 95% interval coverage in both response scenarios.
-These conclusions were similar to those observed when the data were gridded.
+Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four combinations in the random location layout.
+All four combinations have fairly similar interval coverage within each scenario.
+Coverage in the normal response scenarios tended to be near 95% and slightly higher than coverage in the lognormal scenarios.
+Coverage in the lognormal scenarios still generally exceeded 90%.
+Coverage tended to always increase with the sample size.
+At a sample size of 200, all four combinations had approximately 95% interval coverage in both response scenarios and all dependent error proportions.
+These conclusions were similar to those found in the grid location layout.
+Tables for interval coverage in all 36 simulation scenarios are provided in the supplementary material.
 
@@
 The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line in each plot represents 95% coverage.", message = FALSE, warning = FALSE}
+The solid, black line in each plot represents 95\\% coverage.", message = FALSE, warning = FALSE}
 ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
@@
 
-In addition to rMS(P)E and interval coverage, we also recorded average bias.
-The average bias is nearly zero for all approaches in all scenarios, so we omit a visualization of the results here.
-The supplementary material contains tables with mean bias, rMS(P)E, and interval coverage for all 36 simulation scenarios.
+The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a visualization of those results here.
+Tables for average bias in all 36 simulation scenarios are provided in the supplementary material.
 

### 2021-11-12 inst/manuscript/manuscript.tex
section 3 updates

@@
 
-Performance of the four sampling-analysis combinations were evaluated in
+Performance of the four sampling-analysis combinations was evaluated in
 36 different simulation scenarios. The 36 scenarios resulted from the
@@
 \(\sigma^2_1\) and \(\sigma^2_2\) are from
-Equation\(~\)\ref{equation:expcov}. The total variance
-(\(\sigma^2_1 + \sigma^2_2\)) was always 2. The range was always
+Equation\(~\)\ref{equation:expcov}. The total variance,
+\(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
 \(\sqrt{2} / 3\), which means that the correlation in the dependent
-random error decays to nearly zero at the largest possible distance
+random error decayed to nearly zero at the largest possible distance
 between two units in the domain. For the lognormal response type, the
@@
 response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a total variance of 2 on
-the exponentiated scale. The lognormal responses were used to evaluate
+The response was then exponentiated, yielding a random variable whose
+total variance is 2. The lognormal responses were used to evaluate
 performance of the sampling-analysis approaches for data that were
 skewed.
-
-a crossed design with the simulation parameters given in Table
-\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
-exponential correlation with a \(\sqrt{4} / 3\) for \(N = 900\) response
-values simulated on the unit square in either random locations (Layout =
-Random) or gridded locations (Layout = Gridded). The mean for the
-spatial process generating the response was set to zero.
-
-For the lognormal scenarios, the response values were simulated using
-the specified correlation parameters using a normal distribution and
-were subsequently exponentiated. A total variance of 0.6931 and a mean
-of 0 on the normal scale yielded a total variance of 2 and a mean of
-1.414 after exponentiation. Therefore, when the model-based methods were
-used for lognormal response, the correlation was mis-specified. We chose
-to simulate values with a lognormal distribution so that we could test
-the model-based analysis approach with a mis-specified model and so that
-we could test both analysis approaches on data that exhibits a large
-amount of skewness.
 
@@
 All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the mean was zero and the variance was two.}
+In each simualtion scenario, the total variance was two.}
 \end{table}
 
-There were 2000 simulation trials for each of the 36 parameter
-combinations. In each trial, response values were generated from a
-spatial process with the specified parameters, and a GRTS sample and an
-IRS sample were selected. For the GRTS sample, the design-based approach
-using the local neighborhood variance (GRTS-Design) and a model-based
-approach were applied (GRTS-Model). For the IRS sample, the design-based
-approach using the simple random sample variance (IRS-Design) and a
-model-based approach were applied (IRS-Model).
+In each of the 36 simulation scenarios, there were 2000 independent
+simulation trials. In each trial, IRS and GRTS samples were selected and
+then design-based and model-based analyses were used to estimate the
+mean and construct confidence (design-based) or prediction (model-based)
+intervals. We recorded the bias, squared error, and interval coverage
+for all sampling-analysis combinations in each trial. Then we summarized
+the performance of the combinations across trials by calculating average
+bias, rMS(P)E (root-mean-squared error for the design-based approaches
+and root-mean-squared-prediction error for the model-based approaches),
+and the rate at which the true mean is contained in its 95\% interval.
+The GRTS algorithm and the local neighborhood variance estimator are
+available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
+et al., 2021). FPBK is available in the \texttt{sptotal}
+\textbf{\textsf{R}} package (Higham et al., 2021) and covariance
+parameters were estimated using Restricted Maximum Likelihood (Harville,
+1977; Patterson and Thompson, 1971; Wolfinger et al., 1994).
 
-The GRTS algorithm and the local neighborhood variance estimator are
-available in the \textbf{\textsf{R}} package spsurvey (Dumelle et al.,
-2021). FPBK can be readily performed using the \texttt{sptotal}
-\textbf{\textsf{R}} package (Higham et al., 2021). We use sptotal for
-both the simulation analysis and the application, estimating covariance
-parameters with Restricted Maximum Likelihood (REML).
-
-We define rMS(P)E as the root-mean-squared error (design-based) or the
-root-mean-squared-prediction error (model-based).Figure \ref{fig:figeff}
-shows the relative rMS(P)E of the four approaches from Table
-\ref{tab:designanalysis} using the random location layout with
-``IRS-Design'' is the baseline. More formally, the relative rMS(P)E is
-defined as \begin{equation*}
+Figure \ref{fig:figeff} shows the relative rMS(P)E of the four
+approaches from Table \ref{tab:designanalysis} using the random location
+layout with ``IRS-Design'' is the baseline. More formally, the relative
+rMS(P)E is defined as \begin{equation*}
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
-\end{equation*}
-
-When there is no spatial correlation (Figure \ref{fig:figeff}, top row),
-the four approaches have approximately equal rMSPE, even when the
-assumptions of the model-based approaches are violated. So, using GRTS
-or using a spatial model does not result in much, if any, loss in
-efficiency even if the response variable is not spatially correlated.
-When there is high spatial correlation (Figure \ref{fig:figeff}, bottom
-row), the GRTS-Model approach tends to perform best, but difference in
-relative efficiency between GRTS-Model and GRTS-Design is small. In the
-lognormal, high partial sill settings (Figure \ref{fig:figeff}, bottom
-row and right column), IRS-Model outperforms IRS-Design by a large
-margin, suggesting that the poor design properties of IRS are largely
-mitigated by the model-based analysis.
-
-Unsurprisingly, Figure \ref{fig:figeff} also shows that, when the
-assumptions for GRTS-Model are satisfied, the approach outperforms
-GRTS-Design. However, even when the model that generates the data is
-different than the model used to fit the data, as in the lognormal
-response, the model-based approach still outperforms the design-based
-approach when there is a high amount of spatial correlation.
-Additionally, as the sample size increases, IRS-Design performs
-relatively worse compared to the other approaches. These conclusions
-were similar to those observed when the data were gridded.
+\end{equation*} When there is no spatial correlation (Figure
+\ref{fig:figeff}, top row), the four sampling-analysis combinations have
+approximately equal rMS(P)E. So, using GRTS or using a spatial model
+does not result in much, if any, loss in efficiency even when the
+response variable is not spatially correlated. When there is spatial
+correlation (Figure \ref{fig:figeff}, middle and bottom row), the
+GRTS-Model combination tends to perform best, followed by GRTS-Design,
+IRS-Model, and finally IRS-Design, though the difference in relative
+rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively
+small. As the strength of spatial correlation increases, the gap in
+rMS(P)E between IRS-Design and the other combinations widens. Finally we
+note that when there is spatial correlation, IRS-Model outperforms
+IRS-Design by a large margin, suggesting that the poor design properties
+of IRS are largely mitigated by the model-based analysis. These
+conclusions are similar to those observed in the grid location layout.
+Tables for rMS(P)E in all 36 simulation scenarios are provided in the
+supplementary material.
 
@@
 
-We also studied 95\% interval coverage among the approaches. The
+We also studied 95\% interval coverage among the combinations. The
 design-based 95\% confidence intervals and model-based 95\% prediction
@@
 for the design-based and model-based intervals comes from the asymptotic
-normality of totals via the Central Limit Theorem (under some
-assumptions).
+normality of totals via the Central Limit Theorem.
 
 Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
-the four approaches in the random location layout. All four approaches
-have somewhat similar interval coverage in all settings, with
-GRTS-Design having slightly lower coverage when the response is normal.
-Coverage in the normal response settings tended to be near 95\%,
-slightly higher than coverage in the lognormal settings, which still
-generally exceeded 90\%. Coverage for all approaches increased with the
-sample size for both the normal and lognormal scenarios. At a sample
-size of 200, all four sampling-design combinations had approximately
-95\% interval coverage in both response scenarios. These conclusions
-were similar to those observed when the data were gridded.
+the four combinations in the random location layout. All four
+combinations have fairly similar interval coverage within each scenario.
+Coverage in the normal response scenarios tended to be near 95\% and
+slightly higher than coverage in the lognormal scenarios. Coverage in
+the lognormal scenarios still generally exceeded 90\%. Coverage tended
+to always increase with the sample size. At a sample size of 200, all
+four combinations had approximately 95\% interval coverage in both
+response scenarios and all dependent error proportions. These
+conclusions were similar to those found in the grid location layout.
+Tables for interval coverage in all 36 simulation scenarios are provided
+in the supplementary material.
 
-\textbackslash begin\{figure\}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1}
-\textbackslash caption\{Interval coverage for the four sampling-analysis
-combinations. The rows indicate the proportion of dependent error and
-the columns indicate the response type. The solid, black line in each
-plot represents 95\% coverage.\}\label{fig:figconf}
-\textbackslash end\{figure\}
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1} \caption{Interval coverage for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line in each plot represents 95\% coverage.}\label{fig:figconf}
+\end{figure}
 
-In addition to rMS(P)E and interval coverage, we also recorded average
-bias. The average bias is nearly zero for all approaches in all
-scenarios, so we omit a visualization of the results here. The
-supplementary material contains tables with mean bias, rMS(P)E, and
-interval coverage for all 36 simulation scenarios.
+The average bias was nearly zero for all four combinations in all 36
+scenarios, so we omit a visualization of those results here. Tables for
+average bias in all 36 simulation scenarios are provided in the
+supplementary material.
 
@@
 
+\leavevmode\hypertarget{ref-harville1977maximum}{}%
+Harville, D.A., 1977. Maximum likelihood approaches to variance
+component estimation and to related problems. Journal of the American
+statistical association 72, 320--338.
+
 \leavevmode\hypertarget{ref-higham2021sptotal}{}%
@@
 Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.
+
+\leavevmode\hypertarget{ref-patterson1971recovery}{}%
+Patterson, H.D., Thompson, R., 1971. Recovery of inter-block information
+when block sizes are unequal. Biometrika 58, 545--554.
 
@@
 
+\leavevmode\hypertarget{ref-wolfinger1994computing}{}%
+Wolfinger, R., Tobias, R., Sall, J., 1994. Computing gaussian
+likelihoods and their derivatives for general linear mixed models. SIAM
+Journal on Scientific Computing 15, 1294--1310.
+
 \end{CSLReferences}

### 2021-11-12 inst/manuscript/manuscript.Rmd
finished draft for review

@@
 abstract: |
-  This is the abstract.
+  The design-based and model-based approaches to frequentist statistical inference rest on fundamentally different foundations.
+In the design-based approach, inference depends on random sampling.
+In the model-based approach, inference depends on distributional assumptions.
+In this manuscript, we compare the approaches for finite population spatial data.
+We first provide relevant background for design-based and model-based approaches to finite population patial data.
+Then we use a simulation study and an analysis of real mercurcy concentration data to compare them numerically.
+We find that sampling plans that incorporate spatial locations (spatially balanced samples) perform better than sampling plans ignoring spatial locations (non-spatially balanced samples), regardless of whether design-based or model-based approaches were used to analyze the data.
+We also find that within sampling plans, model-based approaches tend to outperform design-based approaches, even for skewed data.
+This gap in performance is small when spatially samples are used but large when non-spatially balanced samples are used.
 
@@
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @wang2012review; @brus2020statistical].
+Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @wang2012review; @brus2021statistical].
 @cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
@@
 @cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@verhoef2002sampling and @brus2020statistical provide similar figures).
+A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@verhoef2002sampling and @brus2021statistical provide similar figures).
 
@@
 We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-Other methods, such as k-nearest-neighbors [@fix1951discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
+Other methods, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
 We choose to use FPBK because it is faster than a Bayesian approach and it was developed with theoretically-based variance estimators of means and totals for spatial data, whereas random forests and k-nearest-neighbors use ad-hoc variance estimators in most cases [@ver2013comparison]; additionally, FBPK outperformed the other methods in most scenarios.
@@
 FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal] and covariance parameters were estimated using Restricted Maximum Likelihood  [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+
+The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+Tables for average bias in all 36 simulation scenarios are provided in the supplementary material.
 
@@
 
-The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a visualization of those results here.
-Tables for average bias in all 36 simulation scenarios are provided in the supplementary material.
-
 # Application {#application}
@@
 The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration.
+We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to study mercury concentration.
 Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
@@
 nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-```
-
-```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.", message = FALSE, warning = FALSE}
-merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(colour = "black", fill = "white", bins = 15) +
-  labs(x = "HG (ng / g)")
-# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
-#   geom_histogram(colour = "black", fill = "white", bins = 15) +
-#   labs(x = "Lakeshore Disturbance")
-merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "hg") +
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank()) +
-  coord_quickmap() +
-  theme_bw(base_size = 14)
-# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-#   geom_point(aes(colour = RDis_IX)) +
-#   scale_colour_viridis_c() +
-#   labs(x = "", y = "", colour = "Dist") +
-#   theme(axis.text.x = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.ticks = element_blank())
-library(gridExtra)
-# grid.arrange(merc, merc_map, nrow = 2)
-merc_map
 ```
@@
 
+```{r figdata, out.width = "49%", fig.cap = "Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).", message = FALSE, warning = FALSE, fig.show='hold', fig.align='center'}
+merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
+  geom_histogram(colour = "black", fill = "white", bins = 15) +
+  labs(x = "HG (ng / g)")
+# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
+#   geom_histogram(colour = "black", fill = "white", bins = 15) +
+#   labs(x = "Lakeshore Disturbance")
+merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
+  geom_point(aes(colour = total_hg)) +
+  scale_colour_viridis_c() +
+  labs(x = "", y = "", colour = "Hg") +
+  lims(x = c(0, 6500)) +
+  coord_quickmap() +
+  theme_bw(base_size = 20) + 
+  theme(axis.text.x = element_blank(),
+      axis.text.y = element_blank(),
+      axis.ticks = element_blank(),
+      legend.position = c(0.85, 0.5),
+      legend.key.size = unit(0.75, "cm"))
+
+merc_hist <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
+  geom_histogram(bins = 50) + 
+  labs(x = "Hg", y = "Count") +
+  theme_bw(base_size = 20)
+
+merc_map
+merc_hist
+```
+
 ```{r apptab, results = "asis"}
@@
 
-Table \ref{tab:appliedtab} shows the application of a design-based analysis of an IRS sample, a model-based analysis of an IRS sample, a design-based analysis of a GRTS sample, and a model-based analysis of a GRTS sample.
-For all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals.
-However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.
-
-<!-- not sure why this is getting cutoff. temporary fix: \pagebreak 
-probably because it is an html table -- I used xtable to fix -->
-
+We selected a single IRS sample and a single GRTS sample and estimated the mean mercury concentration and its standard error using using design-based and model-based approaches; Table \ref{tab:appliedtab} shows the results.
+For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% intervals.
+However, we should not generalize these results to any other data or even to other samples from these data.
 But, we do note a couple of patterns.
@@
 
+To better understand the dependence structure in mercury concentration, the empirical semivariogram and corresponding fit of the model-based approaches can be visualized.
+The empirical semivariogram quantifies the halved squared differences (semivariance) among response values at different distances apart.
+If a process exhibits strong spatial dependence, the empirical semivariogram will have small values at small distances and large values at large distances.
+Figure \ref{fig:figsv} shows the empirical semivariogram for GRTS-Model, displaying the average semivariance for several distances.
+Overlain onto Figure \ref{fig:figsv} is the estimated semivariance obtained using the covariance parmaeters from the REML fit of GRTS-Model.
+Figure \ref{fig:figsv} provides evidence that there is strong correlation in mercury concentration among the sites.
+
+```{r figsv, out.width = "85%", fig.align='center', fig.cap = "The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model."}
+plot(mod_grts)
+```
+
 # Discussion {#sec:discussion}
 
-The design-based and model-based approaches to inference are fundamentally different paradigms by which to samples are selected and data are analyzed.
+The design-based and model-based approaches to inference are fundamentally different paradigms by which samples are selected and data are analyzed.
 The design-based approach incorporates randomness through sampling to estimate a population parameter.
-The model-based approach incorporates randomness through distributional assumptions to predict the realized value of a random variable.
+The model-based approach incorporates randomness through distributional assumptions to predict the realized values of a random process.
 Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
-Aside from the theoretical differences described, a few analytical findings are particularly notable: the design decision (GRTS vs IRS) seems much more important than the analysis decision (design-based vs model-based); independent of the analysis approach, there is no reason to prefer IRS over GRTS for spatial data -- GRTS tends to perform at least as well as IRS when there is no spatial correlation and increasingly better than IRS as the strength of spatial correlation increases; the gap in relative efficiency between GRTS-design and GRTS-model widens as the strength of spatial correlation increases; and when the data are skewed, interval coverage for all approaches improves both as the sample size increases and as the strength of correlation increases.
-
-<!-- The design-based approach relies on few assumptions, while the model-based approach relies on rigid distributional ones. -->
+Aside from the theoretical differences described, a few analytical findings from the simulation study are particularly notable.
+First, the sampling decision (GRTS vs IRS) is most important when using a design-based analysis.
+Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated much of the inefficiency of the IRS sample.
+Second, independent of the analysis approach, there is no reason to use IRS over GRTS for sampling spatial data, as GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial correlation and noticeably better than there IRS counterparts when there was spatial correlation.
+Third, The stronger the spatial correlation, the larger the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations.
+Fourth and finally, interval coverage for the normal response was very close to 95% for all sample sizes, while interval coverage for the lognormal response was not very close to 95% until $n = 200$.
 
@@
 The model-based approach also yields estimated spatial covariance parameters, which help better understand the process of study.
-Model selection is also possible using model-based approaches and criteria such as likelihood ratio tests or AIC [@akaike1974new].
+Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
 Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
 Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
-The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be heavily considered when choosing among them.
+The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be seriously considered when choosing among them.
 This is especially true from an analysis perspective, as we found that using a spatially balanced sampling algorithm benefits both design-based and model-based analyses.
@@
 
-In the supplementary material, we provide summary statistics for all 36 simulation scenarios.
+In the supplementary material, we provide tables presenting summary statistics for all 36 simulation scenarios.
 

### 2021-11-12 inst/manuscript/manuscript.tex
finished draft for review

@@
   \begin{abstract}
-  This is the abstract.
+  The design-based and model-based approaches to frequentist statistical
+  inference rest on fundamentally different foundations. In the
+  design-based approach, inference depends on random sampling. In the
+  model-based approach, inference depends on distributional assumptions.
+  In this manuscript, we compare the approaches for finite population
+  spatial data. We first provide relevant background for design-based
+  and model-based approaches to finite population patial data. Then we
+  use a simulation study and an analysis of real mercurcy concentration
+  data to compare them numerically. We find that sampling plans that
+  incorporate spatial locations (spatially balanced samples) perform
+  better than sampling plans ignoring spatial locations (non-spatially
+  balanced samples), regardless of whether design-based or model-based
+  approaches were used to analyze the data. We also find that within
+  sampling plans, model-based approaches tend to outperform design-based
+  approaches, even for skewed data. This gap in performance is small
+  when spatially samples are used but large when non-spatially balanced
+  samples are used.
   \end{abstract}
@@
 comparisons between design-based and model-based approaches for spatial
-data (Brus and De Gruijter, 1997; Brus, 2020; Ver Hoef, 2002; Ver Hoef,
+data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002; Ver Hoef,
 2008; Wang et al., 2012). Cooper (2006) reviews the two approaches in an
@@
 model-based assumptions is provided in Figure \ref{fig:fig1} (Ver Hoef
-(2002) and Brus (2020) provide similar figures).
+(2002) and Brus (2021) provide similar figures).
 
@@
 comparing the design-based and model-based approaches. Other methods,
-such as k-nearest-neighbors (Fix and Hodges, 1951; Ver Hoef and
+such as k-nearest-neighbors (Fix and Hodges, 1989; Ver Hoef and
 Temesgen, 2013), random forest (Breiman, 2001), Bayesian models
@@
 
+The average bias was nearly zero for all four combinations in all 36
+scenarios, so we omit a more detailed summary of those results here.
+Tables for average bias in all 36 simulation scenarios are provided in
+the supplementary material.
+
 Figure \ref{fig:figeff} shows the relative rMS(P)E of the four
@@
 
-The average bias was nearly zero for all four combinations in all 36
-scenarios, so we omit a visualization of those results here. Tables for
-average bias in all 36 simulation scenarios are provided in the
-supplementary material.
-
 \hypertarget{application}{%
@@
 various aspects of lake health and quality in lakes in the contiguous
-United States, to obtain an interval for mean mercury concentration.
-Although we know the true mean mercury concentration values for the 986
-lakes from the 2012 NLA, we will explore whether or not we obtain an
-adequately precise estimate for the realized mean mercury concentration
-if we sample only 100 of the 986 lakes.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.}\label{fig:figdata}
-\end{figure}
+United States, to study mercury concentration. Although we know the true
+mean mercury concentration values for the 986 lakes from the 2012 NLA,
+we will explore whether or not we obtain an adequately precise estimate
+for the realized mean mercury concentration if we sample only 100 of the
+986 lakes.
 
@@
 103.2 ng / g.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.49\linewidth]{manuscript_files/figure-latex/figdata-1} \includegraphics[width=0.49\linewidth]{manuscript_files/figure-latex/figdata-2}
+
+}
+
+\caption{Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).}\label{fig:figdata}
+\end{figure}
 
@@
 
-Table \ref{tab:appliedtab} shows the application of a design-based
-analysis of an IRS sample, a model-based analysis of an IRS sample, a
-design-based analysis of a GRTS sample, and a model-based analysis of a
-GRTS sample. For all four analyses, the true realized mean mercury
-concentration is within the bounds of the 95\% intervals. However, we
-should not generalize the results of this particular realization to any
-other data set or even to other potential samples of this data set.
+We selected a single IRS sample and a single GRTS sample and estimated
+the mean mercury concentration and its standard error using using
+design-based and model-based approaches; Table \ref{tab:appliedtab}
+shows the results. For all four sampling-analysis combinations, the true
+realized mean mercury concentration is within the bounds of the 95\%
+intervals. However, we should not generalize these results to any other
+data or even to other samples from these data. But, we do note a couple
+of patterns. The design-based IRS analysis shows the largest standard
+error: a likely reason is that this is the only approach that does not
+incorporate any spatial information regarding mercury concentration
+across the contiguous United States. We also see that both approaches
+using the GRTS sample have a lower standard error than the both
+approaches using the IRS sample. We would expect this to be the case for
+most samples because mercury concentration exhibits spatial patterning,
+so a spatially balanced sample should usually yield a lower standard
+error.
 
-But, we do note a couple of patterns. The design-based IRS analysis
-shows the largest standard error: a likely reason is that this is the
-only approach that does not incorporate any spatial information
-regarding mercury concentration across the contiguous United States. We
-also see that both approaches using the GRTS sample have a lower
-standard error than the both approaches using the IRS sample. We would
-expect this to be the case for most samples because mercury
-concentration exhibits spatial patterning, so a spatially balanced
-sample should usually yield a lower standard error.
+To better understand the dependence structure in mercury concentration,
+the empirical semivariogram and corresponding fit of the model-based
+approaches can be visualized. The empirical semivariogram quantifies the
+halved squared differences (semivariance) among response values at
+different distances apart. If a process exhibits strong spatial
+dependence, the empirical semivariogram will have small values at small
+distances and large values at large distances. Figure \ref{fig:figsv}
+shows the empirical semivariogram for GRTS-Model, displaying the average
+semivariance for several distances. Overlain onto Figure \ref{fig:figsv}
+is the estimated semivariance obtained using the covariance parmaeters
+from the REML fit of GRTS-Model. Figure \ref{fig:figsv} provides
+evidence that there is strong correlation in mercury concentration among
+the sites.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.85\linewidth]{manuscript_files/figure-latex/figsv-1}
+
+}
+
+\caption{The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model.}\label{fig:figsv}
+\end{figure}
 
@@
 The design-based and model-based approaches to inference are
-fundamentally different paradigms by which to samples are selected and
-data are analyzed. The design-based approach incorporates randomness
-through sampling to estimate a population parameter. The model-based
-approach incorporates randomness through distributional assumptions to
-predict the realized value of a random variable. Though these approaches
-have often been compared in the literature both from theoretical and
+fundamentally different paradigms by which samples are selected and data
+are analyzed. The design-based approach incorporates randomness through
+sampling to estimate a population parameter. The model-based approach
+incorporates randomness through distributional assumptions to predict
+the realized values of a random process. Though these approaches have
+often been compared in the literature both from theoretical and
 analytical perspectives, our contribution lies in studying them in a
@@
 from the theoretical differences described, a few analytical findings
-are particularly notable: the design decision (GRTS vs IRS) seems much
-more important than the analysis decision (design-based vs model-based);
-independent of the analysis approach, there is no reason to prefer IRS
-over GRTS for spatial data -- GRTS tends to perform at least as well as
-IRS when there is no spatial correlation and increasingly better than
-IRS as the strength of spatial correlation increases; the gap in
-relative efficiency between GRTS-design and GRTS-model widens as the
-strength of spatial correlation increases; and when the data are skewed,
-interval coverage for all approaches improves both as the sample size
-increases and as the strength of correlation increases.
+from the simulation study are particularly notable. First, the sampling
+decision (GRTS vs IRS) is most important when using a design-based
+analysis. Though GRTS-Model still outperformed IRS-Model, the
+model-based analysis mitigated much of the inefficiency of the IRS
+sample. Second, independent of the analysis approach, there is no reason
+to use IRS over GRTS for sampling spatial data, as GRTS-Design and
+GRTS-Model generally performed at least as well as their IRS
+counterparts when there was no spatial correlation and noticeably better
+than there IRS counterparts when there was spatial correlation. Third,
+The stronger the spatial correlation, the larger the gap in rMS(P)E
+between IRS-Design and the other sampling-analysis combinations. Fourth
+and finally, interval coverage for the normal response was very close to
+95\% for all sample sizes, while interval coverage for the lognormal
+response was not very close to 95\% until \(n = 200\).
 
@@
 understand the process of study. Model selection is also possible using
-model-based approaches and criteria such as likelihood ratio tests or
-AIC (Akaike, 1974). Model-based approaches are capable of more efficient
-small-area estimation than design-based approaches by leveraging
-distributional assumptions in areas with few observed sites. Model-based
-approaches can also compute site-by-site predictions at unobserved
-locations and use them to construct informative visualizations. The
-benefits and drawbacks of both approaches, alongside our theoretical and
-analytical comparisons, should be heavily considered when choosing among
-them. This is especially true from an analysis perspective, as we found
-that using a spatially balanced sampling algorithm benefits both
-design-based and model-based analyses.
+model-based approaches and criteria such as cross validation, likelihood
+ratio tests, or AIC (Akaike, 1974). Model-based approaches are capable
+of more efficient small-area estimation than design-based approaches by
+leveraging distributional assumptions in areas with few observed sites.
+Model-based approaches can also compute site-by-site predictions at
+unobserved locations and use them to construct informative
+visualizations. The benefits and drawbacks of both approaches, alongside
+our theoretical and analytical comparisons, should be seriously
+considered when choosing among them. This is especially true from an
+analysis perspective, as we found that using a spatially balanced
+sampling algorithm benefits both design-based and model-based analyses.
 
@@
 
-In the supplementary material, we provide summary statistics for all 36
-simulation scenarios.
+In the supplementary material, we provide tables presenting summary
+statistics for all 36 simulation scenarios.
 
@@
 Akaike, H., 1974. A new look at the statistical model identification.
-IEEE transactions on automatic control 19, 716--723.
+IEEE Transactions on Automatic Control 19, 716--723.
 
@@
 \leavevmode\hypertarget{ref-breiman2001random}{}%
-Breiman, L., 2001. Random forests. Machine learning 45, 5--32.
+Breiman, L., 2001. Random forests. Machine Learning 45, 5--32.
 
@@
 Brus, D., De Gruijter, J., 1997. Random sampling or geostatistical
-modelling? Choosing between design-based and model-based sampling
+modelling? Choosing between design-based and model-dased sampling
 strategies for soil (with discussion). Geoderma 80, 1--44.
 
-\leavevmode\hypertarget{ref-brus2020statistical}{}%
-Brus, D.J., 2020. Statistical approaches for spatial sample survey:
+\leavevmode\hypertarget{ref-brus2021statistical}{}%
+Brus, D.J., 2021. Statistical approaches for spatial sample survey:
 Persistent misconceptions and new developments. European Journal of Soil
-Science.
+Science 72, 686--703.
 
@@
 Cooper, C., 2006. Sampling and variance estimation on continuous
-domains. Environmetrics: The official journal of the International
-Environmetrics Society 17, 539--553.
+domains. Environmetrics 17, 539--553.
 
@@
 samples: A reappraisal of classical sampling theory. Mathematical
-geology 22, 407--415.
+Geology 22, 407--415.
 
@@
 
-\leavevmode\hypertarget{ref-fix1951discriminatory}{}%
-Fix, E., Hodges, J.L., 1951. Discriminatory analysis, nonparametric
-discrimination: Consistency properties. USAF School of Aviation
-Medicine.
+\leavevmode\hypertarget{ref-fix1989discriminatory}{}%
+Fix, E., Hodges, J.L., 1989. Discriminatory analysis. Nonparametric
+discrimination: Consistency properties. International Statistical
+Review/Revue Internationale de Statistique 57, 238--247.
 
@@
 component estimation and to related problems. Journal of the American
-statistical association 72, 320--338.
+Statistical Association 72, 320--338.
 
@@
 without replacement from a finite universe. Journal of the American
-statistical Association 47, 663--685.
+Statistical Association 47, 663--685.
 
@@
 for inference from samples to populations: From polarization to
-integration. Multivariate behavioral research 44, 711--740.
+integration. Multivariate Behavioral Research 44, 711--740.
 
@@
 Stevens, D.L., Olsen, A.R., 2004. Spatially balanced sampling of natural
-resources. Journal of the american Statistical association 99, 262--278.
+resources. Journal of the American Statistical Association 99, 262--278.
 
@@
 Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
-model to nearest neighbor (k-NN) methods for forestry applications. PloS
-one 8, e59129.
+model to nearest neighbor (k-NN) methods for forestry applications. PlOS
+ONE 8, e59129.
 
@@
 Liu, T.-J., Meng, B., 2013. Design-based spatial sampling: Theory and
-implementation. Environmental modelling \& software 40, 280--288.
+implementation. Environmental Modelling \& Software 40, 280--288.
 

### 2021-12-02 inst/manuscript/manuscript.Rmd
added internal update edits

@@
 In this manuscript, we compare the approaches for finite population spatial data.
-We first provide relevant background for design-based and model-based approaches to finite population patial data.
-Then we use a simulation study and an analysis of real mercurcy concentration data to compare them numerically.
+We first provide relevant background for design-based and model-based approaches to finite population spatial data.
+Then we use a simulation study and an analysis of real mercury concentration data to compare them numerically.
 We find that sampling plans that incorporate spatial locations (spatially balanced samples) perform better than sampling plans ignoring spatial locations (non-spatially balanced samples), regardless of whether design-based or model-based approaches were used to analyze the data.
 We also find that within sampling plans, model-based approaches tend to outperform design-based approaches, even for skewed data.
-This gap in performance is small when spatially samples are used but large when non-spatially balanced samples are used.
+This gap in performance is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
 
@@
 In this manuscript, we use the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced samples sampling because the algorithm has several attractive properties.
-It accommodates finite and infinite sampling frames.
-It accommodates equal, unequal, and proportional (to size) inclusion probabilities.
-It accommodates legacy (historical) sampling [@foster2017spatially].
-It accommodates a minimum distance between units in a sample.
-Lastly, it accommodates replacement units in a sample, which are units that can be sampled in place of an original unit that can no longer be sampled.
+It accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units in a sample (which are units that can be sampled in place of an original unit that can no longer be sampled).
 The GRTS algorithm samples from finite and infinite populations by utilizing a mapping between two-dimensional and one-dimensional space.
@@
 where $Z_i$ is the value of the $i$th unit in the sample and $\pi_i$ is the inclusion probability of the $i$th unit in the sample.
-An estimate of the population mean can be obtained by dividing $\hat{\tau}_{ht}$ by number of population units.
+An estimate of the population mean can be obtained by dividing $\hat{\tau}_{ht}$ by the number of population units, $N$.
 
@@
 
-Let $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for teh sampled and unsampled population units, respectively.
-We assume $E(\bm{\delta} = \mathbf{0}$ and that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
+Let $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
+We assume $E(\bm{\delta}) = \mathbf{0}$ and that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
 It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th elment of hte matrix $\cov(\bm{\delta})$ is
+Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th elment of the matrix $\cov(\bm{\delta})$ is
 \mbox{}
@@
 With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
-While details of the derivation are in [@verhoef2008spatial], we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
+While details of the derivation are in @verhoef2008spatial, we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
 
@@
 
-Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" is the baseline.
+Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
 More formally, the relative rMS(P)E is defined as
@@
 colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
+psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
 effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
                               colour = approach)) +
-  facet_grid(psill_ratio ~ resptype) +
+  facet_grid(
+    psill_ratio ~ resptype,
+    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
+  ) +
   geom_jitter(width = 0.24, size = 2.5) +
@@
   geom_jitter(width = 0.24, size = 2.5) + 
-  facet_grid(psill_ratio ~ resptype) +
+  facet_grid(
+    psill_ratio ~ resptype,
+    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
+  ) +
   scale_colour_manual(values = colour_scale) +
@@
 The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to study mercury concentration.
+We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality in lakes in the contiguous United States, to study mercury concentration.
 Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
@@
 
+# Acknowledgements {.unnumbered}
+
+The views expressed in this manuscript are those of the authors and do not necessarily represent the views or policies of the U.S.
+Environmental Protection Agency.
+Any mention of trade names, products, or services does not imply an endorsement by the U.S.
+government or the U.S.
+Environmental Protection Agency.
+The U.S.
+Environmental Protection Agency does not endorse any commercial products, services, or enterprises.
+
 # References {#references .unnumbered}

### 2021-12-02 inst/manuscript/manuscript.tex
added internal update edits

@@
 \usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 
-  \title{A comparison of design-based and model-based approaches for
-finite population spatial data.}
+  \title{A comparison of design-based and model-based approaches for finite
+population spatial data.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200
-SW 35th St, Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics,
-Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
-13617}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
-Center, National Oceanic and Atmospheric Administration, Seattle,
-Washington, 98115}
-    \address[OSU]{Oregon State University Department of Statistics, 239
-Weniger Hall, Corvallis, Oregon, 97331}
-      \cortext[1]{Corresponding Author: Michael Dumelle
-(Dumelle.Michael@epa.gov)}
+      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
+Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
+and Statistics, 23 Romoda Drive, Canton, New York, 13617}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
+Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
+    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
+Corvallis, Oregon, 97331}
+      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}
 
@@
   In this manuscript, we compare the approaches for finite population
-  spatial data. We first provide relevant background for design-based
-  and model-based approaches to finite population patial data. Then we
-  use a simulation study and an analysis of real mercurcy concentration
-  data to compare them numerically. We find that sampling plans that
-  incorporate spatial locations (spatially balanced samples) perform
-  better than sampling plans ignoring spatial locations (non-spatially
-  balanced samples), regardless of whether design-based or model-based
-  approaches were used to analyze the data. We also find that within
-  sampling plans, model-based approaches tend to outperform design-based
-  approaches, even for skewed data. This gap in performance is small
-  when spatially samples are used but large when non-spatially balanced
-  samples are used.
+  spatial data. We first provide relevant background for design-based and
+  model-based approaches to finite population spatial data. Then we use a
+  simulation study and an analysis of real mercury concentration data to
+  compare them numerically. We find that sampling plans that incorporate
+  spatial locations (spatially balanced samples) perform better than
+  sampling plans ignoring spatial locations (non-spatially balanced
+  samples), regardless of whether design-based or model-based approaches
+  were used to analyze the data. We also find that within sampling plans,
+  model-based approaches tend to outperform design-based approaches, even
+  for skewed data. This gap in performance is small when spatially
+  balanced samples are used but large when non-spatially balanced samples
+  are used.
   \end{abstract}
@@
 comparisons between design-based and model-based approaches for spatial
-data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002; Ver Hoef,
-2008; Wang et al., 2012). Cooper (2006) reviews the two approaches in an
-ecological context before introducing a ``model-assisted'' variance
-estimator that combines aspects from each approach. In addition to
-Cooper (2006), there has been substantial research and development into
-estimators that use both design and model-based principles (see e.g.,
-Sterba (2009), Cicchitelli and Montanari (2012), Chan-Golston et al.
-(2020) for a Bayesian approach).
+data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang
+et al., 2012). Cooper (2006) reviews the two approaches in an ecological
+context before introducing a ``model-assisted'' variance estimator that
+combines aspects from each approach. In addition to Cooper (2006), there
+has been substantial research and development into estimators that use
+both design and model-based principles (see e.g., Sterba (2009),
+Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
+Bayesian approach).
 
@@
 sample. An estimate of the population mean can be obtained by dividing
-\(\hat{\tau}_{ht}\) by number of population units.
+\(\hat{\tau}_{ht}\) by the number of population units, \(N\).
 
@@
 Let \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
-\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for teh
+\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
 sampled and unsampled population units, respectively. We assume
-\(E(\bm{\delta} = \mathbf{0}\) and that there is spatial correlation in
+\(E(\bm{\delta}) = \mathbf{0}\) and that there is spatial correlation in
 \(\bm{\delta}\) that can be modeled using a covariance function. It is
@@
 the simulations and applications in this manuscript is the exponential
-covariance function: the \(i,j\)th elment of hte matrix
+covariance function: the \(i,j\)th elment of the matrix
 \(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
@@
 (BLUP) for \(f(\mathbf{b}'\mathbf{z})\) and its prediction variance can
-be computed. While details of the derivation are in (Ver Hoef, 2008), we
+be computed. While details of the derivation are in Ver Hoef (2008), we
 note here that the predictor and its variance are both moment-based,
@@
 approaches from Table \ref{tab:designanalysis} using the random location
-layout with ``IRS-Design'' is the baseline. More formally, the relative
+layout with ``IRS-Design'' as the baseline. More formally, the relative
 rMS(P)E is defined as \begin{equation*}
@@
 
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+The views expressed in this manuscript are those of the authors and do
+not necessarily represent the views or policies of the U.S.
+Environmental Protection Agency. Any mention of trade names, products,
+or services does not imply an endorsement by the U.S. government or the
+U.S. Environmental Protection Agency. The U.S. Environmental Protection
+Agency does not endorse any commercial products, services, or
+enterprises.
+
 \hypertarget{references}{%
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-akaike1974new}{}%
@@
 \leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
-Diggle, P.J., Menezes, R., Su, T., 2010. Geostatistical inference under
-preferential sampling. Journal of the Royal Statistical Society: Series
-C (Applied Statistics) 59, 191--232.
+Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
+under preferential sampling. Journal of the Royal Statistical Society:
+Series C (Applied Statistics) 59, 191--232.
 
@@
 Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
-model to nearest neighbor (k-NN) methods for forestry applications. PlOS
+model to nearest neighbor (k-nn) methods for forestry applications. PlOS
 ONE 8, e59129.

### 2021-12-04 inst/manuscript/manuscript.Rmd
abstract and introduction edits

@@
 abstract: |
-  The design-based and model-based approaches to frequentist statistical inference rest on fundamentally different foundations.
+  The design-based and model-based approaches to frequentist statistical inference lie on fundamentally different foundations.
 In the design-based approach, inference depends on random sampling.
@@
 In this manuscript, we compare the approaches for finite population spatial data.
-We first provide relevant background for design-based and model-based approaches to finite population spatial data.
-Then we use a simulation study and an analysis of real mercury concentration data to compare them numerically.
+We first provide relevant background for the approaches and then use a simulation study and an analysis of real mercury concentration data to numerically compare them.
 We find that sampling plans that incorporate spatial locations (spatially balanced samples) perform better than sampling plans ignoring spatial locations (non-spatially balanced samples), regardless of whether design-based or model-based approaches were used to analyze the data.
-We also find that within sampling plans, model-based approaches tend to outperform design-based approaches, even for skewed data.
+We also find that within sampling plans, the model-based approaches often outperform design-based approaches, even for skewed data.
 This gap in performance is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
@@
 There are two general approaches for using data to make frequentist statistical inferences about a population: design-based and model-based.
-When data cannot be collected for all units in a population (population units), data are collected on a subset of the population units.
+When data cannot be collected for all units in a population (i.e., population units), data are collected on a subset of the population units.
 This subset is called a sample.
-In the design-based approach, inferences about the underlying population are informed via a probabilistic process assigning some population units to the sample.
+In the design-based approach, inferences about the underlying population are informed via a probabilistic process assigning some population units to be part of the sample.
 Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process generating the data.
@@
 <!-- Paragraph about the contribution of the manuscript and outlining the rest -->
-Though comparisons between design-based and model-based approaches to spatial data have been studied, no numerical comparison has been made between design-based approaches that incorporate spatial locations and model-based approaches.
-In this manuscript, we compare design-based approaches that incorporate spatial locations to model-based approaches for spatial data.
+Certainly comparisons between design-based and model-based approaches to spatial data have been studied.
+But no numerical comparison has been made between design-based approaches incorporating spatial information and design-based approaches.
+In this manuscript, we compare design-based approaches incorporating spatial information to model-based approaches for spatial data.
 We focus on finite populations, but these comparisons generalize to infinite populations as well.
@@
 The rest of the manuscript is organized as follows.
-In Section \ref{sec:background}, we introduce and compare several sampling and estimation procedures of the design-based and model-based approaches for finite population spatial data.
-In Section \ref{sec:numstudy}, we use a simulation approach to study the behavior and performance of both approaches.
-In Section \ref{application}, we use both approaches to analyze real data consisting of mercury concentration from lakes in the contiguous United States.
+In Section \ref{sec:background}, we introduce and provide relevant background for the design-based and model-based approaches to finite population spatial data.
+In Section \ref{sec:numstudy}, we use a simulation study to compare the performance of the approaches in a variety of scenarios.
+In Section \ref{application}, we compare the performance of the approaches on real data that contains mercury concentration in lakes from the contiguous United States.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.

### 2021-12-04 inst/manuscript/manuscript.tex
abstract and introduction edits

@@
   The design-based and model-based approaches to frequentist statistical
-  inference rest on fundamentally different foundations. In the
+  inference lie on fundamentally different foundations. In the
   design-based approach, inference depends on random sampling. In the
@@
   In this manuscript, we compare the approaches for finite population
-  spatial data. We first provide relevant background for design-based and
-  model-based approaches to finite population spatial data. Then we use a
-  simulation study and an analysis of real mercury concentration data to
-  compare them numerically. We find that sampling plans that incorporate
-  spatial locations (spatially balanced samples) perform better than
-  sampling plans ignoring spatial locations (non-spatially balanced
-  samples), regardless of whether design-based or model-based approaches
-  were used to analyze the data. We also find that within sampling plans,
-  model-based approaches tend to outperform design-based approaches, even
-  for skewed data. This gap in performance is small when spatially
-  balanced samples are used but large when non-spatially balanced samples
-  are used.
+  spatial data. We first provide relevant background for the approaches
+  and then use a simulation study and an analysis of real mercury
+  concentration data to numerically compare them. We find that sampling
+  plans that incorporate spatial locations (spatially balanced samples)
+  perform better than sampling plans ignoring spatial locations
+  (non-spatially balanced samples), regardless of whether design-based or
+  model-based approaches were used to analyze the data. We also find that
+  within sampling plans, the model-based approaches often outperform
+  design-based approaches, even for skewed data. This gap in performance
+  is small when spatially balanced samples are used but large when
+  non-spatially balanced samples are used.
   \end{abstract}
@@
 statistical inferences about a population: design-based and model-based.
-When data cannot be collected for all units in a population (population
-units), data are collected on a subset of the population units. This
-subset is called a sample. In the design-based approach, inferences
-about the underlying population are informed via a probabilistic process
-assigning some population units to the sample. Alternatively, in the
-model-based approach, inferences are made from specific assumptions
-about the underlying process generating the data. Each paradigm has a
-deep historical context (Sterba, 2009) and its own set of benefits and
-drawbacks (Hansen et al., 1983).
+When data cannot be collected for all units in a population (i.e.,
+population units), data are collected on a subset of the population
+units. This subset is called a sample. In the design-based approach,
+inferences about the underlying population are informed via a
+probabilistic process assigning some population units to be part of the
+sample. Alternatively, in the model-based approach, inferences are made
+from specific assumptions about the underlying process generating the
+data. Each paradigm has a deep historical context (Sterba, 2009) and its
+own set of benefits and drawbacks (Hansen et al., 1983).
 
@@
 
-Though comparisons between design-based and model-based approaches to
-spatial data have been studied, no numerical comparison has been made
-between design-based approaches that incorporate spatial locations and
-model-based approaches. In this manuscript, we compare design-based
-approaches that incorporate spatial locations to model-based approaches
+Certainly comparisons between design-based and model-based approaches to
+spatial data have been studied. But no numerical comparison has been
+made between design-based approaches incorporating spatial information
+and design-based approaches. In this manuscript, we compare design-based
+approaches incorporating spatial information to model-based approaches
 for spatial data. We focus on finite populations, but these comparisons
@@
 The rest of the manuscript is organized as follows. In Section
-\ref{sec:background}, we introduce and compare several sampling and
-estimation procedures of the design-based and model-based approaches for
-finite population spatial data. In Section \ref{sec:numstudy}, we use a
-simulation approach to study the behavior and performance of both
-approaches. In Section \ref{application}, we use both approaches to
-analyze real data consisting of mercury concentration from lakes in the
+\ref{sec:background}, we introduce and provide relevant background for
+the design-based and model-based approaches to finite population spatial
+data. In Section \ref{sec:numstudy}, we use a simulation study to
+compare the performance of the approaches in a variety of scenarios. In
+Section \ref{application}, we compare the performance of the approaches
+on real data that contains mercury concentration in lakes from the
 contiguous United States. And in Section \ref{sec:discussion}, we end
@@
 the algorithm has several attractive properties. It accommodates finite
-and infinite sampling frames. It accommodates equal, unequal, and
-proportional (to size) inclusion probabilities. It accommodates legacy
-(historical) sampling (Foster et al., 2017). It accommodates a minimum
-distance between units in a sample. Lastly, it accommodates replacement
-units in a sample, which are units that can be sampled in place of an
-original unit that can no longer be sampled. The GRTS algorithm samples
+and infinite sampling frames, equal, unequal, and proportional (to size)
+inclusion probabilities, legacy (historical) sampling (Foster et al.,
+2017), a minimum distance between units in a sample, and replacement
+units in a sample (which are units that can be sampled in place of an
+original unit that can no longer be sampled). The GRTS algorithm samples
 from finite and infinite populations by utilizing a mapping between
@@
 will use the 2012 National Lakes Assessment (NLA), which measures
-various aspects of lake health and quality in lakes in the contiguous
-United States, to study mercury concentration. Although we know the true
-mean mercury concentration values for the 986 lakes from the 2012 NLA,
-we will explore whether or not we obtain an adequately precise estimate
-for the realized mean mercury concentration if we sample only 100 of the
-986 lakes.
+various aspects of lake health and water quality in lakes in the
+contiguous United States, to study mercury concentration. Although we
+know the true mean mercury concentration values for the 986 lakes from
+the 2012 NLA, we will explore whether or not we obtain an adequately
+precise estimate for the realized mean mercury concentration if we
+sample only 100 of the 986 lakes.
 

### 2021-12-04 inst/manuscript/manuscript.Rmd
background edits

@@
 The design-based and model-based approaches incorporate randomness in fundamentally different ways.
-In this section, we describe the role of randomness and its effects on subsequent inferences.
-We then discuss specific inference methods of the approaches for spatial data.
+In this section, we describe the role of randomness for each approach and the subsequent effects on statistical inferences for spatial data.
 
@@
 The design-based approach assumes the population is fixed.
-Randomness is incorporated via the selection of units in a smapling frame according to a sampling design.
+Randomness is incorporated via the selection of units in a sampling frame.
 A sampling frame is the set of all units available to be sampled.
-A sampling design assigns a positive probability of inclusion (inclusion probability) to each unit in the sampling frame.
+Units from the sampling frame are selected as part of the sample according to a sampling design, which assigns a positive probability of inclusion (inclusion probability) to each unit from the sampling frame.
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
-If a sampling design selects units from the sampling frame while ignoring their spatial locations, we call them "Independent Random Sampling" (IRS) designs.
-If a sampling design selects units from the sampling frame while incorporating their spatial locations, we call them spatially balanced designs.
-Spatially balanced designs can be obtained using the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
-The design-based approach combines the randomness of the sampling design and the data collected via the sample to estimate fixed, unknown parameters (e.g., means and totals) of a population.
+When sampling designs incorporate spatial locations into sampling, we call the resulting samples "spatially balanced." One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
+When sampling designs do not incorporate spatial locations into sampling, we call the resulting samples "non-spatially balanced."
 
+Fundamentally, the design-based approach combines the randomness of the sampling design with the data collected via the sample to justify the estimation and uncertainty quantification of fixed, unknown parameters of a population (e.g., a population mean).
 Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design.
-Means and totals, for example, are asymptotically normally distributed (normal) by the Central Limit Theorem (under some assumptions).
-If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
+Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible samples.
+Sample means, for example, are asymptotically normal (Gaussian) by the Central Limit Theorem (under some assumptions).
+If we repeatedly select samples from the population, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
 @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
@@
 Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
-To guard against preferential sampling, model-based approaches often still implement random sampling.
+To guard against preferential sampling, model-based approaches often still implement some form of random sampling.
 
-Instead of estimating fixed but unknown parameters like a mean or total (as in the design-based approach), the goal of model-based inference in the spatial context is often to predict a realized variable, or value.
+Instead of estimating fixed, unknown population parameters, as in the design-based approach, often the goal of model-based inference is to predict a realized variable, or value.
 For example, suppose the realized mean of all population units is the value of interest.
 Instead of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the value of the mean, a random variable.
-Prediction intervals are then derived using assumptions of the data generating process.
-If we repeatedly generate the response values from the same spatial process and sample, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
-@cressie1993statistics and @schabenberger2017statistical provide reviews of model-based approaches for spatial data.
-A visual comparison of the design-based and model-based assumptions is provided in Figure \ref{fig:fig1} (@verhoef2002sampling and @brus2021statistical provide similar figures).
+Prediction intervals are then derived using assumptions of the data-generating stochastic process.
+If we repeatedly generate response values from the same data-generating stochastic process and select samples, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
+@cressie1993statistics and @schabenberger2017statistical provide thorough reviews of model-based approaches for spatial data.
+In Figure \ref{fig:fig1}, we provide a visual comparison of the design-based and model-based approaches (@verhoef2002sampling and @brus2021statistical provide similar figures).
 
-```{r fig1, out.width = "100%", fig.cap = "A comparison of sampling under the design-based and model-based frameworks.
-Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of size four.
-The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
-The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels."}
+```{r fig1, out.width = "100%", fig.cap = "A visual comparison of the design-based and model-based approaches.
+In the top row, there is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
+The response values at each site are fixed, but we obtain different estimates for the mean response in each random sample.
+In the bottom row, there are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
+The ata-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations."}
 set.seed(10102021)
@@
 
-The design-based approach can be used to select samples that are "well-spread" in space, or spatially balanced.
+We previously mentioned that the design-based approach can be used to select spatially balanced samples (samples that incorporate spatial locations of the population units and are "well-spread" is space).
 Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-The first spatially balanced sampling algorithm that saw widespread use was the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
-To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voroni polygons.
-After the GRTS algorithm was developed, several other spatially balanced sampling algorithms have emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we use the Generalized Random Tessellation Stratified (GRTS) algorithm to select spatially balanced samples sampling because the algorithm has several attractive properties.
-It accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units in a sample (which are units that can be sampled in place of an original unit that can no longer be sampled).
-The GRTS algorithm samples from finite and infinite populations by utilizing a mapping between two-dimensional and one-dimensional space.
-The units in the two-dimensional sampling frame are divided into cells using a hierarchical address.
-This hierarchical address is then used to map the units from two-dimensional space to a one-dimensional line where each unit's line length equals its inclusion probability.
-A systematic sample is conducted on the line and linked back to a unit in two-dimensional space, which results in the desired sample.
-@stevens2004spatially provides further details.
+The first spatially balanced sampling algorithm seeing widespread use is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
+To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Direchlet Tessellations).
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, such as the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
+In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it has several attractive properties.
+More specifically, the GRTS algorithm accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
+The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
+Via this mapping, units in two-dimensional space are partitioned using a hierarchical address.
+This hierarchical address is used to map population units to a one-dimensional line.
+On the one dimensional line, each population unit's line length equals its inclusion probability.
+Then, a systematic sample of population units is selected on the line, yielding desired sample.
+@stevens2004spatially provides more technical details.
 
@@
 
-After selecting a spatially balanced sample using the GRTS algorithm (i.e., a GRTS sample), data are collected and used to estimate population parameters.
-To unbiasedly estimate population means and totals from sample data, one can use the Horvitz-Thompson estimator [@horvitz1952generalization].
+After selecting a sample and collecting data, unbiased estimates of population means and totals can be obtained using the Horvitz-Thompson estimator [@horvitz1952generalization].
 If $\tau$ is a population total, the Horvitz-Thompson estimate of $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
@@
 \end{align}
-where $Z_i$ is the value of the $i$th unit in the sample and $\pi_i$ is the inclusion probability of the $i$th unit in the sample.
-An estimate of the population mean can be obtained by dividing $\hat{\tau}_{ht}$ by the number of population units, $N$.
+where $Z_i$ is the value of the $i$th population unit in the sample and $\pi_i$ is the inclusion probability of the $i$th population unit in the sample.
+An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by $N$, the number of population units.
 
-While the Horvitz-Thompson estimator is unbiased for population means and totals, it is also important to quantify the uncertainty in these estimates.
+It is also important to quantify uncertainty $\hat{\tau}_{ht}$.
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that unit $i$ and unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
-Second, these estimators ignore the spatial locations of the units in the sampling frame.
+First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and populatoin unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
+Second, these estimators ignore the spatial locations of the population units.
 To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ compared to variance estimators ignoring spatial locations, yielding narrower confidence intervals for $\tau$.
+@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ and yield narrower confidence intervals compared to variance estimators that ignore spatial locations.
 
@@
 Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
-Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial process.
-@verhoef2008spatial gives details on the theory of FPBK, but some of the basic principles are summarized below.
-Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
+Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
+We summarize some of the basic principles of FBPK next (for more technical details, see @verhoef2008spatial) Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
 .
@@
 , $s_N$ that can be measured at the $N$ population units.
-Suppose we want to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights.
-For example, if we want to predict the population total across all population units, then we would use a vector of 1's for the weights.
-
-We often only have a sample of the $N$ population units.
+Suppose we want to use a sample to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights (e.g, the population mean is represented by a weights vector whose elements all equal one).
 Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with subscript \emph{u}, let
@@
 \end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, and $\bm{\beta}$ is the parameter vector of fixed effects.
+where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, $\bm{\beta}$ is the parameter vector of fixed effects, and $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
 
-Let $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
-We assume $E(\bm{\delta}) = \mathbf{0}$ and that there is spatial correlation in $\bm{\delta}$ that can be modeled using a covariance function.
-It is common to assume the covariance function is second-order stationary and isotropic [@cressie1993statistics], and that the spatial covariance decreases as the separation between population units increases.
-Many spatial covariance functions exist, but the primary function we use throughout the simulations and applications in this manuscript is the exponential covariance function: the $i,j$th elment of the matrix $\cov(\bm{\delta})$ is
+FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean zero and a spatial correlation structure that can be modeled using a covariance function.
+This covariance function is commonly assumed to be non-negative (between zero and one), second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
+Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$, though @chiles1999geostatistics, pp.
+80-93 discuss covariance functions that are not second-order stationary, not isotropic, or both.
+A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics]; one example is the exponential covariance function (for a thorough list of spatial covariance functions, see @cressie1993statistics.
+The $i,j$th element of the exponential covariance matrix, $\cov(\bm{\delta})$, is
 \mbox{}
@@
 \end{align}
-where $\sigma^2_{1}$ is dependent random error variance measuring coarse-scale (correlated) variability, $\sigma^2_{2}$ is the independent random error variance measuring fine-scale (independent) variability, $\phi$ is the range parameter measuring the distance-decay rate of the correlation, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-Often $\sigma^2_{1}$ and $\sigma^2_{2}$ are called the partial sill and nugget, respectively.
-Any spatial covariance function could be used in the place of the exponential, including functions that allow for non-stationarity or anisotropy [@chiles1999geostatistics, pp.
-80-93].
+where $\sigma^2_{1}$ is the variance parameter quantifying the variability that is dependent (coarse-scale), $\sigma^2_{2}$ is the variance parameter quantifying the variability that is independent (fine-scale), $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+The proportion of variability attributable to dependent random error is $\sigma^2_{1} / (\sigma^2_{1} + \sigma^2_{2})$.
+Similarly, the proportion of variability attributable to independent random error is $\sigma^2_{2} / (\sigma^2_{1} + \sigma^2_{2})$.
+Finally we note that $\sigma^2_{1}$ and $\sigma^2_{2}$ are often called the partial sill and nugget, respectively.
 
@@
 
-We note that we only use FPBK in this paper in order to focus more on comparing the design-based and model-based approaches.
-Other methods, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison], random forest [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
-We choose to use FPBK because it is faster than a Bayesian approach and it was developed with theoretically-based variance estimators of means and totals for spatial data, whereas random forests and k-nearest-neighbors use ad-hoc variance estimators in most cases [@ver2013comparison]; additionally, FBPK outperformed the other methods in most scenarios.
+Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison], random forests [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
+Compared to the k-nearest-neighbors and random forest approach, we prefer FBPK because it is model-based and relies on theoretically-based variance estimators leveraging the model's spatial covariance structure, whereas k-nearest-neighbors and random forests use ad-hoc variance estimators [@ver2013comparison].
+Additionally, @ver2013comparison studied compared FBPK, k-nearest-neighbors, and random forests in a variety of spatial data contexts, and FBPK tended to perform best.
+Compared to the Bayesian approach, we prefer FPBK mostly because it is much more computationally efficient.
 

### 2021-12-04 inst/manuscript/manuscript.tex
background edits

@@
 fundamentally different ways. In this section, we describe the role of
-randomness and its effects on subsequent inferences. We then discuss
-specific inference methods of the approaches for spatial data.
+randomness for each approach and the subsequent effects on statistical
+inferences for spatial data.
 
@@
 The design-based approach assumes the population is fixed. Randomness is
-incorporated via the selection of units in a smapling frame according to
-a sampling design. A sampling frame is the set of all units available to
-be sampled. A sampling design assigns a positive probability of
-inclusion (inclusion probability) to each unit in the sampling frame.
-Some examples of commonly used sampling designs include simple random
-sampling, stratified random sampling, and cluster sampling. If a
-sampling design selects units from the sampling frame while ignoring
-their spatial locations, we call them ``Independent Random Sampling''
-(IRS) designs. If a sampling design selects units from the sampling
-frame while incorporating their spatial locations, we call them
-spatially balanced designs. Spatially balanced designs can be obtained
-using the Generalized Random Tessellation Stratified (GRTS) algorithm
-(Stevens and Olsen, 2004), which we discuss in more detail in Section
-\ref{subsec:spb_design}. The design-based approach combines the
-randomness of the sampling design and the data collected via the sample
-to estimate fixed, unknown parameters (e.g., means and totals) of a
-population.
+incorporated via the selection of units in a sampling frame. A sampling
+frame is the set of all units available to be sampled. Units from the
+sampling frame are selected as part of the sample according to a
+sampling design, which assigns a positive probability of inclusion
+(inclusion probability) to each unit from the sampling frame. Some
+examples of commonly used sampling designs include simple random
+sampling, stratified random sampling, and cluster sampling. When
+sampling designs incorporate spatial locations into sampling, we call
+the resulting samples ``spatially balanced.'' One approach to selecting
+spatially balanced samples is the Generalized Random Tessellation
+Stratified (GRTS) algorithm (Stevens and Olsen, 2004), which we discuss
+in more detail in Section \ref{subsec:spb_design}. When sampling designs
+do not incorporate spatial locations into sampling, we call the
+resulting samples ``non-spatially balanced.''
 
-Treating the data as fixed and incorporating randomness through the
-sampling design yields estimators having very few other assumptions.
-Confidence intervals for these types of estimators are typically derived
-using limiting arguments that incorporate all possible randomizations of
-sampling units selected via the sampling design. Means and totals, for
-example, are asymptotically normally distributed (normal) by the Central
-Limit Theorem (under some assumptions). If we repeatedly sample the
-surface, then 95\% of all 95\% confidence intervals constructed from a
-procedure with appropriate coverage will contain the true, fixed mean.
-Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
-design-based approach.
+Fundamentally, the design-based approach combines the randomness of the
+sampling design with the data collected via the sample to justify the
+estimation and uncertainty quantification of fixed, unknown parameters
+of a population (e.g., a population mean). Treating the data as fixed
+and incorporating randomness through the sampling design yields
+estimators having very few other assumptions. Confidence intervals for
+these types of estimators are typically derived using limiting arguments
+that incorporate all possible samples. Sample means, for example, are
+asymptotically normal (Gaussian) by the Central Limit Theorem (under
+some assumptions). If we repeatedly select samples from the population,
+then 95\% of all 95\% confidence intervals constructed from a procedure
+with appropriate coverage will contain the true, fixed mean. Särndal et
+al. (2003) and Lohr (2009) provide thorough reviews of the design-based
+approach.
 
@@
 against preferential sampling, model-based approaches often still
-implement random sampling.
+implement some form of random sampling.
 
-Instead of estimating fixed but unknown parameters like a mean or total
-(as in the design-based approach), the goal of model-based inference in
-the spatial context is often to predict a realized variable, or value.
-For example, suppose the realized mean of all population units is the
-value of interest. Instead of \emph{estimating} a fixed, unknown mean,
-we are \emph{predicting} the value of the mean, a random variable.
-Prediction intervals are then derived using assumptions of the data
-generating process. If we repeatedly generate the response values from
-the same spatial process and sample, then 95\% of all 95\% prediction
+Instead of estimating fixed, unknown population parameters, as in the
+design-based approach, often the goal of model-based inference is to
+predict a realized variable, or value. For example, suppose the realized
+mean of all population units is the value of interest. Instead of
+\emph{estimating} a fixed, unknown mean, we are \emph{predicting} the
+value of the mean, a random variable. Prediction intervals are then
+derived using assumptions of the data-generating stochastic process. If
+we repeatedly generate response values from the same data-generating
+stochastic process and select samples, then 95\% of all 95\% prediction
 intervals constructed from a procedure with appropriate coverage will
 contain their respective realized means. Cressie (1993) and
-Schabenberger and Gotway (2017) provide reviews of model-based
-approaches for spatial data. A visual comparison of the design-based and
-model-based assumptions is provided in Figure \ref{fig:fig1} (Ver Hoef
-(2002) and Brus (2021) provide similar figures).
+Schabenberger and Gotway (2017) provide thorough reviews of model-based
+approaches for spatial data. In Figure \ref{fig:fig1}, we provide a
+visual comparison of the design-based and model-based approaches (Ver
+Hoef (2002) and Brus (2021) provide similar figures).
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/fig1-1} \caption{A comparison of sampling under the design-based and model-based frameworks.
-Points circled are those that are sampled.
-In the top row, we have one fixed population, and three random samples of size four.
-The response values at each site are fixed, but we obtain different estimates for the mean response because the randomly sampled sites vary from sample to sample.
-In the bottom row, we have three realizations of the same spatial process sampled at the same locations.
-The spatial process generating the response values has a single mean, but the realized mean is different in each of the three panels.}\label{fig:fig1}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/fig1-1} \caption{A visual comparison of the design-based and model-based approaches.
+In the top row, there is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
+The response values at each site are fixed, but we obtain different estimates for the mean response in each random sample.
+In the bottom row, there are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
+The ata-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations.}\label{fig:fig1}
 \end{figure}
@@
 
-The design-based approach can be used to select samples that are
-``well-spread'' in space, or spatially balanced. Spatially balanced
-samples are useful because parameter estimates from these samples tend
-to vary less than parameter estimates from samples that are not
-spatially balanced (Barabesi and Franceschi, 2011; Benedetti et al.,
-2017; Grafström and Lundström, 2013; Robertson et al., 2013; Stevens and
-Olsen, 2004; Wang et al., 2013). The first spatially balanced sampling
-algorithm that saw widespread use was the Generalized Random
-Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004). To
-quantify the spatial balance of a sample, Stevens and Olsen (2004)
-proposed loss metrics based on Voroni polygons. After the GRTS algorithm
-was developed, several other spatially balanced sampling algorithms have
-emerged, including the Local Pivotal Method (Grafström et al., 2012;
-Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
-(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
-2013), Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017),
-and Halton Iterative Partitioning Sampling (Robertson et al., 2018). In
-this manuscript, we use the Generalized Random Tessellation Stratified
-(GRTS) algorithm to select spatially balanced samples sampling because
-the algorithm has several attractive properties. It accommodates finite
-and infinite sampling frames, equal, unequal, and proportional (to size)
-inclusion probabilities, legacy (historical) sampling (Foster et al.,
-2017), a minimum distance between units in a sample, and replacement
-units in a sample (which are units that can be sampled in place of an
-original unit that can no longer be sampled). The GRTS algorithm samples
-from finite and infinite populations by utilizing a mapping between
-two-dimensional and one-dimensional space. The units in the
-two-dimensional sampling frame are divided into cells using a
-hierarchical address. This hierarchical address is then used to map the
-units from two-dimensional space to a one-dimensional line where each
-unit's line length equals its inclusion probability. A systematic sample
-is conducted on the line and linked back to a unit in two-dimensional
-space, which results in the desired sample. Stevens and Olsen (2004)
-provides further details.
+We previously mentioned that the design-based approach can be used to
+select spatially balanced samples (samples that incorporate spatial
+locations of the population units and are ``well-spread'' is space).
+Spatially balanced samples are useful because parameter estimates from
+these samples tend to vary less than parameter estimates from samples
+that are not spatially balanced (Barabesi and Franceschi, 2011;
+Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
+2013; Stevens and Olsen, 2004; Wang et al., 2013). The first spatially
+balanced sampling algorithm seeing widespread use is the Generalized
+Random Tessellation Stratified (GRTS) algorithm (Stevens and Olsen,
+2004). To quantify the spatial balance of a sample, Stevens and Olsen
+(2004) proposed loss metrics based on Voronoi polygons (Direchlet
+Tessellations). After the GRTS algorithm was developed, several other
+spatially balanced sampling algorithms emerged, such as the Local
+Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
+Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
+Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
+Sampling (Benedetti and Piersimoni, 2017), and Halton Iterative
+Partitioning Sampling (Robertson et al., 2018). In this manuscript, we
+select spatially balanced samples using the Generalized Random
+Tessellation Stratified (GRTS) algorithm because it has several
+attractive properties. More specifically, the GRTS algorithm
+accommodates finite and infinite sampling frames, equal, unequal, and
+proportional (to size) inclusion probabilities, legacy (historical)
+sampling (Foster et al., 2017), a minimum distance between units in a
+sample, and replacement units (replacement units are population units
+that can be sampled when a population unit originally selected can no
+longer be sampled). The GRTS algorithm selects samples by utilizing a
+particular mapping between two-dimensional and one-dimensional space
+that preserves proximity relationships. Via this mapping, units in
+two-dimensional space are partitioned using a hierarchical address. This
+hierarchical address is used to map population units to a
+one-dimensional line. On the one dimensional line, each population
+unit's line length equals its inclusion probability. Then, a systematic
+sample of population units is selected on the line, yielding desired
+sample. Stevens and Olsen (2004) provides more technical details.
 
-After selecting a spatially balanced sample using the GRTS algorithm
-(i.e., a GRTS sample), data are collected and used to estimate
-population parameters. To unbiasedly estimate population means and
-totals from sample data, one can use the Horvitz-Thompson estimator
-(Horvitz and Thompson, 1952). If \(\tau\) is a population total, the
-Horvitz-Thompson estimate of \(\tau\), denoted by \(\hat{\tau}_{ht}\),
-is is given by \begin{align}\label{eq:ht}
+After selecting a sample and collecting data, unbiased estimates of
+population means and totals can be obtained using the Horvitz-Thompson
+estimator (Horvitz and Thompson, 1952). If \(\tau\) is a population
+total, the Horvitz-Thompson estimate of \(\tau\), denoted by
+\(\hat{\tau}_{ht}\), is is given by \begin{align}\label{eq:ht}
   \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align} where \(Z_i\) is the value of the \(i\)th unit in the sample
-and \(\pi_i\) is the inclusion probability of the \(i\)th unit in the
-sample. An estimate of the population mean can be obtained by dividing
-\(\hat{\tau}_{ht}\) by the number of population units, \(N\).
+\end{align} where \(Z_i\) is the value of the \(i\)th population unit in
+the sample and \(\pi_i\) is the inclusion probability of the \(i\)th
+population unit in the sample. An estimate of the population mean is
+obtained by dividing \(\hat{\tau}_{ht}\) by \(N\), the number of
+population units.
 
-While the Horvitz-Thompson estimator is unbiased for population means
-and totals, it is also important to quantify the uncertainty in these
-estimates. Horvitz and Thompson (1952) and Sen (1953) provide variance
-estimators for \(\hat{\tau}_{ht}\), but these estimators have two
-drawbacks. First, they rely on calculating \(\pi_{ij}\), the probability
-that unit \(i\) and unit \(j\) are both in the sample -- this quantity
-can be challenging if not impossible to calculate analytically. Second,
-these estimators ignore the spatial locations of the units in the
-sampling frame. To address these two drawbacks simultaneously, Stevens
-and Olsen (2003) proposed the local neighborhood variance estimator. The
-local neighborhood variance estimator does not rely on \(\pi_{ij}\) and
+It is also important to quantify uncertainty \(\hat{\tau}_{ht}\).
+Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
+for \(\hat{\tau}_{ht}\), but these estimators have two drawbacks. First,
+they rely on calculating \(\pi_{ij}\), the probability that population
+unit \(i\) and populatoin unit \(j\) are both in the sample -- this
+quantity can be challenging if not impossible to calculate analytically.
+Second, these estimators ignore the spatial locations of the population
+units. To address these two drawbacks simultaneously, Stevens and Olsen
+(2003) proposed the local neighborhood variance estimator. The local
+neighborhood variance estimator does not rely on \(\pi_{ij}\) and
 incorporates spatial locations -- for technical details see Stevens and
@@
 variance estimator tends to reduce the estimated variance of
-\(\hat{\tau}\) compared to variance estimators ignoring spatial
-locations, yielding narrower confidence intervals for \(\tau\).
+\(\hat{\tau}\) and yield narrower confidence intervals compared to
+variance estimators that ignore spatial locations.
 
@@
 specific sampling design, we assume the data are generated by a spatial
-process. Ver Hoef (2008) gives details on the theory of FPBK, but some
-of the basic principles are summarized below. Let
+stochastic process. We summarize some of the basic principles of FBPK
+next (for more technical details, see Ver Hoef (2008)) Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 . , \(s_N\) that can be measured at the \(N\) population units. Suppose
-we want to predict some linear function of the response variable,
-\(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
-\(\mathbf{b}^\prime\) is a \(1 \times N\) vector of weights. For
-example, if we want to predict the population total across all
-population units, then we would use a vector of 1's for the weights.
-
-We often only have a sample of the \(N\) population units. Denoting
-quantities that are part of the sampled population units with a
-subscript \emph{s} and quantities that are part of the unsampled
-population units with subscript \emph{u}, let
+we want to use a sample to predict some linear function of the response
+variable, \(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
+\(\mathbf{b}^\prime\) is a \(1 \times N\) vector of weights (e.g, the
+population mean is represented by a weights vector whose elements all
+equal one). Denoting quantities that are part of the sampled population
+units with a subscript \emph{s} and quantities that are part of the
+unsampled population units with subscript \emph{u}, let
 
@@
 design matrices for the sampled and unsampled population units,
-respectively, and \(\bm{\beta}\) is the parameter vector of fixed
-effects.
+respectively, \(\bm{\beta}\) is the parameter vector of fixed effects,
+and \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
+\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
+sampled and unsampled population units, respectively.
 
-Let \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
-\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
-sampled and unsampled population units, respectively. We assume
-\(E(\bm{\delta}) = \mathbf{0}\) and that there is spatial correlation in
-\(\bm{\delta}\) that can be modeled using a covariance function. It is
-common to assume the covariance function is second-order stationary and
-isotropic (Cressie, 1993), and that the spatial covariance decreases as
-the separation between population units increases. Many spatial
-covariance functions exist, but the primary function we use throughout
-the simulations and applications in this manuscript is the exponential
-covariance function: the \(i,j\)th elment of the matrix
-\(\mathop{\mathrm{{cov}}}(\bm{\delta})\) is \mbox{}
+FBPK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
+has mean zero and a spatial correlation structure that can be modeled
+using a covariance function. This covariance function is commonly
+assumed to be non-negative (between zero and one), second-order
+stationary (depending only on the distance between population units),
+isotropic (independent of direction), and decay with distance between
+population units (Cressie, 1993). Henceforth, it is implied that we have
+made these same assumptions regarding \(\bm{\delta}\), though Chiles and
+Delfiner (1999), pp.~80-93 discuss covariance functions that are not
+second-order stationary, not isotropic, or both. A variety of flexible
+covariance functions can be used to model \(\bm{\delta}\) (Cressie,
+1993); one example is the exponential covariance function (for a
+thorough list of spatial covariance functions, see Cressie (1993). The
+\(i,j\)th element of the exponential covariance matrix,
+\(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
 \begin{align}\label{equation:expcov}
@@
 ,
-\end{align} where \(\sigma^2_{1}\) is dependent random error variance
-measuring coarse-scale (correlated) variability, \(\sigma^2_{2}\) is the
-independent random error variance measuring fine-scale (independent)
-variability, \(\phi\) is the range parameter measuring the
-distance-decay rate of the correlation, and \(h_{i,j}\) is the Euclidean
-distance between population units \(i\) and \(j\). Often
-\(\sigma^2_{1}\) and \(\sigma^2_{2}\) are called the partial sill and
-nugget, respectively. Any spatial covariance function could be used in
-the place of the exponential, including functions that allow for
-non-stationarity or anisotropy (Chiles and Delfiner, 1999, pp. 80--93).
+\end{align} where \(\sigma^2_{1}\) is the variance parameter quantifying
+the variability that is dependent (coarse-scale), \(\sigma^2_{2}\) is
+the variance parameter quantifying the variability that is independent
+(fine-scale), \(\phi\) is the range parameter measuring the
+distance-decay rate of the covariance, and \(h_{i,j}\) is the Euclidean
+distance between population units \(i\) and \(j\). The proportion of
+variability attributable to dependent random error is
+\(\sigma^2_{1} / (\sigma^2_{1} + \sigma^2_{2})\). Similarly, the
+proportion of variability attributable to independent random error is
+\(\sigma^2_{2} / (\sigma^2_{1} + \sigma^2_{2})\). Finally we note that
+\(\sigma^2_{1}\) and \(\sigma^2_{2}\) are often called the partial sill
+and nugget, respectively.
 
@@
 
-We note that we only use FPBK in this paper in order to focus more on
-comparing the design-based and model-based approaches. Other methods,
-such as k-nearest-neighbors (Fix and Hodges, 1989; Ver Hoef and
-Temesgen, 2013), random forest (Breiman, 2001), Bayesian models
-(Chan-Golston et al., 2020), among others, could also be used to obtain
-predictions for a mean or total from spatially correlated responses of a
-finite population. We choose to use FPBK because it is faster than a
-Bayesian approach and it was developed with theoretically-based variance
-estimators of means and totals for spatial data, whereas random forests
-and k-nearest-neighbors use ad-hoc variance estimators in most cases
-(Ver Hoef and Temesgen, 2013); additionally, FBPK outperformed the other
-methods in most scenarios.
+Other approaches, such as k-nearest-neighbors (Fix and Hodges, 1989; Ver
+Hoef and Temesgen, 2013), random forests (Breiman, 2001), Bayesian
+models (Chan-Golston et al., 2020), among others, could also be used to
+obtain predictions for a mean or total from spatially correlated
+responses of a finite population. Compared to the k-nearest-neighbors
+and random forest approach, we prefer FBPK because it is model-based and
+relies on theoretically-based variance estimators leveraging the model's
+spatial covariance structure, whereas k-nearest-neighbors and random
+forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
+Additionally, Ver Hoef and Temesgen (2013) studied compared FBPK,
+k-nearest-neighbors, and random forests in a variety of spatial data
+contexts, and FBPK tended to perform best. Compared to the Bayesian
+approach, we prefer FPBK mostly because it is much more computationally
+efficient.
 

### 2021-12-07 inst/manuscript/manuscript.Rmd
section 3,4,5 updates

@@
 
-FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean zero and a spatial correlation structure that can be modeled using a covariance function.
+FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial correlation structure that can be modeled using a covariance function.
 This covariance function is commonly assumed to be non-negative (between zero and one), second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
@@
 
-# Numerical Study {#sec:numstudy}
+# Simulation Study {#sec:numstudy}
 
-We used a simulation study to investigate performance of four sampling-analysis combinations: IRS-Design, IRS with a design-based analysis; IRS-Model, IRS with a model-based analysis; GRTS-Design, GRTS sampling with a design-based analysis; and GRTS-Model, GRTS sampling with a model-based analysis.
+We used a simulation study to investigate performance of four sampling-analysis combinations: IRS with a design-based analysis, called "IRS-Design"; IRS with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; GRTS sampling with a model-based analysis, called "GRTS-Model".
 These combinations are also provided in Table \ref{tab:designanalysis}.
@@
 
-Performance of the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
+Performance for the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
 The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error.
@@
 Samples were always selected from a population size ($N$) of $N = 900$.
-The two location layouts were random and gridded.
-Locations in the random layout were selected randomly from the unit square ($[0, 1] \times [0, 1]$).
-Locations in the gridded layout were selected randomly on a fixed grid from the unit square.
+The two location layouts (of the population units) were random and gridded.
+Locations in the random layout were randomly generated inside the unit square ($[0, 1] \times [0, 1]$).
+Locations in the gridded layout were placed on a fixed, equall spaced grid inside the unit square.
 The two response types were normal and lognormal.
 For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error.
-The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are from Equation$~$\ref{equation:expcov}.
+The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are the dependent random error variance (partial sill) and independent random error variance (nugget), respectively, from Equation$~$\ref{equation:expcov}.
 The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
@@
 The response was then exponentiated, yielding a random variable whose total variance is 2.
-The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed.
+The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed (i.e., not normal).
 
@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate the mean and construct confidence (design-based) or prediction (model-based) intervals.
-We recorded the bias, squared error, and interval coverage for all sampling-analysis combinations in each trial.
-Then we summarized the performance of the combinations across trials by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the rate at which the true mean is contained in its 95% interval.
+In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct confidence (design-based) or prediction (model-based) intervals.
+Then we recorded the bias, squared error, and interval coverage for all sampling-analysis combinations.
+After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of tiems the true mean is contained in its 95% interval.
 The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package `spsurvey` [@dumelle2021spsurvey].
@@
 Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
-More formally, the relative rMS(P)E is defined as
+The relative rMS(P)E is defined as
 \begin{equation*}
@@
 \end{equation*}
-When there is no spatial correlation (Figure \ref{fig:figeff}, top row), the four sampling-analysis combinations have approximately equal rMS(P)E.
-So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even when the response variable is not spatially correlated.
-When there is spatial correlation (Figure \ref{fig:figeff}, middle and bottom row), the GRTS-Model combination tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively small.
-As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other combinations widens.
+When there is no spatial correlation (Figure \ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+So using the GRTS sampling plan or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial correlation.
+When there is spatial correlation (Figure \ref{fig:figeff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
+As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
 Finally we note that when there is spatial correlation, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
-These conclusions are similar to those observed in the grid location layout.
+These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
 Tables for rMS(P)E in all 36 simulation scenarios are provided in the supplementary material.
@@
 
-We also studied 95% interval coverage among the combinations.
-The design-based 95% confidence intervals and model-based 95% prediction intervals were constructed using the normal distribution.
-Justification for the design-based and model-based intervals comes from the asymptotic normality of totals via the Central Limit Theorem.
+We also studied 95% interval coverage among the sampling-analysis combinations.
+The design-based confidence intervals and model-based prediction intervals were constructed using the normal distribution.
+Justification for this comes from the asymptotic normality of means via the Central Limit Theorem.
 
-Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four combinations in the random location layout.
-All four combinations have fairly similar interval coverage within each scenario.
-Coverage in the normal response scenarios tended to be near 95% and slightly higher than coverage in the lognormal scenarios.
-Coverage in the lognormal scenarios still generally exceeded 90%.
+Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
+Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage.
+Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios varied from from 90% to 95%.
 Coverage tended to always increase with the sample size.
-At a sample size of 200, all four combinations had approximately 95% interval coverage in both response scenarios and all dependent error proportions.
-These conclusions were similar to those found in the grid location layout.
+At a sample size of 200, all four sampling-analysis combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
+These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
 Tables for interval coverage in all 36 simulation scenarios are provided in the supplementary material.
@@
 The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality in lakes in the contiguous United States, to study mercury concentration.
+We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality for lakes in the contiguous United States.
+Specifically, we will analyze mercury concentration in lakes.
 Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
@@
 
-We selected a single IRS sample and a single GRTS sample and estimated the mean mercury concentration and its standard error using using design-based and model-based approaches; Table \ref{tab:appliedtab} shows the results.
-For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% intervals.
-However, we should not generalize these results to any other data or even to other samples from these data.
-But, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration across the contiguous United States.
-We also see that both approaches using the GRTS sample have a lower standard error than the both approaches using the IRS sample.
-We would expect this to be the case for most samples because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.
+We selected a single IRS sample and a single GRTS sample and estimated (design-based) or predicted (model-based) the mean mercury concentration and its standard error using using design-based and model-based approaches.
+For the model-based analyses, the exponential covariance was used.
+Table \ref{tab:appliedtab} shows the results from these analyses.
+For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% confidence (design-based) or prediction (model-based) intervals.
+Though we should not generalize these results to other samples from these data, we do note a couple of patterns.
+The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration.
+Both analyses using GRTS sampling have lower standard errors than both analyses using IRS sampling.
+We expect that these patterns are consistent with other samples from these data because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.
 
-To better understand the dependence structure in mercury concentration, the empirical semivariogram and corresponding fit of the model-based approaches can be visualized.
+To better understand the dependence structure in mercury concentration, the empirical semivariogram can be visualized alongside a fit using the model-based approaches.
 The empirical semivariogram quantifies the halved squared differences (semivariance) among response values at different distances apart.
 If a process exhibits strong spatial dependence, the empirical semivariogram will have small values at small distances and large values at large distances.
-Figure \ref{fig:figsv} shows the empirical semivariogram for GRTS-Model, displaying the average semivariance for several distances.
+Figure \ref{fig:figsv} shows the empirical semivariogram quantifying the average semivariance for several distances.
 Overlain onto Figure \ref{fig:figsv} is the estimated semivariance obtained using the covariance parmaeters from the REML fit of GRTS-Model.
@@
 
-The design-based and model-based approaches to inference are fundamentally different paradigms by which samples are selected and data are analyzed.
-The design-based approach incorporates randomness through sampling to estimate a population parameter.
-The model-based approach incorporates randomness through distributional assumptions to predict the realized values of a random process.
+The design-based and model-based approaches to statistical inference are fundamentally different paradigms by which samples are selected and data are analyzed.
+The design-based approach incorporates randomness through sampling to estimate population parameters.
+The model-based approach incorporates randomness through distributional assumptions to predict realized values of a random process.
 Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
@@
 Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated much of the inefficiency of the IRS sample.
-Second, independent of the analysis approach, there is no reason to use IRS over GRTS for sampling spatial data, as GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial correlation and noticeably better than there IRS counterparts when there was spatial correlation.
-Third, The stronger the spatial correlation, the larger the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations.
-Fourth and finally, interval coverage for the normal response was very close to 95% for all sample sizes, while interval coverage for the lognormal response was not very close to 95% until $n = 200$.
+Second, independent of the analysis approach, we found no reason to prefer IRS over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial correlation and noticeably better than their IRS counterparts when there was spatial correlation.
+Third, as the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations also increases.
+Fourth and finally, when the response was normal, interval coverage for all sampling-analysis combinations was very close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling and analysis was between 90% and 95% and closest to 95% when $n = 200$.
 
-There are several benefits and drawbacks of the design-based and model-based approaches for spatial data, some of which we have not yet discussed but are worthy of consideration in future research.
-Design-based approaches are often computationally efficient, while model-based estimation of covariance parameters can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
-The design-based approach also more naturally handles binary data, free from the more complicated logistic regression formulation commonly used to handle binary data in a model-based approach.
-The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
-The model-based approach also yields estimated spatial covariance parameters, which help better understand the process of study.
+There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial data.
+Some we have discuss, but others we have not and they are worthy of consideration in future research.
+Design-based approaches are often computationally efficient, while model-based estimation can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
+The design-based approach also more naturally handles binary data, free from the more complicated logistic regression framework commonly used to analyze binary data in a model-based approach.
+The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and response variable.
+The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the process of study.
 Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
@@
 Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
-The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, should be seriously considered when choosing among them.
+The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, can motive the process of choosing among them.
 This is especially true from an analysis perspective, as we found that using a spatially balanced sampling algorithm benefits both design-based and model-based analyses.

### 2021-12-07 inst/manuscript/manuscript.tex
section 3,4,5 updates

@@
 FBPK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
-has mean zero and a spatial correlation structure that can be modeled
+has mean-zero and a spatial correlation structure that can be modeled
 using a covariance function. This covariance function is commonly
@@
 \hypertarget{sec:numstudy}{%
-\section{Numerical Study}\label{sec:numstudy}}
+\section{Simulation Study}\label{sec:numstudy}}
 
 We used a simulation study to investigate performance of four
-sampling-analysis combinations: IRS-Design, IRS with a design-based
-analysis; IRS-Model, IRS with a model-based analysis; GRTS-Design, GRTS
-sampling with a design-based analysis; and GRTS-Model, GRTS sampling
-with a model-based analysis. These combinations are also provided in
-Table \ref{tab:designanalysis}.
+sampling-analysis combinations: IRS with a design-based analysis, called
+``IRS-Design''; IRS with a model-based analysis, called ``IRS-Model'';
+GRTS sampling with a design-based analysis, called ``GRTS-Design''; GRTS
+sampling with a model-based analysis, called ``GRTS-Model''. These
+combinations are also provided in Table \ref{tab:designanalysis}.
 
@@
 
-Performance of the four sampling-analysis combinations was evaluated in
+Performance for the four sampling-analysis combinations was evaluated in
 36 different simulation scenarios. The 36 scenarios resulted from the
@@
 always selected from a population size (\(N\)) of \(N = 900\). The two
-location layouts were random and gridded. Locations in the random layout
-were selected randomly from the unit square (\([0, 1] \times [0, 1]\)).
-Locations in the gridded layout were selected randomly on a fixed grid
-from the unit square. The two response types were normal and lognormal.
-For the normal response type, the response was simulated using mean-zero
-random errors with the exponential covariance
-(Equation\(~\)\ref{equation:expcov}) for varying proportions of
-dependent random error. The proportion of dependent random error is
-represented by \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where
-\(\sigma^2_1\) and \(\sigma^2_2\) are from
+location layouts (of the population units) were random and gridded.
+Locations in the random layout were randomly generated inside the unit
+square (\([0, 1] \times [0, 1]\)). Locations in the gridded layout were
+placed on a fixed, equall spaced grid inside the unit square. The two
+response types were normal and lognormal. For the normal response type,
+the response was simulated using mean-zero random errors with the
+exponential covariance (Equation\(~\)\ref{equation:expcov}) for varying
+proportions of dependent random error. The proportion of dependent
+random error is represented by
+\(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where \(\sigma^2_1\) and
+\(\sigma^2_2\) are the dependent random error variance (partial sill)
+and independent random error variance (nugget), respectively, from
 Equation\(~\)\ref{equation:expcov}. The total variance,
@@
 performance of the sampling-analysis approaches for data that were
-skewed.
+skewed (i.e., not normal).
 
@@
 simulation trials. In each trial, IRS and GRTS samples were selected and
-then design-based and model-based analyses were used to estimate the
-mean and construct confidence (design-based) or prediction (model-based)
-intervals. We recorded the bias, squared error, and interval coverage
-for all sampling-analysis combinations in each trial. Then we summarized
-the performance of the combinations across trials by calculating average
-bias, rMS(P)E (root-mean-squared error for the design-based approaches
-and root-mean-squared-prediction error for the model-based approaches),
-and the rate at which the true mean is contained in its 95\% interval.
+then design-based and model-based analyses were used to estimate
+(design-based) or predict (model-based) the mean and construct
+confidence (design-based) or prediction (model-based) intervals. Then we
+recorded the bias, squared error, and interval coverage for all
+sampling-analysis combinations. After all 2000 trials, we summarized the
+long-run performance of the combinations by calculating average bias,
+rMS(P)E (root-mean-squared error for the design-based approaches and
+root-mean-squared-prediction error for the model-based approaches), and
+the proportion of tiems the true mean is contained in its 95\% interval.
 The GRTS algorithm and the local neighborhood variance estimator are
@@
 approaches from Table \ref{tab:designanalysis} using the random location
-layout with ``IRS-Design'' as the baseline. More formally, the relative
-rMS(P)E is defined as \begin{equation*}
+layout with ``IRS-Design'' as the baseline. The relative rMS(P)E is
+defined as \begin{equation*}
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*} When there is no spatial correlation (Figure
-\ref{fig:figeff}, top row), the four sampling-analysis combinations have
-approximately equal rMS(P)E. So, using GRTS or using a spatial model
-does not result in much, if any, loss in efficiency even when the
-response variable is not spatially correlated. When there is spatial
-correlation (Figure \ref{fig:figeff}, middle and bottom row), the
-GRTS-Model combination tends to perform best, followed by GRTS-Design,
-IRS-Model, and finally IRS-Design, though the difference in relative
-rMS(P)E among IRS-Model, GRTS-Design, and GRTS-Model is relatively
-small. As the strength of spatial correlation increases, the gap in
-rMS(P)E between IRS-Design and the other combinations widens. Finally we
-note that when there is spatial correlation, IRS-Model outperforms
-IRS-Design by a large margin, suggesting that the poor design properties
-of IRS are largely mitigated by the model-based analysis. These
-conclusions are similar to those observed in the grid location layout.
-Tables for rMS(P)E in all 36 simulation scenarios are provided in the
-supplementary material.
+\ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
+combinations have approximately equal rMS(P)E. So using the GRTS
+sampling plan or a model-based analysis does not result in much, if any,
+loss in efficiency compared to IRS-Design when there is no spatial
+correlation. When there is spatial correlation (Figure \ref{fig:figeff},
+``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
+best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though
+the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and
+IRS-Model is relatively small. As the strength of spatial correlation
+increases, the gap in rMS(P)E between IRS-Design and the other
+sampling-analysis combinations widens. Finally we note that when there
+is spatial correlation, IRS-Model outperforms IRS-Design by a large
+margin, suggesting that the poor design properties of IRS are largely
+mitigated by the model-based analysis. These conclusions are similar to
+those observed in the grid location layout, so we omit a grid location
+layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
+are provided in the supplementary material.
 
@@
 
-We also studied 95\% interval coverage among the combinations. The
-design-based 95\% confidence intervals and model-based 95\% prediction
-intervals were constructed using the normal distribution. Justification
-for the design-based and model-based intervals comes from the asymptotic
-normality of totals via the Central Limit Theorem.
+We also studied 95\% interval coverage among the sampling-analysis
+combinations. The design-based confidence intervals and model-based
+prediction intervals were constructed using the normal distribution.
+Justification for this comes from the asymptotic normality of means via
+the Central Limit Theorem.
 
 Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
-the four combinations in the random location layout. All four
-combinations have fairly similar interval coverage within each scenario.
-Coverage in the normal response scenarios tended to be near 95\% and
-slightly higher than coverage in the lognormal scenarios. Coverage in
-the lognormal scenarios still generally exceeded 90\%. Coverage tended
-to always increase with the sample size. At a sample size of 200, all
-four combinations had approximately 95\% interval coverage in both
-response scenarios and all dependent error proportions. These
-conclusions were similar to those found in the grid location layout.
-Tables for interval coverage in all 36 simulation scenarios are provided
-in the supplementary material.
+the four sampling-analysis combinations in the random location layout.
+Within each scenario, the sampling-analysis combinations tend to have
+fairly similar interval coverage. Coverage in the normal response
+scenarios was usually near 95\%, while coverage in the lognormal
+response scenarios varied from from 90\% to 95\%. Coverage tended to
+always increase with the sample size. At a sample size of 200, all four
+sampling-analysis combinations had approximately 95\% interval coverage
+in both response scenarios for all dependent error proportions. These
+conclusions are similar to those observed in the grid location layout,
+so we omit a grid location layout figure here. Tables for interval
+coverage in all 36 simulation scenarios are provided in the
+supplementary material.
 
@@
 will use the 2012 National Lakes Assessment (NLA), which measures
-various aspects of lake health and water quality in lakes in the
-contiguous United States, to study mercury concentration. Although we
-know the true mean mercury concentration values for the 986 lakes from
-the 2012 NLA, we will explore whether or not we obtain an adequately
-precise estimate for the realized mean mercury concentration if we
-sample only 100 of the 986 lakes.
+various aspects of lake health and water quality for lakes in the
+contiguous United States. Specifically, we will analyze mercury
+concentration in lakes. Although we know the true mean mercury
+concentration values for the 986 lakes from the 2012 NLA, we will
+explore whether or not we obtain an adequately precise estimate for the
+realized mean mercury concentration if we sample only 100 of the 986
+lakes.
 
@@
 We selected a single IRS sample and a single GRTS sample and estimated
-the mean mercury concentration and its standard error using using
-design-based and model-based approaches; Table \ref{tab:appliedtab}
-shows the results. For all four sampling-analysis combinations, the true
-realized mean mercury concentration is within the bounds of the 95\%
-intervals. However, we should not generalize these results to any other
-data or even to other samples from these data. But, we do note a couple
-of patterns. The design-based IRS analysis shows the largest standard
-error: a likely reason is that this is the only approach that does not
-incorporate any spatial information regarding mercury concentration
-across the contiguous United States. We also see that both approaches
-using the GRTS sample have a lower standard error than the both
-approaches using the IRS sample. We would expect this to be the case for
-most samples because mercury concentration exhibits spatial patterning,
-so a spatially balanced sample should usually yield a lower standard
-error.
+(design-based) or predicted (model-based) the mean mercury concentration
+and its standard error using using design-based and model-based
+approaches. For the model-based analyses, the exponential covariance was
+used. Table \ref{tab:appliedtab} shows the results from these analyses.
+For all four sampling-analysis combinations, the true realized mean
+mercury concentration is within the bounds of the 95\% confidence
+(design-based) or prediction (model-based) intervals. Though we should
+not generalize these results to other samples from these data, we do
+note a couple of patterns. The design-based IRS analysis shows the
+largest standard error: a likely reason is that this is the only
+approach that does not incorporate any spatial information regarding
+mercury concentration. Both analyses using GRTS sampling have lower
+standard errors than both analyses using IRS sampling. We expect that
+these patterns are consistent with other samples from these data because
+mercury concentration exhibits spatial patterning, so a spatially
+balanced sample should usually yield a lower standard error.
 
 To better understand the dependence structure in mercury concentration,
-the empirical semivariogram and corresponding fit of the model-based
-approaches can be visualized. The empirical semivariogram quantifies the
+the empirical semivariogram can be visualized alongside a fit using the
+model-based approaches. The empirical semivariogram quantifies the
 halved squared differences (semivariance) among response values at
@@
 distances and large values at large distances. Figure \ref{fig:figsv}
-shows the empirical semivariogram for GRTS-Model, displaying the average
-semivariance for several distances. Overlain onto Figure \ref{fig:figsv}
-is the estimated semivariance obtained using the covariance parmaeters
-from the REML fit of GRTS-Model. Figure \ref{fig:figsv} provides
-evidence that there is strong correlation in mercury concentration among
-the sites.
+shows the empirical semivariogram quantifying the average semivariance
+for several distances. Overlain onto Figure \ref{fig:figsv} is the
+estimated semivariance obtained using the covariance parmaeters from the
+REML fit of GRTS-Model. Figure \ref{fig:figsv} provides evidence that
+there is strong correlation in mercury concentration among the sites.
 
@@
 
-The design-based and model-based approaches to inference are
+The design-based and model-based approaches to statistical inference are
 fundamentally different paradigms by which samples are selected and data
 are analyzed. The design-based approach incorporates randomness through
-sampling to estimate a population parameter. The model-based approach
+sampling to estimate population parameters. The model-based approach
 incorporates randomness through distributional assumptions to predict
-the realized values of a random process. Though these approaches have
-often been compared in the literature both from theoretical and
-analytical perspectives, our contribution lies in studying them in a
-spatial context while implementing spatially balanced sampling. Aside
-from the theoretical differences described, a few analytical findings
-from the simulation study are particularly notable. First, the sampling
-decision (GRTS vs IRS) is most important when using a design-based
-analysis. Though GRTS-Model still outperformed IRS-Model, the
-model-based analysis mitigated much of the inefficiency of the IRS
-sample. Second, independent of the analysis approach, there is no reason
-to use IRS over GRTS for sampling spatial data, as GRTS-Design and
-GRTS-Model generally performed at least as well as their IRS
-counterparts when there was no spatial correlation and noticeably better
-than there IRS counterparts when there was spatial correlation. Third,
-The stronger the spatial correlation, the larger the gap in rMS(P)E
-between IRS-Design and the other sampling-analysis combinations. Fourth
-and finally, interval coverage for the normal response was very close to
-95\% for all sample sizes, while interval coverage for the lognormal
-response was not very close to 95\% until \(n = 200\).
+realized values of a random process. Though these approaches have often
+been compared in the literature both from theoretical and analytical
+perspectives, our contribution lies in studying them in a spatial
+context while implementing spatially balanced sampling. Aside from the
+theoretical differences described, a few analytical findings from the
+simulation study are particularly notable. First, the sampling decision
+(GRTS vs IRS) is most important when using a design-based analysis.
+Though GRTS-Model still outperformed IRS-Model, the model-based analysis
+mitigated much of the inefficiency of the IRS sample. Second,
+independent of the analysis approach, we found no reason to prefer IRS
+over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model
+generally performed at least as well as their IRS counterparts when
+there was no spatial correlation and noticeably better than their IRS
+counterparts when there was spatial correlation. Third, as the strength
+of spatial correlation increases, the gap in rMS(P)E between IRS-Design
+and the other sampling-analysis combinations also increases. Fourth and
+finally, when the response was normal, interval coverage for all
+sampling-analysis combinations was very close to 95\% for all sample
+sizes; when the response was lognormal, interval coverage for all
+sampling and analysis was between 90\% and 95\% and closest to 95\% when
+\(n = 200\).
 
 There are several benefits and drawbacks of the design-based and
-model-based approaches for spatial data, some of which we have not yet
-discussed but are worthy of consideration in future research.
-Design-based approaches are often computationally efficient, while
-model-based estimation of covariance parameters can be computationally
+model-based approaches for finite population spatial data. Some we have
+discuss, but others we have not and they are worthy of consideration in
+future research. Design-based approaches are often computationally
+efficient, while model-based estimation can be computationally
 burdensome, especially for likelihood-based methods such as REML that
@@
 more naturally handles binary data, free from the more complicated
-logistic regression formulation commonly used to handle binary data in a
+logistic regression framework commonly used to analyze binary data in a
 model-based approach. The model-based approach, however, can more
 naturally quantify the relationship between covariates (predictor
-variables) and the response variable. The model-based approach also
-yields estimated spatial covariance parameters, which help better
-understand the process of study. Model selection is also possible using
-model-based approaches and criteria such as cross validation, likelihood
-ratio tests, or AIC (Akaike, 1974). Model-based approaches are capable
-of more efficient small-area estimation than design-based approaches by
-leveraging distributional assumptions in areas with few observed sites.
-Model-based approaches can also compute site-by-site predictions at
-unobserved locations and use them to construct informative
-visualizations. The benefits and drawbacks of both approaches, alongside
-our theoretical and analytical comparisons, should be seriously
-considered when choosing among them. This is especially true from an
-analysis perspective, as we found that using a spatially balanced
+variables) and response variable. The model-based approach also yields
+estimated spatial covariance parameters, which help better understand
+the dependence structure in the process of study. Model selection is
+also possible using model-based approaches and criteria such as cross
+validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
+approaches are capable of more efficient small-area estimation than
+design-based approaches by leveraging distributional assumptions in
+areas with few observed sites. Model-based approaches can also compute
+site-by-site predictions at unobserved locations and use them to
+construct informative visualizations. The benefits and drawbacks of both
+approaches, alongside our theoretical and analytical comparisons, can
+motive the process of choosing among them. This is especially true from
+an analysis perspective, as we found that using a spatially balanced
 sampling algorithm benefits both design-based and model-based analyses.

### 2021-12-07 R/sim_trial.R

@@ -59,7 +59,7 @@ sim_trial <- function(seed = sample.int(1e7, size = 1),
   data_sf <- sf::st_as_sf(data, coords = c("x", "y"), crs = 5070)
 
   ###############################
-  ## IRS Sample
+  ## IRS Sample # must be spsurvey version 5.1.0 or greater
   ###############################
 
   # irs sample

### 2021-12-08 inst/manuscript/manuscript.Rmd
minor updates using plot_grid

@@
 
-```{r figdata, out.width = "49%", fig.cap = "Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).", message = FALSE, warning = FALSE, fig.show='hold', fig.align='center'}
+```{r, message = FALSE, warning = FALSE}
 merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
@@
   labs(x = "", y = "", colour = "Hg") +
-  lims(x = c(0, 6500)) +
+  # lims(x = c(0, 6500)) +
   coord_quickmap() +
@@
       axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      legend.position = c(0.85, 0.5),
-      legend.key.size = unit(0.75, "cm"))
+      axis.ticks = element_blank())#,
+      #legend.position = c(0.85, 0.5),
+      #legend.key.size = unit(0.75, "cm"))
 
@@
   theme_bw(base_size = 20)
+```
 
-merc_map
-merc_hist
+```{r, warning = FALSE, message = FALSE, include = FALSE}
+mod_plot <- plot(mod_grts) +
+  labs(x = "Distance", title = "") +
+  scale_size_continuous(name = "Pairs") +
+  theme_bw(base_size = 20)
+```
+
+```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).
+The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model.", message = FALSE, warning = FALSE, fig.align='center'}
+plot_grid(merc_map, plot_grid(merc_hist, mod_plot), ncol = 1, nrow = 2)
+# plot_grid(merc_map, merc_hist, mod_plot, nrow = 3)
 ```
@@
 Figure \ref{fig:figsv} provides evidence that there is strong correlation in mercury concentration among the sites.
-
-```{r figsv, out.width = "85%", fig.align='center', fig.cap = "The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model."}
-plot(mod_grts)
-```
 

### 2021-12-08 inst/manuscript/manuscript.tex
minor updates using plot_grid

@@
 
-{\centering \includegraphics[width=0.49\linewidth]{manuscript_files/figure-latex/figdata-1} \includegraphics[width=0.49\linewidth]{manuscript_files/figure-latex/figdata-2}
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1}
 
@@
 
-\caption{Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).}\label{fig:figdata}
+\caption{Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).
+The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model.}\label{fig:figdata}
 \end{figure}
@@
 there is strong correlation in mercury concentration among the sites.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{manuscript_files/figure-latex/figsv-1}
-
-}
-
-\caption{The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model.}\label{fig:figsv}
-\end{figure}
 

### 2021-12-18 inst/manuscript/manuscript.Rmd
update to remove figures and application computations from main body of text

--- 2021-12-07 inst/manuscript/manuscript.Rmd
@@
 This gap in performance is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
-
 journal: "Methods in Ecology and Evolution"
@@
   \DeclareMathOperator{\cov}{{cov}}
+  \usepackage{caption}
+  \usepackage{subcaption}
 output: rticles::elsevier_article
@@
 
-```{r fig1, out.width = "100%", fig.cap = "A visual comparison of the design-based and model-based approaches.
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/dvm_comp.jpeg}
+  \caption{A visual comparison of the design-based and model-based approaches.
 In the top row, there is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
@@
 In the bottom row, there are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
-The ata-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations."}
-set.seed(10102021)
-
-source(here("R", "sim_pop.R"))
-N <- 9
-pop_df <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                  psill = 0.9, range = 1, nugget = 0.1)
-
-n <- 4
-samp_df <- pop_df %>% sample_n(n) 
-samp_df2 <- pop_df %>% sample_n(n) 
-samp_df3 <- pop_df %>% sample_n(n)
-
-library(cowplot)
-d1 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df, size = 5.5, shape = 1, 
-             show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d2 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df2, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(),
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-d3 <- ggplot(data = pop_df, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df3, size = 5.5, shape = 1, show.legend = FALSE,
-             stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Design 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-pop_df_mod <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, range = 1, nugget = 0.1)
-pop_df_mod2 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                       psill = 0.9, range = 1, nugget = 0.1)
-pop_df_mod3 <- sim_pop(N = N, gridded = TRUE, cortype = "Exponential",
-                      psill = 0.9, range = 1, nugget = 0.1)
-
-indeces <- sample(1:N, size = n, replace = FALSE)
-samp_df_mod <- pop_df_mod %>% slice(indeces)
-samp_df_mod2 <- pop_df_mod2 %>% slice(indeces)
-samp_df_mod3 <- pop_df_mod3 %>% slice(indeces)
-
-m1 <- ggplot(data = pop_df_mod, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 1",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m2 <- ggplot(data = pop_df_mod2, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 2",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-m3 <- ggplot(data = pop_df_mod3, aes(x = x, y = y)) +
-  geom_point(aes(colour = response), size = 4) +
-  scale_colour_viridis_c() +
-  geom_point(data = samp_df_mod2, size = 5.5, shape = 1, show.legend = FALSE, stroke = 1.5) +
-  theme_bw() +
-  labs(title = "Model 3",
-       colour = "Response") +
-  theme(axis.title = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text = element_blank(), 
-        legend.position = "none",
-        panel.grid.major = element_blank()) +
-  xlim(-0.1, 1.1) +
-  ylim(-0.1, 1.1)
-
-library(ggpubr)
-ggarrange(d1, d2, d3, m1, m2, m3, nrow = 2, ncol = 3,
-          common.legend = TRUE, legend = "right")
-```
+The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations}
+  \label{fig:fig1}
+\end{figure}
 
@@
 The first spatially balanced sampling algorithm seeing widespread use is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
-To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Direchlet Tessellations).
+To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Dirichlet Tessellations).
 After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, such as the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
@@
 @stevens2004spatially provides more technical details.
-
-<!-- This makes it seem like HT is being used to estimate population parameters in GRTS?
-Needs to be in there, but maybe a sentence making it clear that this isn't being used?
-Or just change the order and talk about Local variance first and then mention these for more general settings?
--->
 
@@
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and populatoin unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
+First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
 Second, these estimators ignore the spatial locations of the population units.
@@
 Locations in the random layout were randomly generated inside the unit square ($[0, 1] \times [0, 1]$).
-Locations in the gridded layout were placed on a fixed, equall spaced grid inside the unit square.
+Locations in the gridded layout were placed on a fixed, equally spaced grid inside the unit square.
 The two response types were normal and lognormal.
@@
 Then we recorded the bias, squared error, and interval coverage for all sampling-analysis combinations.
-After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of tiems the true mean is contained in its 95% interval.
+After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% interval.
 The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package `spsurvey` [@dumelle2021spsurvey].
@@
 Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
+<!--- figure 2 -->
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:figeff}
+\end{figure}
 The relative rMS(P)E is defined as
@@
 
-```{r figeff, out.width = "100%", fig.cap = "Relative rMS(P)E for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.", message = FALSE, warning = FALSE}
-library(tidyverse)
-library(here)
-
-files <- list.files(path = here("inst", "output", "simulation_summary"), pattern = "*.csv", full.names = TRUE)
-
-combo_data <- purrr::map_df(files,
-                            ~read_csv(.x) %>% mutate(filename = .x))
-
-combo_data <- combo_data %>%
-  dplyr::filter(gridded == FALSE) %>%
-  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
-  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
-                                            "Design GRTS", "Model GRTS"))) %>%
-  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal")))
-combo_data <- combo_data %>% group_by(sim) %>% mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
-  fill(designirsrmspe, .direction = "downup") %>%
-  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  ungroup() %>%
-  mutate(psill_ratio = 1 - nugget_ratio) %>%
-  mutate(n_factor = factor(n))
-
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
-psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
-effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
-                              colour = approach)) +
-  facet_grid(
-    psill_ratio ~ resptype,
-    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
-  ) +
-  geom_jitter(width = 0.24, size = 2.5) +
-  scale_colour_manual(values = colour_scale) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Relative Efficiency")
-effplot
-```
-
 We also studied 95% interval coverage among the sampling-analysis combinations.
@@
 Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
+<!--- figure 3 -->
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
+  \caption{Interval coverage for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line in each plot represents 95\% coverage.}
+  \label{fig:figconf}
+\end{figure}
 Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage.
@@
 Tables for interval coverage in all 36 simulation scenarios are provided in the supplementary material.
-
-```{r figconf, out.width = "100%", fig.cap = "Interval coverage for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line in each plot represents 95\\% coverage.", message = FALSE, warning = FALSE}
-ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
-  geom_hline(yintercept = 0.95) +
-  geom_jitter(width = 0.24, size = 2.5) + 
-  facet_grid(
-    psill_ratio ~ resptype,
-    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
-  ) +
-  scale_colour_manual(values = colour_scale) +
-  ylim(0.80, 1) +
-  theme_bw(base_size = 14) +
-  labs(x = "Sample Size", colour = "Approach",
-       y = "Interval Coverage")
-```
 
@@
 
-```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
-library(tidyverse)
-library(here)
-library(DvMsp)
-data(nla_df) # may need to reinstall package
-nla_df
-factor(nla_df$TOTALHG_UNITS) ## all same units
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>% summarise(maxnonmiss = max(nonmiss))
-## HG never measured twice at the same site.
-
-nla_df %>% group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-
-## in duplicate sites, fill in the value for RDis_IX
-nla_nomiss <- nla_df %>%
-  group_by(SITE_ID) %>%
-  fill(RDis_IX, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-nla_nomiss_both <- nla_nomiss
-## check
-nla_nomiss_both %>% group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-
-## transform coordinates
-library(sptotal)
-coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
-                     lat = nla_nomiss_both$INDEX_LAT_DD, 
-                     lon = nla_nomiss_both$INDEX_LON_DD)
-nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
-nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-```
-
-```{r, echo = FALSE}
-mean_hg <- mean(nla_nomiss_both$total_hg)
-# mean(nla_nomiss_both$RDis_IX)
-```
-
-Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
+Figure \ref{fig:merc} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
 Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
 The realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
+TALK ABOUT EMPIRICAL SEMIVARIOGRAM IN THIS PARAGRAPH.
 
-```{r, results = "hide", fig.keep = "none"}
-## IRS Sample, Model Analysis
-
-set.seed(080421)
-n <- 100
-lakesobs <- nla_nomiss_both %>% sample_n(n) 
-lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
-lakesunobs$total_hg <- NA
-
-lakes_test <- bind_rows(lakesobs, lakesunobs) 
-lakes_test$wts <- 1 / nrow(lakes_test) ## predicting for mean
-
-slmfitout_exp_lakes <- slmfit(formula = total_hg ~ 1,
-                              data = lakes_test, 
-                              xcoordcol = 'xcoords',
-                              ycoordcol = 'ycoords',
-                              CorModel = "Exponential")
-summary(slmfitout_exp_lakes)
-plot(slmfitout_exp_lakes)
-
-pred_exp_lakes <- predict(slmfitout_exp_lakes, wtscol = "wts",
-                          conf_level = 0.95)
-print(pred_exp_lakes)
-mean_irs_mod <- pred_exp_lakes$FPBK_Prediction
-se_irs_mod <- sqrt(pred_exp_lakes$PredVar)
-lb_irs_mod <- pred_exp_lakes$conf_bounds[1]
-ub_irs_mod <- pred_exp_lakes$conf_bounds[2]
-
-realized_mean <- mean(nla_nomiss_both$total_hg) ## compute realized mean
-```
-
-```{r, results = "hide"}
-## IRS Sample, Design Analysis
-## 
-irs_info <- lakesobs %>% summarise(meandoc = mean(total_hg),
-                                   vardoc = var(total_hg))
-N <- nrow(nla_nomiss_both); n <- nrow(lakesobs)
-irs_se_irs_samp <- sqrt((irs_info$vardoc / n) * (N - n) / N)
-irs_lb_irs_samp <- irs_info$meandoc - 1.96 * irs_se_irs_samp
-irs_ub_irs_samp <- irs_info$meandoc + 1.96 * irs_se_irs_samp
-irs_info$meandoc; sqrt(irs_info$vardoc)
-irs_lb_irs_samp; irs_ub_irs_samp
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Design Analysis
-data_sf <- sf::st_as_sf(nla_nomiss_both, coords = c("xcoords", "ycoords"),
-                        crs = 5070)
-library(spsurvey)
-grts_samp <- grts(data_sf, n_base = n)
-grts_bind <- sp_rbind(grts_samp)
-
-## get coordinates
-grts_coords <- sf::st_coordinates(grts_bind)
-## make data frame
-grts_df <- data.frame(
-  response = grts_bind$total_hg,
-  x = grts_coords[, "X"],
-  y = grts_coords[, "Y"],
-  siteID = grts_bind$siteID,
-  wgt = grts_bind$wgt
-)
-## head(grts_df)
-summary(grts_df$response)
-
-design_analysis <- cont_analysis(
-  grts_df,
-  siteID = "siteID",
-  vars = "response",
-  weight = "wgt",
-  xcoord = "x",
-  ycoord = "y"
-)
-## just return the mean info
-## design_mean <- subset(design_analysis$Pct, Statistic == "Mean")
-design_mean_grts <- design_analysis$Mean
-design_mean_grts$Estimate
-design_mean_grts$StdError
-design_mean_grts$LCB95Pct
-design_mean_grts$UCB95Pct
-```
-
-```{r, results = "hide"}
-## GRTS Sample, Model Analysis
-
-grts_coords_resp <- grts_df %>% select(-siteID, -wgt) %>%
-  rename(total_hg = "response", xcoords = "x", ycoords = "y")
-grts_unsamp <- anti_join(nla_nomiss_both, grts_coords_resp)
-grts_unsamp$total_hg <- NA
-grts_unsamp <- grts_unsamp %>% select(total_hg, xcoords, ycoords)
-grts_full <- dplyr::bind_rows(grts_coords_resp, grts_unsamp)
-
-grts_full$wts <- 1 / nrow(grts_full)
-mod_grts <- slmfit(total_hg ~ 1, data = grts_full,
-                   xcoordcol = "xcoords",
-                   ycoordcol = "ycoords")
-pred_mod_grts <- predict(mod_grts, wtscol = "wts")
-model_mean_grts <- pred_mod_grts$FPBK_Prediction
-model_se_grts <- sqrt(pred_mod_grts$PredVar)
-model_lb_grts <- model_mean_grts + -1 * 1.96 * model_se_grts
-model_ub_grts <- model_mean_grts + 1 * 1.96 * model_se_grts
-model_mean_grts
-model_se_grts
-model_lb_grts
-model_ub_grts
-```
-
-```{r, results = "hide"}
-## combine results
-res_df <- tibble(approach = c("IRS-Design", "GRTS-Design", "IRS-Model", "GRTS-Model"),
-       realized_mean = c(realized_mean, realized_mean, realized_mean,
-                     realized_mean),
-       estimate = c(irs_info$meandoc,
-                    design_mean_grts$Estimate, as.vector(mean_irs_mod),
-                    as.vector(model_mean_grts)),
-       se = c(irs_se_irs_samp, design_mean_grts$StdError, as.vector(se_irs_mod),
-              as.vector(model_se_grts)),
-       lb = c(irs_lb_irs_samp, design_mean_grts$LCB95Pct, as.vector(lb_irs_mod),
-              as.vector(model_lb_grts)),
-       ub = c(irs_ub_irs_samp, design_mean_grts$UCB95Pct, as.vector(ub_irs_mod),
-              as.vector(model_ub_grts)))
-res_df <- res_df %>% slice(1, 3, 2, 4)
-```
-
-```{r figdata, out.width = "49%", fig.cap = "Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).", message = FALSE, warning = FALSE, fig.show='hold', fig.align='center'}
-merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(colour = "black", fill = "white", bins = 15) +
-  labs(x = "HG (ng / g)")
-# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
-#   geom_histogram(colour = "black", fill = "white", bins = 15) +
-#   labs(x = "Lakeshore Disturbance")
-merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
-  geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "Hg") +
-  lims(x = c(0, 6500)) +
-  coord_quickmap() +
-  theme_bw(base_size = 20) + 
-  theme(axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      legend.position = c(0.85, 0.5),
-      legend.key.size = unit(0.75, "cm"))
-
-merc_hist <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
-  geom_histogram(bins = 50) + 
-  labs(x = "Hg", y = "Count") +
-  theme_bw(base_size = 20)
-
-merc_map
-merc_hist
-```
-
-```{r apptab, results = "asis"}
-res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 ng / g.")
-print(xtab_app, include.rownames = FALSE, comment = FALSE)
-```
+\begin{figure}
+\centering
+\begin{subfigure}{0.98\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
+  \caption*{}
+  \label{fig:mercury_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
+  \caption*{}
+  \label{fig:mercury_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/sv_plot.jpeg}
+  \caption*{}
+  \label{fig:sv_plot}
+\end{subfigure}
+\caption{Mercury concentration visualizations for the population (Hg) for 986 lakes in the contiguous United States.
+A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
+\label{fig:merc}
+\end{figure}
 
@@
 
-To better understand the dependence structure in mercury concentration, the empirical semivariogram can be visualized alongside a fit using the model-based approaches.
+```{r apptab, results = "asis"}
+res_df <- readr::read_csv(here("inst", "output", "application", "application.csv"))
+res_df_drop <- res_df[ ,-2]
+names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
+The true mean concentration is 103.2 ng / g.")
+print(xtab_app, include.rownames = FALSE, comment = FALSE)
+```
+
+<!--- To better understand the dependence structure in mercury concentration, the empirical semivariogram can be visualized alongside a fit using the model-based approaches.
 The empirical semivariogram quantifies the halved squared differences (semivariance) among response values at different distances apart.
@@
 Figure \ref{fig:figsv} shows the empirical semivariogram quantifying the average semivariance for several distances.
-Overlain onto Figure \ref{fig:figsv} is the estimated semivariance obtained using the covariance parmaeters from the REML fit of GRTS-Model.
+Overlain onto Figure \ref{fig:figsv} is the estimated semivariance obtained using the covariance parameters from the REML fit of GRTS-Model.
 Figure \ref{fig:figsv} provides evidence that there is strong correlation in mercury concentration among the sites.
-
-```{r figsv, out.width = "85%", fig.align='center', fig.cap = "The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model."}
-plot(mod_grts)
-```
+-->
 
@@
 
-This manuscript has a supplementary R package that contains all of the data and code used.
-Instructions for download at available at [https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
+This manuscript has a supplementary R package that contains all of the data and code used. Instructions for download at available at
+
+[https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
 
@@
 
-# Acknowledgements {.unnumbered}
+# Acknowledgments {.unnumbered}
 
 The views expressed in this manuscript are those of the authors and do not necessarily represent the views or policies of the U.S.
-Environmental Protection Agency.
+Environmental Protection Agency or the National Oceanic and Atmospheric Administration.
 Any mention of trade names, products, or services does not imply an endorsement by the U.S.
-government or the U.S.
-Environmental Protection Agency.
+government, the U.S.
+Environmental Protection Agency, or the National Oceanic and Atmospheric Administration.
 The U.S.
-Environmental Protection Agency does not endorse any commercial products, services, or enterprises.
+Environmental Protection Agency and National Oceanic and Atmospheric Administration do not endorse any commercial products, services, or enterprises.
 

### 2021-12-18 inst/manuscript/manuscript.tex
update to remove figures and application computations from main body of text

@@
 
-\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}}
+\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/fig1-1} \caption{A visual comparison of the design-based and model-based approaches.
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/dvm_comp.jpeg}
+  \caption{A visual comparison of the design-based and model-based approaches.
 In the top row, there is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
@@
 In the bottom row, there are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
-The ata-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations.}\label{fig:fig1}
+The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations}
+  \label{fig:fig1}
 \end{figure}
@@
 2004). To quantify the spatial balance of a sample, Stevens and Olsen
-(2004) proposed loss metrics based on Voronoi polygons (Direchlet
+(2004) proposed loss metrics based on Voronoi polygons (Dirichlet
 Tessellations). After the GRTS algorithm was developed, several other
@@
 they rely on calculating \(\pi_{ij}\), the probability that population
-unit \(i\) and populatoin unit \(j\) are both in the sample -- this
+unit \(i\) and population unit \(j\) are both in the sample -- this
 quantity can be challenging if not impossible to calculate analytically.
@@
 square (\([0, 1] \times [0, 1]\)). Locations in the gridded layout were
-placed on a fixed, equall spaced grid inside the unit square. The two
+placed on a fixed, equally spaced grid inside the unit square. The two
 response types were normal and lognormal. For the normal response type,
@@
 root-mean-squared-prediction error for the model-based approaches), and
-the proportion of tiems the true mean is contained in its 95\% interval.
+the proportion of times the true mean is contained in its 95\% interval.
 The GRTS algorithm and the local neighborhood variance estimator are
@@
 approaches from Table \ref{tab:designanalysis} using the random location
-layout with ``IRS-Design'' as the baseline. The relative rMS(P)E is
-defined as \begin{equation*}
+layout with ``IRS-Design'' as the baseline.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:figeff}
+\end{figure}
+
+The relative rMS(P)E is defined as \begin{equation*}
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figeff-1} \caption{Relative rMS(P)E for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.}\label{fig:figeff}
-\end{figure}
-
 We also studied 95\% interval coverage among the sampling-analysis
@@
 the four sampling-analysis combinations in the random location layout.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
+  \caption{Interval coverage for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line in each plot represents 95\% coverage.}
+  \label{fig:figconf}
+\end{figure}
+
 Within each scenario, the sampling-analysis combinations tend to have
@@
 supplementary material.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figconf-1} \caption{Interval coverage for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line in each plot represents 95\% coverage.}\label{fig:figconf}
-\end{figure}
 
@@
 
-Figure \ref{fig:figdata} shows that mercury concentration is
-right-skewed, with most lakes having a low value of mercury
-concentration but a few having a much higher concentration. Mercury
-concentration exhibits some spatial correlation, with high mercury
-concentrations in lakes in the northeast and north central United
-States. The realized mean mercury concentration in the 986 lakes is
-103.2 ng / g.
+Figure \ref{fig:merc} shows that mercury concentration is right-skewed,
+with most lakes having a low value of mercury concentration but a few
+having a much higher concentration. Mercury concentration exhibits some
+spatial correlation, with high mercury concentrations in lakes in the
+northeast and north central United States. The realized mean mercury
+concentration in the 986 lakes is 103.2 ng / g. TALK ABOUT EMPIRICAL
+SEMIVARIOGRAM IN THIS PARAGRAPH.
 
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1}
-
-}
-
-\caption{Population distribution of mercury concentration (hg) for 986 lakes in the contiguous United States in a spatial layout (left) and a histogram (right).
-The empirical semivariogram (black circles) of mercury concentration against the REML fit using the estimated covariance parameters (black line) from GRTS-Model.}\label{fig:figdata}
+\centering
+\begin{subfigure}{0.98\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
+  \caption*{}
+  \label{fig:mercury_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
+  \caption*{}
+  \label{fig:mercury_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/sv_plot.jpeg}
+  \caption*{}
+  \label{fig:sv_plot}
+\end{subfigure}
+\caption{Mercury concentration visualizations for the population (Hg) for 986 lakes in the contiguous United States.
+A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
+\label{fig:merc}
 \end{figure}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Approach & Estimate & SE & 95\% LB & 95\% UB \\ 
-  \hline
-IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\ 
-  IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\ 
-  GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\ 
-  GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\ 
-   \hline
-\end{tabular}
-\caption{\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.} 
-\end{table}
 
@@
 
-To better understand the dependence structure in mercury concentration,
-the empirical semivariogram can be visualized alongside a fit using the
-model-based approaches. The empirical semivariogram quantifies the
-halved squared differences (semivariance) among response values at
-different distances apart. If a process exhibits strong spatial
-dependence, the empirical semivariogram will have small values at small
-distances and large values at large distances. Figure \ref{fig:figsv}
-shows the empirical semivariogram quantifying the average semivariance
-for several distances. Overlain onto Figure \ref{fig:figsv} is the
-estimated semivariance obtained using the covariance parmaeters from the
-REML fit of GRTS-Model. Figure \ref{fig:figsv} provides evidence that
-there is strong correlation in mercury concentration among the sites.
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Approach & Estimate & SE & 95\% LB & 95\% UB \\ 
+  \hline
+IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\ 
+  IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\ 
+  GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\ 
+  GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\ 
+   \hline
+\end{tabular}
+\caption{\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.} 
+\end{table}
 
@@
 data and code used. Instructions for download at available at
+
 \url{https://github.com/michaeldumelle/DvMsp}.
@@
 
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
 
@@
 not necessarily represent the views or policies of the U.S.
-Environmental Protection Agency. Any mention of trade names, products,
-or services does not imply an endorsement by the U.S. government or the
-U.S. Environmental Protection Agency. The U.S. Environmental Protection
-Agency does not endorse any commercial products, services, or
-enterprises.
+Environmental Protection Agency or the National Oceanic and Atmospheric
+Administration. Any mention of trade names, products, or services does
+not imply an endorsement by the U.S. government, the U.S. Environmental
+Protection Agency, or the National Oceanic and Atmospheric
+Administration. The U.S. Environmental Protection Agency and National
+Oceanic and Atmospheric Administration do not endorse any commercial
+products, services, or enterprises.
 

### 2021-12-18 inst/manuscript/supplementary.Rmd
update to remove figures and application computations from main body of text

@@
   - code: 1
-    text: "Corresponding Author"
-journal: "TBD"
+    text: "Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)"
+journal: "Methods in Ecology and Evolution"
 date: "`r Sys.Date()`"
@@
 )
-print(sim_table_out, include.rownames = FALSE, comment = FALSE)
+print(sim_table_out, include.rownames = FALSE, comment = FALSE, scalebox = "0.85")
 }

### 2021-12-18 inst/manuscript/supplementary.tex
minor updates
reanalyzing simulation data
minor manuscript updates
update to remove figures and application computations from main body of text

@@
 
-  \journal{An awesome journal} % Sets Journal name
+  \journal{Methods in Ecology and Evolution} % Sets Journal name
 
@@
     \author[USEPA]{Michael Dumelle\corref{1}}
-   \ead{Dumelle.Michael@epa.gov} 
+
     \author[STLAW]{Matt Higham}
-   \ead{mhigham@stlaw.edu} 
+
     \author[OSU]{Lisa Madsen}
@@
 Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
-      \cortext[1]{Corresponding Author}
+      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}
 
@@
 \centering
+\scalebox{0.85}{
 \begin{tabular}{lllllrrr}
@@
   \hline
-IRS-Design & Gridded & Normal & 0 & 50 & -0.0001 & 0.1939 & 0.9475 \\ 
-  IRS-Model & Gridded & Normal & 0 & 50 & -0.0001 & 0.1960 & 0.9370 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0010 & 0.1930 & 0.9175 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0016 & 0.1936 & 0.9405 \\ 
+IRS-Design & Gridded & Normal & 0 & 50 & -0.0023 & 0.1968 & 0.9440 \\ 
+  IRS-Model & Gridded & Normal & 0 & 50 & -0.0029 & 0.1988 & 0.9400 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0011 & 0.1946 & 0.9110 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0006 & 0.1949 & 0.9300 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 50 & 0.0077 & 1.0498 & 0.8355 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0274 & 1.2011 & 0.8230 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 50 & 0.0092 & 0.9618 & 0.7910 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0069 & 0.9808 & 0.8100 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 50 & -0.0036 & 0.2007 & 0.9080 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0094 & 0.2071 & 0.9045 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 50 & -0.0007 & 0.1962 & 0.8690 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0022 & 0.1969 & 0.8945 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0 & 50 & 0.0045 & 0.1948 & 0.9410 \\ 
-  IRS-Model & Random & Normal & 0 & 50 & 0.0048 & 0.1966 & 0.9360 \\ 
-  GRTS-Design & Random & Normal & 0 & 50 & 0.0056 & 0.1943 & 0.9000 \\ 
-  GRTS-Model & Random & Normal & 0 & 50 & 0.0057 & 0.1947 & 0.9365 \\ 
+IRS-Design & Random & Normal & 0 & 50 & 0.0019 & 0.1861 & 0.9505 \\ 
+  IRS-Model & Random & Normal & 0 & 50 & 0.0014 & 0.1892 & 0.9445 \\ 
+  GRTS-Design & Random & Normal & 0 & 50 & -0.0040 & 0.1955 & 0.9090 \\ 
+  GRTS-Model & Random & Normal & 0 & 50 & -0.0040 & 0.1965 & 0.9315 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0 & 50 & 0.0116 & 0.9684 & 0.8205 \\ 
-  IRS-Model & Random & Lognormal & 0 & 50 & -0.0235 & 1.0055 & 0.8150 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0005 & 0.9151 & 0.8075 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0175 & 0.9426 & 0.8240 \\ 
+IRS-Design & Random & Lognormal & 0 & 50 & 0.0032 & 0.1843 & 0.9205 \\ 
+  IRS-Model & Random & Lognormal & 0 & 50 & -0.0039 & 0.1945 & 0.9105 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0056 & 0.1944 & 0.8870 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0078 & 0.1967 & 0.9075 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 50 & -0.0006 & 0.1571 & 0.9600 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 50 & -0.0005 & 0.1509 & 0.9435 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0006 & 0.1463 & 0.9240 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 50 & 0.0006 & 0.1443 & 0.9445 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 50 & 0.0025 & 0.1762 & 0.9470 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 50 & 0.0007 & 0.1655 & 0.9305 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0004 & 0.1524 & 0.9115 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 50 & -0.0003 & 0.1499 & 0.9320 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0154 & 0.8578 & 0.8700 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0209 & 0.9091 & 0.8490 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & -0.0046 & 0.8302 & 0.8400 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0167 & 0.8301 & 0.8495 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0004 & 0.1863 & 0.9140 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0076 & 0.1834 & 0.9035 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0016 & 0.1612 & 0.8810 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0017 & 0.1606 & 0.8940 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.5 & 50 & 0.0089 & 0.1720 & 0.9385 \\ 
-  IRS-Model & Random & Normal & 0.5 & 50 & 0.0074 & 0.1580 & 0.9310 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 50 & 0.0060 & 0.1460 & 0.9145 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 50 & 0.0059 & 0.1439 & 0.9365 \\ 
+IRS-Design & Random & Normal & 0.5 & 50 & 0.0037 & 0.1657 & 0.9590 \\ 
+  IRS-Model & Random & Normal & 0.5 & 50 & 0.0027 & 0.1546 & 0.9495 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 50 & -0.0034 & 0.1511 & 0.9170 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 50 & -0.0037 & 0.1504 & 0.9305 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0006 & 0.8479 & 0.8670 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0447 & 0.8845 & 0.8485 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 50 & 0.0022 & 0.8663 & 0.8230 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0044 & 0.8620 & 0.8470 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0023 & 0.1637 & 0.9245 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0055 & 0.1622 & 0.9125 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 50 & -0.0057 & 0.1570 & 0.9000 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0079 & 0.1567 & 0.9100 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 50 & -0.0021 & 0.1185 & 0.9790 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 50 & -0.0023 & 0.0925 & 0.9475 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 50 & -0.0004 & 0.0957 & 0.9280 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 50 & 0.0015 & 0.0874 & 0.9300 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 50 & 0.0053 & 0.1579 & 0.9470 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 50 & 0.0026 & 0.1165 & 0.9315 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 50 & 0.0013 & 0.1074 & 0.9220 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 50 & -0.0007 & 0.0949 & 0.9430 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0191 & 0.6345 & 0.9310 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 50 & 0.0043 & 0.6031 & 0.8850 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & -0.0072 & 0.6496 & 0.8890 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0144 & 0.5873 & 0.8890 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1731 & 0.9220 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0020 & 0.1325 & 0.9135 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1183 & 0.9065 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0002 & 0.1090 & 0.9120 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.9 & 50 & -0.0022 & 0.1451 & 0.9445 \\ 
-  IRS-Model & Random & Normal & 0.9 & 50 & -0.0016 & 0.1018 & 0.9350 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0022 & 0.0963 & 0.9235 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0016 & 0.0881 & 0.9360 \\ 
+IRS-Design & Random & Normal & 0.9 & 50 & 0.0062 & 0.1520 & 0.9525 \\ 
+  IRS-Model & Random & Normal & 0.9 & 50 & 0.0037 & 0.1074 & 0.9525 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0029 & 0.1038 & 0.9340 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0027 & 0.0940 & 0.9360 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 50 & -0.0279 & 0.7417 & 0.9080 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0479 & 0.6787 & 0.8720 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0101 & 0.5712 & 0.8995 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0079 & 0.5143 & 0.8935 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 50 & 0.0053 & 0.1542 & 0.9325 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0012 & 0.1160 & 0.9115 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0033 & 0.1150 & 0.9160 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0031 & 0.1014 & 0.9235 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0 & 100 & -0.0031 & 0.1343 & 0.9430 \\ 
-  IRS-Model & Gridded & Normal & 0 & 100 & -0.0035 & 0.1349 & 0.9405 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0016 & 0.1327 & 0.9200 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0015 & 0.1327 & 0.9395 \\ 
+IRS-Design & Gridded & Normal & 0 & 100 & -0.0029 & 0.1322 & 0.9430 \\ 
+  IRS-Model & Gridded & Normal & 0 & 100 & -0.0033 & 0.1332 & 0.9410 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0041 & 0.1302 & 0.9245 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0038 & 0.1304 & 0.9420 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 100 & 0.0123 & 0.6517 & 0.8595 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 100 & 0.0037 & 0.6698 & 0.8540 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0175 & 0.6309 & 0.8275 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0108 & 0.6450 & 0.8460 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 100 & -0.0052 & 0.1334 & 0.9335 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 100 & -0.0072 & 0.1350 & 0.9300 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0042 & 0.1303 & 0.9035 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0034 & 0.1307 & 0.9265 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0 & 100 & 0.0058 & 0.1363 & 0.9350 \\ 
-  IRS-Model & Random & Normal & 0 & 100 & 0.0062 & 0.1373 & 0.9315 \\ 
-  GRTS-Design & Random & Normal & 0 & 100 & 0.0033 & 0.1351 & 0.9175 \\ 
-  GRTS-Model & Random & Normal & 0 & 100 & 0.0033 & 0.1350 & 0.9460 \\ 
+IRS-Design & Random & Normal & 0 & 100 & -0.0066 & 0.1366 & 0.9405 \\ 
+  IRS-Model & Random & Normal & 0 & 100 & -0.0067 & 0.1370 & 0.9385 \\ 
+  GRTS-Design & Random & Normal & 0 & 100 & 0.0028 & 0.1364 & 0.9180 \\ 
+  GRTS-Model & Random & Normal & 0 & 100 & 0.0029 & 0.1363 & 0.9345 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0 & 100 & -0.0187 & 0.6427 & 0.8595 \\ 
-  IRS-Model & Random & Lognormal & 0 & 100 & -0.0394 & 0.6726 & 0.8565 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0049 & 0.6366 & 0.8445 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 100 & 0.0011 & 0.6431 & 0.8585 \\ 
+IRS-Design & Random & Lognormal & 0 & 100 & -0.0051 & 0.1348 & 0.9280 \\ 
+  IRS-Model & Random & Lognormal & 0 & 100 & -0.0075 & 0.1372 & 0.9230 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0002 & 0.1375 & 0.8840 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 100 & -0.0003 & 0.1377 & 0.9090 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0005 & 0.1070 & 0.9640 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0006 & 0.1041 & 0.9335 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 100 & -0.0010 & 0.1014 & 0.9145 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 100 & -0.0011 & 0.1006 & 0.9365 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0016 & 0.1177 & 0.9540 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0023 & 0.1072 & 0.9470 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 100 & 0.0051 & 0.0998 & 0.9300 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 100 & 0.0039 & 0.0982 & 0.9470 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0156 & 0.5285 & 0.9060 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0020 & 0.5346 & 0.8800 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0061 & 0.4945 & 0.8595 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0015 & 0.4965 & 0.8705 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 100 & -0.0021 & 0.1211 & 0.9445 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 100 & -0.0058 & 0.1153 & 0.9295 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0041 & 0.1090 & 0.9085 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0019 & 0.1090 & 0.9180 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.5 & 100 & 0.0002 & 0.1169 & 0.9440 \\ 
-  IRS-Model & Random & Normal & 0.5 & 100 & 0.0023 & 0.1042 & 0.9450 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0030 & 0.1032 & 0.9140 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0024 & 0.1022 & 0.9335 \\ 
+IRS-Design & Random & Normal & 0.5 & 100 & -0.0064 & 0.1222 & 0.9440 \\ 
+  IRS-Model & Random & Normal & 0.5 & 100 & -0.0049 & 0.1073 & 0.9440 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0013 & 0.1041 & 0.9155 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0014 & 0.1020 & 0.9400 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0020 & 0.5609 & 0.8895 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0209 & 0.5570 & 0.8835 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 100 & -0.0106 & 0.5261 & 0.8800 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0199 & 0.5358 & 0.8875 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0063 & 0.1204 & 0.9355 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0097 & 0.1150 & 0.9275 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 100 & 0.0003 & 0.1092 & 0.8960 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0004 & 0.1088 & 0.9120 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 100 & 0.0008 & 0.0702 & 0.9915 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 100 & 0.0016 & 0.0573 & 0.9510 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0002 & 0.0600 & 0.9315 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 100 & -0.0001 & 0.0540 & 0.9515 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 100 & -0.0007 & 0.1059 & 0.9605 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 100 & -0.0019 & 0.0700 & 0.9445 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0044 & 0.0655 & 0.9435 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 100 & 0.0030 & 0.0585 & 0.9440 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0165 & 0.4844 & 0.9595 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0062 & 0.4274 & 0.8975 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0056 & 0.3959 & 0.9135 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0015 & 0.3764 & 0.9200 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 100 & -0.0011 & 0.1092 & 0.9540 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0033 & 0.0779 & 0.9330 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0035 & 0.0757 & 0.9190 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0017 & 0.0693 & 0.9190 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.9 & 100 & -0.0034 & 0.1021 & 0.9425 \\ 
-  IRS-Model & Random & Normal & 0.9 & 100 & -0.0015 & 0.0611 & 0.9480 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0614 & 0.9275 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 100 & -0.0004 & 0.0548 & 0.9410 \\ 
+IRS-Design & Random & Normal & 0.9 & 100 & -0.0048 & 0.1095 & 0.9430 \\ 
+  IRS-Model & Random & Normal & 0.9 & 100 & -0.0017 & 0.0671 & 0.9535 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0674 & 0.9265 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 100 & 0.0000 & 0.0595 & 0.9420 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0048 & 0.5829 & 0.9275 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0287 & 0.5423 & 0.9205 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 100 & 0.0010 & 0.4351 & 0.9180 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0025 & 0.3975 & 0.9140 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0045 & 0.1110 & 0.9375 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0052 & 0.0771 & 0.9310 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 100 & -0.0009 & 0.0732 & 0.9195 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0007 & 0.0661 & 0.9165 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0 & 200 & 0.0005 & 0.0882 & 0.9455 \\ 
-  IRS-Model & Gridded & Normal & 0 & 200 & 0.0005 & 0.0882 & 0.9435 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 200 & 0.0034 & 0.0882 & 0.9430 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 200 & 0.0034 & 0.0882 & 0.9500 \\ 
+IRS-Design & Gridded & Normal & 0 & 200 & 0.0013 & 0.0882 & 0.9550 \\ 
+  IRS-Model & Gridded & Normal & 0 & 200 & 0.0013 & 0.0886 & 0.9530 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9380 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9460 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0109 & 0.4122 & 0.8815 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 200 & 0.0102 & 0.4128 & 0.8790 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0128 & 0.4355 & 0.8825 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0132 & 0.4362 & 0.8865 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0001 & 0.0880 & 0.9350 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 200 & -0.0002 & 0.0883 & 0.9355 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0034 & 0.0898 & 0.9285 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0035 & 0.0900 & 0.9335 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0 & 200 & -0.0025 & 0.0869 & 0.9500 \\ 
-  IRS-Model & Random & Normal & 0 & 200 & -0.0024 & 0.0872 & 0.9505 \\ 
-  GRTS-Design & Random & Normal & 0 & 200 & 0.0027 & 0.0875 & 0.9470 \\ 
-  GRTS-Model & Random & Normal & 0 & 200 & 0.0027 & 0.0876 & 0.9505 \\ 
+IRS-Design & Random & Normal & 0 & 200 & -0.0014 & 0.0893 & 0.9465 \\ 
+  IRS-Model & Random & Normal & 0 & 200 & -0.0015 & 0.0896 & 0.9465 \\ 
+  GRTS-Design & Random & Normal & 0 & 200 & 0.0007 & 0.0868 & 0.9460 \\ 
+  GRTS-Model & Random & Normal & 0 & 200 & 0.0007 & 0.0867 & 0.9490 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0 & 200 & 0.0166 & 0.4146 & 0.8975 \\ 
-  IRS-Model & Random & Lognormal & 0 & 200 & 0.0074 & 0.4479 & 0.8930 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 200 & -0.0044 & 0.4800 & 0.8870 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 200 & -0.0063 & 0.4820 & 0.8875 \\ 
+IRS-Design & Random & Lognormal & 0 & 200 & -0.0028 & 0.0891 & 0.9425 \\ 
+  IRS-Model & Random & Lognormal & 0 & 200 & -0.0033 & 0.0896 & 0.9395 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 200 & 0.0015 & 0.0860 & 0.9365 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 200 & 0.0012 & 0.0861 & 0.9415 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 200 & -0.0003 & 0.0682 & 0.9735 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 200 & -0.0005 & 0.0664 & 0.9435 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 200 & 0.0002 & 0.0661 & 0.9415 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 200 & 0.0001 & 0.0656 & 0.9450 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 200 & 0.0022 & 0.0788 & 0.9535 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 200 & 0.0008 & 0.0678 & 0.9580 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 200 & -0.0024 & 0.0671 & 0.9335 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 200 & -0.0030 & 0.0661 & 0.9410 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0008 & 0.3556 & 0.9379 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0030 & 0.3540 & 0.9144 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0001 & 0.3305 & 0.9029 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.3287 & 0.9049 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 200 & 0.0016 & 0.0816 & 0.9420 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0000 & 0.0739 & 0.9335 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0033 & 0.0744 & 0.9330 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.0742 & 0.9350 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.5 & 200 & 0.0039 & 0.0771 & 0.9464 \\ 
-  IRS-Model & Random & Normal & 0.5 & 200 & 0.0024 & 0.0684 & 0.9489 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0001 & 0.0658 & 0.9374 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 200 & -0.0003 & 0.0652 & 0.9444 \\ 
+IRS-Design & Random & Normal & 0.5 & 200 & -0.0016 & 0.0790 & 0.9480 \\ 
+  IRS-Model & Random & Normal & 0.5 & 200 & -0.0010 & 0.0690 & 0.9475 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0002 & 0.0652 & 0.9455 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 200 & 0.0001 & 0.0640 & 0.9500 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 200 & 0.0091 & 0.3861 & 0.9164 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0048 & 0.3783 & 0.9069 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 200 & 0.0012 & 0.4088 & 0.9029 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0005 & 0.3941 & 0.9029 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 200 & -0.0027 & 0.0809 & 0.9494 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0037 & 0.0732 & 0.9454 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 200 & -0.0008 & 0.0671 & 0.9434 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0012 & 0.0662 & 0.9479 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0002 & 0.0413 & 0.9965 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 200 & -0.0008 & 0.0356 & 0.9475 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0004 & 0.0381 & 0.9475 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0000 & 0.0349 & 0.9490 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0021 & 0.0704 & 0.9570 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 200 & 0.0005 & 0.0411 & 0.9590 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0005 & 0.0423 & 0.9370 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0012 & 0.0377 & 0.9450 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0026 & 0.2868 & 0.9835 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0016 & 0.2560 & 0.9154 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0038 & 0.2341 & 0.9309 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0062 & 0.2290 & 0.9239 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0014 & 0.0744 & 0.9520 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0002 & 0.0480 & 0.9355 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0014 & 0.0499 & 0.9380 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0022 & 0.0459 & 0.9380 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Normal & 0.9 & 200 & -0.0037 & 0.0673 & 0.9410 \\ 
-  IRS-Model & Random & Normal & 0.9 & 200 & -0.0012 & 0.0385 & 0.9505 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 200 & 0.0002 & 0.0389 & 0.9370 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0005 & 0.0349 & 0.9435 \\ 
+IRS-Design & Random & Normal & 0.9 & 200 & -0.0009 & 0.0725 & 0.9470 \\ 
+  IRS-Model & Random & Normal & 0.9 & 200 & -0.0010 & 0.0409 & 0.9450 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 200 & -0.0005 & 0.0405 & 0.9490 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0001 & 0.0354 & 0.9510 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
@@
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 200 & 0.0154 & 0.3554 & 0.9313 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 200 & 0.0008 & 0.2657 & 0.9238 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 200 & 0.0058 & 0.2299 & 0.9218 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 200 & 0.0043 & 0.2206 & 0.9147 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 200 & -0.0006 & 0.0750 & 0.9393 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 200 & -0.0025 & 0.0461 & 0.9418 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 200 & -0.0013 & 0.0452 & 0.9348 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 200 & -0.0012 & 0.0397 & 0.9418 \\ 
    \hline
 \end{tabular}
+}
 \caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}

### 2021-12-18 inst/scripts/application.R

@@ -1,0 +1,159 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(here)
+library(sptotal)
+library(spsurvey) # depends on sf
+library(DvMsp) # need package installed to load data
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## manipulate data
+
+data(nla_df)
+
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  ungroup() %>%
+  summarise(maxnonmiss = max(nonmiss))
+
+## HG never measured twice at the same site.
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  summarise(sumnomiss = sum(nonmiss == 0))
+
+## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
+## in duplicate sites, fill in the value for RDis_IX
+nla_nomiss <- nla_df %>%
+  group_by(SITE_ID) %>%
+  fill(RDis_IX, .direction = "downup") %>%
+  ungroup() %>%
+  filter(!is.na(TOTALHG_RESULT)) %>%
+  rename(total_hg = "TOTALHG_RESULT")
+
+nla_nomiss_both <- nla_nomiss
+## check
+nla_nomiss_both %>%
+  group_by(SITE_ID) %>% count() %>%
+  ungroup() %>%
+  summarise(not1 = sum(n != 1))
+
+## transform coordinates
+coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
+                     lat = nla_nomiss_both$INDEX_LAT_DD,
+                     lon = nla_nomiss_both$INDEX_LON_DD)
+nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
+nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
+
+## find population mean
+realized_mean <- mean(nla_nomiss_both$total_hg)
+
+## set seed and other preliminaries
+set.seed(080421)
+n <- 100
+
+## IRS Sample, Model Analysis
+lakesobs <- nla_nomiss_both %>% sample_n(n)
+lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
+lakesunobs$total_hg <- NA
+
[…]

### 2021-12-18 inst/scripts/figures/application.R

@@ -1,0 +1,131 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(cowplot)
+library(here)
+library(sptotal)
+library(DvMsp) # need package installed to load data
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## manipulate data
+
+data(nla_df)
+
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  ungroup() %>%
+  summarise(maxnonmiss = max(nonmiss))
+
+## HG never measured twice at the same site.
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
+  summarise(sumnomiss = sum(nonmiss == 0))
+
+## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
+## in duplicate sites, fill in the value for RDis_IX
+nla_nomiss <- nla_df %>%
+  group_by(SITE_ID) %>%
+  fill(RDis_IX, .direction = "downup") %>%
+  ungroup() %>%
+  filter(!is.na(TOTALHG_RESULT)) %>%
+  rename(total_hg = "TOTALHG_RESULT")
+
+nla_nomiss_both <- nla_nomiss
+## check
+nla_nomiss_both %>%
+  group_by(SITE_ID) %>% count() %>%
+  ungroup() %>%
+  summarise(not1 = sum(n != 1))
+
+## transform coordinates
+coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
+                     lat = nla_nomiss_both$INDEX_LAT_DD,
+                     lon = nla_nomiss_both$INDEX_LON_DD)
+nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
+nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
+
+## figure 4a
+mercury_map <- nla_nomiss_both %>%
+  ggplot(aes(x = xcoords, y = ycoords)) +
+  geom_point(aes(colour = total_hg)) +
+  scale_colour_viridis_c() +
+  labs(x = "", y = "", colour = "Hg") +
+  coord_quickmap() +
+  theme_bw(base_size = 10) +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
[…]

### 2021-12-18 inst/scripts/figures/background.R

@@ -1,0 +1,209 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(here)
+library(cowplot)
+library(ggpubr)
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## set base size
+base_size <- 9
+
+## set a reproducible seed
+set.seed(10102021)
+
+## load sim_pop function (if running from within package project)
+source(here("R", "sim_pop.R"))
+
+## load sim_pop function (if running after installation)
+# library(DvMsp)
+
+### simulate populations
+N <- 9
+pop_df <- sim_pop(
+  N = N,
+  gridded = TRUE,
+  cortype = "Exponential",
+  psill = 0.9,
+  range = 1,
+  nugget = 0.1
+)
+
+### select samples
+n <- 4
+samp_df <- pop_df %>% sample_n(n)
+samp_df2 <- pop_df %>% sample_n(n)
+samp_df3 <- pop_df %>% sample_n(n)
+
+
+### create plots
+d1 <- pop_df %>%
+  ggplot(aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(
+    data = samp_df,
+    size = 5.5,
+    shape = 1,
+    show.legend = FALSE,
+    stroke = 1.5
+  ) +
+  labs(title = "Design 1", colour = "Response") +
+  theme_bw(base_size = base_size) +
+  theme(
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text = element_blank(),
+    legend.position = "none",
+    panel.grid.minor = element_blank(),
+    panel.grid.major = element_blank()
+  ) +
+  xlim(-0.1, 1.1) +
+  ylim(-0.1, 1.1)
+
+d2 <- pop_df %>%
+  ggplot(aes(x = x, y = y)) +
+  geom_point(aes(colour = response), size = 4) +
+  scale_colour_viridis_c() +
+  geom_point(
+    data = samp_df2,
+    size = 5.5,
+    shape = 1,
+    show.legend = FALSE,
+    stroke = 1.5
+  ) +
+  labs(title = "Design 2", colour = "Response") +
+  theme_bw(base_size = base_size) +
+  theme(
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
[…]

### 2021-12-18 inst/scripts/figures/simulation.R

@@ -1,0 +1,90 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(here)
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## set base size
+base_size <- 10
+
+## read in files and create data
+files <- list.files(path = here("inst", "output", "simulation_summary"), pattern = "*.csv", full.names = TRUE)
+
+combo_data <- purrr::map_df(files,
+                            ~read_csv(.x) %>% mutate(filename = .x))
+
+combo_data <- combo_data %>%
+  dplyr::filter(gridded == FALSE) %>%
+  mutate(sim = interaction(n, psill, resptype)) %>% ## , gridded
+  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
+                                            "Design GRTS", "Model GRTS"))) %>%
+  mutate(resptype = fct_relevel(resptype, c("normal", "lognormal"))) %>%
+  group_by(sim) %>%
+  mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
+  fill(designirsrmspe, .direction = "downup") %>%
+  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
+  ungroup() %>%
+  mutate(psill_ratio = 1 - nugget_ratio) %>%
+  mutate(n_factor = factor(n))
+
+# figure 2
+colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
+psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
+rmspe_eff <- combo_data %>%
+  ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
+  facet_grid(
+    psill_ratio ~ resptype,
+    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
+  ) +
+  geom_jitter(width = 0.24, size = 2) +
+  scale_colour_manual(values = colour_scale) +
+  theme_bw(base_size = base_size) +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative Efficiency") +
+  theme(
+    panel.grid.major = element_blank(),
[…]

### 2021-12-20 inst/manuscript/manuscript.Rmd
style guide updates

@@
 abstract: |
+  1. 
+  2.
+  3.
+  4.
   The design-based and model-based approaches to frequentist statistical inference lie on fundamentally different foundations.
@@
   \usepackage{subcaption}
+  \usepackage{setspace}
+  \doublespacing
 output: rticles::elsevier_article
 ---
+
+# Keywords {.unnumbered}
+
+Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based inference; Spatially balanced sampling; Spatial covariance;
 
@@
 In Section \ref{sec:background}, we introduce and provide relevant background for the design-based and model-based approaches to finite population spatial data.
-In Section \ref{sec:numstudy}, we use a simulation study to compare the performance of the approaches in a variety of scenarios.
-In Section \ref{application}, we compare the performance of the approaches on real data that contains mercury concentration in lakes from the contiguous United States.
+In Section \ref{sec:mm}, we describe how we compare performance of the approaches wiht a simulation study and an analysis of real data that contains mercury concentration in lakes from the contiguous United States.
+In Section \ref{sec:results}, we present results from the simulation study and the analysis of mercury concentrations.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
 
-# Background {#sec:background}
+## Background {#sec:background}
 
@@
 
-## Comparing Design-Based and Model-Based Approaches {#subsec:dvm_compare}
+### Comparing Design-Based and Model-Based Approaches {#subsec:dvm_compare}
 
@@
 @cressie1993statistics and @schabenberger2017statistical provide thorough reviews of model-based approaches for spatial data.
-In Figure \ref{fig:fig1}, we provide a visual comparison of the design-based and model-based approaches (@verhoef2002sampling and @brus2021statistical provide similar figures).
+In Fig.
+\ref{fig:fig1}, we provide a visual comparison of the design-based and model-based approaches (@verhoef2002sampling and @brus2021statistical provide similar figures).
 
@@
 
-## Spatially Balanced Design and Analysis {#subsec:spb_design}
+### Spatially Balanced Design and Analysis {#subsec:spb_design}
 
@@
 
-## Finite Population Block Kriging
+### Finite Population Block Kriging
 
@@
 
-# Simulation Study {#sec:numstudy}
+# Materials and Methods {#sec:mm}
+
+## Simulation Study {#sec:mm_sim}
 
@@
 
+## Application {#sec:mm_app}
+
+The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
+We will use data from the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality for lakes in the contiguous United States [@USEPA2012NLA].
+Specifically, we will analyze mercury concentration in lakes.
+Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
+For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
+
+# Results {#sec:results}
+
+## Simulation Study {#sec:r_sim}
+
 The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
-Tables for average bias in all 36 simulation scenarios are provided in the supplementary material.
+Tables for average bias in all 36 simulation scenarios are provided in the supporting information.
 
-Figure \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
+Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
 <!--- figure 2 -->
@@
 \end{equation*}
-When there is no spatial correlation (Figure \ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+When there is no spatial correlation (Fig.
+\ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
 So using the GRTS sampling plan or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial correlation.
-When there is spatial correlation (Figure \ref{fig:figeff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
+When there is spatial correlation (Fig.
+\ref{fig:figeff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
 As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
@@
 These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
-Tables for rMS(P)E in all 36 simulation scenarios are provided in the supplementary material.
+Tables for rMS(P)E in all 36 simulation scenarios are provided in the supporting information.
 
@@
 
-Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
+Fig. \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
 <!--- figure 3 -->
@@
 These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
-Tables for interval coverage in all 36 simulation scenarios are provided in the supplementary material.
+Tables for interval coverage in all 36 simulation scenarios are provided in the supporting information.
 
-# Application {#application}
+## Application {#sec:r_app}
 
-The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality for lakes in the contiguous United States.
-Specifically, we will analyze mercury concentration in lakes.
-Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-
-Figure \ref{fig:merc} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States.
-The realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
-TALK ABOUT EMPIRICAL SEMIVARIOGRAM IN THIS PARAGRAPH.
-
+Fig.
+\ref{fig:merc} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
+Mercury concentration exhibits some spatial patterning, with high mercury concentrations in lakes in the northeast and north central United States.
+Fig.
+\ref{fig:merc} also shows the spatial dependence in mercury concentration via the empirical semivariogram.
+The empirical semivariogram can be used as a tool to visualize spatial dependence.
+It quantifes the halved squared differences (semivariance) among mercury concentration at different distances apart.
+When a process is spatially correlated, the semivariance tends to be smaller at small distances and larger at large distances.
+Together, the map, histogram, and semivariogram in Fig.
+\ref{fig:merc} suggest that mercury concentration is skewed and exhibits spatial dependence.
+Lastly we note that the realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
 \begin{figure}
@@
 
-<!--- To better understand the dependence structure in mercury concentration, the empirical semivariogram can be visualized alongside a fit using the model-based approaches.
-The empirical semivariogram quantifies the halved squared differences (semivariance) among response values at different distances apart.
-If a process exhibits strong spatial dependence, the empirical semivariogram will have small values at small distances and large values at large distances.
-Figure \ref{fig:figsv} shows the empirical semivariogram quantifying the average semivariance for several distances.
-Overlain onto Figure \ref{fig:figsv} is the estimated semivariance obtained using the covariance parameters from the REML fit of GRTS-Model.
-Figure \ref{fig:figsv} provides evidence that there is strong correlation in mercury concentration among the sites.
--->
-
 # Discussion {#sec:discussion}
@@
 
-# Data and Code Availability {.unnumbered}
-
-This manuscript has a supplementary R package that contains all of the data and code used. Instructions for download at available at
-
-[https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
-
-# Supplementary Material {.unnumbered}
-
-In the supplementary material, we provide tables presenting summary statistics for all 36 simulation scenarios.
-
 # Acknowledgments {.unnumbered}
@@
 
+# Conflict of Interest Statement {.unnumbered}
+
+There are no conflicts of interest for any of the authors.
+
+# Data and Code Availability {.unnumbered}
+
+This manuscript has a supplementary R package that contains all of the data and code used in its creation. The supplementary R package is hosted on GitHub. Instructions for download at available at
+
+[https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
+
+# Supporting Information {.unnumbered}
+
+In the supporting information, we provide tables presenting summary statistics for all 36 simulation scenarios.
+
+# Author Contributions {.unnumbered}
+
+All authors conceived the ideas; All authors designed methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
+
 # References {#references .unnumbered}

### 2021-12-20 inst/manuscript/manuscript.tex
style guide updates

@@
 
-\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption}
+\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption} \usepackage{setspace} \doublespacing
 
@@
   \begin{abstract}
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \item
+  \item
+  \item
+  \item
     The design-based and model-based approaches to frequentist statistical
@@
     perform better than sampling plans ignoring spatial locations
-  (non-spatially balanced samples), regardless of whether design-based or
-  model-based approaches were used to analyze the data. We also find that
-  within sampling plans, the model-based approaches often outperform
-  design-based approaches, even for skewed data. This gap in performance
-  is small when spatially balanced samples are used but large when
-  non-spatially balanced samples are used.
+    (non-spatially balanced samples), regardless of whether design-based
+    or model-based approaches were used to analyze the data. We also find
+    that within sampling plans, the model-based approaches often
+    outperform design-based approaches, even for skewed data. This gap in
+    performance is small when spatially balanced samples are used but
+    large when non-spatially balanced samples are used.
+  \end{enumerate}
   \end{abstract}
@@
  \end{frontmatter}
+
+\hypertarget{keywords}{%
+\section*{Keywords}\label{keywords}}
+\addcontentsline{toc}{section}{Keywords}
+
+Design-based inference; Finite Population Block Kriging (FPBK);
+Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based
+inference; Spatially balanced sampling; Spatial covariance;
 
@@
 the design-based and model-based approaches to finite population spatial
-data. In Section \ref{sec:numstudy}, we use a simulation study to
-compare the performance of the approaches in a variety of scenarios. In
-Section \ref{application}, we compare the performance of the approaches
-on real data that contains mercury concentration in lakes from the
-contiguous United States. And in Section \ref{sec:discussion}, we end
-with a discussion and provide directions for future research.
+data. In Section \ref{sec:mm}, we describe how we compare performance of
+the approaches wiht a simulation study and an analysis of real data that
+contains mercury concentration in lakes from the contiguous United
+States. In Section \ref{sec:results}, we present results from the
+simulation study and the analysis of mercury concentrations. And in
+Section \ref{sec:discussion}, we end with a discussion and provide
+directions for future research.
 
 \hypertarget{sec:background}{%
-\section{Background}\label{sec:background}}
+\subsection{Background}\label{sec:background}}
 
@@
 \hypertarget{subsec:dvm_compare}{%
-\subsection{Comparing Design-Based and Model-Based
+\subsubsection{Comparing Design-Based and Model-Based
 Approaches}\label{subsec:dvm_compare}}
@@
 Schabenberger and Gotway (2017) provide thorough reviews of model-based
-approaches for spatial data. In Figure \ref{fig:fig1}, we provide a
-visual comparison of the design-based and model-based approaches (Ver
-Hoef (2002) and Brus (2021) provide similar figures).
+approaches for spatial data. In Fig. \ref{fig:fig1}, we provide a visual
+comparison of the design-based and model-based approaches (Ver Hoef
+(2002) and Brus (2021) provide similar figures).
 
@@
 \hypertarget{subsec:spb_design}{%
-\subsection{Spatially Balanced Design and
+\subsubsection{Spatially Balanced Design and
 Analysis}\label{subsec:spb_design}}
@@
 \hypertarget{finite-population-block-kriging}{%
-\subsection{Finite Population Block
+\subsubsection{Finite Population Block
 Kriging}\label{finite-population-block-kriging}}
@@
 
-\hypertarget{sec:numstudy}{%
-\section{Simulation Study}\label{sec:numstudy}}
+\hypertarget{sec:mm}{%
+\section{Materials and Methods}\label{sec:mm}}
+
+\hypertarget{sec:mm_sim}{%
+\subsection{Simulation Study}\label{sec:mm_sim}}
 
@@
 
+\hypertarget{sec:mm_app}{%
+\subsection{Application}\label{sec:mm_app}}
+
+The Environmental Protection Agency (EPA), states, and tribes
+periodically conduct National Aquatic Research Surveys (NARS) in the
+United States to assess the water quality of various bodies of water. We
+will use data from the 2012 National Lakes Assessment (NLA), which
+measures various aspects of lake health and water quality for lakes in
+the contiguous United States (USEPA, 2012). Specifically, we will
+analyze mercury concentration in lakes. Although we know the true mean
+mercury concentration values for the 986 lakes from the 2012 NLA, we
+will explore whether or not we obtain an adequately precise estimate for
+the realized mean mercury concentration if we sample only 100 of the 986
+lakes. For each of the four familiar sampling-analysis combinations
+(IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
+
+\hypertarget{sec:results}{%
+\section{Results}\label{sec:results}}
+
+\hypertarget{sec:r_sim}{%
+\subsection{Simulation Study}\label{sec:r_sim}}
+
 The average bias was nearly zero for all four combinations in all 36
@@
 Tables for average bias in all 36 simulation scenarios are provided in
-the supplementary material.
+the supporting information.
 
-Figure \ref{fig:figeff} shows the relative rMS(P)E of the four
-approaches from Table \ref{tab:designanalysis} using the random location
-layout with ``IRS-Design'' as the baseline.
+Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four approaches
+from Table \ref{tab:designanalysis} using the random location layout
+with ``IRS-Design'' as the baseline.
 
@@
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
-\end{equation*} When there is no spatial correlation (Figure
+\end{equation*} When there is no spatial correlation (Fig.
 \ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
@@
 loss in efficiency compared to IRS-Design when there is no spatial
-correlation. When there is spatial correlation (Figure \ref{fig:figeff},
+correlation. When there is spatial correlation (Fig. \ref{fig:figeff},
 ``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
@@
 layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
-are provided in the supplementary material.
+are provided in the supporting information.
 
@@
 
-Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
-the four sampling-analysis combinations in the random location layout.
+Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
+four sampling-analysis combinations in the random location layout.
 
@@
 so we omit a grid location layout figure here. Tables for interval
-coverage in all 36 simulation scenarios are provided in the
-supplementary material.
+coverage in all 36 simulation scenarios are provided in the supporting
+information.
 
-\hypertarget{application}{%
-\section{Application}\label{application}}
+\hypertarget{sec:r_app}{%
+\subsection{Application}\label{sec:r_app}}
 
-The Environmental Protection Agency (EPA), states, and tribes
-periodically conduct National Aquatic Research Surveys (NARS) in the
-United States to assess the water quality of various bodies of water. We
-will use the 2012 National Lakes Assessment (NLA), which measures
-various aspects of lake health and water quality for lakes in the
-contiguous United States. Specifically, we will analyze mercury
-concentration in lakes. Although we know the true mean mercury
-concentration values for the 986 lakes from the 2012 NLA, we will
-explore whether or not we obtain an adequately precise estimate for the
-realized mean mercury concentration if we sample only 100 of the 986
-lakes.
-
-Figure \ref{fig:merc} shows that mercury concentration is right-skewed,
+Fig. \ref{fig:merc} shows that mercury concentration is right-skewed,
 with most lakes having a low value of mercury concentration but a few
 having a much higher concentration. Mercury concentration exhibits some
-spatial correlation, with high mercury concentrations in lakes in the
-northeast and north central United States. The realized mean mercury
-concentration in the 986 lakes is 103.2 ng / g. TALK ABOUT EMPIRICAL
-SEMIVARIOGRAM IN THIS PARAGRAPH.
+spatial patterning, with high mercury concentrations in lakes in the
+northeast and north central United States. Fig. \ref{fig:merc} also
+shows the spatial dependence in mercury concentration via the empirical
+semivariogram. The empirical semivariogram can be used as a tool to
+visualize spatial dependence. It quantifes the halved squared
+differences (semivariance) among mercury concentration at different
+distances apart. When a process is spatially correlated, the
+semivariance tends to be smaller at small distances and larger at large
+distances. Together, the map, histogram, and semivariogram in Fig.
+\ref{fig:merc} suggest that mercury concentration is skewed and exhibits
+spatial dependence. Lastly we note that the realized mean mercury
+concentration in the 986 lakes is 103.2 ng / g.
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\section*{Data and Code Availability}\label{data-and-code-availability}}
-\addcontentsline{toc}{section}{Data and Code Availability}
-
-This manuscript has a supplementary R package that contains all of the
-data and code used. Instructions for download at available at
-
-\url{https://github.com/michaeldumelle/DvMsp}.
-
-\hypertarget{supplementary-material}{%
-\section*{Supplementary Material}\label{supplementary-material}}
-\addcontentsline{toc}{section}{Supplementary Material}
-
-In the supplementary material, we provide tables presenting summary
-statistics for all 36 simulation scenarios.
-
 \hypertarget{acknowledgments}{%
@@
 products, services, or enterprises.
+
+\hypertarget{conflict-of-interest-statement}{%
+\section*{Conflict of Interest
+Statement}\label{conflict-of-interest-statement}}
+\addcontentsline{toc}{section}{Conflict of Interest Statement}
+
+There are no conflicts of interest for any of the authors.
+
+\hypertarget{data-and-code-availability}{%
+\section*{Data and Code Availability}\label{data-and-code-availability}}
+\addcontentsline{toc}{section}{Data and Code Availability}
+
+This manuscript has a supplementary R package that contains all of the
+data and code used in its creation. The supplementary R package is
+hosted on GitHub. Instructions for download at available at
+
+\url{https://github.com/michaeldumelle/DvMsp}.
+
+\hypertarget{supporting-information}{%
+\section*{Supporting Information}\label{supporting-information}}
+\addcontentsline{toc}{section}{Supporting Information}
+
+In the supporting information, we provide tables presenting summary
+statistics for all 36 simulation scenarios.
+
+\hypertarget{author-contributions}{%
+\section*{Author Contributions}\label{author-contributions}}
+\addcontentsline{toc}{section}{Author Contributions}
+
+All authors conceived the ideas; All authors designed methodology; MD
+and MH performed the simulations and analyzed the data; MD and MH led
+the writing of the manuscript; All authors contributed critically to the
+drafts and gave final approval for publication.
 
@@
 
+\leavevmode\hypertarget{ref-USEPA2012NLA}{}%
+USEPA, 2012. National lakes assessment 2012.
+https://www.epa.gov/national-aquatic-resource-surveys/national-results-and-regional-highlights-national-lakes-assessment.
+
 \leavevmode\hypertarget{ref-verhoef2002sampling}{}%

### 2021-12-20 inst/manuscript/supporting.tex
style guide updates

--- 2021-12-18 inst/manuscript/supplementary.tex
@@
             pdfauthor={},
-            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supplementary Material.},
+            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supporting Information.},
             colorlinks=false,
@@
   \title{A comparison of design-based and model-based approaches for finite
-population spatial data -- Supplementary Material.}
+population spatial data -- Supporting Information.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-\hypertarget{sec:numstudy}{%
-\section{Numerical Study Tables}\label{sec:numstudy}}
+\hypertarget{sec:simtabs}{%
+\section{Simulation Study Tables}\label{sec:simtabs}}
 

### 2021-12-21 inst/manuscript/manuscript.Rmd
updated abstract

@@
   1.
-  2.
-  3.
-  4.
 The design-based and model-based approaches to frequentist statistical inference lie on fundamentally different foundations.
@@
 In the model-based approach, inference depends on distributional assumptions.
-In this manuscript, we compare the approaches for finite population spatial data.
-We first provide relevant background for the approaches and then use a simulation study and an analysis of real mercury concentration data to numerically compare them.
-We find that sampling plans that incorporate spatial locations (spatially balanced samples) perform better than sampling plans ignoring spatial locations (non-spatially balanced samples), regardless of whether design-based or model-based approaches were used to analyze the data.
-We also find that within sampling plans, the model-based approaches often outperform design-based approaches, even for skewed data.
-This gap in performance is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
+We compare the approaches for finite population spatial data.
+  2.
+We provide relevant background for the design-based and model-based approaches and then study their performance using simulations and an analysis of real mercury concentration data.
+In the simulations, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
+In the simualtions and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+  3.
+When studying the simulations and mercury concentration data, we found that regardless of the strength of spatial dependence in the data, sampling plans that incorporate spatial locations (spatially balanced samples) generally outperform sampling plans that ignore spatial locations (non-spatially balanced samples).
+We also found that model-based analyses tend to outperform design-based analyses, even when the data are skewed (and by consequence, the model-based distributional assumptions violated).
+The performance gap between the analysis approaches is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
+This suggests that the sampling choice (spatially balanced samples versus non-spatially balanced samples) is most important when using a design-based analysis.
+  4.
+There are many benefits and drawbacks practitioners must consider when choosing between design-based and model-based approaches for finite population spatial data.
+We provide relevant background contextualizing each approach and study their properties in a variety of scenarios, making recommendations for use based on the practitioner's goals.
 journal: "Methods in Ecology and Evolution"

### 2021-12-21 inst/manuscript/manuscript.tex
updated abstract

@@
   \def\labelenumi{\arabic{enumi}.}
-  \item
-  \item
-  \item
+  \tightlist
   \item
@@
     model-based approach, inference depends on distributional assumptions.
-    In this manuscript, we compare the approaches for finite population
-    spatial data. We first provide relevant background for the approaches
-    and then use a simulation study and an analysis of real mercury
-    concentration data to numerically compare them. We find that sampling
-    plans that incorporate spatial locations (spatially balanced samples)
-    perform better than sampling plans ignoring spatial locations
-    (non-spatially balanced samples), regardless of whether design-based
-    or model-based approaches were used to analyze the data. We also find
-    that within sampling plans, the model-based approaches often
-    outperform design-based approaches, even for skewed data. This gap in
-    performance is small when spatially balanced samples are used but
-    large when non-spatially balanced samples are used.
+    We compare the approaches for finite population spatial data.
+  \item
+    We provide relevant background for the design-based and model-based
+    approaches and then study their performance using simulations and an
+    analysis of real mercury concentration data. In the simulations, a
+    variety of sample sizes, location layouts, dependence structures, and
+    response types are considered. In the simualtions and real data
+    analysis, the population mean is the parameter of interest and
+    performance is measured using statistics like bias, squared error, and
+    interval coverage.
+  \item
+    When studying the simulations and mercury concentration data, we found
+    that regardless of the strength of spatial dependence in the data,
+    sampling plans that incorporate spatial locations (spatially balanced
+    samples) generally outperform sampling plans that ignore spatial
+    locations (non-spatially balanced samples). We also found that
+    model-based analyses tend to outperform design-based analyses, even
+    when the data are skewed (and by consequence, the model-based
+    distributional assumptions violated). The performance gap between the
+    analysis approaches is small when spatially balanced samples are used
+    but large when non-spatially balanced samples are used. This suggests
+    that the sampling choice (spatially balanced samples versus
+    non-spatially balanced samples) is most important when using a
+    design-based analysis.
+  \item
+    There are many benefits and drawbacks practitioners must consider when
+    choosing between design-based and model-based approaches for finite
+    population spatial data. We provide relevant background
+    contextualizing each approach and study their properties in a variety
+    of scenarios, making recommendations for use based on the
+    practitioner's goals.
   \end{enumerate}

### 2021-12-21 inst/manuscript/manuscript.Rmd
minor updates

@@
 When studying the simulations and mercury concentration data, we found that regardless of the strength of spatial dependence in the data, sampling plans that incorporate spatial locations (spatially balanced samples) generally outperform sampling plans that ignore spatial locations (non-spatially balanced samples).
-We also found that model-based analyses tend to outperform design-based analyses, even when the data are skewed (and by consequence, the model-based distributional assumptions violated).
+We also found that model-based approaches tend to outperform design-based approaches, even when the data are skewed (and by consequence, the model-based distributional assumptions violated).
 The performance gap between the analysis approaches is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
-This suggests that the sampling choice (spatially balanced samples versus non-spatially balanced samples) is most important when using a design-based analysis.
+This suggests that the sampling choice (whether to select a sample that is spatially balanced) is most important when performing design-based inference.
   4.
-There are many benefits and drawbacks practitioners must consider when choosing between design-based and model-based approaches for finite population spatial data.
+There are many benefits and drawbacks to the design-based and model-based approaches for finite population spatial data that practitioners must consider when choosing between them.
 We provide relevant background contextualizing each approach and study their properties in a variety of scenarios, making recommendations for use based on the practitioner's goals.
@@
 When data cannot be collected for all units in a population (i.e., population units), data are collected on a subset of the population units.
-This subset is called a sample.
-In the design-based approach, inferences about the underlying population are informed via a probabilistic process assigning some population units to be part of the sample.
+This subset of population units is called a sample.
+In the design-based approach, inferences about the underlying population are informed via a probabilistic process that randomly assigns some population units to be in the sample.
 Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process generating the data.
@@
 @cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g., @sterba2009alternative, @cicchitelli2012model, @chan2020bayesian for a Bayesian approach).
+In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g., @sterba2009alternative and @cicchitelli2012model, and see @chan2020bayesian for a Bayesian approach).
 
@@
 Certainly comparisons between design-based and model-based approaches to spatial data have been studied.
-But no numerical comparison has been made between design-based approaches incorporating spatial information and design-based approaches.
-In this manuscript, we compare design-based approaches incorporating spatial information to model-based approaches for spatial data.
-We focus on finite populations, but these comparisons generalize to infinite populations as well.
-A finite population contains a finite number of population units; an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
+But no numerical comparison has been made between design-based approaches that incorporate spatial information and model-based approaches.
+In this manuscript, we compare design-based approaches that incorporate spatial information to model-based approaches for finite population spatial data.
+A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
+Though we manuscript focuses on finite populations, these comparisons generalize to infinite populations as well.
 An infinite population contains an infinite number of population units; an example is locations within a single lake.
+In this manuscript we assume the number of finite population units is known
 
@@
 In Section \ref{sec:background}, we introduce and provide relevant background for the design-based and model-based approaches to finite population spatial data.
-In Section \ref{sec:mm}, we describe how we compare performance of the approaches wiht a simulation study and an analysis of real data that contains mercury concentration in lakes from the contiguous United States.
-In Section \ref{sec:results}, we present results from the simulation study and the analysis of mercury concentrations.
+In Section \ref{sec:mm}, we describe how we compare performance of the approaches with a simulation study and an analysis of real data that contains mercury concentration in lakes located in the contiguous United States.
+In Section \ref{sec:results}, we present results from the simulation study and the mercury concentration analysis.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 Units from the sampling frame are selected as part of the sample according to a sampling design, which assigns a positive probability of inclusion (inclusion probability) to each unit from the sampling frame.
+These inclusion probabilities are later used to analyze data.
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
+
 When sampling designs incorporate spatial locations into sampling, we call the resulting samples "spatially balanced." One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
@@
 To guard against preferential sampling, model-based approaches often still implement some form of random sampling.
+When model-based approaches implement random sampling, the inclusion probabilities are ignored when analyzing the data (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the data).
 
@@
 
-We previously mentioned that the design-based approach can be used to select spatially balanced samples (samples that incorporate spatial locations of the population units and are "well-spread" is space).
+We previously mentioned that the design-based approach can be used to select spatially balanced samples (samples that incorporate spatial locations of the population units).
 Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-The first spatially balanced sampling algorithm seeing widespread use is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
+The first spatially balanced sampling algorithm to see widespread use was the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
 To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Dirichlet Tessellations).
@@
 On the one dimensional line, each population unit's line length equals its inclusion probability.
-Then, a systematic sample of population units is selected on the line, yielding desired sample.
-@stevens2004spatially provides more technical details.
+Then, a systematic sample of population units is selected on the line and mapped back to two-dimensional space, yielding the desired sample.
+@stevens2004spatially provide more technical details.
 
 After selecting a sample and collecting data, unbiased estimates of population means and totals can be obtained using the Horvitz-Thompson estimator [@horvitz1952generalization].
-If $\tau$ is a population total, the Horvitz-Thompson estimate of $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
+If $\tau$ is a population total, the Horvitz-Thompson estimator for $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
 \begin{align}\label{eq:ht}
@@
 Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
-We summarize some of the basic principles of FBPK next (for more technical details, see @verhoef2008spatial) Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
+We summarize some of the basic principles of FBPK next (for more technical details, see @verhoef2008spatial).
+Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
 .
@@
 
-FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial correlation structure that can be modeled using a covariance function.
-This covariance function is commonly assumed to be non-negative (between zero and one), second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
+FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
+This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
 Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$, though @chiles1999geostatistics, pp.
-80-93 discuss covariance functions that are not second-order stationary, not isotropic, or both.
+80-93 discuss covariance functions that are not second-order stationary, not isotropic, or not either.
 A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics]; one example is the exponential covariance function (for a thorough list of spatial covariance functions, see @cressie1993statistics.
@@
 
-Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison], random forests [@breiman2001random], Bayesian models [@chan2020bayesian], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
+Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison] and random forests [@breiman2001random], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
 Compared to the k-nearest-neighbors and random forest approach, we prefer FBPK because it is model-based and relies on theoretically-based variance estimators leveraging the model's spatial covariance structure, whereas k-nearest-neighbors and random forests use ad-hoc variance estimators [@ver2013comparison].
 Additionally, @ver2013comparison studied compared FBPK, k-nearest-neighbors, and random forests in a variety of spatial data contexts, and FBPK tended to perform best.
-Compared to the Bayesian approach, we prefer FPBK mostly because it is much more computationally efficient.
 
@@
 
-We used a simulation study to investigate performance of four sampling-analysis combinations: IRS with a design-based analysis, called "IRS-Design"; IRS with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; GRTS sampling with a model-based analysis, called "GRTS-Model".
+We used a simulation study to investigate performance of four sampling-analysis combinations: IRS sampling with a design-based analysis, called "IRS-Design"; IRS sampling with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; GRTS sampling with a model-based analysis, called "GRTS-Model".
 These combinations are also provided in Table \ref{tab:designanalysis}.
@@
 The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
-The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decayed to nearly zero at the largest possible distance between two units in the domain.
+The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decayed to nearly zero at the largest possible distance between two population units in the domain.
 For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a random variable whose total variance is 2.
+The response was then exponentiated, yielding a lognormal random variable whose total variance is 2.
 The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed (i.e., not normal).
@@
 All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was two.")
+In each simualtion scenario, the total variance was 2.")
 align(parm_xtab) <- "r|lll"
@@
 
-The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water.
-We will use data from the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality for lakes in the contiguous United States [@USEPA2012NLA].
+The United States Environmental Protection Agency (USEPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) to assess the water quality of various bodies of water in the contiguous United States.
+We will use data from the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality [@USEPA2012NLA].
 Specifically, we will analyze mercury concentration in lakes.
-Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
+Although we can calculate the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
+For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% confidence (design-based) or prediction (model-based) intervals from this sample of 100 lakes, which we compare to the actual mean from all 986 lakes.
 
@@
 
-Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
-<!--- figure 2 -->
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:figeff}
-\end{figure}
+Fig.
+\ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
 The relative rMS(P)E is defined as
@@
 Tables for rMS(P)E in all 36 simulation scenarios are provided in the supporting information.
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:figeff}
+\end{figure}
 
@@
 Justification for this comes from the asymptotic normality of means via the Central Limit Theorem.
-
-Fig. \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
-<!--- figure 3 -->
+Fig.
+\ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
+Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage.
+Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios varied from from 90% to 95% but increased with the sample size.
+At a sample size of 200, all four sampling-analysis combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
+These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+Tables for interval coverage in all 36 simulation scenarios are provided in the supporting information.
 \begin{figure}
@@
   \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
-  \caption{Interval coverage for the four sampling-analysis combinations.
+  \caption{Interval coverage in the simulation study for the four sampling-analysis combinations.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 \end{figure}
-Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage.
-Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios varied from from 90% to 95%.
-Coverage tended to always increase with the sample size.
-At a sample size of 200, all four sampling-analysis combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
-These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
-Tables for interval coverage in all 36 simulation scenarios are provided in the supporting information.
 
@@
 Fig.
-\ref{fig:merc} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Mercury concentration exhibits some spatial patterning, with high mercury concentrations in lakes in the northeast and north central United States.
+\ref{fig:merc} shows a map and histogram of mercury concentration.
+The map shows mercury concentration exhibits some spatial patterning, with high mercury concentrations in lakes in the northeast and north central United States.
+The histogram shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
 Fig.
-\ref{fig:merc} also shows the spatial dependence in mercury concentration via the empirical semivariogram.
+\ref{fig:merc} also shows mercury concentration's empirical semivariogram.
 The empirical semivariogram can be used as a tool to visualize spatial dependence.
-It quantifes the halved squared differences (semivariance) among mercury concentration at different distances apart.
-When a process is spatially correlated, the semivariance tends to be smaller at small distances and larger at large distances.
-Together, the map, histogram, and semivariogram in Fig.
-\ref{fig:merc} suggest that mercury concentration is skewed and exhibits spatial dependence.
-Lastly we note that the realized mean mercury concentration in the 986 lakes is 103.2 ng / g.
+It quantifies the halved squared differences (semivariance) among mercury concentration at different distances apart.
+When a process is spatially correlated (has spatial dependence), the semivariance tends to be smaller at small distances and larger at large distances.
+The empirical semivariogram in Fig.
+\ref{fig:merc} suggests that mercury concentration is exhibits spatial dependence.
+Lastly we note that the realized mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
 \begin{figure}
@@
 \end{subfigure}
-\caption{Mercury concentration visualizations for the population (Hg) for 986 lakes in the contiguous United States.
+\caption{Mercury concentration visualizations for the population (Hg) for 986 lakes in the NLA data.
 A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
@@
 res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration.
-The true mean concentration is 103.2 ng / g.")
+names(res_df_drop) <- c("Approach", "Est/Pred", "SE", "95% LB", "95% UB")
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using the sample of 100 lakes in the NLA data.
+The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.")
 print(xtab_app, include.rownames = FALSE, comment = FALSE)
@@
 
-The design-based and model-based approaches to statistical inference are fundamentally different paradigms by which samples are selected and data are analyzed.
+The design-based and model-based approaches to statistical inference are fundamentally different paradigms that can be used to analyze data.
 The design-based approach incorporates randomness through sampling to estimate population parameters.
@@
 There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial data.
-Some we have discuss, but others we have not and they are worthy of consideration in future research.
-Design-based approaches are often computationally efficient, while model-based estimation can be computationally burdensome, especially for likelihood-based methods such as REML that rely on inverting a covariance matrix.
+Some we have discussed, but others we have not, and they are worthy of consideration in future research.
+Design-based approaches are often computationally efficient, while model-based approaches can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
 The design-based approach also more naturally handles binary data, free from the more complicated logistic regression framework commonly used to analyze binary data in a model-based approach.
@@
 Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
-The benefits and drawbacks of both approaches, alongside our theoretical and analytical comparisons, can motive the process of choosing among them.
-This is especially true from an analysis perspective, as we found that using a spatially balanced sampling algorithm benefits both design-based and model-based analyses.
+The benefits and drawbacks of the design-based and model-based approaches should be considered alongside the particular goals of a study when deciding which approach is most appropriate to implement.
 

### 2021-12-21 inst/manuscript/manuscript.tex
minor updates

@@
     locations (non-spatially balanced samples). We also found that
-    model-based analyses tend to outperform design-based analyses, even
-    when the data are skewed (and by consequence, the model-based
+    model-based approaches tend to outperform design-based approaches,
+    even when the data are skewed (and by consequence, the model-based
     distributional assumptions violated). The performance gap between the
@@
     but large when non-spatially balanced samples are used. This suggests
-    that the sampling choice (spatially balanced samples versus
-    non-spatially balanced samples) is most important when using a
-    design-based analysis.
+    that the sampling choice (whether to select a sample that is spatially
+    balanced) is most important when performing design-based inference.
   \item
-    There are many benefits and drawbacks practitioners must consider when
-    choosing between design-based and model-based approaches for finite
-    population spatial data. We provide relevant background
-    contextualizing each approach and study their properties in a variety
-    of scenarios, making recommendations for use based on the
-    practitioner's goals.
+    There are many benefits and drawbacks to the design-based and
+    model-based approaches for finite population spatial data that
+    practitioners must consider when choosing between them. We provide
+    relevant background contextualizing each approach and study their
+    properties in a variety of scenarios, making recommendations for use
+    based on the practitioner's goals.
   \end{enumerate}
@@
 population units), data are collected on a subset of the population
-units. This subset is called a sample. In the design-based approach,
-inferences about the underlying population are informed via a
-probabilistic process assigning some population units to be part of the
-sample. Alternatively, in the model-based approach, inferences are made
-from specific assumptions about the underlying process generating the
-data. Each paradigm has a deep historical context (Sterba, 2009) and its
-own set of benefits and drawbacks (Hansen et al., 1983).
+units. This subset of population units is called a sample. In the
+design-based approach, inferences about the underlying population are
+informed via a probabilistic process that randomly assigns some
+population units to be in the sample. Alternatively, in the model-based
+approach, inferences are made from specific assumptions about the
+underlying process generating the data. Each paradigm has a deep
+historical context (Sterba, 2009) and its own set of benefits and
+drawbacks (Hansen et al., 1983).
 
@@
 has been substantial research and development into estimators that use
-both design and model-based principles (see e.g., Sterba (2009),
-Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
-Bayesian approach).
+both design and model-based principles (see e.g., Sterba (2009) and
+Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020) for
+a Bayesian approach).
 
@@
 spatial data have been studied. But no numerical comparison has been
-made between design-based approaches incorporating spatial information
-and design-based approaches. In this manuscript, we compare design-based
-approaches incorporating spatial information to model-based approaches
-for spatial data. We focus on finite populations, but these comparisons
-generalize to infinite populations as well. A finite population contains
-a finite number of population units; an example is lakes (treated as a
-whole with the lake centroid representing location) in the contiguous
-United States. An infinite population contains an infinite number of
-population units; an example is locations within a single lake.
+made between design-based approaches that incorporate spatial
+information and model-based approaches. In this manuscript, we compare
+design-based approaches that incorporate spatial information to
+model-based approaches for finite population spatial data. A finite
+population contains a finite number of population units (we assume the
+finite number is known); an example is lakes (treated as a whole with
+the lake centroid representing location) in the contiguous United
+States. Though we manuscript focuses on finite populations, these
+comparisons generalize to infinite populations as well. An infinite
+population contains an infinite number of population units; an example
+is locations within a single lake. In this manuscript we assume the
+number of finite population units is known
 
@@
 data. In Section \ref{sec:mm}, we describe how we compare performance of
-the approaches wiht a simulation study and an analysis of real data that
-contains mercury concentration in lakes from the contiguous United
+the approaches with a simulation study and an analysis of real data that
+contains mercury concentration in lakes located in the contiguous United
 States. In Section \ref{sec:results}, we present results from the
-simulation study and the analysis of mercury concentrations. And in
-Section \ref{sec:discussion}, we end with a discussion and provide
-directions for future research.
+simulation study and the mercury concentration analysis. And in Section
+\ref{sec:discussion}, we end with a discussion and provide directions
+for future research.
 
@@
 sampling design, which assigns a positive probability of inclusion
-(inclusion probability) to each unit from the sampling frame. Some
-examples of commonly used sampling designs include simple random
-sampling, stratified random sampling, and cluster sampling. When
-sampling designs incorporate spatial locations into sampling, we call
-the resulting samples ``spatially balanced.'' One approach to selecting
-spatially balanced samples is the Generalized Random Tessellation
-Stratified (GRTS) algorithm (Stevens and Olsen, 2004), which we discuss
-in more detail in Section \ref{subsec:spb_design}. When sampling designs
-do not incorporate spatial locations into sampling, we call the
-resulting samples ``non-spatially balanced.''
+(inclusion probability) to each unit from the sampling frame. These
+inclusion probabilities are later used to analyze data. Some examples of
+commonly used sampling designs include simple random sampling,
+stratified random sampling, and cluster sampling.
+
+When sampling designs incorporate spatial locations into sampling, we
+call the resulting samples ``spatially balanced.'' One approach to
+selecting spatially balanced samples is the Generalized Random
+Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004),
+which we discuss in more detail in Section \ref{subsec:spb_design}. When
+sampling designs do not incorporate spatial locations into sampling, we
+call the resulting samples ``non-spatially balanced.''
 
@@
 against preferential sampling, model-based approaches often still
-implement some form of random sampling.
+implement some form of random sampling. When model-based approaches
+implement random sampling, the inclusion probabilities are ignored when
+analyzing the data (in contrast to the design-based approach, which
+relies on these inclusion probabilities to analyze the data).
 
@@
 select spatially balanced samples (samples that incorporate spatial
-locations of the population units and are ``well-spread'' is space).
-Spatially balanced samples are useful because parameter estimates from
-these samples tend to vary less than parameter estimates from samples
-that are not spatially balanced (Barabesi and Franceschi, 2011;
-Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
-2013; Stevens and Olsen, 2004; Wang et al., 2013). The first spatially
-balanced sampling algorithm seeing widespread use is the Generalized
-Random Tessellation Stratified (GRTS) algorithm (Stevens and Olsen,
-2004). To quantify the spatial balance of a sample, Stevens and Olsen
-(2004) proposed loss metrics based on Voronoi polygons (Dirichlet
-Tessellations). After the GRTS algorithm was developed, several other
-spatially balanced sampling algorithms emerged, such as the Local
-Pivotal Method (Grafström et al., 2012; Grafström and Matei, 2018),
-Spatially Correlated Poisson Sampling (Grafström, 2012), Balanced
-Acceptance Sampling (Robertson et al., 2013), Within-Sample-Distance
-Sampling (Benedetti and Piersimoni, 2017), and Halton Iterative
-Partitioning Sampling (Robertson et al., 2018). In this manuscript, we
-select spatially balanced samples using the Generalized Random
-Tessellation Stratified (GRTS) algorithm because it has several
+locations of the population units). Spatially balanced samples are
+useful because parameter estimates from these samples tend to vary less
+than parameter estimates from samples that are not spatially balanced
+(Barabesi and Franceschi, 2011; Benedetti et al., 2017; Grafström and
+Lundström, 2013; Robertson et al., 2013; Stevens and Olsen, 2004; Wang
+et al., 2013). The first spatially balanced sampling algorithm to see
+widespread use was the Generalized Random Tessellation Stratified (GRTS)
+algorithm (Stevens and Olsen, 2004). To quantify the spatial balance of
+a sample, Stevens and Olsen (2004) proposed loss metrics based on
+Voronoi polygons (Dirichlet Tessellations). After the GRTS algorithm was
+developed, several other spatially balanced sampling algorithms emerged,
+such as the Local Pivotal Method (Grafström et al., 2012; Grafström and
+Matei, 2018), Spatially Correlated Poisson Sampling (Grafström, 2012),
+Balanced Acceptance Sampling (Robertson et al., 2013),
+Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017), and
+Halton Iterative Partitioning Sampling (Robertson et al., 2018). In this
+manuscript, we select spatially balanced samples using the Generalized
+Random Tessellation Stratified (GRTS) algorithm because it has several
 attractive properties. More specifically, the GRTS algorithm
@@
 unit's line length equals its inclusion probability. Then, a systematic
-sample of population units is selected on the line, yielding desired
-sample. Stevens and Olsen (2004) provides more technical details.
+sample of population units is selected on the line and mapped back to
+two-dimensional space, yielding the desired sample. Stevens and Olsen
+(2004) provide more technical details.
 
@@
 estimator (Horvitz and Thompson, 1952). If \(\tau\) is a population
-total, the Horvitz-Thompson estimate of \(\tau\), denoted by
+total, the Horvitz-Thompson estimator for \(\tau\), denoted by
 \(\hat{\tau}_{ht}\), is is given by \begin{align}\label{eq:ht}
@@
 stochastic process. We summarize some of the basic principles of FBPK
-next (for more technical details, see Ver Hoef (2008)) Let
+next (for more technical details, see Ver Hoef (2008)). Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 FBPK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
-has mean-zero and a spatial correlation structure that can be modeled
+has mean-zero and a spatial dependence structure that can be modeled
 using a covariance function. This covariance function is commonly
-assumed to be non-negative (between zero and one), second-order
-stationary (depending only on the distance between population units),
-isotropic (independent of direction), and decay with distance between
-population units (Cressie, 1993). Henceforth, it is implied that we have
-made these same assumptions regarding \(\bm{\delta}\), though Chiles and
-Delfiner (1999), pp.~80-93 discuss covariance functions that are not
-second-order stationary, not isotropic, or both. A variety of flexible
+assumed to be non-negative, second-order stationary (depending only on
+the distance between population units), isotropic (independent of
+direction), and decay with distance between population units (Cressie,
+1993). Henceforth, it is implied that we have made these same
+assumptions regarding \(\bm{\delta}\), though Chiles and Delfiner
+(1999), pp.~80-93 discuss covariance functions that are not second-order
+stationary, not isotropic, or not either. A variety of flexible
 covariance functions can be used to model \(\bm{\delta}\) (Cressie,
@@
 Other approaches, such as k-nearest-neighbors (Fix and Hodges, 1989; Ver
-Hoef and Temesgen, 2013), random forests (Breiman, 2001), Bayesian
-models (Chan-Golston et al., 2020), among others, could also be used to
-obtain predictions for a mean or total from spatially correlated
-responses of a finite population. Compared to the k-nearest-neighbors
-and random forest approach, we prefer FBPK because it is model-based and
-relies on theoretically-based variance estimators leveraging the model's
-spatial covariance structure, whereas k-nearest-neighbors and random
-forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
-Additionally, Ver Hoef and Temesgen (2013) studied compared FBPK,
-k-nearest-neighbors, and random forests in a variety of spatial data
-contexts, and FBPK tended to perform best. Compared to the Bayesian
-approach, we prefer FPBK mostly because it is much more computationally
-efficient.
+Hoef and Temesgen, 2013) and random forests (Breiman, 2001), among
+others, could also be used to obtain predictions for a mean or total
+from spatially correlated responses of a finite population. Compared to
+the k-nearest-neighbors and random forest approach, we prefer FBPK
+because it is model-based and relies on theoretically-based variance
+estimators leveraging the model's spatial covariance structure, whereas
+k-nearest-neighbors and random forests use ad-hoc variance estimators
+(Ver Hoef and Temesgen, 2013). Additionally, Ver Hoef and Temesgen
+(2013) studied compared FBPK, k-nearest-neighbors, and random forests in
+a variety of spatial data contexts, and FBPK tended to perform best.
 
@@
 We used a simulation study to investigate performance of four
-sampling-analysis combinations: IRS with a design-based analysis, called
-``IRS-Design''; IRS with a model-based analysis, called ``IRS-Model'';
-GRTS sampling with a design-based analysis, called ``GRTS-Design''; GRTS
-sampling with a model-based analysis, called ``GRTS-Model''. These
-combinations are also provided in Table \ref{tab:designanalysis}.
+sampling-analysis combinations: IRS sampling with a design-based
+analysis, called ``IRS-Design''; IRS sampling with a model-based
+analysis, called ``IRS-Model''; GRTS sampling with a design-based
+analysis, called ``GRTS-Design''; GRTS sampling with a model-based
+analysis, called ``GRTS-Model''. These combinations are also provided in
+Table \ref{tab:designanalysis}.
 
@@
 random error decayed to nearly zero at the largest possible distance
-between two units in the domain. For the lognormal response type, the
-response was first simulated using the same approach as for the normal
-response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a random variable whose
-total variance is 2. The lognormal responses were used to evaluate
-performance of the sampling-analysis approaches for data that were
-skewed (i.e., not normal).
+between two population units in the domain. For the lognormal response
+type, the response was first simulated using the same approach as for
+the normal response type, except that the total variance was 0.6931
+instead of 2. The response was then exponentiated, yielding a lognormal
+random variable whose total variance is 2. The lognormal responses were
+used to evaluate performance of the sampling-analysis approaches for
+data that were skewed (i.e., not normal).
 
@@
 All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was two.}
+In each simualtion scenario, the total variance was 2.}
 \end{table}
@@
 
-The Environmental Protection Agency (EPA), states, and tribes
-periodically conduct National Aquatic Research Surveys (NARS) in the
-United States to assess the water quality of various bodies of water. We
-will use data from the 2012 National Lakes Assessment (NLA), which
-measures various aspects of lake health and water quality for lakes in
-the contiguous United States (USEPA, 2012). Specifically, we will
-analyze mercury concentration in lakes. Although we know the true mean
-mercury concentration values for the 986 lakes from the 2012 NLA, we
-will explore whether or not we obtain an adequately precise estimate for
-the realized mean mercury concentration if we sample only 100 of the 986
-lakes. For each of the four familiar sampling-analysis combinations
-(IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
+The United States Environmental Protection Agency (USEPA), states, and
+tribes periodically conduct National Aquatic Research Surveys (NARS) to
+assess the water quality of various bodies of water in the contiguous
+United States. We will use data from the 2012 National Lakes Assessment
+(NLA), which measures various aspects of lake health and water quality
+(USEPA, 2012). Specifically, we will analyze mercury concentration in
+lakes. Although we can calculate the true mean mercury concentration
+values for the 986 lakes from the 2012 NLA, we will explore whether or
+not we obtain an adequately precise estimate for the realized mean
+mercury concentration if we sample only 100 of the 986 lakes. For each
+of the four familiar sampling-analysis combinations (IRS-Design,
+IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or
+predict (model-based) the mean mercury concentration and construct 95\%
+confidence (design-based) or prediction (model-based) intervals from
+this sample of 100 lakes, which we compare to the actual mean from all
+986 lakes.
 
@@
 from Table \ref{tab:designanalysis} using the random location layout
-with ``IRS-Design'' as the baseline.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:figeff}
-\end{figure}
-
-The relative rMS(P)E is defined as \begin{equation*}
+with ``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
+\begin{equation*}
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:figeff}
+\end{figure}
+
 We also studied 95\% interval coverage among the sampling-analysis
@@
 Justification for this comes from the asymptotic normality of means via
-the Central Limit Theorem.
-
-Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
-four sampling-analysis combinations in the random location layout.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
-  \caption{Interval coverage for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line in each plot represents 95\% coverage.}
-  \label{fig:figconf}
-\end{figure}
-
-Within each scenario, the sampling-analysis combinations tend to have
-fairly similar interval coverage. Coverage in the normal response
-scenarios was usually near 95\%, while coverage in the lognormal
-response scenarios varied from from 90\% to 95\%. Coverage tended to
-always increase with the sample size. At a sample size of 200, all four
+the Central Limit Theorem. Fig. \ref{fig:figconf} shows the 95\%
+interval coverage for each of the four sampling-analysis combinations in
+the random location layout. Within each scenario, the sampling-analysis
+combinations tend to have fairly similar interval coverage. Coverage in
+the normal response scenarios was usually near 95\%, while coverage in
+the lognormal response scenarios varied from from 90\% to 95\% but
+increased with the sample size. At a sample size of 200, all four
 sampling-analysis combinations had approximately 95\% interval coverage
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
+  \caption{Interval coverage in the simulation study for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line in each plot represents 95\% coverage.}
+  \label{fig:figconf}
+\end{figure}
+
 \hypertarget{sec:r_app}{%
@@
 
-Fig. \ref{fig:merc} shows that mercury concentration is right-skewed,
-with most lakes having a low value of mercury concentration but a few
-having a much higher concentration. Mercury concentration exhibits some
-spatial patterning, with high mercury concentrations in lakes in the
-northeast and north central United States. Fig. \ref{fig:merc} also
-shows the spatial dependence in mercury concentration via the empirical
+Fig. \ref{fig:merc} shows a map and histogram of mercury concentration.
+The map shows mercury concentration exhibits some spatial patterning,
+with high mercury concentrations in lakes in the northeast and north
+central United States. The histogram shows that mercury concentration is
+right-skewed, with most lakes having a low value of mercury
+concentration but a few having a much higher concentration. Fig.
+\ref{fig:merc} also shows mercury concentration's empirical
 semivariogram. The empirical semivariogram can be used as a tool to
-visualize spatial dependence. It quantifes the halved squared
+visualize spatial dependence. It quantifies the halved squared
 differences (semivariance) among mercury concentration at different
-distances apart. When a process is spatially correlated, the
-semivariance tends to be smaller at small distances and larger at large
-distances. Together, the map, histogram, and semivariogram in Fig.
-\ref{fig:merc} suggest that mercury concentration is skewed and exhibits
-spatial dependence. Lastly we note that the realized mean mercury
-concentration in the 986 lakes is 103.2 ng / g.
+distances apart. When a process is spatially correlated (has spatial
+dependence), the semivariance tends to be smaller at small distances and
+larger at large distances. The empirical semivariogram in Fig.
+\ref{fig:merc} suggests that mercury concentration is exhibits spatial
+dependence. Lastly we note that the realized mean mercury concentration
+in the 986 NLA lakes is 103.2 ng / g.
 
@@
   \hline
-Approach & Estimate & SE & 95\% LB & 95\% UB \\ 
+Approach & Est/Pred & SE & 95\% LB & 95\% UB \\ 
   \hline
@@
 \end{tabular}
-\caption{\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.} 
+\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using the sample of 100 lakes in the NLA data.
+The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.}
 \end{table}
@@
 The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms by which samples are selected and data
-are analyzed. The design-based approach incorporates randomness through
-sampling to estimate population parameters. The model-based approach
-incorporates randomness through distributional assumptions to predict
-realized values of a random process. Though these approaches have often
-been compared in the literature both from theoretical and analytical
-perspectives, our contribution lies in studying them in a spatial
-context while implementing spatially balanced sampling. Aside from the
-theoretical differences described, a few analytical findings from the
-simulation study are particularly notable. First, the sampling decision
-(GRTS vs IRS) is most important when using a design-based analysis.
-Though GRTS-Model still outperformed IRS-Model, the model-based analysis
+fundamentally different paradigms that can be used to analyze data. The
+design-based approach incorporates randomness through sampling to
+estimate population parameters. The model-based approach incorporates
+randomness through distributional assumptions to predict realized values
+of a random process. Though these approaches have often been compared in
+the literature both from theoretical and analytical perspectives, our
+contribution lies in studying them in a spatial context while
+implementing spatially balanced sampling. Aside from the theoretical
+differences described, a few analytical findings from the simulation
+study are particularly notable. First, the sampling decision (GRTS vs
+IRS) is most important when using a design-based analysis. Though
+GRTS-Model still outperformed IRS-Model, the model-based analysis
 mitigated much of the inefficiency of the IRS sample. Second,
@@
 model-based approaches for finite population spatial data. Some we have
-discuss, but others we have not and they are worthy of consideration in
-future research. Design-based approaches are often computationally
-efficient, while model-based estimation can be computationally
-burdensome, especially for likelihood-based methods such as REML that
-rely on inverting a covariance matrix. The design-based approach also
-more naturally handles binary data, free from the more complicated
+discussed, but others we have not, and they are worthy of consideration
+in future research. Design-based approaches are often computationally
+efficient, while model-based approaches can be computationally
+burdensome, especially for likelihood-based estimation methods like REML
+that rely on inverting a covariance matrix. The design-based approach
+also more naturally handles binary data, free from the more complicated
 logistic regression framework commonly used to analyze binary data in a
@@
 site-by-site predictions at unobserved locations and use them to
-construct informative visualizations. The benefits and drawbacks of both
-approaches, alongside our theoretical and analytical comparisons, can
-motive the process of choosing among them. This is especially true from
-an analysis perspective, as we found that using a spatially balanced
-sampling algorithm benefits both design-based and model-based analyses.
+construct informative visualizations. The benefits and drawbacks of the
+design-based and model-based approaches should be considered alongside
+the particular goals of a study when deciding which approach is most
+appropriate to implement.
 

### 2021-12-21 inst/manuscript/supporting.Rmd
style guide updates
minor update to supporting information

--- 2021-12-18 inst/manuscript/supplementary.Rmd
@@
 ---
-title: A comparison of design-based and model-based approaches for finite population spatial data -- Supplementary Material.
+title: A comparison of design-based and model-based approaches for finite population spatial data -- Supporting Information.
 author:
@@
 
-# Numerical Study Tables {#sec:numstudy}
+# Simulation Study Tables {#sec:simtabs}
 
@@
 
-Recall the four design-analysis combinations and the parameter configurations, summarized in Table \ref{tab:designanalysis} and \ref{tab:parmtab}, respectively.
+Recall the four sampling-analysis combinations from the simulation study, summarized in Table \ref{tab:designanalysis}.
+Also recall the 36 parameter configurations from the simulation study, summarized in Table \ref{tab:parmtab}.
 
@@
 #       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.")
+parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the total variance was 2.")
 align(parm_xtab) <- "r|lll"
@@
 
-```{r, results = "asis", eval = FALSE}
-sim_one <- read_csv(here("inst", "output", "sim_one", "summ_output.csv"))
-colnames(sim_one) <- c("Approach", "Bias", "RMSE", "MedAE", "Coverage", "PClose", "MedIL")
-sim_one_table <- xtable(sim_one, digits = 4, caption = "Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)", type = "latex", latex.environments = "center", label = "tab:sim_one")
-print(sim_one_table, include.rownames = FALSE, comment = FALSE)
-```
-
-Next we present a summary table for each of the 36 simulation scenarios.
+Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
 ```{r}
@@
   digits = 4,
-  caption = "Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\\% interval coverage (Coverage) for a simulation scenario.",
+  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
    type = "latex",
-  latex.environments = "center"
+  latex.environments = "center",
+  align = c("llllrrrrr")
 )

### 2021-12-21 inst/manuscript/supporting.tex
minor update to supporting information

@@
 
-Recall the four design-analysis combinations and the parameter
-configurations, summarized in Table \ref{tab:designanalysis} and
-\ref{tab:parmtab}, respectively.
+Recall the four sampling-analysis combinations from the simulation
+study, summarized in Table \ref{tab:designanalysis}. Also recall the 36
+parameter configurations from the simulation study, summarized in Table
+\ref{tab:parmtab}.
 
@@
 \end{tabular}
-\caption{\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study.
-The rows give the two types of sampling designs while the columns give the two types of analyses.}
+\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
 \end{table}
@@
 \end{tabular}
-\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.} 
+\caption{\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the total variance was 2.}
 \end{table}
 
-Next we present a summary table for each of the 36 simulation scenarios.
+Next we present tables summarizing mean estimation bias (design-based)
+or mean prediction bias (model-based), root-mean-squared error
+(design-based) or root-mean-squared-prediction error (model-based), and
+95\% confidence interval coverage (design-based) or 95\% prediction
+interval coverage (model-based) for all 36 simulation scenarios.
 
@@
 \scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\begin{tabular}{lllrrrrr}
   \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 1.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 2.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 3.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 4.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 5.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 6.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 7.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 8.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 9.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 10.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 11.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 12.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 13.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 14.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 15.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 16.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 17.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 18.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 19.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 20.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 21.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 22.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 23.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 24.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 25.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 26.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 27.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 28.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 29.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 30.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 31.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 32.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 33.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 34.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 35.}
 \end{table}
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 36.}
 \end{table}

### 2021-12-21 inst/manuscript/manuscript.Rmd
minor updates

@@
 We also found that model-based approaches tend to outperform design-based approaches, even when the data are skewed (and by consequence, the model-based distributional assumptions violated).
-The performance gap between the analysis approaches is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
+The performance gap between these approaches is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
 This suggests that the sampling choice (whether to select a sample that is spatially balanced) is most important when performing design-based inference.

### 2021-12-22 inst/manuscript/manuscript.Rmd
more minor updates

@@
 In the simulations, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
-In the simualtions and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
   3.
@@
 
-Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based inference; Spatially balanced sampling; Spatial covariance;
+Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based inference; Spatially balanced sampling; Spatial covariance
 
@@
 Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as data that incorporates the specific locations of the population units into either the design or estimation process.
+We define spatial data as data that incorporates the specific locations of the population units into either the sampling or estimation process.
 @de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
@@
 A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
-Though we manuscript focuses on finite populations, these comparisons generalize to infinite populations as well.
+Though we focus on finite populations, these comparisons generalize to infinite populations as well.
 An infinite population contains an infinite number of population units; an example is locations within a single lake.
-In this manuscript we assume the number of finite population units is known
 
@@
 The design-based approach assumes the population is fixed.
-Randomness is incorporated via the selection of units in a sampling frame.
-A sampling frame is the set of all units available to be sampled.
-Units from the sampling frame are selected as part of the sample according to a sampling design, which assigns a positive probability of inclusion (inclusion probability) to each unit from the sampling frame.
+Randomness is incorporated via the selection of population units according to a sampling design.
+A sampling design assigns a positive probability of inclusion (inclusion probability) in the sample to each population unit.
 These inclusion probabilities are later used to analyze data.
@@
   \caption{A visual comparison of the design-based and model-based approaches.
-In the top row, there is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
+In the top row, the design-based approach is highlighted.
+There is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
 The response values at each site are fixed, but we obtain different estimates for the mean response in each random sample.
-In the bottom row, there are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
-The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations}
+In the bottom row, the model-based approach is highlighted.
+There are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
+The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations.}
   \label{fig:fig1}
@@
 To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Dirichlet Tessellations).
-After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, such as the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it has several attractive properties.
-More specifically, the GRTS algorithm accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
+In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it has several attractive properties: the GRTS algorithm accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
 The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
@@
 \end{align}
-where $Z_i$ is the value of the $i$th population unit in the sample and $\pi_i$ is the inclusion probability of the $i$th population unit in the sample.
+where $Z_i$ is the value of the $i$th population unit in the sample, $\pi_i$ is the inclusion probability of the $i$th population unit in the sample, and $n$ is the sample size.
 An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by $N$, the number of population units.
 
-It is also important to quantify uncertainty $\hat{\tau}_{ht}$.
+It is also important to quantify the uncertainty in $\hat{\tau}_{ht}$.
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
@@
 Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
-We summarize some of the basic principles of FBPK next (for more technical details, see @verhoef2008spatial).
+We summarize some of the basic principles of FBPK next -- for technical details, see @verhoef2008spatial.
 Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
@@
 Suppose we want to use a sample to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights (e.g, the population mean is represented by a weights vector whose elements all equal one).
-Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with subscript \emph{u}, let
+Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u}, let
 
@@
 80-93 discuss covariance functions that are not second-order stationary, not isotropic, or not either.
-A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics]; one example is the exponential covariance function (for a thorough list of spatial covariance functions, see @cressie1993statistics.
+A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics]; one example is the exponential covariance function (@cressie1993statistics provides a thorough list of spatial covariance functions).
 The $i,j$th element of the exponential covariance matrix, $\cov(\bm{\delta})$, is
@@
 
-Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison] and random forests [@breiman2001random], among others, could also be used to obtain predictions for a mean or total from spatially correlated responses of a finite population.
+Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison] and random forest [@breiman2001random], among others, could also be used to obtain predictions for a mean or total from finite population spatial data.
 Compared to the k-nearest-neighbors and random forest approach, we prefer FBPK because it is model-based and relies on theoretically-based variance estimators leveraging the model's spatial covariance structure, whereas k-nearest-neighbors and random forests use ad-hoc variance estimators [@ver2013comparison].
-Additionally, @ver2013comparison studied compared FBPK, k-nearest-neighbors, and random forests in a variety of spatial data contexts, and FBPK tended to perform best.
+Additionally, @ver2013comparison studied compared FBPK, k-nearest-neighbors, and random forest in a variety of spatial data contexts, and FBPK tended to perform best.
 
@@
 
-We used a simulation study to investigate performance of four sampling-analysis combinations: IRS sampling with a design-based analysis, called "IRS-Design"; IRS sampling with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; GRTS sampling with a model-based analysis, called "GRTS-Model".
+We used a simulation study to investigate performance of four sampling-analysis combinations: Independent Random Sampling (IRS) with a design-based analysis, called "IRS-Design" (IRS samples are not spatially balanced because population units are selected independent of their spatial locations); IRS with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; and GRTS sampling with a model-based analysis, called "GRTS-Model".
 These combinations are also provided in Table \ref{tab:designanalysis}.
@@
 Performance for the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
-The 36 scenarios resulted from the crossing of three sample sizes, two location layouts, two response types, and three proportions of dependent random error.
+The 36 scenarios resulted from the crossing of three sample sizes, two location layouts (of the population units), two response types, and three proportions of dependent random error.
 The three sample sizes ($n$) were $n = 50, n = 100,$ and $n = 200$.
 Samples were always selected from a population size ($N$) of $N = 900$.
-The two location layouts (of the population units) were random and gridded.
+The two location layouts were random and gridded.
 Locations in the random layout were randomly generated inside the unit square ($[0, 1] \times [0, 1]$).
@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct confidence (design-based) or prediction (model-based) intervals.
+In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
 Then we recorded the bias, squared error, and interval coverage for all sampling-analysis combinations.
 After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% interval.
-The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package `spsurvey` [@dumelle2021spsurvey].
+The GRTS algorithm and the local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
 FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal] and covariance parameters were estimated using Restricted Maximum Likelihood  [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
@@
 Although we can calculate the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% confidence (design-based) or prediction (model-based) intervals from this sample of 100 lakes, which we compare to the actual mean from all 986 lakes.
+For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% confidence (design-based) or 95% prediction (model-based) intervals from this sample of 100 lakes, which we compare to the true mean mercury concentration from all 986 lakes.
 
@@
 
-The average bias was nearly zero for all four combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+The average bias was nearly zero for all four sampling-analysis combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
 Tables for average bias in all 36 simulation scenarios are provided in the supporting information.
 
-Fig.
-\ref{fig:figeff} shows the relative rMS(P)E of the four approaches from Table \ref{tab:designanalysis} using the random location layout with "IRS-Design" as the baseline.
-The relative rMS(P)E is defined as
+Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative rMS(P)E is defined as
 \begin{equation*}
@@
 \end{equation*}
-When there is no spatial correlation (Fig.
+When there is no spatial covariance (Fig.
 \ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
-So using the GRTS sampling plan or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial correlation.
-When there is spatial correlation (Fig.
+So using the GRTS sampling plan or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial covariance.
+When there is spatial covariance (Fig.
 \ref{fig:figeff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
-As the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
-Finally we note that when there is spatial correlation, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
+As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
+Finally we note that when there is spatial covariance, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
 These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
@@
 The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line in each plot represents 95\% coverage.}
+The solid, black line represents 95\% coverage.}
   \label{fig:figconf}
@@
 Fig.
-\ref{fig:merc} shows a map and histogram of mercury concentration.
-The map shows mercury concentration exhibits some spatial patterning, with high mercury concentrations in lakes in the northeast and north central United States.
+\ref{fig:merc} shows a map and histogram of mercury concentration in all 986 NLA lakes.
+The map shows mercury concentration exhibits some spatial patterning, with high mercury concentrations in the northeast and north central United States.
 The histogram shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
@@
 It quantifies the halved squared differences (semivariance) among mercury concentration at different distances apart.
-When a process is spatially correlated (has spatial dependence), the semivariance tends to be smaller at small distances and larger at large distances.
+When a process has spatial covariance (exhibits spatial dependence), the semivariance tends to be smaller at small distances and larger at large distances.
 The empirical semivariogram in Fig.
@@
 
-We selected a single IRS sample and a single GRTS sample and estimated (design-based) or predicted (model-based) the mean mercury concentration and its standard error using using design-based and model-based approaches.
+We selected a single IRS sample and a single GRTS sample and estimated (design-based) or predicted (model-based) the mean mercury concentration and constructed 95% confidence (design-based) and 95% (model-based) prediction intervals.
 For the model-based analyses, the exponential covariance was used.
 Table \ref{tab:appliedtab} shows the results from these analyses.
-For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% confidence (design-based) or prediction (model-based) intervals.
+For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% confidence (design-based) or 95% prediction (model-based) intervals.
 Though we should not generalize these results to other samples from these data, we do note a couple of patterns.
@@
 Both analyses using GRTS sampling have lower standard errors than both analyses using IRS sampling.
-We expect that these patterns are consistent with other samples from these data because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.
+We expect that these patterns are consistent with other samples from these data because mercury concentration exhibits spatial dependence, so a spatially balanced sample should usually yield a lower standard error.
 
@@
 names(res_df_drop) <- c("Approach", "Est/Pred", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using the sample of 100 lakes in the NLA data.
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.
 The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.")
@@
 
-The design-based and model-based approaches to statistical inference are fundamentally different paradigms that can be used to analyze data.
+The design-based and model-based approaches to statistical inference are fundamentally different paradigms.
 The design-based approach incorporates randomness through sampling to estimate population parameters.
-The model-based approach incorporates randomness through distributional assumptions to predict realized values of a random process.
-Though these approaches have often been compared in the literature both from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling.
+The model-based approach incorporates randomness through distributional assumptions to predict realized values of a stochastic process.
+Though these approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling and the local neighborhood variance estimator (in the design-based approach).
 Aside from the theoretical differences described, a few analytical findings from the simulation study are particularly notable.
-First, the sampling decision (GRTS vs IRS) is most important when using a design-based analysis.
+First, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
 Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated much of the inefficiency of the IRS sample.
-Second, independent of the analysis approach, we found no reason to prefer IRS over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial correlation and noticeably better than their IRS counterparts when there was spatial correlation.
-Third, as the strength of spatial correlation increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations also increases.
+Second, independent of the analysis approach, we found no reason to prefer IRS over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial covariance and noticeably better than their IRS counterparts when there was spatial covariance.
+Third, as the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations also increases.
 Fourth and finally, when the response was normal, interval coverage for all sampling-analysis combinations was very close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling and analysis was between 90% and 95% and closest to 95% when $n = 200$.
@@
 The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and response variable.
-The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the process of study.
+The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the stochastic process of study.
 Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
@@
 Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
-The benefits and drawbacks of the design-based and model-based approaches should be considered alongside the particular goals of a study when deciding which approach is most appropriate to implement.
+In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.
 
@@
 
+# Author Contribution Statement {.unnumbered}
+
+All authors conceived the ideas; All authors designed methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
+
 # Data and Code Availability {.unnumbered}
@@
 
+If the manuscript is accepted, this repository will be archived in Zenodo.
+
 # Supporting Information {.unnumbered}
 
-In the supporting information, we provide tables presenting summary statistics for all 36 simulation scenarios.
-
-# Author Contributions {.unnumbered}
-
-All authors conceived the ideas; All authors designed methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
+In the supporting information, we provide tables of summary statistics for all 36 simulation scenarios.
 

### 2021-12-22 inst/manuscript/manuscript.tex
more minor updates

@@
     variety of sample sizes, location layouts, dependence structures, and
-    response types are considered. In the simualtions and real data
+    response types are considered. In the simualations and real data
     analysis, the population mean is the parameter of interest and
@@
     even when the data are skewed (and by consequence, the model-based
-    distributional assumptions violated). The performance gap between the
-    analysis approaches is small when spatially balanced samples are used
-    but large when non-spatially balanced samples are used. This suggests
-    that the sampling choice (whether to select a sample that is spatially
+    distributional assumptions violated). The performance gap between
+    these approaches is small when spatially balanced samples are used but
+    large when non-spatially balanced samples are used. This suggests that
+    the sampling choice (whether to select a sample that is spatially
     balanced) is most important when performing design-based inference.
@@
 Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based
-inference; Spatially balanced sampling; Spatial covariance;
+inference; Spatially balanced sampling; Spatial covariance
 
@@
 spatial data. We define spatial data as data that incorporates the
-specific locations of the population units into either the design or
+specific locations of the population units into either the sampling or
 estimation process. De Gruijter and Ter Braak (1990) give an early
@@
 the lake centroid representing location) in the contiguous United
-States. Though we manuscript focuses on finite populations, these
-comparisons generalize to infinite populations as well. An infinite
-population contains an infinite number of population units; an example
-is locations within a single lake. In this manuscript we assume the
-number of finite population units is known
+States. Though we focus on finite populations, these comparisons
+generalize to infinite populations as well. An infinite population
+contains an infinite number of population units; an example is locations
+within a single lake.
 
@@
 The design-based approach assumes the population is fixed. Randomness is
-incorporated via the selection of units in a sampling frame. A sampling
-frame is the set of all units available to be sampled. Units from the
-sampling frame are selected as part of the sample according to a
-sampling design, which assigns a positive probability of inclusion
-(inclusion probability) to each unit from the sampling frame. These
-inclusion probabilities are later used to analyze data. Some examples of
-commonly used sampling designs include simple random sampling,
-stratified random sampling, and cluster sampling.
+incorporated via the selection of population units according to a
+sampling design. A sampling design assigns a positive probability of
+inclusion (inclusion probability) in the sample to each population unit.
+These inclusion probabilities are later used to analyze data. Some
+examples of commonly used sampling designs include simple random
+sampling, stratified random sampling, and cluster sampling.
 
@@
 developed, several other spatially balanced sampling algorithms emerged,
-such as the Local Pivotal Method (Grafström et al., 2012; Grafström and
-Matei, 2018), Spatially Correlated Poisson Sampling (Grafström, 2012),
-Balanced Acceptance Sampling (Robertson et al., 2013),
+including the Local Pivotal Method (Grafström et al., 2012; Grafström
+and Matei, 2018), Spatially Correlated Poisson Sampling (Grafström,
+2012), Balanced Acceptance Sampling (Robertson et al., 2013),
 Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017), and
@@
 Random Tessellation Stratified (GRTS) algorithm because it has several
-attractive properties. More specifically, the GRTS algorithm
-accommodates finite and infinite sampling frames, equal, unequal, and
-proportional (to size) inclusion probabilities, legacy (historical)
-sampling (Foster et al., 2017), a minimum distance between units in a
-sample, and replacement units (replacement units are population units
-that can be sampled when a population unit originally selected can no
-longer be sampled). The GRTS algorithm selects samples by utilizing a
-particular mapping between two-dimensional and one-dimensional space
-that preserves proximity relationships. Via this mapping, units in
-two-dimensional space are partitioned using a hierarchical address. This
-hierarchical address is used to map population units to a
-one-dimensional line. On the one dimensional line, each population
-unit's line length equals its inclusion probability. Then, a systematic
-sample of population units is selected on the line and mapped back to
-two-dimensional space, yielding the desired sample. Stevens and Olsen
-(2004) provide more technical details.
+attractive properties: the GRTS algorithm accommodates finite and
+infinite sampling frames, equal, unequal, and proportional (to size)
+inclusion probabilities, legacy (historical) sampling (Foster et al.,
+2017), a minimum distance between units in a sample, and replacement
+units (replacement units are population units that can be sampled when a
+population unit originally selected can no longer be sampled). The GRTS
+algorithm selects samples by utilizing a particular mapping between
+two-dimensional and one-dimensional space that preserves proximity
+relationships. Via this mapping, units in two-dimensional space are
+partitioned using a hierarchical address. This hierarchical address is
+used to map population units to a one-dimensional line. On the one
+dimensional line, each population unit's line length equals its
+inclusion probability. Then, a systematic sample of population units is
+selected on the line and mapped back to two-dimensional space, yielding
+the desired sample. Stevens and Olsen (2004) provide more technical
+details.
 
@@
 \end{align} where \(Z_i\) is the value of the \(i\)th population unit in
-the sample and \(\pi_i\) is the inclusion probability of the \(i\)th
-population unit in the sample. An estimate of the population mean is
-obtained by dividing \(\hat{\tau}_{ht}\) by \(N\), the number of
-population units.
-
-It is also important to quantify uncertainty \(\hat{\tau}_{ht}\).
+the sample, \(\pi_i\) is the inclusion probability of the \(i\)th
+population unit in the sample, and \(n\) is the sample size. An estimate
+of the population mean is obtained by dividing \(\hat{\tau}_{ht}\) by
+\(N\), the number of population units.
+
+It is also important to quantify the uncertainty in \(\hat{\tau}_{ht}\).
 Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
@@
 stochastic process. We summarize some of the basic principles of FBPK
-next (for more technical details, see Ver Hoef (2008)). Let
+next -- for technical details, see Ver Hoef (2008). Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 units with a subscript \emph{s} and quantities that are part of the
-unsampled population units with subscript \emph{u}, let
+unsampled population units with a subscript \emph{u}, let
 
@@
 covariance functions can be used to model \(\bm{\delta}\) (Cressie,
-1993); one example is the exponential covariance function (for a
-thorough list of spatial covariance functions, see Cressie (1993). The
+1993); one example is the exponential covariance function (Cressie
+(1993) provides a thorough list of spatial covariance functions). The
 \(i,j\)th element of the exponential covariance matrix,
@@
 Other approaches, such as k-nearest-neighbors (Fix and Hodges, 1989; Ver
-Hoef and Temesgen, 2013) and random forests (Breiman, 2001), among
+Hoef and Temesgen, 2013) and random forest (Breiman, 2001), among
 others, could also be used to obtain predictions for a mean or total
-from spatially correlated responses of a finite population. Compared to
-the k-nearest-neighbors and random forest approach, we prefer FBPK
-because it is model-based and relies on theoretically-based variance
-estimators leveraging the model's spatial covariance structure, whereas
-k-nearest-neighbors and random forests use ad-hoc variance estimators
-(Ver Hoef and Temesgen, 2013). Additionally, Ver Hoef and Temesgen
-(2013) studied compared FBPK, k-nearest-neighbors, and random forests in
-a variety of spatial data contexts, and FBPK tended to perform best.
+from finite population spatial data. Compared to the k-nearest-neighbors
+and random forest approach, we prefer FBPK because it is model-based and
+relies on theoretically-based variance estimators leveraging the model's
+spatial covariance structure, whereas k-nearest-neighbors and random
+forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
+Additionally, Ver Hoef and Temesgen (2013) studied compared FBPK,
+k-nearest-neighbors, and random forest in a variety of spatial data
+contexts, and FBPK tended to perform best.
 
@@
 We used a simulation study to investigate performance of four
-sampling-analysis combinations: IRS sampling with a design-based
-analysis, called ``IRS-Design''; IRS sampling with a model-based
-analysis, called ``IRS-Model''; GRTS sampling with a design-based
-analysis, called ``GRTS-Design''; GRTS sampling with a model-based
-analysis, called ``GRTS-Model''. These combinations are also provided in
-Table \ref{tab:designanalysis}.
+sampling-analysis combinations: Independent Random Sampling (IRS) with a
+design-based analysis, called ``IRS-Design'' (IRS samples are not
+spatially balanced because population units are selected independent of
+their spatial locations); IRS with a model-based analysis, called
+``IRS-Model''; GRTS sampling with a design-based analysis, called
+``GRTS-Design''; and GRTS sampling with a model-based analysis, called
+``GRTS-Model''. These combinations are also provided in Table
+\ref{tab:designanalysis}.
 
@@
 36 different simulation scenarios. The 36 scenarios resulted from the
-crossing of three sample sizes, two location layouts, two response
-types, and three proportions of dependent random error. The three sample
-sizes (\(n\)) were \(n = 50, n = 100,\) and \(n = 200\). Samples were
-always selected from a population size (\(N\)) of \(N = 900\). The two
-location layouts (of the population units) were random and gridded.
+crossing of three sample sizes, two location layouts (of the population
+units), two response types, and three proportions of dependent random
+error. The three sample sizes (\(n\)) were \(n = 50, n = 100,\) and
+\(n = 200\). Samples were always selected from a population size (\(N\))
+of \(N = 900\). The two location layouts were random and gridded.
 Locations in the random layout were randomly generated inside the unit
@@
 then design-based and model-based analyses were used to estimate
-(design-based) or predict (model-based) the mean and construct
-confidence (design-based) or prediction (model-based) intervals. Then we
-recorded the bias, squared error, and interval coverage for all
+(design-based) or predict (model-based) the mean and construct 95\%
+confidence (design-based) or 95\% prediction (model-based) intervals.
+Then we recorded the bias, squared error, and interval coverage for all
 sampling-analysis combinations. After all 2000 trials, we summarized the
@@
 The GRTS algorithm and the local neighborhood variance estimator are
-available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
+available in the \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle
 et al., 2021). FPBK is available in the \texttt{sptotal}
@@
 predict (model-based) the mean mercury concentration and construct 95\%
-confidence (design-based) or prediction (model-based) intervals from
-this sample of 100 lakes, which we compare to the actual mean from all
-986 lakes.
+confidence (design-based) or 95\% prediction (model-based) intervals
+from this sample of 100 lakes, which we compare to the true mean mercury
+concentration from all 986 lakes.
 
@@
 
-The average bias was nearly zero for all four combinations in all 36
-scenarios, so we omit a more detailed summary of those results here.
-Tables for average bias in all 36 simulation scenarios are provided in
-the supporting information.
-
-Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four approaches
-from Table \ref{tab:designanalysis} using the random location layout
-with ``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
+The average bias was nearly zero for all four sampling-analysis
+combinations in all 36 scenarios, so we omit a more detailed summary of
+those results here. Tables for average bias in all 36 simulation
+scenarios are provided in the supporting information.
+
+Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling
+analysis combinations using the random location layout with
+``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
 \begin{equation*}
 \frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
-\end{equation*} When there is no spatial correlation (Fig.
+\end{equation*} When there is no spatial covariance (Fig.
 \ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
@@
 loss in efficiency compared to IRS-Design when there is no spatial
-correlation. When there is spatial correlation (Fig. \ref{fig:figeff},
+covariance. When there is spatial covariance (Fig. \ref{fig:figeff},
 ``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
@@
 the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and
-IRS-Model is relatively small. As the strength of spatial correlation
+IRS-Model is relatively small. As the strength of spatial covariance
 increases, the gap in rMS(P)E between IRS-Design and the other
 sampling-analysis combinations widens. Finally we note that when there
-is spatial correlation, IRS-Model outperforms IRS-Design by a large
+is spatial covariance, IRS-Model outperforms IRS-Design by a large
 margin, suggesting that the poor design properties of IRS are largely
@@
 
-Fig. \ref{fig:merc} shows a map and histogram of mercury concentration.
-The map shows mercury concentration exhibits some spatial patterning,
-with high mercury concentrations in lakes in the northeast and north
-central United States. The histogram shows that mercury concentration is
-right-skewed, with most lakes having a low value of mercury
-concentration but a few having a much higher concentration. Fig.
+Fig. \ref{fig:merc} shows a map and histogram of mercury concentration
+in all 986 NLA lakes. The map shows mercury concentration exhibits some
+spatial patterning, with high mercury concentrations in the northeast
+and north central United States. The histogram shows that mercury
+concentration is right-skewed, with most lakes having a low value of
+mercury concentration but a few having a much higher concentration. Fig.
 \ref{fig:merc} also shows mercury concentration's empirical
@@
 differences (semivariance) among mercury concentration at different
-distances apart. When a process is spatially correlated (has spatial
+distances apart. When a process has spatial covariance (exhibits spatial
 dependence), the semivariance tends to be smaller at small distances and
@@
 (design-based) or predicted (model-based) the mean mercury concentration
-and its standard error using using design-based and model-based
-approaches. For the model-based analyses, the exponential covariance was
-used. Table \ref{tab:appliedtab} shows the results from these analyses.
-For all four sampling-analysis combinations, the true realized mean
-mercury concentration is within the bounds of the 95\% confidence
-(design-based) or prediction (model-based) intervals. Though we should
-not generalize these results to other samples from these data, we do
-note a couple of patterns. The design-based IRS analysis shows the
-largest standard error: a likely reason is that this is the only
-approach that does not incorporate any spatial information regarding
-mercury concentration. Both analyses using GRTS sampling have lower
-standard errors than both analyses using IRS sampling. We expect that
-these patterns are consistent with other samples from these data because
-mercury concentration exhibits spatial patterning, so a spatially
-balanced sample should usually yield a lower standard error.
+and constructed 95\% confidence (design-based) and 95\% (model-based)
+prediction intervals. For the model-based analyses, the exponential
+covariance was used. Table \ref{tab:appliedtab} shows the results from
+these analyses. For all four sampling-analysis combinations, the true
+realized mean mercury concentration is within the bounds of the 95\%
+confidence (design-based) or 95\% prediction (model-based) intervals.
+Though we should not generalize these results to other samples from
+these data, we do note a couple of patterns. The design-based IRS
+analysis shows the largest standard error: a likely reason is that this
+is the only approach that does not incorporate any spatial information
+regarding mercury concentration. Both analyses using GRTS sampling have
+lower standard errors than both analyses using IRS sampling. We expect
+that these patterns are consistent with other samples from these data
+because mercury concentration exhibits spatial dependence, so a
+spatially balanced sample should usually yield a lower standard error.
 
@@
 \end{tabular}
-\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using the sample of 100 lakes in the NLA data.
+\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.
 The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.}
@@
 The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms that can be used to analyze data. The
-design-based approach incorporates randomness through sampling to
-estimate population parameters. The model-based approach incorporates
-randomness through distributional assumptions to predict realized values
-of a random process. Though these approaches have often been compared in
-the literature both from theoretical and analytical perspectives, our
+fundamentally different paradigms. The design-based approach
+incorporates randomness through sampling to estimate population
+parameters. The model-based approach incorporates randomness through
+distributional assumptions to predict realized values of a stochastic
+process. Though these approaches have often been compared in the
+literature from theoretical and analytical perspectives, our
 contribution lies in studying them in a spatial context while
-implementing spatially balanced sampling. Aside from the theoretical
-differences described, a few analytical findings from the simulation
-study are particularly notable. First, the sampling decision (GRTS vs
-IRS) is most important when using a design-based analysis. Though
-GRTS-Model still outperformed IRS-Model, the model-based analysis
+implementing spatially balanced sampling and the local neighborhood
+variance estimator (in the design-based approach). Aside from the
+theoretical differences described, a few analytical findings from the
+simulation study are particularly notable. First, the sampling decision
+(IRS vs GRTS) is most important when using a design-based analysis.
+Though GRTS-Model still outperformed IRS-Model, the model-based analysis
 mitigated much of the inefficiency of the IRS sample. Second,
@@
 generally performed at least as well as their IRS counterparts when
-there was no spatial correlation and noticeably better than their IRS
-counterparts when there was spatial correlation. Third, as the strength
-of spatial correlation increases, the gap in rMS(P)E between IRS-Design
+there was no spatial covariance and noticeably better than their IRS
+counterparts when there was spatial covariance. Third, as the strength
+of spatial covariance increases, the gap in rMS(P)E between IRS-Design
 and the other sampling-analysis combinations also increases. Fourth and
@@
 estimated spatial covariance parameters, which help better understand
-the dependence structure in the process of study. Model selection is
-also possible using model-based approaches and criteria such as cross
-validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
-approaches are capable of more efficient small-area estimation than
-design-based approaches by leveraging distributional assumptions in
-areas with few observed sites. Model-based approaches can also compute
-site-by-site predictions at unobserved locations and use them to
-construct informative visualizations. The benefits and drawbacks of the
-design-based and model-based approaches should be considered alongside
-the particular goals of a study when deciding which approach is most
-appropriate to implement.
+the dependence structure in the stochastic process of study. Model
+selection is also possible using model-based approaches and criteria
+such as cross validation, likelihood ratio tests, or AIC (Akaike, 1974).
+Model-based approaches are capable of more efficient small-area
+estimation than design-based approaches by leveraging distributional
+assumptions in areas with few observed sites. Model-based approaches can
+also compute site-by-site predictions at unobserved locations and use
+them to construct informative visualizations. In short, when deciding
+whether the design-based or model-based approach is more appropriate to
+implement, the benefits and drawbacks of each approach should be
+considered alongside the particular goals of the study.
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\section*{Data and Code Availability}\label{data-and-code-availability}}
-\addcontentsline{toc}{section}{Data and Code Availability}
-
-This manuscript has a supplementary R package that contains all of the
-data and code used in its creation. The supplementary R package is
-hosted on GitHub. Instructions for download at available at
-
-\url{https://github.com/michaeldumelle/DvMsp}.
-
-\hypertarget{supporting-information}{%
-\section*{Supporting Information}\label{supporting-information}}
-\addcontentsline{toc}{section}{Supporting Information}
-
-In the supporting information, we provide tables presenting summary
-statistics for all 36 simulation scenarios.
-
-\hypertarget{author-contributions}{%
-\section*{Author Contributions}\label{author-contributions}}
-\addcontentsline{toc}{section}{Author Contributions}
+\hypertarget{author-contribution-statement}{%
+\section*{Author Contribution
+Statement}\label{author-contribution-statement}}
+\addcontentsline{toc}{section}{Author Contribution Statement}
 
@@
 drafts and gave final approval for publication.
+
+\hypertarget{data-and-code-availability}{%
+\section*{Data and Code Availability}\label{data-and-code-availability}}
+\addcontentsline{toc}{section}{Data and Code Availability}
+
+This manuscript has a supplementary R package that contains all of the
+data and code used in its creation. The supplementary R package is
+hosted on GitHub. Instructions for download at available at
+
+\url{https://github.com/michaeldumelle/DvMsp}.
+
+If the manuscript is accepted, this repository will be archived in
+Zenodo.
+
+\hypertarget{supporting-information}{%
+\section*{Supporting Information}\label{supporting-information}}
+\addcontentsline{toc}{section}{Supporting Information}
+
+In the supporting information, we provide tables of summary statistics
+for all 36 simulation scenarios.
 

### 2021-12-22 inst/scripts/figures/simulation.R
slight update to include relative standard error values

@@ -24,8 +24,11 @@ combo_data <- combo_data %>%
   mutate(resptype = fct_relevel(resptype, c("normal", "lognormal"))) %>%
   group_by(sim) %>%
   mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
+  mutate(designirsmil = if_else(approach == "Design IRS", true = med_ci_len, false = NA_real_)) %>%
   fill(designirsrmspe, .direction = "downup") %>%
+  fill(designirsmil, .direction = "downup") %>%
   mutate(rel_efficiency = rmspe / designirsrmspe) %>%
+  mutate(mil_efficiency = med_ci_len / designirsmil) %>%
   ungroup() %>%
   mutate(psill_ratio = 1 - nugget_ratio) %>%
   mutate(n_factor = factor(n))

### 2021-12-22 inst/manuscript/manuscript.Rmd
more minor updates

@@
 
-We used a simulation study to investigate performance of four sampling-analysis combinations: Independent Random Sampling (IRS) with a design-based analysis, called "IRS-Design" (IRS samples are not spatially balanced because population units are selected independent of their spatial locations); IRS with a model-based analysis, called "IRS-Model"; GRTS sampling with a design-based analysis, called "GRTS-Design"; and GRTS sampling with a model-based analysis, called "GRTS-Model".
-These combinations are also provided in Table \ref{tab:designanalysis}.
+We used a simulation study to investigate performance of four sampling-analysis combinations.
+The first sampling-analysis combination is IRS-Design.
+In IRS-Design, samples are selected using the Independent Random Sampling (IRS) algorithm.
+The IRS algorithm ignores the spatial locations of the population units, which implies IRS samples are not spatially balanced.
+In IRS-Design, samples are analyzed using the design-based approach with an IRS variance estimator that does not incorporate the spatial locations of the units in the sample.
+The second sampling-analysis combination is IRS-Model, where samples are selected using the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The third sampling-analysis combination is GRTS-Design, where samples are selected using the GRTS algorithm and analyzed using the design-based approach with the local neighborhood variance estimator.
+The fourth and final sampling-analysis combination is GRTS-Model, where samples are selected using the GRTS algorithm and analyzed using the model-based approach via REML estimation.
+These sampling-analysis combinations are also provided in Table \ref{tab:designanalysis}.
+Lastly we note that for both the IRS and GRTS samples, equal inclusion probabilities were assumed for all population units.
+When IRS assumes equal inclusion probabilities for all population units, the algorithm is equivalent to "simple random sampling."
 
@@
 For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a lognormal random variable whose total variance is 2.
+The response was then exponentiated, yielding a lognormal random variable whose total variance was 2.
 The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed (i.e., not normal).
@@
 After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% interval.
-The GRTS algorithm and the local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
-FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal] and covariance parameters were estimated using Restricted Maximum Likelihood  [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
+FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
 
@@
 The United States Environmental Protection Agency (USEPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) to assess the water quality of various bodies of water in the contiguous United States.
-We will use data from the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and water quality [@USEPA2012NLA].
-Specifically, we will analyze mercury concentration in lakes.
-Although we can calculate the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
+One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality [@USEPA2012NLA].
+We will analyze mercury concentration data collected at 986 lakes as part of the 2012 NLA.
+Although we can calculate the true mean mercury concentration values for these 986 lakes, here we will explore whether or not we can obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
 For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% confidence (design-based) or 95% prediction (model-based) intervals from this sample of 100 lakes, which we compare to the true mean mercury concentration from all 986 lakes.
@@
 \ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
-So using the GRTS sampling plan or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial covariance.
+So using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial covariance.
 When there is spatial covariance (Fig.
@@
 Though we should not generalize these results to other samples from these data, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration.
-Both analyses using GRTS sampling have lower standard errors than both analyses using IRS sampling.
-We expect that these patterns are consistent with other samples from these data because mercury concentration exhibits spatial dependence, so a spatially balanced sample should usually yield a lower standard error.
+The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial locations.
+Additionally, both analyses using GRTS sampling have lower standard errors than both analyses using IRS sampling.
 
@@
 First, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
-Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated much of the inefficiency of the IRS sample.
+Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated most of the inefficiencies that result from the IRS samples lacking spatial balance.
 Second, independent of the analysis approach, we found no reason to prefer IRS over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial covariance and noticeably better than their IRS counterparts when there was spatial covariance.
@@
 Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
-Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations.
+Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
 In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.

### 2021-12-22 inst/manuscript/manuscript.tex
more minor updates

@@
 We used a simulation study to investigate performance of four
-sampling-analysis combinations: Independent Random Sampling (IRS) with a
-design-based analysis, called ``IRS-Design'' (IRS samples are not
-spatially balanced because population units are selected independent of
-their spatial locations); IRS with a model-based analysis, called
-``IRS-Model''; GRTS sampling with a design-based analysis, called
-``GRTS-Design''; and GRTS sampling with a model-based analysis, called
-``GRTS-Model''. These combinations are also provided in Table
-\ref{tab:designanalysis}.
+sampling-analysis combinations. The first sampling-analysis combination
+is IRS-Design. In IRS-Design, samples are selected using the Independent
+Random Sampling (IRS) algorithm. The IRS algorithm ignores the spatial
+locations of the population units, which implies IRS samples are not
+spatially balanced. In IRS-Design, samples are analyzed using the
+design-based approach with an IRS variance estimator that does not
+incorporate the spatial locations of the units in the sample. The second
+sampling-analysis combination is IRS-Model, where samples are selected
+using the IRS algorithm and analyzed using the model-based approach via
+Restricted Maximum Likelihood (REML) estimation (Harville, 1977;
+Patterson and Thompson, 1971; Wolfinger et al., 1994). The third
+sampling-analysis combination is GRTS-Design, where samples are selected
+using the GRTS algorithm and analyzed using the design-based approach
+with the local neighborhood variance estimator. The fourth and final
+sampling-analysis combination is GRTS-Model, where samples are selected
+using the GRTS algorithm and analyzed using the model-based approach via
+REML estimation. These sampling-analysis combinations are also provided
+in Table \ref{tab:designanalysis}. Lastly we note that for both the IRS
+and GRTS samples, equal inclusion probabilities were assumed for all
+population units. When IRS assumes equal inclusion probabilities for all
+population units, the algorithm is equivalent to ``simple random
+sampling.''
 
@@
 instead of 2. The response was then exponentiated, yielding a lognormal
-random variable whose total variance is 2. The lognormal responses were
+random variable whose total variance was 2. The lognormal responses were
 used to evaluate performance of the sampling-analysis approaches for
@@
 the proportion of times the true mean is contained in its 95\% interval.
-The GRTS algorithm and the local neighborhood variance estimator are
-available in the \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle
-et al., 2021). FPBK is available in the \texttt{sptotal}
-\textbf{\textsf{R}} package (Higham et al., 2021) and covariance
-parameters were estimated using Restricted Maximum Likelihood (Harville,
-1977; Patterson and Thompson, 1971; Wolfinger et al., 1994).
+The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
+neighborhood variance estimator are available in the \texttt{spsurvey}
+\textbf{\textsf{R}} package (Dumelle et al., 2021). FPBK is available in
+the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).
 
@@
 assess the water quality of various bodies of water in the contiguous
-United States. We will use data from the 2012 National Lakes Assessment
+United States. One component of NARS is the National Lakes Assessment
 (NLA), which measures various aspects of lake health and water quality
-(USEPA, 2012). Specifically, we will analyze mercury concentration in
-lakes. Although we can calculate the true mean mercury concentration
-values for the 986 lakes from the 2012 NLA, we will explore whether or
-not we obtain an adequately precise estimate for the realized mean
-mercury concentration if we sample only 100 of the 986 lakes. For each
-of the four familiar sampling-analysis combinations (IRS-Design,
-IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or
-predict (model-based) the mean mercury concentration and construct 95\%
-confidence (design-based) or 95\% prediction (model-based) intervals
-from this sample of 100 lakes, which we compare to the true mean mercury
-concentration from all 986 lakes.
+(USEPA, 2012). We will analyze mercury concentration data collected at
+986 lakes as part of the 2012 NLA. Although we can calculate the true
+mean mercury concentration values for these 986 lakes, here we will
+explore whether or not we can obtain an adequately precise estimate for
+the realized mean mercury concentration if we sample only 100 of the 986
+lakes. For each of the four familiar sampling-analysis combinations
+(IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
+(design-based) or predict (model-based) the mean mercury concentration
+and construct 95\% confidence (design-based) or 95\% prediction
+(model-based) intervals from this sample of 100 lakes, which we compare
+to the true mean mercury concentration from all 986 lakes.
 
@@
 combinations have approximately equal rMS(P)E. So using the GRTS
-sampling plan or a model-based analysis does not result in much, if any,
+algorithm or a model-based analysis does not result in much, if any,
 loss in efficiency compared to IRS-Design when there is no spatial
@@
 analysis shows the largest standard error: a likely reason is that this
-is the only approach that does not incorporate any spatial information
-regarding mercury concentration. Both analyses using GRTS sampling have
-lower standard errors than both analyses using IRS sampling. We expect
-that these patterns are consistent with other samples from these data
-because mercury concentration exhibits spatial dependence, so a
-spatially balanced sample should usually yield a lower standard error.
+is the only approach that does not incorporate any spatial locations.
+Additionally, both analyses using GRTS sampling have lower standard
+errors than both analyses using IRS sampling.
 
@@
 Though GRTS-Model still outperformed IRS-Model, the model-based analysis
-mitigated much of the inefficiency of the IRS sample. Second,
-independent of the analysis approach, we found no reason to prefer IRS
-over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model
-generally performed at least as well as their IRS counterparts when
-there was no spatial covariance and noticeably better than their IRS
-counterparts when there was spatial covariance. Third, as the strength
-of spatial covariance increases, the gap in rMS(P)E between IRS-Design
-and the other sampling-analysis combinations also increases. Fourth and
-finally, when the response was normal, interval coverage for all
-sampling-analysis combinations was very close to 95\% for all sample
-sizes; when the response was lognormal, interval coverage for all
-sampling and analysis was between 90\% and 95\% and closest to 95\% when
-\(n = 200\).
+mitigated most of the inefficiencies that result from the IRS samples
+lacking spatial balance. Second, independent of the analysis approach,
+we found no reason to prefer IRS over GRTS for sampling spatial data --
+GRTS-Design and GRTS-Model generally performed at least as well as their
+IRS counterparts when there was no spatial covariance and noticeably
+better than their IRS counterparts when there was spatial covariance.
+Third, as the strength of spatial covariance increases, the gap in
+rMS(P)E between IRS-Design and the other sampling-analysis combinations
+also increases. Fourth and finally, when the response was normal,
+interval coverage for all sampling-analysis combinations was very close
+to 95\% for all sample sizes; when the response was lognormal, interval
+coverage for all sampling and analysis was between 90\% and 95\% and
+closest to 95\% when \(n = 200\).
 
@@
 also compute site-by-site predictions at unobserved locations and use
-them to construct informative visualizations. In short, when deciding
-whether the design-based or model-based approach is more appropriate to
-implement, the benefits and drawbacks of each approach should be
-considered alongside the particular goals of the study.
+them to construct informative visualizations like smoothed maps. In
+short, when deciding whether the design-based or model-based approach is
+more appropriate to implement, the benefits and drawbacks of each
+approach should be considered alongside the particular goals of the
+study.
 

### 2021-12-22 inst/scripts/figures/simulation.R

@@ -46,7 +46,7 @@ rmspe_eff <- combo_data %>%
   geom_jitter(width = 0.24, size = 2) +
   scale_colour_manual(values = colour_scale) +
   theme_bw(base_size = base_size) +
-  labs(x = "Sample Size", colour = "Approach", y = "Relative Efficiency") +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
   theme(
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),

### 2021-12-22 inst/manuscript/manuscript.tex
minor updates to include estimated standard errors

@@
 root-mean-squared-prediction error for the model-based approaches), and
-the proportion of times the true mean is contained in its 95\% interval.
-The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
-neighborhood variance estimator are available in the \texttt{spsurvey}
-\textbf{\textsf{R}} package (Dumelle et al., 2021). FPBK is available in
-the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).
+the proportion of times the true mean is contained in its 95\%
+confidence (design-based) or 95\% prediction (model-based) interval. The
+95\% confidence intervals (design-based) and 95\% prediction intervals
+(model-based) were constructed using the normal distribution.
+Justification for this comes from the asymptotic normality of means via
+the Central Limit Theorem (under some assumptions). The IRS algorithm,
+IRS variance estimator, GRTS algorithm, and local neighborhood variance
+estimator are available in the \texttt{spsurvey} \textbf{\textsf{R}}
+package (Dumelle et al., 2021). FPBK is available in the
+\texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).
 
@@
 
-Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling
+Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling
 analysis combinations using the random location layout with
@@
 \end{equation*} When there is no spatial covariance (Fig.
-\ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
+\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
 combinations have approximately equal rMS(P)E. So using the GRTS
@@
 loss in efficiency compared to IRS-Design when there is no spatial
-covariance. When there is spatial covariance (Fig. \ref{fig:figeff},
+covariance. When there is spatial covariance (Fig. \ref{fig:rmspe_eff},
 ``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
@@
   \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:figeff}
+  \label{fig:rmspe_eff}
 \end{figure}
 
-We also studied 95\% interval coverage among the sampling-analysis
-combinations. The design-based confidence intervals and model-based
-prediction intervals were constructed using the normal distribution.
-Justification for this comes from the asymptotic normality of means via
-the Central Limit Theorem. Fig. \ref{fig:figconf} shows the 95\%
-interval coverage for each of the four sampling-analysis combinations in
-the random location layout. Within each scenario, the sampling-analysis
-combinations tend to have fairly similar interval coverage. Coverage in
-the normal response scenarios was usually near 95\%, while coverage in
-the lognormal response scenarios varied from from 90\% to 95\% but
-increased with the sample size. At a sample size of 200, all four
-sampling-analysis combinations had approximately 95\% interval coverage
-in both response scenarios for all dependent error proportions. These
+Fig. \ref{fig:mse_eff} shows the relative mean standard errors (MStdE)
+of the four sampling-analysis combinations using the random location
+layout with ``IRS-Design'' as the baseline. The MStdE is defined as
+\begin{equation*}
+\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
+\end{equation*} Many general takeaways regarding MStdE are similar to
+general takeaways regarding rMS(P)E: there seems to be no benefit to
+using IRS, even when there is no spatial covariance; as the strength of
+spatial covariance increases, the gap in MStdE between IRS-Design and
+the other sampling-analysis combinations widens; and IRS-Model
+outperforms IRS-Design by a large margin. This is not surprising because
+all sampling-analysis combinations had nearly zero average bias, thus
+rMS(P)E is driven by the variance of the estimators (design-based) or
+predictors (model-based). We do note that between GRTS-Design and
+GRTS-Model, GRTS-Design had lower MStdE when there was no spatial
+covariance or a medium amount of spatial covariance (Fig.
+\ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows) and
+GRTS-Model had lower MStdE when there was a high amount of spatial
+covariance Fig. \ref{fig:mse_eff}, ``Prop DE: 0.9'' row). These
 conclusions are similar to those observed in the grid location layout,
-so we omit a grid location layout figure here. Tables for interval
-coverage in all 36 simulation scenarios are provided in the supporting
-information.
+so we omit a grid location layout figure here. Tables for MStdE in all
+36 simulation scenarios are provided in the supporting information.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
+  \caption{Relative standard errors in the simulation study for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:mse_eff}
+\end{figure}
+
+Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
+four sampling-analysis combinations in the random location layout.
+Within each scenario, the sampling-analysis combinations tend to have
+fairly similar interval coverage. Coverage in the normal response
+scenarios was usually near 95\%, while coverage in the lognormal
+response scenarios varied from from 90\% to 95\% but increased with the
+sample size. At a sample size of 200, all four sampling-analysis
+combinations had approximately 95\% interval coverage in both response
+scenarios for all dependent error proportions. These conclusions are
+similar to those observed in the grid location layout, so we omit a grid
+location layout figure here. Tables for interval coverage in all 36
+simulation scenarios are provided in the supporting information.
 

### 2021-12-22 inst/scripts/figures/simulation.R

@@ -47,6 +47,7 @@ rmspe_eff <- combo_data %>%
   scale_colour_manual(values = colour_scale) +
   theme_bw(base_size = base_size) +
   labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
+  lims(y = c(0.4, 1.1)) +
   theme(
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
@@ -63,7 +64,40 @@ if (write_out) {
   )
 }
 
+
 # figure 3
+colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
+psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
+mse_eff <- combo_data %>%
+  ggplot(aes(x = n_factor, y = mil_efficiency, colour = approach)) +
+  facet_grid(
+    psill_ratio ~ resptype,
+    labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
+  ) +
+  geom_jitter(width = 0.24, size = 2) +
+  scale_colour_manual(values = colour_scale) +
+  theme_bw(base_size = base_size) +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative MStdE") +
+  lims(y = c(0.4, 1.1)) +
+  theme(
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+  )
+
+### rmspe_eff
+if (write_out) {
+  ggsave(
+    plot = mse_eff,
+    file = here("inst", "manuscript", "figures", "mse_eff.jpeg"),
+    dpi = 300,
+    width = 5.07,
+    height = 4.39
+  )
+}
+
+
+# figure 4
 coverage <- combo_data %>%
   ggplot(aes(x = n_factor, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95) +

### 2021-12-22 inst/scripts/sim_runs.R
update simulations to include standard errors in summaries

@@ -25,12 +25,7 @@ parm_df <- expand_grid(N, n, psill_ratio,
          psill = total_var * psill_ratio,
          nugget = total_var * nugget_ratio)
 
-# source functions for now
-source("R/sim_pop.R")
-source("R/sim_trial.R")
-
-# library for later
-# library(SpatialDesignVModel)
+library(DvMsp)
 
 ## loop through each row of parm_df (could be done with purrr::pmap() instead)
 for (i in 1:nrow(parm_df)) {
@@ -38,7 +33,7 @@ for (i in 1:nrow(parm_df)) {
   seed <- 1:n_trials
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
-  clusterExport(cluster, varlist = c("sim_pop", "covmx_exp"))
+  clusterEvalQ(cluster, library(DvMsp)) # export DvMsp to cluster
   clusterEvalQ(cluster, library(spsurvey)) # export spsurvey to cluster v >= 5.1.0
   clusterEvalQ(cluster, library(sptotal)) # export sptotal to cluster
   clusterEvalQ(cluster, library(dplyr)) # export dplyr to cluster
@@ -89,6 +84,7 @@ for (i in 1:nrow(parm_df)) {
       mpbias = mean(resid),
       rmspe = sqrt(mean(resid^2)),
       medae = median(abs(resid)),
+      mse = mean(sd),
       coverage = mean(cover),
       prcloser = mean(closer),
       med_ci_len = median(ub - lb)

### 2021-12-22 inst/manuscript/manuscript.Rmd

@@
 Then we recorded the bias, squared error, and interval coverage for all sampling-analysis combinations.
-After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% interval.
+After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% confidence (design-based) or 95% prediction (model-based) interval.
+The 95% confidence intervals (design-based) and 95% prediction intervals (model-based) were constructed using the normal distribution.
+Justification for this comes from the asymptotic normality of means via the Central Limit Theorem (under some assumptions).
 The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
@@
 
-Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative rMS(P)E is defined as
+Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative rMS(P)E is defined as
 \begin{equation*}
@@
 When there is no spatial covariance (Fig.
-\ref{fig:figeff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
 So using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial covariance.
 When there is spatial covariance (Fig.
-\ref{fig:figeff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
+\ref{fig:rmspe_eff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
 As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
@@
   \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:figeff}
+  \label{fig:rmspe_eff}
 \end{figure}
 
-We also studied 95% interval coverage among the sampling-analysis combinations.
-The design-based confidence intervals and model-based prediction intervals were constructed using the normal distribution.
-Justification for this comes from the asymptotic normality of means via the Central Limit Theorem.
+Fig.
+\ref{fig:mse_eff} shows the relative mean standard errors (MStdE) of the four sampling-analysis combinations using the random location layout with "IRS-Design" as the baseline.
+The MStdE is defined as
+\begin{equation*}
+\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
+\end{equation*}
+Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms IRS-Design by a large margin.
+This is not surprising because all sampling-analysis combinations had nearly zero average bias, thus rMS(P)E is driven by the variance of the estimators (design-based) or predictors (model-based).
+We do note that between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was no spatial covariance or a medium amount of spatial covariance (Fig.
+\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows) and GRTS-Model had lower MStdE when there was a high amount of spatial covariance Fig.
+\ref{fig:mse_eff}, "Prop DE: 0.9" row).
+These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+Tables for MStdE in all 36 simulation scenarios are provided in the supporting information.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
+  \caption{Relative standard errors in the simulation study for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:mse_eff}
+\end{figure}
+
 Fig.

### 2021-12-22 inst/manuscript/manuscript.Rmd
more minor updates to incorporate mean standard error addition

@@
   1.
-The design-based and model-based approaches to frequentist statistical inference lie on fundamentally different foundations.
-In the design-based approach, inference depends on random sampling.
-In the model-based approach, inference depends on distributional assumptions.
+The design-based and model-based approaches to frequentist statistical inference rest on fundamentally different foundations.
+In the design-based approach, inference relies on random sampling.
+In the model-based approach, inference relies on distributional assumptions.
 We compare the approaches for finite population spatial data.
@@
 In the simulations, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
-In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval length and coverage.
   3.
@@
 
-Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based inference; Spatially balanced sampling; Spatial covariance
+Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Local neighborhood variance estimator; Model-based inference; Restricted Maximum Likelihood (REML) estimation; Spatially balanced sampling; Spatial covariance
 
@@
 The first sampling-analysis combination is IRS-Design.
-In IRS-Design, samples are selected using the Independent Random Sampling (IRS) algorithm.
-The IRS algorithm ignores the spatial locations of the population units, which implies IRS samples are not spatially balanced.
-In IRS-Design, samples are analyzed using the design-based approach with an IRS variance estimator that does not incorporate the spatial locations of the units in the sample.
-The second sampling-analysis combination is IRS-Model, where samples are selected using the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
-The third sampling-analysis combination is GRTS-Design, where samples are selected using the GRTS algorithm and analyzed using the design-based approach with the local neighborhood variance estimator.
-The fourth and final sampling-analysis combination is GRTS-Model, where samples are selected using the GRTS algorithm and analyzed using the model-based approach via REML estimation.
+In IRS-Design, samples are selected with the Independent Random Sampling (IRS) algorithm.
+The IRS algorithm ignores the spatial locations of the population units, thus IRS samples are not spatially balanced.
+In IRS-Design, samples are analyzed using the design-based approach via the Horvitz-Thompson mean estimator and an IRS variance estimator that does not incorporate the spatial locations of the units in the sample.
+The second sampling-analysis combination is IRS-Model, where samples are selected with the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The third sampling-analysis combination is GRTS-Design, where samples are selected with the GRTS algorithm and analyzed using the design-based approach via the Horvitz-Thompson mean estimator and the local neighborhood variance estimator (which does incorporate the spatial locations of the units in the sample).
+The fourth and final sampling-analysis combination is GRTS-Model, where samples are selected with the GRTS algorithm and analyzed using the model-based approach via REML estimation.
 These sampling-analysis combinations are also provided in Table \ref{tab:designanalysis}.
 Lastly we note that for both the IRS and GRTS samples, equal inclusion probabilities were assumed for all population units.
-When IRS assumes equal inclusion probabilities for all population units, the algorithm is equivalent to "simple random sampling."
+When IRS assumes equal inclusion probabilities for all population units, the algorithm is equivalent to simple random sampling (SRS).
 
@@
 For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error.
-The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are the dependent random error variance (partial sill) and independent random error variance (nugget), respectively, from Equation$~$\ref{equation:expcov}.
+The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are the dependent random error variance (partial sill) and independent random error variance (nugget) from Equation$~$\ref{equation:expcov}, respectively.
 The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
@@
 In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
-Then we recorded the bias, squared error, and interval coverage for all sampling-analysis combinations.
-After all 2000 trials, we summarized the long-run performance of the combinations by calculating average bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), and the proportion of times the true mean is contained in its 95% confidence (design-based) or 95% prediction (model-based) interval.
-The 95% confidence intervals (design-based) and 95% prediction intervals (model-based) were constructed using the normal distribution.
+Then we recorded the bias, squared error, standard error, and interval coverage for all sampling-analysis combinations.
+After all 2000 trials, we summarized the long-run performance of the combinations by calculating mean bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), MStdE (mean standard error), and the proportion of times the true mean is contained in its 95% confidence (design-based) or 95% prediction (model-based) interval.
+The 95% intervals were constructed using the normal distribution.
 Justification for this comes from the asymptotic normality of means via the Central Limit Theorem (under some assumptions).
+Quantifying mean bias and rMS(P)E is important because they help us understand how far (under different loss metrics) the estimates (design-based) or predictions (model-based) tend to be from the true mean.
+Quantifying MStdE is important because it helps us understand how precise intervals tend to be.
+Quantifying interval coverage is important because it helps us understand how often our 95% intervals actually contain the true mean.
+
 The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
@@
 One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality [@USEPA2012NLA].
-We will analyze mercury concentration data collected at 986 lakes as part of the 2012 NLA.
-Although we can calculate the true mean mercury concentration values for these 986 lakes, here we will explore whether or not we can obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% confidence (design-based) or 95% prediction (model-based) intervals from this sample of 100 lakes, which we compare to the true mean mercury concentration from all 986 lakes.
+We will analyze mercury concentration data collected at 986 lakes from the 2012 NLA.
+Although we can calculate the true mean mercury concentration values for these 986 lakes, here we will explore whether or not we can obtain an adequately precise estimate (design-based) or prediction (model-based) for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
+For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% intervals from this sample of 100 lakes and compare to the true mean mercury concentration from all 986 lakes.
 
@@
 
-The average bias was nearly zero for all four sampling-analysis combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
-Tables for average bias in all 36 simulation scenarios are provided in the supporting information.
+The mean bias was nearly zero for all four sampling-analysis combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+Tables for mean bias in all 36 simulation scenarios are provided in the supporting information.
 
@@
 \ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
-So using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design when there is no spatial covariance.
+So when there is no spatial covariance, using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design.
 When there is spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to perform best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
+\ref{fig:rmspe_eff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to have the lowest rMS(P)E, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
 As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
-Finally we note that when there is spatial covariance, IRS-Model outperforms IRS-Design by a large margin, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
-These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+Finally we note that when there is spatial covariance, IRS-Model has a much lower rMS(P)E than IRS-Design, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
+These rMS(P)E conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
 Tables for rMS(P)E in all 36 simulation scenarios are provided in the supporting information.
@@
 
-Fig.
-\ref{fig:mse_eff} shows the relative mean standard errors (MStdE) of the four sampling-analysis combinations using the random location layout with "IRS-Design" as the baseline.
-The MStdE is defined as
+Fig. \ref{fig:mse_eff} shows the relative MStdE of the four sampling-analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative MStdE is defined as
 \begin{equation*}
@@
 \end{equation*}
-Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms IRS-Design by a large margin.
-This is not surprising because all sampling-analysis combinations had nearly zero average bias, thus rMS(P)E is driven by the variance of the estimators (design-based) or predictors (model-based).
+Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms (has a much lower MStdE) IRS-Design by a large margin.
+These fact that the rMS(P)E and MStdE findings are similar is not particularly surprising because the mean bias for all sampling-analysis combinations was nearly zero, thus rMS(P)E is driven by the standard error of the estimators (design-based) or predictors (model-based).
 We do note that between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was no spatial covariance or a medium amount of spatial covariance (Fig.
-\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows) and GRTS-Model had lower MStdE when there was a high amount of spatial covariance Fig.
+\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows) and GRTS-Model had lower MStdE when there was a high amount of spatial covariance (Fig.
 \ref{fig:mse_eff}, "Prop DE: 0.9" row).
-These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+These MStdE conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
 Tables for MStdE in all 36 simulation scenarios are provided in the supporting information.
@@
   \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
-  \caption{Relative standard errors in the simulation study for the four sampling-analysis combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative MStdE in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
   \label{fig:mse_eff}
@@
 \ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
-Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage.
-Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios varied from from 90% to 95% but increased with the sample size.
+Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-Design coverage is usually a few percentage points lower than the other combinations.
+Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios usually varied from 90% to 95% but increased with the sample size.
 At a sample size of 200, all four sampling-analysis combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
-These conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
 Tables for interval coverage in all 36 simulation scenarios are provided in the supporting information.
@@
 The empirical semivariogram can be used as a tool to visualize spatial dependence.
-It quantifies the halved squared differences (semivariance) among mercury concentration at different distances apart.
-When a process has spatial covariance (exhibits spatial dependence), the semivariance tends to be smaller at small distances and larger at large distances.
+It quantifies the mean of the halved squared differences (semivariance) among all pairs of mercury concentrations at different distances apart.
+When a process has spatial covariance (exhibits spatial dependence), the mean semivariance tends to be smaller at small distances and larger at large distances.
 The empirical semivariogram in Fig.
-\ref{fig:merc} suggests that mercury concentration is exhibits spatial dependence.
-Lastly we note that the realized mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
+\ref{fig:merc} suggests that mercury concentration exhibits spatial dependence.
+Lastly we note that the true mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
 \begin{figure}
@@
 \end{subfigure}
-\caption{Mercury concentration visualizations for the population (Hg) for 986 lakes in the NLA data.
+\caption{Mercury concentration (Hg) visualizations for all 986 lakes in the NLA data.
 A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
@@
 Table \ref{tab:appliedtab} shows the results from these analyses.
-For all four sampling-analysis combinations, the true realized mean mercury concentration is within the bounds of the 95% confidence (design-based) or 95% prediction (model-based) intervals.
-Though we should not generalize these results to other samples from these data, we do note a couple of patterns.
-The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial locations.
-Additionally, both analyses using GRTS sampling have lower standard errors than both analyses using IRS sampling.
+Though we should not generalize these results to other samples from these data, we do mention a few findings.
+First, IRS-Design has the largest standard error.
+Second, compared to IRS-Design and IRS-Model, GRTS-Design and GRTS-Model are much closer to the true mean mercury concentration (have bias closer to zero) and have much lower standard errors (more precise intervals).
+Third, GRTS-Model has the least amount of bias and the lowest standard error (most precise interval).
+Finally, we note that for all sampling-analysis combinations, the true mean mercury concentration (103.2 ng / g) is within the bounds of the combination's 95% interval.
 
@@
 res_df <- readr::read_csv(here("inst", "output", "application", "application.csv"))
-res_df_drop <- res_df[ ,-2]
-names(res_df_drop) <- c("Approach", "Est/Pred", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.
-The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.")
+# res_df$bias <- res_df$realized_mean - res_df$estimate
+# res_df_drop <- res_df[ , c("approach", "estimate", "bias", "se", "lb", "ub")]
+res_df_drop <- res_df[, ]
+names(res_df_drop) <- c("Approach", "True Mean", "Est/Pred", "SE", "95% LB", "95% UB")
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
 print(xtab_app, include.rownames = FALSE, comment = FALSE)
@@
 The design-based and model-based approaches to statistical inference are fundamentally different paradigms.
-The design-based approach incorporates randomness through sampling to estimate population parameters.
-The model-based approach incorporates randomness through distributional assumptions to predict realized values of a stochastic process.
-Though these approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling and the local neighborhood variance estimator (in the design-based approach).
+The design-based approach relies on random sampling to estimate population parameters.
+The model-based approach relies on distributional assumptions to predict realized values of a stochastic process.
+Though the model-based approach does not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
+While the design-based and model-based approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling and the design-based, local neighborhood variance estimator.
 Aside from the theoretical differences described, a few analytical findings from the simulation study are particularly notable.
-First, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
-Though GRTS-Model still outperformed IRS-Model, the model-based analysis mitigated most of the inefficiencies that result from the IRS samples lacking spatial balance.
-Second, independent of the analysis approach, we found no reason to prefer IRS over GRTS for sampling spatial data -- GRTS-Design and GRTS-Model generally performed at least as well as their IRS counterparts when there was no spatial covariance and noticeably better than their IRS counterparts when there was spatial covariance.
-Third, as the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations also increases.
-Fourth and finally, when the response was normal, interval coverage for all sampling-analysis combinations was very close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling and analysis was between 90% and 95% and closest to 95% when $n = 200$.
+First, independent of the analysis approach, we found no reason to prefer IRS over GRTS when sampling spatial data -- GRTS-Design and GRTS-Model generally had similar rMS(P)E as their IRS counterparts when there was no spatial covariance and lower rMS(P)E than their IRS counterparts when there was spatial covariance.
+Second, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
+Though GRTS-Model still had lower rMS(P)E than IRS-Model, the model-based analysis mitigated most of the rMS(P)E inefficiencies that result from the IRS samples lacking spatial balance.
+Third, as the strength of spatial covariance increases, the gap in rMS(P)E and MStdE between IRS-Design and the other sampling-analysis combinations also increases, likely because IRS-Design is the only combination that ignores spatial locations in sampling and analysis.
+Fourth and finally, when the response was normal, interval coverage for all sampling-analysis combinations was usually close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling-analysis combinations was usually between 90% and 95% and closest to 95% when $n = 200$.
 

### 2021-12-22 inst/manuscript/manuscript.tex
more minor updates to incorporate mean standard error addition

--- 2021-12-22 inst/manuscript/manuscript.tex
@@
     performance is measured using statistics like bias, squared error, and
-    interval coverage.
+    interval length and coverage.
   \item
@@
 Design-based inference; Finite Population Block Kriging (FPBK);
-Generalized Random Tessellation Stratified (GRTS) algorithm; Model-based
-inference; Spatially balanced sampling; Spatial covariance
+Generalized Random Tessellation Stratified (GRTS) algorithm; Local
+neighborhood variance estimator; Model-based inference; Restricted
+Maximum Likelihood (REML) estimation; Spatially balanced sampling;
+Spatial covariance
 
@@
 sampling-analysis combinations. The first sampling-analysis combination
-is IRS-Design. In IRS-Design, samples are selected using the Independent
+is IRS-Design. In IRS-Design, samples are selected with the Independent
 Random Sampling (IRS) algorithm. The IRS algorithm ignores the spatial
-locations of the population units, which implies IRS samples are not
-spatially balanced. In IRS-Design, samples are analyzed using the
-design-based approach with an IRS variance estimator that does not
-incorporate the spatial locations of the units in the sample. The second
-sampling-analysis combination is IRS-Model, where samples are selected
-using the IRS algorithm and analyzed using the model-based approach via
-Restricted Maximum Likelihood (REML) estimation (Harville, 1977;
-Patterson and Thompson, 1971; Wolfinger et al., 1994). The third
-sampling-analysis combination is GRTS-Design, where samples are selected
-using the GRTS algorithm and analyzed using the design-based approach
-with the local neighborhood variance estimator. The fourth and final
+locations of the population units, thus IRS samples are not spatially
+balanced. In IRS-Design, samples are analyzed using the design-based
+approach via the Horvitz-Thompson mean estimator and an IRS variance
+estimator that does not incorporate the spatial locations of the units
+in the sample. The second sampling-analysis combination is IRS-Model,
+where samples are selected with the IRS algorithm and analyzed using the
+model-based approach via Restricted Maximum Likelihood (REML) estimation
+(Harville, 1977; Patterson and Thompson, 1971; Wolfinger et al., 1994).
+The third sampling-analysis combination is GRTS-Design, where samples
+are selected with the GRTS algorithm and analyzed using the design-based
+approach via the Horvitz-Thompson mean estimator and the local
+neighborhood variance estimator (which does incorporate the spatial
+locations of the units in the sample). The fourth and final
 sampling-analysis combination is GRTS-Model, where samples are selected
-using the GRTS algorithm and analyzed using the model-based approach via
+with the GRTS algorithm and analyzed using the model-based approach via
 REML estimation. These sampling-analysis combinations are also provided
@@
 population units. When IRS assumes equal inclusion probabilities for all
-population units, the algorithm is equivalent to ``simple random
-sampling.''
+population units, the algorithm is equivalent to simple random sampling
+(SRS).
 
@@
 \(\sigma^2_2\) are the dependent random error variance (partial sill)
-and independent random error variance (nugget), respectively, from
-Equation\(~\)\ref{equation:expcov}. The total variance,
+and independent random error variance (nugget) from
+Equation\(~\)\ref{equation:expcov}, respectively. The total variance,
 \(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
@@
 confidence (design-based) or 95\% prediction (model-based) intervals.
-Then we recorded the bias, squared error, and interval coverage for all
-sampling-analysis combinations. After all 2000 trials, we summarized the
-long-run performance of the combinations by calculating average bias,
-rMS(P)E (root-mean-squared error for the design-based approaches and
-root-mean-squared-prediction error for the model-based approaches), and
-the proportion of times the true mean is contained in its 95\% interval.
+Then we recorded the bias, squared error, standard error, and interval
+coverage for all sampling-analysis combinations. After all 2000 trials,
+we summarized the long-run performance of the combinations by
+calculating mean bias, rMS(P)E (root-mean-squared error for the
+design-based approaches and root-mean-squared-prediction error for the
+model-based approaches), MStdE (mean standard error), and the proportion
+of times the true mean is contained in its 95\% confidence
+(design-based) or 95\% prediction (model-based) interval. The 95\%
+intervals were constructed using the normal distribution. Justification
+for this comes from the asymptotic normality of means via the Central
+Limit Theorem (under some assumptions). Quantifying mean bias and
+rMS(P)E is important because they help us understand how far (under
+different loss metrics) the estimates (design-based) or predictions
+(model-based) tend to be from the true mean. Quantifying MStdE is
+important because it helps us understand how precise intervals tend to
+be. Quantifying interval coverage is important because it helps us
+understand how often our 95\% intervals actually contain the true mean.
+
 The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
@@
 (USEPA, 2012). We will analyze mercury concentration data collected at
-986 lakes as part of the 2012 NLA. Although we can calculate the true
-mean mercury concentration values for these 986 lakes, here we will
-explore whether or not we can obtain an adequately precise estimate for
-the realized mean mercury concentration if we sample only 100 of the 986
-lakes. For each of the four familiar sampling-analysis combinations
-(IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate
-(design-based) or predict (model-based) the mean mercury concentration
-and construct 95\% confidence (design-based) or 95\% prediction
-(model-based) intervals from this sample of 100 lakes, which we compare
-to the true mean mercury concentration from all 986 lakes.
+986 lakes from the 2012 NLA. Although we can calculate the true mean
+mercury concentration values for these 986 lakes, here we will explore
+whether or not we can obtain an adequately precise estimate
+(design-based) or prediction (model-based) for the realized mean mercury
+concentration if we sample only 100 of the 986 lakes. For each of the
+four familiar sampling-analysis combinations (IRS-Design, IRS-Model,
+GRTS-Design, and GRTS-Model), we estimate (design-based) or predict
+(model-based) the mean mercury concentration and construct 95\%
+intervals from this sample of 100 lakes and compare to the true mean
+mercury concentration from all 986 lakes.
 
@@
 
-The average bias was nearly zero for all four sampling-analysis
+The mean bias was nearly zero for all four sampling-analysis
 combinations in all 36 scenarios, so we omit a more detailed summary of
-those results here. Tables for average bias in all 36 simulation
-scenarios are provided in the supporting information.
-
-Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling
+those results here. Tables for mean bias in all 36 simulation scenarios
+are provided in the supporting information.
+
+Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling
 analysis combinations using the random location layout with
@@
 \end{equation*} When there is no spatial covariance (Fig.
-\ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
-combinations have approximately equal rMS(P)E. So using the GRTS
-algorithm or a model-based analysis does not result in much, if any,
-loss in efficiency compared to IRS-Design when there is no spatial
-covariance. When there is spatial covariance (Fig. \ref{fig:figeff},
-``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
-best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though
-the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and
-IRS-Model is relatively small. As the strength of spatial covariance
-increases, the gap in rMS(P)E between IRS-Design and the other
-sampling-analysis combinations widens. Finally we note that when there
-is spatial covariance, IRS-Model outperforms IRS-Design by a large
-margin, suggesting that the poor design properties of IRS are largely
-mitigated by the model-based analysis. These conclusions are similar to
-those observed in the grid location layout, so we omit a grid location
-layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
-are provided in the supporting information.
+\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
+combinations have approximately equal rMS(P)E. So when there is no
+spatial covariance, using the GRTS algorithm or a model-based analysis
+does not result in much, if any, loss in efficiency compared to
+IRS-Design. When there is spatial covariance (Fig. \ref{fig:rmspe_eff},
+``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to have
+the lowest rMS(P)E, followed by GRTS-Design, IRS-Model, and finally
+IRS-Design, though the difference in relative rMS(P)E among GRTS-Model,
+GRTS-Design, and IRS-Model is relatively small. As the strength of
+spatial covariance increases, the gap in rMS(P)E between IRS-Design and
+the other sampling-analysis combinations widens. Finally we note that
+when there is spatial covariance, IRS-Model has a much lower rMS(P)E
+than IRS-Design, suggesting that the poor design properties of IRS are
+largely mitigated by the model-based analysis. These rMS(P)E conclusions
+are similar to those observed in the grid location layout, so we omit a
+grid location layout figure here. Tables for rMS(P)E in all 36
+simulation scenarios are provided in the supporting information.
 
@@
   \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:figeff}
+  \label{fig:rmspe_eff}
 \end{figure}
 
-We also studied 95\% interval coverage among the sampling-analysis
-combinations. The design-based confidence intervals and model-based
-prediction intervals were constructed using the normal distribution.
-Justification for this comes from the asymptotic normality of means via
-the Central Limit Theorem. Fig. \ref{fig:figconf} shows the 95\%
-interval coverage for each of the four sampling-analysis combinations in
-the random location layout. Within each scenario, the sampling-analysis
-combinations tend to have fairly similar interval coverage. Coverage in
-the normal response scenarios was usually near 95\%, while coverage in
-the lognormal response scenarios varied from from 90\% to 95\% but
-increased with the sample size. At a sample size of 200, all four
-sampling-analysis combinations had approximately 95\% interval coverage
-in both response scenarios for all dependent error proportions. These
+Fig. \ref{fig:mse_eff} shows the relative MStdE of the four
+sampling-analysis combinations using the random location layout with
+``IRS-Design'' as the baseline. The relative MStdE is defined as
+\begin{equation*}
+\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
+\end{equation*} Many general takeaways regarding MStdE are similar to
+general takeaways regarding rMS(P)E: there seems to be no benefit to
+using IRS, even when there is no spatial covariance; as the strength of
+spatial covariance increases, the gap in MStdE between IRS-Design and
+the other sampling-analysis combinations widens; and IRS-Model
+outperforms (has a much lower MStdE) IRS-Design by a large margin. These
+fact that the rMS(P)E and MStdE findings are similar is not particularly
+surprising because the mean bias for all sampling-analysis combinations
+was nearly zero, thus rMS(P)E is driven by the standard error of the
+estimators (design-based) or predictors (model-based). We do note that
+between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when
+there was no spatial covariance or a medium amount of spatial covariance
+(Fig. \ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows) and
+GRTS-Model had lower MStdE when there was a high amount of spatial
+covariance (Fig. \ref{fig:mse_eff}, ``Prop DE: 0.9'' row). These MStdE
 conclusions are similar to those observed in the grid location layout,
-so we omit a grid location layout figure here. Tables for interval
-coverage in all 36 simulation scenarios are provided in the supporting
-information.
+so we omit a grid location layout figure here. Tables for MStdE in all
+36 simulation scenarios are provided in the supporting information.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
+  \caption{Relative MStdE in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:mse_eff}
+\end{figure}
+
+Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
+four sampling-analysis combinations in the random location layout.
+Within each scenario, the sampling-analysis combinations tend to have
+fairly similar interval coverage, though when \(n = 50\) or \(n = 100\),
+GRTS-Design coverage is usually a few percentage points lower than the
+other combinations. Coverage in the normal response scenarios was
+usually near 95\%, while coverage in the lognormal response scenarios
+usually varied from 90\% to 95\% but increased with the sample size. At
+a sample size of 200, all four sampling-analysis combinations had
+approximately 95\% interval coverage in both response scenarios for all
+dependent error proportions. These interval coverage conclusions are
+similar to those observed in the grid location layout, so we omit a grid
+location layout figure here. Tables for interval coverage in all 36
+simulation scenarios are provided in the supporting information.
 
@@
 semivariogram. The empirical semivariogram can be used as a tool to
-visualize spatial dependence. It quantifies the halved squared
-differences (semivariance) among mercury concentration at different
-distances apart. When a process has spatial covariance (exhibits spatial
-dependence), the semivariance tends to be smaller at small distances and
-larger at large distances. The empirical semivariogram in Fig.
-\ref{fig:merc} suggests that mercury concentration is exhibits spatial
-dependence. Lastly we note that the realized mean mercury concentration
-in the 986 NLA lakes is 103.2 ng / g.
+visualize spatial dependence. It quantifies the mean of the halved
+squared differences (semivariance) among all pairs of mercury
+concentrations at different distances apart. When a process has spatial
+covariance (exhibits spatial dependence), the mean semivariance tends to
+be smaller at small distances and larger at large distances. The
+empirical semivariogram in Fig. \ref{fig:merc} suggests that mercury
+concentration exhibits spatial dependence. Lastly we note that the true
+mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
 
@@
 covariance was used. Table \ref{tab:appliedtab} shows the results from
-these analyses. For all four sampling-analysis combinations, the true
-realized mean mercury concentration is within the bounds of the 95\%
-confidence (design-based) or 95\% prediction (model-based) intervals.
-Though we should not generalize these results to other samples from
-these data, we do note a couple of patterns. The design-based IRS
-analysis shows the largest standard error: a likely reason is that this
-is the only approach that does not incorporate any spatial locations.
-Additionally, both analyses using GRTS sampling have lower standard
-errors than both analyses using IRS sampling.
+these analyses. Though we should not generalize these results to other
+samples from these data, we do mention a few findings. First, IRS-Design
+has the largest standard error. Second, compared to IRS-Design and
+IRS-Model, GRTS-Design and GRTS-Model are much closer to the true mean
+mercury concentration (have bias closer to zero) and have much lower
+standard errors (more precise intervals). Third, GRTS-Model has the
+least amount of bias and the lowest standard error (most precise
+interval). Finally, we note that for all sampling-analysis combinations,
+the true mean mercury concentration (103.2 ng / g) is within the bounds
+of the combination's 95\% interval.
 
@@
 \centering
-\begin{tabular}{lrrrr}
+\begin{tabular}{lrrrrr}
   \hline
-Approach & Est/Pred & SE & 95\% LB & 95\% UB \\ 
+Approach & True Mean & Est/Pred & SE & 95\% LB & 95\% UB \\ 
   \hline
-IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\ 
-  IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\ 
-  GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\ 
-  GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\ 
+IRS-Design & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\ 
+  IRS-Model & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\ 
+  GRTS-Design & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\ 
+  GRTS-Model & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\ 
    \hline
 \end{tabular}
-\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.
-The true mean concentration of all 986 lakes in the NLA data is 103.2 ng / g.}
+\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.}
 \end{table}
@@
 The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms. The design-based approach
-incorporates randomness through sampling to estimate population
-parameters. The model-based approach incorporates randomness through
-distributional assumptions to predict realized values of a stochastic
-process. Though these approaches have often been compared in the
-literature from theoretical and analytical perspectives, our
-contribution lies in studying them in a spatial context while
-implementing spatially balanced sampling and the local neighborhood
-variance estimator (in the design-based approach). Aside from the
+fundamentally different paradigms. The design-based approach relies on
+random sampling to estimate population parameters. The model-based
+approach relies on distributional assumptions to predict realized values
+of a stochastic process. Though the model-based approach does not rely
+on random sampling, it can still be beneficial as a way to guard against
+preferential sampling. The design-based and model-based approaches have
+often been compared in the literature from theoretical and analytical
+perspectives, but our contribution lies in studying them in a spatial
+context while implementing spatially balanced sampling and the
+design-based, local neighborhood variance estimator. Aside from the
 theoretical differences described, a few analytical findings from the
-simulation study are particularly notable. First, the sampling decision
-(IRS vs GRTS) is most important when using a design-based analysis.
-Though GRTS-Model still outperformed IRS-Model, the model-based analysis
-mitigated most of the inefficiencies that result from the IRS samples
-lacking spatial balance. Second, independent of the analysis approach,
-we found no reason to prefer IRS over GRTS for sampling spatial data --
-GRTS-Design and GRTS-Model generally performed at least as well as their
-IRS counterparts when there was no spatial covariance and noticeably
-better than their IRS counterparts when there was spatial covariance.
-Third, as the strength of spatial covariance increases, the gap in
-rMS(P)E between IRS-Design and the other sampling-analysis combinations
-also increases. Fourth and finally, when the response was normal,
-interval coverage for all sampling-analysis combinations was very close
-to 95\% for all sample sizes; when the response was lognormal, interval
-coverage for all sampling and analysis was between 90\% and 95\% and
-closest to 95\% when \(n = 200\).
+simulation study are particularly notable. First, independent of the
+analysis approach, we found no reason to prefer IRS over GRTS when
+sampling spatial data -- GRTS-Design and GRTS-Model generally had
+similar rMS(P)E as their IRS counterparts when there was no spatial
+covariance and lower rMS(P)E than their IRS counterparts when there was
+spatial covariance. Second, the sampling decision (IRS vs GRTS) is most
+important when using a design-based analysis. Though GRTS-Model still
+had lower rMS(P)E than IRS-Model, the model-based analysis mitigated
+most of the rMS(P)E inefficiencies that result from the IRS samples
+lacking spatial balance. Third, as the strength of spatial covariance
+increases, the gap in rMS(P)E and MStdE between IRS-Design and the other
+sampling-analysis combinations also increases, likely because IRS-Design
+is the only combination that ignores spatial locations in sampling and
+analysis. Fourth and finally, when the response was normal, interval
+coverage for all sampling-analysis combinations was usually close to
+95\% for all sample sizes; when the response was lognormal, interval
+coverage for all sampling-analysis combinations was usually between 90\%
+and 95\% and closest to 95\% when \(n = 200\).
 

### 2021-12-22 inst/manuscript/supporting.Rmd
more minor updates to incorporate mean standard error addition

@@
 
-Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
+Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), mean standard errors, root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
 ```{r}
@@
                   Response = resptype,
-                  `DE Prop` = psill_ratio,
+                  PDE = psill_ratio,
                   n = n,
@@
                   `rMS(P)E` = rmspe,
-                  Coverage = coverage) %>%
+                  MStdE = mse,
+                  Coverage = coverage
+                  ) %>%
     mutate(
@@
      ),
-     across(c(`DE Prop`, n), as.character)
+     across(c(PDE, n), as.character)
     )
@@
   digits = 4,
-  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
+  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
    type = "latex",
   latex.environments = "center",
-  align = c("llllrrrrr")
+  align = c("llllrrrrrr")
 )
-print(sim_table_out, include.rownames = FALSE, comment = FALSE, scalebox = "0.85")
+print(sim_table_out, include.rownames = FALSE, comment = FALSE, scalebox = "0.8")
 }

### 2021-12-22 inst/manuscript/supporting.tex
update simulations to include standard errors in summaries
more minor updates to incorporate mean standard error addition

--- 2021-12-20 inst/manuscript/supporting.tex
@@
 \end{tabular}
-\caption{\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2.} 
+\caption{\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the total variance was 2.}
 \end{table}
 
-Next we present a summary table for each of the 36 simulation scenarios.
+Next we present tables summarizing mean estimation bias (design-based)
+or mean prediction bias (model-based), mean standard errors,
+root-mean-squared error (design-based) or root-mean-squared-prediction
+error (model-based), and 95\% confidence interval coverage
+(design-based) or 95\% prediction interval coverage (model-based) for
+all 36 simulation scenarios.
 
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0 & 50 & -0.0023 & 0.1968 & 0.9440 \\ 
-  IRS-Model & Gridded & Normal & 0 & 50 & -0.0029 & 0.1988 & 0.9400 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0011 & 0.1946 & 0.9110 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0006 & 0.1949 & 0.9300 \\ 
+IRS-Design & Gridded & Normal & 0 & 50 & -0.0023 & 0.1968 & 0.1936 & 0.9440 \\ 
+  IRS-Model & Gridded & Normal & 0 & 50 & -0.0029 & 0.1988 & 0.1917 & 0.9400 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0011 & 0.1946 & 0.1703 & 0.9110 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0006 & 0.1949 & 0.1875 & 0.9300 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 1.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 50 & -0.0036 & 0.2007 & 0.9080 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0094 & 0.2071 & 0.9045 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 50 & -0.0007 & 0.1962 & 0.8690 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0022 & 0.1969 & 0.8945 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 50 & -0.0036 & 0.2007 & 0.1855 & 0.9080 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0094 & 0.2071 & 0.1833 & 0.9045 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 50 & -0.0007 & 0.1962 & 0.1629 & 0.8690 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0022 & 0.1969 & 0.1787 & 0.8945 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 2.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0 & 50 & 0.0019 & 0.1861 & 0.9505 \\ 
-  IRS-Model & Random & Normal & 0 & 50 & 0.0014 & 0.1892 & 0.9445 \\ 
-  GRTS-Design & Random & Normal & 0 & 50 & -0.0040 & 0.1955 & 0.9090 \\ 
-  GRTS-Model & Random & Normal & 0 & 50 & -0.0040 & 0.1965 & 0.9315 \\ 
+IRS-Design & Random & Normal & 0 & 50 & 0.0019 & 0.1861 & 0.1935 & 0.9505 \\ 
+  IRS-Model & Random & Normal & 0 & 50 & 0.0014 & 0.1892 & 0.1912 & 0.9445 \\ 
+  GRTS-Design & Random & Normal & 0 & 50 & -0.0040 & 0.1955 & 0.1709 & 0.9090 \\ 
+  GRTS-Model & Random & Normal & 0 & 50 & -0.0040 & 0.1965 & 0.1879 & 0.9315 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 3.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0 & 50 & 0.0032 & 0.1843 & 0.9205 \\ 
-  IRS-Model & Random & Lognormal & 0 & 50 & -0.0039 & 0.1945 & 0.9105 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0056 & 0.1944 & 0.8870 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0078 & 0.1967 & 0.9075 \\ 
+IRS-Design & Random & Lognormal & 0 & 50 & 0.0032 & 0.1843 & 0.1833 & 0.9205 \\ 
+  IRS-Model & Random & Lognormal & 0 & 50 & -0.0039 & 0.1945 & 0.1808 & 0.9105 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0056 & 0.1944 & 0.1634 & 0.8870 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0078 & 0.1967 & 0.1789 & 0.9075 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 4.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 50 & 0.0025 & 0.1762 & 0.9470 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 50 & 0.0007 & 0.1655 & 0.9305 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0004 & 0.1524 & 0.9115 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 50 & -0.0003 & 0.1499 & 0.9320 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 50 & 0.0025 & 0.1762 & 0.1748 & 0.9470 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 50 & 0.0007 & 0.1655 & 0.1567 & 0.9305 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0004 & 0.1524 & 0.1357 & 0.9115 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 50 & -0.0003 & 0.1499 & 0.1453 & 0.9320 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 5.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0004 & 0.1863 & 0.9140 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0076 & 0.1834 & 0.9035 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0016 & 0.1612 & 0.8810 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0017 & 0.1606 & 0.8940 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0004 & 0.1863 & 0.1627 & 0.9140 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0076 & 0.1834 & 0.1495 & 0.9035 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0016 & 0.1612 & 0.1303 & 0.8810 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0017 & 0.1606 & 0.1385 & 0.8940 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 6.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.5 & 50 & 0.0037 & 0.1657 & 0.9590 \\ 
-  IRS-Model & Random & Normal & 0.5 & 50 & 0.0027 & 0.1546 & 0.9495 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 50 & -0.0034 & 0.1511 & 0.9170 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 50 & -0.0037 & 0.1504 & 0.9305 \\ 
+IRS-Design & Random & Normal & 0.5 & 50 & 0.0037 & 0.1657 & 0.1737 & 0.9590 \\ 
+  IRS-Model & Random & Normal & 0.5 & 50 & 0.0027 & 0.1546 & 0.1559 & 0.9495 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 50 & -0.0034 & 0.1511 & 0.1355 & 0.9170 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 50 & -0.0037 & 0.1504 & 0.1447 & 0.9305 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 7.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0023 & 0.1637 & 0.9245 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0055 & 0.1622 & 0.9125 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 50 & -0.0057 & 0.1570 & 0.9000 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0079 & 0.1567 & 0.9100 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0023 & 0.1637 & 0.1574 & 0.9245 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0055 & 0.1622 & 0.1443 & 0.9125 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 50 & -0.0057 & 0.1570 & 0.1287 & 0.9000 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0079 & 0.1567 & 0.1362 & 0.9100 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 8.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 50 & 0.0053 & 0.1579 & 0.9470 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 50 & 0.0026 & 0.1165 & 0.9315 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 50 & 0.0013 & 0.1074 & 0.9220 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 50 & -0.0007 & 0.0949 & 0.9430 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 50 & 0.0053 & 0.1579 & 0.1570 & 0.9470 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 50 & 0.0026 & 0.1165 & 0.1094 & 0.9315 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 50 & 0.0013 & 0.1074 & 0.0996 & 0.9220 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 50 & -0.0007 & 0.0949 & 0.0938 & 0.9430 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 9.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1731 & 0.9220 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0020 & 0.1325 & 0.9135 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1183 & 0.9065 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0002 & 0.1090 & 0.9120 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1731 & 0.1412 & 0.9220 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0020 & 0.1325 & 0.1051 & 0.9135 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1183 & 0.0954 & 0.9065 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0002 & 0.1090 & 0.0909 & 0.9120 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 10.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.9 & 50 & 0.0062 & 0.1520 & 0.9525 \\ 
-  IRS-Model & Random & Normal & 0.9 & 50 & 0.0037 & 0.1074 & 0.9525 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0029 & 0.1038 & 0.9340 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0027 & 0.0940 & 0.9360 \\ 
+IRS-Design & Random & Normal & 0.9 & 50 & 0.0062 & 0.1520 & 0.1545 & 0.9525 \\ 
+  IRS-Model & Random & Normal & 0.9 & 50 & 0.0037 & 0.1074 & 0.1078 & 0.9525 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0029 & 0.1038 & 0.0984 & 0.9340 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0027 & 0.0940 & 0.0921 & 0.9360 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 11.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 50 & 0.0053 & 0.1542 & 0.9325 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0012 & 0.1160 & 0.9115 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0033 & 0.1150 & 0.9160 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0031 & 0.1014 & 0.9235 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 50 & 0.0053 & 0.1542 & 0.1351 & 0.9325 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0012 & 0.1160 & 0.0997 & 0.9115 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0033 & 0.1150 & 0.0923 & 0.9160 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0031 & 0.1014 & 0.0872 & 0.9235 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 12.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0 & 100 & -0.0029 & 0.1322 & 0.9430 \\ 
-  IRS-Model & Gridded & Normal & 0 & 100 & -0.0033 & 0.1332 & 0.9410 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0041 & 0.1302 & 0.9245 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0038 & 0.1304 & 0.9420 \\ 
+IRS-Design & Gridded & Normal & 0 & 100 & -0.0029 & 0.1322 & 0.1333 & 0.9430 \\ 
+  IRS-Model & Gridded & Normal & 0 & 100 & -0.0033 & 0.1332 & 0.1326 & 0.9410 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0041 & 0.1302 & 0.1208 & 0.9245 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0038 & 0.1304 & 0.1309 & 0.9420 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 13.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 100 & -0.0052 & 0.1334 & 0.9335 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 100 & -0.0072 & 0.1350 & 0.9300 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0042 & 0.1303 & 0.9035 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0034 & 0.1307 & 0.9265 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 100 & -0.0052 & 0.1334 & 0.1309 & 0.9335 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 100 & -0.0072 & 0.1350 & 0.1301 & 0.9300 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0042 & 0.1303 & 0.1170 & 0.9035 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0034 & 0.1307 & 0.1266 & 0.9265 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 14.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0 & 100 & -0.0066 & 0.1366 & 0.9405 \\ 
-  IRS-Model & Random & Normal & 0 & 100 & -0.0067 & 0.1370 & 0.9385 \\ 
-  GRTS-Design & Random & Normal & 0 & 100 & 0.0028 & 0.1364 & 0.9180 \\ 
-  GRTS-Model & Random & Normal & 0 & 100 & 0.0029 & 0.1363 & 0.9345 \\ 
+IRS-Design & Random & Normal & 0 & 100 & -0.0066 & 0.1366 & 0.1329 & 0.9405 \\ 
+  IRS-Model & Random & Normal & 0 & 100 & -0.0067 & 0.1370 & 0.1322 & 0.9385 \\ 
+  GRTS-Design & Random & Normal & 0 & 100 & 0.0028 & 0.1364 & 0.1210 & 0.9180 \\ 
+  GRTS-Model & Random & Normal & 0 & 100 & 0.0029 & 0.1363 & 0.1310 & 0.9345 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 15.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0 & 100 & -0.0051 & 0.1348 & 0.9280 \\ 
-  IRS-Model & Random & Lognormal & 0 & 100 & -0.0075 & 0.1372 & 0.9230 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0002 & 0.1375 & 0.8840 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 100 & -0.0003 & 0.1377 & 0.9090 \\ 
+IRS-Design & Random & Lognormal & 0 & 100 & -0.0051 & 0.1348 & 0.1298 & 0.9280 \\ 
+  IRS-Model & Random & Lognormal & 0 & 100 & -0.0075 & 0.1372 & 0.1290 & 0.9230 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0002 & 0.1375 & 0.1181 & 0.8840 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 100 & -0.0003 & 0.1377 & 0.1275 & 0.9090 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 16.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0016 & 0.1177 & 0.9540 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0023 & 0.1072 & 0.9470 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 100 & 0.0051 & 0.0998 & 0.9300 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 100 & 0.0039 & 0.0982 & 0.9470 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0016 & 0.1177 & 0.1201 & 0.9540 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0023 & 0.1072 & 0.1057 & 0.9470 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 100 & 0.0051 & 0.0998 & 0.0935 & 0.9300 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 100 & 0.0039 & 0.0982 & 0.0990 & 0.9470 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 17.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 100 & -0.0021 & 0.1211 & 0.9445 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 100 & -0.0058 & 0.1153 & 0.9295 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0041 & 0.1090 & 0.9085 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0019 & 0.1090 & 0.9180 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 100 & -0.0021 & 0.1211 & 0.1137 & 0.9445 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 100 & -0.0058 & 0.1153 & 0.1032 & 0.9295 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0041 & 0.1090 & 0.0910 & 0.9085 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0019 & 0.1090 & 0.0959 & 0.9180 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 18.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.5 & 100 & -0.0064 & 0.1222 & 0.9440 \\ 
-  IRS-Model & Random & Normal & 0.5 & 100 & -0.0049 & 0.1073 & 0.9440 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0013 & 0.1041 & 0.9155 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0014 & 0.1020 & 0.9400 \\ 
+IRS-Design & Random & Normal & 0.5 & 100 & -0.0064 & 0.1222 & 0.1190 & 0.9440 \\ 
+  IRS-Model & Random & Normal & 0.5 & 100 & -0.0049 & 0.1073 & 0.1051 & 0.9440 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0013 & 0.1041 & 0.0932 & 0.9155 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0014 & 0.1020 & 0.0986 & 0.9400 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 19.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0063 & 0.1204 & 0.9355 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0097 & 0.1150 & 0.9275 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 100 & 0.0003 & 0.1092 & 0.8960 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0004 & 0.1088 & 0.9120 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0063 & 0.1204 & 0.1108 & 0.9355 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0097 & 0.1150 & 0.1005 & 0.9275 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 100 & 0.0003 & 0.1092 & 0.0891 & 0.8960 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0004 & 0.1088 & 0.0937 & 0.9120 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 20.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 100 & -0.0007 & 0.1059 & 0.9605 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 100 & -0.0019 & 0.0700 & 0.9445 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0044 & 0.0655 & 0.9435 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 100 & 0.0030 & 0.0585 & 0.9440 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 100 & -0.0007 & 0.1059 & 0.1078 & 0.9605 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 100 & -0.0019 & 0.0700 & 0.0686 & 0.9445 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0044 & 0.0655 & 0.0637 & 0.9435 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 100 & 0.0030 & 0.0585 & 0.0594 & 0.9440 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 21.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 100 & -0.0011 & 0.1092 & 0.9540 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0033 & 0.0779 & 0.9330 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0035 & 0.0757 & 0.9190 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0017 & 0.0693 & 0.9190 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 100 & -0.0011 & 0.1092 & 0.0981 & 0.9540 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0033 & 0.0779 & 0.0674 & 0.9330 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0035 & 0.0757 & 0.0621 & 0.9190 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0017 & 0.0693 & 0.0585 & 0.9190 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 22.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.9 & 100 & -0.0048 & 0.1095 & 0.9430 \\ 
-  IRS-Model & Random & Normal & 0.9 & 100 & -0.0017 & 0.0671 & 0.9535 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0674 & 0.9265 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 100 & 0.0000 & 0.0595 & 0.9420 \\ 
+IRS-Design & Random & Normal & 0.9 & 100 & -0.0048 & 0.1095 & 0.1058 & 0.9430 \\ 
+  IRS-Model & Random & Normal & 0.9 & 100 & -0.0017 & 0.0671 & 0.0674 & 0.9535 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0674 & 0.0627 & 0.9265 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 100 & 0.0000 & 0.0595 & 0.0580 & 0.9420 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 23.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0045 & 0.1110 & 0.9375 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0052 & 0.0771 & 0.9310 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 100 & -0.0009 & 0.0732 & 0.9195 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0007 & 0.0661 & 0.9165 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0045 & 0.1110 & 0.0944 & 0.9375 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0052 & 0.0771 & 0.0645 & 0.9310 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 100 & -0.0009 & 0.0732 & 0.0596 & 0.9195 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0007 & 0.0661 & 0.0555 & 0.9165 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 24.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0 & 200 & 0.0013 & 0.0882 & 0.9550 \\ 
-  IRS-Model & Gridded & Normal & 0 & 200 & 0.0013 & 0.0886 & 0.9530 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9380 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9460 \\ 
+IRS-Design & Gridded & Normal & 0 & 200 & 0.0013 & 0.0882 & 0.0882 & 0.9550 \\ 
+  IRS-Model & Gridded & Normal & 0 & 200 & 0.0013 & 0.0886 & 0.0880 & 0.9530 \\ 
+  GRTS-Design & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.0854 & 0.9380 \\ 
+  GRTS-Model & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.0874 & 0.9460 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 25.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0001 & 0.0880 & 0.9350 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 200 & -0.0002 & 0.0883 & 0.9355 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0034 & 0.0898 & 0.9285 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0035 & 0.0900 & 0.9335 \\ 
+IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0001 & 0.0880 & 0.0869 & 0.9350 \\ 
+  IRS-Model & Gridded & Lognormal & 0 & 200 & -0.0002 & 0.0883 & 0.0866 & 0.9355 \\ 
+  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0034 & 0.0898 & 0.0843 & 0.9285 \\ 
+  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0035 & 0.0900 & 0.0862 & 0.9335 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 26.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0 & 200 & -0.0014 & 0.0893 & 0.9465 \\ 
-  IRS-Model & Random & Normal & 0 & 200 & -0.0015 & 0.0896 & 0.9465 \\ 
-  GRTS-Design & Random & Normal & 0 & 200 & 0.0007 & 0.0868 & 0.9460 \\ 
-  GRTS-Model & Random & Normal & 0 & 200 & 0.0007 & 0.0867 & 0.9490 \\ 
+IRS-Design & Random & Normal & 0 & 200 & -0.0014 & 0.0893 & 0.0882 & 0.9465 \\ 
+  IRS-Model & Random & Normal & 0 & 200 & -0.0015 & 0.0896 & 0.0880 & 0.9465 \\ 
+  GRTS-Design & Random & Normal & 0 & 200 & 0.0007 & 0.0868 & 0.0855 & 0.9460 \\ 
+  GRTS-Model & Random & Normal & 0 & 200 & 0.0007 & 0.0867 & 0.0873 & 0.9490 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 27.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0 & 200 & -0.0028 & 0.0891 & 0.9425 \\ 
-  IRS-Model & Random & Lognormal & 0 & 200 & -0.0033 & 0.0896 & 0.9395 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 200 & 0.0015 & 0.0860 & 0.9365 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 200 & 0.0012 & 0.0861 & 0.9415 \\ 
+IRS-Design & Random & Lognormal & 0 & 200 & -0.0028 & 0.0891 & 0.0872 & 0.9425 \\ 
+  IRS-Model & Random & Lognormal & 0 & 200 & -0.0033 & 0.0896 & 0.0870 & 0.9395 \\ 
+  GRTS-Design & Random & Lognormal & 0 & 200 & 0.0015 & 0.0860 & 0.0840 & 0.9365 \\ 
+  GRTS-Model & Random & Lognormal & 0 & 200 & 0.0012 & 0.0861 & 0.0857 & 0.9415 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 28.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 200 & 0.0022 & 0.0788 & 0.9535 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 200 & 0.0008 & 0.0678 & 0.9580 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 200 & -0.0024 & 0.0671 & 0.9335 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 200 & -0.0030 & 0.0661 & 0.9410 \\ 
+IRS-Design & Gridded & Normal & 0.5 & 200 & 0.0022 & 0.0788 & 0.0796 & 0.9535 \\ 
+  IRS-Model & Gridded & Normal & 0.5 & 200 & 0.0008 & 0.0678 & 0.0687 & 0.9580 \\ 
+  GRTS-Design & Gridded & Normal & 0.5 & 200 & -0.0024 & 0.0671 & 0.0646 & 0.9335 \\ 
+  GRTS-Model & Gridded & Normal & 0.5 & 200 & -0.0030 & 0.0661 & 0.0651 & 0.9410 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 29.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 200 & 0.0016 & 0.0816 & 0.9420 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0000 & 0.0739 & 0.9335 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0033 & 0.0744 & 0.9330 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.0742 & 0.9350 \\ 
+IRS-Design & Gridded & Lognormal & 0.5 & 200 & 0.0016 & 0.0816 & 0.0759 & 0.9420 \\ 
+  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0000 & 0.0739 & 0.0680 & 0.9335 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0033 & 0.0744 & 0.0646 & 0.9330 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.0742 & 0.0650 & 0.9350 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 30.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.5 & 200 & -0.0016 & 0.0790 & 0.9480 \\ 
-  IRS-Model & Random & Normal & 0.5 & 200 & -0.0010 & 0.0690 & 0.9475 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0002 & 0.0652 & 0.9455 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 200 & 0.0001 & 0.0640 & 0.9500 \\ 
+IRS-Design & Random & Normal & 0.5 & 200 & -0.0016 & 0.0790 & 0.0790 & 0.9480 \\ 
+  IRS-Model & Random & Normal & 0.5 & 200 & -0.0010 & 0.0690 & 0.0684 & 0.9475 \\ 
+  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0002 & 0.0652 & 0.0644 & 0.9455 \\ 
+  GRTS-Model & Random & Normal & 0.5 & 200 & 0.0001 & 0.0640 & 0.0647 & 0.9500 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 31.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.5 & 200 & -0.0027 & 0.0809 & 0.9494 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0037 & 0.0732 & 0.9454 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 200 & -0.0008 & 0.0671 & 0.9434 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0012 & 0.0662 & 0.9479 \\ 
+IRS-Design & Random & Lognormal & 0.5 & 200 & -0.0027 & 0.0809 & 0.0738 & 0.9494 \\ 
+  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0037 & 0.0732 & 0.0662 & 0.9454 \\ 
+  GRTS-Design & Random & Lognormal & 0.5 & 200 & -0.0008 & 0.0671 & 0.0622 & 0.9434 \\ 
+  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0012 & 0.0662 & 0.0625 & 0.9479 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 32.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0021 & 0.0704 & 0.9570 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 200 & 0.0005 & 0.0411 & 0.9590 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0005 & 0.0423 & 0.9370 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0012 & 0.0377 & 0.9450 \\ 
+IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0021 & 0.0704 & 0.0716 & 0.9570 \\ 
+  IRS-Model & Gridded & Normal & 0.9 & 200 & 0.0005 & 0.0411 & 0.0417 & 0.9590 \\ 
+  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0005 & 0.0423 & 0.0410 & 0.9370 \\ 
+  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0012 & 0.0377 & 0.0370 & 0.9450 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 33.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0014 & 0.0744 & 0.9520 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0002 & 0.0480 & 0.9355 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0014 & 0.0499 & 0.9380 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0022 & 0.0459 & 0.9380 \\ 
+IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0014 & 0.0744 & 0.0655 & 0.9520 \\ 
+  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0002 & 0.0480 & 0.0418 & 0.9355 \\ 
+  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0014 & 0.0499 & 0.0409 & 0.9380 \\ 
+  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0022 & 0.0459 & 0.0373 & 0.9380 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 34.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0.9 & 200 & -0.0009 & 0.0725 & 0.9470 \\ 
-  IRS-Model & Random & Normal & 0.9 & 200 & -0.0010 & 0.0409 & 0.9450 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 200 & -0.0005 & 0.0405 & 0.9490 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0001 & 0.0354 & 0.9510 \\ 
+IRS-Design & Random & Normal & 0.9 & 200 & -0.0009 & 0.0725 & 0.0702 & 0.9470 \\ 
+  IRS-Model & Random & Normal & 0.9 & 200 & -0.0010 & 0.0409 & 0.0409 & 0.9450 \\ 
+  GRTS-Design & Random & Normal & 0.9 & 200 & -0.0005 & 0.0405 & 0.0401 & 0.9490 \\ 
+  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0001 & 0.0354 & 0.0358 & 0.9510 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 35.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllllrrr}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0.9 & 200 & -0.0006 & 0.0750 & 0.9393 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 200 & -0.0025 & 0.0461 & 0.9418 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 200 & -0.0013 & 0.0452 & 0.9348 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 200 & -0.0012 & 0.0397 & 0.9418 \\ 
+IRS-Design & Random & Lognormal & 0.9 & 200 & -0.0006 & 0.0750 & 0.0626 & 0.9393 \\ 
+  IRS-Model & Random & Lognormal & 0.9 & 200 & -0.0025 & 0.0461 & 0.0397 & 0.9418 \\ 
+  GRTS-Design & Random & Lognormal & 0.9 & 200 & -0.0013 & 0.0452 & 0.0387 & 0.9348 \\ 
+  GRTS-Model & Random & Lognormal & 0.9 & 200 & -0.0012 & 0.0397 & 0.0347 & 0.9418 \\ 
    \hline
@@
 }
-\caption{Approach, layout (Layout), response type (Response) roportion of dependent error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) for a simulation scenario.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 36.}
 \end{table}

### 2021-12-22 inst/scripts/figures/simulation.R

@@ -24,11 +24,11 @@ combo_data <- combo_data %>%
   mutate(resptype = fct_relevel(resptype, c("normal", "lognormal"))) %>%
   group_by(sim) %>%
   mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
-  mutate(designirsmil = if_else(approach == "Design IRS", true = med_ci_len, false = NA_real_)) %>%
+  mutate(designirsmse = if_else(approach == "Design IRS", true = mse, false = NA_real_)) %>%
   fill(designirsrmspe, .direction = "downup") %>%
-  fill(designirsmil, .direction = "downup") %>%
+  fill(designirsmse, .direction = "downup") %>%
   mutate(rel_efficiency = rmspe / designirsrmspe) %>%
-  mutate(mil_efficiency = med_ci_len / designirsmil) %>%
+  mutate(mse_efficiency = mse / designirsmse) %>%
   ungroup() %>%
   mutate(psill_ratio = 1 - nugget_ratio) %>%
   mutate(n_factor = factor(n))
@@ -70,7 +70,7 @@ colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
 psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
 mse_eff <- combo_data %>%
-  ggplot(aes(x = n_factor, y = mil_efficiency, colour = approach)) +
+  ggplot(aes(x = n_factor, y = mse_efficiency, colour = approach)) +
   facet_grid(
     psill_ratio ~ resptype,
     labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
@@ -85,7 +85,7 @@ mse_eff <- combo_data %>%
     panel.grid.minor = element_blank(),
   )
 
-### rmspe_eff
+### mse_eff
 if (write_out) {
   ggsave(
     plot = mse_eff,

### 2021-12-22 inst/manuscript/manuscript.Rmd
minor updates before submission

@@
 In the simulations, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
-In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval length and coverage.
+In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
   3.
@@
 
-There are two general approaches for using data to make frequentist statistical inferences about a population: design-based and model-based.
-When data cannot be collected for all units in a population (i.e., population units), data are collected on a subset of the population units.
-This subset of population units is called a sample.
-In the design-based approach, inferences about the underlying population are informed via a probabilistic process that randomly assigns some population units to be in the sample.
-Alternatively, in the model-based approach, inferences are made from specific assumptions about the underlying process generating the data.
+When data cannot be collected for all units in a population (i.e., population units), data are collected on a subset of the population units -- this subset is called a sample.
+There are two general approaches for using samples to make frequentist statistical inferences about a population: design-based and model-based.
+In the design-based approach, inference relies on randomly assigning some population units to be in the sample (e.g., random sampling).
+Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying stochastic process that generated the sample.
 Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation].
@@
 @cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design and model-based principles (see e.g., @sterba2009alternative and @cicchitelli2012model, and see @chan2020bayesian for a Bayesian approach).
+In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design-based and model-based principles (see e.g., @sterba2009alternative and @cicchitelli2012model, and see @chan2020bayesian for a Bayesian approach).
 
 <!-- Paragraph about the contribution of the manuscript and outlining the rest -->
-Certainly comparisons between design-based and model-based approaches to spatial data have been studied.
-But no numerical comparison has been made between design-based approaches that incorporate spatial information and model-based approaches.
-In this manuscript, we compare design-based approaches that incorporate spatial information to model-based approaches for finite population spatial data.
+Certainly comparisons between design-based and model-based approaches have been studied in spatial contexts.
+But no numerical comparison has been made between design-based approaches that incorporate spatial locations into sampling and analysis and model-based approaches.
+In this manuscript, we compare design-based approaches that incorporate spatial locations into sampling and analysis to model-based approaches for finite population spatial data.
 A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
-Though we focus on finite populations, these comparisons generalize to infinite populations as well.
+Though here we focus on finite populations, the comparisons we discuss generalize to infinite populations as well.
 An infinite population contains an infinite number of population units; an example is locations within a single lake.
@@
 A sampling design assigns a positive probability of inclusion (inclusion probability) in the sample to each population unit.
-These inclusion probabilities are later used to analyze data.
+These inclusion probabilities are later used to estimate population parameters.
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
@@
 Sample means, for example, are asymptotically normal (Gaussian) by the Central Limit Theorem (under some assumptions).
-If we repeatedly select samples from the population, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean.
+If we repeatedly select samples from the population, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true fixed population mean.
 @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.
 
-The model-based approach assumes the data are a random realization of a data-generating stochastic process.
-Randomness is incorporated through distributional assumptions on this process.
+The model-based approach assumes the sample is a random realization of a data-generating stochastic process.
+Randomness is formally incorporated through distributional assumptions on this process.
 Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling.
@@
 To guard against preferential sampling, model-based approaches often still implement some form of random sampling.
-When model-based approaches implement random sampling, the inclusion probabilities are ignored when analyzing the data (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the data).
+When model-based approaches implement random sampling, the inclusion probabilities are ignored when analyzing the sample (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the sample).
 
@@
 For example, suppose the realized mean of all population units is the value of interest.
-Instead of \emph{estimating} a fixed, unknown mean, we are \emph{predicting} the value of the mean, a random variable.
+Instead of a fixed, unknown mean, we are the value of the mean, a random variable.
 Prediction intervals are then derived using assumptions of the data-generating stochastic process.
-If we repeatedly generate response values from the same data-generating stochastic process and select samples, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
+If we repeatedly generate response values from the same process and select samples, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
 @cressie1993statistics and @schabenberger2017statistical provide thorough reviews of model-based approaches for spatial data.
@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ and yield narrower confidence intervals compared to variance estimators that ignore spatial locations.
+@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ and yield more precise (narrower) confidence intervals compared to variance estimators that ignore spatial locations.
 
@@
 , $s_N$ that can be measured at the $N$ population units.
-Suppose we want to use a sample to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights (e.g, the population mean is represented by a weights vector whose elements all equal one).
+Suppose we want to use a sample to predict some linear function of the response variable, $f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}$, where $\mathbf{b}^\prime$ is a $1 \times N$ vector of weights (e.g, the population mean is represented by a weights vector whose elements all equal $1 / N$).
 Denoting quantities that are part of the sampled population units with a subscript \emph{s} and quantities that are part of the unsampled population units with a subscript \emph{u}, let
@@
 While details of the derivation are in @verhoef2008spatial, we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
+Distributional assumptions are used, however, when constructing prediction intervals.
 
@@
 We used a simulation study to investigate performance of four sampling-analysis combinations.
-The first sampling-analysis combination is IRS-Design.
-In IRS-Design, samples are selected with the Independent Random Sampling (IRS) algorithm.
-The IRS algorithm ignores the spatial locations of the population units, thus IRS samples are not spatially balanced.
-In IRS-Design, samples are analyzed using the design-based approach via the Horvitz-Thompson mean estimator and an IRS variance estimator that does not incorporate the spatial locations of the units in the sample.
-The second sampling-analysis combination is IRS-Model, where samples are selected with the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
-The third sampling-analysis combination is GRTS-Design, where samples are selected with the GRTS algorithm and analyzed using the design-based approach via the Horvitz-Thompson mean estimator and the local neighborhood variance estimator (which does incorporate the spatial locations of the units in the sample).
-The fourth and final sampling-analysis combination is GRTS-Model, where samples are selected with the GRTS algorithm and analyzed using the model-based approach via REML estimation.
+The first sampling-analysis combination was IRS-Design.
+In IRS-Design, samples were selected with the Independent Random Sampling (IRS) algorithm.
+The IRS algorithm ignores the spatial locations of the population units, thus the IRS samples were not spatially balanced.
+In IRS-Design, samples were analyzed using the design-based approach via the Horvitz-Thompson mean estimator and an IRS variance estimator that ignored the spatial locations of the units in the sample.
+The second sampling-analysis combination was IRS-Model, where samples were selected with the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The third sampling-analysis combination was GRTS-Design, where samples were selected with the GRTS algorithm and analyzed using the design-based approach via the Horvitz-Thompson mean estimator and the local neighborhood variance estimator (which does incorporate the spatial locations of the units in the sample).
+The fourth and final sampling-analysis combination was GRTS-Model, where samples were selected with the GRTS algorithm and analyzed using the model-based approach via REML estimation.
 These sampling-analysis combinations are also provided in Table \ref{tab:designanalysis}.
@@
 The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
-The range was always $\sqrt{2} / 3$, which means that the correlation in the dependent random error decayed to nearly zero at the largest possible distance between two population units in the domain.
+The range was always $\sqrt{2} / 3$, chosen so that the correlation in the dependent random error decayed to nearly zero at $\sqrt{2}$, the largest possible distance between two population units in the domain.
 For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
@@
 When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
-So when there is no spatial covariance, using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design.
+\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E and using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design.
 When there is spatial covariance (Fig.
@@
 \end{equation*}
-Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms (has a much lower MStdE) IRS-Design by a large margin.
+Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms IRS-Design by a noticeable margin.
 These fact that the rMS(P)E and MStdE findings are similar is not particularly surprising because the mean bias for all sampling-analysis combinations was nearly zero, thus rMS(P)E is driven by the standard error of the estimators (design-based) or predictors (model-based).
 We do note that between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was no spatial covariance or a medium amount of spatial covariance (Fig.
-\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows) and GRTS-Model had lower MStdE when there was a high amount of spatial covariance (Fig.
+\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows), and GRTS-Model had lower MStdE when there was a high amount of spatial covariance (Fig.
 \ref{fig:mse_eff}, "Prop DE: 0.9" row).
@@
 Table \ref{tab:appliedtab} shows the results from these analyses.
-Though we should not generalize these results to other samples from these data, we do mention a few findings.
+Though we should not generalize these results to other samples from this population, we do mention a few findings.
 First, IRS-Design has the largest standard error.
@@
 The design-based approach also more naturally handles binary data, free from the more complicated logistic regression framework commonly used to analyze binary data in a model-based approach.
-The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and response variable.
+The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
 The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the stochastic process of study.
 Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
-Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed sites.
-Model-based approaches can also compute site-by-site predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
+Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed units.
+Model-based approaches can also compute unit-by-unit predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
 In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.
@@
 
-All authors conceived the ideas; All authors designed methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
+All authors conceived the ideas; All authors designed the methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
 
@@
 
-This manuscript has a supplementary R package that contains all of the data and code used in its creation. The supplementary R package is hosted on GitHub. Instructions for download at available at
+This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data and code used in its creation.
+The supplementary \textbf{\textsf{R}} package is hosted on GitHub.
+Instructions for download at available at
 

### 2021-12-22 inst/manuscript/manuscript.tex
minor updates before submission

@@
     The design-based and model-based approaches to frequentist statistical
-    inference lie on fundamentally different foundations. In the
-    design-based approach, inference depends on random sampling. In the
-    model-based approach, inference depends on distributional assumptions.
+    inference rest on fundamentally different foundations. In the
+    design-based approach, inference relies on random sampling. In the
+    model-based approach, inference relies on distributional assumptions.
     We compare the approaches for finite population spatial data.
@@
     performance is measured using statistics like bias, squared error, and
-    interval length and coverage.
+    interval coverage.
   \item
@@
 
-There are two general approaches for using data to make frequentist
-statistical inferences about a population: design-based and model-based.
 When data cannot be collected for all units in a population (i.e.,
 population units), data are collected on a subset of the population
-units. This subset of population units is called a sample. In the
-design-based approach, inferences about the underlying population are
-informed via a probabilistic process that randomly assigns some
-population units to be in the sample. Alternatively, in the model-based
-approach, inferences are made from specific assumptions about the
-underlying process generating the data. Each paradigm has a deep
-historical context (Sterba, 2009) and its own set of benefits and
-drawbacks (Hansen et al., 1983).
+units -- this subset is called a sample. There are two general
+approaches for using samples to make frequentist statistical inferences
+about a population: design-based and model-based. In the design-based
+approach, inference relies on randomly assigning some population units
+to be in the sample (e.g., random sampling). Alternatively, in the
+model-based approach, inference relies on distributional assumptions
+about the underlying stochastic process that generated the sample. Each
+paradigm has a deep historical context (Sterba, 2009) and its own set of
+benefits and drawbacks (Hansen et al., 1983).
 
@@
 has been substantial research and development into estimators that use
-both design and model-based principles (see e.g., Sterba (2009) and
-Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020) for
-a Bayesian approach).
-
-Certainly comparisons between design-based and model-based approaches to
-spatial data have been studied. But no numerical comparison has been
-made between design-based approaches that incorporate spatial
-information and model-based approaches. In this manuscript, we compare
-design-based approaches that incorporate spatial information to
-model-based approaches for finite population spatial data. A finite
-population contains a finite number of population units (we assume the
-finite number is known); an example is lakes (treated as a whole with
-the lake centroid representing location) in the contiguous United
-States. Though we focus on finite populations, these comparisons
-generalize to infinite populations as well. An infinite population
-contains an infinite number of population units; an example is locations
-within a single lake.
+both design-based and model-based principles (see e.g., Sterba (2009)
+and Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020)
+for a Bayesian approach).
+
+Certainly comparisons between design-based and model-based approaches
+have been studied in spatial contexts. But no numerical comparison has
+been made between design-based approaches that incorporate spatial
+locations into sampling and analysis and model-based approaches. In this
+manuscript, we compare design-based approaches that incorporate spatial
+locations into sampling and analysis to model-based approaches for
+finite population spatial data. A finite population contains a finite
+number of population units (we assume the finite number is known); an
+example is lakes (treated as a whole with the lake centroid representing
+location) in the contiguous United States. Though here we focus on
+finite populations, the comparisons we discuss generalize to infinite
+populations as well. An infinite population contains an infinite number
+of population units; an example is locations within a single lake.
 
@@
 inclusion (inclusion probability) in the sample to each population unit.
-These inclusion probabilities are later used to analyze data. Some
-examples of commonly used sampling designs include simple random
-sampling, stratified random sampling, and cluster sampling.
+These inclusion probabilities are later used to estimate population
+parameters. Some examples of commonly used sampling designs include
+simple random sampling, stratified random sampling, and cluster
+sampling.
 
@@
 then 95\% of all 95\% confidence intervals constructed from a procedure
-with appropriate coverage will contain the true, fixed mean. Särndal et
-al. (2003) and Lohr (2009) provide thorough reviews of the design-based
-approach.
-
-The model-based approach assumes the data are a random realization of a
-data-generating stochastic process. Randomness is incorporated through
-distributional assumptions on this process. Strictly speaking,
+with appropriate coverage will contain the true fixed population mean.
+Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
+design-based approach.
+
+The model-based approach assumes the sample is a random realization of a
+data-generating stochastic process. Randomness is formally incorporated
+through distributional assumptions on this process. Strictly speaking,
 randomness need not be incorporated through random sampling, though
@@
 implement random sampling, the inclusion probabilities are ignored when
-analyzing the data (in contrast to the design-based approach, which
-relies on these inclusion probabilities to analyze the data).
+analyzing the sample (in contrast to the design-based approach, which
+relies on these inclusion probabilities to analyze the sample).
 
@@
 predict a realized variable, or value. For example, suppose the realized
-mean of all population units is the value of interest. Instead of
-\emph{estimating} a fixed, unknown mean, we are \emph{predicting} the
-value of the mean, a random variable. Prediction intervals are then
-derived using assumptions of the data-generating stochastic process. If
-we repeatedly generate response values from the same data-generating
-stochastic process and select samples, then 95\% of all 95\% prediction
-intervals constructed from a procedure with appropriate coverage will
-contain their respective realized means. Cressie (1993) and
-Schabenberger and Gotway (2017) provide thorough reviews of model-based
-approaches for spatial data. In Fig. \ref{fig:fig1}, we provide a visual
-comparison of the design-based and model-based approaches (Ver Hoef
-(2002) and Brus (2021) provide similar figures).
+mean of all population units is the value of interest. Instead of a
+fixed, unknown mean, we are the value of the mean, a random variable.
+Prediction intervals are then derived using assumptions of the
+data-generating stochastic process. If we repeatedly generate response
+values from the same process and select samples, then 95\% of all 95\%
+prediction intervals constructed from a procedure with appropriate
+coverage will contain their respective realized means. Cressie (1993)
+and Schabenberger and Gotway (2017) provide thorough reviews of
+model-based approaches for spatial data. In Fig. \ref{fig:fig1}, we
+provide a visual comparison of the design-based and model-based
+approaches (Ver Hoef (2002) and Brus (2021) provide similar figures).
 
@@
 variance estimator tends to reduce the estimated variance of
-\(\hat{\tau}\) and yield narrower confidence intervals compared to
-variance estimators that ignore spatial locations.
+\(\hat{\tau}\) and yield more precise (narrower) confidence intervals
+compared to variance estimators that ignore spatial locations.
 
@@
 population mean is represented by a weights vector whose elements all
-equal one). Denoting quantities that are part of the sampled population
-units with a subscript \emph{s} and quantities that are part of the
-unsampled population units with a subscript \emph{u}, let
+equal \(1 / N\)). Denoting quantities that are part of the sampled
+population units with a subscript \emph{s} and quantities that are part
+of the unsampled population units with a subscript \emph{u}, let
 
@@
 meaning that they do not rely on any distributional assumptions.
+Distributional assumptions are used, however, when constructing
+prediction intervals.
 
@@
 sampling-analysis combinations. The first sampling-analysis combination
-is IRS-Design. In IRS-Design, samples are selected with the Independent
-Random Sampling (IRS) algorithm. The IRS algorithm ignores the spatial
-locations of the population units, thus IRS samples are not spatially
-balanced. In IRS-Design, samples are analyzed using the design-based
-approach via the Horvitz-Thompson mean estimator and an IRS variance
-estimator that does not incorporate the spatial locations of the units
-in the sample. The second sampling-analysis combination is IRS-Model,
-where samples are selected with the IRS algorithm and analyzed using the
-model-based approach via Restricted Maximum Likelihood (REML) estimation
-(Harville, 1977; Patterson and Thompson, 1971; Wolfinger et al., 1994).
-The third sampling-analysis combination is GRTS-Design, where samples
-are selected with the GRTS algorithm and analyzed using the design-based
-approach via the Horvitz-Thompson mean estimator and the local
-neighborhood variance estimator (which does incorporate the spatial
-locations of the units in the sample). The fourth and final
-sampling-analysis combination is GRTS-Model, where samples are selected
-with the GRTS algorithm and analyzed using the model-based approach via
-REML estimation. These sampling-analysis combinations are also provided
-in Table \ref{tab:designanalysis}. Lastly we note that for both the IRS
-and GRTS samples, equal inclusion probabilities were assumed for all
-population units. When IRS assumes equal inclusion probabilities for all
-population units, the algorithm is equivalent to simple random sampling
-(SRS).
+was IRS-Design. In IRS-Design, samples were selected with the
+Independent Random Sampling (IRS) algorithm. The IRS algorithm ignores
+the spatial locations of the population units, thus the IRS samples were
+not spatially balanced. In IRS-Design, samples were analyzed using the
+design-based approach via the Horvitz-Thompson mean estimator and an IRS
+variance estimator that ignored the spatial locations of the units in
+the sample. The second sampling-analysis combination was IRS-Model,
+where samples were selected with the IRS algorithm and analyzed using
+the model-based approach via Restricted Maximum Likelihood (REML)
+estimation (Harville, 1977; Patterson and Thompson, 1971; Wolfinger et
+al., 1994). The third sampling-analysis combination was GRTS-Design,
+where samples were selected with the GRTS algorithm and analyzed using
+the design-based approach via the Horvitz-Thompson mean estimator and
+the local neighborhood variance estimator (which does incorporate the
+spatial locations of the units in the sample). The fourth and final
+sampling-analysis combination was GRTS-Model, where samples were
+selected with the GRTS algorithm and analyzed using the model-based
+approach via REML estimation. These sampling-analysis combinations are
+also provided in Table \ref{tab:designanalysis}. Lastly we note that for
+both the IRS and GRTS samples, equal inclusion probabilities were
+assumed for all population units. When IRS assumes equal inclusion
+probabilities for all population units, the algorithm is equivalent to
+simple random sampling (SRS).
 
@@
 \(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
-\(\sqrt{2} / 3\), which means that the correlation in the dependent
-random error decayed to nearly zero at the largest possible distance
-between two population units in the domain. For the lognormal response
-type, the response was first simulated using the same approach as for
-the normal response type, except that the total variance was 0.6931
-instead of 2. The response was then exponentiated, yielding a lognormal
-random variable whose total variance was 2. The lognormal responses were
-used to evaluate performance of the sampling-analysis approaches for
-data that were skewed (i.e., not normal).
+\(\sqrt{2} / 3\), chosen so that the correlation in the dependent random
+error decayed to nearly zero at \(\sqrt{2}\), the largest possible
+distance between two population units in the domain. For the lognormal
+response type, the response was first simulated using the same approach
+as for the normal response type, except that the total variance was
+0.6931 instead of 2. The response was then exponentiated, yielding a
+lognormal random variable whose total variance was 2. The lognormal
+responses were used to evaluate performance of the sampling-analysis
+approaches for data that were skewed (i.e., not normal).
 
@@
 \ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
-combinations have approximately equal rMS(P)E. So when there is no
-spatial covariance, using the GRTS algorithm or a model-based analysis
-does not result in much, if any, loss in efficiency compared to
-IRS-Design. When there is spatial covariance (Fig. \ref{fig:rmspe_eff},
-``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to have
-the lowest rMS(P)E, followed by GRTS-Design, IRS-Model, and finally
-IRS-Design, though the difference in relative rMS(P)E among GRTS-Model,
-GRTS-Design, and IRS-Model is relatively small. As the strength of
-spatial covariance increases, the gap in rMS(P)E between IRS-Design and
-the other sampling-analysis combinations widens. Finally we note that
-when there is spatial covariance, IRS-Model has a much lower rMS(P)E
-than IRS-Design, suggesting that the poor design properties of IRS are
-largely mitigated by the model-based analysis. These rMS(P)E conclusions
-are similar to those observed in the grid location layout, so we omit a
-grid location layout figure here. Tables for rMS(P)E in all 36
-simulation scenarios are provided in the supporting information.
+combinations have approximately equal rMS(P)E and using the GRTS
+algorithm or a model-based analysis does not result in much, if any,
+loss in efficiency compared to IRS-Design. When there is spatial
+covariance (Fig. \ref{fig:rmspe_eff}, ``Prop DE: 0.5'' and ``Prop DE:
+0.9'' rows), GRTS-Model tends to have the lowest rMS(P)E, followed by
+GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in
+relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is
+relatively small. As the strength of spatial covariance increases, the
+gap in rMS(P)E between IRS-Design and the other sampling-analysis
+combinations widens. Finally we note that when there is spatial
+covariance, IRS-Model has a much lower rMS(P)E than IRS-Design,
+suggesting that the poor design properties of IRS are largely mitigated
+by the model-based analysis. These rMS(P)E conclusions are similar to
+those observed in the grid location layout, so we omit a grid location
+layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
+are provided in the supporting information.
 
@@
 the other sampling-analysis combinations widens; and IRS-Model
-outperforms (has a much lower MStdE) IRS-Design by a large margin. These
-fact that the rMS(P)E and MStdE findings are similar is not particularly
-surprising because the mean bias for all sampling-analysis combinations
-was nearly zero, thus rMS(P)E is driven by the standard error of the
-estimators (design-based) or predictors (model-based). We do note that
-between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when
-there was no spatial covariance or a medium amount of spatial covariance
-(Fig. \ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows) and
+outperforms IRS-Design by a noticeable margin. These fact that the
+rMS(P)E and MStdE findings are similar is not particularly surprising
+because the mean bias for all sampling-analysis combinations was nearly
+zero, thus rMS(P)E is driven by the standard error of the estimators
+(design-based) or predictors (model-based). We do note that between
+GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was
+no spatial covariance or a medium amount of spatial covariance (Fig.
+\ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows), and
 GRTS-Model had lower MStdE when there was a high amount of spatial
@@
 these analyses. Though we should not generalize these results to other
-samples from these data, we do mention a few findings. First, IRS-Design
-has the largest standard error. Second, compared to IRS-Design and
-IRS-Model, GRTS-Design and GRTS-Model are much closer to the true mean
-mercury concentration (have bias closer to zero) and have much lower
-standard errors (more precise intervals). Third, GRTS-Model has the
-least amount of bias and the lowest standard error (most precise
+samples from this population, we do mention a few findings. First,
+IRS-Design has the largest standard error. Second, compared to
+IRS-Design and IRS-Model, GRTS-Design and GRTS-Model are much closer to
+the true mean mercury concentration (have bias closer to zero) and have
+much lower standard errors (more precise intervals). Third, GRTS-Model
+has the least amount of bias and the lowest standard error (most precise
 interval). Finally, we note that for all sampling-analysis combinations,
@@
 on random sampling, it can still be beneficial as a way to guard against
-preferential sampling. The design-based and model-based approaches have
-often been compared in the literature from theoretical and analytical
-perspectives, but our contribution lies in studying them in a spatial
-context while implementing spatially balanced sampling and the
+preferential sampling. While the design-based and model-based approaches
+have often been compared in the literature from theoretical and
+analytical perspectives, our contribution lies in studying them in a
+spatial context while implementing spatially balanced sampling and the
 design-based, local neighborhood variance estimator. Aside from the
@@
 naturally quantify the relationship between covariates (predictor
-variables) and response variable. The model-based approach also yields
-estimated spatial covariance parameters, which help better understand
-the dependence structure in the stochastic process of study. Model
-selection is also possible using model-based approaches and criteria
-such as cross validation, likelihood ratio tests, or AIC (Akaike, 1974).
-Model-based approaches are capable of more efficient small-area
-estimation than design-based approaches by leveraging distributional
-assumptions in areas with few observed sites. Model-based approaches can
-also compute site-by-site predictions at unobserved locations and use
-them to construct informative visualizations like smoothed maps. In
-short, when deciding whether the design-based or model-based approach is
-more appropriate to implement, the benefits and drawbacks of each
-approach should be considered alongside the particular goals of the
-study.
+variables) and the response variable. The model-based approach also
+yields estimated spatial covariance parameters, which help better
+understand the dependence structure in the stochastic process of study.
+Model selection is also possible using model-based approaches and
+criteria such as cross validation, likelihood ratio tests, or AIC
+(Akaike, 1974). Model-based approaches are capable of more efficient
+small-area estimation than design-based approaches by leveraging
+distributional assumptions in areas with few observed units. Model-based
+approaches can also compute unit-by-unit predictions at unobserved
+locations and use them to construct informative visualizations like
+smoothed maps. In short, when deciding whether the design-based or
+model-based approach is more appropriate to implement, the benefits and
+drawbacks of each approach should be considered alongside the particular
+goals of the study.
 
@@
 
-All authors conceived the ideas; All authors designed methodology; MD
-and MH performed the simulations and analyzed the data; MD and MH led
+All authors conceived the ideas; All authors designed the methodology;
+MD and MH performed the simulations and analyzed the data; MD and MH led
 the writing of the manuscript; All authors contributed critically to the
@@
 
-This manuscript has a supplementary R package that contains all of the
-data and code used in its creation. The supplementary R package is
-hosted on GitHub. Instructions for download at available at
+This manuscript has a supplementary \textbf{\textsf{R}} package that
+contains all of the data and code used in its creation. The
+supplementary \textbf{\textsf{R}} package is hosted on GitHub.
+Instructions for download at available at
 

### 2021-12-22 inst/manuscript/supporting.Rmd
minor updates before submission

@@
 ---
-title: A comparison of design-based and model-based approaches for finite population spatial data -- Supporting Information.
+title: A comparison of design-based and model-based approaches for finite population spatial data -- Supporting information.
 author:
@@
 
-# Simulation Study Tables {#sec:simtabs}
+# Simulation study tables {#sec:simtabs}
 
@@
 
-Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), mean standard errors, root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
+Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), mean standard errors, and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
 ```{r}

### 2021-12-22 inst/manuscript/supporting.tex
minor updates before submission

@@
             pdfauthor={},
-            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supporting Information.},
+            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supporting information.},
             colorlinks=false,
@@
   \title{A comparison of design-based and model-based approaches for finite
-population spatial data -- Supporting Information.}
+population spatial data -- Supporting information.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 \hypertarget{sec:simtabs}{%
-\section{Simulation Study Tables}\label{sec:simtabs}}
+\section{Simulation study tables}\label{sec:simtabs}}
 
@@
 Next we present tables summarizing mean estimation bias (design-based)
-or mean prediction bias (model-based), mean standard errors,
-root-mean-squared error (design-based) or root-mean-squared-prediction
-error (model-based), and 95\% confidence interval coverage
-(design-based) or 95\% prediction interval coverage (model-based) for
-all 36 simulation scenarios.
+or mean prediction bias (model-based), root-mean-squared error
+(design-based) or root-mean-squared-prediction error (model-based), mean
+standard errors, and 95\% confidence interval coverage (design-based) or
+95\% prediction interval coverage (model-based) for all 36 simulation
+scenarios.
 

### 2021-12-22 inst/manuscript/manuscript.tex
added running header

@@
 
-\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption} \usepackage{setspace} \doublespacing
+\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption} \usepackage{setspace} \usepackage{fancyhdr}
+
+\pagestyle{fancy} \fancyhf{} \chead{Spatial design-based vs model-based}
+\doublespacing
 

### 2022-04-05 inst/manuscript/manuscript.Rmd
added running header
data analysis updates

@@
   \usepackage{setspace}
+  \usepackage{fancyhdr}
+  \pagestyle{fancy}
+  \fancyhf{}
+  \chead{Spatial design-based vs model-based}
   \doublespacing
@@
 
+## New Application
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.98\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_map.jpeg}
+  \caption*{}
+  \label{fig:bmmi_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_hist.jpeg}
+  \caption*{}
+  \label{fig:bmmi_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_sv_plot.jpeg}
+  \caption*{}
+  \label{fig:bmmi_sv_plot}
+\end{subfigure}
+\caption{BMMI visualizations for all 986 lakes in the NLA data.
+A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
+\label{fig:merc}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:rmspe_eff}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_mse_eff.jpeg}
+  \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:mse_eff}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
+  \caption{Interval coverage in the data study for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line represents 95\% coverage.}
+  \label{fig:figconf}
+\end{figure}
+
 # Discussion {#sec:discussion}

### 2022-04-05 inst/manuscript/manuscript.tex
data analysis updates

@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
-and Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{Saint Lawrence University Department of Mathematics,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+      \cortext[1]{Corresponding Author: Michael Dumelle
+(Dumelle.Michael@epa.gov)}
 
@@
   \item
-    The design-based and model-based approaches to frequentist statistical
-    inference rest on fundamentally different foundations. In the
-    design-based approach, inference relies on random sampling. In the
-    model-based approach, inference relies on distributional assumptions.
-    We compare the approaches for finite population spatial data.
+    The design-based and model-based approaches to frequentist
+    statistical inference rest on fundamentally different foundations.
+    In the design-based approach, inference relies on random sampling.
+    In the model-based approach, inference relies on distributional
+    assumptions. We compare the approaches for finite population spatial
+    data.
   \item
@@
     analysis of real mercury concentration data. In the simulations, a
-    variety of sample sizes, location layouts, dependence structures, and
-    response types are considered. In the simualations and real data
+    variety of sample sizes, location layouts, dependence structures,
+    and response types are considered. In the simualations and real data
     analysis, the population mean is the parameter of interest and
-    performance is measured using statistics like bias, squared error, and
-    interval coverage.
+    performance is measured using statistics like bias, squared error,
+    and interval coverage.
   \item
-    When studying the simulations and mercury concentration data, we found
-    that regardless of the strength of spatial dependence in the data,
-    sampling plans that incorporate spatial locations (spatially balanced
-    samples) generally outperform sampling plans that ignore spatial
-    locations (non-spatially balanced samples). We also found that
-    model-based approaches tend to outperform design-based approaches,
-    even when the data are skewed (and by consequence, the model-based
-    distributional assumptions violated). The performance gap between
-    these approaches is small when spatially balanced samples are used but
-    large when non-spatially balanced samples are used. This suggests that
-    the sampling choice (whether to select a sample that is spatially
-    balanced) is most important when performing design-based inference.
+    When studying the simulations and mercury concentration data, we
+    found that regardless of the strength of spatial dependence in the
+    data, sampling plans that incorporate spatial locations (spatially
+    balanced samples) generally outperform sampling plans that ignore
+    spatial locations (non-spatially balanced samples). We also found
+    that model-based approaches tend to outperform design-based
+    approaches, even when the data are skewed (and by consequence, the
+    model-based distributional assumptions violated). The performance
+    gap between these approaches is small when spatially balanced
+    samples are used but large when non-spatially balanced samples are
+    used. This suggests that the sampling choice (whether to select a
+    sample that is spatially balanced) is most important when performing
+    design-based inference.
   \item
@@
 comparisons between design-based and model-based approaches for spatial
-data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang
-et al., 2012). Cooper (2006) reviews the two approaches in an ecological
-context before introducing a ``model-assisted'' variance estimator that
-combines aspects from each approach. In addition to Cooper (2006), there
-has been substantial research and development into estimators that use
-both design-based and model-based principles (see e.g., Sterba (2009)
-and Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020)
-for a Bayesian approach).
+data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002; Ver Hoef,
+2008; Wang et al., 2012). Cooper (2006) reviews the two approaches in an
+ecological context before introducing a ``model-assisted'' variance
+estimator that combines aspects from each approach. In addition to
+Cooper (2006), there has been substantial research and development into
+estimators that use both design-based and model-based principles (see
+e.g., Sterba (2009) and Cicchitelli and Montanari (2012), and see
+Chan-Golston et al. (2020) for a Bayesian approach).
 
@@
 
+\hypertarget{new-application}{%
+\subsection{New Application}\label{new-application}}
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.98\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_map.jpeg}
+  \caption*{}
+  \label{fig:bmmi_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_hist.jpeg}
+  \caption*{}
+  \label{fig:bmmi_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/bmmi_sv_plot.jpeg}
+  \caption*{}
+  \label{fig:bmmi_sv_plot}
+\end{subfigure}
+\caption{BMMI visualizations for all 986 lakes in the NLA data.
+A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
+\label{fig:merc}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
+  \caption{Relative rMS(P)E in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:rmspe_eff}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_mse_eff.jpeg}
+  \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:mse_eff}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
+  \caption{Interval coverage in the data study for the four sampling-analysis combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line represents 95\% coverage.}
+  \label{fig:figconf}
+\end{figure}
+
 \hypertarget{sec:discussion}{%
@@
 Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
-model to nearest neighbor (k-nn) methods for forestry applications. PlOS
+model to nearest neighbor (k-NN) methods for forestry applications. PlOS
 ONE 8, e59129.

### 2022-04-05 inst/scripts/figures/data_analysis.R

@@ -1,0 +1,125 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(here)
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## set base size
+base_size <- 10
+
+## read in files and create data
+files <- list.files(path = here("inst", "output", "data_summary"), pattern = "*.csv", full.names = TRUE)
+
+combo_data <- purrr::map_df(files,
+                            ~read_csv(.x) %>% mutate(filename = .x))
+
+combo_data <- combo_data %>%
+  mutate(sim = interaction(n, var)) %>% ## , gridded
+  mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
+                                            "Design GRTS", "Model GRTS"))) %>%
+  mutate(resptype = fct_relevel(var, c("MMI_BENT_NLA12", "TOTALHG_RESULT"))) %>%
+  group_by(sim) %>%
+  mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
+  mutate(designirsmse = if_else(approach == "Design IRS", true = mse, false = NA_real_)) %>%
+  fill(designirsrmspe, .direction = "downup") %>%
+  fill(designirsmse, .direction = "downup") %>%
+  mutate(rel_efficiency = rmspe / designirsrmspe) %>%
+  mutate(mse_efficiency = mse / designirsmse) %>%
+  ungroup() %>%
+  mutate(n_factor = factor(n))
+
+# figure 2
+colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+resptype_labs <- c(MMI_BENT_NLA12 = "Response: MMI_BENT_NLA12", TOTALHG_RESULT = "Response: TOTALHG_RESULT")
+rmspe_eff <- combo_data %>%
+  ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
+  facet_grid(
+    rows = vars(var)
+    # labeller = labeller(var = resptype_labs)
+  ) +
+  geom_jitter(width = 0.24, size = 2) +
+  scale_colour_manual(values = colour_scale) +
+  theme_bw(base_size = base_size) +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
[…]

### 2022-04-05 inst/scripts/figures/data_exploratory.R

@@ -129,3 +129,126 @@ if (write_out) {
     height = 4.39
   )
 }
+
+data(nla_df)
+
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(MMI_BENT_NLA12))) %>%
+  ungroup() %>%
+  summarise(maxnonmiss = max(nonmiss))
+
+## bmmi never measured twice at the same site.
+nla_df %>%
+  group_by(SITE_ID) %>%
+  summarise(nonmiss = sum(!is.na(MMI_BENT_NLA12))) %>%
+  summarise(sumnomiss = sum(nonmiss == 0))
+
+## some sites never have HG measured. These will be dropped as well as the duplicate sites.
+## in duplicate sites, fill in the value for RDis_IX
+nla_nomiss <- nla_df %>%
+  group_by(SITE_ID) %>%
+  fill(RDis_IX, .direction = "downup") %>%
+  ungroup() %>%
+  filter(!is.na(MMI_BENT_NLA12))
+
+nla_nomiss_both <- nla_nomiss
+## check
+nla_nomiss_both %>%
+  group_by(SITE_ID) %>% count() %>%
+  ungroup() %>%
+  summarise(not1 = sum(n != 1))
+
+# average
+nla_nomiss_both <- nla_nomiss_both %>%
+  group_by(SITE_ID) %>%
+  summarize(MMI_BENT_NLA12 = mean(MMI_BENT_NLA12), INDEX_LON_DD = mean(INDEX_LON_DD), INDEX_LAT_DD = INDEX_LAT_DD) %>%
+  ungroup()
+
+## transform coordinates
+coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
+                     lat = nla_nomiss_both$INDEX_LAT_DD,
+                     lon = nla_nomiss_both$INDEX_LON_DD)
+nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
+nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
+
+## figure 4a
+bmmi_map <- nla_nomiss_both %>%
+  ggplot(aes(x = xcoords, y = ycoords)) +
+  geom_point(aes(colour = MMI_BENT_NLA12)) +
+  scale_colour_viridis_c() +
+  labs(x = "", y = "", colour = "BMMI") +
+  coord_quickmap() +
+  theme_bw(base_size = 10) +
+  theme(
[…]

### 2022-04-06 inst/scripts/data_runs.R
seed and zmmi update

@@ -6,15 +6,18 @@ library(DvMsp)
 write_out <- TRUE
 
 n <- c(50, 100, 200)
-vars <- c("TOTALHG_RESULT", "MMI_BENT_NLA12")
+vars <- c("TOTALHG_RESULT", "MMI_ZOOP_NLA6")
 parm_df <- expand_grid(n = n, var = vars)
+
+n_trials <- 2000
+# set overall seed
+set.seed(2)
+parm_df$seed <- lapply(seq_len(NROW(parm_df)), function(x) sample(1e9, size = n_trials))
 
 data("nla_df")
 
 ## loop through each row of parm_df (could be done with purrr::pmap() instead)
 for (i in 1:nrow(parm_df)) {
-  n_trials <- 2000
-  seed <- 1:n_trials
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
   clusterEvalQ(cluster, library(DvMsp)) # export DvMsp to cluster
@@ -25,7 +28,7 @@ for (i in 1:nrow(parm_df)) {
   parm_df_sim <- parm_df %>% slice(i) ## grab row for ith parameter combo
   sim_output <- parLapply(
     cluster, # the cluster
-    seed, # some seeds
+    unlist(parm_df_sim %>% pull(seed)), # some seeds
     safely(data_trial), # see if any errors
     data = nla_df,
     ## pull each parameter from the single row data frame

### 2022-04-06 inst/scripts/figures/data_exploratory.R

@@ -134,14 +134,14 @@ data(nla_df)
 
 nla_df %>%
   group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(MMI_BENT_NLA12))) %>%
+  summarise(nonmiss = sum(!is.na(MMI_ZOOP_NLA6))) %>%
   ungroup() %>%
   summarise(maxnonmiss = max(nonmiss))
 
-## bmmi never measured twice at the same site.
+## zmmi measured twice at the same site?
 nla_df %>%
   group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(MMI_BENT_NLA12))) %>%
+  summarise(nonmiss = sum(!is.na(MMI_ZOOP_NLA6))) %>%
   summarise(sumnomiss = sum(nonmiss == 0))
 
 ## some sites never have HG measured. These will be dropped as well as the duplicate sites.
@@ -150,7 +150,7 @@ nla_nomiss <- nla_df %>%
   group_by(SITE_ID) %>%
   fill(RDis_IX, .direction = "downup") %>%
   ungroup() %>%
-  filter(!is.na(MMI_BENT_NLA12))
+  filter(!is.na(MMI_ZOOP_NLA6))
 
 nla_nomiss_both <- nla_nomiss
 ## check
@@ -162,7 +162,7 @@ nla_nomiss_both %>%
 # average
 nla_nomiss_both <- nla_nomiss_both %>%
   group_by(SITE_ID) %>%
-  summarize(MMI_BENT_NLA12 = mean(MMI_BENT_NLA12), INDEX_LON_DD = mean(INDEX_LON_DD), INDEX_LAT_DD = INDEX_LAT_DD) %>%
+  summarize(MMI_ZOOP_NLA6 = mean(MMI_ZOOP_NLA6), INDEX_LON_DD = mean(INDEX_LON_DD), INDEX_LAT_DD = INDEX_LAT_DD) %>%
   ungroup()
 
 ## transform coordinates
@@ -173,11 +173,11 @@ nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
 nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
 
 ## figure 4a
-bmmi_map <- nla_nomiss_both %>%
+zmmi_map <- nla_nomiss_both %>%
   ggplot(aes(x = xcoords, y = ycoords)) +
[…]

### 2022-04-06 inst/scripts/sim_runs.R

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(parallel)
+library(DvMsp)
 
 # write results?
 write_out <- TRUE
@@ -25,12 +26,13 @@ parm_df <- expand_grid(N, n, psill_ratio,
          psill = total_var * psill_ratio,
          nugget = total_var * nugget_ratio)
 
-library(DvMsp)
+n_trials <- 2000
+# set overall seed
+set.seed(2)
+parm_df$seed <- lapply(seq_len(NROW(parm_df)), function(x) sample(1e9, size = n_trials))
 
 ## loop through each row of parm_df (could be done with purrr::pmap() instead)
 for (i in 1:nrow(parm_df)) {
-  n_trials <- 2000
-  seed <- 1:n_trials
   n_cluster <- detectCores() # find cores (48 on mine)
   cluster <- makeCluster(n_cluster) # make cluster
   clusterEvalQ(cluster, library(DvMsp)) # export DvMsp to cluster
@@ -41,7 +43,7 @@ for (i in 1:nrow(parm_df)) {
   parm_df_sim <- parm_df %>% slice(i) ## grab row for ith parameter combo
   sim_output <- parLapply(
     cluster, # the cluster
-    seed, # some seeds
+    unlist(parm_df_sim %>% pull(seed)), # some seeds
     safely(sim_trial), # see if any errors
     ## pull each parameter from the single row data frame
     N = parm_df_sim %>% pull(N), # pop size

### 2022-04-06 inst/scripts/figures/data_analysis.R
figure updates

@@ -20,7 +20,7 @@ combo_data <- combo_data %>%
   mutate(sim = interaction(n, var)) %>% ## , gridded
   mutate(approach = fct_relevel(approach, c("Design IRS", "Model IRS",
                                             "Design GRTS", "Model GRTS"))) %>%
-  mutate(resptype = fct_relevel(var, c("MMI_BENT_NLA12", "TOTALHG_RESULT"))) %>%
+  mutate(resptype = fct_relevel(var, c("MMI_ZOOP_NLA6", "TOTALHG_RESULT"))) %>%
   group_by(sim) %>%
   mutate(designirsrmspe = if_else(approach == "Design IRS", true = rmspe, false = NA_real_)) %>%
   mutate(designirsmse = if_else(approach == "Design IRS", true = mse, false = NA_real_)) %>%
@@ -33,12 +33,12 @@ combo_data <- combo_data %>%
 
 # figure 2
 colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-resptype_labs <- c(MMI_BENT_NLA12 = "Response: MMI_BENT_NLA12", TOTALHG_RESULT = "Response: TOTALHG_RESULT")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
 rmspe_eff <- combo_data %>%
   ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
   facet_grid(
-    rows = vars(var)
-    # labeller = labeller(var = resptype_labs)
+    rows = vars(var),
+    labeller = labeller(var = resptype_labs)
   ) +
   geom_jitter(width = 0.24, size = 2) +
   scale_colour_manual(values = colour_scale) +
@@ -64,12 +64,12 @@ if (write_out) {
 
 # figure 3
 colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-resptype_labs <- c(MMI_BENT_NLA12 = "Response: MMI_BENT_NLA12", TOTALHG_RESULT = "Response: TOTALHG_RESULT")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
 mse_eff <- combo_data %>%
[…]

### 2022-04-08 inst/manuscript/manuscript.Rmd
seed and zmmi update
some review updates

@@
 Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation].
+In this manuscript, we compare the design-based and model-based approaches for finite population spatial data.
 
@@
 
-Though the design-based and model-based approaches apply to statistical inference in a broad sense, we focus on comparing these approaches for spatial data.
-We define spatial data as data that incorporates the specific locations of the population units into either the sampling or estimation process.
-@de1990model give an early comparison of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
+Spatial data are data that incorporate the locations of the population units into either the sampling or estimation process.
+@de1990model and @brus1993design give early comparisons of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
 Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @wang2012review; @brus2021statistical].
@@
 Certainly comparisons between design-based and model-based approaches have been studied in spatial contexts.
-But no numerical comparison has been made between design-based approaches that incorporate spatial locations into sampling and analysis and model-based approaches.
-In this manuscript, we compare design-based approaches that incorporate spatial locations into sampling and analysis to model-based approaches for finite population spatial data.
+Our contribution is comparing design-based approaches that incorporate spatial locations into sampling and analysis to model-based approaches.
+Though the broad comparisons we draw between design-based and model-based approaches generalize to finite and infinite populations, we focus on finite populations.
 A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
-Though here we focus on finite populations, the comparisons we discuss generalize to infinite populations as well.
 An infinite population contains an infinite number of population units; an example is locations within a single lake.
@@
 After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it has several attractive properties: the GRTS algorithm accommodates finite and infinite sampling frames, equal, unequal, and proportional (to size) inclusion probabilities, legacy (historical) sampling [@foster2017spatially], a minimum distance between units in a sample, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
+In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
+
+The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
+First the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells.
+Each level-one is randomly assigned an level-one address of 0, 1, 2, or 3.
+The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$.
+Within each level-one cell, the inclusion probability for each population unit is summed, and if any of these sums exceed one, a second level of cells is added.
+Then each level-one cell is split into four distinct, equally sized squares called level-two cells.
+Each level-two cell is randomly assigned a level-two address of 0, 1, 2, or 3.
+The set of level-two cells is denoted by $\mathcal{A}_2$ and defined as $\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}$.
+The inclusion probabilities within each level-two cell are summed, and if any of these sums exceed one, a third level of cells is added.
+This process continues for $k$ steps, until all level-$k$ cells have inclusion probability sums no larger than one.
+Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$.
+
+After determining $\mathcal{A}_k$, it is placed into hierarchical order.
+Hierarchical order is a numeric order that first sorts $\mathcal{A}_k$ by the level-one addresses from smallest to largest, then sorts $\mathcal{A}_k$ by the level-two addresses from smallest to largest, and so on.
+For example, $\mathcal{A}_2$ in hierarchical order is the set $\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}$.
+Because hierarchical ordering sorts by level-one cells, then level-two cells, and so on, population units that have similar hierarchical addresses tend to be nearby one another in space.Next each population unit is mapped to a one-dimensional line in hierarchical order where each population unit's inclusion probability equals its line-length.
+If a level-$k$ cell has multiple population units in it, they are randomly placed within the cell's respective line segment.
+A uniform random variable is then simulated in $[0, 1]$ and a systematic sample is selected on the line, yielding $n$ sample points for a sample size $n$.
+Each element in this systematic sample falls on some population unit's line segment, and thus that population unit is selected in the sample.
+
 The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
@@
 
-The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2021spsurvey].
+The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey].
 FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
@@
   \centering
-  \includegraphics[width = 1\linewidth]{figures/bmmi_map.jpeg}
+  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
   \caption*{}
-  \label{fig:bmmi_map}
+  \label{fig:zmmi_map}
 \end{subfigure} \\
@@
   \centering
-  \includegraphics[width = 1\linewidth]{figures/bmmi_hist.jpeg}
+  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
   \caption*{}
-  \label{fig:bmmi_hist}
+  \label{fig:zmmi_hist}
 \end{subfigure}
@@
   \centering
-  \includegraphics[width = 1\linewidth]{figures/bmmi_sv_plot.jpeg}
+  \includegraphics[width = 1\linewidth]{figures/zmmi_sv_plot.jpeg}
   \caption*{}
-  \label{fig:bmmi_sv_plot}
+  \label{fig:zmmi_sv_plot}
 \end{subfigure}
-\caption{BMMI visualizations for all 986 lakes in the NLA data.
+\caption{zmmi visualizations for all 986 lakes in the NLA data.
 A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
-\label{fig:merc}
+\label{fig:zmmi}
 \end{figure}
@@
   \caption{Relative rMS(P)E in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:rmspe_eff}
+  \label{fig:data_rmspe_eff}
 \end{figure}
@@
   \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:mse_eff}
+  \label{fig:data_mse_eff}
 \end{figure}
@@
 The solid, black line represents 95\% coverage.}
-  \label{fig:figconf}
+  \label{fig:data_figconf}
 \end{figure}

### 2022-04-08 inst/manuscript/manuscript.tex
seed and zmmi update
elsevier style update
some review updates

@@
   \linenumbers % turns line numbering on
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
 \usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 
-% Pandoc citation processing
-
-% Pandoc header
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \begin{document}
+
 \begin{frontmatter}
@@
 paradigm has a deep historical context (Sterba, 2009) and its own set of
-benefits and drawbacks (Hansen et al., 1983).
-
-Though the design-based and model-based approaches apply to statistical
-inference in a broad sense, we focus on comparing these approaches for
-spatial data. We define spatial data as data that incorporates the
-specific locations of the population units into either the sampling or
-estimation process. De Gruijter and Ter Braak (1990) give an early
-comparison of design-based and model-based approaches for spatial data,
-quashing the belief that design-based approaches could not be used for
-spatially correlated data. Since then, there have been several general
-comparisons between design-based and model-based approaches for spatial
-data (Brus and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang
-et al., 2012). Cooper (2006) reviews the two approaches in an ecological
-context before introducing a ``model-assisted'' variance estimator that
-combines aspects from each approach. In addition to Cooper (2006), there
-has been substantial research and development into estimators that use
-both design-based and model-based principles (see e.g., Sterba (2009)
-and Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020)
-for a Bayesian approach).
+benefits and drawbacks (Hansen et al., 1983). In this manuscript, we
+compare the design-based and model-based approaches for finite
+population spatial data.
+
+Spatial data are data that incorporate the locations of the population
+units into either the sampling or estimation process. De Gruijter and
+Ter Braak (1990) and Brus and DeGruijter (1993) give early comparisons
+of design-based and model-based approaches for spatial data, quashing
+the belief that design-based approaches could not be used for spatially
+correlated data. Since then, there have been several general comparisons
+between design-based and model-based approaches for spatial data (Brus
+and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang et al.,
+2012). Cooper (2006) reviews the two approaches in an ecological context
+before introducing a ``model-assisted'' variance estimator that combines
+aspects from each approach. In addition to Cooper (2006), there has been
+substantial research and development into estimators that use both
+design-based and model-based principles (see e.g., Sterba (2009) and
+Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020) for
+a Bayesian approach).
 
 Certainly comparisons between design-based and model-based approaches
-have been studied in spatial contexts. But no numerical comparison has
-been made between design-based approaches that incorporate spatial
-locations into sampling and analysis and model-based approaches. In this
-manuscript, we compare design-based approaches that incorporate spatial
-locations into sampling and analysis to model-based approaches for
-finite population spatial data. A finite population contains a finite
-number of population units (we assume the finite number is known); an
-example is lakes (treated as a whole with the lake centroid representing
-location) in the contiguous United States. Though here we focus on
-finite populations, the comparisons we discuss generalize to infinite
-populations as well. An infinite population contains an infinite number
-of population units; an example is locations within a single lake.
+have been studied in spatial contexts. Our contribution is comparing
+design-based approaches that incorporate spatial locations into sampling
+and analysis to model-based approaches. Though the broad comparisons we
+draw between design-based and model-based approaches generalize to
+finite and infinite populations, we focus on finite populations. A
+finite population contains a finite number of population units (we
+assume the finite number is known); an example is lakes (treated as a
+whole with the lake centroid representing location) in the contiguous
+United States. An infinite population contains an infinite number of
+population units; an example is locations within a single lake.
 
@@
 manuscript, we select spatially balanced samples using the Generalized
-Random Tessellation Stratified (GRTS) algorithm because it has several
-attractive properties: the GRTS algorithm accommodates finite and
-infinite sampling frames, equal, unequal, and proportional (to size)
-inclusion probabilities, legacy (historical) sampling (Foster et al.,
-2017), a minimum distance between units in a sample, and replacement
-units (replacement units are population units that can be sampled when a
-population unit originally selected can no longer be sampled). The GRTS
-algorithm selects samples by utilizing a particular mapping between
-two-dimensional and one-dimensional space that preserves proximity
-relationships. Via this mapping, units in two-dimensional space are
-partitioned using a hierarchical address. This hierarchical address is
-used to map population units to a one-dimensional line. On the one
-dimensional line, each population unit's line length equals its
+Random Tessellation Stratified (GRTS) algorithm because it is readily
+available in the \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle
+et al., 2022) and naturally accommodates finite and infinite sampling
+frames, unequal inclusion probabilities, and replacement units
+(replacement units are population units that can be sampled when a
+population unit originally selected can no longer be sampled).
+
+The GRTS algorithm selects samples by utilizing a particular mapping
+between two-dimensional and one-dimensional space that preserves
+proximity relationships. First the bounding box of the domain is split
+up into four distinct, equally sized squares called level-one cells.
+Each level-one is randomly assigned an level-one address of 0, 1, 2, or
+3. The set of level-one cells is denoted by \(\mathcal{A}_1\) and
+defined as \(\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}\). Within
+each level-one cell, the inclusion probability for each population unit
+is summed, and if any of these sums exceed one, a second level of cells
+is added. Then each level-one cell is split into four distinct, equally
+sized squares called level-two cells. Each level-two cell is randomly
+assigned a level-two address of 0, 1, 2, or 3. The set of level-two
+cells is denoted by \(\mathcal{A}_2\) and defined as
+\(\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}\).
+The inclusion probabilities within each level-two cell are summed, and
+if any of these sums exceed one, a third level of cells is added. This
+process continues for \(k\) steps, until all level-\(k\) cells have
+inclusion probability sums no larger than one. Then
+\(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).
+
+After determining \(\mathcal{A}_k\), it is placed into hierarchical
+order. Hierarchical order is a numeric order that first sorts
+\(\mathcal{A}_k\) by the level-one addresses from smallest to largest,
+then sorts \(\mathcal{A}_k\) by the level-two addresses from smallest to
+largest, and so on. For example, \(\mathcal{A}_2\) in hierarchical order
+is the set
+\(\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}\). Because
+hierarchical ordering sorts by level-one cells, then level-two cells,
+and so on, population units that have similar hierarchical addresses
+tend to be nearby one another in space.Next each population unit is
+mapped to a one-dimensional line in hierarchical order where each
+population unit's inclusion probability equals its line-length. If a
+level-\(k\) cell has multiple population units in it, they are randomly
+placed within the cell's respective line segment. A uniform random
+variable is then simulated in \([0, 1]\) and a systematic sample is
+selected on the line, yielding \(n\) sample points for a sample size
+\(n\). Each element in this systematic sample falls on some population
+unit's line segment, and thus that population unit is selected in the
+sample.
+
+The GRTS algorithm selects samples by utilizing a particular mapping
+between two-dimensional and one-dimensional space that preserves
+proximity relationships. Via this mapping, units in two-dimensional
+space are partitioned using a hierarchical address. This hierarchical
+address is used to map population units to a one-dimensional line. On
+the one dimensional line, each population unit's line length equals its
 inclusion probability. Then, a systematic sample of population units is
@@
 neighborhood variance estimator are available in the \texttt{spsurvey}
-\textbf{\textsf{R}} package (Dumelle et al., 2021). FPBK is available in
+\textbf{\textsf{R}} package (Dumelle et al., 2022). FPBK is available in
 the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).
@@
 
+\leavevmode\hypertarget{ref-brus1993design}{}%
+Brus, D.J., DeGruijter, J.J., 1993. Design-based versus model-based
+estimates of spatial means: Theory and application in environmental soil
+science. Environmetrics 4, 123--152.
+
 \leavevmode\hypertarget{ref-chan2020bayesian}{}%
@@
 
-\leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
-Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2021. Spsurvey:
+\leavevmode\hypertarget{ref-dumelle2022spsurvey}{}%
+Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2022. Spsurvey:
 Spatial sampling design and analysis.
@@
 
-\leavevmode\hypertarget{ref-foster2017spatially}{}%
-Foster, S.D., Hosack, G.R., Lawrence, E., Przeslawski, R., Hedge, P.,
-Caley, M.J., Barrett, N.S., Williams, A., Li, J., Lynch, T., others,
-2017. Spatially balanced designs that incorporate legacy sites. Methods
-in Ecology and Evolution 8, 1433--1442.
-
 \leavevmode\hypertarget{ref-grafstrom2012spatiallypoisson}{}%

### 2022-04-11 inst/manuscript/manuscript.Rmd
minor updates

@@
   2.
-We provide relevant background for the design-based and model-based approaches and then study their performance using simulations and an analysis of real mercury concentration data.
-In the simulations, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
-In the simualations and real data analysis, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+We provide relevant background for the design-based and model-based approaches and then study their performance using simulated and real data.
+In the simulated data, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
+In the simulated and real data, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
   3.
-When studying the simulations and mercury concentration data, we found that regardless of the strength of spatial dependence in the data, sampling plans that incorporate spatial locations (spatially balanced samples) generally outperform sampling plans that ignore spatial locations (non-spatially balanced samples).
-We also found that model-based approaches tend to outperform design-based approaches, even when the data are skewed (and by consequence, the model-based distributional assumptions violated).
-The performance gap between these approaches is small when spatially balanced samples are used but large when non-spatially balanced samples are used.
-This suggests that the sampling choice (whether to select a sample that is spatially balanced) is most important when performing design-based inference.
+When studying the simulated and real data, we found that regardless of the strength of spatial dependence in the data, the Generalized Random Tessellation Stratified (GRTS) algorithm, which explicitly incorporates spatial locations into sampling, tends to outperform the Simple Random Sampling (SRS) algorithm, which does not explicitly incorporate spatial locations into sampling.
+We also found that model-based approaches tend to outperform design-based approaches, even for skewed data where the model-based distributional assumptions are violated.
+The performance gap between these approaches is small GRTS samples are used but large when SRS samples are used.
+This suggests that the sampling choice (whether to use GRTS or SRS) is most important when performing design-based inference.
   4.
@@
 Randomness is incorporated via the selection of population units according to a sampling design.
-A sampling design assigns a positive probability of inclusion (inclusion probability) in the sample to each population unit.
+A sampling design assigns a non-zero probability of inclusion (inclusion probability) in the sample to each population unit.
 These inclusion probabilities are later used to estimate population parameters.
@@
 
-When sampling designs incorporate spatial locations into sampling, we call the resulting samples "spatially balanced." One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
+When samples are chosen in a manner such that the layout of sampled units reflects the layout of the population units, we call the resulting sample "spatially balanced." By "reflecting the layout of the population units", we mean that if population units are concentrated in specific areas, the units in the sample should be concentrated in the same areas.
+Because spatially balanced samples reflect the layout of the population units, they are not necessarily "spread out" in space in some equidistant manner.
+
+One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
 When sampling designs do not incorporate spatial locations into sampling, we call the resulting samples "non-spatially balanced."
@@
 For example, suppose the realized mean of all population units is the value of interest.
-Instead of a fixed, unknown mean, we are the value of the mean, a random variable.
+Instead of a fixed, unknown mean, we are predicting the value of the mean, a random variable.
 Prediction intervals are then derived using assumptions of the data-generating stochastic process.
@@
 
-We previously mentioned that the design-based approach can be used to select spatially balanced samples (samples that incorporate spatial locations of the population units).
+We previously mentioned that the design-based approach can be used to select spatially balanced samples.
 Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
@@
 Each element in this systematic sample falls on some population unit's line segment, and thus that population unit is selected in the sample.
-
-The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
-Via this mapping, units in two-dimensional space are partitioned using a hierarchical address.
-This hierarchical address is used to map population units to a one-dimensional line.
-On the one dimensional line, each population unit's line length equals its inclusion probability.
-Then, a systematic sample of population units is selected on the line and mapped back to two-dimensional space, yielding the desired sample.
-@stevens2004spatially provide more technical details.
+For further details regarding the GRTS algorithm, see @stevens2004spatially.
 
@@
 @horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
-First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically.
+First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically for GRTS samples.
 Second, these estimators ignore the spatial locations of the population units.
 To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$ and incorporates spatial locations -- for technical details see @stevens2003variance.
-@stevens2003variance show the local neighborhood variance estimator tends to reduce the estimated variance of $\hat{\tau}$ and yield more precise (narrower) confidence intervals compared to variance estimators that ignore spatial locations.
+The local neighborhood variance estimator does not rely on $\pi_{ij}$ and estimates the variance of $\hat{\tau}$ conditional on the random properties of the GRTS sample -- the idea being that this conditioning should yield a more precise estimate of $\hat{\tau}$.
+They show that the each observation's contribution to the overall variance is dominated by local variation.
+Thus the local neighborhood variance estimator is a weighted sum of variance estimates from each observation's local neighborhood.
+These local neighborhoods contain observation itself and its three nearest neighbors.
+For more details, see @stevens2003variance.
 
@@
 Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
-We summarize some of the basic principles of FBPK next -- for technical details, see @verhoef2008spatial.
+We summarize some of the basic principles of FPBK next -- for technical details, see @verhoef2008spatial.
 Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
@@
 
-FBPK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
+FPBK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
 This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
@@
 \end{align}
-where $\sigma^2_{1}$ is the variance parameter quantifying the variability that is dependent (coarse-scale), $\sigma^2_{2}$ is the variance parameter quantifying the variability that is independent (fine-scale), $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-The proportion of variability attributable to dependent random error is $\sigma^2_{1} / (\sigma^2_{1} + \sigma^2_{2})$.
-Similarly, the proportion of variability attributable to independent random error is $\sigma^2_{2} / (\sigma^2_{1} + \sigma^2_{2})$.
-Finally we note that $\sigma^2_{1}$ and $\sigma^2_{2}$ are often called the partial sill and nugget, respectively.
+where $\sigma^2_{1}$ is the variance parameter that quantifies the spatially dependent variability, $\sigma^2_{2}$ is the variance parameter the quantifies that spatially independent variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+In geostatistical literature, $\sigma^2_{1}$ is often called the partial sill and $\sigma^2_{2}$ is often called the nugget.
 
-With the above model formulation, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
+The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on using restricted maximum likelihood [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+REML is preferred over maximum likelihood (ML) because ML estimates can be badly biased for small sample sizes, due to the fact that ML makes no adjustment for the simultaneous estimation of $\bm{\beta}$ and $\bm{\delta}$ [@patterson1971recovery].
+Minus twice the REML log-likelihood of the sampled sites is given by
+\begin{equation}\label{equation:reml}
+  \ln|\bm{\Sigma}| + (\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}})^T \bm{\Sigma}_{ss}^{-1}(\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}}) + \ln|\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s| + (n - p) \ln(2 \pi) ,
+\end{equation}
+where $\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s$ and $\bm{\Sigma}_{ss}$  is the covariance matrix of the sampled sites.
+Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\delta}}_{reml}$, the REML estimates of $\bm{\delta}$.
+Then $\bm{\beta}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\delta}}_{reml}$.
+
+With the model formulation in Equation$~$\ref{equation:Zmarginal}, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
 While details of the derivation are in @verhoef2008spatial, we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
@@
 Other approaches, such as k-nearest-neighbors [@fix1989discriminatory; @ver2013comparison] and random forest [@breiman2001random], among others, could also be used to obtain predictions for a mean or total from finite population spatial data.
-Compared to the k-nearest-neighbors and random forest approach, we prefer FBPK because it is model-based and relies on theoretically-based variance estimators leveraging the model's spatial covariance structure, whereas k-nearest-neighbors and random forests use ad-hoc variance estimators [@ver2013comparison].
-Additionally, @ver2013comparison studied compared FBPK, k-nearest-neighbors, and random forest in a variety of spatial data contexts, and FBPK tended to perform best.
+Compared to the k-nearest-neighbors and random forest approach, we prefer FPBK because it is model-based and relies on theoretically-based variance estimators leveraging the model's spatial covariance structure, whereas k-nearest-neighbors and random forests use ad-hoc variance estimators [@ver2013comparison].
+Additionally, @ver2013comparison compared FPBK, k-nearest-neighbors, and random forest in a variety of spatial data contexts, and FPBK tended to perform best.
 
@@
 
-## Simulation Study {#sec:mm_sim}
+## Simulated Data {#sec:mm_sim}
 
@@
 In IRS-Design, samples were analyzed using the design-based approach via the Horvitz-Thompson mean estimator and an IRS variance estimator that ignored the spatial locations of the units in the sample.
-The second sampling-analysis combination was IRS-Model, where samples were selected with the IRS algorithm and analyzed using the model-based approach via Restricted Maximum Likelihood (REML) estimation [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The second sampling-analysis combination was IRS-Model, where samples were selected with the IRS algorithm and analyzed using the model-based approach while estimating the covariance parameters ($\bm{\delta}$) and fixed effects ($\bm{\beta}$ using restricted maximum likelihood (REML).
 The third sampling-analysis combination was GRTS-Design, where samples were selected with the GRTS algorithm and analyzed using the design-based approach via the Horvitz-Thompson mean estimator and the local neighborhood variance estimator (which does incorporate the spatial locations of the units in the sample).
-The fourth and final sampling-analysis combination was GRTS-Model, where samples were selected with the GRTS algorithm and analyzed using the model-based approach via REML estimation.
+The fourth and final sampling-analysis combination was GRTS-Model, where samples were selected with the GRTS algorithm and analyzed using the model-based approach while estimating the covariance parameters ($\bm{\delta}$) and fixed effects ($\bm{\beta}$ using restricted maximum likelihood (REML).
 These sampling-analysis combinations are also provided in Table \ref{tab:designanalysis}.
@@
 ```
+
+CHANGE LOGNORMAL VERBAGE TO SKEWED
 
@@
 
-## Application {#sec:mm_app}
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
+  \caption{}
+  \label{fig:symm_pop_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
+  \caption{}
+  \label{fig:skew_pop_hist}
+\end{subfigure}
+\caption{Histograms of an example population realization for the normal (a) and skewed (b) scenarios for the simulated data.}
+\label{fig:sim_pops}
+\end{figure}
+
+## National Lakes Assessment Data {#sec:mm_app}
 
@@
 
-## Simulation Study {#sec:r_sim}
+## Simulated Data {#sec:r_sim}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
-  \caption{Relative MStdE in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:mse_eff}
-\end{figure}
-
 Fig.
@@
 
-## Application {#sec:r_app}
+## National Lakes Assessment DAta {#sec:r_app}
 
@@
 Design-based approaches are often computationally efficient, while model-based approaches can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
-The design-based approach also more naturally handles binary data, free from the more complicated logistic regression framework commonly used to analyze binary data in a model-based approach.
+The design-based approach easily handles binary data through a straightforward application of the Horvitz-Thompson estimator.
+In contrast, analyzing binary data using a model-based approach generally requires a logistic mixed regression model, which can be challenging to estimate and interpret [@bolker2009generalized].
 The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
@@
 # Acknowledgments {.unnumbered}
+
+We would like to thank the editors and anonymous reviewers for their thoughtful comments which we feel greatly improved the manuscript.
 

### 2022-04-11 inst/manuscript/manuscript.tex
minor updates

@@
     We provide relevant background for the design-based and model-based
-    approaches and then study their performance using simulations and an
-    analysis of real mercury concentration data. In the simulations, a
-    variety of sample sizes, location layouts, dependence structures, and
-    response types are considered. In the simualations and real data
-    analysis, the population mean is the parameter of interest and
-    performance is measured using statistics like bias, squared error, and
-    interval coverage.
+    approaches and then study their performance using simulated and real
+    data. In the simulated data, a variety of sample sizes, location
+    layouts, dependence structures, and response types are considered. In
+    the simulated and real data, the population mean is the parameter of
+    interest and performance is measured using statistics like bias,
+    squared error, and interval coverage.
   \item
-    When studying the simulations and mercury concentration data, we found
-    that regardless of the strength of spatial dependence in the data,
-    sampling plans that incorporate spatial locations (spatially balanced
-    samples) generally outperform sampling plans that ignore spatial
-    locations (non-spatially balanced samples). We also found that
+    When studying the simulated and real data, we found that regardless of
+    the strength of spatial dependence in the data, the Generalized Random
+    Tessellation Stratified (GRTS) algorithm, which explicitly
+    incorporates spatial locations into sampling, tends to outperform the
+    Simple Random Sampling (SRS) algorithm, which does not explicitly
+    incorporate spatial locations into sampling. We also found that
     model-based approaches tend to outperform design-based approaches,
-    even when the data are skewed (and by consequence, the model-based
-    distributional assumptions violated). The performance gap between
-    these approaches is small when spatially balanced samples are used but
-    large when non-spatially balanced samples are used. This suggests that
-    the sampling choice (whether to select a sample that is spatially
-    balanced) is most important when performing design-based inference.
+    even for skewed data where the model-based distributional assumptions
+    are violated. The performance gap between these approaches is small
+    GRTS samples are used but large when SRS samples are used. This
+    suggests that the sampling choice (whether to use GRTS or SRS) is most
+    important when performing design-based inference.
   \item
@@
 incorporated via the selection of population units according to a
-sampling design. A sampling design assigns a positive probability of
+sampling design. A sampling design assigns a non-zero probability of
 inclusion (inclusion probability) in the sample to each population unit.
@@
 
-When sampling designs incorporate spatial locations into sampling, we
-call the resulting samples ``spatially balanced.'' One approach to
-selecting spatially balanced samples is the Generalized Random
-Tessellation Stratified (GRTS) algorithm (Stevens and Olsen, 2004),
-which we discuss in more detail in Section \ref{subsec:spb_design}. When
-sampling designs do not incorporate spatial locations into sampling, we
-call the resulting samples ``non-spatially balanced.''
+When samples are chosen in a manner such that the layout of sampled
+units reflects the layout of the population units, we call the resulting
+sample ``spatially balanced.'' By ``reflecting the layout of the
+population units'', we mean that if population units are concentrated in
+specific areas, the units in the sample should be concentrated in the
+same areas. Because spatially balanced samples reflect the layout of the
+population units, they are not necessarily ``spread out'' in space in
+some equidistant manner.
+
+One approach to selecting spatially balanced samples is the Generalized
+Random Tessellation Stratified (GRTS) algorithm (Stevens and Olsen,
+2004), which we discuss in more detail in Section
+\ref{subsec:spb_design}. When sampling designs do not incorporate
+spatial locations into sampling, we call the resulting samples
+``non-spatially balanced.''
 
@@
 mean of all population units is the value of interest. Instead of a
-fixed, unknown mean, we are the value of the mean, a random variable.
-Prediction intervals are then derived using assumptions of the
+fixed, unknown mean, we are predicting the value of the mean, a random
+variable. Prediction intervals are then derived using assumptions of the
 data-generating stochastic process. If we repeatedly generate response
@@
 We previously mentioned that the design-based approach can be used to
-select spatially balanced samples (samples that incorporate spatial
-locations of the population units). Spatially balanced samples are
-useful because parameter estimates from these samples tend to vary less
-than parameter estimates from samples that are not spatially balanced
+select spatially balanced samples. Spatially balanced samples are useful
+because parameter estimates from these samples tend to vary less than
+parameter estimates from samples that are not spatially balanced
 (Barabesi and Franceschi, 2011; Benedetti et al., 2017; Grafström and
@@
 unit's line segment, and thus that population unit is selected in the
-sample.
-
-The GRTS algorithm selects samples by utilizing a particular mapping
-between two-dimensional and one-dimensional space that preserves
-proximity relationships. Via this mapping, units in two-dimensional
-space are partitioned using a hierarchical address. This hierarchical
-address is used to map population units to a one-dimensional line. On
-the one dimensional line, each population unit's line length equals its
-inclusion probability. Then, a systematic sample of population units is
-selected on the line and mapped back to two-dimensional space, yielding
-the desired sample. Stevens and Olsen (2004) provide more technical
-details.
+sample. For further details regarding the GRTS algorithm, see Stevens
+and Olsen (2004).
 
@@
 unit \(i\) and population unit \(j\) are both in the sample -- this
-quantity can be challenging if not impossible to calculate analytically.
-Second, these estimators ignore the spatial locations of the population
-units. To address these two drawbacks simultaneously, Stevens and Olsen
-(2003) proposed the local neighborhood variance estimator. The local
-neighborhood variance estimator does not rely on \(\pi_{ij}\) and
-incorporates spatial locations -- for technical details see Stevens and
-Olsen (2003). Stevens and Olsen (2003) show the local neighborhood
-variance estimator tends to reduce the estimated variance of
-\(\hat{\tau}\) and yield more precise (narrower) confidence intervals
-compared to variance estimators that ignore spatial locations.
+quantity can be challenging if not impossible to calculate analytically
+for GRTS samples. Second, these estimators ignore the spatial locations
+of the population units. To address these two drawbacks simultaneously,
+Stevens and Olsen (2003) proposed the local neighborhood variance
+estimator. The local neighborhood variance estimator does not rely on
+\(\pi_{ij}\) and estimates the variance of \(\hat{\tau}\) conditional on
+the random properties of the GRTS sample -- the idea being that this
+conditioning should yield a more precise estimate of \(\hat{\tau}\).
+They show that the each observation's contribution to the overall
+variance is dominated by local variation. Thus the local neighborhood
+variance estimator is a weighted sum of variance estimates from each
+observation's local neighborhood. These local neighborhoods contain
+observation itself and its three nearest neighbors. For more details,
+see Stevens and Olsen (2003).
 
@@
 specific sampling design, we assume the data are generated by a spatial
-stochastic process. We summarize some of the basic principles of FBPK
+stochastic process. We summarize some of the basic principles of FPBK
 next -- for technical details, see Ver Hoef (2008). Let
@@
 
-FBPK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
+FPBK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
 has mean-zero and a spatial dependence structure that can be modeled
@@
 ,
-\end{align} where \(\sigma^2_{1}\) is the variance parameter quantifying
-the variability that is dependent (coarse-scale), \(\sigma^2_{2}\) is
-the variance parameter quantifying the variability that is independent
-(fine-scale), \(\phi\) is the range parameter measuring the
+\end{align} where \(\sigma^2_{1}\) is the variance parameter that
+quantifies the spatially dependent variability, \(\sigma^2_{2}\) is the
+variance parameter the quantifies that spatially independent
+variability, \(\phi\) is the range parameter measuring the
 distance-decay rate of the covariance, and \(h_{i,j}\) is the Euclidean
-distance between population units \(i\) and \(j\). The proportion of
-variability attributable to dependent random error is
-\(\sigma^2_{1} / (\sigma^2_{1} + \sigma^2_{2})\). Similarly, the
-proportion of variability attributable to independent random error is
-\(\sigma^2_{2} / (\sigma^2_{1} + \sigma^2_{2})\). Finally we note that
-\(\sigma^2_{1}\) and \(\sigma^2_{2}\) are often called the partial sill
-and nugget, respectively.
-
-With the above model formulation, the Best Linear Unbiased Predictor
-(BLUP) for \(f(\mathbf{b}'\mathbf{z})\) and its prediction variance can
-be computed. While details of the derivation are in Ver Hoef (2008), we
-note here that the predictor and its variance are both moment-based,
-meaning that they do not rely on any distributional assumptions.
-Distributional assumptions are used, however, when constructing
-prediction intervals.
+distance between population units \(i\) and \(j\). In geostatistical
+literature, \(\sigma^2_{1}\) is often called the partial sill and
+\(\sigma^2_{2}\) is often called the nugget.
+
+The parameters in Equation\(~\)\ref{equation:Zmarginal} can be estimated
+using a variety of techniques, but we focus on using restricted maximum
+likelihood (Harville, 1977; Patterson and Thompson, 1971; Wolfinger et
+al., 1994). REML is preferred over maximum likelihood (ML) because ML
+estimates can be badly biased for small sample sizes, due to the fact
+that ML makes no adjustment for the simultaneous estimation of
+\(\bm{\beta}\) and \(\bm{\delta}\) (Patterson and Thompson, 1971). Minus
+twice the REML log-likelihood of the sampled sites is given by
+\begin{equation}\label{equation:reml}
+  \ln|\bm{\Sigma}| + (\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}})^T \bm{\Sigma}_{ss}^{-1}(\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}}) + \ln|\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s| + (n - p) \ln(2 \pi) ,
+\end{equation} where
+\(\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s\)
+and \(\bm{\Sigma}_{ss}\) is the covariance matrix of the sampled sites.
+Minimizing Equation\(~\)\ref{equation:reml} yields
+\(\bm{\hat{\delta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
+Then \(\bm{\beta}_{reml}\), the REML estimate of \(\bm{\beta}\), is
+given by
+\((\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s\),
+where \(\bm{\hat{\Sigma}}_{ss}\) is \(\bm{\Sigma}_{ss}\) evaluated at
+\(\bm{\hat{\delta}}_{reml}\).
+
+With the model formulation in Equation\(~\)\ref{equation:Zmarginal}, the
+Best Linear Unbiased Predictor (BLUP) for \(f(\mathbf{b}'\mathbf{z})\)
+and its prediction variance can be computed. While details of the
+derivation are in Ver Hoef (2008), we note here that the predictor and
+its variance are both moment-based, meaning that they do not rely on any
+distributional assumptions. Distributional assumptions are used,
+however, when constructing prediction intervals.
 
@@
 from finite population spatial data. Compared to the k-nearest-neighbors
-and random forest approach, we prefer FBPK because it is model-based and
+and random forest approach, we prefer FPBK because it is model-based and
 relies on theoretically-based variance estimators leveraging the model's
@@
 forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
-Additionally, Ver Hoef and Temesgen (2013) studied compared FBPK,
+Additionally, Ver Hoef and Temesgen (2013) compared FPBK,
 k-nearest-neighbors, and random forest in a variety of spatial data
-contexts, and FBPK tended to perform best.
+contexts, and FPBK tended to perform best.
 
@@
 \hypertarget{sec:mm_sim}{%
-\subsection{Simulation Study}\label{sec:mm_sim}}
+\subsection{Simulated Data}\label{sec:mm_sim}}
 
@@
 where samples were selected with the IRS algorithm and analyzed using
-the model-based approach via Restricted Maximum Likelihood (REML)
-estimation (Harville, 1977; Patterson and Thompson, 1971; Wolfinger et
-al., 1994). The third sampling-analysis combination was GRTS-Design,
-where samples were selected with the GRTS algorithm and analyzed using
-the design-based approach via the Horvitz-Thompson mean estimator and
-the local neighborhood variance estimator (which does incorporate the
-spatial locations of the units in the sample). The fourth and final
-sampling-analysis combination was GRTS-Model, where samples were
-selected with the GRTS algorithm and analyzed using the model-based
-approach via REML estimation. These sampling-analysis combinations are
-also provided in Table \ref{tab:designanalysis}. Lastly we note that for
-both the IRS and GRTS samples, equal inclusion probabilities were
-assumed for all population units. When IRS assumes equal inclusion
-probabilities for all population units, the algorithm is equivalent to
-simple random sampling (SRS).
+the model-based approach while estimating the covariance parameters
+(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
+maximum likelihood (REML). The third sampling-analysis combination was
+GRTS-Design, where samples were selected with the GRTS algorithm and
+analyzed using the design-based approach via the Horvitz-Thompson mean
+estimator and the local neighborhood variance estimator (which does
+incorporate the spatial locations of the units in the sample). The
+fourth and final sampling-analysis combination was GRTS-Model, where
+samples were selected with the GRTS algorithm and analyzed using the
+model-based approach while estimating the covariance parameters
+(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
+maximum likelihood (REML). These sampling-analysis combinations are also
+provided in Table \ref{tab:designanalysis}. Lastly we note that for both
+the IRS and GRTS samples, equal inclusion probabilities were assumed for
+all population units. When IRS assumes equal inclusion probabilities for
+all population units, the algorithm is equivalent to simple random
+sampling (SRS).
 
@@
 \end{table}
+
+CHANGE LOGNORMAL VERBAGE TO SKEWED
 
@@
 
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
+  \caption{}
+  \label{fig:symm_pop_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
+  \caption{}
+  \label{fig:skew_pop_hist}
+\end{subfigure}
+\caption{Histograms of an example population realization for the normal (a) and skewed (b) scenarios for the simulated data.}
+\label{fig:sim_pops}
+\end{figure}
+
 \hypertarget{sec:mm_app}{%
-\subsection{Application}\label{sec:mm_app}}
+\subsection{National Lakes Assessment Data}\label{sec:mm_app}}
 
@@
 \hypertarget{sec:r_sim}{%
-\subsection{Simulation Study}\label{sec:r_sim}}
+\subsection{Simulated Data}\label{sec:r_sim}}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
-  \caption{Relative MStdE in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:mse_eff}
-\end{figure}
-
 Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
@@
 \hypertarget{sec:r_app}{%
-\subsection{Application}\label{sec:r_app}}
+\subsection{National Lakes Assessment DAta}\label{sec:r_app}}
 
@@
 that rely on inverting a covariance matrix. The design-based approach
-also more naturally handles binary data, free from the more complicated
-logistic regression framework commonly used to analyze binary data in a
-model-based approach. The model-based approach, however, can more
-naturally quantify the relationship between covariates (predictor
-variables) and the response variable. The model-based approach also
-yields estimated spatial covariance parameters, which help better
-understand the dependence structure in the stochastic process of study.
-Model selection is also possible using model-based approaches and
-criteria such as cross validation, likelihood ratio tests, or AIC
-(Akaike, 1974). Model-based approaches are capable of more efficient
-small-area estimation than design-based approaches by leveraging
-distributional assumptions in areas with few observed units. Model-based
-approaches can also compute unit-by-unit predictions at unobserved
-locations and use them to construct informative visualizations like
-smoothed maps. In short, when deciding whether the design-based or
-model-based approach is more appropriate to implement, the benefits and
-drawbacks of each approach should be considered alongside the particular
-goals of the study.
+easily handles binary data through a straightforward application of the
+Horvitz-Thompson estimator. In contrast, analyzing binary data using a
+model-based approach generally requires a logistic mixed regression
+model, which can be challenging to estimate and interpret (Bolker et
+al., 2009). The model-based approach, however, can more naturally
+quantify the relationship between covariates (predictor variables) and
+the response variable. The model-based approach also yields estimated
+spatial covariance parameters, which help better understand the
+dependence structure in the stochastic process of study. Model selection
+is also possible using model-based approaches and criteria such as cross
+validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
+approaches are capable of more efficient small-area estimation than
+design-based approaches by leveraging distributional assumptions in
+areas with few observed units. Model-based approaches can also compute
+unit-by-unit predictions at unobserved locations and use them to
+construct informative visualizations like smoothed maps. In short, when
+deciding whether the design-based or model-based approach is more
+appropriate to implement, the benefits and drawbacks of each approach
+should be considered alongside the particular goals of the study.
 
@@
 \addcontentsline{toc}{section}{Acknowledgments}
+
+We would like to thank the editors and anonymous reviewers for their
+thoughtful comments which we feel greatly improved the manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-bolker2009generalized}{}%
+Bolker, B.M., Brooks, M.E., Clark, C.J., Geange, S.W., Poulsen, J.R.,
+Stevens, M.H.H., White, J.-S.S., 2009. Generalized linear mixed models:
+A practical guide for ecology and evolution. Trends in ecology \&
+evolution 24, 127--135.
+
 \leavevmode\hypertarget{ref-breiman2001random}{}%

### 2022-04-11 inst/scripts/figures/data.R

@@ -31,6 +31,49 @@ combo_data <- combo_data %>%
   ungroup() %>%
   mutate(n_factor = factor(n))
 
+# create adjusted factor
+combo_data <- combo_data %>%
+  mutate(n_factor_adj = case_when(
+    n == 50 & approach == "Design IRS" ~ 35,
+    n == 50 & approach == "Model IRS" ~ 45,
+    n == 50 & approach == "Design GRTS" ~ 55,
+    n == 50 & approach == "Model GRTS" ~ 65,
+    n == 100 & approach == "Design IRS" ~ 85,
+    n == 100 & approach == "Model IRS" ~ 95,
+    n == 100 & approach == "Design GRTS" ~ 105,
+    n == 100 & approach == "Model GRTS" ~ 115,
+    n == 200 & approach == "Design IRS" ~ 135,
+    n == 200 & approach == "Model IRS" ~ 145,
+    n == 200 & approach == "Design GRTS" ~ 155,
+    n == 200 & approach == "Model GRTS" ~ 165,
+    TRUE ~ NA_real_
+  ))
+
+combo_data <- combo_data %>%
+  pivot_longer(cols = c(rel_efficiency, coverage), names_to = "stat", values_to = "value")
+colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+stat_labs <- c(rel_efficiency = "Relative rMS(P)E", coverage = "Interval Coverage")
+rmspe_eff <- combo_data %>%
+  ggplot(aes(x = n_factor_adj, y = value, colour = approach)) +
+  facet_grid(
+    stat ~ var,
+    labeller = labeller(var = resptype_labs, stat = stat_labs)
+  ) +
+  scale_colour_manual(values = colour_scale) +
+  geom_point(size = 2) +
+  geom_vline(xintercept = c(75, 125), lty = "solid") +
+  scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
+  theme_bw(base_size = base_size) +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
+  # lims(y = c(0.4, 1.1)) +
+  theme(
+    panel.grid.major = element_blank(),
[…]

### 2022-04-11 inst/scripts/figures/simulation.R

@@ -33,18 +33,38 @@ combo_data <- combo_data %>%
   mutate(psill_ratio = 1 - nugget_ratio) %>%
   mutate(n_factor = factor(n))
 
+# create adjusted factor
+combo_data <- combo_data %>%
+  mutate(n_factor_adj = case_when(
+    n == 50 & approach == "Design IRS" ~ 35,
+    n == 50 & approach == "Model IRS" ~ 45,
+    n == 50 & approach == "Design GRTS" ~ 55,
+    n == 50 & approach == "Model GRTS" ~ 65,
+    n == 100 & approach == "Design IRS" ~ 85,
+    n == 100 & approach == "Model IRS" ~ 95,
+    n == 100 & approach == "Design GRTS" ~ 105,
+    n == 100 & approach == "Model GRTS" ~ 115,
+    n == 200 & approach == "Design IRS" ~ 135,
+    n == 200 & approach == "Model IRS" ~ 145,
+    n == 200 & approach == "Design GRTS" ~ 155,
+    n == 200 & approach == "Model GRTS" ~ 165,
+    TRUE ~ NA_real_
+  ))
+
 # figure 2
 colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
+resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Skewed")
 psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
 rmspe_eff <- combo_data %>%
-  ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
+  ggplot(aes(x = n_factor_adj, y = rel_efficiency, colour = approach)) +
   facet_grid(
     psill_ratio ~ resptype,
     labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
   ) +
-  geom_jitter(width = 0.24, size = 2) +
   scale_colour_manual(values = colour_scale) +
+  geom_point(size = 2) +
+  geom_vline(xintercept = c(75, 125), lty = "solid") +
+  scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
   theme_bw(base_size = base_size) +
[…]

### 2022-04-11 inst/scripts/figures/simulation_exploratory.R

@@ -1,0 +1,63 @@
+# preliminaries
+
+## load packages
+library(ggplot2)
+library(here)
+library(DvMsp)
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## histogram for the simultation
+set.seed(2)
+symm_pop_var <- 2
+symm_pop <- sim_pop(N = 900, gridded = FALSE, cortype = "Exponential", psill = symm_pop_var * 0.5,
+                    range = sqrt(2) / 3, nugget = 0.5 * symm_pop_var, resptype = "normal", mu = 0)
+symm_plot <- ggplot(symm_pop, aes(x = response)) +
+    geom_histogram(bins = 20) +
+    labs(x = "Response", y = "Count") +
+    theme_bw(base_size = 18) +
+    theme(
+      panel.grid.major = element_blank(),
+      panel.grid.minor = element_blank(),
+      panel.border = element_blank(),
+    )
+
+
+# write out the image
+if (write_out) {
+  ggsave(
+    plot = symm_plot,
+    file = here("inst", "manuscript", "figures", "symm_pop_hist.jpeg"),
+    dpi = 300,
+    width = 5.07,
+    height = 4.39
+  )
+}
+
+
+
+skew_pop_var <- uniroot(function(x) (exp(x) - 1) * exp(x) - 2, interval = c(0, 10))$root
+skew_pop <- sim_pop(N = 900, gridded = FALSE, cortype = "Exponential", psill = skew_pop_var * 0.5,
+                    range = sqrt(2) / 3, nugget = 0.5 * skew_pop_var, resptype = "lognormal", mu = 0)
+skew_plot <- ggplot(skew_pop, aes(x = response)) +
+  geom_histogram(bins = 20) +
+  labs(x = "Response", y = "Count") +
+  theme_bw(base_size = 18) +
+  theme(
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    panel.border = element_blank(),
+  )
+
+
+# write out the image
+if (write_out) {
+  ggsave(
+    plot = skew_plot,
+    file = here("inst", "manuscript", "figures", "skew_pop_hist.jpeg"),
+    dpi = 300,
+    width = 5.07,
+    height = 4.39
+  )
+}

### 2022-04-12 inst/scripts/figures/data.R
image updates

@@ -2,6 +2,7 @@
 
 ## load packages
 library(tidyverse)
+library(viridis)
 library(here)
 
 ## do you want to write out? default of FALSE
@@ -49,24 +50,49 @@ combo_data <- combo_data %>%
     TRUE ~ NA_real_
   ))
 
-combo_data <- combo_data %>%
-  pivot_longer(cols = c(rel_efficiency, coverage), names_to = "stat", values_to = "value")
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# combo_data <- combo_data %>%
+#   pivot_longer(cols = c(rel_efficiency, coverage), names_to = "stat", values_to = "value")
+# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+# stat_labs <- c(rel_efficiency = "Relative rMS(P)E", coverage = "Interval Coverage")
+# rmspe_eff <- combo_data %>%
+#   ggplot(aes(x = n_factor_adj, y = value, colour = approach)) +
+#   facet_grid(
+#     stat ~ var,
+#     labeller = labeller(var = resptype_labs, stat = stat_labs)
+#   ) +
+#   scale_colour_manual(values = colour_scale) +
+#   geom_point(size = 2) +
+#   geom_vline(xintercept = c(75, 125), lty = "solid") +
+#   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
+#   theme_bw(base_size = base_size) +
+#   labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
+#   # lims(y = c(0.4, 1.1)) +
+#   theme(
+#     panel.grid.major = element_blank(),
+#     panel.grid.minor = element_blank(),
+#   )
+
+
+
+# figure 2
+# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# colour_scale <- palette(hcl.colors(4, "viridis"))
+colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
[…]

### 2022-04-12 inst/scripts/figures/simulation.R

@@ -52,7 +52,9 @@ combo_data <- combo_data %>%
   ))
 
 # figure 2
-colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# colour_scale <- palette(hcl.colors(4, "viridis"))
+colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Skewed")
 psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
 rmspe_eff <- combo_data %>%
@@ -65,9 +67,10 @@ rmspe_eff <- combo_data %>%
   geom_point(size = 2) +
   geom_vline(xintercept = c(75, 125), lty = "solid") +
   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
+  scale_y_continuous(breaks = c(0.5, 0.75, 1), labels = c("0.50", "0.75", "1.00"), limits = c(0.5, 1.1)) +
   theme_bw(base_size = base_size) +
   labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
-  lims(y = c(0.4, 1.1)) +
+  # lims(y = c(0.5, 1.1)) +
   theme(
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
@@ -159,6 +162,9 @@ if (write_out) {
 # }
 
 # figure 4
+# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
+# colour_scale <- palette(hcl.colors(4, "viridis"))
+colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 coverage <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95, lty = "dashed") +

### 2022-04-12 inst/manuscript/manuscript.Rmd
image updates
more review updates

@@
 There are two general approaches for using samples to make frequentist statistical inferences about a population: design-based and model-based.
-In the design-based approach, inference relies on randomly assigning some population units to be in the sample (e.g., random sampling).
+In the design-based approach, inference relies on randomly assigning some population units to be in the sample (random sampling).
 Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying stochastic process that generated the sample.
@@
 @cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
-In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design-based and model-based principles (see e.g., @sterba2009alternative and @cicchitelli2012model, and see @chan2020bayesian for a Bayesian approach).
+In addition to @cooper2006sampling, there has been substantial research and development into estimators that use both design-based and model-based principles (see e.g., @sterba2009alternative and @cicchitelli2012model, and for Bayesian approaches, see @chan2020bayesian and @hofman2021many).
 
@@
 
-Instead of estimating fixed, unknown population parameters, as in the design-based approach, often the goal of model-based inference is to predict a realized variable, or value.
-For example, suppose the realized mean of all population units is the value of interest.
+Instead of estimating fixed, unknown population parameters, as in the design-based approach, often the goal of model-based inference is to predict a realized variable.
+For example, suppose the realized mean of all population units (the realized population mean) is the variable of interest.
 Instead of a fixed, unknown mean, we are predicting the value of the mean, a random variable.
@@
 FPBK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
-This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the distance between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
+This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
 Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$, though @chiles1999geostatistics, pp.
@@
 \end{align}
-where $\sigma^2_{1}$ is the variance parameter that quantifies the spatially dependent variability, $\sigma^2_{2}$ is the variance parameter the quantifies that spatially independent variability, $\phi$ is the range parameter measuring the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-In geostatistical literature, $\sigma^2_{1}$ is often called the partial sill and $\sigma^2_{2}$ is often called the nugget.
+where $\sigma^2_{1}$ is the variance parameter that quantifies the spatially dependent variability, $\sigma^2_{2}$ is the variance parameter the quantifies that spatially independent variability, $\phi$ is the distance parameter that measures the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+In geostatistical literature, $\sigma^2_{1}$ is often called the partial sill, $\sigma^2_{2}$ is often called the nugget, and $\phi$ is often called the range.
 
@@
 
-CHANGE LOGNORMAL VERBAGE TO SKEWED
+CHANGE LOGNORMAL VERBAGE TO SKEWED -- look for DRE acronym
 
 Performance for the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
-The 36 scenarios resulted from the crossing of three sample sizes, two location layouts (of the population units), two response types, and three proportions of dependent random error.
+The 36 scenarios resulted from the crossing of three sample sizes, two location layouts (of the population units), two response types, and three proportions of dependent random error (DRE).
 The three sample sizes ($n$) were $n = 50, n = 100,$ and $n = 200$.
@@
   \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
-  \caption{}
+  \caption{Histogram of a single realized population for the normal response.}
   \label{fig:symm_pop_hist}
@@
   \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
-  \caption{}
+  \caption{Histogram of a single realized population for the skewed response.}
   \label{fig:skew_pop_hist}
 \end{subfigure}
-\caption{Histograms of an example population realization for the normal (a) and skewed (b) scenarios for the simulated data.}
+\caption{Histograms realized populations for the simulated data.}
 \label{fig:sim_pops}
@@
 Lastly we note that the true mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
-\begin{figure}
-\centering
-\begin{subfigure}{0.98\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
-  \caption*{}
-  \label{fig:mercury_map}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
-  \caption*{}
-  \label{fig:mercury_hist}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/sv_plot.jpeg}
-  \caption*{}
-  \label{fig:sv_plot}
-\end{subfigure}
-\caption{Mercury concentration (Hg) visualizations for all 986 lakes in the NLA data.
-A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
-\label{fig:merc}
-\end{figure}
 
@@
 \centering
-\begin{subfigure}{0.98\textwidth}
+\begin{subfigure}{0.49\textwidth}
   \centering
   \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
-  \caption*{}
+  \caption{Spatial map of the ZMMI population.}
   \label{fig:zmmi_map}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
+  \caption{Spatial map of mercury (Hg) population.}
+  \label{fig:mercury_map}
 \end{subfigure} \\
@@
   \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
-  \caption*{}
+  \caption{Histogram of the ZMMI population.}
   \label{fig:zmmi_hist}
@@
   \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_sv_plot.jpeg}
-  \caption*{}
+  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
+  \caption{Histogram of the mercury (Hg) population.}
+  \label{fig:mercury_hist}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
+  \caption{Semivariogram of the ZMMI population.}
   \label{fig:zmmi_sv_plot}
 \end{subfigure}
-\caption{zmmi visualizations for all 986 lakes in the NLA data.
-A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
+  \caption{Semivariogram of the mercury (Hg) population.}
+  \label{fig:mercury_sv_plot}
+\end{subfigure}
+\caption{Exploratory graphics of the ZMMI and mercury (Hg) populations in the National Lakes Assessment (NLA) 2012 data.}
 \label{fig:zmmi}
@@
   \centering
-  \includegraphics[width = 1\linewidth]{figures/data_mse_eff.jpeg}
-  \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:data_mse_eff}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
@@
 # Discussion {#sec:discussion}
+
+ADD EXTRAS LIKE ANISOTROPY AND UNEQUAL INCLUSION PROBABILITIES
 
@@
 
-We would like to thank the editors and anonymous reviewers for their thoughtful comments which we feel greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for their thoughtful comments which greatly improved the manuscript.
 

### 2022-04-12 inst/manuscript/manuscript.tex
image updates
more review updates

@@
 approach, inference relies on randomly assigning some population units
-to be in the sample (e.g., random sampling). Alternatively, in the
-model-based approach, inference relies on distributional assumptions
-about the underlying stochastic process that generated the sample. Each
-paradigm has a deep historical context (Sterba, 2009) and its own set of
-benefits and drawbacks (Hansen et al., 1983). In this manuscript, we
-compare the design-based and model-based approaches for finite
-population spatial data.
+to be in the sample (random sampling). Alternatively, in the model-based
+approach, inference relies on distributional assumptions about the
+underlying stochastic process that generated the sample. Each paradigm
+has a deep historical context (Sterba, 2009) and its own set of benefits
+and drawbacks (Hansen et al., 1983). In this manuscript, we compare the
+design-based and model-based approaches for finite population spatial
+data.
 
@@
 design-based and model-based principles (see e.g., Sterba (2009) and
-Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020) for
-a Bayesian approach).
+Cicchitelli and Montanari (2012), and for Bayesian approaches, see
+Chan-Golston et al. (2020) and Hofman and Brus (2021)).
 
@@
 design-based approach, often the goal of model-based inference is to
-predict a realized variable, or value. For example, suppose the realized
-mean of all population units is the value of interest. Instead of a
-fixed, unknown mean, we are predicting the value of the mean, a random
-variable. Prediction intervals are then derived using assumptions of the
-data-generating stochastic process. If we repeatedly generate response
-values from the same process and select samples, then 95\% of all 95\%
-prediction intervals constructed from a procedure with appropriate
-coverage will contain their respective realized means. Cressie (1993)
-and Schabenberger and Gotway (2017) provide thorough reviews of
-model-based approaches for spatial data. In Fig. \ref{fig:fig1}, we
-provide a visual comparison of the design-based and model-based
-approaches (Ver Hoef (2002) and Brus (2021) provide similar figures).
+predict a realized variable. For example, suppose the realized mean of
+all population units (the realized population mean) is the variable of
+interest. Instead of a fixed, unknown mean, we are predicting the value
+of the mean, a random variable. Prediction intervals are then derived
+using assumptions of the data-generating stochastic process. If we
+repeatedly generate response values from the same process and select
+samples, then 95\% of all 95\% prediction intervals constructed from a
+procedure with appropriate coverage will contain their respective
+realized means. Cressie (1993) and Schabenberger and Gotway (2017)
+provide thorough reviews of model-based approaches for spatial data. In
+Fig. \ref{fig:fig1}, we provide a visual comparison of the design-based
+and model-based approaches (Ver Hoef (2002) and Brus (2021) provide
+similar figures).
 
@@
 assumed to be non-negative, second-order stationary (depending only on
-the distance between population units), isotropic (independent of
-direction), and decay with distance between population units (Cressie,
-1993). Henceforth, it is implied that we have made these same
-assumptions regarding \(\bm{\delta}\), though Chiles and Delfiner
-(1999), pp.~80-93 discuss covariance functions that are not second-order
-stationary, not isotropic, or not either. A variety of flexible
-covariance functions can be used to model \(\bm{\delta}\) (Cressie,
-1993); one example is the exponential covariance function (Cressie
-(1993) provides a thorough list of spatial covariance functions). The
-\(i,j\)th element of the exponential covariance matrix,
+the separation vector (e.g., distance) between population units),
+isotropic (independent of direction), and decay with distance between
+population units (Cressie, 1993). Henceforth, it is implied that we have
+made these same assumptions regarding \(\bm{\delta}\), though Chiles and
+Delfiner (1999), pp.~80-93 discuss covariance functions that are not
+second-order stationary, not isotropic, or not either. A variety of
+flexible covariance functions can be used to model \(\bm{\delta}\)
+(Cressie, 1993); one example is the exponential covariance function
+(Cressie (1993) provides a thorough list of spatial covariance
+functions). The \(i,j\)th element of the exponential covariance matrix,
 \(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
@@
 variance parameter the quantifies that spatially independent
-variability, \(\phi\) is the range parameter measuring the
+variability, \(\phi\) is the distance parameter that measures the
 distance-decay rate of the covariance, and \(h_{i,j}\) is the Euclidean
 distance between population units \(i\) and \(j\). In geostatistical
-literature, \(\sigma^2_{1}\) is often called the partial sill and
-\(\sigma^2_{2}\) is often called the nugget.
+literature, \(\sigma^2_{1}\) is often called the partial sill,
+\(\sigma^2_{2}\) is often called the nugget, and \(\phi\) is often
+called the range.
 
@@
 
-CHANGE LOGNORMAL VERBAGE TO SKEWED
+CHANGE LOGNORMAL VERBAGE TO SKEWED -- look for DRE acronym
 
@@
 units), two response types, and three proportions of dependent random
-error. The three sample sizes (\(n\)) were \(n = 50, n = 100,\) and
-\(n = 200\). Samples were always selected from a population size (\(N\))
-of \(N = 900\). The two location layouts were random and gridded.
-Locations in the random layout were randomly generated inside the unit
-square (\([0, 1] \times [0, 1]\)). Locations in the gridded layout were
-placed on a fixed, equally spaced grid inside the unit square. The two
-response types were normal and lognormal. For the normal response type,
-the response was simulated using mean-zero random errors with the
-exponential covariance (Equation\(~\)\ref{equation:expcov}) for varying
-proportions of dependent random error. The proportion of dependent
-random error is represented by
+error (DRE). The three sample sizes (\(n\)) were \(n = 50, n = 100,\)
+and \(n = 200\). Samples were always selected from a population size
+(\(N\)) of \(N = 900\). The two location layouts were random and
+gridded. Locations in the random layout were randomly generated inside
+the unit square (\([0, 1] \times [0, 1]\)). Locations in the gridded
+layout were placed on a fixed, equally spaced grid inside the unit
+square. The two response types were normal and lognormal. For the normal
+response type, the response was simulated using mean-zero random errors
+with the exponential covariance (Equation\(~\)\ref{equation:expcov}) for
+varying proportions of dependent random error. The proportion of
+dependent random error is represented by
 \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where \(\sigma^2_1\) and
@@
 
-\begin{figure}
-\centering
-\begin{subfigure}{0.98\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
-  \caption*{}
-  \label{fig:mercury_map}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
-  \caption*{}
-  \label{fig:mercury_hist}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/sv_plot.jpeg}
-  \caption*{}
-  \label{fig:sv_plot}
-\end{subfigure}
-\caption{Mercury concentration (Hg) visualizations for all 986 lakes in the NLA data.
-A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
-\label{fig:merc}
-\end{figure}
-
 We selected a single IRS sample and a single GRTS sample and estimated
@@
   \label{fig:data_rmspe_eff}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/data_mse_eff.jpeg}
-  \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:data_mse_eff}
 \end{figure}
@@
 \section{Discussion}\label{sec:discussion}}
+
+ADD EXTRAS LIKE ANISOTROPY AND UNEQUAL INCLUSION PROBABILITIES
 
@@
 We would like to thank the editors and anonymous reviewers for their
-thoughtful comments which we feel greatly improved the manuscript.
+thoughtful comments which greatly improved the manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-hofman2021many}{}%
+Hofman, S.C., Brus, D., 2021. How many sampling points are needed to
+estimate the mean nitrate-n content of agricultural fields? A
+geostatistical simulation approach with uncertain variograms. Geoderma
+385, 114816.
+
 \leavevmode\hypertarget{ref-horvitz1952generalization}{}%

### 2022-04-12 inst/scripts/figures/data.R

@@ -79,7 +79,8 @@ combo_data <- combo_data %>%
 # colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
-resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+# resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg")
 rmspe_eff <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = rel_efficiency, colour = approach)) +
   facet_grid(
@@ -166,7 +167,8 @@ if (write_out) {
 # colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
-resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+# resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg")
 coverage <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95, lty = "dashed") +

### 2022-04-12 inst/scripts/figures/data_exploratory.R

@@ -53,10 +53,10 @@ nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
 mercury_map <- nla_nomiss_both %>%
   ggplot(aes(x = xcoords, y = ycoords)) +
   geom_point(aes(colour = total_hg)) +
-  scale_colour_viridis_c() +
+  scale_colour_viridis_c(breaks = c(250, 500, 750)) +
   labs(x = "", y = "", colour = "Hg") +
   coord_quickmap() +
-  theme_bw(base_size = 10) +
+  theme_bw(base_size = 18) +
   theme(
     axis.text.x = element_blank(),
     axis.text.y = element_blank(),
@@ -107,7 +107,7 @@ sv_val <- sv(nla_nomiss_both, "xcoords", "ycoords", "total_hg") # same as residu
 sv_plot <- sv_val %>%
   ggplot(aes(x = dist, y = gamma * (1/1000))) +
   geom_point(size = 3) +
-  labs(x = "Distance (Kilometers)", y = "Semi-Variance (Thousands)", title = "") +
+  labs(x = "Distance (Kilometers)", y = "Hg Semi-Variance (Thousands)", title = "") +
   ylim(c(0, 15)) +
   # scale_size_continuous(name = "Pairs") +
   theme_bw(base_size = 18) +
@@ -123,7 +123,7 @@ sv_plot <- sv_val %>%
 if (write_out) {
   ggsave(
     plot = sv_plot,
-    file = here("inst", "manuscript", "figures", "sv_plot.jpeg"),
+    file = here("inst", "manuscript", "figures", "mercury_sv.jpeg"),
     dpi = 300,
     width = 5.07,
     height = 4.39
@@ -177,9 +177,9 @@ zmmi_map <- nla_nomiss_both %>%
   ggplot(aes(x = xcoords, y = ycoords)) +
   geom_point(aes(colour = MMI_ZOOP_NLA6)) +
   scale_colour_viridis_c() +
-  labs(x = "", y = "", colour = "zmmi") +
+  labs(x = "", y = "", colour = "ZMMI") +
   coord_quickmap() +
-  theme_bw(base_size = 10) +
+  theme_bw(base_size = 18) +
   theme(
     axis.text.x = element_blank(),
     axis.text.y = element_blank(),
@@ -205,7 +205,7 @@ if (write_out) {
 zmmi_hist <- nla_nomiss_both %>%
[…]

### 2022-04-12 inst/scripts/figures/simulation.R

@@ -56,7 +56,7 @@ combo_data <- combo_data %>%
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Skewed")
-psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
+psill_ratio_labs <- c("0" = "Prop DRE: 0", "0.5" = "Prop DRE: 0.5", "0.9" = "Prop DRE: 0.9")
 rmspe_eff <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = rel_efficiency, colour = approach)) +
   facet_grid(

### 2022-04-12 inst/scripts/figures/simulation_exploratory.R

@@ -15,7 +15,7 @@ symm_pop <- sim_pop(N = 900, gridded = FALSE, cortype = "Exponential", psill = s
                     range = sqrt(2) / 3, nugget = 0.5 * symm_pop_var, resptype = "normal", mu = 0)
 symm_plot <- ggplot(symm_pop, aes(x = response)) +
     geom_histogram(bins = 20) +
-    labs(x = "Response", y = "Count") +
+    labs(x = "Normal Response", y = "Count") +
     theme_bw(base_size = 18) +
     theme(
       panel.grid.major = element_blank(),
@@ -42,7 +42,7 @@ skew_pop <- sim_pop(N = 900, gridded = FALSE, cortype = "Exponential", psill = s
                     range = sqrt(2) / 3, nugget = 0.5 * skew_pop_var, resptype = "lognormal", mu = 0)
 skew_plot <- ggplot(skew_pop, aes(x = response)) +
   geom_histogram(bins = 20) +
-  labs(x = "Response", y = "Count") +
+  labs(x = "Skewed Response", y = "Count") +
   theme_bw(base_size = 18) +
   theme(
     panel.grid.major = element_blank(),
@@ -61,3 +61,16 @@ if (write_out) {
     height = 4.39
   )
 }
+
+
+# # population simulations
+#
+# ## normal
+# df_norm <- data.frame(x = seq(-5, 5, length.out = 1000))
+# df_norm$y <- dnorm(df_norm$x, mean = 0, sd = sqrt(2))
+# ggplot(df_norm, aes(x = x, ymin = 0, ymax = y)) + geom_ribbon()
+#
+# ## lognormal
+# df_lnorm <- data.frame(x = seq(0, 10, length.out = 1000))
+# df_lnorm$y <- dlnorm(df_lnorm$x, meanlog = 0, sdlog = sqrt(2))
+# ggplot(df_lnorm, aes(x = x, ymin = 0, ymax = y)) + geom_ribbon()

### 2022-04-13 inst/manuscript/manuscript.Rmd
more review updates

@@
 @de1990model and @brus1993design give early comparisons of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
-Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @wang2012review; @brus2021statistical].
+Since then, there have been several general comparisons between design-based and model-based approaches for spatial data [@brus1997random; @verhoef2002sampling; @verhoef2008spatial; @brus2021statistical].
 @cooper2006sampling reviews the two approaches in an ecological context before introducing a "model-assisted" variance estimator that combines aspects from each approach.
@@
 
-The design-based approach assumes the population is fixed.
-Randomness is incorporated via the selection of population units according to a sampling design.
-A sampling design assigns a non-zero probability of inclusion (inclusion probability) in the sample to each population unit.
-These inclusion probabilities are later used to estimate population parameters.
+The design-based approach assumes the population is fixed. Randomness is incorporated via the selection of population units according to a sampling design. A sampling
+design assigns a probability of selection to each sample (a subset of population
+units).
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
+The inclusion probability of a population unit follows by summing each population unit's probability of selection in each sample containing that population unit.
+Inclusion probabilities are later used to estimate population parameters.
 
@@
 Because spatially balanced samples reflect the layout of the population units, they are not necessarily "spread out" in space in some equidistant manner.
-
 One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
-When sampling designs do not incorporate spatial locations into sampling, we call the resulting samples "non-spatially balanced."
 
@@
 
-The model-based approach assumes the sample is a random realization of a data-generating stochastic process.
-Randomness is formally incorporated through distributional assumptions on this process.
+The model-based approach assumes the population is a random realization of a data-generating stochastic process called a superpopulation.
+Randomness is formally incorporated through distributional assumptions on the superpopulation.
 Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling.
 Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
-To guard against preferential sampling, model-based approaches often still implement some form of random sampling.
-When model-based approaches implement random sampling, the inclusion probabilities are ignored when analyzing the sample (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the sample).
+To guard against preferential sampling, model-based approaches can implement some form of random sampling, though it is common for model-based approaches to not implement random sampling.
+When model-based approaches do implement random sampling, the inclusion probabilities are ignored when analyzing the sample (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the sample).
 
@@
 To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Dirichlet Tessellations).
-After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata @walvoort2010r, the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
 In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
@@
 First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically for GRTS samples.
-Second, these estimators ignore the spatial locations of the population units.
+Second, these estimators tend to ignore the spatial locations of the population units.
 To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
@@
 ## Simulated Data {#sec:mm_sim}
+
+REWRITE AS SIMPLE RANDOM SAMPLING AND WITHOUT REPLACEMENT
+
+USE SRS / GRTS - DB / MB
+
+ADD LOHR REFERENCE AND FORM FOR SRS VARIANCE WITHOUT REPLACEMENT WITH FPC
 
@@
 When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E and using the GRTS algorithm or a model-based analysis does not result in much, if any, loss in efficiency compared to IRS-Design.
+\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+In these scenarios, using the GRTS algorithm or a model-based analysis does not increase efficiency compared to IRS-Design.
 When there is spatial covariance (Fig.
@@
 ## National Lakes Assessment DAta {#sec:r_app}
+
+USE MERCURY UNITS
 
@@
 
+AT LEAST HAVE DISCUSSION ABOUT MODEL BASED ASSUMPTIONS AND MOVE VALIDITY COMMENTS TO RESULTS SECTION.
+
 There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial data.
@@
 In contrast, analyzing binary data using a model-based approach generally requires a logistic mixed regression model, which can be challenging to estimate and interpret [@bolker2009generalized].
+The design-based approach yields valid results because the sampling plan and inclusion probabilities are specified directly by the researcher, while the model-based approach may not yield valid results if the assumptions made do not not accurately capture reality.
 The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
-The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the stochastic process of study.
+The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the process in study.
 Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
@@
 Model-based approaches can also compute unit-by-unit predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
+@brus1997random provide a more thorough discussion regarding the benefits and drawbacks of the two approaches.
 In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.

### 2022-04-13 inst/manuscript/manuscript.tex
more review updates

@@
 between design-based and model-based approaches for spatial data (Brus
-and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang et al.,
-2012). Cooper (2006) reviews the two approaches in an ecological context
-before introducing a ``model-assisted'' variance estimator that combines
-aspects from each approach. In addition to Cooper (2006), there has been
-substantial research and development into estimators that use both
-design-based and model-based principles (see e.g., Sterba (2009) and
-Cicchitelli and Montanari (2012), and for Bayesian approaches, see
-Chan-Golston et al. (2020) and Hofman and Brus (2021)).
+and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006)
+reviews the two approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. In addition to Cooper (2006), there has been substantial
+research and development into estimators that use both design-based and
+model-based principles (see e.g., Sterba (2009) and Cicchitelli and
+Montanari (2012), and for Bayesian approaches, see Chan-Golston et al.
+(2020) and Hofman and Brus (2021)).
 
@@
 incorporated via the selection of population units according to a
-sampling design. A sampling design assigns a non-zero probability of
-inclusion (inclusion probability) in the sample to each population unit.
-These inclusion probabilities are later used to estimate population
-parameters. Some examples of commonly used sampling designs include
-simple random sampling, stratified random sampling, and cluster
-sampling.
+sampling design. A sampling design assigns a probability of selection to
+each sample (a subset of population units). Some examples of commonly
+used sampling designs include simple random sampling, stratified random
+sampling, and cluster sampling. The inclusion probability of a
+population unit follows by summing each population unit's probability of
+selection in each sample containing that population unit. Inclusion
+probabilities are later used to estimate population parameters.
 
@@
 population units, they are not necessarily ``spread out'' in space in
-some equidistant manner.
-
-One approach to selecting spatially balanced samples is the Generalized
-Random Tessellation Stratified (GRTS) algorithm (Stevens and Olsen,
-2004), which we discuss in more detail in Section
-\ref{subsec:spb_design}. When sampling designs do not incorporate
-spatial locations into sampling, we call the resulting samples
-``non-spatially balanced.''
+some equidistant manner. One approach to selecting spatially balanced
+samples is the Generalized Random Tessellation Stratified (GRTS)
+algorithm (Stevens and Olsen, 2004), which we discuss in more detail in
+Section \ref{subsec:spb_design}.
 
@@
 
-The model-based approach assumes the sample is a random realization of a
-data-generating stochastic process. Randomness is formally incorporated
-through distributional assumptions on this process. Strictly speaking,
-randomness need not be incorporated through random sampling, though
-Diggle et al. (2010) warn against preferential sampling. Preferential
-sampling occurs when the process generating the data locations and the
-process being modeled are not independent of one another. To guard
-against preferential sampling, model-based approaches often still
-implement some form of random sampling. When model-based approaches
-implement random sampling, the inclusion probabilities are ignored when
-analyzing the sample (in contrast to the design-based approach, which
-relies on these inclusion probabilities to analyze the sample).
+The model-based approach assumes the population is a random realization
+of a data-generating stochastic process called a superpopulation.
+Randomness is formally incorporated through distributional assumptions
+on the superpopulation. Strictly speaking, randomness need not be
+incorporated through random sampling, though Diggle et al. (2010) warn
+against preferential sampling. Preferential sampling occurs when the
+process generating the data locations and the process being modeled are
+not independent of one another. To guard against preferential sampling,
+model-based approaches can implement some form of random sampling,
+though it is common for model-based approaches to not implement random
+sampling. When model-based approaches do implement random sampling, the
+inclusion probabilities are ignored when analyzing the sample (in
+contrast to the design-based approach, which relies on these inclusion
+probabilities to analyze the sample).
 
@@
 developed, several other spatially balanced sampling algorithms emerged,
-including the Local Pivotal Method (Grafström et al., 2012; Grafström
-and Matei, 2018), Spatially Correlated Poisson Sampling (Grafström,
-2012), Balanced Acceptance Sampling (Robertson et al., 2013),
-Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017), and
-Halton Iterative Partitioning Sampling (Robertson et al., 2018). In this
-manuscript, we select spatially balanced samples using the Generalized
-Random Tessellation Stratified (GRTS) algorithm because it is readily
-available in the \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle
-et al., 2022) and naturally accommodates finite and infinite sampling
-frames, unequal inclusion probabilities, and replacement units
-(replacement units are population units that can be sampled when a
-population unit originally selected can no longer be sampled).
+including stratified sampling with compact geographical strata Walvoort
+et al. (2010), the Local Pivotal Method (Grafström et al., 2012;
+Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
+(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
+2013), Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017),
+and Halton Iterative Partitioning Sampling (Robertson et al., 2018). In
+this manuscript, we select spatially balanced samples using the
+Generalized Random Tessellation Stratified (GRTS) algorithm because it
+is readily available in the \texttt{spsurvey} \textbf{\textsf{R}}
+package (Dumelle et al., 2022) and naturally accommodates finite and
+infinite sampling frames, unequal inclusion probabilities, and
+replacement units (replacement units are population units that can be
+sampled when a population unit originally selected can no longer be
+sampled).
 
@@
 quantity can be challenging if not impossible to calculate analytically
-for GRTS samples. Second, these estimators ignore the spatial locations
-of the population units. To address these two drawbacks simultaneously,
-Stevens and Olsen (2003) proposed the local neighborhood variance
-estimator. The local neighborhood variance estimator does not rely on
-\(\pi_{ij}\) and estimates the variance of \(\hat{\tau}\) conditional on
-the random properties of the GRTS sample -- the idea being that this
-conditioning should yield a more precise estimate of \(\hat{\tau}\).
-They show that the each observation's contribution to the overall
-variance is dominated by local variation. Thus the local neighborhood
-variance estimator is a weighted sum of variance estimates from each
-observation's local neighborhood. These local neighborhoods contain
-observation itself and its three nearest neighbors. For more details,
-see Stevens and Olsen (2003).
+for GRTS samples. Second, these estimators tend to ignore the spatial
+locations of the population units. To address these two drawbacks
+simultaneously, Stevens and Olsen (2003) proposed the local neighborhood
+variance estimator. The local neighborhood variance estimator does not
+rely on \(\pi_{ij}\) and estimates the variance of \(\hat{\tau}\)
+conditional on the random properties of the GRTS sample -- the idea
+being that this conditioning should yield a more precise estimate of
+\(\hat{\tau}\). They show that the each observation's contribution to
+the overall variance is dominated by local variation. Thus the local
+neighborhood variance estimator is a weighted sum of variance estimates
+from each observation's local neighborhood. These local neighborhoods
+contain observation itself and its three nearest neighbors. For more
+details, see Stevens and Olsen (2003).
 
@@
 
+REWRITE AS SIMPLE RANDOM SAMPLING AND WITHOUT REPLACEMENT
+
+USE SRS / GRTS - DB / MB
+
+ADD LOHR REFERENCE AND FORM FOR SRS VARIANCE WITHOUT REPLACEMENT WITH
+FPC
+
 We used a simulation study to investigate performance of four
@@
 \subsection{National Lakes Assessment DAta}\label{sec:r_app}}
+
+USE MERCURY UNITS
 
@@
 model, which can be challenging to estimate and interpret (Bolker et
-al., 2009). The model-based approach, however, can more naturally
-quantify the relationship between covariates (predictor variables) and
-the response variable. The model-based approach also yields estimated
-spatial covariance parameters, which help better understand the
-dependence structure in the stochastic process of study. Model selection
-is also possible using model-based approaches and criteria such as cross
-validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
-approaches are capable of more efficient small-area estimation than
-design-based approaches by leveraging distributional assumptions in
-areas with few observed units. Model-based approaches can also compute
-unit-by-unit predictions at unobserved locations and use them to
-construct informative visualizations like smoothed maps. In short, when
-deciding whether the design-based or model-based approach is more
-appropriate to implement, the benefits and drawbacks of each approach
-should be considered alongside the particular goals of the study.
+al., 2009). The design-based approach yields valid results because the
+sampling plan and inclusion probabilities are specified directly by the
+researcher, while the model-based approach may not yield valid results
+if the assumptions made do not not accurately capture reailty. The
+model-based approach, however, can more naturally quantify the
+relationship between covariates (predictor variables) and the response
+variable. The model-based approach also yields estimated spatial
+covariance parameters, which help better understand the dependence
+structure in the process in study. Model selection is also possible
+using model-based approaches and criteria such as cross validation,
+likelihood ratio tests, or AIC (Akaike, 1974). Model-based approaches
+are capable of more efficient small-area estimation than design-based
+approaches by leveraging distributional assumptions in areas with few
+observed units. Model-based approaches can also compute unit-by-unit
+predictions at unobserved locations and use them to construct
+informative visualizations like smoothed maps. Brus and De Gruijter
+(1997) provide a more thorough discussion regarding the benefits and
+drawbacks of the two approaches. In short, when deciding whether the
+design-based or model-based approach is more appropriate to implement,
+the benefits and drawbacks of each approach should be considered
+alongside the particular goals of the study.
 
@@
 
+\leavevmode\hypertarget{ref-walvoort2010r}{}%
+Walvoort, D.J., Brus, D., De Gruijter, J., 2010. An r package for
+spatial coverage sampling and random sampling from compact geographical
+strata by k-means. Computers \& geosciences 36, 1261--1267.
+
 \leavevmode\hypertarget{ref-wang2013design}{}%
@@
 
-\leavevmode\hypertarget{ref-wang2012review}{}%
-Wang, J.-F., Stein, A., Gao, B.-B., Ge, Y., 2012. A review of spatial
-sampling. Spatial Statistics 2, 1--14.
-
 \leavevmode\hypertarget{ref-wolfinger1994computing}{}%

### 2022-04-13 inst/scripts/figures/background.R

@@ -51,7 +51,7 @@ d1 <- pop_df %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Design 1", colour = "Response") +
+  labs(title = "Sample 1", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),
@@ -75,7 +75,7 @@ d2 <- pop_df %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Design 2", colour = "Response") +
+  labs(title = "Sample 2", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),
@@ -99,7 +99,7 @@ d3 <- pop_df %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Design 3", colour = "Response") +
+  labs(title = "Sample 3", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),
@@ -132,7 +132,7 @@ m1 <- pop_df_mod %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Model 1", colour = "Response") +
+  labs(title = "Population 1", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),
@@ -156,7 +156,7 @@ m2 <- pop_df_mod2 %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Model 2", colour = "Response") +
+  labs(title = "Population 2", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),
@@ -180,7 +180,7 @@ m3 <- pop_df_mod3 %>%
     show.legend = FALSE,
     stroke = 1.5
   ) +
-  labs(title = "Model 3", colour = "Response") +
+  labs(title = "Population 3", colour = "Response") +
   theme_bw(base_size = base_size) +
   theme(
     axis.title = element_blank(),

### 2022-04-13 inst/scripts/figures/data.R

@@ -50,6 +50,16 @@ combo_data <- combo_data %>%
     TRUE ~ NA_real_
   ))
 
+# new approach variable
+combo_data <- combo_data %>%
+  mutate(approach = case_when(
+    approach == "Design IRS" ~ "SRS-DB",
+    approach == "Model IRS" ~ "SRS-MB",
+    approach == "Design GRTS" ~ "GRTS-DB",
+    approach == "Model GRTS" ~ "GRTS-MB"
+  ) %>% factor(levels = c("SRS-DB", "SRS-MB", "GRTS-DB", "GRTS-MB")))
+
+
 # combo_data <- combo_data %>%
 #   pivot_longer(cols = c(rel_efficiency, coverage), names_to = "stat", values_to = "value")
 # colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")

### 2022-04-13 inst/scripts/figures/simulation.R

@@ -51,6 +51,15 @@ combo_data <- combo_data %>%
     TRUE ~ NA_real_
   ))
 
+# new approach variable
+combo_data <- combo_data %>%
+  mutate(approach = case_when(
+    approach == "Design IRS" ~ "SRS-DB",
+    approach == "Model IRS" ~ "SRS-MB",
+    approach == "Design GRTS" ~ "GRTS-DB",
+    approach == "Model GRTS" ~ "GRTS-MB"
+  ) %>% factor(levels = c("SRS-DB", "SRS-MB", "GRTS-DB", "GRTS-MB")))
+
 # figure 2
 # colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 # colour_scale <- palette(hcl.colors(4, "viridis"))

### 2022-04-13 inst/manuscript/manuscript.Rmd
review edits through 2.1

@@
 We provide relevant background for the design-based and model-based approaches and then study their performance using simulated and real data.
-In the simulated data, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
-In the simulated and real data, the population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+In the simulated and real data, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
+The population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
   3.
-When studying the simulated and real data, we found that regardless of the strength of spatial dependence in the data, the Generalized Random Tessellation Stratified (GRTS) algorithm, which explicitly incorporates spatial locations into sampling, tends to outperform the Simple Random Sampling (SRS) algorithm, which does not explicitly incorporate spatial locations into sampling.
+When studying the simulated and real data, we found that regardless of the strength of spatial dependence in the data, the generalized random tessellation stratified (GRTS) algorithm, which explicitly incorporates spatial locations into sampling, tends to outperform the simple random sampling (SRS) algorithm, which does not explicitly incorporate spatial locations into sampling.
 We also found that model-based approaches tend to outperform design-based approaches, even for skewed data where the model-based distributional assumptions are violated.
@@
 
-Design-based inference; Finite Population Block Kriging (FPBK); Generalized Random Tessellation Stratified (GRTS) algorithm; Local neighborhood variance estimator; Model-based inference; Restricted Maximum Likelihood (REML) estimation; Spatially balanced sampling; Spatial covariance
+Design-based inference; Finite population block kriging (FPBK); Generalized random tessellation stratified (GRTS) algorithm; Local neighborhood variance estimator; Model-based inference; Restricted maximum likelihood (REML) estimation; Spatially balanced sampling; Spatial covariance
 
@@
 In the design-based approach, inference relies on randomly assigning some population units to be in the sample (random sampling).
-Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying stochastic process that generated the sample.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation].
+Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying stochastic process generating the sample.
+Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation, @brus1997random].
 In this manuscript, we compare the design-based and model-based approaches for finite population spatial data.
@@
 
-Spatial data are data that incorporate the locations of the population units into either the sampling or estimation process.
+Spatial data are data that have some sort of spatial index, usually via coordinates.
 @de1990model and @brus1993design give early comparisons of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
@@
 <!-- Paragraph about the contribution of the manuscript and outlining the rest -->
-Certainly comparisons between design-based and model-based approaches have been studied in spatial contexts.
-Our contribution is comparing design-based approaches that incorporate spatial locations into sampling and analysis to model-based approaches.
+While comparisons between design-based and model-based approaches have been studied in spatial contexts, our contribution is comparing design-based approaches specifically built for spatial data to model-based approaches.
 Though the broad comparisons we draw between design-based and model-based approaches generalize to finite and infinite populations, we focus on finite populations.
-A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the contiguous United States.
+A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the conterminous United States.
 An infinite population contains an infinite number of population units; an example is locations within a single lake.
@@
 The rest of the manuscript is organized as follows.
-In Section \ref{sec:background}, we introduce and provide relevant background for the design-based and model-based approaches to finite population spatial data.
-In Section \ref{sec:mm}, we describe how we compare performance of the approaches with a simulation study and an analysis of real data that contains mercury concentration in lakes located in the contiguous United States.
-In Section \ref{sec:results}, we present results from the simulation study and the mercury concentration analysis.
+In Section \ref{sec:background}, we introduce and provide relevant background for design-based and model-based approaches to finite population spatial data.
+In Section \ref{sec:mm}, we describe how we intend to compare performance of the approaches using simulated and real data.
+In Section \ref{sec:results}, we present analysis reslts for the simulated and real data.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 The design-based approach assumes the population is fixed. Randomness is incorporated via the selection of population units according to a sampling design. A sampling
-design assigns a probability of selection to each sample (a subset of population
+design assigns a probability of selection to each sample (subset of population
 units).
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
-The inclusion probability of a population unit follows by summing each population unit's probability of selection in each sample containing that population unit.
+The inclusion probability of a population unit follows by summing each sample's probability of selection over all samples that contain the population unit.
 Inclusion probabilities are later used to estimate population parameters.
 
-When samples are chosen in a manner such that the layout of sampled units reflects the layout of the population units, we call the resulting sample "spatially balanced." By "reflecting the layout of the population units", we mean that if population units are concentrated in specific areas, the units in the sample should be concentrated in the same areas.
-Because spatially balanced samples reflect the layout of the population units, they are not necessarily "spread out" in space in some equidistant manner.
-One approach to selecting spatially balanced samples is the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
+When samples are chosen in a manner such that the layout of sampled units reflects the layout of the population units, we call the resulting sample spatially balanced.
+By "reflecting the layout of the population units", we mean that if population units are concentrated in specific areas, the units in the sample should be concentrated in the same areas.
+Because spatially balanced samples reflect the layout of the population units, they are not necessarily spread out in space in some equidistant manner.
+One approach to selecting spatially balanced samples is the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
 
@@
 
-The model-based approach assumes the population is a random realization of a data-generating stochastic process called a superpopulation.
-Randomness is formally incorporated through distributional assumptions on the superpopulation.
+The model-based approach assumes the population is a random realization of a data-generating stochastic process.
+Randomness is formally incorporated through distributional assumptions on this process.
 Strictly speaking, randomness need not be incorporated through random sampling, though @diggle2010geostatistical warn against preferential sampling.
 Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
-To guard against preferential sampling, model-based approaches can implement some form of random sampling, though it is common for model-based approaches to not implement random sampling.
+To guard against preferential sampling, model-based approaches can implement some form of random sampling, though it is common for model-based approaches to sample non-randomly.
 When model-based approaches do implement random sampling, the inclusion probabilities are ignored when analyzing the sample (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the sample).
@@
 Prediction intervals are then derived using assumptions of the data-generating stochastic process.
-If we repeatedly generate response values from the same process and select samples, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
+If we repeatedly generate realizations from the same process and select samples, then 95% of all 95% prediction intervals constructed from a procedure with appropriate coverage will contain their respective realized means.
 @cressie1993statistics and @schabenberger2017statistical provide thorough reviews of model-based approaches for spatial data.
@@
 \ref{fig:fig1}, we provide a visual comparison of the design-based and model-based approaches (@verhoef2002sampling and @brus2021statistical provide similar figures).
+This figure contrasts the design-based approach with a fixed population and random sampling to the model-based approach with random populations and non-random sampling.
 
@@
 There is one fixed population with nine population units and three random samples of size four (points circled are those sampled).
-The response values at each site are fixed, but we obtain different estimates for the mean response in each random sample.
+The response values at each site are fixed.
 In the bottom row, the model-based approach is highlighted.
 There are three realizations of the same data-generating stochastic process that are all sampled at the same four locations.
-The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations.}
+The response values at each site are random.}
   \label{fig:fig1}
@@
 We previously mentioned that the design-based approach can be used to select spatially balanced samples.
-Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples that are not spatially balanced [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
-The first spatially balanced sampling algorithm to see widespread use was the Generalized Random Tessellation Stratified (GRTS) algorithm [@stevens2004spatially].
-To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (Dirichlet Tessellations).
-After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata @walvoort2010r, the Local Pivotal Method [@grafstrom2012spatially; @grafstrom2018spatially], Spatially Correlated Poisson Sampling [@grafstrom2012spatiallypoisson], Balanced Acceptance Sampling [@robertson2013bas], Within-Sample-Distance Sampling [@benedetti2017spatially], and Halton Iterative Partitioning Sampling [@robertson2018halton].
-In this manuscript, we select spatially balanced samples using the Generalized Random Tessellation Stratified (GRTS) algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units (replacement units are population units that can be sampled when a population unit originally selected can no longer be sampled).
+Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
+The first spatially balanced sampling algorithm to see widespread use was the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially].
+To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (i.e., Dirichlet Tessellations).
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata @walvoort2010r, the local pivotal method [@grafstrom2012spatially; @grafstrom2018spatially], spatially correlated Poisson sampling [@grafstrom2012spatiallypoisson], balanced acceptance sampling [@robertson2013bas], within-sample-distance sampling [@benedetti2017spatially], and Halton iterative partitioning sampling [@robertson2018halton].
+In this manuscript, we select spatially balanced samples using the GRTS algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units.
+Replacement units are additional population units that can be sampled when a population unit originally selected can no longer be sampled.
+A couple reasons why an originally selected site can no longer be sampled include its location being physically inaccessible or on private land that the researcher does not have permission to access.
 
@@
 First the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells.
-Each level-one is randomly assigned an level-one address of 0, 1, 2, or 3.
+Each level-one cell is randomly assigned a level-one address of 0, 1, 2, or 3.
 The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$.
@@
 Hierarchical order is a numeric order that first sorts $\mathcal{A}_k$ by the level-one addresses from smallest to largest, then sorts $\mathcal{A}_k$ by the level-two addresses from smallest to largest, and so on.
-For example, $\mathcal{A}_2$ in hierarchical order is the set $\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}$.
-Because hierarchical ordering sorts by level-one cells, then level-two cells, and so on, population units that have similar hierarchical addresses tend to be nearby one another in space.Next each population unit is mapped to a one-dimensional line in hierarchical order where each population unit's inclusion probability equals its line-length.
+For example, $\mathcal{A}_2$ in hierarchical order is the set \linebreak $\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}$.
+Because hierarchical ordering sorts by level-one cells, then level-two cells, and so on, population units that have similar hierarchical addresses tend to be nearby one another in space.
+Next each population unit is mapped to a one-dimensional line in hierarchical order where each population unit's inclusion probability equals its line-length.
 If a level-$k$ cell has multiple population units in it, they are randomly placed within the cell's respective line segment.
 A uniform random variable is then simulated in $[0, 1]$ and a systematic sample is selected on the line, yielding $n$ sample points for a sample size $n$.
-Each element in this systematic sample falls on some population unit's line segment, and thus that population unit is selected in the sample.
+Each of these sample points falls on some population unit's line segment, and thus that population unit is selected in the sample.
 For further details regarding the GRTS algorithm, see @stevens2004spatially.
@@
 \begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
+  \hat{\tau}_{ht} = \sum_{i = 1}^n z_i \pi_i^{-1},
 \end{align}
-where $Z_i$ is the value of the $i$th population unit in the sample, $\pi_i$ is the inclusion probability of the $i$th population unit in the sample, and $n$ is the sample size.
+where $z_i$ is the value of the $i$th population unit in the sample, $\pi_i$ is the inclusion probability of the $i$th population unit in the sample, and $n$ is the sample size.
 An estimate of the population mean is obtained by dividing $\hat{\tau}_{ht}$ by $N$, the number of population units.
@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and estimates the variance of $\hat{\tau}$ conditional on the random properties of the GRTS sample -- the idea being that this conditioning should yield a more precise estimate of $\hat{\tau}$.
-They show that the each observation's contribution to the overall variance is dominated by local variation.
-Thus the local neighborhood variance estimator is a weighted sum of variance estimates from each observation's local neighborhood.
-These local neighborhoods contain observation itself and its three nearest neighbors.
+They show that the contribution from each sample unit (population unit in the sample) to the overall variance is dominated by local variation.
+Thus the local neighborhood variance estimator is a weighted sum of variance estimates from each sample unit's local neighborhood.
+These local neighborhoods contain the sample unit itself and its three nearest neighbors among all other sample units.
 For more details, see @stevens2003variance.
@@
 
-Finite Population Block Kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
+Finite population block kriging (FPBK) is a model-based approach that expands the geostatistical Kriging framework to the finite population setting [@verhoef2008spatial].
 Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
-We summarize some of the basic principles of FPBK next -- for technical details, see @verhoef2008spatial.
+We summarize some of the basic principles of FPBK next -- for more details, see @verhoef2008spatial.
 Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
@@
 
+In this section we describe how we used simulated and real data to investigate performance between simple random sampling without replacement (SRS) and GRTS sampling as well as performance between design-based (DB) and model-based (MB) inference.
+In SRS and GRTS sampling, all population units had equal inclusion probabilities.
+The important distinction between SRS and GRTS is that SRS ignores spatial locations while sampling but GRTS explicitly incorporates them.
+Together, the two sampling plans (SRS and GRTS) combined with the two inference approaches (DB and MB) yielded four sampling-inference combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB.
+For SRS-DB, the Horvitz-Thompson estimator \eqref{eq:ht} was used to estimate means and the commonly-used SRS variance formula [@sarndal2003model; @lohr2009sampling] was used to estimate the variance.
+This variance formula is given by
+\begin{equation}\label{equation:srs_var}
+ \frac{f[\sum_{i = 1}^n (z_i - \bar{z})^2]}{n(n - 1)},
+\end{equation}
+where $z_i$ is the $i$th response value, $\bar{z}$ is the mean of all $z_i$, $n$ is the sample size, $N$ is the population size, and $f = (1 - n / N)$ ($f$ is often called the finite population correction factor).
+For GRTS-DB, the Horvitz-Thompson esetimator was used to estimate means and the local neighborhood variance was used to estimate variances.
+For SRS-MB and GRTS-MB, FPBK was used to estimate means and variances and parameters were estimated using restricted maximum likelihood.
+
+We used simulated data to compare the sampling-inference combinations across many realized populations from the same data-generating stochastic process.
+With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
+We used real data from the 2012 National Lakes Assessment [@USEPA2012NLA] to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
+With the real data, we were in control of only the random sampling process.
+
 ## Simulated Data {#sec:mm_sim}
 
-REWRITE AS SIMPLE RANDOM SAMPLING AND WITHOUT REPLACEMENT
-
-USE SRS / GRTS - DB / MB
-
-ADD LOHR REFERENCE AND FORM FOR SRS VARIANCE WITHOUT REPLACEMENT WITH FPC
-
-We used a simulation study to investigate performance of four sampling-analysis combinations.
-The first sampling-analysis combination was IRS-Design.
-In IRS-Design, samples were selected with the Independent Random Sampling (IRS) algorithm.
-The IRS algorithm ignores the spatial locations of the population units, thus the IRS samples were not spatially balanced.
-In IRS-Design, samples were analyzed using the design-based approach via the Horvitz-Thompson mean estimator and an IRS variance estimator that ignored the spatial locations of the units in the sample.
-The second sampling-analysis combination was IRS-Model, where samples were selected with the IRS algorithm and analyzed using the model-based approach while estimating the covariance parameters ($\bm{\delta}$) and fixed effects ($\bm{\beta}$ using restricted maximum likelihood (REML).
-The third sampling-analysis combination was GRTS-Design, where samples were selected with the GRTS algorithm and analyzed using the design-based approach via the Horvitz-Thompson mean estimator and the local neighborhood variance estimator (which does incorporate the spatial locations of the units in the sample).
-The fourth and final sampling-analysis combination was GRTS-Model, where samples were selected with the GRTS algorithm and analyzed using the model-based approach while estimating the covariance parameters ($\bm{\delta}$) and fixed effects ($\bm{\beta}$ using restricted maximum likelihood (REML).
-These sampling-analysis combinations are also provided in Table \ref{tab:designanalysis}.
-Lastly we note that for both the IRS and GRTS samples, equal inclusion probabilities were assumed for all population units.
-When IRS assumes equal inclusion probabilities for all population units, the algorithm is equivalent to simple random sampling (SRS).
-
 ```{r, echo = FALSE, results = "asis"}
-tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-       nrow = 2, byrow = TRUE) %>%
-  as.table()
-row.names(tab) <- c("IRS", "GRTS")
-xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-analysis combinations in the simulation study.
+# tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
+#        nrow = 2, byrow = TRUE) %>%
+#   as.table()
+# row.names(tab) <- c("IRS", "GRTS")
+# xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-inference combinations in the simulation study.
 The rows give the two types of sampling designs and the columns give the two types of analyses.")
-align(xtab1) <- "r|ll"
-names(xtab1) <- c("Design", "Model")
-print(xtab1, comment = FALSE)
-## I really want "Analysis" to appear as a row header and "Sampling Design" 
-## to appear as a column header, but this isn't easy to figure out.
-## I couldn't figure this out with xtable either so it might need to be
-## made "by hand."
+# align(xtab1) <- "r|ll"
+# names(xtab1) <- c("Design", "Model")
+# print(xtab1, comment = FALSE)
+# ## I really want "Analysis" to appear as a row header and "Sampling Design" 
+# ## to appear as a column header, but this isn't easy to figure out.
+# ## I couldn't figure this out with xtable either so it might need to be
+# ## made "by hand."
 ```
@@
 
-Performance for the four sampling-analysis combinations was evaluated in 36 different simulation scenarios.
+We evaluated performance of the four sampling-inference combinations in 36 different simulation scenarios.
 The 36 scenarios resulted from the crossing of three sample sizes, two location layouts (of the population units), two response types, and three proportions of dependent random error (DRE).
@@
 Locations in the gridded layout were placed on a fixed, equally spaced grid inside the unit square.
-The two response types were normal and lognormal.
-For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for varying proportions of dependent random error.
-The proportion of dependent random error is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are the dependent random error variance (partial sill) and independent random error variance (nugget) from Equation$~$\ref{equation:expcov}, respectively.
+The two response types were normal and skewed.
+For the normal response type, the response was simulated using mean-zero random errors with the exponential covariance (Equation$~$\ref{equation:expcov}) for three proportions of dependent random error (DRE): 0% DRE, 50% DRE, and 90% DRE.
+Recall the proportion of DRE is represented by $\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)$, where $\sigma^2_1$ and $\sigma^2_2$ are the DRE variance and independent random error (IRE) variance from Equation$~$\ref{equation:expcov}, respectively.
 The total variance, $\sigma^2_1 + \sigma^2_2$, was always 2.
-The range was always $\sqrt{2} / 3$, chosen so that the correlation in the dependent random error decayed to nearly zero at $\sqrt{2}$, the largest possible distance between two population units in the domain.
-For the lognormal response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
-The response was then exponentiated, yielding a lognormal random variable whose total variance was 2.
-The lognormal responses were used to evaluate performance of the sampling-analysis approaches for data that were skewed (i.e., not normal).
-
-```{r, echo = FALSE, results = "asis"}
-n <- c("50", "100", "200")
-site_locations <- c("Random", "Gridded", "")
-psill_ratio <- c("0", "0.5", "0.9")
-resp_type <- c("Normal", "Lognormal", "")
-parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-parm_tab[c(2, 4), 3] <- "-"
-row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
-                         "Proportion of Dependent Error", "Response Type")
-# parm_tab <- parm_tab[c(1, 2, 4, 3), ]
-parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
-All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was 2.")
-align(parm_xtab) <- "r|lll"
-print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
-```
-
-In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-In each trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
-Then we recorded the bias, squared error, standard error, and interval coverage for all sampling-analysis combinations.
-After all 2000 trials, we summarized the long-run performance of the combinations by calculating mean bias, rMS(P)E (root-mean-squared error for the design-based approaches and root-mean-squared-prediction error for the model-based approaches), MStdE (mean standard error), and the proportion of times the true mean is contained in its 95% confidence (design-based) or 95% prediction (model-based) interval.
-The 95% intervals were constructed using the normal distribution.
-Justification for this comes from the asymptotic normality of means via the Central Limit Theorem (under some assumptions).
-Quantifying mean bias and rMS(P)E is important because they help us understand how far (under different loss metrics) the estimates (design-based) or predictions (model-based) tend to be from the true mean.
-Quantifying MStdE is important because it helps us understand how precise intervals tend to be.
-Quantifying interval coverage is important because it helps us understand how often our 95% intervals actually contain the true mean.
-
-The IRS algorithm, IRS variance estimator, GRTS algorithm, and local neighborhood variance estimator are available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey].
-FPBK is available in the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
+The distance parameter was always $\sqrt{2} / 3$, chosen so that the correlation in the DRE decayed to nearly zero at $\sqrt{2}$, the largest possible distance between two population units in the domain.
+For the skewed response type, the response was first simulated using the same approach as for the normal response type, except that the total variance was 0.6931 instead of 2.
+The response was then exponentiated, yielding a skewed random variable whose total variance was 2.
+The skewed responses were used to evaluate performance of the sampling-inference approaches for data that were not normal but were still estimated using REML, which relies on a normal log-likelihood.
+Figure \ref{fig:sim_pops} shows an example of a realized population for the normal and skewed responses using the random layout and 50% DRE.
 
@@
   \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
-  \caption{Histogram of a single realized population for the normal response.}
+  \caption{Histogram of a realized population for the normal response.}
   \label{fig:symm_pop_hist}
@@
   \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
-  \caption{Histogram of a single realized population for the skewed response.}
+  \caption{Histogram of a realized population for the skewed response.}
   \label{fig:skew_pop_hist}
 \end{subfigure}
-\caption{Histograms realized populations for the simulated data.}
+\caption{Histograms of realized populations simulated for the normal and skewed resposnes using the random layout and 50\% DRE.}
 \label{fig:sim_pops}
 \end{figure}
+
+```{r, echo = FALSE, results = "asis"}
+# n <- c("50", "100", "200")
+# site_locations <- c("Random", "Gridded", "")
+# psill_ratio <- c("0", "0.5", "0.9")
+# resp_type <- c("Normal", "Lognormal", "")
+# parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
+# parm_tab[c(2, 4), 3] <- "-"
+# row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
+#                          "Proportion of Dependent Error", "Response Type")
+# # parm_tab <- parm_tab[c(1, 2, 4, 3), ]
+# parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
+All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
+In each simualtion scenario, the total variance was 2.")
+# align(parm_xtab) <- "r|lll"
+# print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
+```
+
+In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
+Within each simulation scenario and trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
+With the model-based analyses, covariance parameters were estimated (using REML) separately for each trial.
+After all 2000 trials, we summarized the long-run performance of the sampling-inference combination in each scenario by calculating mean bias, root-mean-squared error, and interval coverage.
+Mean bias is taken as the average deviation between each trial's estimated (or predicted) mean and its realized mean: $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes simulation trials.
+Root-mean-squared error is taken as the square root of the average squared deviation between each trial's estimated (or predicted) mean and its realized mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}$.
+Interval coverage is taken as the proportion of simulation trials where the realized mean is contained in its 95% confidence (or prediction) interval.
+These intervals are constructed using the normal distribution -- justification comes from the asymptotic normality of means via the central limit theorem (under some assumptions).
+Quantifying these metrics is important because together, they give us an idea of the accuracy (mean bias), spread (RMSE), and validity (interval coverage) of the sampling-inference combinations.
 
@@
 The United States Environmental Protection Agency (USEPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) to assess the water quality of various bodies of water in the contiguous United States.
-One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality [@USEPA2012NLA].
+One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality.
 We will analyze mercury concentration data collected at 986 lakes from the 2012 NLA.
 Although we can calculate the true mean mercury concentration values for these 986 lakes, here we will explore whether or not we can obtain an adequately precise estimate (design-based) or prediction (model-based) for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-For each of the four familiar sampling-analysis combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% intervals from this sample of 100 lakes and compare to the true mean mercury concentration from all 986 lakes.
+For each of the four familiar sampling-inference combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% intervals from this sample of 100 lakes and compare to the true mean mercury concentration from all 986 lakes.
 
@@
 
-The mean bias was nearly zero for all four sampling-analysis combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+The mean bias was nearly zero for all four sampling-inference combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
 Tables for mean bias in all 36 simulation scenarios are provided in the supporting information.
@@
 \begin{equation*}
-\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
+\frac{\text{rMS(P)E of sampling-inference combination}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*}
 When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-analysis combinations have approximately equal rMS(P)E.
+\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-inference combinations have approximately equal rMS(P)E.
 In these scenarios, using the GRTS algorithm or a model-based analysis does not increase efficiency compared to IRS-Design.
@@
 \ref{fig:rmspe_eff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to have the lowest rMS(P)E, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
-As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-analysis combinations widens.
+As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-inference combinations widens.
 Finally we note that when there is spatial covariance, IRS-Model has a much lower rMS(P)E than IRS-Design, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
@@
   \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative rMS(P)E in the simulation study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
   \label{fig:rmspe_eff}
@@
 
-Fig. \ref{fig:mse_eff} shows the relative MStdE of the four sampling-analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative MStdE is defined as
+Fig. \ref{fig:mse_eff} shows the relative MStdE of the four sampling-inference combinations using the random location layout with "IRS-Design" as the baseline. The relative MStdE is defined as
 \begin{equation*}
-\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
+\frac{\text{MStdE of sampling-inference combination}}{\text{MStdE of IRS-Design}},
 \end{equation*}
-Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-analysis combinations widens; and IRS-Model outperforms IRS-Design by a noticeable margin.
-These fact that the rMS(P)E and MStdE findings are similar is not particularly surprising because the mean bias for all sampling-analysis combinations was nearly zero, thus rMS(P)E is driven by the standard error of the estimators (design-based) or predictors (model-based).
+Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-inference combinations widens; and IRS-Model outperforms IRS-Design by a noticeable margin.
+These fact that the rMS(P)E and MStdE findings are similar is not particularly surprising because the mean bias for all sampling-inference combinations was nearly zero, thus rMS(P)E is driven by the standard error of the estimators (design-based) or predictors (model-based).
 We do note that between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was no spatial covariance or a medium amount of spatial covariance (Fig.
@@
 Fig.
-\ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-analysis combinations in the random location layout.
-Within each scenario, the sampling-analysis combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-Design coverage is usually a few percentage points lower than the other combinations.
+\ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-inference combinations in the random location layout.
+Within each scenario, the sampling-inference combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-Design coverage is usually a few percentage points lower than the other combinations.
 Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios usually varied from 90% to 95% but increased with the sample size.
-At a sample size of 200, all four sampling-analysis combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
+At a sample size of 200, all four sampling-inference combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
 These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
@@
   \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
-  \caption{Interval coverage in the simulation study for the four sampling-analysis combinations.
+  \caption{Interval coverage in the simulation study for the four sampling-inference combinations.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 Third, GRTS-Model has the least amount of bias and the lowest standard error (most precise interval).
-Finally, we note that for all sampling-analysis combinations, the true mean mercury concentration (103.2 ng / g) is within the bounds of the combination's 95% interval.
+Finally, we note that for all sampling-inference combinations, the true mean mercury concentration (103.2 ng / g) is within the bounds of the combination's 95% interval.
 
@@
 names(res_df_drop) <- c("Approach", "True Mean", "Est/Pred", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-analysis combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
+xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
 print(xtab_app, include.rownames = FALSE, comment = FALSE)
@@
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative rMS(P)E in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
   \label{fig:data_rmspe_eff}
@@
   \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
-  \caption{Interval coverage in the data study for the four sampling-analysis combinations.
+  \caption{Interval coverage in the data study for the four sampling-inference combinations.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 Though GRTS-Model still had lower rMS(P)E than IRS-Model, the model-based analysis mitigated most of the rMS(P)E inefficiencies that result from the IRS samples lacking spatial balance.
-Third, as the strength of spatial covariance increases, the gap in rMS(P)E and MStdE between IRS-Design and the other sampling-analysis combinations also increases, likely because IRS-Design is the only combination that ignores spatial locations in sampling and analysis.
-Fourth and finally, when the response was normal, interval coverage for all sampling-analysis combinations was usually close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling-analysis combinations was usually between 90% and 95% and closest to 95% when $n = 200$.
+Third, as the strength of spatial covariance increases, the gap in rMS(P)E and MStdE between IRS-Design and the other sampling-inference combinations also increases, likely because IRS-Design is the only combination that ignores spatial locations in sampling and analysis.
+Fourth and finally, when the response was normal, interval coverage for all sampling-inference combinations was usually close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling-inference combinations was usually between 90% and 95% and closest to 95% when $n = 200$.
 

### 2022-04-13 inst/manuscript/manuscript.tex
review edits through 2.1

@@
     approaches and then study their performance using simulated and real
-    data. In the simulated data, a variety of sample sizes, location
-    layouts, dependence structures, and response types are considered. In
-    the simulated and real data, the population mean is the parameter of
-    interest and performance is measured using statistics like bias,
-    squared error, and interval coverage.
+    data. In the simulated and real data, a variety of sample sizes,
+    location layouts, dependence structures, and response types are
+    considered. The population mean is the parameter of interest and
+    performance is measured using statistics like bias, squared error, and
+    interval coverage.
   \item
     When studying the simulated and real data, we found that regardless of
-    the strength of spatial dependence in the data, the Generalized Random
-    Tessellation Stratified (GRTS) algorithm, which explicitly
+    the strength of spatial dependence in the data, the generalized random
+    tessellation stratified (GRTS) algorithm, which explicitly
     incorporates spatial locations into sampling, tends to outperform the
-    Simple Random Sampling (SRS) algorithm, which does not explicitly
+    simple random sampling (SRS) algorithm, which does not explicitly
     incorporate spatial locations into sampling. We also found that
@@
 
-Design-based inference; Finite Population Block Kriging (FPBK);
-Generalized Random Tessellation Stratified (GRTS) algorithm; Local
+Design-based inference; Finite population block kriging (FPBK);
+Generalized random tessellation stratified (GRTS) algorithm; Local
 neighborhood variance estimator; Model-based inference; Restricted
-Maximum Likelihood (REML) estimation; Spatially balanced sampling;
+maximum likelihood (REML) estimation; Spatially balanced sampling;
 Spatial covariance
@@
 approach, inference relies on distributional assumptions about the
-underlying stochastic process that generated the sample. Each paradigm
-has a deep historical context (Sterba, 2009) and its own set of benefits
-and drawbacks (Hansen et al., 1983). In this manuscript, we compare the
-design-based and model-based approaches for finite population spatial
-data.
-
-Spatial data are data that incorporate the locations of the population
-units into either the sampling or estimation process. De Gruijter and
-Ter Braak (1990) and Brus and DeGruijter (1993) give early comparisons
-of design-based and model-based approaches for spatial data, quashing
-the belief that design-based approaches could not be used for spatially
-correlated data. Since then, there have been several general comparisons
-between design-based and model-based approaches for spatial data (Brus
-and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006)
-reviews the two approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. In addition to Cooper (2006), there has been substantial
-research and development into estimators that use both design-based and
-model-based principles (see e.g., Sterba (2009) and Cicchitelli and
-Montanari (2012), and for Bayesian approaches, see Chan-Golston et al.
-(2020) and Hofman and Brus (2021)).
-
-Certainly comparisons between design-based and model-based approaches
-have been studied in spatial contexts. Our contribution is comparing
-design-based approaches that incorporate spatial locations into sampling
-and analysis to model-based approaches. Though the broad comparisons we
-draw between design-based and model-based approaches generalize to
-finite and infinite populations, we focus on finite populations. A
-finite population contains a finite number of population units (we
-assume the finite number is known); an example is lakes (treated as a
-whole with the lake centroid representing location) in the contiguous
-United States. An infinite population contains an infinite number of
-population units; an example is locations within a single lake.
+underlying stochastic process generating the sample. Each paradigm has a
+deep historical context (Sterba, 2009) and its own set of benefits and
+drawbacks (Hansen et al., 1983, p. @brus1997random). In this manuscript,
+we compare the design-based and model-based approaches for finite
+population spatial data.
+
+Spatial data are data that have some sort of spatial index, usually via
+coordinates. De Gruijter and Ter Braak (1990) and Brus and DeGruijter
+(1993) give early comparisons of design-based and model-based approaches
+for spatial data, quashing the belief that design-based approaches could
+not be used for spatially correlated data. Since then, there have been
+several general comparisons between design-based and model-based
+approaches for spatial data (Brus and De Gruijter, 1997; Brus, 2021; Ver
+Hoef, 2002, 2008). Cooper (2006) reviews the two approaches in an
+ecological context before introducing a ``model-assisted'' variance
+estimator that combines aspects from each approach. In addition to
+Cooper (2006), there has been substantial research and development into
+estimators that use both design-based and model-based principles (see
+e.g., Sterba (2009) and Cicchitelli and Montanari (2012), and for
+Bayesian approaches, see Chan-Golston et al. (2020) and Hofman and Brus
+(2021)).
+
+While comparisons between design-based and model-based approaches have
+been studied in spatial contexts, our contribution is comparing
+design-based approaches specifically built for spatial data to
+model-based approaches. Though the broad comparisons we draw between
+design-based and model-based approaches generalize to finite and
+infinite populations, we focus on finite populations. A finite
+population contains a finite number of population units (we assume the
+finite number is known); an example is lakes (treated as a whole with
+the lake centroid representing location) in the conterminous United
+States. An infinite population contains an infinite number of population
+units; an example is locations within a single lake.
 
@@
 \ref{sec:background}, we introduce and provide relevant background for
-the design-based and model-based approaches to finite population spatial
-data. In Section \ref{sec:mm}, we describe how we compare performance of
-the approaches with a simulation study and an analysis of real data that
-contains mercury concentration in lakes located in the contiguous United
-States. In Section \ref{sec:results}, we present results from the
-simulation study and the mercury concentration analysis. And in Section
-\ref{sec:discussion}, we end with a discussion and provide directions
-for future research.
+design-based and model-based approaches to finite population spatial
+data. In Section \ref{sec:mm}, we describe how we intend to compare
+performance of the approaches using simulated and real data. In Section
+\ref{sec:results}, we present analysis reslts for the simulated and real
+data. And in Section \ref{sec:discussion}, we end with a discussion and
+provide directions for future research.
 
@@
 sampling design. A sampling design assigns a probability of selection to
-each sample (a subset of population units). Some examples of commonly
-used sampling designs include simple random sampling, stratified random
+each sample (subset of population units). Some examples of commonly used
+sampling designs include simple random sampling, stratified random
 sampling, and cluster sampling. The inclusion probability of a
-population unit follows by summing each population unit's probability of
-selection in each sample containing that population unit. Inclusion
+population unit follows by summing each sample's probability of
+selection over all samples that contain the population unit. Inclusion
 probabilities are later used to estimate population parameters.
@@
 units reflects the layout of the population units, we call the resulting
-sample ``spatially balanced.'' By ``reflecting the layout of the
-population units'', we mean that if population units are concentrated in
-specific areas, the units in the sample should be concentrated in the
-same areas. Because spatially balanced samples reflect the layout of the
-population units, they are not necessarily ``spread out'' in space in
-some equidistant manner. One approach to selecting spatially balanced
-samples is the Generalized Random Tessellation Stratified (GRTS)
-algorithm (Stevens and Olsen, 2004), which we discuss in more detail in
-Section \ref{subsec:spb_design}.
+sample spatially balanced. By ``reflecting the layout of the population
+units'', we mean that if population units are concentrated in specific
+areas, the units in the sample should be concentrated in the same areas.
+Because spatially balanced samples reflect the layout of the population
+units, they are not necessarily spread out in space in some equidistant
+manner. One approach to selecting spatially balanced samples is the
+generalized random tessellation stratified (GRTS) algorithm (Stevens and
+Olsen, 2004), which we discuss in more detail in Section
+\ref{subsec:spb_design}.
 
@@
 The model-based approach assumes the population is a random realization
-of a data-generating stochastic process called a superpopulation.
-Randomness is formally incorporated through distributional assumptions
-on the superpopulation. Strictly speaking, randomness need not be
-incorporated through random sampling, though Diggle et al. (2010) warn
-against preferential sampling. Preferential sampling occurs when the
-process generating the data locations and the process being modeled are
-not independent of one another. To guard against preferential sampling,
-model-based approaches can implement some form of random sampling,
-though it is common for model-based approaches to not implement random
-sampling. When model-based approaches do implement random sampling, the
-inclusion probabilities are ignored when analyzing the sample (in
-contrast to the design-based approach, which relies on these inclusion
-probabilities to analyze the sample).
+of a data-generating stochastic process. Randomness is formally
+incorporated through distributional assumptions on this process.
+Strictly speaking, randomness need not be incorporated through random
+sampling, though Diggle et al. (2010) warn against preferential
+sampling. Preferential sampling occurs when the process generating the
+data locations and the process being modeled are not independent of one
+another. To guard against preferential sampling, model-based approaches
+can implement some form of random sampling, though it is common for
+model-based approaches to sample non-randomly. When model-based
+approaches do implement random sampling, the inclusion probabilities are
+ignored when analyzing the sample (in contrast to the design-based
+approach, which relies on these inclusion probabilities to analyze the
+sample).
 
@@
 using assumptions of the data-generating stochastic process. If we
-repeatedly generate response values from the same process and select
+repeatedly generate realizations from the same process and select
 samples, then 95\% of all 95\% prediction intervals constructed from a
@@
 and model-based approaches (Ver Hoef (2002) and Brus (2021) provide
-similar figures).
+similar figures). This figure contrasts the design-based approach with a
+fixed population and random sampling to the model-based approach with
+random populations and non-random sampling.
 
@@
 because parameter estimates from these samples tend to vary less than
-parameter estimates from samples that are not spatially balanced
-(Barabesi and Franceschi, 2011; Benedetti et al., 2017; Grafström and
-Lundström, 2013; Robertson et al., 2013; Stevens and Olsen, 2004; Wang
-et al., 2013). The first spatially balanced sampling algorithm to see
-widespread use was the Generalized Random Tessellation Stratified (GRTS)
-algorithm (Stevens and Olsen, 2004). To quantify the spatial balance of
-a sample, Stevens and Olsen (2004) proposed loss metrics based on
-Voronoi polygons (Dirichlet Tessellations). After the GRTS algorithm was
-developed, several other spatially balanced sampling algorithms emerged,
-including stratified sampling with compact geographical strata Walvoort
-et al. (2010), the Local Pivotal Method (Grafström et al., 2012;
-Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
-(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
-2013), Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017),
-and Halton Iterative Partitioning Sampling (Robertson et al., 2018). In
-this manuscript, we select spatially balanced samples using the
-Generalized Random Tessellation Stratified (GRTS) algorithm because it
-is readily available in the \texttt{spsurvey} \textbf{\textsf{R}}
-package (Dumelle et al., 2022) and naturally accommodates finite and
-infinite sampling frames, unequal inclusion probabilities, and
-replacement units (replacement units are population units that can be
-sampled when a population unit originally selected can no longer be
-sampled).
+parameter estimates from samples lacking spatial balance (Barabesi and
+Franceschi, 2011; Benedetti et al., 2017; Grafström and Lundström, 2013;
+Robertson et al., 2013; Stevens and Olsen, 2004; Wang et al., 2013). The
+first spatially balanced sampling algorithm to see widespread use was
+the generalized random tessellation stratified (GRTS) algorithm (Stevens
+and Olsen, 2004). To quantify the spatial balance of a sample, Stevens
+and Olsen (2004) proposed loss metrics based on Voronoi polygons (i.e.,
+Dirichlet Tessellations). After the GRTS algorithm was developed,
+several other spatially balanced sampling algorithms emerged, including
+stratified sampling with compact geographical strata Walvoort et al.
+(2010), the local pivotal method (Grafström et al., 2012; Grafström and
+Matei, 2018), spatially correlated Poisson sampling (Grafström, 2012),
+balanced acceptance sampling (Robertson et al., 2013),
+within-sample-distance sampling (Benedetti and Piersimoni, 2017), and
+Halton iterative partitioning sampling (Robertson et al., 2018). In this
+manuscript, we select spatially balanced samples using the GRTS
+algorithm because it is readily available in the \texttt{spsurvey}
+\textbf{\textsf{R}} package (Dumelle et al., 2022) and naturally
+accommodates finite and infinite sampling frames, unequal inclusion
+probabilities, and replacement units. Replacement units are additional
+population units that can be sampled when a population unit originally
+selected can no longer be sampled. A couple reasons why an originally
+selected site can no longer be sampled include its location being
+physically inaccessible or on private land that the researcher does not
+have permission to access.
 
@@
 up into four distinct, equally sized squares called level-one cells.
-Each level-one is randomly assigned an level-one address of 0, 1, 2, or
-3. The set of level-one cells is denoted by \(\mathcal{A}_1\) and
+Each level-one cell is randomly assigned a level-one address of 0, 1, 2,
+or 3. The set of level-one cells is denoted by \(\mathcal{A}_1\) and
 defined as \(\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}\). Within
@@
 is the set
-\(\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}\). Because
-hierarchical ordering sorts by level-one cells, then level-two cells,
-and so on, population units that have similar hierarchical addresses
-tend to be nearby one another in space.Next each population unit is
-mapped to a one-dimensional line in hierarchical order where each
-population unit's inclusion probability equals its line-length. If a
-level-\(k\) cell has multiple population units in it, they are randomly
-placed within the cell's respective line segment. A uniform random
-variable is then simulated in \([0, 1]\) and a systematic sample is
-selected on the line, yielding \(n\) sample points for a sample size
-\(n\). Each element in this systematic sample falls on some population
-unit's line segment, and thus that population unit is selected in the
-sample. For further details regarding the GRTS algorithm, see Stevens
-and Olsen (2004).
+\linebreak \(\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}\).
+Because hierarchical ordering sorts by level-one cells, then level-two
+cells, and so on, population units that have similar hierarchical
+addresses tend to be nearby one another in space. Next each population
+unit is mapped to a one-dimensional line in hierarchical order where
+each population unit's inclusion probability equals its line-length. If
+a level-\(k\) cell has multiple population units in it, they are
+randomly placed within the cell's respective line segment. A uniform
+random variable is then simulated in \([0, 1]\) and a systematic sample
+is selected on the line, yielding \(n\) sample points for a sample size
+\(n\). Each of these sample points falls on some population unit's line
+segment, and thus that population unit is selected in the sample. For
+further details regarding the GRTS algorithm, see Stevens and Olsen
+(2004).
 
@@
 \(\hat{\tau}_{ht}\), is is given by \begin{align}\label{eq:ht}
-  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
-\end{align} where \(Z_i\) is the value of the \(i\)th population unit in
+  \hat{\tau}_{ht} = \sum_{i = 1}^n z_i \pi_i^{-1},
+\end{align} where \(z_i\) is the value of the \(i\)th population unit in
 the sample, \(\pi_i\) is the inclusion probability of the \(i\)th
@@
 being that this conditioning should yield a more precise estimate of
-\(\hat{\tau}\). They show that the each observation's contribution to
-the overall variance is dominated by local variation. Thus the local
-neighborhood variance estimator is a weighted sum of variance estimates
-from each observation's local neighborhood. These local neighborhoods
-contain observation itself and its three nearest neighbors. For more
+\(\hat{\tau}\). They show that the contribution from each sample unit
+(population unit in the sample) to the overall variance is dominated by
+local variation. Thus the local neighborhood variance estimator is a
+weighted sum of variance estimates from each sample unit's local
+neighborhood. These local neighborhoods contain the sample unit itself
+and its three nearest neighbors among all other sample units. For more
 details, see Stevens and Olsen (2003).
@@
 
-Finite Population Block Kriging (FPBK) is a model-based approach that
+Finite population block kriging (FPBK) is a model-based approach that
 expands the geostatistical Kriging framework to the finite population
@@
 stochastic process. We summarize some of the basic principles of FPBK
-next -- for technical details, see Ver Hoef (2008). Let
+next -- for more details, see Ver Hoef (2008). Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 
+In this section we describe how we used simulated and real data to
+investigate performance between simple random sampling without
+replacement (SRS) and GRTS sampling as well as performance between
+design-based (DB) and model-based (MB) inference. In SRS and GRTS
+sampling, all population units had equal inclusion probabilities. The
+important distinction between SRS and GRTS is that SRS ignores spatial
+locations while sampling but GRTS explicitly incorporates them.
+Together, the two sampling plans (SRS and GRTS) combined with the two
+inference approaches (DB and MB) yielded four sampling-inference
+combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB. For SRS-DB, the
+Horvitz-Thompson estimator \eqref{eq:ht} was used to estimate means and
+the commonly-used SRS variance formula (Lohr, 2009; Särndal et al.,
+2003) was used to estimate the variance. This variance formula is given
+by \begin{equation}\label{equation:srs_var}
+ \frac{f[\sum_{i = 1}^n (z_i - \bar{z})^2]}{n(n - 1)},
+\end{equation} where \(z_i\) is the \(i\)th response value, \(\bar{z}\)
+is the mean of all \(z_i\), \(n\) is the sample size, \(N\) is the
+population size, and \(f = (1 - n / N)\) (\(f\) is often called the
+finite population correction factor). For GRTS-DB, the Horvitz-Thompson
+esetimator was used to estimate means and the local neighborhood
+variance was used to estimate variances. For SRS-MB and GRTS-MB, FPBK
+was used to estimate means and variances and parameters were estimated
+using restricted maximum likelihood.
+
+We used simulated data to compare the sampling-inference combinations
+across many realized populations from the same data-generating
+stochastic process. With the simulated data, we were in control of the
+data-generating stochastic process and the random sampling process. We
+used real data from the 2012 National Lakes Assessment (USEPA, 2012) to
+compare the sampling-inference combinations within a single realized
+population (which is typically the case in reality). With the real data,
+we were in control of only the random sampling process.
+
 \hypertarget{sec:mm_sim}{%
@@
 
-REWRITE AS SIMPLE RANDOM SAMPLING AND WITHOUT REPLACEMENT
-
-USE SRS / GRTS - DB / MB
-
-ADD LOHR REFERENCE AND FORM FOR SRS VARIANCE WITHOUT REPLACEMENT WITH
-FPC
-
-We used a simulation study to investigate performance of four
-sampling-analysis combinations. The first sampling-analysis combination
-was IRS-Design. In IRS-Design, samples were selected with the
-Independent Random Sampling (IRS) algorithm. The IRS algorithm ignores
-the spatial locations of the population units, thus the IRS samples were
-not spatially balanced. In IRS-Design, samples were analyzed using the
-design-based approach via the Horvitz-Thompson mean estimator and an IRS
-variance estimator that ignored the spatial locations of the units in
-the sample. The second sampling-analysis combination was IRS-Model,
-where samples were selected with the IRS algorithm and analyzed using
-the model-based approach while estimating the covariance parameters
-(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
-maximum likelihood (REML). The third sampling-analysis combination was
-GRTS-Design, where samples were selected with the GRTS algorithm and
-analyzed using the design-based approach via the Horvitz-Thompson mean
-estimator and the local neighborhood variance estimator (which does
-incorporate the spatial locations of the units in the sample). The
-fourth and final sampling-analysis combination was GRTS-Model, where
-samples were selected with the GRTS algorithm and analyzed using the
-model-based approach while estimating the covariance parameters
-(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
-maximum likelihood (REML). These sampling-analysis combinations are also
-provided in Table \ref{tab:designanalysis}. Lastly we note that for both
-the IRS and GRTS samples, equal inclusion probabilities were assumed for
-all population units. When IRS assumes equal inclusion probabilities for
-all population units, the algorithm is equivalent to simple random
-sampling (SRS).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{r|ll}
-  \hline
- & Design & Model \\ 
-  \hline
-IRS & IRS-Design & IRS-Model \\ 
-  GRTS & GRTS-Design & GRTS-Model \\ 
-   \hline
-\end{tabular}
-\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
-\end{table}
-
 CHANGE LOGNORMAL VERBAGE TO SKEWED -- look for DRE acronym
 
-Performance for the four sampling-analysis combinations was evaluated in
+We evaluated performance of the four sampling-inference combinations in
 36 different simulation scenarios. The 36 scenarios resulted from the
@@
 layout were placed on a fixed, equally spaced grid inside the unit
-square. The two response types were normal and lognormal. For the normal
+square. The two response types were normal and skewed. For the normal
 response type, the response was simulated using mean-zero random errors
 with the exponential covariance (Equation\(~\)\ref{equation:expcov}) for
-varying proportions of dependent random error. The proportion of
-dependent random error is represented by
+three proportions of dependent random error (DRE): 0\% DRE, 50\% DRE,
+and 90\% DRE. Recall the proportion of DRE is represented by
 \(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where \(\sigma^2_1\) and
-\(\sigma^2_2\) are the dependent random error variance (partial sill)
-and independent random error variance (nugget) from
-Equation\(~\)\ref{equation:expcov}, respectively. The total variance,
-\(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
-\(\sqrt{2} / 3\), chosen so that the correlation in the dependent random
-error decayed to nearly zero at \(\sqrt{2}\), the largest possible
-distance between two population units in the domain. For the lognormal
+\(\sigma^2_2\) are the DRE variance and independent random error (IRE)
+variance from Equation\(~\)\ref{equation:expcov}, respectively. The
+total variance, \(\sigma^2_1 + \sigma^2_2\), was always 2. The distance
+parameter was always \(\sqrt{2} / 3\), chosen so that the correlation in
+the DRE decayed to nearly zero at \(\sqrt{2}\), the largest possible
+distance between two population units in the domain. For the skewed
 response type, the response was first simulated using the same approach
@@
 0.6931 instead of 2. The response was then exponentiated, yielding a
-lognormal random variable whose total variance was 2. The lognormal
-responses were used to evaluate performance of the sampling-analysis
-approaches for data that were skewed (i.e., not normal).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{r|lll}
-   \hline
-Sample Size (n) & 50 & 100 & 200 \\ 
-  Location Layout & Random & Gridded & - \\ 
-  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
-  Response Type & Normal & Lognormal & - \\ 
-   \hline
-\end{tabular}
-\caption{\label{tab:parmtab} Simulation scenario options.
-All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was 2.}
-\end{table}
-
-In each of the 36 simulation scenarios, there were 2000 independent
-simulation trials. In each trial, IRS and GRTS samples were selected and
-then design-based and model-based analyses were used to estimate
-(design-based) or predict (model-based) the mean and construct 95\%
-confidence (design-based) or 95\% prediction (model-based) intervals.
-Then we recorded the bias, squared error, standard error, and interval
-coverage for all sampling-analysis combinations. After all 2000 trials,
-we summarized the long-run performance of the combinations by
-calculating mean bias, rMS(P)E (root-mean-squared error for the
-design-based approaches and root-mean-squared-prediction error for the
-model-based approaches), MStdE (mean standard error), and the proportion
-of times the true mean is contained in its 95\% confidence
-(design-based) or 95\% prediction (model-based) interval. The 95\%
-intervals were constructed using the normal distribution. Justification
-for this comes from the asymptotic normality of means via the Central
-Limit Theorem (under some assumptions). Quantifying mean bias and
-rMS(P)E is important because they help us understand how far (under
-different loss metrics) the estimates (design-based) or predictions
-(model-based) tend to be from the true mean. Quantifying MStdE is
-important because it helps us understand how precise intervals tend to
-be. Quantifying interval coverage is important because it helps us
-understand how often our 95\% intervals actually contain the true mean.
-
-The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
-neighborhood variance estimator are available in the \texttt{spsurvey}
-\textbf{\textsf{R}} package (Dumelle et al., 2022). FPBK is available in
-the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).
+skewed random variable whose total variance was 2. The skewed responses
+were used to evaluate performance of the sampling-inference approaches
+for data that were not normal but were still estimated using REML, which
+relies on a normal log-likelihood. Figure \ref{fig:sim_pops} shows an
+example of a realized population for the normal and skewed responses
+using the random layout and 50\% DRE.
 
@@
   \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
-  \caption{Histogram of a single realized population for the skewed response.}
+  \caption{Histogram of a realized population for the skewed response.}
   \label{fig:skew_pop_hist}
 \end{subfigure}
-\caption{Histograms realized populations for the simulated data.}
+\caption{Histograms of realized populations simulated for the normal and skewed resposnes using the random layout and 50\% DRE.}
 \label{fig:sim_pops}
 \end{figure}
+
+In each of the 36 simulation scenarios, there were 2000 independent
+simulation trials. Within each simulation scenario and trial, IRS and
+GRTS samples were selected and then design-based and model-based
+analyses were used to estimate (design-based) or predict (model-based)
+the mean and construct 95\% confidence (design-based) or 95\% prediction
+(model-based) intervals. With the model-based analyses, covariance
+parameters were estimated (using REML) separately for each trial. After
+all 2000 trials, we summarized the long-run performance of the
+sampling-inference combination in each scenario by calculating mean
+bias, root-mean-squared error, and interval coverage. Mean bias is taken
+as the average deviation between each trial's estimated (or predicted)
+mean and its realized mean:
+\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
+indexes simulation trials. Root-mean-squared error is taken as the
+square root of the average squared deviation between each trial's
+estimated (or predicted) mean and its realized mean:
+\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}\).
+Interval coverage is taken as the proportion of simulation trials where
+the realized mean is contained in its 95\% confidence (or prediction)
+interval. These intervals are constructed using the normal distribution
+-- justification comes from the asymptotic normality of means via the
+central limit theorem (under some assumptions). Quantifying these
+metrics is important because together, they give us an idea of the
+accuracy (mean bias), spread (RMSE), and validity (interval coverage) of
+the sampling-inference combinations.
 
@@
 United States. One component of NARS is the National Lakes Assessment
-(NLA), which measures various aspects of lake health and water quality
-(USEPA, 2012). We will analyze mercury concentration data collected at
-986 lakes from the 2012 NLA. Although we can calculate the true mean
-mercury concentration values for these 986 lakes, here we will explore
-whether or not we can obtain an adequately precise estimate
-(design-based) or prediction (model-based) for the realized mean mercury
-concentration if we sample only 100 of the 986 lakes. For each of the
-four familiar sampling-analysis combinations (IRS-Design, IRS-Model,
-GRTS-Design, and GRTS-Model), we estimate (design-based) or predict
-(model-based) the mean mercury concentration and construct 95\%
-intervals from this sample of 100 lakes and compare to the true mean
-mercury concentration from all 986 lakes.
+(NLA), which measures various aspects of lake health and water quality.
+We will analyze mercury concentration data collected at 986 lakes from
+the 2012 NLA. Although we can calculate the true mean mercury
+concentration values for these 986 lakes, here we will explore whether
+or not we can obtain an adequately precise estimate (design-based) or
+prediction (model-based) for the realized mean mercury concentration if
+we sample only 100 of the 986 lakes. For each of the four familiar
+sampling-inference combinations (IRS-Design, IRS-Model, GRTS-Design, and
+GRTS-Model), we estimate (design-based) or predict (model-based) the
+mean mercury concentration and construct 95\% intervals from this sample
+of 100 lakes and compare to the true mean mercury concentration from all
+986 lakes.
 
@@
 
-The mean bias was nearly zero for all four sampling-analysis
+The mean bias was nearly zero for all four sampling-inference
 combinations in all 36 scenarios, so we omit a more detailed summary of
@@
 \begin{equation*}
-\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
+\frac{\text{rMS(P)E of sampling-inference combination}}{\text{rMS(P)E of IRS-Design}},
 \end{equation*} When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
-combinations have approximately equal rMS(P)E and using the GRTS
-algorithm or a model-based analysis does not result in much, if any,
-loss in efficiency compared to IRS-Design. When there is spatial
-covariance (Fig. \ref{fig:rmspe_eff}, ``Prop DE: 0.5'' and ``Prop DE:
-0.9'' rows), GRTS-Model tends to have the lowest rMS(P)E, followed by
-GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in
-relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is
-relatively small. As the strength of spatial covariance increases, the
-gap in rMS(P)E between IRS-Design and the other sampling-analysis
-combinations widens. Finally we note that when there is spatial
-covariance, IRS-Model has a much lower rMS(P)E than IRS-Design,
-suggesting that the poor design properties of IRS are largely mitigated
-by the model-based analysis. These rMS(P)E conclusions are similar to
-those observed in the grid location layout, so we omit a grid location
-layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
-are provided in the supporting information.
+\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-inference
+combinations have approximately equal rMS(P)E. In these scenarios, using
+the GRTS algorithm or a model-based analysis does not increase
+efficiency compared to IRS-Design. When there is spatial covariance
+(Fig. \ref{fig:rmspe_eff}, ``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows),
+GRTS-Model tends to have the lowest rMS(P)E, followed by GRTS-Design,
+IRS-Model, and finally IRS-Design, though the difference in relative
+rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively
+small. As the strength of spatial covariance increases, the gap in
+rMS(P)E between IRS-Design and the other sampling-inference combinations
+widens. Finally we note that when there is spatial covariance, IRS-Model
+has a much lower rMS(P)E than IRS-Design, suggesting that the poor
+design properties of IRS are largely mitigated by the model-based
+analysis. These rMS(P)E conclusions are similar to those observed in the
+grid location layout, so we omit a grid location layout figure here.
+Tables for rMS(P)E in all 36 simulation scenarios are provided in the
+supporting information.
 
@@
 Fig. \ref{fig:mse_eff} shows the relative MStdE of the four
-sampling-analysis combinations using the random location layout with
+sampling-inference combinations using the random location layout with
 ``IRS-Design'' as the baseline. The relative MStdE is defined as
 \begin{equation*}
-\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
+\frac{\text{MStdE of sampling-inference combination}}{\text{MStdE of IRS-Design}},
 \end{equation*} Many general takeaways regarding MStdE are similar to
@@
 spatial covariance increases, the gap in MStdE between IRS-Design and
-the other sampling-analysis combinations widens; and IRS-Model
+the other sampling-inference combinations widens; and IRS-Model
 outperforms IRS-Design by a noticeable margin. These fact that the
 rMS(P)E and MStdE findings are similar is not particularly surprising
-because the mean bias for all sampling-analysis combinations was nearly
+because the mean bias for all sampling-inference combinations was nearly
 zero, thus rMS(P)E is driven by the standard error of the estimators
@@
 Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
-four sampling-analysis combinations in the random location layout.
-Within each scenario, the sampling-analysis combinations tend to have
+four sampling-inference combinations in the random location layout.
+Within each scenario, the sampling-inference combinations tend to have
 fairly similar interval coverage, though when \(n = 50\) or \(n = 100\),
@@
 usually varied from 90\% to 95\% but increased with the sample size. At
-a sample size of 200, all four sampling-analysis combinations had
+a sample size of 200, all four sampling-inference combinations had
 approximately 95\% interval coverage in both response scenarios for all
@@
 has the least amount of bias and the lowest standard error (most precise
-interval). Finally, we note that for all sampling-analysis combinations,
-the true mean mercury concentration (103.2 ng / g) is within the bounds
-of the combination's 95\% interval.
+interval). Finally, we note that for all sampling-inference
+combinations, the true mean mercury concentration (103.2 ng / g) is
+within the bounds of the combination's 95\% interval.
 
@@
 \end{tabular}
-\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.}
+\caption{\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.}
 \end{table}
@@
 increases, the gap in rMS(P)E and MStdE between IRS-Design and the other
-sampling-analysis combinations also increases, likely because IRS-Design
-is the only combination that ignores spatial locations in sampling and
-analysis. Fourth and finally, when the response was normal, interval
-coverage for all sampling-analysis combinations was usually close to
-95\% for all sample sizes; when the response was lognormal, interval
-coverage for all sampling-analysis combinations was usually between 90\%
-and 95\% and closest to 95\% when \(n = 200\).
+sampling-inference combinations also increases, likely because
+IRS-Design is the only combination that ignores spatial locations in
+sampling and analysis. Fourth and finally, when the response was normal,
+interval coverage for all sampling-inference combinations was usually
+close to 95\% for all sample sizes; when the response was lognormal,
+interval coverage for all sampling-inference combinations was usually
+between 90\% and 95\% and closest to 95\% when \(n = 200\).
+
+AT LEAST HAVE DISCUSSION ABOUT MODEL BASED ASSUMPTIONS AND MOVE VALIDITY
+COMMENTS TO RESULTS SECTION.
 
@@
 researcher, while the model-based approach may not yield valid results
-if the assumptions made do not not accurately capture reailty. The
+if the assumptions made do not not accurately capture reality. The
 model-based approach, however, can more naturally quantify the
@@
 
-\leavevmode\hypertarget{ref-higham2021sptotal}{}%
-Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
-Predicting totals and weighted sums from spatial data.
-
 \leavevmode\hypertarget{ref-hofman2021many}{}%

### 2022-04-13 inst/manuscript/manuscript.Rmd
section 2 and supporting updates

@@
 
-## National Lakes Assessment Data {#sec:mm_app}
+## National Lakes Assessment (Real) Data {#sec:mm_app}
 
@@
 One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality.
-We will analyze mercury concentration data collected at 986 lakes from the 2012 NLA.
-Although we can calculate the true mean mercury concentration values for these 986 lakes, here we will explore whether or not we can obtain an adequately precise estimate (design-based) or prediction (model-based) for the realized mean mercury concentration if we sample only 100 of the 986 lakes.
-For each of the four familiar sampling-inference combinations (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model), we estimate (design-based) or predict (model-based) the mean mercury concentration and construct 95% intervals from this sample of 100 lakes and compare to the true mean mercury concentration from all 986 lakes.
+We focus on analyzing zooplankton multi-metric indices (ZMMI) and mercury concentrations in parts per billion (Hg ppb) from the 2012 NLA.
+For ZMMI, data were collected at 1035 unique lakes.
+At less than 10% of lakes, two ZMMI replicates were collected.
+These were averaged for the purposes of our study so that each lake had one measurement for ZMMI.
+For Hg ppb, data were collected at 995 unique lakes (and there were no replicates like for ZMMI).
+The ZMMI and Hg ppb data are shown spatially and as a histogram in Figure \ref{fig:zmmi}.
+The ZMMI data are relatively symmetric, and the Hg ppb data are skewed.
+Also in Figure \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
+The spatial semivariogram quantifies the average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
+The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
+Both ZMMI and Hg ppb seem to have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as the semivariance increases steadily with distance (meaning that observations nearby one another tend to be more similar than observations far apart from one another).
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
+  \caption{Spatial map of the ZMMI population.}
+  \label{fig:zmmi_map}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
+  \caption{Spatial map of mercury (Hg ppb) population.}
+  \label{fig:mercury_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
+  \caption{Histogram of the ZMMI population.}
+  \label{fig:zmmi_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
+  \caption{Histogram of the mercury (Hg ppb) population.}
+  \label{fig:mercury_hist}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
+  \caption{Semivariogram of the ZMMI population.}
+  \label{fig:zmmi_sv_plot}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
+  \caption{Semivariogram of the mercury (Hg ppb) population.}
+  \label{fig:mercury_sv_plot}
+\end{subfigure}
+\caption{Exploratory graphics of the ZMMI and mercury (Hg ppb) populations in the National Lakes Assessment (NLA) 2012 data.}
+\label{fig:zmmi}
+\end{figure}
+
+We studied performance of the four sampling-inference combinations by selecting 2000 random IRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
+We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
+Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean and the population mean (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials and $\mu$ is the population mean.
+Root-mean-squared error is taken as the square root of the average squared deviation between each sample's estimated (or predicted) mean and its population mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}$.
+Interval coverage is taken as the proportion of simulation trials where the population mean is contained in its 95% confidence (or prediction) interval.
+These intervals are constructed using the normal distribution.
 
@@
   \centering
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
-  \caption{Spatial map of the ZMMI population.}
-  \label{fig:zmmi_map}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
-  \caption{Spatial map of mercury (Hg) population.}
-  \label{fig:mercury_map}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
-  \caption{Histogram of the ZMMI population.}
-  \label{fig:zmmi_hist}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
-  \caption{Histogram of the mercury (Hg) population.}
-  \label{fig:mercury_hist}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
-  \caption{Semivariogram of the ZMMI population.}
-  \label{fig:zmmi_sv_plot}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
-  \caption{Semivariogram of the mercury (Hg) population.}
-  \label{fig:mercury_sv_plot}
-\end{subfigure}
-\caption{Exploratory graphics of the ZMMI and mercury (Hg) populations in the National Lakes Assessment (NLA) 2012 data.}
-\label{fig:zmmi}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}

### 2022-04-13 inst/manuscript/manuscript.tex
section 2 and supporting updates

@@
 \hypertarget{sec:mm_app}{%
-\subsection{National Lakes Assessment Data}\label{sec:mm_app}}
+\subsection{National Lakes Assessment (Real) Data}\label{sec:mm_app}}
 
@@
 (NLA), which measures various aspects of lake health and water quality.
-We will analyze mercury concentration data collected at 986 lakes from
-the 2012 NLA. Although we can calculate the true mean mercury
-concentration values for these 986 lakes, here we will explore whether
-or not we can obtain an adequately precise estimate (design-based) or
-prediction (model-based) for the realized mean mercury concentration if
-we sample only 100 of the 986 lakes. For each of the four familiar
-sampling-inference combinations (IRS-Design, IRS-Model, GRTS-Design, and
-GRTS-Model), we estimate (design-based) or predict (model-based) the
-mean mercury concentration and construct 95\% intervals from this sample
-of 100 lakes and compare to the true mean mercury concentration from all
-986 lakes.
+We focus on analyzing zooplankton multi-metric indices (ZMMI) and
+mercury concentrations in parts per billion (Hg ppb) from the 2012 NLA.
+For ZMMI, data were collected at 1035 unique lakes. At less than 10\% of
+lakes, two ZMMI replicates were collected. These were averaged for the
+purposes of our study so that each lake had one measurement for ZMMI.
+For Hg ppb, data were collected at 995 unique lakes (and there were no
+replicates like for ZMMI). The ZMMI and Hg ppb data are shown spatially
+and as a histogram in Figure \ref{fig:zmmi}. The ZMMI data are
+relatively symmetric, and the Hg ppb data are skewed. Also in Figure
+\ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
+The spatial semivariogram quantifies the average squared differences
+(semivariance) of responses whose separation (distance) falls within
+some distance class. The spatial semivariance is closely related to the
+spatial covariance, and spatial semivariograms are often used to gauge
+the strength of spatial dependence in data. Both ZMMI and Hg ppb seem to
+have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as
+the semivariance increases steadily with distance (meaning that
+observations nearby one another tend to be more similar than
+observations far apart from one another).
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
+  \caption{Spatial map of the ZMMI population.}
+  \label{fig:zmmi_map}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
+  \caption{Spatial map of mercury (Hg ppb) population.}
+  \label{fig:mercury_map}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
+  \caption{Histogram of the ZMMI population.}
+  \label{fig:zmmi_hist}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
+  \caption{Histogram of the mercury (Hg ppb) population.}
+  \label{fig:mercury_hist}
+\end{subfigure} \\
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
+  \caption{Semivariogram of the ZMMI population.}
+  \label{fig:zmmi_sv_plot}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
+  \caption{Semivariogram of the mercury (Hg ppb) population.}
+  \label{fig:mercury_sv_plot}
+\end{subfigure}
+\caption{Exploratory graphics of the ZMMI and mercury (Hg ppb) populations in the National Lakes Assessment (NLA) 2012 data.}
+\label{fig:zmmi}
+\end{figure}
+
+We studied performance of the four sampling-inference combinations by
+selecting 2000 random IRS and GRTS samples of size \(n = 50\),
+\(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
+populations and then analyzing the samples using MB and DB inference. We
+used the same evaluation metrics as for the simulated data: mean bias,
+RMSE, and interval coverage. Mean bias is taken as the average deviation
+between each sample's estimated (or predicted) mean and the population
+mean (of ZMMI or Hg ppb):
+\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
+indexes simulation trials and \(\mu\) is the population mean.
+Root-mean-squared error is taken as the square root of the average
+squared deviation between each sample's estimated (or predicted) mean
+and its population mean:
+\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
+Interval coverage is taken as the proportion of simulation trials where
+the population mean is contained in its 95\% confidence (or prediction)
+interval. These intervals are constructed using the normal distribution.
 
@@
 \subsection{New Application}\label{new-application}}
-
-\begin{figure}
-\centering
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
-  \caption{Spatial map of the ZMMI population.}
-  \label{fig:zmmi_map}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
-  \caption{Spatial map of mercury (Hg) population.}
-  \label{fig:mercury_map}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
-  \caption{Histogram of the ZMMI population.}
-  \label{fig:zmmi_hist}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
-  \caption{Histogram of the mercury (Hg) population.}
-  \label{fig:mercury_hist}
-\end{subfigure} \\
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
-  \caption{Semivariogram of the ZMMI population.}
-  \label{fig:zmmi_sv_plot}
-\end{subfigure}
-\begin{subfigure}{0.49\textwidth}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
-  \caption{Semivariogram of the mercury (Hg) population.}
-  \label{fig:mercury_sv_plot}
-\end{subfigure}
-\caption{Exploratory graphics of the ZMMI and mercury (Hg) populations in the National Lakes Assessment (NLA) 2012 data.}
-\label{fig:zmmi}
-\end{figure}
 

### 2022-04-13 inst/manuscript/supporting.Rmd
section 2 and supporting updates

@@
 
-# Simulation study tables {#sec:simtabs}
-
 ```{r, include = FALSE}
@@
 
-Recall the four sampling-analysis combinations from the simulation study, summarized in Table \ref{tab:designanalysis}.
-Also recall the 36 parameter configurations from the simulation study, summarized in Table \ref{tab:parmtab}.
+Recall the four sampling-analysis combinations from the simulated and real data: simple random sampling with design-based inference (SRS-DB), simple random sapling with model-based inference (SRS-MB), GRTS sampling with design-based inference (GRTS-DB), and GRTS sampling with model-based inference (GRTS-MB).
+
+# Simulated Data {#sec:simtabs}
+
+For the simulated data, we considered 36 parameter configurations -- the crossing of three sample sizes ($n = 50$, $n = 100$, $n = 200)$, two location layouts (random and gridded), three proportions of dependent random error (DRE) (0%, 50%, 90%), and two response types (normal and skewed).
 
 ```{r, echo = FALSE, results = "asis"}
-tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-       nrow = 2, byrow = TRUE) %>%
-  as.table()
-row.names(tab) <- c("IRS", "GRTS")
-xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-analysis combinations in the simulation study.
+# tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
+#        nrow = 2, byrow = TRUE) %>%
+#   as.table()
+# row.names(tab) <- c("IRS", "GRTS")
+# xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-analysis combinations in the simulation study.
 The rows give the two types of sampling designs and the columns give the two types of analyses.")
-align(xtab1) <- "r|ll"
-names(xtab1) <- c("Design", "Model")
-print(xtab1, comment = FALSE)
-## I really want "Analysis" to appear as a row header and "Sampling Design" 
-## to appear as a column header, but this isn't easy to figure out.
-## I couldn't figure this out with xtable either so it might need to be
-## made "by hand."
+# align(xtab1) <- "r|ll"
+# names(xtab1) <- c("Design", "Model")
+# print(xtab1, comment = FALSE)
+# ## I really want "Analysis" to appear as a row header and "Sampling Design" 
+# ## to appear as a column header, but this isn't easy to figure out.
+# ## I couldn't figure this out with xtable either so it might need to be
+# ## made "by hand."
 ```
@@
 ```{r, echo = FALSE, results = "asis"}
-n <- c("50", "100", "200")
-site_locations <- c("Random", "Gridded", "")
-psill_ratio <- c("0", "0.5", "0.9")
-resp_type <- c("Normal", "Lognormal", "")
-parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-parm_tab[c(2, 4), 3] <- "-"
-row.names(parm_tab) <- c("Sample Size (n)", "Layout", 
-                         "Proportion of Dependent Error", "Response Type")
-# library(knitr) kable with htlm has trouble with page breaks -- kable with latex harder for me than xtable
-# kable(parm_tab, col.names = c("", "", ""),
-#       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
+# n <- c("50", "100", "200")
+# site_locations <- c("Random", "Gridded", "")
+# psill_ratio <- c("0", "0.5", "0.9")
+# resp_type <- c("Normal", "Lognormal", "")
+# parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
+# parm_tab[c(2, 4), 3] <- "-"
+# row.names(parm_tab) <- c("Sample Size (n)", "Layout", 
+#                          "Proportion of Dependent Error", "Response Type")
+# # library(knitr) kable with htlm has trouble with page breaks -- kable with latex harder for me than xtable
+# # kable(parm_tab, col.names = c("", "", ""),
+# #       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
+# parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
 All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
 In each simualtion scenario, the total variance was 2.")
-align(parm_xtab) <- "r|lll"
-print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
+# align(parm_xtab) <- "r|lll"
+# print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
 ```
 
-Next we present tables summarizing mean estimation bias (design-based) or mean prediction bias (model-based), root-mean-squared error (design-based) or root-mean-squared-prediction error (model-based), mean standard errors, and 95% confidence interval coverage (design-based) or 95% prediction interval coverage (model-based) for all 36 simulation scenarios.
+Next we present tables summarizing mean bias, RMSE, and interval coverage for all 36 simulation scenarios.
 ```{r}
 make_sim_table <- function(table_value) {
+  # browser()
   sim_table <- read_csv(here("inst", "output", "simulation_summary", paste0("allsims", table_value, ".csv")))
   sim_table <- sim_table %>%
+    mutate(psill_ratio = paste0(100 * psill_ratio, "%")) %>%
     dplyr::select(Approach = approach,
@@
                   Response = resptype,
-                  PDE = psill_ratio,
+                  `DRE%` = psill_ratio,
                   n = n,
-                  Bias = mpbias,
-                  `rMS(P)E` = rmspe,
-                  MStdE = mse,
+                  MB = mpbias,
+                  RMSE = rmspe,
                   Coverage = coverage
@@
      Approach = case_when(
-       Approach == "Design GRTS" ~ "GRTS-Design",
-       Approach == "Design IRS" ~ "IRS-Design",
-       Approach == "Model GRTS" ~ "GRTS-Model",
-       Approach == "Model IRS" ~ "IRS-Model"
+       Approach == "Design GRTS" ~ "GRTS-DB",
+       Approach == "Design IRS" ~ "GRTS-MB",
+       Approach == "Model GRTS" ~ "SRS-DB",
+       Approach == "Model IRS" ~ "SRS-MB"
      ),
@@
        Response == "normal" ~ "Normal",
-       Response == "lognormal" ~ "Lognormal"
+       Response == "lognormal" ~ "Skewed"
      ),
-     across(c(PDE, n), as.character)
+     across(c(`DRE%`, n), as.character)
     )
-sim_table <- sim_table[c(2, 4, 1, 3), ]
+sim_table <- sim_table[c(3, 4, 1, 2), ]
 sim_table_out <- xtable(
@@
   digits = 4,
-  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
+  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
    type = "latex",
   latex.environments = "center",
-  align = c("llllrrrrrr")
+  align = c("llllrrrrr")
 )
@@
 ```
+
+# National Lakes Assessment (Real) Data {#sec:datatabs}
+
+Next we present tables summarizing mean bias, RMSE, and interval coverage for ZMMI and Hg ppb data.
+```{r}
+make_data_table <- function(table_value) {
+  # browser()
+  data_table <- read_csv(here("inst", "output", "data_summary", paste0("datasims", table_value, ".csv")))
+  data_table <- data_table %>%
+    dplyr::select(Approach = approach,
+                  Response = var,
+                  n = n,
+                  MB = mpbias,
+                  RMSE = rmspe,
+                  Coverage = coverage
+                  ) %>%
+    mutate(
+     Approach = case_when(
+       Approach == "Design GRTS" ~ "GRTS-DB",
+       Approach == "Design IRS" ~ "GRTS-MB",
+       Approach == "Model GRTS" ~ "SRS-DB",
+       Approach == "Model IRS" ~ "SRS-MB"
+     ),
+     Response = case_when(
+       Response == "MMI_ZOOP_NLA6" ~ "ZMMI",
+       Response == "TOTALHG_RESULT" ~ "Hg ppb"
+     ),
+     across(c(n), as.character)
+    )
+data_table <- data_table[c(3, 4, 1, 2), ]
+data_table_out <- xtable(
+  data_table,
+  digits = 4,
+  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\\% interval coverage (Coverage) in a data application scenario ", table_value, "."),
+   type = "latex",
+  latex.environments = "center",
+  align = c("lllrrrr")
+)
+print(data_table_out, include.rownames = FALSE, comment = FALSE, scalebox = "0.8")
+}
+```
+
+```{r, results = "asis"}
+for (i in 1:6) {
+  make_data_table(i)
+}
+```

### 2022-04-13 inst/manuscript/supporting.tex
section 2 and supporting updates

--- 2021-12-21 inst/manuscript/supporting.tex
@@
 
+Recall the four sampling-analysis combinations from the simulated and
+real data: simple random sampling with design-based inference (SRS-DB),
+simple random sapling with model-based inference (SRS-MB), GRTS sampling
+with design-based inference (GRTS-DB), and GRTS sampling with
+model-based inference (GRTS-MB).
+
 \hypertarget{sec:simtabs}{%
-\section{Simulation Study Tables}\label{sec:simtabs}}
+\section{Simulated Data}\label{sec:simtabs}}
 
-Recall the four sampling-analysis combinations from the simulation
-study, summarized in Table \ref{tab:designanalysis}. Also recall the 36
-parameter configurations from the simulation study, summarized in Table
-\ref{tab:parmtab}.
+For the simulated data, we considered 36 parameter configurations -- the
+crossing of three sample sizes (\(n = 50\), \(n = 100\), \(n = 200)\),
+two location layouts (random and gridded), three proportions of
+dependent random error (DRE) (0\%, 50\%, 90\%), and two response types
+(normal and skewed).
+
+Next we present tables summarizing mean bias, RMSE, and interval
+coverage for all 36 simulation scenarios.
 
@@
 \centering
-\begin{tabular}{r|ll}
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
   \hline
- & Design & Model \\ 
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
   \hline
-IRS & IRS-Design & IRS-Model \\ 
-  GRTS & GRTS-Design & GRTS-Model \\ 
+SRS-DB & Gridded & Normal & 0\% & 50 & 0.0004 & 0.1922 & 0.9345 \\ 
+  SRS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1930 & 0.9420 \\ 
+  GRTS-DB & Gridded & Normal & 0\% & 50 & 0.0005 & 0.1914 & 0.9170 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1905 & 0.9490 \\ 
    \hline
 \end{tabular}
-\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 1.}
 \end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 0\% & 50 & 0.0004 & 0.1965 & 0.8955 \\ 
+  SRS-MB & Gridded & Skewed & 0\% & 50 & -0.0110 & 0.2055 & 0.9045 \\ 
+  GRTS-DB & Gridded & Skewed & 0\% & 50 & 0.0025 & 0.1947 & 0.8710 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 50 & -0.0048 & 0.1988 & 0.9110 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 2.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 0\% & 50 & -0.0002 & 0.1964 & 0.9295 \\ 
+  SRS-MB & Random & Normal & 0\% & 50 & 0.0045 & 0.1953 & 0.9395 \\ 
+  GRTS-DB & Random & Normal & 0\% & 50 & -0.0004 & 0.1951 & 0.9065 \\ 
+  GRTS-MB & Random & Normal & 0\% & 50 & 0.0031 & 0.1940 & 0.9425 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 3.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 0\% & 50 & 0.0016 & 0.1941 & 0.8985 \\ 
+  SRS-MB & Random & Skewed & 0\% & 50 & -0.0059 & 0.1932 & 0.9130 \\ 
+  GRTS-DB & Random & Skewed & 0\% & 50 & 0.0041 & 0.1910 & 0.8740 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 50 & 0.0000 & 0.1861 & 0.9225 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 4.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 50\% & 50 & -0.0031 & 0.1472 & 0.9400 \\ 
+  SRS-MB & Gridded & Normal & 50\% & 50 & -0.0007 & 0.1617 & 0.9355 \\ 
+  GRTS-DB & Gridded & Normal & 50\% & 50 & -0.0030 & 0.1501 & 0.9205 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 50 & -0.0013 & 0.1775 & 0.9460 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 5.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 50\% & 50 & -0.0060 & 0.1605 & 0.9085 \\ 
+  SRS-MB & Gridded & Skewed & 50\% & 50 & -0.0083 & 0.1787 & 0.8990 \\ 
+  GRTS-DB & Gridded & Skewed & 50\% & 50 & -0.0041 & 0.1610 & 0.8995 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 50 & -0.0047 & 0.1891 & 0.9100 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 6.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 50\% & 50 & -0.0097 & 0.1507 & 0.9315 \\ 
+  SRS-MB & Random & Normal & 50\% & 50 & 0.0042 & 0.1615 & 0.9350 \\ 
+  GRTS-DB & Random & Normal & 50\% & 50 & -0.0092 & 0.1513 & 0.9165 \\ 
+  GRTS-MB & Random & Normal & 50\% & 50 & 0.0061 & 0.1744 & 0.9445 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 7.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 50\% & 50 & -0.0021 & 0.1624 & 0.9000 \\ 
+  SRS-MB & Random & Skewed & 50\% & 50 & -0.0051 & 0.1808 & 0.9065 \\ 
+  GRTS-DB & Random & Skewed & 50\% & 50 & 0.0021 & 0.1618 & 0.8880 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 50 & 0.0014 & 0.1804 & 0.9195 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 8.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 90\% & 50 & -0.0037 & 0.0960 & 0.9400 \\ 
+  SRS-MB & Gridded & Normal & 90\% & 50 & 0.0044 & 0.1137 & 0.9435 \\ 
+  GRTS-DB & Gridded & Normal & 90\% & 50 & -0.0021 & 0.1072 & 0.9280 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 50 & 0.0055 & 0.1621 & 0.9355 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 9.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 90\% & 50 & -0.0039 & 0.1039 & 0.9205 \\ 
+  SRS-MB & Gridded & Skewed & 90\% & 50 & -0.0011 & 0.1217 & 0.9130 \\ 
+  GRTS-DB & Gridded & Skewed & 90\% & 50 & -0.0017 & 0.1129 & 0.9180 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 50 & 0.0007 & 0.1546 & 0.9345 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 10.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 90\% & 50 & -0.0011 & 0.0940 & 0.9410 \\ 
+  SRS-MB & Random & Normal & 90\% & 50 & 0.0045 & 0.1093 & 0.9445 \\ 
+  GRTS-DB & Random & Normal & 90\% & 50 & -0.0017 & 0.1054 & 0.9295 \\ 
+  GRTS-MB & Random & Normal & 90\% & 50 & 0.0058 & 0.1575 & 0.9470 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 11.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1127 & 0.9225 \\ 
+  SRS-MB & Random & Skewed & 90\% & 50 & -0.0046 & 0.1301 & 0.9180 \\ 
+  GRTS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1250 & 0.9175 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 50 & -0.0022 & 0.1772 & 0.9305 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 12.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 0\% & 100 & -0.0010 & 0.1352 & 0.9445 \\ 
+  SRS-MB & Gridded & Normal & 0\% & 100 & -0.0026 & 0.1362 & 0.9405 \\ 
+  GRTS-DB & Gridded & Normal & 0\% & 100 & -0.0011 & 0.1345 & 0.9210 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 100 & -0.0028 & 0.1352 & 0.9475 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 13.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 0\% & 100 & 0.0077 & 0.1307 & 0.9155 \\ 
+  SRS-MB & Gridded & Skewed & 0\% & 100 & -0.0024 & 0.1412 & 0.9150 \\ 
+  GRTS-DB & Gridded & Skewed & 0\% & 100 & 0.0082 & 0.1304 & 0.8945 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 100 & -0.0001 & 0.1384 & 0.9210 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 14.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 0\% & 100 & 0.0019 & 0.1318 & 0.9460 \\ 
+  SRS-MB & Random & Normal & 0\% & 100 & -0.0007 & 0.1344 & 0.9435 \\ 
+  GRTS-DB & Random & Normal & 0\% & 100 & 0.0021 & 0.1315 & 0.9280 \\ 
+  GRTS-MB & Random & Normal & 0\% & 100 & -0.0004 & 0.1339 & 0.9450 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 15.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 0\% & 100 & 0.0039 & 0.1315 & 0.9230 \\ 
+  SRS-MB & Random & Skewed & 0\% & 100 & 0.0005 & 0.1342 & 0.9280 \\ 
+  GRTS-DB & Random & Skewed & 0\% & 100 & 0.0043 & 0.1311 & 0.8990 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 100 & 0.0021 & 0.1322 & 0.9280 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 16.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 50\% & 100 & -0.0003 & 0.0993 & 0.9435 \\ 
+  SRS-MB & Gridded & Normal & 50\% & 100 & -0.0007 & 0.1058 & 0.9425 \\ 
+  GRTS-DB & Gridded & Normal & 50\% & 100 & 0.0000 & 0.1003 & 0.9300 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 100 & -0.0025 & 0.1207 & 0.9420 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 17.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 50\% & 100 & -0.0019 & 0.1127 & 0.9175 \\ 
+  SRS-MB & Gridded & Skewed & 50\% & 100 & 0.0002 & 0.1135 & 0.9160 \\ 
+  GRTS-DB & Gridded & Skewed & 50\% & 100 & 0.0000 & 0.1113 & 0.9025 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 100 & 0.0023 & 0.1221 & 0.9315 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 18.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 50\% & 100 & -0.0010 & 0.0997 & 0.9380 \\ 
+  SRS-MB & Random & Normal & 50\% & 100 & -0.0001 & 0.1079 & 0.9440 \\ 
+  GRTS-DB & Random & Normal & 50\% & 100 & -0.0012 & 0.1012 & 0.9235 \\ 
+  GRTS-MB & Random & Normal & 50\% & 100 & -0.0016 & 0.1214 & 0.9440 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 19.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 50\% & 100 & -0.0002 & 0.1105 & 0.9235 \\ 
+  SRS-MB & Random & Skewed & 50\% & 100 & -0.0020 & 0.1097 & 0.9290 \\ 
+  GRTS-DB & Random & Skewed & 50\% & 100 & 0.0013 & 0.1104 & 0.9090 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 100 & 0.0020 & 0.1188 & 0.9395 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 20.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0602 & 0.9425 \\ 
+  SRS-MB & Gridded & Normal & 90\% & 100 & -0.0003 & 0.0701 & 0.9440 \\ 
+  GRTS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0670 & 0.9325 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 100 & -0.0027 & 0.1083 & 0.9460 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 21.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 90\% & 100 & 0.0004 & 0.0695 & 0.9280 \\ 
+  SRS-MB & Gridded & Skewed & 90\% & 100 & -0.0038 & 0.0767 & 0.9265 \\ 
+  GRTS-DB & Gridded & Skewed & 90\% & 100 & 0.0018 & 0.0746 & 0.9270 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 100 & -0.0015 & 0.1108 & 0.9345 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 22.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 90\% & 100 & -0.0007 & 0.0570 & 0.9480 \\ 
+  SRS-MB & Random & Normal & 90\% & 100 & 0.0011 & 0.0675 & 0.9475 \\ 
+  GRTS-DB & Random & Normal & 90\% & 100 & 0.0009 & 0.0651 & 0.9430 \\ 
+  GRTS-MB & Random & Normal & 90\% & 100 & 0.0002 & 0.1071 & 0.9490 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 23.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 90\% & 100 & 0.0023 & 0.0671 & 0.9230 \\ 
+  SRS-MB & Random & Skewed & 90\% & 100 & -0.0031 & 0.0757 & 0.9315 \\ 
+  GRTS-DB & Random & Skewed & 90\% & 100 & 0.0026 & 0.0751 & 0.9225 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 100 & -0.0019 & 0.1165 & 0.9325 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 24.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0843 & 0.9610 \\ 
+  SRS-MB & Gridded & Normal & 0\% & 200 & 0.0003 & 0.0894 & 0.9440 \\ 
+  GRTS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0842 & 0.9520 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 200 & 0.0005 & 0.0891 & 0.9465 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 25.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 0\% & 200 & 0.0005 & 0.0892 & 0.9405 \\ 
+  SRS-MB & Gridded & Skewed & 0\% & 200 & -0.0023 & 0.0895 & 0.9445 \\ 
+  GRTS-DB & Gridded & Skewed & 0\% & 200 & 0.0006 & 0.0891 & 0.9360 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 200 & -0.0018 & 0.0894 & 0.9440 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 26.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 0\% & 200 & 0.0009 & 0.0882 & 0.9480 \\ 
+  SRS-MB & Random & Normal & 0\% & 200 & 0.0017 & 0.0907 & 0.9375 \\ 
+  GRTS-DB & Random & Normal & 0\% & 200 & 0.0008 & 0.0881 & 0.9400 \\ 
+  GRTS-MB & Random & Normal & 0\% & 200 & 0.0019 & 0.0905 & 0.9385 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 27.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9395 \\ 
+  SRS-MB & Random & Skewed & 0\% & 200 & 0.0005 & 0.0869 & 0.9390 \\ 
+  GRTS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9345 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 200 & 0.0007 & 0.0867 & 0.9400 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 28.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 50\% & 200 & 0.0039 & 0.0679 & 0.9375 \\ 
+  SRS-MB & Gridded & Normal & 50\% & 200 & -0.0004 & 0.0678 & 0.9510 \\ 
+  GRTS-DB & Gridded & Normal & 50\% & 200 & 0.0036 & 0.0688 & 0.9330 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 200 & -0.0016 & 0.0777 & 0.9520 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 29.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 50\% & 200 & 0.0018 & 0.0691 & 0.9360 \\ 
+  SRS-MB & Gridded & Skewed & 50\% & 200 & 0.0016 & 0.0747 & 0.9400 \\ 
+  GRTS-DB & Gridded & Skewed & 50\% & 200 & 0.0019 & 0.0698 & 0.9325 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 200 & 0.0013 & 0.0816 & 0.9400 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 30.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 50\% & 200 & 0.0009 & 0.0643 & 0.9510 \\ 
+  SRS-MB & Random & Normal & 50\% & 200 & 0.0005 & 0.0713 & 0.9444 \\ 
+  GRTS-DB & Random & Normal & 50\% & 200 & 0.0005 & 0.0655 & 0.9474 \\ 
+  GRTS-MB & Random & Normal & 50\% & 200 & 0.0008 & 0.0816 & 0.9389 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 31.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 50\% & 200 & -0.0007 & 0.0679 & 0.9385 \\ 
+  SRS-MB & Random & Skewed & 50\% & 200 & -0.0048 & 0.0729 & 0.9350 \\ 
+  GRTS-DB & Random & Skewed & 50\% & 200 & -0.0001 & 0.0680 & 0.9400 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 200 & -0.0041 & 0.0789 & 0.9510 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 32.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0376 & 0.9475 \\ 
+  SRS-MB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0418 & 0.9445 \\ 
+  GRTS-DB & Gridded & Normal & 90\% & 200 & -0.0010 & 0.0418 & 0.9455 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 200 & -0.0011 & 0.0730 & 0.9450 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 33.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Gridded & Skewed & 90\% & 200 & 0.0010 & 0.0423 & 0.9375 \\ 
+  SRS-MB & Gridded & Skewed & 90\% & 200 & -0.0007 & 0.0475 & 0.9365 \\ 
+  GRTS-DB & Gridded & Skewed & 90\% & 200 & 0.0018 & 0.0482 & 0.9370 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 200 & 0.0038 & 0.0768 & 0.9430 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 34.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Normal & 90\% & 200 & 0.0014 & 0.0368 & 0.9475 \\ 
+  SRS-MB & Random & Normal & 90\% & 200 & 0.0011 & 0.0408 & 0.9480 \\ 
+  GRTS-DB & Random & Normal & 90\% & 200 & 0.0016 & 0.0418 & 0.9450 \\ 
+  GRTS-MB & Random & Normal & 90\% & 200 & 0.0026 & 0.0713 & 0.9495 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 35.}
+\end{table}
+\begin{table}[ht]
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lllrrrrr}
+  \hline
+Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+  \hline
+SRS-DB & Random & Skewed & 90\% & 200 & -0.0001 & 0.0396 & 0.9419 \\ 
+  SRS-MB & Random & Skewed & 90\% & 200 & 0.0011 & 0.0437 & 0.9414 \\ 
+  GRTS-DB & Random & Skewed & 90\% & 200 & -0.0008 & 0.0431 & 0.9504 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 200 & 0.0028 & 0.0708 & 0.9504 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 36.}
+\end{table}
+
+\hypertarget{sec:datatabs}{%
+\section{National Lakes Assessment (Real) Data}\label{sec:datatabs}}
+
+Next we present tables summarizing mean bias, RMSE, and interval
+coverage for ZMMI and Hg ppb data.
 
@@
 \centering
-\begin{tabular}{r|lll}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Sample Size (n) & 50 & 100 & 200 \\ 
-  Layout & Random & Gridded & - \\ 
-  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
-  Response Type & Normal & Lognormal & - \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-\end{tabular}
-\caption{\label{tab:parmtab} Simulation scenario options.
-All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was 2.}
-\end{table}
-
-Next we present tables summarizing mean estimation bias (design-based)
-or mean prediction bias (model-based), root-mean-squared error
-(design-based) or root-mean-squared-prediction error (model-based), and
-95\% confidence interval coverage (design-based) or 95\% prediction
-interval coverage (model-based) for all 36 simulation scenarios.
-
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0 & 50 & -0.0023 & 0.1968 & 0.9440 \\ 
-  IRS-Model & Gridded & Normal & 0 & 50 & -0.0029 & 0.1988 & 0.9400 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0011 & 0.1946 & 0.9110 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0006 & 0.1949 & 0.9300 \\ 
+SRS-DB & Hg ppb & 50 & -0.5604 & 10.4434 & 0.9070 \\ 
+  SRS-MB & Hg ppb & 50 & 0.0573 & 11.3760 & 0.9020 \\ 
+  GRTS-DB & Hg ppb & 50 & -0.1880 & 10.5558 & 0.8970 \\ 
+  GRTS-MB & Hg ppb & 50 & 0.2539 & 12.9467 & 0.9215 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 1.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 1.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0 & 50 & -0.0036 & 0.2007 & 0.9080 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0094 & 0.2071 & 0.9045 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 50 & -0.0007 & 0.1962 & 0.8690 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0022 & 0.1969 & 0.8945 \\ 
+SRS-DB & ZMMI & 50 & -0.1191 & 1.9608 & 0.9355 \\ 
+  SRS-MB & ZMMI & 50 & 0.0000 & 2.1477 & 0.9395 \\ 
+  GRTS-DB & ZMMI & 50 & -0.0047 & 1.9953 & 0.9210 \\ 
+  GRTS-MB & ZMMI & 50 & 0.0461 & 2.4502 & 0.9440 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 2.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 2.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-IRS-Design & Random & Normal & 0 & 50 & 0.0019 & 0.1861 & 0.9505 \\ 
-  IRS-Model & Random & Normal & 0 & 50 & 0.0014 & 0.1892 & 0.9445 \\ 
-  GRTS-Design & Random & Normal & 0 & 50 & -0.0040 & 0.1955 & 0.9090 \\ 
-  GRTS-Model & Random & Normal & 0 & 50 & -0.0040 & 0.1965 & 0.9315 \\ 
+SRS-DB & Hg ppb & 100 & -0.4156 & 7.0344 & 0.9230 \\ 
+  SRS-MB & Hg ppb & 100 & -0.4213 & 7.4820 & 0.9190 \\ 
+  GRTS-DB & Hg ppb & 100 & 0.0075 & 7.1415 & 0.9045 \\ 
+  GRTS-MB & Hg ppb & 100 & -0.1390 & 8.9667 & 0.9290 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 3.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 3.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-IRS-Design & Random & Lognormal & 0 & 50 & 0.0032 & 0.1843 & 0.9205 \\ 
-  IRS-Model & Random & Lognormal & 0 & 50 & -0.0039 & 0.1945 & 0.9105 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0056 & 0.1944 & 0.8870 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0078 & 0.1967 & 0.9075 \\ 
+SRS-DB & ZMMI & 100 & -0.0968 & 1.2958 & 0.9450 \\ 
+  SRS-MB & ZMMI & 100 & -0.0942 & 1.3920 & 0.9445 \\ 
+  GRTS-DB & ZMMI & 100 & -0.0066 & 1.3219 & 0.9205 \\ 
+  GRTS-MB & ZMMI & 100 & -0.0541 & 1.6901 & 0.9460 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 4.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 4.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Normal & 0.5 & 50 & 0.0025 & 0.1762 & 0.9470 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 50 & 0.0007 & 0.1655 & 0.9305 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0004 & 0.1524 & 0.9115 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 50 & -0.0003 & 0.1499 & 0.9320 \\ 
+SRS-DB & Hg ppb & 200 & -0.2762 & 4.3998 & 0.9410 \\ 
+  SRS-MB & Hg ppb & 200 & -0.1775 & 4.6970 & 0.9400 \\ 
+  GRTS-DB & Hg ppb & 200 & -0.0538 & 4.5051 & 0.9325 \\ 
+  GRTS-MB & Hg ppb & 200 & -0.0876 & 5.9176 & 0.9450 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 5.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 5.}
 \end{table}
@@
 \centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
+\scalebox{0.8}{
+\begin{tabular}{llrrrr}
   \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
+Approach & Response & n & MB & RMSE & Coverage \\ 
   \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0004 & 0.1863 & 0.9140 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0076 & 0.1834 & 0.9035 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0016 & 0.1612 & 0.8810 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0017 & 0.1606 & 0.8940 \\ 
+SRS-DB & ZMMI & 200 & -0.0311 & 0.8280 & 0.9450 \\ 
+  SRS-MB & ZMMI & 200 & -0.0736 & 0.9090 & 0.9450 \\ 
+  GRTS-DB & ZMMI & 200 & 0.0193 & 0.8524 & 0.9460 \\ 
+  GRTS-MB & ZMMI & 200 & -0.0093 & 1.1311 & 0.9455 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 6.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.5 & 50 & 0.0037 & 0.1657 & 0.9590 \\ 
-  IRS-Model & Random & Normal & 0.5 & 50 & 0.0027 & 0.1546 & 0.9495 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 50 & -0.0034 & 0.1511 & 0.9170 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 50 & -0.0037 & 0.1504 & 0.9305 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 7.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0023 & 0.1637 & 0.9245 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0055 & 0.1622 & 0.9125 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 50 & -0.0057 & 0.1570 & 0.9000 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0079 & 0.1567 & 0.9100 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 8.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0.9 & 50 & 0.0053 & 0.1579 & 0.9470 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 50 & 0.0026 & 0.1165 & 0.9315 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 50 & 0.0013 & 0.1074 & 0.9220 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 50 & -0.0007 & 0.0949 & 0.9430 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 9.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1731 & 0.9220 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0020 & 0.1325 & 0.9135 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1183 & 0.9065 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0002 & 0.1090 & 0.9120 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 10.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.9 & 50 & 0.0062 & 0.1520 & 0.9525 \\ 
-  IRS-Model & Random & Normal & 0.9 & 50 & 0.0037 & 0.1074 & 0.9525 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0029 & 0.1038 & 0.9340 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0027 & 0.0940 & 0.9360 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 11.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.9 & 50 & 0.0053 & 0.1542 & 0.9325 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0012 & 0.1160 & 0.9115 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0033 & 0.1150 & 0.9160 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0031 & 0.1014 & 0.9235 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 12.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0 & 100 & -0.0029 & 0.1322 & 0.9430 \\ 
-  IRS-Model & Gridded & Normal & 0 & 100 & -0.0033 & 0.1332 & 0.9410 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0041 & 0.1302 & 0.9245 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0038 & 0.1304 & 0.9420 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 13.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0 & 100 & -0.0052 & 0.1334 & 0.9335 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 100 & -0.0072 & 0.1350 & 0.9300 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0042 & 0.1303 & 0.9035 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0034 & 0.1307 & 0.9265 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 14.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0 & 100 & -0.0066 & 0.1366 & 0.9405 \\ 
-  IRS-Model & Random & Normal & 0 & 100 & -0.0067 & 0.1370 & 0.9385 \\ 
-  GRTS-Design & Random & Normal & 0 & 100 & 0.0028 & 0.1364 & 0.9180 \\ 
-  GRTS-Model & Random & Normal & 0 & 100 & 0.0029 & 0.1363 & 0.9345 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 15.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0 & 100 & -0.0051 & 0.1348 & 0.9280 \\ 
-  IRS-Model & Random & Lognormal & 0 & 100 & -0.0075 & 0.1372 & 0.9230 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0002 & 0.1375 & 0.8840 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 100 & -0.0003 & 0.1377 & 0.9090 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 16.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0016 & 0.1177 & 0.9540 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0023 & 0.1072 & 0.9470 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 100 & 0.0051 & 0.0998 & 0.9300 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 100 & 0.0039 & 0.0982 & 0.9470 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 17.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 100 & -0.0021 & 0.1211 & 0.9445 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 100 & -0.0058 & 0.1153 & 0.9295 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0041 & 0.1090 & 0.9085 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0019 & 0.1090 & 0.9180 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 18.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.5 & 100 & -0.0064 & 0.1222 & 0.9440 \\ 
-  IRS-Model & Random & Normal & 0.5 & 100 & -0.0049 & 0.1073 & 0.9440 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0013 & 0.1041 & 0.9155 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0014 & 0.1020 & 0.9400 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 19.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0063 & 0.1204 & 0.9355 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0097 & 0.1150 & 0.9275 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 100 & 0.0003 & 0.1092 & 0.8960 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0004 & 0.1088 & 0.9120 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 20.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0.9 & 100 & -0.0007 & 0.1059 & 0.9605 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 100 & -0.0019 & 0.0700 & 0.9445 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0044 & 0.0655 & 0.9435 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 100 & 0.0030 & 0.0585 & 0.9440 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 21.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 100 & -0.0011 & 0.1092 & 0.9540 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0033 & 0.0779 & 0.9330 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0035 & 0.0757 & 0.9190 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0017 & 0.0693 & 0.9190 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 22.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.9 & 100 & -0.0048 & 0.1095 & 0.9430 \\ 
-  IRS-Model & Random & Normal & 0.9 & 100 & -0.0017 & 0.0671 & 0.9535 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0674 & 0.9265 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 100 & 0.0000 & 0.0595 & 0.9420 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 23.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0045 & 0.1110 & 0.9375 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0052 & 0.0771 & 0.9310 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 100 & -0.0009 & 0.0732 & 0.9195 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0007 & 0.0661 & 0.9165 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 24.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0 & 200 & 0.0013 & 0.0882 & 0.9550 \\ 
-  IRS-Model & Gridded & Normal & 0 & 200 & 0.0013 & 0.0886 & 0.9530 \\ 
-  GRTS-Design & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9380 \\ 
-  GRTS-Model & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.9460 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 25.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0001 & 0.0880 & 0.9350 \\ 
-  IRS-Model & Gridded & Lognormal & 0 & 200 & -0.0002 & 0.0883 & 0.9355 \\ 
-  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0034 & 0.0898 & 0.9285 \\ 
-  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0035 & 0.0900 & 0.9335 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 26.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0 & 200 & -0.0014 & 0.0893 & 0.9465 \\ 
-  IRS-Model & Random & Normal & 0 & 200 & -0.0015 & 0.0896 & 0.9465 \\ 
-  GRTS-Design & Random & Normal & 0 & 200 & 0.0007 & 0.0868 & 0.9460 \\ 
-  GRTS-Model & Random & Normal & 0 & 200 & 0.0007 & 0.0867 & 0.9490 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 27.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0 & 200 & -0.0028 & 0.0891 & 0.9425 \\ 
-  IRS-Model & Random & Lognormal & 0 & 200 & -0.0033 & 0.0896 & 0.9395 \\ 
-  GRTS-Design & Random & Lognormal & 0 & 200 & 0.0015 & 0.0860 & 0.9365 \\ 
-  GRTS-Model & Random & Lognormal & 0 & 200 & 0.0012 & 0.0861 & 0.9415 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 28.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0.5 & 200 & 0.0022 & 0.0788 & 0.9535 \\ 
-  IRS-Model & Gridded & Normal & 0.5 & 200 & 0.0008 & 0.0678 & 0.9580 \\ 
-  GRTS-Design & Gridded & Normal & 0.5 & 200 & -0.0024 & 0.0671 & 0.9335 \\ 
-  GRTS-Model & Gridded & Normal & 0.5 & 200 & -0.0030 & 0.0661 & 0.9410 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 29.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0.5 & 200 & 0.0016 & 0.0816 & 0.9420 \\ 
-  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0000 & 0.0739 & 0.9335 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0033 & 0.0744 & 0.9330 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.0742 & 0.9350 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 30.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.5 & 200 & -0.0016 & 0.0790 & 0.9480 \\ 
-  IRS-Model & Random & Normal & 0.5 & 200 & -0.0010 & 0.0690 & 0.9475 \\ 
-  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0002 & 0.0652 & 0.9455 \\ 
-  GRTS-Model & Random & Normal & 0.5 & 200 & 0.0001 & 0.0640 & 0.9500 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 31.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.5 & 200 & -0.0027 & 0.0809 & 0.9494 \\ 
-  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0037 & 0.0732 & 0.9454 \\ 
-  GRTS-Design & Random & Lognormal & 0.5 & 200 & -0.0008 & 0.0671 & 0.9434 \\ 
-  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0012 & 0.0662 & 0.9479 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 32.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0021 & 0.0704 & 0.9570 \\ 
-  IRS-Model & Gridded & Normal & 0.9 & 200 & 0.0005 & 0.0411 & 0.9590 \\ 
-  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0005 & 0.0423 & 0.9370 \\ 
-  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0012 & 0.0377 & 0.9450 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 33.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0014 & 0.0744 & 0.9520 \\ 
-  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0002 & 0.0480 & 0.9355 \\ 
-  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0014 & 0.0499 & 0.9380 \\ 
-  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0022 & 0.0459 & 0.9380 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 34.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Normal & 0.9 & 200 & -0.0009 & 0.0725 & 0.9470 \\ 
-  IRS-Model & Random & Normal & 0.9 & 200 & -0.0010 & 0.0409 & 0.9450 \\ 
-  GRTS-Design & Random & Normal & 0.9 & 200 & -0.0005 & 0.0405 & 0.9490 \\ 
-  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0001 & 0.0354 & 0.9510 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 35.}
-\end{table}
-\begin{table}[ht]
-\centering
-\scalebox{0.85}{
-\begin{tabular}{lllrrrrr}
-  \hline
-Approach & Layout & Response & DE Prop & n & Bias & rMS(P)E & Coverage \\ 
-  \hline
-IRS-Design & Random & Lognormal & 0.9 & 200 & -0.0006 & 0.0750 & 0.9393 \\ 
-  IRS-Model & Random & Lognormal & 0.9 & 200 & -0.0025 & 0.0461 & 0.9418 \\ 
-  GRTS-Design & Random & Lognormal & 0.9 & 200 & -0.0013 & 0.0452 & 0.9348 \\ 
-  GRTS-Model & Random & Lognormal & 0.9 & 200 & -0.0012 & 0.0397 & 0.9418 \\ 
-   \hline
-\end{tabular}
-}
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DE Prop), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 36.}
+\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 6.}
 \end{table}

### 2022-04-13 inst/scripts/figures/data.R

@@ -90,7 +90,7 @@ combo_data <- combo_data %>%
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 # resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
-resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg ppb")
 rmspe_eff <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = rel_efficiency, colour = approach)) +
   facet_grid(
@@ -103,7 +103,7 @@ rmspe_eff <- combo_data %>%
   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
   scale_y_continuous(breaks = c(0.5, 0.75, 1), labels = c("0.50", "0.75", "1.00"), limits = c(0.5, 1.1)) +
   theme_bw(base_size = base_size) +
-  labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative RMSE") +
   theme(
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
@@ -178,7 +178,7 @@ if (write_out) {
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 # resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
-resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg")
+resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: ZMMI", TOTALHG_RESULT = "Response: Hg ppb")
 coverage <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = coverage, colour = approach)) +
   geom_hline(yintercept = 0.95, lty = "dashed") +

### 2022-04-13 inst/scripts/figures/data_exploratory.R

@@ -54,7 +54,7 @@ mercury_map <- nla_nomiss_both %>%
   ggplot(aes(x = xcoords, y = ycoords)) +
   geom_point(aes(colour = total_hg)) +
   scale_colour_viridis_c(breaks = c(250, 500, 750)) +
-  labs(x = "", y = "", colour = "Hg") +
+  labs(x = "", y = "", colour = "Hg ppb") +
   coord_quickmap() +
   theme_bw(base_size = 18) +
   theme(
@@ -82,7 +82,7 @@ if (write_out) {
 mercury_hist <- nla_nomiss_both %>%
   ggplot(aes(x = total_hg)) +
   geom_histogram(bins = 20) +
-  labs(x = "Hg", y = "Count") +
+  labs(x = "Hg ppb", y = "Count") +
   theme_bw(base_size = 18) +
   theme(
     panel.grid.major = element_blank(),
@@ -107,7 +107,7 @@ sv_val <- sv(nla_nomiss_both, "xcoords", "ycoords", "total_hg") # same as residu
 sv_plot <- sv_val %>%
   ggplot(aes(x = dist, y = gamma * (1/1000))) +
   geom_point(size = 3) +
-  labs(x = "Distance (Kilometers)", y = "Hg Semi-Variance (Thousands)", title = "") +
+  labs(x = "Distance (Kilometers)", y = "Hg ppb Semi-Var (Thousands)", title = "") +
   ylim(c(0, 15)) +
   # scale_size_continuous(name = "Pairs") +
   theme_bw(base_size = 18) +
@@ -230,7 +230,7 @@ sv_val <- sv(nla_nomiss_both, "xcoords", "ycoords", "MMI_ZOOP_NLA6") # same as r
 sv_plot <- sv_val %>%
   ggplot(aes(x = dist, y = gamma * (1/10))) +
   geom_point(size = 3) +
-  labs(x = "Distance (Kilometers)", y = "ZMMI Semi-Variance (Tens)", title = "") +
+  labs(x = "Distance (Kilometers)", y = "ZMMI Semi-Var (Tens)", title = "") +
   ylim(c(0, 45)) +
   # scale_size_continuous(name = "Pairs") +
   theme_bw(base_size = 18) +

### 2022-04-13 inst/scripts/figures/simulation.R

@@ -65,7 +65,7 @@ combo_data <- combo_data %>%
 # colour_scale <- palette(hcl.colors(4, "viridis"))
 colour_scale <- unname(palette.colors(palette = "Okabe-Ito")[1:4])
 resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Skewed")
-psill_ratio_labs <- c("0" = "Prop DRE: 0", "0.5" = "Prop DRE: 0.5", "0.9" = "Prop DRE: 0.9")
+psill_ratio_labs <- c("0" = "DRE%: 0%", "0.5" = "DRE%: 50%", "0.9" = "DRE%: 90%")
 rmspe_eff <- combo_data %>%
   ggplot(aes(x = n_factor_adj, y = rel_efficiency, colour = approach)) +
   facet_grid(
@@ -78,7 +78,7 @@ rmspe_eff <- combo_data %>%
   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
   scale_y_continuous(breaks = c(0.5, 0.75, 1), labels = c("0.50", "0.75", "1.00"), limits = c(0.5, 1.1)) +
   theme_bw(base_size = base_size) +
-  labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
+  labs(x = "Sample Size", colour = "Approach", y = "Relative RMSE") +
   # lims(y = c(0.5, 1.1)) +
   theme(
     panel.grid.major = element_blank(),

### 2022-04-13 inst/manuscript/manuscript.Rmd
section 3.1 updates

@@
 ```{r, echo = FALSE, results = "asis"}
-# tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
+# tab <- matrix(c("SRS-DB", "SRS-MB", "GRTS-DB", "GRTS-MB"),
 #        nrow = 2, byrow = TRUE) %>%
@@
 ```
-
-CHANGE LOGNORMAL VERBAGE TO SKEWED -- look for DRE acronym
 
@@
 For Hg ppb, data were collected at 995 unique lakes (and there were no replicates like for ZMMI).
-The ZMMI and Hg ppb data are shown spatially and as a histogram in Figure \ref{fig:zmmi}.
-The ZMMI data are relatively symmetric, and the Hg ppb data are skewed.
+The ZMMI and Hg ppb data are shown as spatial maps and as histograms in Figure \ref{fig:zmmi}.
+The ZMMI data tend to be highest near the coasts, lowest in the Central United States, higher near the coasts, are relatively symmetric, and have a mean of 55.05.
+The Hg ppb data tend to be highest in the Northeastern United States, lowest elsewhere, are skewed, and have a mean of 103.16 ppb.
 Also in Figure \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
-The spatial semivariogram quantifies the average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
+The spatial semivariogram quantifies the the halved average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
 The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
@@
 We studied performance of the four sampling-inference combinations by selecting 2000 random IRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
+In total, there were six separate scenarios (two responses and three sample sizes).
 We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
@@
 
-The mean bias was nearly zero for all four sampling-inference combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+Mean bias was nearly zero for all four sampling-inference combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
 Tables for mean bias in all 36 simulation scenarios are provided in the supporting information.
 
-Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling analysis combinations using the random location layout with "IRS-Design" as the baseline. The relative rMS(P)E is defined as
-\begin{equation*}
-\frac{\text{rMS(P)E of sampling-inference combination}}{\text{rMS(P)E of IRS-Design}},
-\end{equation*}
+We define the relative RMSE as a ratio with numerator given by the RMSE for a sampling-inference combination and the denominator given by the RMSE for SRS-DB.
+Relative RMSEs for the random location layout are provided in Fig.
+\ref{fig:rmspe_eff}.
 When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0" row), the four sampling-inference combinations have approximately equal rMS(P)E.
-In these scenarios, using the GRTS algorithm or a model-based analysis does not increase efficiency compared to IRS-Design.
+\ref{fig:rmspe_eff}, "DRE%: 0%"), the four sampling-inference combinations have approximately equal RMSE.
+In these scenarios, using GRTS sampling or model-based inference does not generally increase efficiency compared to SRS-DB.
 When there is spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "Prop DE: 0.5" and "Prop DE: 0.9" rows), GRTS-Model tends to have the lowest rMS(P)E, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively small.
-As the strength of spatial covariance increases, the gap in rMS(P)E between IRS-Design and the other sampling-inference combinations widens.
-Finally we note that when there is spatial covariance, IRS-Model has a much lower rMS(P)E than IRS-Design, suggesting that the poor design properties of IRS are largely mitigated by the model-based analysis.
-These rMS(P)E conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
-Tables for rMS(P)E in all 36 simulation scenarios are provided in the supporting information.
+\ref{fig:rmspe_eff}, "DRE%: 0%" and "DRE%: 50%"), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB, SRS-MB, and finally SRS-DB, though the difference in relative RMSE among GRTS-MB, GRTS-DB, and SRS-MB is small.
+As the strength of spatial covariance increases, the gap in RMSE between SRS-DB and the other sampling-inference combinations widens.
+Finally we note that when there is spatial covariance, SRS-MB has a much lower RMSE than SRS-DB, suggesting that the lack of efficiency from SRS is largely mitigated by model-based inference.
+These RMSE conclusions are similar to those observed in the grid location layout, so we omit a figure and discussion regarding the grid location layout here.
+Tables for RMSE in all 36 simulation scenarios are provided in the supporting information.
 \begin{figure}
@@
   \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E in the simulation study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative RMSE in the simulation study for the four sampling-inference combinations and three sample sizes in the random location layout.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black lines separate the sample sizes}
   \label{fig:rmspe_eff}
@@
 
-Fig. \ref{fig:mse_eff} shows the relative MStdE of the four sampling-inference combinations using the random location layout with "IRS-Design" as the baseline. The relative MStdE is defined as
-\begin{equation*}
-\frac{\text{MStdE of sampling-inference combination}}{\text{MStdE of IRS-Design}},
-\end{equation*}
-Many general takeaways regarding MStdE are similar to general takeaways regarding rMS(P)E: there seems to be no benefit to using IRS, even when there is no spatial covariance; as the strength of spatial covariance increases, the gap in MStdE between IRS-Design and the other sampling-inference combinations widens; and IRS-Model outperforms IRS-Design by a noticeable margin.
-These fact that the rMS(P)E and MStdE findings are similar is not particularly surprising because the mean bias for all sampling-inference combinations was nearly zero, thus rMS(P)E is driven by the standard error of the estimators (design-based) or predictors (model-based).
-We do note that between GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was no spatial covariance or a medium amount of spatial covariance (Fig.
-\ref{fig:mse_eff}, "Prop DE: 0" and "Prop DE: 0.5" rows), and GRTS-Model had lower MStdE when there was a high amount of spatial covariance (Fig.
-\ref{fig:mse_eff}, "Prop DE: 0.9" row).
-These MStdE conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
-Tables for MStdE in all 36 simulation scenarios are provided in the supporting information.
-
 Fig.
-\ref{fig:figconf} shows the 95% interval coverage for each of the four sampling-inference combinations in the random location layout.
-Within each scenario, the sampling-inference combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-Design coverage is usually a few percentage points lower than the other combinations.
-Coverage in the normal response scenarios was usually near 95%, while coverage in the lognormal response scenarios usually varied from 90% to 95% but increased with the sample size.
+\ref{fig:figconf} shows 95% interval coverage for each of the four sampling-inference combinations in the random location layout.
+Within each simulation scenario, all sampling-inference combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-DB coverage is usually a few percentage points lower than the other combinations, which suggests that the local neighborhood variance estimate may be slightly too small for small $n$.
+Coverage in the normal response scenarios was usually near 95%, while coverage in the skewed response scenarios usually varied from 90% to 95% but increased with the sample size.
 At a sample size of 200, all four sampling-inference combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
-These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a grid location layout figure here.
+These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a figure and discussion regarding the grid location layout here.
 Tables for interval coverage in all 36 simulation scenarios are provided in the supporting information.
@@
   \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
-  \caption{Interval coverage in the simulation study for the four sampling-inference combinations.
+  \caption{Interval coverage in the simulation study for the four sampling-inference combinations and three sample sizes in the random location layout.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line represents 95\% coverage.}
+The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
   \label{fig:figconf}
@@
 
-## National Lakes Assessment DAta {#sec:r_app}
+## National Lakes Assessment (Real) Data {#sec:r_app}
 
-USE MERCURY UNITS
+Mean bias was nearly zero for all four sampling-inference combinations in all six scenarios, so we omit a more detailed summary of those results here.
+Tables for mean bias in all six simulations scenarios are provided in the supporting information.
 
-Fig.
-\ref{fig:merc} shows a map and histogram of mercury concentration in all 986 NLA lakes.
-The map shows mercury concentration exhibits some spatial patterning, with high mercury concentrations in the northeast and north central United States.
-The histogram shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration.
-Fig.
-\ref{fig:merc} also shows mercury concentration's empirical semivariogram.
-The empirical semivariogram can be used as a tool to visualize spatial dependence.
-It quantifies the mean of the halved squared differences (semivariance) among all pairs of mercury concentrations at different distances apart.
-When a process has spatial covariance (exhibits spatial dependence), the mean semivariance tends to be smaller at small distances and larger at large distances.
-The empirical semivariogram in Fig.
-\ref{fig:merc} suggests that mercury concentration exhibits spatial dependence.
-Lastly we note that the true mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
+Talk about RMSE
 
-We selected a single IRS sample and a single GRTS sample and estimated (design-based) or predicted (model-based) the mean mercury concentration and constructed 95% confidence (design-based) and 95% (model-based) prediction intervals.
-For the model-based analyses, the exponential covariance was used.
-Table \ref{tab:appliedtab} shows the results from these analyses.
-Though we should not generalize these results to other samples from this population, we do mention a few findings.
-First, IRS-Design has the largest standard error.
-Second, compared to IRS-Design and IRS-Model, GRTS-Design and GRTS-Model are much closer to the true mean mercury concentration (have bias closer to zero) and have much lower standard errors (more precise intervals).
-Third, GRTS-Model has the least amount of bias and the lowest standard error (most precise interval).
-Finally, we note that for all sampling-inference combinations, the true mean mercury concentration (103.2 ng / g) is within the bounds of the combination's 95% interval.
+Talk about interval coverage -- mention how the model-based assumptions of coverage are not really satisfied here, but they can be thought of as "approximate".
 
 ```{r apptab, results = "asis"}
-res_df <- readr::read_csv(here("inst", "output", "application", "application.csv"))
-# res_df$bias <- res_df$realized_mean - res_df$estimate
-# res_df_drop <- res_df[ , c("approach", "estimate", "bias", "se", "lb", "ub")]
-res_df_drop <- res_df[, ]
-names(res_df_drop) <- c("Approach", "True Mean", "Est/Pred", "SE", "95% LB", "95% UB")
-xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
-print(xtab_app, include.rownames = FALSE, comment = FALSE)
+# res_df <- readr::read_csv(here("inst", "output", "application", "application.csv"))
+# # res_df$bias <- res_df$realized_mean - res_df$estimate
+# # res_df_drop <- res_df[ , c("approach", "estimate", "bias", "se", "lb", "ub")]
+# res_df_drop <- res_df[, ]
+# names(res_df_drop) <- c("Approach", "True Mean", "Est/Pred", "SE", "95% LB", "95% UB")
+# xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
+# print(xtab_app, include.rownames = FALSE, comment = FALSE)
 ```
@@
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{Relative rMS(P)E in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative RMSE in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
   \label{fig:data_rmspe_eff}
@@
 Aside from the theoretical differences described, a few analytical findings from the simulation study are particularly notable.
-First, independent of the analysis approach, we found no reason to prefer IRS over GRTS when sampling spatial data -- GRTS-Design and GRTS-Model generally had similar rMS(P)E as their IRS counterparts when there was no spatial covariance and lower rMS(P)E than their IRS counterparts when there was spatial covariance.
+First, independent of the analysis approach, we found no reason to prefer IRS over GRTS when sampling spatial data -- GRTS-DB and GRTS-MB generally had similar RMSE as their IRS counterparts when there was no spatial covariance and lower RMSE than their IRS counterparts when there was spatial covariance.
 Second, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
-Though GRTS-Model still had lower rMS(P)E than IRS-Model, the model-based analysis mitigated most of the rMS(P)E inefficiencies that result from the IRS samples lacking spatial balance.
-Third, as the strength of spatial covariance increases, the gap in rMS(P)E and MStdE between IRS-Design and the other sampling-inference combinations also increases, likely because IRS-Design is the only combination that ignores spatial locations in sampling and analysis.
+Though GRTS-MB still had lower RMSE than SRS-MB, the model-based analysis mitigated most of the RMSE inefficiencies that result from the IRS samples lacking spatial balance.
+Third, as the strength of spatial covariance increases, the gap in RMSE and MStdE between SRS-DB and the other sampling-inference combinations also increases, likely because SRS-DB is the only combination that ignores spatial locations in sampling and analysis.
 Fourth and finally, when the response was normal, interval coverage for all sampling-inference combinations was usually close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling-inference combinations was usually between 90% and 95% and closest to 95% when $n = 200$.

### 2022-04-13 inst/manuscript/manuscript.tex
section 3.1 updates

@@
 
-CHANGE LOGNORMAL VERBAGE TO SKEWED -- look for DRE acronym
-
 We evaluated performance of the four sampling-inference combinations in
@@
 For Hg ppb, data were collected at 995 unique lakes (and there were no
-replicates like for ZMMI). The ZMMI and Hg ppb data are shown spatially
-and as a histogram in Figure \ref{fig:zmmi}. The ZMMI data are
-relatively symmetric, and the Hg ppb data are skewed. Also in Figure
+replicates like for ZMMI). The ZMMI and Hg ppb data are shown as spatial
+maps and as histograms in Figure \ref{fig:zmmi}. The ZMMI data tend to
+be highest near the coasts, lowest in the Central United States, higher
+near the coasts, are relatively symmetric, and have a mean of 55.05. The
+Hg ppb data tend to be highest in the Northeastern United States, lowest
+elsewhere, are skewed, and have a mean of 103.16 ppb. Also in Figure
 \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
-The spatial semivariogram quantifies the average squared differences
-(semivariance) of responses whose separation (distance) falls within
-some distance class. The spatial semivariance is closely related to the
-spatial covariance, and spatial semivariograms are often used to gauge
-the strength of spatial dependence in data. Both ZMMI and Hg ppb seem to
-have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as
-the semivariance increases steadily with distance (meaning that
-observations nearby one another tend to be more similar than
-observations far apart from one another).
+The spatial semivariogram quantifies the the halved average squared
+differences (semivariance) of responses whose separation (distance)
+falls within some distance class. The spatial semivariance is closely
+related to the spatial covariance, and spatial semivariograms are often
+used to gauge the strength of spatial dependence in data. Both ZMMI and
+Hg ppb seem to have moderately strong spatial dependence (Figure
+\ref{fig:zmmi}), as the semivariance increases steadily with distance
+(meaning that observations nearby one another tend to be more similar
+than observations far apart from one another).
 
@@
 \(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
-populations and then analyzing the samples using MB and DB inference. We
-used the same evaluation metrics as for the simulated data: mean bias,
-RMSE, and interval coverage. Mean bias is taken as the average deviation
-between each sample's estimated (or predicted) mean and the population
-mean (of ZMMI or Hg ppb):
+populations and then analyzing the samples using MB and DB inference. In
+total, there were six separate scenarios (two responses and three sample
+sizes). We used the same evaluation metrics as for the simulated data:
+mean bias, RMSE, and interval coverage. Mean bias is taken as the
+average deviation between each sample's estimated (or predicted) mean
+and the population mean (of ZMMI or Hg ppb):
 \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
@@
 
-The mean bias was nearly zero for all four sampling-inference
-combinations in all 36 scenarios, so we omit a more detailed summary of
-those results here. Tables for mean bias in all 36 simulation scenarios
-are provided in the supporting information.
-
-Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling
-analysis combinations using the random location layout with
-``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
-\begin{equation*}
-\frac{\text{rMS(P)E of sampling-inference combination}}{\text{rMS(P)E of IRS-Design}},
-\end{equation*} When there is no spatial covariance (Fig.
-\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-inference
-combinations have approximately equal rMS(P)E. In these scenarios, using
-the GRTS algorithm or a model-based analysis does not increase
-efficiency compared to IRS-Design. When there is spatial covariance
-(Fig. \ref{fig:rmspe_eff}, ``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows),
-GRTS-Model tends to have the lowest rMS(P)E, followed by GRTS-Design,
-IRS-Model, and finally IRS-Design, though the difference in relative
-rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is relatively
-small. As the strength of spatial covariance increases, the gap in
-rMS(P)E between IRS-Design and the other sampling-inference combinations
-widens. Finally we note that when there is spatial covariance, IRS-Model
-has a much lower rMS(P)E than IRS-Design, suggesting that the poor
-design properties of IRS are largely mitigated by the model-based
-analysis. These rMS(P)E conclusions are similar to those observed in the
-grid location layout, so we omit a grid location layout figure here.
-Tables for rMS(P)E in all 36 simulation scenarios are provided in the
-supporting information.
+Mean bias was nearly zero for all four sampling-inference combinations
+in all 36 scenarios, so we omit a more detailed summary of those results
+here. Tables for mean bias in all 36 simulation scenarios are provided
+in the supporting information.
+
+We define the relative RMSE as a ratio with numerator given by the RMSE
+for a sampling-inference combination and the denominator given by the
+RMSE for SRS-DB. Relative RMSEs for the random location layout are
+provided in Fig. \ref{fig:rmspe_eff}. When there is no spatial
+covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 0\%''), the four
+sampling-inference combinations have approximately equal RMSE. In these
+scenarios, using GRTS sampling or model-based inference does not
+generally increase efficiency compared to SRS-DB. When there is spatial
+covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 0\%'' and ``DRE\%:
+50\%''), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB,
+SRS-MB, and finally SRS-DB, though the difference in relative RMSE among
+GRTS-MB, GRTS-DB, and SRS-MB is small. As the strength of spatial
+covariance increases, the gap in RMSE between SRS-DB and the other
+sampling-inference combinations widens. Finally we note that when there
+is spatial covariance, SRS-MB has a much lower RMSE than SRS-DB,
+suggesting that the lack of efficiency from SRS is largely mitigated by
+model-based inference. These RMSE conclusions are similar to those
+observed in the grid location layout, so we omit a figure and discussion
+regarding the grid location layout here. Tables for RMSE in all 36
+simulation scenarios are provided in the supporting information.
 
@@
 
-Fig. \ref{fig:mse_eff} shows the relative MStdE of the four
-sampling-inference combinations using the random location layout with
-``IRS-Design'' as the baseline. The relative MStdE is defined as
-\begin{equation*}
-\frac{\text{MStdE of sampling-inference combination}}{\text{MStdE of IRS-Design}},
-\end{equation*} Many general takeaways regarding MStdE are similar to
-general takeaways regarding rMS(P)E: there seems to be no benefit to
-using IRS, even when there is no spatial covariance; as the strength of
-spatial covariance increases, the gap in MStdE between IRS-Design and
-the other sampling-inference combinations widens; and IRS-Model
-outperforms IRS-Design by a noticeable margin. These fact that the
-rMS(P)E and MStdE findings are similar is not particularly surprising
-because the mean bias for all sampling-inference combinations was nearly
-zero, thus rMS(P)E is driven by the standard error of the estimators
-(design-based) or predictors (model-based). We do note that between
-GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was
-no spatial covariance or a medium amount of spatial covariance (Fig.
-\ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows), and
-GRTS-Model had lower MStdE when there was a high amount of spatial
-covariance (Fig. \ref{fig:mse_eff}, ``Prop DE: 0.9'' row). These MStdE
-conclusions are similar to those observed in the grid location layout,
-so we omit a grid location layout figure here. Tables for MStdE in all
-36 simulation scenarios are provided in the supporting information.
-
-Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
-four sampling-inference combinations in the random location layout.
-Within each scenario, the sampling-inference combinations tend to have
-fairly similar interval coverage, though when \(n = 50\) or \(n = 100\),
-GRTS-Design coverage is usually a few percentage points lower than the
-other combinations. Coverage in the normal response scenarios was
-usually near 95\%, while coverage in the lognormal response scenarios
-usually varied from 90\% to 95\% but increased with the sample size. At
-a sample size of 200, all four sampling-inference combinations had
-approximately 95\% interval coverage in both response scenarios for all
-dependent error proportions. These interval coverage conclusions are
-similar to those observed in the grid location layout, so we omit a grid
-location layout figure here. Tables for interval coverage in all 36
-simulation scenarios are provided in the supporting information.
+Fig. \ref{fig:figconf} shows 95\% interval coverage for each of the four
+sampling-inference combinations in the random location layout. Within
+each simulation scenario, all sampling-inference combinations tend to
+have fairly similar interval coverage, though when \(n = 50\) or
+\(n = 100\), GRTS-DB coverage is usually a few percentage points lower
+than the other combinations, which suggests that the local neighborhood
+variance estimate may be slightly too small for small \(n\). Coverage in
+the normal response scenarios was usually near 95\%, while coverage in
+the skewed response scenarios usually varied from 90\% to 95\% but
+increased with the sample size. At a sample size of 200, all four
+sampling-inference combinations had approximately 95\% interval coverage
+in both response scenarios for all dependent error proportions. These
+interval coverage conclusions are similar to those observed in the grid
+location layout, so we omit a figure and discussion regarding the grid
+location layout here. Tables for interval coverage in all 36 simulation
+scenarios are provided in the supporting information.
 
@@
 \hypertarget{sec:r_app}{%
-\subsection{National Lakes Assessment DAta}\label{sec:r_app}}
-
-USE MERCURY UNITS
-
-Fig. \ref{fig:merc} shows a map and histogram of mercury concentration
-in all 986 NLA lakes. The map shows mercury concentration exhibits some
-spatial patterning, with high mercury concentrations in the northeast
-and north central United States. The histogram shows that mercury
-concentration is right-skewed, with most lakes having a low value of
-mercury concentration but a few having a much higher concentration. Fig.
-\ref{fig:merc} also shows mercury concentration's empirical
-semivariogram. The empirical semivariogram can be used as a tool to
-visualize spatial dependence. It quantifies the mean of the halved
-squared differences (semivariance) among all pairs of mercury
-concentrations at different distances apart. When a process has spatial
-covariance (exhibits spatial dependence), the mean semivariance tends to
-be smaller at small distances and larger at large distances. The
-empirical semivariogram in Fig. \ref{fig:merc} suggests that mercury
-concentration exhibits spatial dependence. Lastly we note that the true
-mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.
-
-We selected a single IRS sample and a single GRTS sample and estimated
-(design-based) or predicted (model-based) the mean mercury concentration
-and constructed 95\% confidence (design-based) and 95\% (model-based)
-prediction intervals. For the model-based analyses, the exponential
-covariance was used. Table \ref{tab:appliedtab} shows the results from
-these analyses. Though we should not generalize these results to other
-samples from this population, we do mention a few findings. First,
-IRS-Design has the largest standard error. Second, compared to
-IRS-Design and IRS-Model, GRTS-Design and GRTS-Model are much closer to
-the true mean mercury concentration (have bias closer to zero) and have
-much lower standard errors (more precise intervals). Third, GRTS-Model
-has the least amount of bias and the lowest standard error (most precise
-interval). Finally, we note that for all sampling-inference
-combinations, the true mean mercury concentration (103.2 ng / g) is
-within the bounds of the combination's 95\% interval.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrrr}
-  \hline
-Approach & True Mean & Est/Pred & SE & 95\% LB & 95\% UB \\ 
-  \hline
-IRS-Design & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\ 
-  IRS-Model & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\ 
-  GRTS-Design & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\ 
-  GRTS-Model & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\ 
-   \hline
-\end{tabular}
-\caption{\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.}
-\end{table}
+\subsection{National Lakes Assessment (Real) Data}\label{sec:r_app}}
+
+Mean bias was nearly zero for all four sampling-inference combinations
+in all six scenarios, so we omit a more detailed summary of those
+results here. Tables for mean bias in all six simulations scenarios are
+provided in the supporting information.
+
+Talk about RMSE
+
+Talk about interval coverage -- mention how the model-based assumptions
+of coverage are not really satisfied here, but they can be thought of as
+``approximate''.
 
@@
 analysis approach, we found no reason to prefer IRS over GRTS when
-sampling spatial data -- GRTS-Design and GRTS-Model generally had
-similar rMS(P)E as their IRS counterparts when there was no spatial
-covariance and lower rMS(P)E than their IRS counterparts when there was
-spatial covariance. Second, the sampling decision (IRS vs GRTS) is most
-important when using a design-based analysis. Though GRTS-Model still
-had lower rMS(P)E than IRS-Model, the model-based analysis mitigated
-most of the rMS(P)E inefficiencies that result from the IRS samples
-lacking spatial balance. Third, as the strength of spatial covariance
-increases, the gap in rMS(P)E and MStdE between IRS-Design and the other
-sampling-inference combinations also increases, likely because
-IRS-Design is the only combination that ignores spatial locations in
-sampling and analysis. Fourth and finally, when the response was normal,
-interval coverage for all sampling-inference combinations was usually
-close to 95\% for all sample sizes; when the response was lognormal,
-interval coverage for all sampling-inference combinations was usually
-between 90\% and 95\% and closest to 95\% when \(n = 200\).
+sampling spatial data -- GRTS-DB and GRTS-MB generally had similar RMSE
+as their IRS counterparts when there was no spatial covariance and lower
+RMSE than their IRS counterparts when there was spatial covariance.
+Second, the sampling decision (IRS vs GRTS) is most important when using
+a design-based analysis. Though GRTS-MB still had lower RMSE than
+SRS-MB, the model-based analysis mitigated most of the RMSE
+inefficiencies that result from the IRS samples lacking spatial balance.
+Third, as the strength of spatial covariance increases, the gap in RMSE
+and MStdE between SRS-DB and the other sampling-inference combinations
+also increases, likely because SRS-DB is the only combination that
+ignores spatial locations in sampling and analysis. Fourth and finally,
+when the response was normal, interval coverage for all
+sampling-inference combinations was usually close to 95\% for all sample
+sizes; when the response was lognormal, interval coverage for all
+sampling-inference combinations was usually between 90\% and 95\% and
+closest to 95\% when \(n = 200\).
 

### 2022-04-14 inst/manuscript/manuscript.Rmd
section 3.2 updates

@@
 
-The model-based approach assumes the population is a random realization of a data-generating stochastic process.
+The model-based approach assumes the population is a random realization of a data-generating stochastic process (superpopulation).
 Randomness is formally incorporated through distributional assumptions on this process.
@@
 
-Fig.
-\ref{fig:figconf} shows 95% interval coverage for each of the four sampling-inference combinations in the random location layout.
+95% interval coverage for each of the four sampling-inference combinations in the random location layout is shown in Fig.
+\ref{fig:figconf}.
 Within each simulation scenario, all sampling-inference combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-DB coverage is usually a few percentage points lower than the other combinations, which suggests that the local neighborhood variance estimate may be slightly too small for small $n$.
@@
 
-Talk about RMSE
+The relative RMSE of both ZMMI (symmetric response) and Hg ppb (skewed response) for all four sampling-inference combinations are shown in Fig.$~$\ref{fig:data_rmspe_eff}.
+GRTS-MB has the lowest RMSE, followed by GRTS-DB, SRS-MB, and then SRS-DB.
+The difference in RMSE among GRTS-MB and GRTS-DB tends to be quite small.
+When $n = 50$, SRS-MB RMSE is approximately evenly between IRS-DB RMSE and GRTS-MB RMSE, but for the larger sample sizes ($n = 100$, $n = 200$), SRS-MB RMSE is closer to GRTS-MB RMSE.
+Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB all have noticeably lower RMSE than SRS-DB.
 
-Talk about interval coverage -- mention how the model-based assumptions of coverage are not really satisfied here, but they can be thought of as "approximate".
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
+  \caption{Relative RMSE in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \label{fig:data_rmspe_eff}
+\end{figure}
+
+95% interval coverage of both ZMMI and Hg ppb for all four sampling-inference combinations is shown in Fig.
+\ref{fig:figconf}.
+When $n = 50$, interval coverage for both responses is too low, though interval coverage is higher for ZMMI (symmetric response) than for Hg ppb (skewed response).
+When $n = 100$, ZMMI interval coverage is approximately 95% except for GRTS-DB, which has coverage around 92%, while Hg ppb interval coverage ranges from approximately 90% (GRTS-DB) to 93% (GRTS-MB).
+When $n = 200$, ZMMI interval coverage is approximately 95% while Hg ppb interval coverage ranges from approximately 93% (GRTS-DB) to 95% (GRTS-MB).
+As with the simulated data, coverages for the NLA data tended to increase with the sample sizes, coverages tended to be higher symmteric responses than skewed responses, and regularly the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
+Recall that model-based inference defines interval coverage properties across realized populations.
+With the simulated data, we evaluated interval coverage across realization populations, but for the NLA data, we evaluated interval coverage within a single realization for different samples.
+We did find that model-based coverages were similar to the design-based coverages, however, suggesting that in many cases it is reasonable to heuristically view data from separate samples as being from approximately separate realized populations.
+But generally, if model-based intervals constructed from many random samples of a single realized population show improper coverage, this does not necessarily imply a deficiency in model-based inference.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
+  \caption{Interval coverage in the data study for the four sampling-inference combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black line represents 95\% coverage.}
+  \label{fig:data_figconf}
+\end{figure}
 
@@
 ```
-
-## New Application
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{Relative RMSE in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
-  \label{fig:data_rmspe_eff}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
-  \caption{Interval coverage in the data study for the four sampling-inference combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line represents 95\% coverage.}
-  \label{fig:data_figconf}
-\end{figure}
 

### 2022-04-14 inst/manuscript/manuscript.tex
section 3.2 updates

@@
 The model-based approach assumes the population is a random realization
-of a data-generating stochastic process. Randomness is formally
-incorporated through distributional assumptions on this process.
-Strictly speaking, randomness need not be incorporated through random
-sampling, though Diggle et al. (2010) warn against preferential
+of a data-generating stochastic process (superpopulation). Randomness is
+formally incorporated through distributional assumptions on this
+process. Strictly speaking, randomness need not be incorporated through
+random sampling, though Diggle et al. (2010) warn against preferential
 sampling. Preferential sampling occurs when the process generating the
@@
 
-Fig. \ref{fig:figconf} shows 95\% interval coverage for each of the four
-sampling-inference combinations in the random location layout. Within
-each simulation scenario, all sampling-inference combinations tend to
-have fairly similar interval coverage, though when \(n = 50\) or
-\(n = 100\), GRTS-DB coverage is usually a few percentage points lower
-than the other combinations, which suggests that the local neighborhood
-variance estimate may be slightly too small for small \(n\). Coverage in
-the normal response scenarios was usually near 95\%, while coverage in
-the skewed response scenarios usually varied from 90\% to 95\% but
-increased with the sample size. At a sample size of 200, all four
-sampling-inference combinations had approximately 95\% interval coverage
-in both response scenarios for all dependent error proportions. These
-interval coverage conclusions are similar to those observed in the grid
-location layout, so we omit a figure and discussion regarding the grid
-location layout here. Tables for interval coverage in all 36 simulation
-scenarios are provided in the supporting information.
+95\% interval coverage for each of the four sampling-inference
+combinations in the random location layout is shown in Fig.
+\ref{fig:figconf}. Within each simulation scenario, all
+sampling-inference combinations tend to have fairly similar interval
+coverage, though when \(n = 50\) or \(n = 100\), GRTS-DB coverage is
+usually a few percentage points lower than the other combinations, which
+suggests that the local neighborhood variance estimate may be slightly
+too small for small \(n\). Coverage in the normal response scenarios was
+usually near 95\%, while coverage in the skewed response scenarios
+usually varied from 90\% to 95\% but increased with the sample size. At
+a sample size of 200, all four sampling-inference combinations had
+approximately 95\% interval coverage in both response scenarios for all
+dependent error proportions. These interval coverage conclusions are
+similar to those observed in the grid location layout, so we omit a
+figure and discussion regarding the grid location layout here. Tables
+for interval coverage in all 36 simulation scenarios are provided in the
+supporting information.
 
@@
 
-Talk about RMSE
-
-Talk about interval coverage -- mention how the model-based assumptions
-of coverage are not really satisfied here, but they can be thought of as
-``approximate''.
-
-\hypertarget{new-application}{%
-\subsection{New Application}\label{new-application}}
+The relative RMSE of both ZMMI (symmetric response) and Hg ppb (skewed
+response) for all four sampling-inference combinations are shown in
+Fig.\(~\)\ref{fig:data_rmspe_eff}. GRTS-MB has the lowest RMSE, followed
+by GRTS-DB, SRS-MB, and then SRS-DB. The difference in RMSE among
+GRTS-MB and GRTS-DB tends to be quite small. When \(n = 50\), SRS-MB
+RMSE is approximately evenly between IRS-DB RMSE and GRTS-MB RMSE, but
+for the larger sample sizes (\(n = 100\), \(n = 200\)), SRS-MB RMSE is
+closer to GRTS-MB RMSE. Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB
+all have noticeably lower RMSE than SRS-DB.
 
@@
 \end{figure}
+
+95\% interval coverage of both ZMMI and Hg ppb for all four
+sampling-inference combinations is shown in Fig. \ref{fig:figconf}. When
+\(n = 50\), interval coverage for both responses is too low, though
+interval coverage is higher for ZMMI (symmetric response) than for Hg
+ppb (skewed response). When \(n = 100\), ZMMI interval coverage is
+approximately 95\% except for GRTS-DB, which has coverage around 92\%,
+while Hg ppb interval coverage ranges from approximately 90\% (GRTS-DB)
+to 93\% (GRTS-MB). When \(n = 200\), ZMMI interval coverage is
+approximately 95\% while Hg ppb interval coverage ranges from
+approximately 93\% (GRTS-DB) to 95\% (GRTS-MB). As with the simulated
+data, coverages for the NLA data tended to increase with the sample
+sizes, coverages tended to be higher symmteric responses than skewed
+responses, and regularly the local neighborhood variance was slightly
+too small for small \(n\), yielding slightly lower interval coverages
+than the other sampling-inference combinations. Recall that model-based
+inference defines interval coverage properties across realized
+populations. With the simulated data, we evaluated interval coverage
+across realization populations, but for the NLA data, we evaluated
+interval coverage within a single realization for different samples. We
+did find that model-based coverages were similar to the design-based
+coverages, however, suggesting that in many cases it is reasonable to
+heuristically view data from separate samples as being from
+approximately separate realized populations. But generally, if
+model-based intervals constructed from many random samples of a single
+realized population show improper coverage, this does not necessarily
+imply a deficiency in model-based inference.
 

### 2022-04-14 inst/manuscript/manuscript.Rmd
section 4 updates

@@
 The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black lines separate the sample sizes}
+The solid, black lines separate the sample sizes.}
   \label{fig:rmspe_eff}
@@
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{Relative RMSE in the data study for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
+  \caption{Relative RMSE in the data study for the four sampling-inference combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid, black lines separate the sample sizes.}
   \label{fig:data_rmspe_eff}
@@
   \caption{Interval coverage in the data study for the four sampling-inference combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black line represents 95\% coverage.}
+The rows indicate the proportion of dependent error and the columns indicate the response type.The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
   \label{fig:data_figconf}
@@
 The design-based and model-based approaches to statistical inference are fundamentally different paradigms.
-The design-based approach relies on random sampling to estimate population parameters.
-The model-based approach relies on distributional assumptions to predict realized values of a stochastic process.
-Though the model-based approach does not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
-While the design-based and model-based approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them in a spatial context while implementing spatially balanced sampling and the design-based, local neighborhood variance estimator.
-Aside from the theoretical differences described, a few analytical findings from the simulation study are particularly notable.
-First, independent of the analysis approach, we found no reason to prefer IRS over GRTS when sampling spatial data -- GRTS-DB and GRTS-MB generally had similar RMSE as their IRS counterparts when there was no spatial covariance and lower RMSE than their IRS counterparts when there was spatial covariance.
-Second, the sampling decision (IRS vs GRTS) is most important when using a design-based analysis.
-Though GRTS-MB still had lower RMSE than SRS-MB, the model-based analysis mitigated most of the RMSE inefficiencies that result from the IRS samples lacking spatial balance.
-Third, as the strength of spatial covariance increases, the gap in RMSE and MStdE between SRS-DB and the other sampling-inference combinations also increases, likely because SRS-DB is the only combination that ignores spatial locations in sampling and analysis.
-Fourth and finally, when the response was normal, interval coverage for all sampling-inference combinations was usually close to 95% for all sample sizes; when the response was lognormal, interval coverage for all sampling-inference combinations was usually between 90% and 95% and closest to 95% when $n = 200$.
-
-AT LEAST HAVE DISCUSSION ABOUT MODEL BASED ASSUMPTIONS AND MOVE VALIDITY COMMENTS TO RESULTS SECTION.
+Design-based inference relies on random sampling to estimate population parameters.
+Model-based inference relies on distributional assumptions to predict realized values of a data-generating stochastic process.
+Though model-based inference does not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
+While design-based inference and model-based inference have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them for finite population spatial data while implementing GRTS sampling and the local neighborhood variance estimator.
+Aside from the theoretical differences described throughout the manuscript, a few analytical findings from the simulated and real data studies were particularly notable.
+All sampling-inference combinations had approximately zero mean bias.
+Independent of the inference approach, GRTS-DB and GRTS-MB had lower RMSE than their SRS counterparts.
+Though GRTS-DB and GRTS-MB generally had very similar RMSE, SRS-MB tended to have much lower RMSE than SRS-DB, suggesting that the model-based inference mitigated much of the inefficiency in RMSE from SRS.
+As the proportion of dependent random error in the simulated data increased, SRS-MB, GRTS-DB, and GRTS-MB become increasingly more efficient (lower RMSE) than SRS-DB.
+Interval coverage tended to be higher for the symmetric responses than skewed responses and tended to increase with the sample size.
+At a sample size of $n = 200$, generally all interval coverages were near the desired value of 95%.
 
@@
 Some we have discussed, but others we have not, and they are worthy of consideration in future research.
-Design-based approaches are often computationally efficient, while model-based approaches can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
-The design-based approach easily handles binary data through a straightforward application of the Horvitz-Thompson estimator.
-In contrast, analyzing binary data using a model-based approach generally requires a logistic mixed regression model, which can be challenging to estimate and interpret [@bolker2009generalized].
-The design-based approach yields valid results because the sampling plan and inclusion probabilities are specified directly by the researcher, while the model-based approach may not yield valid results if the assumptions made do not not accurately capture reality.
-The model-based approach, however, can more naturally quantify the relationship between covariates (predictor variables) and the response variable.
-The model-based approach also yields estimated spatial covariance parameters, which help better understand the dependence structure in the process in study.
-Model selection is also possible using model-based approaches and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
-Model-based approaches are capable of more efficient small-area estimation than design-based approaches by leveraging distributional assumptions in areas with few observed units.
-Model-based approaches can also compute unit-by-unit predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
+First we discuss advantages of design-based inference.
+Design-based inference is often computationally efficient, while model-based inference can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
+Design-based inference easily handles binary data through a straightforward application of the Horvitz-Thompson estimator.
+In contrast, analyzing binary data using model-based inference generally requires a logistic mixed regression model, which can be difficult to estimate and interpret [@bolker2009generalized].
+An advantage of design-based inference is that interval coverage is valid (has the proper coverage rate) as long as 1) the sample is sufficiently large to ensure the statistic's sampling distribution is approximately normal and 2) the variance estimator is consistent [@brus1997random; @sarndal2003model].
+This is because with design-based inference, the sampling plan and inclusion probabilities are specified directly by the researcher.
+With the model-based approach, however, interval coverage is unlikely to be valid if the model's assumptions made do not not accurately reflect reality.
+Whether a model's assumptions accurately reflect reality can be a challenging and sometimes impossible question to answer definitively.
+Now we discuss advantages of model-based inference.
+Model-based inference can more naturally quantify the relationship between covariates (predictor variables) and the response variable than design-based inference.
+Model-based inference also yields estimated spatial covariance parameters, which help better understand the dependence structure of the process in study.
+Model selection is also possible using model-based inference and criteria such as cross validation, likelihood ratio tests, or AIC [@akaike1974new].
+Model-based inference is capable of more efficient small-area estimation than design-based inference because model-based inference can leverage distributional assumptions in areas with few observed population units.
+Model-based inference can also compute unit-by-unit predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
 @brus1997random provide a more thorough discussion regarding the benefits and drawbacks of the two approaches.
 In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.
+
+There are many extensions of this research worthy of future consideration, some of which we discuss next.
+Some extensions to study include sampling with unequal inclusion probabilities, different spatially balanced sampling approaches (instead of GRTS), different spatial data configurations, different spatial domains like stream networks [@ver2010moving], different response or covariance structures, the effect of spatial or external mean trends (which can be defined through covariates), and more.
 

### 2022-04-14 inst/manuscript/manuscript.tex
section 4 updates

@@
 The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms. The design-based approach relies on
-random sampling to estimate population parameters. The model-based
-approach relies on distributional assumptions to predict realized values
-of a stochastic process. Though the model-based approach does not rely
-on random sampling, it can still be beneficial as a way to guard against
-preferential sampling. While the design-based and model-based approaches
-have often been compared in the literature from theoretical and
-analytical perspectives, our contribution lies in studying them in a
-spatial context while implementing spatially balanced sampling and the
-design-based, local neighborhood variance estimator. Aside from the
-theoretical differences described, a few analytical findings from the
-simulation study are particularly notable. First, independent of the
-analysis approach, we found no reason to prefer IRS over GRTS when
-sampling spatial data -- GRTS-DB and GRTS-MB generally had similar RMSE
-as their IRS counterparts when there was no spatial covariance and lower
-RMSE than their IRS counterparts when there was spatial covariance.
-Second, the sampling decision (IRS vs GRTS) is most important when using
-a design-based analysis. Though GRTS-MB still had lower RMSE than
-SRS-MB, the model-based analysis mitigated most of the RMSE
-inefficiencies that result from the IRS samples lacking spatial balance.
-Third, as the strength of spatial covariance increases, the gap in RMSE
-and MStdE between SRS-DB and the other sampling-inference combinations
-also increases, likely because SRS-DB is the only combination that
-ignores spatial locations in sampling and analysis. Fourth and finally,
-when the response was normal, interval coverage for all
-sampling-inference combinations was usually close to 95\% for all sample
-sizes; when the response was lognormal, interval coverage for all
-sampling-inference combinations was usually between 90\% and 95\% and
-closest to 95\% when \(n = 200\).
-
-AT LEAST HAVE DISCUSSION ABOUT MODEL BASED ASSUMPTIONS AND MOVE VALIDITY
-COMMENTS TO RESULTS SECTION.
+fundamentally different paradigms. Design-based inference relies on
+random sampling to estimate population parameters. Model-based inference
+relies on distributional assumptions to predict realized values of a
+data-generating stochastic process. Though model-based inference does
+not rely on random sampling, it can still be beneficial as a way to
+guard against preferential sampling. While design-based inference and
+model-based inference have often been compared in the literature from
+theoretical and analytical perspectives, our contribution lies in
+studying them for finite population spatial data while implementing GRTS
+sampling and the local neighborhood variance estimator. Aside from the
+theoretical differences described throughout the manuscript, a few
+analytical findings from the simulated and real data studies were
+particularly notable. All sampling-inference combinations had
+approximately zero mean bias. Independent of the inference approach,
+GRTS-DB and GRTS-MB had lower RMSE than their SRS counterparts. Though
+GRTS-DB and GRTS-MB generally had very similar RMSE, SRS-MB tended to
+have much lower RMSE than SRS-DB, suggesting that the model-based
+inference mitigated much of the inefficiency in RMSE from SRS. As the
+proportion of dependent random error in the simulated data increased,
+SRS-MB, GRTS-DB, and GRTS-MB become increasingly more efficient (lower
+RMSE) than SRS-DB. Interval coverage tended to be higher for the
+symmetric responses than skewed responses and tended to increase with
+the sample size. At a sample size of \(n = 200\), generally all interval
+coverages were near the desired value of 95\%.
 
@@
 discussed, but others we have not, and they are worthy of consideration
-in future research. Design-based approaches are often computationally
-efficient, while model-based approaches can be computationally
-burdensome, especially for likelihood-based estimation methods like REML
-that rely on inverting a covariance matrix. The design-based approach
-easily handles binary data through a straightforward application of the
-Horvitz-Thompson estimator. In contrast, analyzing binary data using a
-model-based approach generally requires a logistic mixed regression
-model, which can be challenging to estimate and interpret (Bolker et
-al., 2009). The design-based approach yields valid results because the
-sampling plan and inclusion probabilities are specified directly by the
-researcher, while the model-based approach may not yield valid results
-if the assumptions made do not not accurately capture reality. The
-model-based approach, however, can more naturally quantify the
-relationship between covariates (predictor variables) and the response
-variable. The model-based approach also yields estimated spatial
-covariance parameters, which help better understand the dependence
-structure in the process in study. Model selection is also possible
-using model-based approaches and criteria such as cross validation,
-likelihood ratio tests, or AIC (Akaike, 1974). Model-based approaches
-are capable of more efficient small-area estimation than design-based
-approaches by leveraging distributional assumptions in areas with few
-observed units. Model-based approaches can also compute unit-by-unit
-predictions at unobserved locations and use them to construct
-informative visualizations like smoothed maps. Brus and De Gruijter
-(1997) provide a more thorough discussion regarding the benefits and
-drawbacks of the two approaches. In short, when deciding whether the
-design-based or model-based approach is more appropriate to implement,
-the benefits and drawbacks of each approach should be considered
-alongside the particular goals of the study.
+in future research. First we discuss advantages of design-based
+inference. Design-based inference is often computationally efficient,
+while model-based inference can be computationally burdensome,
+especially for likelihood-based estimation methods like REML that rely
+on inverting a covariance matrix. Design-based inference easily handles
+binary data through a straightforward application of the
+Horvitz-Thompson estimator. In contrast, analyzing binary data using
+model-based inference generally requires a logistic mixed regression
+model, which can be difficult to estimate and interpret (Bolker et al.,
+2009). An advantage of design-based inference is that interval coverage
+is valid (has the proper coverage rate) as long as 1) the sample is
+sufficiently large to ensure the statistic's sampling distribution is
+approximately normal and 2) the variance estimator is consistent (Brus
+and De Gruijter, 1997; Särndal et al., 2003). This is because with
+design-based inference, the sampling plan and inclusion probabilities
+are specified directly by the researcher. With the model-based approach,
+however, interval coverage is unlikely to be valid if the model's
+assumptions made do not not accurately reflect reality. Whether a
+model's assumptions accurately reflect reality can be a challenging and
+sometimes impossible question to answer definitively. Now we discuss
+advantages of model-based inference. Model-based inference can more
+naturally quantify the relationship between covariates (predictor
+variables) and the response variable than design-based inference.
+Model-based inference also yields estimated spatial covariance
+parameters, which help better understand the dependence structure of the
+process in study. Model selection is also possible using model-based
+inference and criteria such as cross validation, likelihood ratio tests,
+or AIC (Akaike, 1974). Model-based inference is capable of more
+efficient small-area estimation than design-based inference because
+model-based inference can leverage distributional assumptions in areas
+with few observed population units. Model-based inference can also
+compute unit-by-unit predictions at unobserved locations and use them to
+construct informative visualizations like smoothed maps. Brus and De
+Gruijter (1997) provide a more thorough discussion regarding the
+benefits and drawbacks of the two approaches. In short, when deciding
+whether the design-based or model-based approach is more appropriate to
+implement, the benefits and drawbacks of each approach should be
+considered alongside the particular goals of the study.
+
+There are many extensions of this research worthy of future
+consideration, some of which we discuss next. Some extensions to study
+include sampling with unequal inclusion probabilities, different
+spatially balanced sampling approaches (instead of GRTS), different
+spatial data configurations, different spatial domains like stream
+networks (Ver Hoef and Peterson, 2010), different response or covariance
+structures, the effect of spatial or external mean trends (which can be
+defined through covariates), and more.
 
@@
 
+\leavevmode\hypertarget{ref-ver2010moving}{}%
+Ver Hoef, J.M., Peterson, E.E., 2010. A moving average approach for
+spatial statistical models of stream networks. Journal of the American
+Statistical Association 105, 6--18.
+
 \leavevmode\hypertarget{ref-ver2013comparison}{}%

### 2022-04-14 inst/manuscript/manuscript.Rmd
more minor updates

@@
 ---
-title: A comparison of design-based and model-based approaches for finite population spatial data.
+title: A comparison of design-based and model-based approaches for finite population spatial sampling and inference.
 author:
@@
   2.
-We provide relevant background for the design-based and model-based approaches and then study their performance using simulated and real data.
-In the simulated and real data, a variety of sample sizes, location layouts, dependence structures, and response types are considered.
+We provide relevant background for the design-based and model-based approaches and then study their performance using simulated and real data from the United States Environmental Protection Agency's 2012 National Lakes Assessment.
+A variety of sample sizes, location layouts, dependence structures, and response types are considered.
 The population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
@@
 When studying the simulated and real data, we found that regardless of the strength of spatial dependence in the data, the generalized random tessellation stratified (GRTS) algorithm, which explicitly incorporates spatial locations into sampling, tends to outperform the simple random sampling (SRS) algorithm, which does not explicitly incorporate spatial locations into sampling.
-We also found that model-based approaches tend to outperform design-based approaches, even for skewed data where the model-based distributional assumptions are violated.
-The performance gap between these approaches is small GRTS samples are used but large when SRS samples are used.
+We also found that model-based inference tends to outperform design-based inference, even for skewed data where the model-based distributional assumptions are violated.
+The performance gap between these design-based inference and model-based inference is small GRTS samples are used but large when SRS samples are used.
 This suggests that the sampling choice (whether to use GRTS or SRS) is most important when performing design-based inference.
   4.
-There are many benefits and drawbacks to the design-based and model-based approaches for finite population spatial data that practitioners must consider when choosing between them.
+There are many benefits and drawbacks to the design-based and model-based approaches for finite population spatial sampling and inference that practitioners must consider when choosing between them.
 We provide relevant background contextualizing each approach and study their properties in a variety of scenarios, making recommendations for use based on the practitioner's goals.
@@
 In the design-based approach, inference relies on randomly assigning some population units to be in the sample (random sampling).
-Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying stochastic process generating the sample.
-Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation, @brus1997random].
-In this manuscript, we compare the design-based and model-based approaches for finite population spatial data.
+Alternatively, in the model-based approach, inference relies on distributional assumptions about the underlying data-generating stochastic process (superpopulation).
+Each paradigm has a deep historical context [@sterba2009alternative] and its own set of benefits and drawbacks [@hansen1983evaluation; @brus1997random].
+In this manuscript, we compare design-based and model-based approaches for finite population spatial sampling and inference.
 
@@
 
-Spatial data are data that have some sort of spatial index, usually via coordinates.
+Spatial data are data that have some sort of spatial index --usually this index is specified via coordinates.
 @de1990model and @brus1993design give early comparisons of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
@@
 Though the broad comparisons we draw between design-based and model-based approaches generalize to finite and infinite populations, we focus on finite populations.
-A finite population contains a finite number of population units (we assume the finite number is known); an example is lakes (treated as a whole with the lake centroid representing location) in the conterminous United States.
-An infinite population contains an infinite number of population units; an example is locations within a single lake.
+A finite population contains a finite number of population units (we assume the finite number is known) -- an example is lakes (treated as a whole with the lake centroid representing location) in the conterminous United States.
+An infinite population contains an infinite number of population units -- an example is locations within a single lake.
 
 The rest of the manuscript is organized as follows.
-In Section \ref{sec:background}, we introduce and provide relevant background for design-based and model-based approaches to finite population spatial data.
+In Section \ref{sec:background}, we introduce and provide relevant background for design-based and model-based approaches to finite population spatial sampling and inference.
 In Section \ref{sec:mm}, we describe how we intend to compare performance of the approaches using simulated and real data.
-In Section \ref{sec:results}, we present analysis reslts for the simulated and real data.
+In Section \ref{sec:results}, we present analysis reslts for the simulated data and real data from the United States Environmental Protection Agency's 2012 National Lakes Assessment [@USEPA2012NLA].
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 Because spatially balanced samples reflect the layout of the population units, they are not necessarily spread out in space in some equidistant manner.
-One approach to selecting spatially balanced samples is the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
+One method of selecting spatially balanced samples is the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially], which we discuss in more detail in Section \ref{subsec:spb_design}.
+To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (i.e., Dirichlet Tessellations).
 
@@
 
-The model-based approach assumes the population is a random realization of a data-generating stochastic process (superpopulation).
+The model-based approach assumes the population is a random realization of a data-generating stochastic process.
 Randomness is formally incorporated through distributional assumptions on this process.
@@
 Preferential sampling occurs when the process generating the data locations and the process being modeled are not independent of one another.
-To guard against preferential sampling, model-based approaches can implement some form of random sampling, though it is common for model-based approaches to sample non-randomly.
+To guard against preferential sampling, model-based approaches can implement some form of random sampling.
+It is common, however, for model-based approaches to sample non-randomly.
 When model-based approaches do implement random sampling, the inclusion probabilities are ignored when analyzing the sample (in contrast to the design-based approach, which relies on these inclusion probabilities to analyze the sample).
 
-Instead of estimating fixed, unknown population parameters, as in the design-based approach, often the goal of model-based inference is to predict a realized variable.
+Instead of estimating fixed, unknown population parameters, as in the design-based approach, often the goal of model-based inference is to predict the value of a realized variable.
 For example, suppose the realized mean of all population units (the realized population mean) is the variable of interest.
@@
 \ref{fig:fig1}, we provide a visual comparison of the design-based and model-based approaches (@verhoef2002sampling and @brus2021statistical provide similar figures).
-This figure contrasts the design-based approach with a fixed population and random sampling to the model-based approach with random populations and non-random sampling.
+Fig.
+\ref{fig:fig1} contrasts the design-based approach with a fixed population and random sampling to the model-based approach with random populations and non-random sampling.
 
@@
 The first spatially balanced sampling algorithm to see widespread use was the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially].
-To quantify the spatial balance of a sample, @stevens2004spatially proposed loss metrics based on Voronoi polygons (i.e., Dirichlet Tessellations).
 After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata @walvoort2010r, the local pivotal method [@grafstrom2012spatially; @grafstrom2018spatially], spatially correlated Poisson sampling [@grafstrom2012spatiallypoisson], balanced acceptance sampling [@robertson2013bas], within-sample-distance sampling [@benedetti2017spatially], and Halton iterative partitioning sampling [@robertson2018halton].
@@
 This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
-Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$, though @chiles1999geostatistics, pp.
+Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$.
+@chiles1999geostatistics, pp.
 80-93 discuss covariance functions that are not second-order stationary, not isotropic, or not either.
@@
 Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\delta}}_{reml}$, the REML estimates of $\bm{\delta}$.
-Then $\bm{\beta}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\delta}}_{reml}$.
+Then $\bm{\hat{\beta}}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\delta}}_{reml}$.
 
-With the model formulation in Equation$~$\ref{equation:Zmarginal}, the Best Linear Unbiased Predictor (BLUP) for $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
+With the model formulation in Equation$~$\ref{equation:Zmarginal}, the best linear unbiased predictor (BLUP) of $f(\mathbf{b}'\mathbf{z})$ and its prediction variance can be computed.
 While details of the derivation are in @verhoef2008spatial, we note here that the predictor and its variance are both moment-based, meaning that they do not rely on any distributional assumptions.
@@
 
-In this section we describe how we used simulated and real data to investigate performance between simple random sampling without replacement (SRS) and GRTS sampling as well as performance between design-based (DB) and model-based (MB) inference.
-In SRS and GRTS sampling, all population units had equal inclusion probabilities.
+In this section we describe how we used simulated and real data to investigate performance between simple random sampling (SRS) and GRTS sampling as well as performance between design-based (DB) and model-based (MB) inference.
+In SRS and GRTS sampling, all population units had equal inclusion probabilities and were selected without replacement.
 The important distinction between SRS and GRTS is that SRS ignores spatial locations while sampling but GRTS explicitly incorporates them.
@@
 For GRTS-DB, the Horvitz-Thompson esetimator was used to estimate means and the local neighborhood variance was used to estimate variances.
-For SRS-MB and GRTS-MB, FPBK was used to estimate means and variances and parameters were estimated using restricted maximum likelihood.
+For SRS-MB and GRTS-MB, FPBK was used to estimate means and variances using restricted maximum likelihood.
 
@@
 With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
-We used real data from the 2012 National Lakes Assessment [@USEPA2012NLA] to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
+We used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
 With the real data, we were in control of only the random sampling process.
@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-Within each simulation scenario and trial, IRS and GRTS samples were selected and then design-based and model-based analyses were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
-With the model-based analyses, covariance parameters were estimated (using REML) separately for each trial.
+Within each simulation scenario and trial, IRS and GRTS samples were selected and then design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
+With model-based inference, covariance parameters and the mean were estimated (using REML) separately for each trial.
 After all 2000 trials, we summarized the long-run performance of the sampling-inference combination in each scenario by calculating mean bias, root-mean-squared error, and interval coverage.
-Mean bias is taken as the average deviation between each trial's estimated (or predicted) mean and its realized mean: $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes simulation trials.
+Mean bias is taken as the average deviation between each trial's estimated (or predicted) mean ($\hat{\mu}_i$) and its realized mean ($\mu_i$): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes simulation trials.
 Root-mean-squared error is taken as the square root of the average squared deviation between each trial's estimated (or predicted) mean and its realized mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}$.
@@
 These were averaged for the purposes of our study so that each lake had one measurement for ZMMI.
-For Hg ppb, data were collected at 995 unique lakes (and there were no replicates like for ZMMI).
+For Hg ppb, data were collected at 995 unique lakes (there were no replicates).
 The ZMMI and Hg ppb data are shown as spatial maps and as histograms in Figure \ref{fig:zmmi}.
-The ZMMI data tend to be highest near the coasts, lowest in the Central United States, higher near the coasts, are relatively symmetric, and have a mean of 55.05.
+The ZMMI data tend to be highest near the coasts, lowest in the Central United States, are relatively symmetric, and have a mean of 55.05.
 The Hg ppb data tend to be highest in the Northeastern United States, lowest elsewhere, are skewed, and have a mean of 103.16 ppb.
@@
 \end{subfigure}
-\caption{Exploratory graphics of the ZMMI and mercury (Hg ppb) populations in the National Lakes Assessment (NLA) 2012 data.}
+\caption{Exploratory graphics representing populations for the zooplankton multi-metric indices (ZMMI) and mercury concentration in parts per billion (Hg ppb) in the 2012 National Lakes Assessment (NLA) data.}
 \label{fig:zmmi}
@@
 We studied performance of the four sampling-inference combinations by selecting 2000 random IRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
-In total, there were six separate scenarios (two responses and three sample sizes).
+In total, there were six separate scenarios (two responses crossed with three sample sizes).
 We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
-Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean and the population mean (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials and $\mu$ is the population mean.
+Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean ($\hat{\mu}_i$) and the population mean ($\mu$) (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials.
 Root-mean-squared error is taken as the square root of the average squared deviation between each sample's estimated (or predicted) mean and its population mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}$.
@@
 When there is spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "DRE%: 0%" and "DRE%: 50%"), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB, SRS-MB, and finally SRS-DB, though the difference in relative RMSE among GRTS-MB, GRTS-DB, and SRS-MB is small.
+\ref{fig:rmspe_eff}, "DRE%: 50%" and "DRE%: 90%"), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB, SRS-MB, and finally SRS-DB, though the difference in relative RMSE among GRTS-MB, GRTS-DB, and SRS-MB is small.
 As the strength of spatial covariance increases, the gap in RMSE between SRS-DB and the other sampling-inference combinations widens.
@@
 Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB all have noticeably lower RMSE than SRS-DB.
+Tables for RMSE in all six simulations scenarios are provided in the supporting information.
 
@@
 95% interval coverage of both ZMMI and Hg ppb for all four sampling-inference combinations is shown in Fig.
-\ref{fig:figconf}.
+\ref{fig:data_figconf}.
 When $n = 50$, interval coverage for both responses is too low, though interval coverage is higher for ZMMI (symmetric response) than for Hg ppb (skewed response).
@@
 When $n = 200$, ZMMI interval coverage is approximately 95% while Hg ppb interval coverage ranges from approximately 93% (GRTS-DB) to 95% (GRTS-MB).
-As with the simulated data, coverages for the NLA data tended to increase with the sample sizes, coverages tended to be higher symmteric responses than skewed responses, and regularly the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
+As with the simulated data, coverages for the NLA data tended to increase with the sample sizes, coverages tended to be higher for symmteric responses than skewed responses, and regularly the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
 Recall that model-based inference defines interval coverage properties across realized populations.
-With the simulated data, we evaluated interval coverage across realization populations, but for the NLA data, we evaluated interval coverage within a single realization for different samples.
-We did find that model-based coverages were similar to the design-based coverages, however, suggesting that in many cases it is reasonable to heuristically view data from separate samples as being from approximately separate realized populations.
+With the simulated data, we evaluated interval coverage across realization populations, but for the NLA data, we evaluated interval coverage within a single realized population for different samples.
+We did find that model-based coverages were similar to the design-based coverages, however, suggesting that for some realized populations it is reasonable to heuristically view data from separate random samples as being from approximately separate realized populations.
 But generally, if model-based intervals constructed from many random samples of a single realized population show improper coverage, this does not necessarily imply a deficiency in model-based inference.
+Tables for interval coverage in all six simulations scenarios are provided in the supporting information.
 
@@
 
-ADD EXTRAS LIKE ANISOTROPY AND UNEQUAL INCLUSION PROBABILITIES
-
 The design-based and model-based approaches to statistical inference are fundamentally different paradigms.
-Design-based inference relies on random sampling to estimate population parameters.
-Model-based inference relies on distributional assumptions to predict realized values of a data-generating stochastic process.
-Though model-based inference does not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
-While design-based inference and model-based inference have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them for finite population spatial data while implementing GRTS sampling and the local neighborhood variance estimator.
+Design-based approaches rely on random sampling to estimate population parameters.
+Model-based approaches rely on distributional assumptions to predict realized values of a data-generating stochastic process.
+Though model-based approaches do not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
+While design-based and model-based approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them for finite population spatial data while implementing GRTS sampling and the local neighborhood variance estimator.
 Aside from the theoretical differences described throughout the manuscript, a few analytical findings from the simulated and real data studies were particularly notable.
@@
 
-There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial data.
+There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial sampling and inference.
 Some we have discussed, but others we have not, and they are worthy of consideration in future research.
-First we discuss advantages of design-based inference.
+First we discuss advantages of the design-based approach.
 Design-based inference is often computationally efficient, while model-based inference can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
@@
 An advantage of design-based inference is that interval coverage is valid (has the proper coverage rate) as long as 1) the sample is sufficiently large to ensure the statistic's sampling distribution is approximately normal and 2) the variance estimator is consistent [@brus1997random; @sarndal2003model].
-This is because with design-based inference, the sampling plan and inclusion probabilities are specified directly by the researcher.
+This is because with the design-based approach, the sampling plan and inclusion probabilities are specified directly by the researcher.
+An advantage of SRS-DB not previously mentioned is that it is likely to be valid given the consistency of its variance estimator [@sarndal2003model].
 With the model-based approach, however, interval coverage is unlikely to be valid if the model's assumptions made do not not accurately reflect reality.
 Whether a model's assumptions accurately reflect reality can be a challenging and sometimes impossible question to answer definitively.
-Now we discuss advantages of model-based inference.
-Model-based inference can more naturally quantify the relationship between covariates (predictor variables) and the response variable than design-based inference.
+Now we discuss advantages of the model-based approach.
+The model-based approach can more naturally quantify the relationship between covariates (predictor variables) and the response variable than design-based approaches.
 Model-based inference also yields estimated spatial covariance parameters, which help better understand the dependence structure of the process in study.
@@
 Model-based inference is capable of more efficient small-area estimation than design-based inference because model-based inference can leverage distributional assumptions in areas with few observed population units.
-Model-based inference can also compute unit-by-unit predictions at unobserved locations and use them to construct informative visualizations like smoothed maps.
+Model-based approaches also accommodate unit-by-unit predictions at unobserved locations that can be used to construct informative visualizations like smoothed maps.
 @brus1997random provide a more thorough discussion regarding the benefits and drawbacks of the two approaches.
@@
 
-There are many extensions of this research worthy of future consideration, some of which we discuss next.
-Some extensions to study include sampling with unequal inclusion probabilities, different spatially balanced sampling approaches (instead of GRTS), different spatial data configurations, different spatial domains like stream networks [@ver2010moving], different response or covariance structures, the effect of spatial or external mean trends (which can be defined through covariates), and more.
+There are many extensions of this research worthy of future consideration.
+Some (though not all) extensions to study include sampling with unequal inclusion probabilities, using different spatially balanced sampling approaches (instead of GRTS), using different spatial data configurations, using different spatial domains like stream networks [@ver2010moving], using different response or covariance structures, and using spatial or external mean trends (which can be defined through covariates).
 
@@
 
-We would like to thank the editors and anonymous reviewers for their thoughtful comments which greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for hard work and time spent providing us with thoughtful, valuable feedback that which greatly improved the manuscript.
 
@@
 
-In the supporting information, we provide tables of summary statistics for all 36 simulation scenarios.
+In the supporting information, we provide tables of summary statistics for all 36 simulation scenarios and all six real data scenarios.
 

### 2022-04-14 inst/manuscript/manuscript.tex
more minor updates

@@
             pdfauthor={},
-            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data.},
+            pdftitle={A comparison of design-based and model-based approaches for finite population spatial sampling and inference.},
             colorlinks=false,
@@
   \title{A comparison of design-based and model-based approaches for finite
-population spatial data.}
+population spatial sampling and inference.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
     approaches and then study their performance using simulated and real
-    data. In the simulated and real data, a variety of sample sizes,
-    location layouts, dependence structures, and response types are
-    considered. The population mean is the parameter of interest and
-    performance is measured using statistics like bias, squared error, and
-    interval coverage.
+    data from the United States Environmental Protection Agency's 2012
+    National Lakes Assessment. A variety of sample sizes, location
+    layouts, dependence structures, and response types are considered. The
+    population mean is the parameter of interest and performance is
+    measured using statistics like bias, squared error, and interval
+    coverage.
   \item
@@
     incorporate spatial locations into sampling. We also found that
-    model-based approaches tend to outperform design-based approaches,
-    even for skewed data where the model-based distributional assumptions
-    are violated. The performance gap between these approaches is small
-    GRTS samples are used but large when SRS samples are used. This
-    suggests that the sampling choice (whether to use GRTS or SRS) is most
-    important when performing design-based inference.
+    model-based inference tends to outperform design-based inference, even
+    for skewed data where the model-based distributional assumptions are
+    violated. The performance gap between these design-based inference and
+    model-based inference is small GRTS samples are used but large when
+    SRS samples are used. This suggests that the sampling choice (whether
+    to use GRTS or SRS) is most important when performing design-based
+    inference.
   \item
     There are many benefits and drawbacks to the design-based and
-    model-based approaches for finite population spatial data that
-    practitioners must consider when choosing between them. We provide
-    relevant background contextualizing each approach and study their
-    properties in a variety of scenarios, making recommendations for use
-    based on the practitioner's goals.
+    model-based approaches for finite population spatial sampling and
+    inference that practitioners must consider when choosing between them.
+    We provide relevant background contextualizing each approach and study
+    their properties in a variety of scenarios, making recommendations for
+    use based on the practitioner's goals.
   \end{enumerate}
@@
 approach, inference relies on distributional assumptions about the
-underlying stochastic process generating the sample. Each paradigm has a
-deep historical context (Sterba, 2009) and its own set of benefits and
-drawbacks (Hansen et al., 1983, p. @brus1997random). In this manuscript,
-we compare the design-based and model-based approaches for finite
-population spatial data.
-
-Spatial data are data that have some sort of spatial index, usually via
-coordinates. De Gruijter and Ter Braak (1990) and Brus and DeGruijter
-(1993) give early comparisons of design-based and model-based approaches
-for spatial data, quashing the belief that design-based approaches could
-not be used for spatially correlated data. Since then, there have been
-several general comparisons between design-based and model-based
-approaches for spatial data (Brus and De Gruijter, 1997; Brus, 2021; Ver
-Hoef, 2002, 2008). Cooper (2006) reviews the two approaches in an
-ecological context before introducing a ``model-assisted'' variance
-estimator that combines aspects from each approach. In addition to
-Cooper (2006), there has been substantial research and development into
-estimators that use both design-based and model-based principles (see
-e.g., Sterba (2009) and Cicchitelli and Montanari (2012), and for
-Bayesian approaches, see Chan-Golston et al. (2020) and Hofman and Brus
-(2021)).
+underlying data-generating stochastic process (superpopulation). Each
+paradigm has a deep historical context (Sterba, 2009) and its own set of
+benefits and drawbacks (Brus and De Gruijter, 1997; Hansen et al.,
+1983). In this manuscript, we compare design-based and model-based
+approaches for finite population spatial sampling and inference.
+
+Spatial data are data that have some sort of spatial index --usually
+this index is specified via coordinates. De Gruijter and Ter Braak
+(1990) and Brus and DeGruijter (1993) give early comparisons of
+design-based and model-based approaches for spatial data, quashing the
+belief that design-based approaches could not be used for spatially
+correlated data. Since then, there have been several general comparisons
+between design-based and model-based approaches for spatial data (Brus
+and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006)
+reviews the two approaches in an ecological context before introducing a
+``model-assisted'' variance estimator that combines aspects from each
+approach. In addition to Cooper (2006), there has been substantial
+research and development into estimators that use both design-based and
+model-based principles (see e.g., Sterba (2009) and Cicchitelli and
+Montanari (2012), and for Bayesian approaches, see Chan-Golston et al.
+(2020) and Hofman and Brus (2021)).
 
@@
 population contains a finite number of population units (we assume the
-finite number is known); an example is lakes (treated as a whole with
+finite number is known) -- an example is lakes (treated as a whole with
 the lake centroid representing location) in the conterminous United
 States. An infinite population contains an infinite number of population
-units; an example is locations within a single lake.
+units -- an example is locations within a single lake.
 
@@
 design-based and model-based approaches to finite population spatial
-data. In Section \ref{sec:mm}, we describe how we intend to compare
-performance of the approaches using simulated and real data. In Section
-\ref{sec:results}, we present analysis reslts for the simulated and real
-data. And in Section \ref{sec:discussion}, we end with a discussion and
-provide directions for future research.
+sampling and inference. In Section \ref{sec:mm}, we describe how we
+intend to compare performance of the approaches using simulated and real
+data. In Section \ref{sec:results}, we present analysis reslts for the
+simulated data and real data from the United States Environmental
+Protection Agency's 2012 National Lakes Assessment (USEPA, 2012). And in
+Section \ref{sec:discussion}, we end with a discussion and provide
+directions for future research.
 
@@
 units, they are not necessarily spread out in space in some equidistant
-manner. One approach to selecting spatially balanced samples is the
+manner. One method of selecting spatially balanced samples is the
 generalized random tessellation stratified (GRTS) algorithm (Stevens and
 Olsen, 2004), which we discuss in more detail in Section
-\ref{subsec:spb_design}.
+\ref{subsec:spb_design}. To quantify the spatial balance of a sample,
+Stevens and Olsen (2004) proposed loss metrics based on Voronoi polygons
+(i.e., Dirichlet Tessellations).
 
@@
 The model-based approach assumes the population is a random realization
-of a data-generating stochastic process (superpopulation). Randomness is
-formally incorporated through distributional assumptions on this
-process. Strictly speaking, randomness need not be incorporated through
-random sampling, though Diggle et al. (2010) warn against preferential
+of a data-generating stochastic process. Randomness is formally
+incorporated through distributional assumptions on this process.
+Strictly speaking, randomness need not be incorporated through random
+sampling, though Diggle et al. (2010) warn against preferential
 sampling. Preferential sampling occurs when the process generating the
@@
 another. To guard against preferential sampling, model-based approaches
-can implement some form of random sampling, though it is common for
+can implement some form of random sampling. It is common, however, for
 model-based approaches to sample non-randomly. When model-based
@@
 design-based approach, often the goal of model-based inference is to
-predict a realized variable. For example, suppose the realized mean of
-all population units (the realized population mean) is the variable of
-interest. Instead of a fixed, unknown mean, we are predicting the value
-of the mean, a random variable. Prediction intervals are then derived
-using assumptions of the data-generating stochastic process. If we
-repeatedly generate realizations from the same process and select
-samples, then 95\% of all 95\% prediction intervals constructed from a
-procedure with appropriate coverage will contain their respective
-realized means. Cressie (1993) and Schabenberger and Gotway (2017)
-provide thorough reviews of model-based approaches for spatial data. In
-Fig. \ref{fig:fig1}, we provide a visual comparison of the design-based
-and model-based approaches (Ver Hoef (2002) and Brus (2021) provide
-similar figures). This figure contrasts the design-based approach with a
-fixed population and random sampling to the model-based approach with
-random populations and non-random sampling.
+predict the value of a realized variable. For example, suppose the
+realized mean of all population units (the realized population mean) is
+the variable of interest. Instead of a fixed, unknown mean, we are
+predicting the value of the mean, a random variable. Prediction
+intervals are then derived using assumptions of the data-generating
+stochastic process. If we repeatedly generate realizations from the same
+process and select samples, then 95\% of all 95\% prediction intervals
+constructed from a procedure with appropriate coverage will contain
+their respective realized means. Cressie (1993) and Schabenberger and
+Gotway (2017) provide thorough reviews of model-based approaches for
+spatial data. In Fig. \ref{fig:fig1}, we provide a visual comparison of
+the design-based and model-based approaches (Ver Hoef (2002) and Brus
+(2021) provide similar figures). Fig. \ref{fig:fig1} contrasts the
+design-based approach with a fixed population and random sampling to the
+model-based approach with random populations and non-random sampling.
 
@@
 the generalized random tessellation stratified (GRTS) algorithm (Stevens
-and Olsen, 2004). To quantify the spatial balance of a sample, Stevens
-and Olsen (2004) proposed loss metrics based on Voronoi polygons (i.e.,
-Dirichlet Tessellations). After the GRTS algorithm was developed,
-several other spatially balanced sampling algorithms emerged, including
-stratified sampling with compact geographical strata Walvoort et al.
-(2010), the local pivotal method (Grafström et al., 2012; Grafström and
-Matei, 2018), spatially correlated Poisson sampling (Grafström, 2012),
-balanced acceptance sampling (Robertson et al., 2013),
-within-sample-distance sampling (Benedetti and Piersimoni, 2017), and
-Halton iterative partitioning sampling (Robertson et al., 2018). In this
-manuscript, we select spatially balanced samples using the GRTS
-algorithm because it is readily available in the \texttt{spsurvey}
-\textbf{\textsf{R}} package (Dumelle et al., 2022) and naturally
-accommodates finite and infinite sampling frames, unequal inclusion
-probabilities, and replacement units. Replacement units are additional
-population units that can be sampled when a population unit originally
-selected can no longer be sampled. A couple reasons why an originally
-selected site can no longer be sampled include its location being
-physically inaccessible or on private land that the researcher does not
-have permission to access.
+and Olsen, 2004). After the GRTS algorithm was developed, several other
+spatially balanced sampling algorithms emerged, including stratified
+sampling with compact geographical strata Walvoort et al. (2010), the
+local pivotal method (Grafström et al., 2012; Grafström and Matei,
+2018), spatially correlated Poisson sampling (Grafström, 2012), balanced
+acceptance sampling (Robertson et al., 2013), within-sample-distance
+sampling (Benedetti and Piersimoni, 2017), and Halton iterative
+partitioning sampling (Robertson et al., 2018). In this manuscript, we
+select spatially balanced samples using the GRTS algorithm because it is
+readily available in the \texttt{spsurvey} \textbf{\textsf{R}} package
+(Dumelle et al., 2022) and naturally accommodates finite and infinite
+sampling frames, unequal inclusion probabilities, and replacement units.
+Replacement units are additional population units that can be sampled
+when a population unit originally selected can no longer be sampled. A
+couple reasons why an originally selected site can no longer be sampled
+include its location being physically inaccessible or on private land
+that the researcher does not have permission to access.
 
@@
 population units (Cressie, 1993). Henceforth, it is implied that we have
-made these same assumptions regarding \(\bm{\delta}\), though Chiles and
+made these same assumptions regarding \(\bm{\delta}\). Chiles and
 Delfiner (1999), pp.~80-93 discuss covariance functions that are not
@@
 \(\bm{\hat{\delta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
-Then \(\bm{\beta}_{reml}\), the REML estimate of \(\bm{\beta}\), is
-given by
+Then \(\bm{\hat{\beta}}_{reml}\), the REML estimate of \(\bm{\beta}\),
+is given by
 \((\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s\),
@@
 With the model formulation in Equation\(~\)\ref{equation:Zmarginal}, the
-Best Linear Unbiased Predictor (BLUP) for \(f(\mathbf{b}'\mathbf{z})\)
+best linear unbiased predictor (BLUP) of \(f(\mathbf{b}'\mathbf{z})\)
 and its prediction variance can be computed. While details of the
@@
 In this section we describe how we used simulated and real data to
-investigate performance between simple random sampling without
-replacement (SRS) and GRTS sampling as well as performance between
-design-based (DB) and model-based (MB) inference. In SRS and GRTS
-sampling, all population units had equal inclusion probabilities. The
-important distinction between SRS and GRTS is that SRS ignores spatial
-locations while sampling but GRTS explicitly incorporates them.
-Together, the two sampling plans (SRS and GRTS) combined with the two
-inference approaches (DB and MB) yielded four sampling-inference
-combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB. For SRS-DB, the
-Horvitz-Thompson estimator \eqref{eq:ht} was used to estimate means and
-the commonly-used SRS variance formula (Lohr, 2009; Särndal et al.,
-2003) was used to estimate the variance. This variance formula is given
-by \begin{equation}\label{equation:srs_var}
+investigate performance between simple random sampling (SRS) and GRTS
+sampling as well as performance between design-based (DB) and
+model-based (MB) inference. In SRS and GRTS sampling, all population
+units had equal inclusion probabilities and were selected without
+replacement. The important distinction between SRS and GRTS is that SRS
+ignores spatial locations while sampling but GRTS explicitly
+incorporates them. Together, the two sampling plans (SRS and GRTS)
+combined with the two inference approaches (DB and MB) yielded four
+sampling-inference combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB.
+For SRS-DB, the Horvitz-Thompson estimator \eqref{eq:ht} was used to
+estimate means and the commonly-used SRS variance formula (Lohr, 2009;
+Särndal et al., 2003) was used to estimate the variance. This variance
+formula is given by \begin{equation}\label{equation:srs_var}
  \frac{f[\sum_{i = 1}^n (z_i - \bar{z})^2]}{n(n - 1)},
@@
 variance was used to estimate variances. For SRS-MB and GRTS-MB, FPBK
-was used to estimate means and variances and parameters were estimated
-using restricted maximum likelihood.
+was used to estimate means and variances using restricted maximum
+likelihood.
 
@@
 data-generating stochastic process and the random sampling process. We
-used real data from the 2012 National Lakes Assessment (USEPA, 2012) to
-compare the sampling-inference combinations within a single realized
-population (which is typically the case in reality). With the real data,
-we were in control of only the random sampling process.
+used real data from the 2012 National Lakes Assessment to compare the
+sampling-inference combinations within a single realized population
+(which is typically the case in reality). With the real data, we were in
+control of only the random sampling process.
 
@@
 GRTS samples were selected and then design-based and model-based
-analyses were used to estimate (design-based) or predict (model-based)
+inferences were used to estimate (design-based) or predict (model-based)
 the mean and construct 95\% confidence (design-based) or 95\% prediction
-(model-based) intervals. With the model-based analyses, covariance
-parameters were estimated (using REML) separately for each trial. After
-all 2000 trials, we summarized the long-run performance of the
-sampling-inference combination in each scenario by calculating mean
+(model-based) intervals. With model-based inference, covariance
+parameters and the mean were estimated (using REML) separately for each
+trial. After all 2000 trials, we summarized the long-run performance of
+the sampling-inference combination in each scenario by calculating mean
 bias, root-mean-squared error, and interval coverage. Mean bias is taken
 as the average deviation between each trial's estimated (or predicted)
-mean and its realized mean:
+mean (\(\hat{\mu}_i\)) and its realized mean (\(\mu_i\)):
 \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
@@
 purposes of our study so that each lake had one measurement for ZMMI.
-For Hg ppb, data were collected at 995 unique lakes (and there were no
-replicates like for ZMMI). The ZMMI and Hg ppb data are shown as spatial
-maps and as histograms in Figure \ref{fig:zmmi}. The ZMMI data tend to
-be highest near the coasts, lowest in the Central United States, higher
-near the coasts, are relatively symmetric, and have a mean of 55.05. The
-Hg ppb data tend to be highest in the Northeastern United States, lowest
-elsewhere, are skewed, and have a mean of 103.16 ppb. Also in Figure
-\ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
-The spatial semivariogram quantifies the the halved average squared
-differences (semivariance) of responses whose separation (distance)
-falls within some distance class. The spatial semivariance is closely
-related to the spatial covariance, and spatial semivariograms are often
-used to gauge the strength of spatial dependence in data. Both ZMMI and
-Hg ppb seem to have moderately strong spatial dependence (Figure
-\ref{fig:zmmi}), as the semivariance increases steadily with distance
-(meaning that observations nearby one another tend to be more similar
-than observations far apart from one another).
+For Hg ppb, data were collected at 995 unique lakes (there were no
+replicates). The ZMMI and Hg ppb data are shown as spatial maps and as
+histograms in Figure \ref{fig:zmmi}. The ZMMI data tend to be highest
+near the coasts, lowest in the Central United States, are relatively
+symmetric, and have a mean of 55.05. The Hg ppb data tend to be highest
+in the Northeastern United States, lowest elsewhere, are skewed, and
+have a mean of 103.16 ppb. Also in Figure \ref{fig:zmmi} are separate
+spatial semivariograms for ZMMI and Hg ppb. The spatial semivariogram
+quantifies the the halved average squared differences (semivariance) of
+responses whose separation (distance) falls within some distance class.
+The spatial semivariance is closely related to the spatial covariance,
+and spatial semivariograms are often used to gauge the strength of
+spatial dependence in data. Both ZMMI and Hg ppb seem to have moderately
+strong spatial dependence (Figure \ref{fig:zmmi}), as the semivariance
+increases steadily with distance (meaning that observations nearby one
+another tend to be more similar than observations far apart from one
+another).
 
@@
 populations and then analyzing the samples using MB and DB inference. In
-total, there were six separate scenarios (two responses and three sample
-sizes). We used the same evaluation metrics as for the simulated data:
-mean bias, RMSE, and interval coverage. Mean bias is taken as the
-average deviation between each sample's estimated (or predicted) mean
-and the population mean (of ZMMI or Hg ppb):
-\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
-indexes simulation trials and \(\mu\) is the population mean.
-Root-mean-squared error is taken as the square root of the average
-squared deviation between each sample's estimated (or predicted) mean
-and its population mean:
+total, there were six separate scenarios (two responses crossed with
+three sample sizes). We used the same evaluation metrics as for the
+simulated data: mean bias, RMSE, and interval coverage. Mean bias is
+taken as the average deviation between each sample's estimated (or
+predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
+ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
+where \(i\) indexes simulation trials. Root-mean-squared error is taken
+as the square root of the average squared deviation between each
+sample's estimated (or predicted) mean and its population mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
@@
 generally increase efficiency compared to SRS-DB. When there is spatial
-covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 0\%'' and ``DRE\%:
-50\%''), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB,
+covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 50\%'' and ``DRE\%:
+90\%''), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB,
 SRS-MB, and finally SRS-DB, though the difference in relative RMSE among
@@
 closer to GRTS-MB RMSE. Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB
-all have noticeably lower RMSE than SRS-DB.
+all have noticeably lower RMSE than SRS-DB. Tables for RMSE in all six
+simulations scenarios are provided in the supporting information.
 
@@
 95\% interval coverage of both ZMMI and Hg ppb for all four
-sampling-inference combinations is shown in Fig. \ref{fig:figconf}. When
-\(n = 50\), interval coverage for both responses is too low, though
+sampling-inference combinations is shown in Fig. \ref{fig:data_figconf}.
+When \(n = 50\), interval coverage for both responses is too low, though
 interval coverage is higher for ZMMI (symmetric response) than for Hg
@@
 data, coverages for the NLA data tended to increase with the sample
-sizes, coverages tended to be higher symmteric responses than skewed
+sizes, coverages tended to be higher for symmteric responses than skewed
 responses, and regularly the local neighborhood variance was slightly
@@
 across realization populations, but for the NLA data, we evaluated
-interval coverage within a single realization for different samples. We
-did find that model-based coverages were similar to the design-based
-coverages, however, suggesting that in many cases it is reasonable to
-heuristically view data from separate samples as being from
-approximately separate realized populations. But generally, if
-model-based intervals constructed from many random samples of a single
-realized population show improper coverage, this does not necessarily
-imply a deficiency in model-based inference.
+interval coverage within a single realized population for different
+samples. We did find that model-based coverages were similar to the
+design-based coverages, however, suggesting that for some realized
+populations it is reasonable to heuristically view data from separate
+random samples as being from approximately separate realized
+populations. But generally, if model-based intervals constructed from
+many random samples of a single realized population show improper
+coverage, this does not necessarily imply a deficiency in model-based
+inference. Tables for interval coverage in all six simulations scenarios
+are provided in the supporting information.
 
@@
 
-ADD EXTRAS LIKE ANISOTROPY AND UNEQUAL INCLUSION PROBABILITIES
-
 The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms. Design-based inference relies on
-random sampling to estimate population parameters. Model-based inference
-relies on distributional assumptions to predict realized values of a
-data-generating stochastic process. Though model-based inference does
-not rely on random sampling, it can still be beneficial as a way to
-guard against preferential sampling. While design-based inference and
-model-based inference have often been compared in the literature from
-theoretical and analytical perspectives, our contribution lies in
-studying them for finite population spatial data while implementing GRTS
-sampling and the local neighborhood variance estimator. Aside from the
-theoretical differences described throughout the manuscript, a few
-analytical findings from the simulated and real data studies were
-particularly notable. All sampling-inference combinations had
-approximately zero mean bias. Independent of the inference approach,
-GRTS-DB and GRTS-MB had lower RMSE than their SRS counterparts. Though
-GRTS-DB and GRTS-MB generally had very similar RMSE, SRS-MB tended to
-have much lower RMSE than SRS-DB, suggesting that the model-based
-inference mitigated much of the inefficiency in RMSE from SRS. As the
-proportion of dependent random error in the simulated data increased,
-SRS-MB, GRTS-DB, and GRTS-MB become increasingly more efficient (lower
-RMSE) than SRS-DB. Interval coverage tended to be higher for the
-symmetric responses than skewed responses and tended to increase with
-the sample size. At a sample size of \(n = 200\), generally all interval
-coverages were near the desired value of 95\%.
+fundamentally different paradigms. Design-based approaches rely on
+random sampling to estimate population parameters. Model-based
+approaches rely on distributional assumptions to predict realized values
+of a data-generating stochastic process. Though model-based approaches
+do not rely on random sampling, it can still be beneficial as a way to
+guard against preferential sampling. While design-based and model-based
+approaches have often been compared in the literature from theoretical
+and analytical perspectives, our contribution lies in studying them for
+finite population spatial data while implementing GRTS sampling and the
+local neighborhood variance estimator. Aside from the theoretical
+differences described throughout the manuscript, a few analytical
+findings from the simulated and real data studies were particularly
+notable. All sampling-inference combinations had approximately zero mean
+bias. Independent of the inference approach, GRTS-DB and GRTS-MB had
+lower RMSE than their SRS counterparts. Though GRTS-DB and GRTS-MB
+generally had very similar RMSE, SRS-MB tended to have much lower RMSE
+than SRS-DB, suggesting that the model-based inference mitigated much of
+the inefficiency in RMSE from SRS. As the proportion of dependent random
+error in the simulated data increased, SRS-MB, GRTS-DB, and GRTS-MB
+become increasingly more efficient (lower RMSE) than SRS-DB. Interval
+coverage tended to be higher for the symmetric responses than skewed
+responses and tended to increase with the sample size. At a sample size
+of \(n = 200\), generally all interval coverages were near the desired
+value of 95\%.
 
 There are several benefits and drawbacks of the design-based and
-model-based approaches for finite population spatial data. Some we have
-discussed, but others we have not, and they are worthy of consideration
-in future research. First we discuss advantages of design-based
-inference. Design-based inference is often computationally efficient,
-while model-based inference can be computationally burdensome,
-especially for likelihood-based estimation methods like REML that rely
-on inverting a covariance matrix. Design-based inference easily handles
-binary data through a straightforward application of the
-Horvitz-Thompson estimator. In contrast, analyzing binary data using
-model-based inference generally requires a logistic mixed regression
-model, which can be difficult to estimate and interpret (Bolker et al.,
-2009). An advantage of design-based inference is that interval coverage
-is valid (has the proper coverage rate) as long as 1) the sample is
-sufficiently large to ensure the statistic's sampling distribution is
-approximately normal and 2) the variance estimator is consistent (Brus
-and De Gruijter, 1997; Särndal et al., 2003). This is because with
-design-based inference, the sampling plan and inclusion probabilities
-are specified directly by the researcher. With the model-based approach,
-however, interval coverage is unlikely to be valid if the model's
-assumptions made do not not accurately reflect reality. Whether a
-model's assumptions accurately reflect reality can be a challenging and
-sometimes impossible question to answer definitively. Now we discuss
-advantages of model-based inference. Model-based inference can more
-naturally quantify the relationship between covariates (predictor
-variables) and the response variable than design-based inference.
-Model-based inference also yields estimated spatial covariance
-parameters, which help better understand the dependence structure of the
-process in study. Model selection is also possible using model-based
-inference and criteria such as cross validation, likelihood ratio tests,
-or AIC (Akaike, 1974). Model-based inference is capable of more
-efficient small-area estimation than design-based inference because
-model-based inference can leverage distributional assumptions in areas
-with few observed population units. Model-based inference can also
-compute unit-by-unit predictions at unobserved locations and use them to
-construct informative visualizations like smoothed maps. Brus and De
-Gruijter (1997) provide a more thorough discussion regarding the
-benefits and drawbacks of the two approaches. In short, when deciding
-whether the design-based or model-based approach is more appropriate to
-implement, the benefits and drawbacks of each approach should be
-considered alongside the particular goals of the study.
+model-based approaches for finite population spatial sampling and
+inference. Some we have discussed, but others we have not, and they are
+worthy of consideration in future research. First we discuss advantages
+of the design-based approach. Design-based inference is often
+computationally efficient, while model-based inference can be
+computationally burdensome, especially for likelihood-based estimation
+methods like REML that rely on inverting a covariance matrix.
+Design-based inference easily handles binary data through a
+straightforward application of the Horvitz-Thompson estimator. In
+contrast, analyzing binary data using model-based inference generally
+requires a logistic mixed regression model, which can be difficult to
+estimate and interpret (Bolker et al., 2009). An advantage of
+design-based inference is that interval coverage is valid (has the
+proper coverage rate) as long as 1) the sample is sufficiently large to
+ensure the statistic's sampling distribution is approximately normal and
+2) the variance estimator is consistent (Brus and De Gruijter, 1997;
+Särndal et al., 2003). This is because with the design-based approach,
+the sampling plan and inclusion probabilities are specified directly by
+the researcher. An advantage of SRS-DB not previously mentioned is that
+it is likely to be valid given the consistency of its variance estimator
+(Särndal et al., 2003). With the model-based approach, however, interval
+coverage is unlikely to be valid if the model's assumptions made do not
+not accurately reflect reality. Whether a model's assumptions accurately
+reflect reality can be a challenging and sometimes impossible question
+to answer definitively. Now we discuss advantages of the model-based
+approach. The model-based approach can more naturally quantify the
+relationship between covariates (predictor variables) and the response
+variable than design-based approaches. Model-based inference also yields
+estimated spatial covariance parameters, which help better understand
+the dependence structure of the process in study. Model selection is
+also possible using model-based inference and criteria such as cross
+validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
+inference is capable of more efficient small-area estimation than
+design-based inference because model-based inference can leverage
+distributional assumptions in areas with few observed population units.
+Model-based approaches also accommodate unit-by-unit predictions at
+unobserved locations that can be used to construct informative
+visualizations like smoothed maps. Brus and De Gruijter (1997) provide a
+more thorough discussion regarding the benefits and drawbacks of the two
+approaches. In short, when deciding whether the design-based or
+model-based approach is more appropriate to implement, the benefits and
+drawbacks of each approach should be considered alongside the particular
+goals of the study.
 
 There are many extensions of this research worthy of future
-consideration, some of which we discuss next. Some extensions to study
-include sampling with unequal inclusion probabilities, different
-spatially balanced sampling approaches (instead of GRTS), different
-spatial data configurations, different spatial domains like stream
-networks (Ver Hoef and Peterson, 2010), different response or covariance
-structures, the effect of spatial or external mean trends (which can be
-defined through covariates), and more.
+consideration. Some (though not all) extensions to study include
+sampling with unequal inclusion probabilities, using different spatially
+balanced sampling approaches (instead of GRTS), using different spatial
+data configurations, using different spatial domains like stream
+networks (Ver Hoef and Peterson, 2010), using different response or
+covariance structures, and using spatial or external mean trends (which
+can be defined through covariates).
 
@@
 
-We would like to thank the editors and anonymous reviewers for their
-thoughtful comments which greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for hard work
+and time spent providing us with thoughtful, valuable feedback that
+which greatly improved the manuscript.
 
@@
 In the supporting information, we provide tables of summary statistics
-for all 36 simulation scenarios.
+for all 36 simulation scenarios and all six real data scenarios.
 

### 2022-04-18 inst/manuscript/manuscript.Rmd
minor review updates

@@
 We also found that model-based inference tends to outperform design-based inference, even for skewed data where the model-based distributional assumptions are violated.
-The performance gap between these design-based inference and model-based inference is small GRTS samples are used but large when SRS samples are used.
-This suggests that the sampling choice (whether to use GRTS or SRS) is most important when performing design-based inference.
+The performance gap between design-based inference and model-based inference is small when GRTS samples are used but large when SRS samples are used, suggesting that the sampling choice (whether to use GRTS or SRS) is most important when performing design-based inference.
   4.
@@
 
-Spatial data are data that have some sort of spatial index --usually this index is specified via coordinates.
+Spatial data are data that have some sort of spatial index (usually specified via coordinates).
 @de1990model and @brus1993design give early comparisons of design-based and model-based approaches for spatial data, quashing the belief that design-based approaches could not be used for spatially correlated data.
@@
 Some examples of commonly used sampling designs include simple random sampling, stratified random sampling, and cluster sampling.
-The inclusion probability of a population unit follows by summing each sample's probability of selection over all samples that contain the population unit.
-Inclusion probabilities are later used to estimate population parameters.
+The inclusion probability of a population unit is calculated by summing each sample's probability of selection over all samples that contain the population unit.
+Inclusion probabilities are often when selecting samples and estimating population parameters.
 
@@
 The first spatially balanced sampling algorithm to see widespread use was the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially].
-After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata @walvoort2010r, the local pivotal method [@grafstrom2012spatially; @grafstrom2018spatially], spatially correlated Poisson sampling [@grafstrom2012spatiallypoisson], balanced acceptance sampling [@robertson2013bas], within-sample-distance sampling [@benedetti2017spatially], and Halton iterative partitioning sampling [@robertson2018halton].
+After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata [@walvoort2010r], the local pivotal method [@grafstrom2012spatially; @grafstrom2018spatially], spatially correlated Poisson sampling [@grafstrom2012spatiallypoisson], balanced acceptance sampling [@robertson2013bas], within-sample-distance sampling [@benedetti2017spatially], and Halton iterative partitioning sampling [@robertson2018halton].
 In this manuscript, we select spatially balanced samples using the GRTS algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units.
 Replacement units are additional population units that can be sampled when a population unit originally selected can no longer be sampled.
-A couple reasons why an originally selected site can no longer be sampled include its location being physically inaccessible or on private land that the researcher does not have permission to access.
+A couple of reasons why an originally selected site can no longer be sampled include its location being physically inaccessible or it is on private land that the researcher does not have permission to access.
 
 The GRTS algorithm selects samples by utilizing a particular mapping between two-dimensional and one-dimensional space that preserves proximity relationships.
-First the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells.
+First, the bounding box of the domain is split up into four distinct, equally sized squares called level-one cells.
 Each level-one cell is randomly assigned a level-one address of 0, 1, 2, or 3.
 The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$.
-Within each level-one cell, the inclusion probability for each population unit is summed, and if any of these sums exceed one, a second level of cells is added.
+Within each level-one cell, the inclusion probability for each population unit (which is pre-specified) is summed, and if any of these sums exceed one, a second level of cells is added.
 Then each level-one cell is split into four distinct, equally sized squares called level-two cells.
@@
 Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$.
+Figure$~$\ref{fig:grts} provides some intuition regarding the assignment of level-one and level-two cells.
 
-After determining $\mathcal{A}_k$, it is placed into hierarchical order.
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/grts_level1.jpeg}
+  \caption{Assignment of level-one cells to the spatial domain. Grey circles indicate population units.}
+  \label{fig:grts_level1}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/grts_level2.jpeg}
+  \caption{Assignment of level-two cells to the spatial domain. Grey circles indicate population units.}
+  \label{fig:grts_level2}
+\end{subfigure}
+\caption{Assignment of level-one and level-two cells to the spatial domain.
+In (a), each level-one cells is randomly given a level-one address of 0, 1, 2, or 3.
+In (b), each level-two cell within each level-one cell is randomly given a level-two address of 0, 1, 2, or 3.}
+\label{fig:grts}
+\end{figure}
+
+After determining $\mathcal{A}_k$, the set is placed into hierarchical order.
 Hierarchical order is a numeric order that first sorts $\mathcal{A}_k$ by the level-one addresses from smallest to largest, then sorts $\mathcal{A}_k$ by the level-two addresses from smallest to largest, and so on.
@@
 Because hierarchical ordering sorts by level-one cells, then level-two cells, and so on, population units that have similar hierarchical addresses tend to be nearby one another in space.
-Next each population unit is mapped to a one-dimensional line in hierarchical order where each population unit's inclusion probability equals its line-length.
+Next, each population unit is mapped to a one-dimensional line in hierarchical order where each population unit's inclusion probability equals its line-length.
 If a level-$k$ cell has multiple population units in it, they are randomly placed within the cell's respective line segment.
@@
 After selecting a sample and collecting data, unbiased estimates of population means and totals can be obtained using the Horvitz-Thompson estimator [@horvitz1952generalization].
-If $\tau$ is a population total, the Horvitz-Thompson estimator for $\tau$, denoted by $\hat{\tau}_{ht}$, is is given by
+If $\tau$ is a population total, the Horvitz-Thompson estimator for $\tau$, denoted by $\hat{\tau}_{ht}$, is given by
 \begin{align}\label{eq:ht}
@@
 Instead of developing inference based on a specific sampling design, we assume the data are generated by a spatial stochastic process.
-We summarize some of the basic principles of FPBK next -- for more details, see @verhoef2008spatial.
+We summarize some of the basic principles of FPBK next -- see @verhoef2008spatial for technical details and see @higham2021adjusting for an extension to cases of imperfect detection among population units.
 Let ${\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), .
@@
 
-FPBK assumes $\bm{\delta}$ in Equation$~$\ref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
-This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), isotropic (independent of direction), and decay with distance between population units [@cressie1993statistics].
+FPBK assumes $\bm{\delta}$ in Equation$~$\eqref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
+This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), isotropic (independent of direction), and decays with distance between population units [@cressie1993statistics].
 Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$.
@@
 \end{align}
-where $\sigma^2_{1}$ is the variance parameter that quantifies the spatially dependent variability, $\sigma^2_{2}$ is the variance parameter the quantifies that spatially independent variability, $\phi$ is the distance parameter that measures the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
-In geostatistical literature, $\sigma^2_{1}$ is often called the partial sill, $\sigma^2_{2}$ is often called the nugget, and $\phi$ is often called the range.
+where $\sigma^2_{1}$ is the variance parameter that quantifies the spatially dependent (correlated) variability, $\sigma^2_{2}$ is the variance parameter the quantifies that spatially independent (not correlated) variability, $\phi$ is the distance parameter that measures the distance-decay rate of the covariance, and $h_{i,j}$ is the Euclidean distance between population units $i$ and $j$.
+In geostatistical literature, $\sigma^2_{1}$ is called the partial sill, $\sigma^2_{2}$ is called the nugget, and $\phi$ is called the range.
+We denote $\bm{\theta}$ as the vector of covariance parameters that composes $\bm{\delta}$.
+In Equation$~$\ref{equation:expcov}, $\bm{\theta} = \{\sigma^2_1, \sigma^2_2, \phi \}$.
 
-The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on using restricted maximum likelihood [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on using restricted maximum likelihood (REML) [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
 REML is preferred over maximum likelihood (ML) because ML estimates can be badly biased for small sample sizes, due to the fact that ML makes no adjustment for the simultaneous estimation of $\bm{\beta}$ and $\bm{\delta}$ [@patterson1971recovery].
@@
 where $\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s$ and $\bm{\Sigma}_{ss}$  is the covariance matrix of the sampled sites.
-Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\delta}}_{reml}$, the REML estimates of $\bm{\delta}$.
-Then $\bm{\hat{\beta}}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\delta}}_{reml}$.
+Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\theta}}_{reml}$, the REML estimates of $\bm{\delta}$.
+Then $\bm{\hat{\beta}}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\theta}}_{reml}$.
 
@@
 With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
-We used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
+We then used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
 With the real data, we were in control of only the random sampling process.
@@
 Also in Figure \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
-The spatial semivariogram quantifies the the halved average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
+The spatial semivariogram quantifies the halved average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
 The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
@@
 
-Mean bias was nearly zero for all four sampling-inference combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
+Mean bias is nearly zero for all four sampling-inference combinations in all 36 scenarios, so we omit a more detailed summary of those results here.
 Tables for mean bias in all 36 simulation scenarios are provided in the supporting information.
@@
 When there is spatial covariance (Fig.
-\ref{fig:rmspe_eff}, "DRE%: 50%" and "DRE%: 90%"), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB, SRS-MB, and finally SRS-DB, though the difference in relative RMSE among GRTS-MB, GRTS-DB, and SRS-MB is small.
+\ref{fig:rmspe_eff}, "DRE%: 50%" and "DRE%: 90%"), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB, SRS-MB, and finally SRS-DB.
 As the strength of spatial covariance increases, the gap in RMSE between SRS-DB and the other sampling-inference combinations widens.
@@
 Within each simulation scenario, all sampling-inference combinations tend to have fairly similar interval coverage, though when $n = 50$ or $n = 100$, GRTS-DB coverage is usually a few percentage points lower than the other combinations, which suggests that the local neighborhood variance estimate may be slightly too small for small $n$.
-Coverage in the normal response scenarios was usually near 95%, while coverage in the skewed response scenarios usually varied from 90% to 95% but increased with the sample size.
-At a sample size of 200, all four sampling-inference combinations had approximately 95% interval coverage in both response scenarios for all dependent error proportions.
+Coverage in the normal response scenarios is usually near 95%, while coverage in the skewed response scenarios usually varies from 90% to 95% but increases with the sample size.
+At a sample size of 200, all four sampling-inference combinations have approximately 95% interval coverage in both response scenarios for all dependent random error proportions.
 These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a figure and discussion regarding the grid location layout here.
@@
 
-Mean bias was nearly zero for all four sampling-inference combinations in all six scenarios, so we omit a more detailed summary of those results here.
-Tables for mean bias in all six simulations scenarios are provided in the supporting information.
+Mean bias is nearly zero for all four sampling-inference combinations in all six scenarios, so we omit a more detailed summary of those results here.
+Tables for mean bias in all six simulation scenarios are provided in the supporting information.
 
@@
 Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB all have noticeably lower RMSE than SRS-DB.
-Tables for RMSE in all six simulations scenarios are provided in the supporting information.
+Tables for RMSE in all six simulation scenarios are provided in the supporting information.
 
@@
 When $n = 200$, ZMMI interval coverage is approximately 95% while Hg ppb interval coverage ranges from approximately 93% (GRTS-DB) to 95% (GRTS-MB).
-As with the simulated data, coverages for the NLA data tended to increase with the sample sizes, coverages tended to be higher for symmteric responses than skewed responses, and regularly the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
+As with the simulated data, coverages for the NLA data tend to increase with the sample sizes, coverages tend to be higher for symmetric responses than for skewed responses, and the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
 Recall that model-based inference defines interval coverage properties across realized populations.
@@
 But generally, if model-based intervals constructed from many random samples of a single realized population show improper coverage, this does not necessarily imply a deficiency in model-based inference.
-Tables for interval coverage in all six simulations scenarios are provided in the supporting information.
+Tables for interval coverage in all six simulation scenarios are provided in the supporting information.
 
@@
 Some we have discussed, but others we have not, and they are worthy of consideration in future research.
-First we discuss advantages of the design-based approach.
+First, we discuss advantages of the design-based approach.
 Design-based inference is often computationally efficient, while model-based inference can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
@@
 An advantage of SRS-DB not previously mentioned is that it is likely to be valid given the consistency of its variance estimator [@sarndal2003model].
-With the model-based approach, however, interval coverage is unlikely to be valid if the model's assumptions made do not not accurately reflect reality.
-Whether a model's assumptions accurately reflect reality can be a challenging and sometimes impossible question to answer definitively.
-Now we discuss advantages of the model-based approach.
+With the model-based approach, however, interval coverage is unlikely to be valid if the model assumptions made do not not accurately reflect reality.
+Whether model assumptions accurately reflect reality can be a challenging and sometimes impossible question to answer definitively.
+
+Now, we discuss advantages of the model-based approach.
 The model-based approach can more naturally quantify the relationship between covariates (predictor variables) and the response variable than design-based approaches.
@@
 
-There are many extensions of this research worthy of future consideration.
-Some (though not all) extensions to study include sampling with unequal inclusion probabilities, using different spatially balanced sampling approaches (instead of GRTS), using different spatial data configurations, using different spatial domains like stream networks [@ver2010moving], using different response or covariance structures, and using spatial or external mean trends (which can be defined through covariates).
+There are many extensions of this research worthy of future consideration that include sampling with unequal inclusion probabilities, using different spatially balanced sampling approaches (instead of GRTS), using different spatial data configurations, using different spatial domains like stream networks [@ver2010moving], using different response or covariance structures, and using spatial or external mean trends (which can be defined through covariates).
 
@@
 
-We would like to thank the editors and anonymous reviewers for hard work and time spent providing us with thoughtful, valuable feedback that which greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for hard work and time spent providing us with thoughtful, valuable feedback which greatly improved the manuscript.
 

### 2022-04-18 inst/manuscript/manuscript.tex
minor review updates

@@
     for skewed data where the model-based distributional assumptions are
-    violated. The performance gap between these design-based inference and
-    model-based inference is small GRTS samples are used but large when
-    SRS samples are used. This suggests that the sampling choice (whether
-    to use GRTS or SRS) is most important when performing design-based
-    inference.
+    violated. The performance gap between design-based inference and
+    model-based inference is small when GRTS samples are used but large
+    when SRS samples are used, suggesting that the sampling choice
+    (whether to use GRTS or SRS) is most important when performing
+    design-based inference.
   \item
@@
 
-Spatial data are data that have some sort of spatial index --usually
-this index is specified via coordinates. De Gruijter and Ter Braak
-(1990) and Brus and DeGruijter (1993) give early comparisons of
-design-based and model-based approaches for spatial data, quashing the
-belief that design-based approaches could not be used for spatially
-correlated data. Since then, there have been several general comparisons
-between design-based and model-based approaches for spatial data (Brus
-and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006)
-reviews the two approaches in an ecological context before introducing a
+Spatial data are data that have some sort of spatial index (usually
+specified via coordinates). De Gruijter and Ter Braak (1990) and Brus
+and DeGruijter (1993) give early comparisons of design-based and
+model-based approaches for spatial data, quashing the belief that
+design-based approaches could not be used for spatially correlated data.
+Since then, there have been several general comparisons between
+design-based and model-based approaches for spatial data (Brus and De
+Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006) reviews
+the two approaches in an ecological context before introducing a
 ``model-assisted'' variance estimator that combines aspects from each
@@
 sampling, and cluster sampling. The inclusion probability of a
-population unit follows by summing each sample's probability of
+population unit is calculated by summing each sample's probability of
 selection over all samples that contain the population unit. Inclusion
-probabilities are later used to estimate population parameters.
+probabilities are often when selecting samples and estimating population
+parameters.
 
@@
 spatially balanced sampling algorithms emerged, including stratified
-sampling with compact geographical strata Walvoort et al. (2010), the
+sampling with compact geographical strata (Walvoort et al., 2010), the
 local pivotal method (Grafström et al., 2012; Grafström and Matei,
@@
 when a population unit originally selected can no longer be sampled. A
-couple reasons why an originally selected site can no longer be sampled
-include its location being physically inaccessible or on private land
-that the researcher does not have permission to access.
+couple of reasons why an originally selected site can no longer be
+sampled include its location being physically inaccessible or it is on
+private land that the researcher does not have permission to access.
 
@@
 between two-dimensional and one-dimensional space that preserves
-proximity relationships. First the bounding box of the domain is split
+proximity relationships. First, the bounding box of the domain is split
 up into four distinct, equally sized squares called level-one cells.
@@
 each level-one cell, the inclusion probability for each population unit
-is summed, and if any of these sums exceed one, a second level of cells
-is added. Then each level-one cell is split into four distinct, equally
-sized squares called level-two cells. Each level-two cell is randomly
-assigned a level-two address of 0, 1, 2, or 3. The set of level-two
-cells is denoted by \(\mathcal{A}_2\) and defined as
+(which is pre-specified) is summed, and if any of these sums exceed one,
+a second level of cells is added. Then each level-one cell is split into
+four distinct, equally sized squares called level-two cells. Each
+level-two cell is randomly assigned a level-two address of 0, 1, 2, or
+3. The set of level-two cells is denoted by \(\mathcal{A}_2\) and
+defined as
 \(\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}\).
@@
 \(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).
-
-After determining \(\mathcal{A}_k\), it is placed into hierarchical
+Figure\(~\)\ref{fig:grts} provides some intuition regarding the
+assignment of level-one and level-two cells.
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/grts_level1.jpeg}
+  \caption{Assignment of level-one cells to the spatial domain. Grey circles indicate population units.}
+  \label{fig:grts_level1}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+  \centering
+  \includegraphics[width = 1\linewidth]{figures/grts_level2.jpeg}
+  \caption{Assignment of level-two cells to the spatial domain. Grey circles indicate population units.}
+  \label{fig:grts_level2}
+\end{subfigure}
+\caption{Assignment of level-one and level-two cells to the spatial domain.
+In (a), each level-one cells is randomly given a level-one address of 0, 1, 2, or 3.
+In (b), each level-two cell within each level-one cell is randomly given a level-two address of 0, 1, 2, or 3.}
+\label{fig:grts}
+\end{figure}
+
+After determining \(\mathcal{A}_k\), the set is placed into hierarchical
 order. Hierarchical order is a numeric order that first sorts
@@
 cells, and so on, population units that have similar hierarchical
-addresses tend to be nearby one another in space. Next each population
+addresses tend to be nearby one another in space. Next, each population
 unit is mapped to a one-dimensional line in hierarchical order where
@@
 total, the Horvitz-Thompson estimator for \(\tau\), denoted by
-\(\hat{\tau}_{ht}\), is is given by \begin{align}\label{eq:ht}
+\(\hat{\tau}_{ht}\), is given by \begin{align}\label{eq:ht}
   \hat{\tau}_{ht} = \sum_{i = 1}^n z_i \pi_i^{-1},
@@
 stochastic process. We summarize some of the basic principles of FPBK
-next -- for more details, see Ver Hoef (2008). Let
+next -- see Ver Hoef (2008) for technical details and see Higham et al.
+(2021) for an extension to cases of imperfect detection among population
+units. Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 
-FPBK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
+FPBK assumes \(\bm{\delta}\) in Equation\(~\)\eqref{equation:Zmarginal}
 has mean-zero and a spatial dependence structure that can be modeled
@@
 the separation vector (e.g., distance) between population units),
-isotropic (independent of direction), and decay with distance between
+isotropic (independent of direction), and decays with distance between
 population units (Cressie, 1993). Henceforth, it is implied that we have
@@
 \end{align} where \(\sigma^2_{1}\) is the variance parameter that
-quantifies the spatially dependent variability, \(\sigma^2_{2}\) is the
-variance parameter the quantifies that spatially independent
-variability, \(\phi\) is the distance parameter that measures the
-distance-decay rate of the covariance, and \(h_{i,j}\) is the Euclidean
-distance between population units \(i\) and \(j\). In geostatistical
-literature, \(\sigma^2_{1}\) is often called the partial sill,
-\(\sigma^2_{2}\) is often called the nugget, and \(\phi\) is often
-called the range.
+quantifies the spatially dependent (correlated) variability,
+\(\sigma^2_{2}\) is the variance parameter the quantifies that spatially
+independent (not correlated) variability, \(\phi\) is the distance
+parameter that measures the distance-decay rate of the covariance, and
+\(h_{i,j}\) is the Euclidean distance between population units \(i\) and
+\(j\). In geostatistical literature, \(\sigma^2_{1}\) is called the
+partial sill, \(\sigma^2_{2}\) is called the nugget, and \(\phi\) is
+called the range. We denote \(\bm{\theta}\) as the vector of covariance
+parameters that composes \(\bm{\delta}\). In
+Equation\(~\)\ref{equation:expcov},
+\(\bm{\theta} = \{\sigma^2_1, \sigma^2_2, \phi \}\).
 
@@
 using a variety of techniques, but we focus on using restricted maximum
-likelihood (Harville, 1977; Patterson and Thompson, 1971; Wolfinger et
-al., 1994). REML is preferred over maximum likelihood (ML) because ML
-estimates can be badly biased for small sample sizes, due to the fact
-that ML makes no adjustment for the simultaneous estimation of
+likelihood (REML) (Harville, 1977; Patterson and Thompson, 1971;
+Wolfinger et al., 1994). REML is preferred over maximum likelihood (ML)
+because ML estimates can be badly biased for small sample sizes, due to
+the fact that ML makes no adjustment for the simultaneous estimation of
 \(\bm{\beta}\) and \(\bm{\delta}\) (Patterson and Thompson, 1971). Minus
@@
 Minimizing Equation\(~\)\ref{equation:reml} yields
-\(\bm{\hat{\delta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
+\(\bm{\hat{\theta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
 Then \(\bm{\hat{\beta}}_{reml}\), the REML estimate of \(\bm{\beta}\),
@@
 where \(\bm{\hat{\Sigma}}_{ss}\) is \(\bm{\Sigma}_{ss}\) evaluated at
-\(\bm{\hat{\delta}}_{reml}\).
+\(\bm{\hat{\theta}}_{reml}\).
 
@@
 data-generating stochastic process and the random sampling process. We
-used real data from the 2012 National Lakes Assessment to compare the
-sampling-inference combinations within a single realized population
+then used real data from the 2012 National Lakes Assessment to compare
+the sampling-inference combinations within a single realized population
 (which is typically the case in reality). With the real data, we were in
@@
 spatial semivariograms for ZMMI and Hg ppb. The spatial semivariogram
-quantifies the the halved average squared differences (semivariance) of
+quantifies the halved average squared differences (semivariance) of
 responses whose separation (distance) falls within some distance class.
@@
 
-Mean bias was nearly zero for all four sampling-inference combinations
-in all 36 scenarios, so we omit a more detailed summary of those results
+Mean bias is nearly zero for all four sampling-inference combinations in
+all 36 scenarios, so we omit a more detailed summary of those results
 here. Tables for mean bias in all 36 simulation scenarios are provided
@@
 90\%''), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB,
-SRS-MB, and finally SRS-DB, though the difference in relative RMSE among
-GRTS-MB, GRTS-DB, and SRS-MB is small. As the strength of spatial
-covariance increases, the gap in RMSE between SRS-DB and the other
+SRS-MB, and finally SRS-DB. As the strength of spatial covariance
+increases, the gap in RMSE between SRS-DB and the other
 sampling-inference combinations widens. Finally we note that when there
@@
 suggests that the local neighborhood variance estimate may be slightly
-too small for small \(n\). Coverage in the normal response scenarios was
+too small for small \(n\). Coverage in the normal response scenarios is
 usually near 95\%, while coverage in the skewed response scenarios
-usually varied from 90\% to 95\% but increased with the sample size. At
-a sample size of 200, all four sampling-inference combinations had
+usually varies from 90\% to 95\% but increases with the sample size. At
+a sample size of 200, all four sampling-inference combinations have
 approximately 95\% interval coverage in both response scenarios for all
-dependent error proportions. These interval coverage conclusions are
-similar to those observed in the grid location layout, so we omit a
+dependent random error proportions. These interval coverage conclusions
+are similar to those observed in the grid location layout, so we omit a
 figure and discussion regarding the grid location layout here. Tables
@@
 
-Mean bias was nearly zero for all four sampling-inference combinations
-in all six scenarios, so we omit a more detailed summary of those
-results here. Tables for mean bias in all six simulations scenarios are
-provided in the supporting information.
+Mean bias is nearly zero for all four sampling-inference combinations in
+all six scenarios, so we omit a more detailed summary of those results
+here. Tables for mean bias in all six simulation scenarios are provided
+in the supporting information.
 
@@
 all have noticeably lower RMSE than SRS-DB. Tables for RMSE in all six
-simulations scenarios are provided in the supporting information.
+simulation scenarios are provided in the supporting information.
 
@@
 approximately 93\% (GRTS-DB) to 95\% (GRTS-MB). As with the simulated
-data, coverages for the NLA data tended to increase with the sample
-sizes, coverages tended to be higher for symmteric responses than skewed
-responses, and regularly the local neighborhood variance was slightly
-too small for small \(n\), yielding slightly lower interval coverages
-than the other sampling-inference combinations. Recall that model-based
-inference defines interval coverage properties across realized
-populations. With the simulated data, we evaluated interval coverage
-across realization populations, but for the NLA data, we evaluated
-interval coverage within a single realized population for different
-samples. We did find that model-based coverages were similar to the
-design-based coverages, however, suggesting that for some realized
-populations it is reasonable to heuristically view data from separate
-random samples as being from approximately separate realized
-populations. But generally, if model-based intervals constructed from
-many random samples of a single realized population show improper
-coverage, this does not necessarily imply a deficiency in model-based
-inference. Tables for interval coverage in all six simulations scenarios
-are provided in the supporting information.
+data, coverages for the NLA data tend to increase with the sample sizes,
+coverages tend to be higher for symmetric responses than for skewed
+responses, and the local neighborhood variance was slightly too small
+for small \(n\), yielding slightly lower interval coverages than the
+other sampling-inference combinations. Recall that model-based inference
+defines interval coverage properties across realized populations. With
+the simulated data, we evaluated interval coverage across realization
+populations, but for the NLA data, we evaluated interval coverage within
+a single realized population for different samples. We did find that
+model-based coverages were similar to the design-based coverages,
+however, suggesting that for some realized populations it is reasonable
+to heuristically view data from separate random samples as being from
+approximately separate realized populations. But generally, if
+model-based intervals constructed from many random samples of a single
+realized population show improper coverage, this does not necessarily
+imply a deficiency in model-based inference. Tables for interval
+coverage in all six simulation scenarios are provided in the supporting
+information.
 
@@
 inference. Some we have discussed, but others we have not, and they are
-worthy of consideration in future research. First we discuss advantages
+worthy of consideration in future research. First, we discuss advantages
 of the design-based approach. Design-based inference is often
@@
 (Särndal et al., 2003). With the model-based approach, however, interval
-coverage is unlikely to be valid if the model's assumptions made do not
-not accurately reflect reality. Whether a model's assumptions accurately
+coverage is unlikely to be valid if the model assumptions made do not
+not accurately reflect reality. Whether model assumptions accurately
 reflect reality can be a challenging and sometimes impossible question
-to answer definitively. Now we discuss advantages of the model-based
-approach. The model-based approach can more naturally quantify the
-relationship between covariates (predictor variables) and the response
-variable than design-based approaches. Model-based inference also yields
-estimated spatial covariance parameters, which help better understand
-the dependence structure of the process in study. Model selection is
-also possible using model-based inference and criteria such as cross
-validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
-inference is capable of more efficient small-area estimation than
-design-based inference because model-based inference can leverage
-distributional assumptions in areas with few observed population units.
-Model-based approaches also accommodate unit-by-unit predictions at
-unobserved locations that can be used to construct informative
-visualizations like smoothed maps. Brus and De Gruijter (1997) provide a
-more thorough discussion regarding the benefits and drawbacks of the two
-approaches. In short, when deciding whether the design-based or
-model-based approach is more appropriate to implement, the benefits and
-drawbacks of each approach should be considered alongside the particular
-goals of the study.
+to answer definitively.
+
+Now, we discuss advantages of the model-based approach. The model-based
+approach can more naturally quantify the relationship between covariates
+(predictor variables) and the response variable than design-based
+approaches. Model-based inference also yields estimated spatial
+covariance parameters, which help better understand the dependence
+structure of the process in study. Model selection is also possible
+using model-based inference and criteria such as cross validation,
+likelihood ratio tests, or AIC (Akaike, 1974). Model-based inference is
+capable of more efficient small-area estimation than design-based
+inference because model-based inference can leverage distributional
+assumptions in areas with few observed population units. Model-based
+approaches also accommodate unit-by-unit predictions at unobserved
+locations that can be used to construct informative visualizations like
+smoothed maps. Brus and De Gruijter (1997) provide a more thorough
+discussion regarding the benefits and drawbacks of the two approaches.
+In short, when deciding whether the design-based or model-based approach
+is more appropriate to implement, the benefits and drawbacks of each
+approach should be considered alongside the particular goals of the
+study.
 
 There are many extensions of this research worthy of future
-consideration. Some (though not all) extensions to study include
-sampling with unequal inclusion probabilities, using different spatially
-balanced sampling approaches (instead of GRTS), using different spatial
-data configurations, using different spatial domains like stream
-networks (Ver Hoef and Peterson, 2010), using different response or
-covariance structures, and using spatial or external mean trends (which
-can be defined through covariates).
+consideration that include sampling with unequal inclusion
+probabilities, using different spatially balanced sampling approaches
+(instead of GRTS), using different spatial data configurations, using
+different spatial domains like stream networks (Ver Hoef and Peterson,
+2010), using different response or covariance structures, and using
+spatial or external mean trends (which can be defined through
+covariates).
 
@@
 We would like to thank the editors and anonymous reviewers for hard work
-and time spent providing us with thoughtful, valuable feedback that
-which greatly improved the manuscript.
+and time spent providing us with thoughtful, valuable feedback which
+greatly improved the manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-higham2021adjusting}{}%
+Higham, M., Ver Hoef, J., Madsen, L., Aderman, A., 2021. Adjusting a
+finite population block kriging estimator for imperfect detection.
+Environmetrics 32, e2654.
+
 \leavevmode\hypertarget{ref-hofman2021many}{}%

### 2022-04-18 inst/manuscript/supporting.Rmd
minor review updates

@@
        Approach == "Design GRTS" ~ "GRTS-DB",
-       Approach == "Design IRS" ~ "GRTS-MB",
-       Approach == "Model GRTS" ~ "SRS-DB",
+       Approach == "Design IRS" ~ "SRS-DB",
+       Approach == "Model GRTS" ~ "GRTS-MB",
        Approach == "Model IRS" ~ "SRS-MB"
@@
     )
-sim_table <- sim_table[c(3, 4, 1, 2), ]
+sim_table <- sim_table[c(2, 4, 1, 3), ]
 sim_table_out <- xtable(
@@
   digits = 4,
-  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
+  caption = paste0("Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
    type = "latex",
@@
     )
-data_table <- data_table[c(3, 4, 1, 2), ]
+data_table <- data_table[c(2, 4, 1, 3), ]
 data_table_out <- xtable(
@@
   digits = 4,
-  caption = paste0("Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\\% interval coverage (Coverage) in a data application scenario ", table_value, "."),
+  caption = paste0("Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in a data application scenario ", table_value, "."),
    type = "latex",

### 2022-04-18 inst/manuscript/supporting.tex
minor review updates

@@
 
-    \author[OSU]{Lisa Madsen}
+    \author[NOAA]{Jay M. Ver Hoef}
 
@@
 
-    \author[NOAA]{Jay M. Ver Hoef}
+    \author[OSU]{Lisa Madsen}
 
@@
   \hline
-SRS-DB & Gridded & Normal & 0\% & 50 & 0.0004 & 0.1922 & 0.9345 \\ 
+SRS-DB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1905 & 0.9490 \\ 
   SRS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1930 & 0.9420 \\ 
   GRTS-DB & Gridded & Normal & 0\% & 50 & 0.0005 & 0.1914 & 0.9170 \\ 
-  GRTS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1905 & 0.9490 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 50 & 0.0004 & 0.1922 & 0.9345 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 1.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 1.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 0\% & 50 & 0.0004 & 0.1965 & 0.8955 \\ 
+SRS-DB & Gridded & Skewed & 0\% & 50 & -0.0048 & 0.1988 & 0.9110 \\ 
   SRS-MB & Gridded & Skewed & 0\% & 50 & -0.0110 & 0.2055 & 0.9045 \\ 
   GRTS-DB & Gridded & Skewed & 0\% & 50 & 0.0025 & 0.1947 & 0.8710 \\ 
-  GRTS-MB & Gridded & Skewed & 0\% & 50 & -0.0048 & 0.1988 & 0.9110 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 50 & 0.0004 & 0.1965 & 0.8955 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 2.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 2.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 0\% & 50 & -0.0002 & 0.1964 & 0.9295 \\ 
+SRS-DB & Random & Normal & 0\% & 50 & 0.0031 & 0.1940 & 0.9425 \\ 
   SRS-MB & Random & Normal & 0\% & 50 & 0.0045 & 0.1953 & 0.9395 \\ 
   GRTS-DB & Random & Normal & 0\% & 50 & -0.0004 & 0.1951 & 0.9065 \\ 
-  GRTS-MB & Random & Normal & 0\% & 50 & 0.0031 & 0.1940 & 0.9425 \\ 
+  GRTS-MB & Random & Normal & 0\% & 50 & -0.0002 & 0.1964 & 0.9295 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 3.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 3.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 0\% & 50 & 0.0016 & 0.1941 & 0.8985 \\ 
+SRS-DB & Random & Skewed & 0\% & 50 & 0.0000 & 0.1861 & 0.9225 \\ 
   SRS-MB & Random & Skewed & 0\% & 50 & -0.0059 & 0.1932 & 0.9130 \\ 
   GRTS-DB & Random & Skewed & 0\% & 50 & 0.0041 & 0.1910 & 0.8740 \\ 
-  GRTS-MB & Random & Skewed & 0\% & 50 & 0.0000 & 0.1861 & 0.9225 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 50 & 0.0016 & 0.1941 & 0.8985 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 4.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 4.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 50\% & 50 & -0.0031 & 0.1472 & 0.9400 \\ 
+SRS-DB & Gridded & Normal & 50\% & 50 & -0.0013 & 0.1775 & 0.9460 \\ 
   SRS-MB & Gridded & Normal & 50\% & 50 & -0.0007 & 0.1617 & 0.9355 \\ 
   GRTS-DB & Gridded & Normal & 50\% & 50 & -0.0030 & 0.1501 & 0.9205 \\ 
-  GRTS-MB & Gridded & Normal & 50\% & 50 & -0.0013 & 0.1775 & 0.9460 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 50 & -0.0031 & 0.1472 & 0.9400 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 5.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 5.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 50\% & 50 & -0.0060 & 0.1605 & 0.9085 \\ 
+SRS-DB & Gridded & Skewed & 50\% & 50 & -0.0047 & 0.1891 & 0.9100 \\ 
   SRS-MB & Gridded & Skewed & 50\% & 50 & -0.0083 & 0.1787 & 0.8990 \\ 
   GRTS-DB & Gridded & Skewed & 50\% & 50 & -0.0041 & 0.1610 & 0.8995 \\ 
-  GRTS-MB & Gridded & Skewed & 50\% & 50 & -0.0047 & 0.1891 & 0.9100 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 50 & -0.0060 & 0.1605 & 0.9085 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 6.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 6.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 50\% & 50 & -0.0097 & 0.1507 & 0.9315 \\ 
+SRS-DB & Random & Normal & 50\% & 50 & 0.0061 & 0.1744 & 0.9445 \\ 
   SRS-MB & Random & Normal & 50\% & 50 & 0.0042 & 0.1615 & 0.9350 \\ 
   GRTS-DB & Random & Normal & 50\% & 50 & -0.0092 & 0.1513 & 0.9165 \\ 
-  GRTS-MB & Random & Normal & 50\% & 50 & 0.0061 & 0.1744 & 0.9445 \\ 
+  GRTS-MB & Random & Normal & 50\% & 50 & -0.0097 & 0.1507 & 0.9315 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 7.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 7.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 50\% & 50 & -0.0021 & 0.1624 & 0.9000 \\ 
+SRS-DB & Random & Skewed & 50\% & 50 & 0.0014 & 0.1804 & 0.9195 \\ 
   SRS-MB & Random & Skewed & 50\% & 50 & -0.0051 & 0.1808 & 0.9065 \\ 
   GRTS-DB & Random & Skewed & 50\% & 50 & 0.0021 & 0.1618 & 0.8880 \\ 
-  GRTS-MB & Random & Skewed & 50\% & 50 & 0.0014 & 0.1804 & 0.9195 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 50 & -0.0021 & 0.1624 & 0.9000 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 8.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 8.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 90\% & 50 & -0.0037 & 0.0960 & 0.9400 \\ 
+SRS-DB & Gridded & Normal & 90\% & 50 & 0.0055 & 0.1621 & 0.9355 \\ 
   SRS-MB & Gridded & Normal & 90\% & 50 & 0.0044 & 0.1137 & 0.9435 \\ 
   GRTS-DB & Gridded & Normal & 90\% & 50 & -0.0021 & 0.1072 & 0.9280 \\ 
-  GRTS-MB & Gridded & Normal & 90\% & 50 & 0.0055 & 0.1621 & 0.9355 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 50 & -0.0037 & 0.0960 & 0.9400 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 9.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 9.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 90\% & 50 & -0.0039 & 0.1039 & 0.9205 \\ 
+SRS-DB & Gridded & Skewed & 90\% & 50 & 0.0007 & 0.1546 & 0.9345 \\ 
   SRS-MB & Gridded & Skewed & 90\% & 50 & -0.0011 & 0.1217 & 0.9130 \\ 
   GRTS-DB & Gridded & Skewed & 90\% & 50 & -0.0017 & 0.1129 & 0.9180 \\ 
-  GRTS-MB & Gridded & Skewed & 90\% & 50 & 0.0007 & 0.1546 & 0.9345 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 50 & -0.0039 & 0.1039 & 0.9205 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 10.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 10.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 90\% & 50 & -0.0011 & 0.0940 & 0.9410 \\ 
+SRS-DB & Random & Normal & 90\% & 50 & 0.0058 & 0.1575 & 0.9470 \\ 
   SRS-MB & Random & Normal & 90\% & 50 & 0.0045 & 0.1093 & 0.9445 \\ 
   GRTS-DB & Random & Normal & 90\% & 50 & -0.0017 & 0.1054 & 0.9295 \\ 
-  GRTS-MB & Random & Normal & 90\% & 50 & 0.0058 & 0.1575 & 0.9470 \\ 
+  GRTS-MB & Random & Normal & 90\% & 50 & -0.0011 & 0.0940 & 0.9410 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 11.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 11.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1127 & 0.9225 \\ 
+SRS-DB & Random & Skewed & 90\% & 50 & -0.0022 & 0.1772 & 0.9305 \\ 
   SRS-MB & Random & Skewed & 90\% & 50 & -0.0046 & 0.1301 & 0.9180 \\ 
   GRTS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1250 & 0.9175 \\ 
-  GRTS-MB & Random & Skewed & 90\% & 50 & -0.0022 & 0.1772 & 0.9305 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1127 & 0.9225 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 12.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 12.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 0\% & 100 & -0.0010 & 0.1352 & 0.9445 \\ 
+SRS-DB & Gridded & Normal & 0\% & 100 & -0.0028 & 0.1352 & 0.9475 \\ 
   SRS-MB & Gridded & Normal & 0\% & 100 & -0.0026 & 0.1362 & 0.9405 \\ 
   GRTS-DB & Gridded & Normal & 0\% & 100 & -0.0011 & 0.1345 & 0.9210 \\ 
-  GRTS-MB & Gridded & Normal & 0\% & 100 & -0.0028 & 0.1352 & 0.9475 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 100 & -0.0010 & 0.1352 & 0.9445 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 13.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 13.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 0\% & 100 & 0.0077 & 0.1307 & 0.9155 \\ 
+SRS-DB & Gridded & Skewed & 0\% & 100 & -0.0001 & 0.1384 & 0.9210 \\ 
   SRS-MB & Gridded & Skewed & 0\% & 100 & -0.0024 & 0.1412 & 0.9150 \\ 
   GRTS-DB & Gridded & Skewed & 0\% & 100 & 0.0082 & 0.1304 & 0.8945 \\ 
-  GRTS-MB & Gridded & Skewed & 0\% & 100 & -0.0001 & 0.1384 & 0.9210 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 100 & 0.0077 & 0.1307 & 0.9155 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 14.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 14.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 0\% & 100 & 0.0019 & 0.1318 & 0.9460 \\ 
+SRS-DB & Random & Normal & 0\% & 100 & -0.0004 & 0.1339 & 0.9450 \\ 
   SRS-MB & Random & Normal & 0\% & 100 & -0.0007 & 0.1344 & 0.9435 \\ 
   GRTS-DB & Random & Normal & 0\% & 100 & 0.0021 & 0.1315 & 0.9280 \\ 
-  GRTS-MB & Random & Normal & 0\% & 100 & -0.0004 & 0.1339 & 0.9450 \\ 
+  GRTS-MB & Random & Normal & 0\% & 100 & 0.0019 & 0.1318 & 0.9460 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 15.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 15.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 0\% & 100 & 0.0039 & 0.1315 & 0.9230 \\ 
+SRS-DB & Random & Skewed & 0\% & 100 & 0.0021 & 0.1322 & 0.9280 \\ 
   SRS-MB & Random & Skewed & 0\% & 100 & 0.0005 & 0.1342 & 0.9280 \\ 
   GRTS-DB & Random & Skewed & 0\% & 100 & 0.0043 & 0.1311 & 0.8990 \\ 
-  GRTS-MB & Random & Skewed & 0\% & 100 & 0.0021 & 0.1322 & 0.9280 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 100 & 0.0039 & 0.1315 & 0.9230 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 16.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 16.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 50\% & 100 & -0.0003 & 0.0993 & 0.9435 \\ 
+SRS-DB & Gridded & Normal & 50\% & 100 & -0.0025 & 0.1207 & 0.9420 \\ 
   SRS-MB & Gridded & Normal & 50\% & 100 & -0.0007 & 0.1058 & 0.9425 \\ 
   GRTS-DB & Gridded & Normal & 50\% & 100 & 0.0000 & 0.1003 & 0.9300 \\ 
-  GRTS-MB & Gridded & Normal & 50\% & 100 & -0.0025 & 0.1207 & 0.9420 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 100 & -0.0003 & 0.0993 & 0.9435 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 17.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 17.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 50\% & 100 & -0.0019 & 0.1127 & 0.9175 \\ 
+SRS-DB & Gridded & Skewed & 50\% & 100 & 0.0023 & 0.1221 & 0.9315 \\ 
   SRS-MB & Gridded & Skewed & 50\% & 100 & 0.0002 & 0.1135 & 0.9160 \\ 
   GRTS-DB & Gridded & Skewed & 50\% & 100 & 0.0000 & 0.1113 & 0.9025 \\ 
-  GRTS-MB & Gridded & Skewed & 50\% & 100 & 0.0023 & 0.1221 & 0.9315 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 100 & -0.0019 & 0.1127 & 0.9175 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 18.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 18.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 50\% & 100 & -0.0010 & 0.0997 & 0.9380 \\ 
+SRS-DB & Random & Normal & 50\% & 100 & -0.0016 & 0.1214 & 0.9440 \\ 
   SRS-MB & Random & Normal & 50\% & 100 & -0.0001 & 0.1079 & 0.9440 \\ 
   GRTS-DB & Random & Normal & 50\% & 100 & -0.0012 & 0.1012 & 0.9235 \\ 
-  GRTS-MB & Random & Normal & 50\% & 100 & -0.0016 & 0.1214 & 0.9440 \\ 
+  GRTS-MB & Random & Normal & 50\% & 100 & -0.0010 & 0.0997 & 0.9380 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 19.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 19.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 50\% & 100 & -0.0002 & 0.1105 & 0.9235 \\ 
+SRS-DB & Random & Skewed & 50\% & 100 & 0.0020 & 0.1188 & 0.9395 \\ 
   SRS-MB & Random & Skewed & 50\% & 100 & -0.0020 & 0.1097 & 0.9290 \\ 
   GRTS-DB & Random & Skewed & 50\% & 100 & 0.0013 & 0.1104 & 0.9090 \\ 
-  GRTS-MB & Random & Skewed & 50\% & 100 & 0.0020 & 0.1188 & 0.9395 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 100 & -0.0002 & 0.1105 & 0.9235 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 20.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 20.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0602 & 0.9425 \\ 
+SRS-DB & Gridded & Normal & 90\% & 100 & -0.0027 & 0.1083 & 0.9460 \\ 
   SRS-MB & Gridded & Normal & 90\% & 100 & -0.0003 & 0.0701 & 0.9440 \\ 
   GRTS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0670 & 0.9325 \\ 
-  GRTS-MB & Gridded & Normal & 90\% & 100 & -0.0027 & 0.1083 & 0.9460 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0602 & 0.9425 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 21.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 21.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 90\% & 100 & 0.0004 & 0.0695 & 0.9280 \\ 
+SRS-DB & Gridded & Skewed & 90\% & 100 & -0.0015 & 0.1108 & 0.9345 \\ 
   SRS-MB & Gridded & Skewed & 90\% & 100 & -0.0038 & 0.0767 & 0.9265 \\ 
   GRTS-DB & Gridded & Skewed & 90\% & 100 & 0.0018 & 0.0746 & 0.9270 \\ 
-  GRTS-MB & Gridded & Skewed & 90\% & 100 & -0.0015 & 0.1108 & 0.9345 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 100 & 0.0004 & 0.0695 & 0.9280 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 22.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 22.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 90\% & 100 & -0.0007 & 0.0570 & 0.9480 \\ 
+SRS-DB & Random & Normal & 90\% & 100 & 0.0002 & 0.1071 & 0.9490 \\ 
   SRS-MB & Random & Normal & 90\% & 100 & 0.0011 & 0.0675 & 0.9475 \\ 
   GRTS-DB & Random & Normal & 90\% & 100 & 0.0009 & 0.0651 & 0.9430 \\ 
-  GRTS-MB & Random & Normal & 90\% & 100 & 0.0002 & 0.1071 & 0.9490 \\ 
+  GRTS-MB & Random & Normal & 90\% & 100 & -0.0007 & 0.0570 & 0.9480 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 23.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 23.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 90\% & 100 & 0.0023 & 0.0671 & 0.9230 \\ 
+SRS-DB & Random & Skewed & 90\% & 100 & -0.0019 & 0.1165 & 0.9325 \\ 
   SRS-MB & Random & Skewed & 90\% & 100 & -0.0031 & 0.0757 & 0.9315 \\ 
   GRTS-DB & Random & Skewed & 90\% & 100 & 0.0026 & 0.0751 & 0.9225 \\ 
-  GRTS-MB & Random & Skewed & 90\% & 100 & -0.0019 & 0.1165 & 0.9325 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 100 & 0.0023 & 0.0671 & 0.9230 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 24.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 24.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0843 & 0.9610 \\ 
+SRS-DB & Gridded & Normal & 0\% & 200 & 0.0005 & 0.0891 & 0.9465 \\ 
   SRS-MB & Gridded & Normal & 0\% & 200 & 0.0003 & 0.0894 & 0.9440 \\ 
   GRTS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0842 & 0.9520 \\ 
-  GRTS-MB & Gridded & Normal & 0\% & 200 & 0.0005 & 0.0891 & 0.9465 \\ 
+  GRTS-MB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0843 & 0.9610 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 25.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 25.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 0\% & 200 & 0.0005 & 0.0892 & 0.9405 \\ 
+SRS-DB & Gridded & Skewed & 0\% & 200 & -0.0018 & 0.0894 & 0.9440 \\ 
   SRS-MB & Gridded & Skewed & 0\% & 200 & -0.0023 & 0.0895 & 0.9445 \\ 
   GRTS-DB & Gridded & Skewed & 0\% & 200 & 0.0006 & 0.0891 & 0.9360 \\ 
-  GRTS-MB & Gridded & Skewed & 0\% & 200 & -0.0018 & 0.0894 & 0.9440 \\ 
+  GRTS-MB & Gridded & Skewed & 0\% & 200 & 0.0005 & 0.0892 & 0.9405 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 26.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 26.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 0\% & 200 & 0.0009 & 0.0882 & 0.9480 \\ 
+SRS-DB & Random & Normal & 0\% & 200 & 0.0019 & 0.0905 & 0.9385 \\ 
   SRS-MB & Random & Normal & 0\% & 200 & 0.0017 & 0.0907 & 0.9375 \\ 
   GRTS-DB & Random & Normal & 0\% & 200 & 0.0008 & 0.0881 & 0.9400 \\ 
-  GRTS-MB & Random & Normal & 0\% & 200 & 0.0019 & 0.0905 & 0.9385 \\ 
+  GRTS-MB & Random & Normal & 0\% & 200 & 0.0009 & 0.0882 & 0.9480 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 27.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 27.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9395 \\ 
+SRS-DB & Random & Skewed & 0\% & 200 & 0.0007 & 0.0867 & 0.9400 \\ 
   SRS-MB & Random & Skewed & 0\% & 200 & 0.0005 & 0.0869 & 0.9390 \\ 
   GRTS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9345 \\ 
-  GRTS-MB & Random & Skewed & 0\% & 200 & 0.0007 & 0.0867 & 0.9400 \\ 
+  GRTS-MB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9395 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 28.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 28.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 50\% & 200 & 0.0039 & 0.0679 & 0.9375 \\ 
+SRS-DB & Gridded & Normal & 50\% & 200 & -0.0016 & 0.0777 & 0.9520 \\ 
   SRS-MB & Gridded & Normal & 50\% & 200 & -0.0004 & 0.0678 & 0.9510 \\ 
   GRTS-DB & Gridded & Normal & 50\% & 200 & 0.0036 & 0.0688 & 0.9330 \\ 
-  GRTS-MB & Gridded & Normal & 50\% & 200 & -0.0016 & 0.0777 & 0.9520 \\ 
+  GRTS-MB & Gridded & Normal & 50\% & 200 & 0.0039 & 0.0679 & 0.9375 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 29.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 29.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 50\% & 200 & 0.0018 & 0.0691 & 0.9360 \\ 
+SRS-DB & Gridded & Skewed & 50\% & 200 & 0.0013 & 0.0816 & 0.9400 \\ 
   SRS-MB & Gridded & Skewed & 50\% & 200 & 0.0016 & 0.0747 & 0.9400 \\ 
   GRTS-DB & Gridded & Skewed & 50\% & 200 & 0.0019 & 0.0698 & 0.9325 \\ 
-  GRTS-MB & Gridded & Skewed & 50\% & 200 & 0.0013 & 0.0816 & 0.9400 \\ 
+  GRTS-MB & Gridded & Skewed & 50\% & 200 & 0.0018 & 0.0691 & 0.9360 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 30.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 30.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 50\% & 200 & 0.0009 & 0.0643 & 0.9510 \\ 
+SRS-DB & Random & Normal & 50\% & 200 & 0.0008 & 0.0816 & 0.9389 \\ 
   SRS-MB & Random & Normal & 50\% & 200 & 0.0005 & 0.0713 & 0.9444 \\ 
   GRTS-DB & Random & Normal & 50\% & 200 & 0.0005 & 0.0655 & 0.9474 \\ 
-  GRTS-MB & Random & Normal & 50\% & 200 & 0.0008 & 0.0816 & 0.9389 \\ 
+  GRTS-MB & Random & Normal & 50\% & 200 & 0.0009 & 0.0643 & 0.9510 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 31.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 31.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 50\% & 200 & -0.0007 & 0.0679 & 0.9385 \\ 
+SRS-DB & Random & Skewed & 50\% & 200 & -0.0041 & 0.0789 & 0.9510 \\ 
   SRS-MB & Random & Skewed & 50\% & 200 & -0.0048 & 0.0729 & 0.9350 \\ 
   GRTS-DB & Random & Skewed & 50\% & 200 & -0.0001 & 0.0680 & 0.9400 \\ 
-  GRTS-MB & Random & Skewed & 50\% & 200 & -0.0041 & 0.0789 & 0.9510 \\ 
+  GRTS-MB & Random & Skewed & 50\% & 200 & -0.0007 & 0.0679 & 0.9385 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 32.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 32.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0376 & 0.9475 \\ 
+SRS-DB & Gridded & Normal & 90\% & 200 & -0.0011 & 0.0730 & 0.9450 \\ 
   SRS-MB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0418 & 0.9445 \\ 
   GRTS-DB & Gridded & Normal & 90\% & 200 & -0.0010 & 0.0418 & 0.9455 \\ 
-  GRTS-MB & Gridded & Normal & 90\% & 200 & -0.0011 & 0.0730 & 0.9450 \\ 
+  GRTS-MB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0376 & 0.9475 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 33.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 33.}
 \end{table}
@@
   \hline
-SRS-DB & Gridded & Skewed & 90\% & 200 & 0.0010 & 0.0423 & 0.9375 \\ 
+SRS-DB & Gridded & Skewed & 90\% & 200 & 0.0038 & 0.0768 & 0.9430 \\ 
   SRS-MB & Gridded & Skewed & 90\% & 200 & -0.0007 & 0.0475 & 0.9365 \\ 
   GRTS-DB & Gridded & Skewed & 90\% & 200 & 0.0018 & 0.0482 & 0.9370 \\ 
-  GRTS-MB & Gridded & Skewed & 90\% & 200 & 0.0038 & 0.0768 & 0.9430 \\ 
+  GRTS-MB & Gridded & Skewed & 90\% & 200 & 0.0010 & 0.0423 & 0.9375 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 34.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 34.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Normal & 90\% & 200 & 0.0014 & 0.0368 & 0.9475 \\ 
+SRS-DB & Random & Normal & 90\% & 200 & 0.0026 & 0.0713 & 0.9495 \\ 
   SRS-MB & Random & Normal & 90\% & 200 & 0.0011 & 0.0408 & 0.9480 \\ 
   GRTS-DB & Random & Normal & 90\% & 200 & 0.0016 & 0.0418 & 0.9450 \\ 
-  GRTS-MB & Random & Normal & 90\% & 200 & 0.0026 & 0.0713 & 0.9495 \\ 
+  GRTS-MB & Random & Normal & 90\% & 200 & 0.0014 & 0.0368 & 0.9475 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 35.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 35.}
 \end{table}
@@
   \hline
-SRS-DB & Random & Skewed & 90\% & 200 & -0.0001 & 0.0396 & 0.9419 \\ 
+SRS-DB & Random & Skewed & 90\% & 200 & 0.0028 & 0.0708 & 0.9504 \\ 
   SRS-MB & Random & Skewed & 90\% & 200 & 0.0011 & 0.0437 & 0.9414 \\ 
   GRTS-DB & Random & Skewed & 90\% & 200 & -0.0008 & 0.0431 & 0.9504 \\ 
-  GRTS-MB & Random & Skewed & 90\% & 200 & 0.0028 & 0.0708 & 0.9504 \\ 
+  GRTS-MB & Random & Skewed & 90\% & 200 & -0.0001 & 0.0396 & 0.9419 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 36.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 36.}
 \end{table}
@@
   \hline
-SRS-DB & Hg ppb & 50 & -0.5604 & 10.4434 & 0.9070 \\ 
+SRS-DB & Hg ppb & 50 & 0.2539 & 12.9467 & 0.9215 \\ 
   SRS-MB & Hg ppb & 50 & 0.0573 & 11.3760 & 0.9020 \\ 
   GRTS-DB & Hg ppb & 50 & -0.1880 & 10.5558 & 0.8970 \\ 
-  GRTS-MB & Hg ppb & 50 & 0.2539 & 12.9467 & 0.9215 \\ 
+  GRTS-MB & Hg ppb & 50 & -0.5604 & 10.4434 & 0.9070 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 1.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 1.}
 \end{table}
@@
   \hline
-SRS-DB & ZMMI & 50 & -0.1191 & 1.9608 & 0.9355 \\ 
+SRS-DB & ZMMI & 50 & 0.0461 & 2.4502 & 0.9440 \\ 
   SRS-MB & ZMMI & 50 & 0.0000 & 2.1477 & 0.9395 \\ 
   GRTS-DB & ZMMI & 50 & -0.0047 & 1.9953 & 0.9210 \\ 
-  GRTS-MB & ZMMI & 50 & 0.0461 & 2.4502 & 0.9440 \\ 
+  GRTS-MB & ZMMI & 50 & -0.1191 & 1.9608 & 0.9355 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 2.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 2.}
 \end{table}
@@
   \hline
-SRS-DB & Hg ppb & 100 & -0.4156 & 7.0344 & 0.9230 \\ 
+SRS-DB & Hg ppb & 100 & -0.1390 & 8.9667 & 0.9290 \\ 
   SRS-MB & Hg ppb & 100 & -0.4213 & 7.4820 & 0.9190 \\ 
   GRTS-DB & Hg ppb & 100 & 0.0075 & 7.1415 & 0.9045 \\ 
-  GRTS-MB & Hg ppb & 100 & -0.1390 & 8.9667 & 0.9290 \\ 
+  GRTS-MB & Hg ppb & 100 & -0.4156 & 7.0344 & 0.9230 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 3.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 3.}
 \end{table}
@@
   \hline
-SRS-DB & ZMMI & 100 & -0.0968 & 1.2958 & 0.9450 \\ 
+SRS-DB & ZMMI & 100 & -0.0541 & 1.6901 & 0.9460 \\ 
   SRS-MB & ZMMI & 100 & -0.0942 & 1.3920 & 0.9445 \\ 
   GRTS-DB & ZMMI & 100 & -0.0066 & 1.3219 & 0.9205 \\ 
-  GRTS-MB & ZMMI & 100 & -0.0541 & 1.6901 & 0.9460 \\ 
+  GRTS-MB & ZMMI & 100 & -0.0968 & 1.2958 & 0.9450 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 4.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 4.}
 \end{table}
@@
   \hline
-SRS-DB & Hg ppb & 200 & -0.2762 & 4.3998 & 0.9410 \\ 
+SRS-DB & Hg ppb & 200 & -0.0876 & 5.9176 & 0.9450 \\ 
   SRS-MB & Hg ppb & 200 & -0.1775 & 4.6970 & 0.9400 \\ 
   GRTS-DB & Hg ppb & 200 & -0.0538 & 4.5051 & 0.9325 \\ 
-  GRTS-MB & Hg ppb & 200 & -0.0876 & 5.9176 & 0.9450 \\ 
+  GRTS-MB & Hg ppb & 200 & -0.2762 & 4.3998 & 0.9410 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 5.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 5.}
 \end{table}
@@
   \hline
-SRS-DB & ZMMI & 200 & -0.0311 & 0.8280 & 0.9450 \\ 
+SRS-DB & ZMMI & 200 & -0.0093 & 1.1311 & 0.9455 \\ 
   SRS-MB & ZMMI & 200 & -0.0736 & 0.9090 & 0.9450 \\ 
   GRTS-DB & ZMMI & 200 & 0.0193 & 0.8524 & 0.9460 \\ 
-  GRTS-MB & ZMMI & 200 & -0.0093 & 1.1311 & 0.9455 \\ 
+  GRTS-MB & ZMMI & 200 & -0.0311 & 0.8280 & 0.9450 \\ 
    \hline
@@
 }
-\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 6.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 6.}
 \end{table}

### 2022-04-18 inst/scripts/figures/grts.R

@@ -1,0 +1,121 @@
+# preliminaries
+
+## load packages
+library(tidyverse)
+library(here)
+library(cowplot)
+
+## do you want to write out? default of FALSE
+write_out <- TRUE
+
+## set base size
+base_size <- 9
+
+# set some graphical parameters
+ht <-  4
+wid <- 4 / 3 * ht
+
+## grid preliminaries ---------------------------------------------------------
+set.seed(5)
+n <- 20
+grts_data <- data.frame(lon = runif(n), lat = runif(n))
+min_lon <- 0
+max_lon <- 1
+min_lat <- 0
+max_lat <- 1
+grts_box <- data.frame(lon = c(0, 1, 1, 0, 0), lat = c(0, 0, 1, 1, 0))
+low <- 0
+q1 <- 0.25
+mid <- 0.5
+q3 <- 0.75
+high <- 1
+q18 <- 1 / 8
+q38 <- 3 / 8
+q58 <- 5 / 8
+q78 <- 7 / 8
+
+# figure a     ----------------------------------------------------------------
+annotate_size <- 30
+sample(0:3)
+grts_level1 <- ggplot() +
+  geom_path(grts_box, mapping = aes(x = lon, y = lat), col = "black") +
+  geom_point(data = grts_data, mapping = aes(x = lon, y = lat), size = 5, col = "grey35") +
+  geom_segment(aes(x = low, xend = high, y = low, yend = low), linetype = "dashed", col = "black") +
+  geom_segment(aes(x = low, xend = high, y = mid, yend = mid), linetype = "dashed", col = "black") +
+  geom_segment(aes(x = low, xend = high, y = high, yend = high), linetype = "dashed", col = "black") +
+  geom_segment(aes(x = low, xend = low, y = low, yend = high), linetype = "dashed", col = "black") +
+  geom_segment(aes(x = mid, xend = mid, y = low, yend = high), linetype = "dashed", col = "black") +
+  geom_segment(aes(x = high, xend = high, y = low, yend = high), linetype = "dashed", col = "black") +
+  annotate(geom = "text", x = q1, y = q1, label = "2", size = annotate_size, col = "black") +
+  annotate(geom = "text", x = q1, y = q3, label = "1", size = annotate_size, col = "black") +
+  annotate(geom = "text", x = q3, y = q3, label = "3", size = annotate_size, col = "black") +
+  annotate(geom = "text", x = q3, y = q1, label = "0", size = annotate_size, col = "black") +
[…]

### 2022-04-18 R/data_trial.R
more minor updates

@@ -51,10 +51,10 @@ data_trial <- function(seed = sample.int(1e7, size = 1), data, var,
 
   # subset data
   data <- data %>%
-  dplyr::select(SITE_ID, response = var, x = INDEX_LON_DD, y = INDEX_LAT_DD) %>%
+  dplyr::select(.data$SITE_ID, response = var, x = .data$INDEX_LON_DD, y = .data$INDEX_LAT_DD) %>%
     na.omit() %>%
-    dplyr::group_by(SITE_ID) %>%
-    dplyr::summarize(response = mean(response), x = mean(x), y = mean(y)) %>%
+    dplyr::group_by(.data$SITE_ID) %>%
+    dplyr::summarize(response = mean(.data$response), x = mean(.data$x), y = mean(.data$y)) %>%
     dplyr::ungroup()
 
   # set N

### 2022-04-18 inst/scripts/application.R

@@ -1,159 +1,0 @@
-# preliminaries
-
-## load packages
-library(tidyverse)
-library(here)
-library(sptotal)
-library(spsurvey) # depends on sf
-library(DvMsp) # need package installed to load data
-
-## do you want to write out? default of FALSE
-write_out <- TRUE
-
-## manipulate data
-
-data(nla_df)
-
-nla_df %>%
-  group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  ungroup() %>%
-  summarise(maxnonmiss = max(nonmiss))
-
-## HG never measured twice at the same site.
-nla_df %>%
-  group_by(SITE_ID) %>%
-  summarise(nonmiss = sum(!is.na(TOTALHG_RESULT))) %>%
-  summarise(sumnomiss = sum(nonmiss == 0))
-
-## 35 sites never have HG measured. These will be dropped as well as the duplicate sites.
-## in duplicate sites, fill in the value for RDis_IX
-nla_nomiss <- nla_df %>%
-  group_by(SITE_ID) %>%
-  fill(RDis_IX, .direction = "downup") %>%
-  ungroup() %>%
-  filter(!is.na(TOTALHG_RESULT)) %>%
-  rename(total_hg = "TOTALHG_RESULT")
-
-nla_nomiss_both <- nla_nomiss
-## check
-nla_nomiss_both %>%
-  group_by(SITE_ID) %>% count() %>%
-  ungroup() %>%
-  summarise(not1 = sum(n != 1))
-
-## transform coordinates
-coord_list <- LLtoTM(cm = mean(nla_nomiss_both$INDEX_LON_DD),
-                     lat = nla_nomiss_both$INDEX_LAT_DD,
-                     lon = nla_nomiss_both$INDEX_LON_DD)
-nla_nomiss_both$xcoords <- coord_list$xy[ ,1]
-nla_nomiss_both$ycoords <- coord_list$xy[ ,2]
-
-## find population mean
-realized_mean <- mean(nla_nomiss_both$total_hg)
-
-## set seed and other preliminaries
-set.seed(080421)
-n <- 100
-
-## IRS Sample, Model Analysis
-lakesobs <- nla_nomiss_both %>% sample_n(n)
-lakesunobs <- anti_join(nla_nomiss_both, lakesobs)
-lakesunobs$total_hg <- NA
-
[…]

### 2022-04-18 inst/scripts/sessionInfo.R

@@ -1,0 +1,64 @@
+library(cowplot)
+library(devtools)
+library(ggpubr)
+library(here)
+library(knitr)
+library(rticles)
+library(parallel)
+library(sf)
+library(spsurvey)
+library(sptotal)
+library(tidyverse)
+library(viridis)
+sessionInfo()
+#> R version 4.0.3 (2020-10-10)
+#> Platform: x86_64-w64-mingw32/x64 (64-bit)
+#> Running under: Windows 10 x64 (build 19042)
+#>
+#> Matrix products: default
+#>
+#> locale:
+#> [1] LC_COLLATE=English_United States.1252
+#> [2] LC_CTYPE=English_United States.1252
+#> [3] LC_MONETARY=English_United States.1252
+#> [4] LC_NUMERIC=C
+#> [5] LC_TIME=English_United States.1252
+#>
+#> attached base packages:
+#> [1] grid      parallel  stats     graphics  grDevices utils     datasets
+#> [8] methods   base
+#>
+#> other attached packages:
+#>  [1] viridis_0.6.1     viridisLite_0.4.0 forcats_0.5.1     stringr_1.4.0
+#>  [5] dplyr_1.0.5       purrr_0.3.4       readr_1.4.0       tidyr_1.1.3
+#>  [9] tibble_3.1.2      tidyverse_1.3.1   sptotal_1.0.0     spsurvey_5.4.0
+#> [13] survey_4.2        survival_3.2-7    Matrix_1.2-18     sf_1.0-1
+#> [17] rticles_0.22      knitr_1.36        here_1.0.1        ggpubr_0.4.0
+#> [21] ggplot2_3.3.5     devtools_2.4.0    usethis_2.0.1     cowplot_1.1.1
+#>
+#> loaded via a namespace (and not attached):
+#>  [1] minqa_1.2.4        colorspace_2.0-1   ggsignif_0.6.3     deldir_1.0-0
+#>  [5] ellipsis_0.3.2     class_7.3-17       rio_0.5.27         crossdes_1.1-1
+#>  [9] rprojroot_2.0.2    fs_1.5.0           rstudioapi_0.13    proxy_0.4-25
+#> [13] remotes_2.4.1      lubridate_1.7.10   fansi_0.5.0        xml2_1.3.2
+#> [17] splines_4.0.3      cachem_1.0.4       pkgload_1.2.1      jsonlite_1.7.2
[…]

### 2022-04-19 inst/manuscript/manuscript.Rmd
more minor updates

@@
 In the model-based approach, inference relies on distributional assumptions.
-We compare the approaches for finite population spatial data.
+We compare the approaches in a finite population spatial context.
   2.
@@
 
-When data cannot be collected for all units in a population (i.e., population units), data are collected on a subset of the population units -- this subset is called a sample.
+When data cannot be collected for all units in a population (population units), data are collected on a subset of the population units -- this subset is called a sample.
 There are two general approaches for using samples to make frequentist statistical inferences about a population: design-based and model-based.
@@
 In Section \ref{sec:mm}, we describe how we intend to compare performance of the approaches using simulated and real data.
-In Section \ref{sec:results}, we present analysis reslts for the simulated data and real data from the United States Environmental Protection Agency's 2012 National Lakes Assessment [@USEPA2012NLA].
+The real data is from the United States Environmental Protection Agency's 2012 National Lakes Assessment (NLA) [@USEPA2012NLA].
+In Section \ref{sec:results}, we present analysis results for the simulated data and NLA data.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 The inclusion probability of a population unit is calculated by summing each sample's probability of selection over all samples that contain the population unit.
-Inclusion probabilities are often when selecting samples and estimating population parameters.
+Inclusion probabilities are often used when selecting samples and estimating population parameters.
 
@@
 Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible samples.
-Sample means, for example, are asymptotically normal (Gaussian) by the Central Limit Theorem (under some assumptions).
+Sample means, for example, are asymptotically normal (Gaussian) by the central limit theorem (under some assumptions).
 If we repeatedly select samples from the population, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true fixed population mean.
@@
 It is also important to quantify the uncertainty in $\hat{\tau}_{ht}$.
-@horvitz1952generalization and @sen1953estimate provide variance estimators for $\hat{\tau}_{ht}$, but these estimators have two drawbacks.
+The Horvitz-Thompson [@horvitz1952generalization] and Sen-Yates-Grundy [@sen1953estimate; @yates1953selection] variance estimators are often used to estimate $\text{Var}(\hat{\tau}_{ht})$, but these estimators have two drawbacks.
 First, they rely on calculating $\pi_{ij}$, the probability that population unit $i$ and population unit $j$ are both in the sample -- this quantity can be challenging if not impossible to calculate analytically for GRTS samples.
@@
 The local neighborhood variance estimator does not rely on $\pi_{ij}$ and estimates the variance of $\hat{\tau}$ conditional on the random properties of the GRTS sample -- the idea being that this conditioning should yield a more precise estimate of $\hat{\tau}$.
-They show that the contribution from each sample unit (population unit in the sample) to the overall variance is dominated by local variation.
-Thus the local neighborhood variance estimator is a weighted sum of variance estimates from each sample unit's local neighborhood.
-These local neighborhoods contain the sample unit itself and its three nearest neighbors among all other sample units.
+They show that the contribution from each sampled population unit to the overall variance is dominated by local variation.
+Thus the local neighborhood variance estimator is a weighted sum of variance estimates from each sampled population unit's local neighborhood.
+These local neighborhoods contain the sampled population unit itself and its three nearest neighbors (among all other sampled population units).
 For more details, see @stevens2003variance.
@@
 80-93 discuss covariance functions that are not second-order stationary, not isotropic, or not either.
-A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics]; one example is the exponential covariance function (@cressie1993statistics provides a thorough list of spatial covariance functions).
+A variety of flexible covariance functions can be used to model $\bm{\delta}$ [@cressie1993statistics] -- one example is the exponential covariance function.
+@cressie1993statistics provides a thorough list of spatial covariance functions.
 The $i,j$th element of the exponential covariance matrix, $\cov(\bm{\delta})$, is
@@
 The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on using restricted maximum likelihood (REML) [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
-REML is preferred over maximum likelihood (ML) because ML estimates can be badly biased for small sample sizes, due to the fact that ML makes no adjustment for the simultaneous estimation of $\bm{\beta}$ and $\bm{\delta}$ [@patterson1971recovery].
+REML is preferred over maximum likelihood (ML) because ML estimates can be badly biased for small sample sizes, due to the fact that ML makes no adjustment for the simultaneous estimation of $\bm{\beta}$ and $\bm{\theta}$ [@patterson1971recovery].
 Minus twice the REML log-likelihood of the sampled sites is given by
@@
 where $\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s$ and $\bm{\Sigma}_{ss}$  is the covariance matrix of the sampled sites.
-Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\theta}}_{reml}$, the REML estimates of $\bm{\delta}$.
+Minimizing Equation$~$\ref{equation:reml} yields $\bm{\hat{\theta}}_{reml}$, the REML estimates of $\bm{\theta}$.
 Then $\bm{\hat{\beta}}_{reml}$, the REML estimate of $\bm{\beta}$, is given by $(\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s$, where $\bm{\hat{\Sigma}}_{ss}$ is $\bm{\Sigma}_{ss}$ evaluated at $\bm{\hat{\theta}}_{reml}$.
@@
 where $z_i$ is the $i$th response value, $\bar{z}$ is the mean of all $z_i$, $n$ is the sample size, $N$ is the population size, and $f = (1 - n / N)$ ($f$ is often called the finite population correction factor).
-For GRTS-DB, the Horvitz-Thompson esetimator was used to estimate means and the local neighborhood variance was used to estimate variances.
+For GRTS-DB, the Horvitz-Thompson estimator was used to estimate means and the local neighborhood variance was used to estimate variances.
 For SRS-MB and GRTS-MB, FPBK was used to estimate means and variances using restricted maximum likelihood.
+SRS, GRTS sampling, and design-based inference were implemented using the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey].
+FPBK was implemented using the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
 
 We used simulated data to compare the sampling-inference combinations across many realized populations from the same data-generating stochastic process.
+We then used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
 With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
-We then used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
 With the real data, we were in control of only the random sampling process.
@@
 
-## National Lakes Assessment (Real) Data {#sec:mm_app}
+## National Lakes Assessment Data {#sec:mm_app}
 
-The United States Environmental Protection Agency (USEPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) to assess the water quality of various bodies of water in the contiguous United States.
+The United States Environmental Protection Agency (USEPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) to assess the water quality of various bodies of water in the conterminous United States.
 One component of NARS is the National Lakes Assessment (NLA), which measures various aspects of lake health and water quality.
@@
   \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
-  \caption{Spatial map of mercury (Hg ppb) population.}
+  \caption{Spatial map of the Hg ppb population.}
   \label{fig:mercury_map}
@@
   \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
-  \caption{Histogram of the mercury (Hg ppb) population.}
+  \caption{Histogram of the Hg ppb population.}
   \label{fig:mercury_hist}
@@
   \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
-  \caption{Semivariogram of the mercury (Hg ppb) population.}
+  \caption{Semivariogram of the Hg ppb population.}
   \label{fig:mercury_sv_plot}
@@
 In total, there were six separate scenarios (two responses crossed with three sample sizes).
-We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
+For each sample from each scenario, parameters were estimated using the same approaches as described for simulated data.
+We also used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
 Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean ($\hat{\mu}_i$) and the population mean ($\mu$) (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials.
@@
   \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
-  \caption{Relative RMSE in the simulation study for the four sampling-inference combinations and three sample sizes in the random location layout.
+  \caption{Simulated data relative RMSE for the four sampling-inference combinations and three sample sizes in the random location layout.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 Coverage in the normal response scenarios is usually near 95%, while coverage in the skewed response scenarios usually varies from 90% to 95% but increases with the sample size.
-At a sample size of 200, all four sampling-inference combinations have approximately 95% interval coverage in both response scenarios for all dependent random error proportions.
+At a sample size of 200, all four sampling-inference combinations have approximately 95% interval coverage in both response scenarios for all DRE proportions.
 These interval coverage conclusions are similar to those observed in the grid location layout, so we omit a figure and discussion regarding the grid location layout here.
@@
   \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
-  \caption{Interval coverage in the simulation study for the four sampling-inference combinations and three sample sizes in the random location layout.
+  \caption{Simulated data interval coverage for the four sampling-inference combinations and three sample sizes in the random location layout.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 
-## National Lakes Assessment (Real) Data {#sec:r_app}
+## National Lakes Assessment Data {#sec:r_app}
 
@@
 Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB all have noticeably lower RMSE than SRS-DB.
-Tables for RMSE in all six simulation scenarios are provided in the supporting information.
+Tables for RMSE in all six scenarios are provided in the supporting information.
 
@@
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{Relative RMSE in the data study for the four sampling-inference combinations.
+  \caption{NLA data relative RMSE for the four sampling-inference combinations.
 The rows indicate the proportion of dependent error and the columns indicate the response type.
@@
 Recall that model-based inference defines interval coverage properties across realized populations.
-With the simulated data, we evaluated interval coverage across realization populations, but for the NLA data, we evaluated interval coverage within a single realized population for different samples.
+With the simulated data, we evaluated interval coverage across realized populations, but for the NLA data, we evaluated interval coverage within a single realized population for different samples.
 We did find that model-based coverages were similar to the design-based coverages, however, suggesting that for some realized populations it is reasonable to heuristically view data from separate random samples as being from approximately separate realized populations.
@@
 
+<!-- Model-based approach is an approximation to the randomization distribution -->
 \begin{figure}
@@
   \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
-  \caption{Interval coverage in the data study for the four sampling-inference combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
+  \caption{NLA data interval coverage for the four sampling-inference combinations.
+The rows indicate the proportion of dependent error and the columns indicate the response type.
+The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
   \label{fig:data_figconf}
@@
 
-The design-based and model-based approaches to statistical inference are fundamentally different paradigms.
+The design-based and model-based approaches to frequentist statistical inference rest on fundamentally different foundations.
 Design-based approaches rely on random sampling to estimate population parameters.
@@
 First, we discuss advantages of the design-based approach.
-Design-based inference is often computationally efficient, while model-based inference can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on inverting a covariance matrix.
+Design-based inference is often computationally efficient, while model-based inference can be computationally burdensome, especially for likelihood-based estimation methods like REML that rely on the inverse of a covariance matrix.
 Design-based inference easily handles binary data through a straightforward application of the Horvitz-Thompson estimator.
-In contrast, analyzing binary data using model-based inference generally requires a logistic mixed regression model, which can be difficult to estimate and interpret [@bolker2009generalized].
+In contrast, analyzing binary data using model-based inference generally requires a logistic mixed regression model, the parameters of which can be difficult to estimate and interpret [@bolker2009generalized].
 An advantage of design-based inference is that interval coverage is valid (has the proper coverage rate) as long as 1) the sample is sufficiently large to ensure the statistic's sampling distribution is approximately normal and 2) the variance estimator is consistent [@brus1997random; @sarndal2003model].
@@
 The supplementary \textbf{\textsf{R}} package is hosted on GitHub.
-Instructions for download at available at
+Instructions for download are available at
 

### 2022-04-19 inst/manuscript/manuscript.tex
more minor updates

@@
     model-based approach, inference relies on distributional assumptions.
-    We compare the approaches for finite population spatial data.
+    We compare the approaches in a finite population spatial context.
   \item
@@
 
-When data cannot be collected for all units in a population (i.e.,
-population units), data are collected on a subset of the population
-units -- this subset is called a sample. There are two general
-approaches for using samples to make frequentist statistical inferences
-about a population: design-based and model-based. In the design-based
-approach, inference relies on randomly assigning some population units
-to be in the sample (random sampling). Alternatively, in the model-based
-approach, inference relies on distributional assumptions about the
-underlying data-generating stochastic process (superpopulation). Each
-paradigm has a deep historical context (Sterba, 2009) and its own set of
-benefits and drawbacks (Brus and De Gruijter, 1997; Hansen et al.,
-1983). In this manuscript, we compare design-based and model-based
-approaches for finite population spatial sampling and inference.
+When data cannot be collected for all units in a population (population
+units), data are collected on a subset of the population units -- this
+subset is called a sample. There are two general approaches for using
+samples to make frequentist statistical inferences about a population:
+design-based and model-based. In the design-based approach, inference
+relies on randomly assigning some population units to be in the sample
+(random sampling). Alternatively, in the model-based approach, inference
+relies on distributional assumptions about the underlying
+data-generating stochastic process (superpopulation). Each paradigm has
+a deep historical context (Sterba, 2009) and its own set of benefits and
+drawbacks (Brus and De Gruijter, 1997; Hansen et al., 1983). In this
+manuscript, we compare design-based and model-based approaches for
+finite population spatial sampling and inference.
 
@@
 intend to compare performance of the approaches using simulated and real
-data. In Section \ref{sec:results}, we present analysis reslts for the
-simulated data and real data from the United States Environmental
-Protection Agency's 2012 National Lakes Assessment (USEPA, 2012). And in
-Section \ref{sec:discussion}, we end with a discussion and provide
-directions for future research.
+data. The real data is from the United States Environmental Protection
+Agency's 2012 National Lakes Assessment (NLA) (USEPA, 2012). In Section
+\ref{sec:results}, we present analysis results for the simulated data
+and NLA data. And in Section \ref{sec:discussion}, we end with a
+discussion and provide directions for future research.
 
@@
 selection over all samples that contain the population unit. Inclusion
-probabilities are often when selecting samples and estimating population
-parameters.
+probabilities are often used when selecting samples and estimating
+population parameters.
 
@@
 that incorporate all possible samples. Sample means, for example, are
-asymptotically normal (Gaussian) by the Central Limit Theorem (under
+asymptotically normal (Gaussian) by the central limit theorem (under
 some assumptions). If we repeatedly select samples from the population,
@@
 It is also important to quantify the uncertainty in \(\hat{\tau}_{ht}\).
-Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
-for \(\hat{\tau}_{ht}\), but these estimators have two drawbacks. First,
-they rely on calculating \(\pi_{ij}\), the probability that population
-unit \(i\) and population unit \(j\) are both in the sample -- this
-quantity can be challenging if not impossible to calculate analytically
-for GRTS samples. Second, these estimators tend to ignore the spatial
-locations of the population units. To address these two drawbacks
-simultaneously, Stevens and Olsen (2003) proposed the local neighborhood
-variance estimator. The local neighborhood variance estimator does not
-rely on \(\pi_{ij}\) and estimates the variance of \(\hat{\tau}\)
-conditional on the random properties of the GRTS sample -- the idea
-being that this conditioning should yield a more precise estimate of
-\(\hat{\tau}\). They show that the contribution from each sample unit
-(population unit in the sample) to the overall variance is dominated by
+The Horvitz-Thompson (Horvitz and Thompson, 1952) and Sen-Yates-Grundy
+(Sen, 1953; Yates and Grundy, 1953) variance estimators are often used
+to estimate \(\text{Var}(\hat{\tau}_{ht})\), but these estimators have
+two drawbacks. First, they rely on calculating \(\pi_{ij}\), the
+probability that population unit \(i\) and population unit \(j\) are
+both in the sample -- this quantity can be challenging if not impossible
+to calculate analytically for GRTS samples. Second, these estimators
+tend to ignore the spatial locations of the population units. To address
+these two drawbacks simultaneously, Stevens and Olsen (2003) proposed
+the local neighborhood variance estimator. The local neighborhood
+variance estimator does not rely on \(\pi_{ij}\) and estimates the
+variance of \(\hat{\tau}\) conditional on the random properties of the
+GRTS sample -- the idea being that this conditioning should yield a more
+precise estimate of \(\hat{\tau}\). They show that the contribution from
+each sampled population unit to the overall variance is dominated by
 local variation. Thus the local neighborhood variance estimator is a
-weighted sum of variance estimates from each sample unit's local
-neighborhood. These local neighborhoods contain the sample unit itself
-and its three nearest neighbors among all other sample units. For more
-details, see Stevens and Olsen (2003).
+weighted sum of variance estimates from each sampled population unit's
+local neighborhood. These local neighborhoods contain the sampled
+population unit itself and its three nearest neighbors (among all other
+sampled population units). For more details, see Stevens and Olsen
+(2003).
 
@@
 stochastic process. We summarize some of the basic principles of FPBK
-next -- see Ver Hoef (2008) for technical details and see Higham et al.
-(2021) for an extension to cases of imperfect detection among population
-units. Let
+next -- see Ver Hoef (2008) for technical details and see Higham, J. Ver
+Hoef, et al. (2021) for an extension to cases of imperfect detection
+among population units. Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 flexible covariance functions can be used to model \(\bm{\delta}\)
-(Cressie, 1993); one example is the exponential covariance function
-(Cressie (1993) provides a thorough list of spatial covariance
-functions). The \(i,j\)th element of the exponential covariance matrix,
+(Cressie, 1993) -- one example is the exponential covariance function.
+Cressie (1993) provides a thorough list of spatial covariance functions.
+The \(i,j\)th element of the exponential covariance matrix,
 \(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
@@
 the fact that ML makes no adjustment for the simultaneous estimation of
-\(\bm{\beta}\) and \(\bm{\delta}\) (Patterson and Thompson, 1971). Minus
+\(\bm{\beta}\) and \(\bm{\theta}\) (Patterson and Thompson, 1971). Minus
 twice the REML log-likelihood of the sampled sites is given by
@@
 Minimizing Equation\(~\)\ref{equation:reml} yields
-\(\bm{\hat{\theta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
+\(\bm{\hat{\theta}}_{reml}\), the REML estimates of \(\bm{\theta}\).
 Then \(\bm{\hat{\beta}}_{reml}\), the REML estimate of \(\bm{\beta}\),
@@
 finite population correction factor). For GRTS-DB, the Horvitz-Thompson
-esetimator was used to estimate means and the local neighborhood
-variance was used to estimate variances. For SRS-MB and GRTS-MB, FPBK
-was used to estimate means and variances using restricted maximum
-likelihood.
+estimator was used to estimate means and the local neighborhood variance
+was used to estimate variances. For SRS-MB and GRTS-MB, FPBK was used to
+estimate means and variances using restricted maximum likelihood. SRS,
+GRTS sampling, and design-based inference were implemented using the
+\texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle et al., 2022).
+FPBK was implemented using the \texttt{sptotal} \textbf{\textsf{R}}
+package (Higham, Ver Hoef, et al., 2021).
 
@@
 across many realized populations from the same data-generating
-stochastic process. With the simulated data, we were in control of the
-data-generating stochastic process and the random sampling process. We
-then used real data from the 2012 National Lakes Assessment to compare
-the sampling-inference combinations within a single realized population
-(which is typically the case in reality). With the real data, we were in
-control of only the random sampling process.
+stochastic process. We then used real data from the 2012 National Lakes
+Assessment to compare the sampling-inference combinations within a
+single realized population (which is typically the case in reality).
+With the simulated data, we were in control of the data-generating
+stochastic process and the random sampling process. With the real data,
+we were in control of only the random sampling process.
 
@@
 \hypertarget{sec:mm_app}{%
-\subsection{National Lakes Assessment (Real) Data}\label{sec:mm_app}}
+\subsection{National Lakes Assessment Data}\label{sec:mm_app}}
 
@@
 tribes periodically conduct National Aquatic Research Surveys (NARS) to
-assess the water quality of various bodies of water in the contiguous
+assess the water quality of various bodies of water in the conterminous
 United States. One component of NARS is the National Lakes Assessment
@@
 total, there were six separate scenarios (two responses crossed with
-three sample sizes). We used the same evaluation metrics as for the
-simulated data: mean bias, RMSE, and interval coverage. Mean bias is
-taken as the average deviation between each sample's estimated (or
-predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
-ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
-where \(i\) indexes simulation trials. Root-mean-squared error is taken
-as the square root of the average squared deviation between each
-sample's estimated (or predicted) mean and its population mean:
+three sample sizes). For each sample from each scenario, parameters were
+estimated using the same approaches as described for simulated data. We
+also used the same evaluation metrics as for the simulated data: mean
+bias, RMSE, and interval coverage. Mean bias is taken as the average
+deviation between each sample's estimated (or predicted) mean
+(\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of ZMMI or Hg ppb):
+\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
+indexes simulation trials. Root-mean-squared error is taken as the
+square root of the average squared deviation between each sample's
+estimated (or predicted) mean and its population mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
@@
 approximately 95\% interval coverage in both response scenarios for all
-dependent random error proportions. These interval coverage conclusions
-are similar to those observed in the grid location layout, so we omit a
-figure and discussion regarding the grid location layout here. Tables
-for interval coverage in all 36 simulation scenarios are provided in the
-supporting information.
+DRE proportions. These interval coverage conclusions are similar to
+those observed in the grid location layout, so we omit a figure and
+discussion regarding the grid location layout here. Tables for interval
+coverage in all 36 simulation scenarios are provided in the supporting
+information.
 
@@
 \hypertarget{sec:r_app}{%
-\subsection{National Lakes Assessment (Real) Data}\label{sec:r_app}}
+\subsection{National Lakes Assessment Data}\label{sec:r_app}}
 
@@
 all have noticeably lower RMSE than SRS-DB. Tables for RMSE in all six
-simulation scenarios are provided in the supporting information.
+scenarios are provided in the supporting information.
 
@@
 defines interval coverage properties across realized populations. With
-the simulated data, we evaluated interval coverage across realization
+the simulated data, we evaluated interval coverage across realized
 populations, but for the NLA data, we evaluated interval coverage within
@@
 
-The design-based and model-based approaches to statistical inference are
-fundamentally different paradigms. Design-based approaches rely on
-random sampling to estimate population parameters. Model-based
-approaches rely on distributional assumptions to predict realized values
-of a data-generating stochastic process. Though model-based approaches
-do not rely on random sampling, it can still be beneficial as a way to
-guard against preferential sampling. While design-based and model-based
-approaches have often been compared in the literature from theoretical
-and analytical perspectives, our contribution lies in studying them for
-finite population spatial data while implementing GRTS sampling and the
-local neighborhood variance estimator. Aside from the theoretical
-differences described throughout the manuscript, a few analytical
-findings from the simulated and real data studies were particularly
-notable. All sampling-inference combinations had approximately zero mean
-bias. Independent of the inference approach, GRTS-DB and GRTS-MB had
-lower RMSE than their SRS counterparts. Though GRTS-DB and GRTS-MB
-generally had very similar RMSE, SRS-MB tended to have much lower RMSE
-than SRS-DB, suggesting that the model-based inference mitigated much of
-the inefficiency in RMSE from SRS. As the proportion of dependent random
-error in the simulated data increased, SRS-MB, GRTS-DB, and GRTS-MB
-become increasingly more efficient (lower RMSE) than SRS-DB. Interval
-coverage tended to be higher for the symmetric responses than skewed
-responses and tended to increase with the sample size. At a sample size
-of \(n = 200\), generally all interval coverages were near the desired
-value of 95\%.
+The design-based and model-based approaches to frequentist statistical
+inference rest on fundamentally different foundations. Design-based
+approaches rely on random sampling to estimate population parameters.
+Model-based approaches rely on distributional assumptions to predict
+realized values of a data-generating stochastic process. Though
+model-based approaches do not rely on random sampling, it can still be
+beneficial as a way to guard against preferential sampling. While
+design-based and model-based approaches have often been compared in the
+literature from theoretical and analytical perspectives, our
+contribution lies in studying them for finite population spatial data
+while implementing GRTS sampling and the local neighborhood variance
+estimator. Aside from the theoretical differences described throughout
+the manuscript, a few analytical findings from the simulated and real
+data studies were particularly notable. All sampling-inference
+combinations had approximately zero mean bias. Independent of the
+inference approach, GRTS-DB and GRTS-MB had lower RMSE than their SRS
+counterparts. Though GRTS-DB and GRTS-MB generally had very similar
+RMSE, SRS-MB tended to have much lower RMSE than SRS-DB, suggesting that
+the model-based inference mitigated much of the inefficiency in RMSE
+from SRS. As the proportion of dependent random error in the simulated
+data increased, SRS-MB, GRTS-DB, and GRTS-MB become increasingly more
+efficient (lower RMSE) than SRS-DB. Interval coverage tended to be
+higher for the symmetric responses than skewed responses and tended to
+increase with the sample size. At a sample size of \(n = 200\),
+generally all interval coverages were near the desired value of 95\%.
 
@@
 computationally burdensome, especially for likelihood-based estimation
-methods like REML that rely on inverting a covariance matrix.
+methods like REML that rely on the inverse of a covariance matrix.
 Design-based inference easily handles binary data through a
@@
 contrast, analyzing binary data using model-based inference generally
-requires a logistic mixed regression model, which can be difficult to
-estimate and interpret (Bolker et al., 2009). An advantage of
-design-based inference is that interval coverage is valid (has the
-proper coverage rate) as long as 1) the sample is sufficiently large to
-ensure the statistic's sampling distribution is approximately normal and
-2) the variance estimator is consistent (Brus and De Gruijter, 1997;
-Särndal et al., 2003). This is because with the design-based approach,
-the sampling plan and inclusion probabilities are specified directly by
-the researcher. An advantage of SRS-DB not previously mentioned is that
-it is likely to be valid given the consistency of its variance estimator
-(Särndal et al., 2003). With the model-based approach, however, interval
-coverage is unlikely to be valid if the model assumptions made do not
-not accurately reflect reality. Whether model assumptions accurately
-reflect reality can be a challenging and sometimes impossible question
-to answer definitively.
+requires a logistic mixed regression model, the parameters of which can
+be difficult to estimate and interpret (Bolker et al., 2009). An
+advantage of design-based inference is that interval coverage is valid
+(has the proper coverage rate) as long as 1) the sample is sufficiently
+large to ensure the statistic's sampling distribution is approximately
+normal and 2) the variance estimator is consistent (Brus and De
+Gruijter, 1997; Särndal et al., 2003). This is because with the
+design-based approach, the sampling plan and inclusion probabilities are
+specified directly by the researcher. An advantage of SRS-DB not
+previously mentioned is that it is likely to be valid given the
+consistency of its variance estimator (Särndal et al., 2003). With the
+model-based approach, however, interval coverage is unlikely to be valid
+if the model assumptions made do not not accurately reflect reality.
+Whether model assumptions accurately reflect reality can be a
+challenging and sometimes impossible question to answer definitively.
 
@@
 supplementary \textbf{\textsf{R}} package is hosted on GitHub.
-Instructions for download at available at
+Instructions for download are available at
 
@@
 
+\leavevmode\hypertarget{ref-higham2021sptotal}{}%
+Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
+Predicting totals and weighted sums from spatial data.
+
 \leavevmode\hypertarget{ref-higham2021adjusting}{}%
@@
 
+\leavevmode\hypertarget{ref-yates1953selection}{}%
+Yates, F., Grundy, P.M., 1953. Selection without replacement from within
+strata with probability proportional to size. Journal of the Royal
+Statistical Society: Series B (Methodological) 15, 253--261.
+
 \end{document}

### 2022-04-19 inst/manuscript/supporting.Rmd
more minor updates

@@
 
-# National Lakes Assessment (Real) Data {#sec:datatabs}
+\clearpage
 
-Next we present tables summarizing mean bias, RMSE, and interval coverage for ZMMI and Hg ppb data.
+# National Lakes Assessment Data {#sec:datatabs}
+
+Next we present tables summarizing mean bias, RMSE, and interval coverage for all six data scenarios.
 ```{r}

### 2022-04-19 inst/manuscript/supporting.tex
more minor updates

@@
 
+\clearpage
+
 \hypertarget{sec:datatabs}{%
-\section{National Lakes Assessment (Real) Data}\label{sec:datatabs}}
+\section{National Lakes Assessment Data}\label{sec:datatabs}}
 
 Next we present tables summarizing mean bias, RMSE, and interval
-coverage for ZMMI and Hg ppb data.
+coverage for all six data scenarios.
 

### 2022-04-19 inst/manuscript/manuscript.Rmd
more minor updates

@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-Within each simulation scenario and trial, IRS and GRTS samples were selected and then design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
-With model-based inference, covariance parameters and the mean were estimated (using REML) separately for each trial.
+Within each simulation scenario and trial, SRS and GRTS samples were selected and then design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the realized mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
+With model-based inference, covariance parameters and the realized mean were estimated (using REML) separately for each trial.
 After all 2000 trials, we summarized the long-run performance of the sampling-inference combination in each scenario by calculating mean bias, root-mean-squared error, and interval coverage.
@@
 
-We studied performance of the four sampling-inference combinations by selecting 2000 random IRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
+We studied performance of the four sampling-inference combinations by selecting 2000 random SRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
 In total, there were six separate scenarios (two responses crossed with three sample sizes).
-For each sample from each scenario, parameters were estimated using the same approaches as described for simulated data.
-We also used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
+Within each SRS and GRTS sample, design-based and model-based inferences were used to estimate or predict the population mean and construct 95% coverage intervals.
+With model-based inference, the exponential covariance was assumed, and covariance parameters and the population mean were estimated using REML (separately for each SRS and GRTS sample).
+We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
 Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean ($\hat{\mu}_i$) and the population mean ($\mu$) (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials.
@@
 The difference in RMSE among GRTS-MB and GRTS-DB tends to be quite small.
-When $n = 50$, SRS-MB RMSE is approximately evenly between IRS-DB RMSE and GRTS-MB RMSE, but for the larger sample sizes ($n = 100$, $n = 200$), SRS-MB RMSE is closer to GRTS-MB RMSE.
+When $n = 50$, SRS-MB RMSE is approximately evenly between SRS-DB RMSE and GRTS-MB RMSE, but for the larger sample sizes ($n = 100$, $n = 200$), SRS-MB RMSE is closer to GRTS-MB RMSE.
 Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB all have noticeably lower RMSE than SRS-DB.

### 2022-04-19 inst/manuscript/manuscript.tex
more minor updates

@@
 In each of the 36 simulation scenarios, there were 2000 independent
-simulation trials. Within each simulation scenario and trial, IRS and
+simulation trials. Within each simulation scenario and trial, SRS and
 GRTS samples were selected and then design-based and model-based
 inferences were used to estimate (design-based) or predict (model-based)
-the mean and construct 95\% confidence (design-based) or 95\% prediction
-(model-based) intervals. With model-based inference, covariance
-parameters and the mean were estimated (using REML) separately for each
-trial. After all 2000 trials, we summarized the long-run performance of
-the sampling-inference combination in each scenario by calculating mean
-bias, root-mean-squared error, and interval coverage. Mean bias is taken
-as the average deviation between each trial's estimated (or predicted)
-mean (\(\hat{\mu}_i\)) and its realized mean (\(\mu_i\)):
+the realized mean and construct 95\% confidence (design-based) or 95\%
+prediction (model-based) intervals. With model-based inference,
+covariance parameters and the realized mean were estimated (using REML)
+separately for each trial. After all 2000 trials, we summarized the
+long-run performance of the sampling-inference combination in each
+scenario by calculating mean bias, root-mean-squared error, and interval
+coverage. Mean bias is taken as the average deviation between each
+trial's estimated (or predicted) mean (\(\hat{\mu}_i\)) and its realized
+mean (\(\mu_i\)):
 \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
@@
 We studied performance of the four sampling-inference combinations by
-selecting 2000 random IRS and GRTS samples of size \(n = 50\),
+selecting 2000 random SRS and GRTS samples of size \(n = 50\),
 \(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
@@
 total, there were six separate scenarios (two responses crossed with
-three sample sizes). For each sample from each scenario, parameters were
-estimated using the same approaches as described for simulated data. We
-also used the same evaluation metrics as for the simulated data: mean
-bias, RMSE, and interval coverage. Mean bias is taken as the average
-deviation between each sample's estimated (or predicted) mean
-(\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of ZMMI or Hg ppb):
-\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
-indexes simulation trials. Root-mean-squared error is taken as the
-square root of the average squared deviation between each sample's
-estimated (or predicted) mean and its population mean:
+three sample sizes). Within each SRS and GRTS sample, design-based and
+model-based inferences were used to estimate or predict the population
+mean and construct 95\% coverage intervals. With model-based inference,
+the exponential covariance was assumed, and covariance parameters and
+the population mean were estimated (using REML) (separately for each SRS
+and GRTS sample). We used the same evaluation metrics as for the
+simulated data: mean bias, RMSE, and interval coverage. Mean bias is
+taken as the average deviation between each sample's estimated (or
+predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
+ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
+where \(i\) indexes simulation trials. Root-mean-squared error is taken
+as the square root of the average squared deviation between each
+sample's estimated (or predicted) mean and its population mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
@@
 GRTS-MB and GRTS-DB tends to be quite small. When \(n = 50\), SRS-MB
-RMSE is approximately evenly between IRS-DB RMSE and GRTS-MB RMSE, but
+RMSE is approximately evenly between SRS-DB RMSE and GRTS-MB RMSE, but
 for the larger sample sizes (\(n = 100\), \(n = 200\)), SRS-MB RMSE is

### 2022-04-19 inst/manuscript/manuscript.tex
more minor updates

@@
 the exponential covariance was assumed, and covariance parameters and
-the population mean were estimated (using REML) (separately for each SRS
+the population mean were estimated using REML (separately for each SRS
 and GRTS sample). We used the same evaluation metrics as for the

### 2022-05-17 inst/manuscript/manuscript.Rmd
minor revision updates

@@
 A variety of sample sizes, location layouts, dependence structures, and response types are considered.
-The population mean is the parameter of interest and performance is measured using statistics like bias, squared error, and interval coverage.
+The population mean is the parameter of interest, and performance is measured using statistics like bias, squared error, and interval coverage.
   3.
@@
 FPBK assumes $\bm{\delta}$ in Equation$~$\eqref{equation:Zmarginal} has mean-zero and a spatial dependence structure that can be modeled using a covariance function.
-This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), isotropic (independent of direction), and decays with distance between population units [@cressie1993statistics].
+This covariance function is commonly assumed to be non-negative, second-order stationary (depending only on the separation vector (e.g., distance) between population units), and isotropic (independent of direction) [@cressie1993statistics].
 Henceforth, it is implied that we have made these same assumptions regarding $\bm{\delta}$.
@@
 
-The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on using restricted maximum likelihood (REML) [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
+The parameters in Equation$~$\ref{equation:Zmarginal} can be estimated using a variety of techniques, but we focus on restricted maximum likelihood (REML) [@patterson1971recovery; @harville1977maximum; @wolfinger1994computing].
 REML is preferred over maximum likelihood (ML) because ML estimates can be badly biased for small sample sizes, due to the fact that ML makes no adjustment for the simultaneous estimation of $\bm{\beta}$ and $\bm{\theta}$ [@patterson1971recovery].
@@
 Together, the two sampling plans (SRS and GRTS) combined with the two inference approaches (DB and MB) yielded four sampling-inference combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB.
-For SRS-DB, the Horvitz-Thompson estimator \eqref{eq:ht} was used to estimate means and the commonly-used SRS variance formula [@sarndal2003model; @lohr2009sampling] was used to estimate the variance.
+For SRS-DB, the Horvitz-Thompson estimator \eqref{eq:ht} was used to estimate means and the commonly-used SRS variance formula [@sarndal2003model; @lohr2009sampling] was used to estimate variances.
 This variance formula is given by
@@
 \end{equation}
-where $z_i$ is the $i$th response value, $\bar{z}$ is the mean of all $z_i$, $n$ is the sample size, $N$ is the population size, and $f = (1 - n / N)$ ($f$ is often called the finite population correction factor).
+where $z_i$ is the response value of the $i$th sampled population unit, $\bar{z}$ is the mean of all $z_i$, $n$ is the sample size, $N$ is the population size, and $f = (1 - n / N)$ ($f$ is often called the finite population correction factor).
 For GRTS-DB, the Horvitz-Thompson estimator was used to estimate means and the local neighborhood variance was used to estimate variances.
@@
 
-We used simulated data to compare the sampling-inference combinations across many realized populations from the same data-generating stochastic process.
-We then used real data from the 2012 National Lakes Assessment to compare the sampling-inference combinations within a single realized population (which is typically the case in reality).
+The simulated and real data were used in distinct ways to answer distinct questions.
+We used the simulated data to compare the sampling-inference combinations across many realized populations (from the same data-generating stochastic process) and random samples.
+We used the real data to compare the sampling-inference combinations within a single realized population but across random samples.
 With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
-With the real data, we were in control of only the random sampling process.
+With the real data, we were only in control of the random sampling process, which is typically the case in reality.
 
@@
   \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
-  \caption{Histogram of a realized population for the normal response.}
+  \caption{Histogram of a single realized population for the normal response.}
   \label{fig:symm_pop_hist}
@@
   \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
-  \caption{Histogram of a realized population for the skewed response.}
+  \caption{Histogram of a single realized population for the skewed response.}
   \label{fig:skew_pop_hist}
 \end{subfigure}
-\caption{Histograms of realized populations simulated for the normal and skewed resposnes using the random layout and 50\% DRE.}
+\caption{Histograms of single realized populations simulated for the normal and skewed respones using the random layout and 50\% DRE.}
 \label{fig:sim_pops}
@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-Within each simulation scenario and trial, SRS and GRTS samples were selected and then design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the realized mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
+First, a population was simulated according to the specifications of the particular simulation scenario (for the random location layout, locations were simulated separately for each trial).
+Next, SRS and GRTS samples were selected.
+Then, design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the realized mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
 With model-based inference, covariance parameters and the realized mean were estimated (using REML) separately for each trial.
 After all 2000 trials, we summarized the long-run performance of the sampling-inference combination in each scenario by calculating mean bias, root-mean-squared error, and interval coverage.
-Mean bias is taken as the average deviation between each trial's estimated (or predicted) mean ($\hat{\mu}_i$) and its realized mean ($\mu_i$): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes simulation trials.
-Root-mean-squared error is taken as the square root of the average squared deviation between each trial's estimated (or predicted) mean and its realized mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}$.
-Interval coverage is taken as the proportion of simulation trials where the realized mean is contained in its 95% confidence (or prediction) interval.
-These intervals are constructed using the normal distribution -- justification comes from the asymptotic normality of means via the central limit theorem (under some assumptions).
+Mean bias was taken as the average deviation between each trial's estimated (or predicted) mean ($\hat{\mu}_i$) and its realized mean ($\mu_i$): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes the simulation trials.
+Because each trial had different realized populations, $\mu_i$ changed with $i$.
+Root-mean-squared error was taken as the square root of the average squared deviation between each trial's estimated (or predicted) mean and its realized mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}$.
+Interval coverage was taken as the proportion of simulation trials where the realized mean was contained in its 95% confidence (or prediction) interval.
+These intervals were constructed using the normal distribution -- justification comes from the asymptotic normality of means via the central limit theorem (under some assumptions).
 Quantifying these metrics is important because together, they give us an idea of the accuracy (mean bias), spread (RMSE), and validity (interval coverage) of the sampling-inference combinations.
@@
 Also in Figure \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
-The spatial semivariogram quantifies the halved average squared differences (semivariance) of responses whose separation (distance) falls within some distance class.
+The spatial semivariogram quantifies the halved average squared differences (semivariance) of responses whose separation (distance) falls within a separation class.
 The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
@@
 
-We studied performance of the four sampling-inference combinations by selecting 2000 random SRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
+We studied performance of the four sampling-inference combinations by selecting 2000 SRS and GRTS samples of size $n = 50$, $n = 100$, and $n = 200$ from the realized ZMMI and Hg ppb populations and then analyzing the samples using MB and DB inference.
 In total, there were six separate scenarios (two responses crossed with three sample sizes).
@@
 We used the same evaluation metrics as for the simulated data: mean bias, RMSE, and interval coverage.
-Mean bias is taken as the average deviation between each sample's estimated (or predicted) mean ($\hat{\mu}_i$) and the population mean ($\mu$) (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes simulation trials.
-Root-mean-squared error is taken as the square root of the average squared deviation between each sample's estimated (or predicted) mean and its population mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}$.
-Interval coverage is taken as the proportion of simulation trials where the population mean is contained in its 95% confidence (or prediction) interval.
-These intervals are constructed using the normal distribution.
+Mean bias was taken as the average deviation between each sample's estimated (or predicted) mean ($\hat{\mu}_i$) and the population mean ($\mu$) (of ZMMI or Hg ppb): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)$, where $i$ indexes the simulation trials.
+Because each trial had the same realized population, $\mu$ did not change with $i$ (in contrast to the simulated data, where the realized mean changed with $i$).
+Root-mean-squared error was taken as the square root of the average squared deviation between each sample's estimated (or predicted) mean and its population mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}$.
+Interval coverage was taken as the proportion of simulation trials where the population mean was contained in its 95% confidence (or prediction) interval.
+These intervals were constructed using the normal distribution.
 
@@
 Recall that model-based inference defines interval coverage properties across realized populations.
-With the simulated data, we evaluated interval coverage across realized populations, but for the NLA data, we evaluated interval coverage within a single realized population for different samples.
+With the simulated data, we evaluated interval coverage across realized populations, but for the NLA data, we evaluated interval coverage within a single realized population.
 We did find that model-based coverages were similar to the design-based coverages, however, suggesting that for some realized populations it is reasonable to heuristically view data from separate random samples as being from approximately separate realized populations.
@@
 Model-based approaches rely on distributional assumptions to predict realized values of a data-generating stochastic process.
-Though model-based approaches do not rely on random sampling, it can still be beneficial as a way to guard against preferential sampling.
+Though model-based approaches do not rely on random sampling, random sampling can still be beneficial as a way to guard against preferential sampling.
 While design-based and model-based approaches have often been compared in the literature from theoretical and analytical perspectives, our contribution lies in studying them for finite population spatial data while implementing GRTS sampling and the local neighborhood variance estimator.
@@
 All sampling-inference combinations had approximately zero mean bias.
-Independent of the inference approach, GRTS-DB and GRTS-MB had lower RMSE than their SRS counterparts.
+Independent of the inference approach, the GRTS samples yielded lower RMSE than their SRS counterparts.
 Though GRTS-DB and GRTS-MB generally had very similar RMSE, SRS-MB tended to have much lower RMSE than SRS-DB, suggesting that the model-based inference mitigated much of the inefficiency in RMSE from SRS.
@@
 @brus1997random provide a more thorough discussion regarding the benefits and drawbacks of the two approaches.
-In short, when deciding whether the design-based or model-based approach is more appropriate to implement, the benefits and drawbacks of each approach should be considered alongside the particular goals of the study.
+In short, when deciding whether the design-based or model-based approach is more appropriate to implement, these benefits and drawbacks should be considered alongside the particular goals of the study.
 
@@
 
-We would like to thank the editors and anonymous reviewers for hard work and time spent providing us with thoughtful, valuable feedback which greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for their hard work and time spent providing us with thoughtful, valuable feedback which greatly improved the manuscript.
 

### 2022-05-18 inst/manuscript/manuscript.Rmd
minor formatting updates

@@
   2.
-We provide relevant background for the design-based and model-based approaches and then study their performance using simulated and real data from the United States Environmental Protection Agency's 2012 National Lakes Assessment.
+We provide relevant background for the design-based and model-based approaches and then study their performance using simulated data and real data.
+The real data is from the United States Environmental Protection Agency's 2012 National Lakes Assessment.
 A variety of sample sizes, location layouts, dependence structures, and response types are considered.
@@
 In each of the 36 simulation scenarios, there were 2000 independent simulation trials.
-First, a population was simulated according to the specifications of the particular simulation scenario (for the random location layout, locations were simulated separately for each trial).
+Within each trial, a population was simulated according to the specifications of the particular simulation scenario (for the random location layout, locations were simulated separately for each trial).
 Next, SRS and GRTS samples were selected.

### 2022-05-18 inst/manuscript/manuscript.tex
minor formatting updates

@@
     We provide relevant background for the design-based and model-based
-    approaches and then study their performance using simulated and real
-    data from the United States Environmental Protection Agency's 2012
-    National Lakes Assessment. A variety of sample sizes, location
-    layouts, dependence structures, and response types are considered. The
-    population mean is the parameter of interest and performance is
-    measured using statistics like bias, squared error, and interval
-    coverage.
+    approaches and then study their performance using simulated data and
+    real data. The real data is from the United States Environmental
+    Protection Agency's 2012 National Lakes Assessment. A variety of
+    sample sizes, location layouts, dependence structures, and response
+    types are considered. The population mean is the parameter of
+    interest, and performance is measured using statistics like bias,
+    squared error, and interval coverage.
   \item
@@
 stochastic process. We summarize some of the basic principles of FPBK
-next -- see Ver Hoef (2008) for technical details and see Higham, J. Ver
-Hoef, et al. (2021) for an extension to cases of imperfect detection
-among population units. Let
+next -- see Ver Hoef (2008) for technical details and see Higham, Ver
+Hoef, Madsen, et al. (2021) for an extension to cases of imperfect
+detection among population units. Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 assumed to be non-negative, second-order stationary (depending only on
-the separation vector (e.g., distance) between population units),
-isotropic (independent of direction), and decays with distance between
-population units (Cressie, 1993). Henceforth, it is implied that we have
-made these same assumptions regarding \(\bm{\delta}\). Chiles and
-Delfiner (1999), pp.~80-93 discuss covariance functions that are not
-second-order stationary, not isotropic, or not either. A variety of
-flexible covariance functions can be used to model \(\bm{\delta}\)
-(Cressie, 1993) -- one example is the exponential covariance function.
-Cressie (1993) provides a thorough list of spatial covariance functions.
-The \(i,j\)th element of the exponential covariance matrix,
-\(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
-\begin{align}\label{equation:expcov}
+the separation vector (e.g., distance) between population units), and
+isotropic (independent of direction) (Cressie, 1993). Henceforth, it is
+implied that we have made these same assumptions regarding
+\(\bm{\delta}\). Chiles and Delfiner (1999), pp.~80-93 discuss
+covariance functions that are not second-order stationary, not
+isotropic, or not either. A variety of flexible covariance functions can
+be used to model \(\bm{\delta}\) (Cressie, 1993) -- one example is the
+exponential covariance function. Cressie (1993) provides a thorough list
+of spatial covariance functions. The \(i,j\)th element of the
+exponential covariance matrix, \(\mathop{\mathrm{{cov}}}(\bm{\delta})\),
+is \mbox{} \begin{align}\label{equation:expcov}
 \mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = 
@@
 The parameters in Equation\(~\)\ref{equation:Zmarginal} can be estimated
-using a variety of techniques, but we focus on using restricted maximum
+using a variety of techniques, but we focus on restricted maximum
 likelihood (REML) (Harville, 1977; Patterson and Thompson, 1971;
@@
 estimate means and the commonly-used SRS variance formula (Lohr, 2009;
-Särndal et al., 2003) was used to estimate the variance. This variance
+Särndal et al., 2003) was used to estimate variances. This variance
 formula is given by \begin{equation}\label{equation:srs_var}
  \frac{f[\sum_{i = 1}^n (z_i - \bar{z})^2]}{n(n - 1)},
-\end{equation} where \(z_i\) is the \(i\)th response value, \(\bar{z}\)
-is the mean of all \(z_i\), \(n\) is the sample size, \(N\) is the
-population size, and \(f = (1 - n / N)\) (\(f\) is often called the
-finite population correction factor). For GRTS-DB, the Horvitz-Thompson
-estimator was used to estimate means and the local neighborhood variance
-was used to estimate variances. For SRS-MB and GRTS-MB, FPBK was used to
-estimate means and variances using restricted maximum likelihood. SRS,
-GRTS sampling, and design-based inference were implemented using the
+\end{equation} where \(z_i\) is the response value of the \(i\)th
+sampled population unit, \(\bar{z}\) is the mean of all \(z_i\), \(n\)
+is the sample size, \(N\) is the population size, and
+\(f = (1 - n / N)\) (\(f\) is often called the finite population
+correction factor). For GRTS-DB, the Horvitz-Thompson estimator was used
+to estimate means and the local neighborhood variance was used to
+estimate variances. For SRS-MB and GRTS-MB, FPBK was used to estimate
+means and variances using restricted maximum likelihood. SRS, GRTS
+sampling, and design-based inference were implemented using the
 \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle et al., 2022).
 FPBK was implemented using the \texttt{sptotal} \textbf{\textsf{R}}
-package (Higham, Ver Hoef, et al., 2021).
-
-We used simulated data to compare the sampling-inference combinations
-across many realized populations from the same data-generating
-stochastic process. We then used real data from the 2012 National Lakes
-Assessment to compare the sampling-inference combinations within a
-single realized population (which is typically the case in reality).
-With the simulated data, we were in control of the data-generating
-stochastic process and the random sampling process. With the real data,
-we were in control of only the random sampling process.
+package (Higham, Ver Hoef, Frank, et al., 2021).
+
+The simulated and real data were used in distinct ways to answer
+distinct questions. We used the simulated data to compare the
+sampling-inference combinations across many realized populations (from
+the same data-generating stochastic process) and random samples. We used
+the real data to compare the sampling-inference combinations within a
+single realized population but across random samples. With the simulated
+data, we were in control of the data-generating stochastic process and
+the random sampling process. With the real data, we were only in control
+of the random sampling process, which is typically the case in reality.
 
@@
 In each of the 36 simulation scenarios, there were 2000 independent
-simulation trials. Within each simulation scenario and trial, SRS and
-GRTS samples were selected and then design-based and model-based
-inferences were used to estimate (design-based) or predict (model-based)
-the realized mean and construct 95\% confidence (design-based) or 95\%
-prediction (model-based) intervals. With model-based inference,
-covariance parameters and the realized mean were estimated (using REML)
-separately for each trial. After all 2000 trials, we summarized the
-long-run performance of the sampling-inference combination in each
-scenario by calculating mean bias, root-mean-squared error, and interval
-coverage. Mean bias is taken as the average deviation between each
-trial's estimated (or predicted) mean (\(\hat{\mu}_i\)) and its realized
-mean (\(\mu_i\)):
+simulation trials. Within each trial, a population was simulated
+according to the specifications of the particular simulation scenario
+(for the random location layout, locations were simulated separately for
+each trial). Next, SRS and GRTS samples were selected. Then,
+design-based and model-based inferences were used to estimate
+(design-based) or predict (model-based) the realized mean and construct
+95\% confidence (design-based) or 95\% prediction (model-based)
+intervals. With model-based inference, covariance parameters and the
+realized mean were estimated (using REML) separately for each trial.
+After all 2000 trials, we summarized the long-run performance of the
+sampling-inference combination in each scenario by calculating mean
+bias, root-mean-squared error, and interval coverage. Mean bias was
+taken as the average deviation between each trial's estimated (or
+predicted) mean (\(\hat{\mu}_i\)) and its realized mean (\(\mu_i\)):
 \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
-indexes simulation trials. Root-mean-squared error is taken as the
-square root of the average squared deviation between each trial's
-estimated (or predicted) mean and its realized mean:
+indexes the simulation trials. Because each trial had different realized
+populations, \(\mu_i\) changed with \(i\). Root-mean-squared error was
+taken as the square root of the average squared deviation between each
+trial's estimated (or predicted) mean and its realized mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}\).
-Interval coverage is taken as the proportion of simulation trials where
-the realized mean is contained in its 95\% confidence (or prediction)
-interval. These intervals are constructed using the normal distribution
+Interval coverage was taken as the proportion of simulation trials where
+the realized mean was contained in its 95\% confidence (or prediction)
+interval. These intervals were constructed using the normal distribution
 -- justification comes from the asymptotic normality of means via the
@@
 quantifies the halved average squared differences (semivariance) of
-responses whose separation (distance) falls within some distance class.
+responses whose separation (distance) falls within a separation class.
 The spatial semivariance is closely related to the spatial covariance,
@@
 We studied performance of the four sampling-inference combinations by
-selecting 2000 random SRS and GRTS samples of size \(n = 50\),
-\(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
-populations and then analyzing the samples using MB and DB inference. In
-total, there were six separate scenarios (two responses crossed with
-three sample sizes). Within each SRS and GRTS sample, design-based and
-model-based inferences were used to estimate or predict the population
-mean and construct 95\% coverage intervals. With model-based inference,
-the exponential covariance was assumed, and covariance parameters and
-the population mean were estimated using REML (separately for each SRS
-and GRTS sample). We used the same evaluation metrics as for the
-simulated data: mean bias, RMSE, and interval coverage. Mean bias is
-taken as the average deviation between each sample's estimated (or
-predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
-ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
-where \(i\) indexes simulation trials. Root-mean-squared error is taken
-as the square root of the average squared deviation between each
-sample's estimated (or predicted) mean and its population mean:
+selecting 2000 SRS and GRTS samples of size \(n = 50\), \(n = 100\), and
+\(n = 200\) from the realized ZMMI and Hg ppb populations and then
+analyzing the samples using MB and DB inference. In total, there were
+six separate scenarios (two responses crossed with three sample sizes).
+Within each SRS and GRTS sample, design-based and model-based inferences
+were used to estimate or predict the population mean and construct 95\%
+coverage intervals. With model-based inference, the exponential
+covariance was assumed, and covariance parameters and the population
+mean were estimated using REML (separately for each SRS and GRTS
+sample). We used the same evaluation metrics as for the simulated data:
+mean bias, RMSE, and interval coverage. Mean bias was taken as the
+average deviation between each sample's estimated (or predicted) mean
+(\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of ZMMI or Hg ppb):
+\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
+indexes the simulation trials. Because each trial had the same realized
+population, \(\mu\) did not change with \(i\) (in contrast to the
+simulated data, where the realized mean changed with \(i\)).
+Root-mean-squared error was taken as the square root of the average
+squared deviation between each sample's estimated (or predicted) mean
+and its population mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
-Interval coverage is taken as the proportion of simulation trials where
-the population mean is contained in its 95\% confidence (or prediction)
-interval. These intervals are constructed using the normal distribution.
+Interval coverage was taken as the proportion of simulation trials where
+the population mean was contained in its 95\% confidence (or prediction)
+interval. These intervals were constructed using the normal
+distribution.
 
@@
 populations, but for the NLA data, we evaluated interval coverage within
-a single realized population for different samples. We did find that
-model-based coverages were similar to the design-based coverages,
-however, suggesting that for some realized populations it is reasonable
-to heuristically view data from separate random samples as being from
-approximately separate realized populations. But generally, if
-model-based intervals constructed from many random samples of a single
-realized population show improper coverage, this does not necessarily
-imply a deficiency in model-based inference. Tables for interval
-coverage in all six simulation scenarios are provided in the supporting
-information.
+a single realized population. We did find that model-based coverages
+were similar to the design-based coverages, however, suggesting that for
+some realized populations it is reasonable to heuristically view data
+from separate random samples as being from approximately separate
+realized populations. But generally, if model-based intervals
+constructed from many random samples of a single realized population
+show improper coverage, this does not necessarily imply a deficiency in
+model-based inference. Tables for interval coverage in all six
+simulation scenarios are provided in the supporting information.
 
@@
 realized values of a data-generating stochastic process. Though
-model-based approaches do not rely on random sampling, it can still be
-beneficial as a way to guard against preferential sampling. While
-design-based and model-based approaches have often been compared in the
-literature from theoretical and analytical perspectives, our
+model-based approaches do not rely on random sampling, random sampling
+can still be beneficial as a way to guard against preferential sampling.
+While design-based and model-based approaches have often been compared
+in the literature from theoretical and analytical perspectives, our
 contribution lies in studying them for finite population spatial data
@@
 combinations had approximately zero mean bias. Independent of the
-inference approach, GRTS-DB and GRTS-MB had lower RMSE than their SRS
+inference approach, the GRTS samples yielded lower RMSE than their SRS
 counterparts. Though GRTS-DB and GRTS-MB generally had very similar
@@
 In short, when deciding whether the design-based or model-based approach
-is more appropriate to implement, the benefits and drawbacks of each
-approach should be considered alongside the particular goals of the
-study.
+is more appropriate to implement, these benefits and drawbacks should be
+considered alongside the particular goals of the study.
 
@@
 
-We would like to thank the editors and anonymous reviewers for hard work
-and time spent providing us with thoughtful, valuable feedback which
-greatly improved the manuscript.
+We would like to thank the editors and anonymous reviewers for their
+hard work and time spent providing us with thoughtful, valuable feedback
+which greatly improved the manuscript.
 

### 2022-05-18 inst/manuscript/manuscript.Rmd
minor review updates

@@
   - code: STLAW
-    address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
+    address: St. Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
   - code: NOAA
@@
 
-The simulated and real data were used in distinct ways to answer distinct questions.
-We used the simulated data to compare the sampling-inference combinations across many realized populations (from the same data-generating stochastic process) and random samples.
-We used the real data to compare the sampling-inference combinations within a single realized population but across random samples.
+The simulated and real data were used for distinct objectives.
+The simulated data was used to compare the sampling-inference combinations across many realized populations (from the same data-generating stochastic process) and random samples.
+The real data was used to compare the sampling-inference combinations within a single realized population but across random samples.
 With the simulated data, we were in control of the data-generating stochastic process and the random sampling process.
-With the real data, we were only in control of the random sampling process, which is typically the case in reality.
+With the real data, we were only in control of the random sampling process (which is typically the case in practice).
 

### 2022-05-18 inst/manuscript/manuscript.tex
minor review updates

@@
 Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
-and Statistics, 23 Romoda Drive, Canton, New York, 13617}
+    \address[STLAW]{St.~Lawrence University Department of Mathematics, Computer Science, and
+Statistics, 23 Romoda Drive, Canton, New York, 13617}
     \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
@@
 
-The simulated and real data were used in distinct ways to answer
-distinct questions. We used the simulated data to compare the
-sampling-inference combinations across many realized populations (from
-the same data-generating stochastic process) and random samples. We used
-the real data to compare the sampling-inference combinations within a
-single realized population but across random samples. With the simulated
-data, we were in control of the data-generating stochastic process and
-the random sampling process. With the real data, we were only in control
-of the random sampling process, which is typically the case in reality.
+The simulated and real data were used for distinct objectives. The
+simulated data was used to compare the sampling-inference combinations
+across many realized populations (from the same data-generating
+stochastic process) and random samples. The real data was used to
+compare the sampling-inference combinations within a single realized
+population but across random samples. With the simulated data, we were
+in control of the data-generating stochastic process and the random
+sampling process. With the real data, we were only in control of the
+random sampling process (which is typically the case in practice).
 

### 2022-05-18 inst/manuscript/supporting.Rmd
minor review updates

@@
   - code: STLAW
-    address: Saint Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
+    address: St. Lawrence University Department of Mathematics, Computer Science, and Statistics, 23 Romoda Drive, Canton, New York, 13617
   - code: OSU

### 2022-05-19 inst/manuscript/manuscript.Rmd
minor updates

@@
 The real data is from the United States Environmental Protection Agency's 2012 National Lakes Assessment (NLA) [@USEPA2012NLA].
-In Section \ref{sec:results}, we present analysis results for the simulated data and NLA data.
+In Section \ref{sec:results}, we present analysis results for the simulated data and real data.
 And in Section \ref{sec:discussion}, we end with a discussion and provide directions for future research.
@@
 We previously mentioned that the design-based approach can be used to select spatially balanced samples.
-Spatially balanced samples are useful because parameter estimates from these samples tend to vary less than parameter estimates from samples lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
+Spatially balanced samples are useful because parameter estimates from these samples tend to vary less (be more precise) than parameter estimates from samples lacking spatial balance [@stevens2004spatially; @barabesi2011sampling; @grafstrom2013well; @robertson2013bas; @wang2013design; @benedetti2017spatiallyreview].
 The first spatially balanced sampling algorithm to see widespread use was the generalized random tessellation stratified (GRTS) algorithm [@stevens2004spatially].
@@
 Within each trial, a population was simulated according to the specifications of the particular simulation scenario (for the random location layout, locations were simulated separately for each trial).
-Next, SRS and GRTS samples were selected.
+Next, a random SRS sample and a random GRTS sample were selected.
 Then, design-based and model-based inferences were used to estimate (design-based) or predict (model-based) the realized mean and construct 95% confidence (design-based) or 95% prediction (model-based) intervals.
@@
 Mean bias was taken as the average deviation between each trial's estimated (or predicted) mean ($\hat{\mu}_i$) and its realized mean ($\mu_i$): $\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)$, where $i$ indexes the simulation trials.
-Because each trial had different realized populations, $\mu_i$ changed with $i$.
+Because each trial had a different realized population, $\mu_i$ changed with $i$.
 Root-mean-squared error was taken as the square root of the average squared deviation between each trial's estimated (or predicted) mean and its realized mean: $\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}$.

### 2022-05-19 inst/manuscript/manuscript.tex
minor updates

@@
 \ref{sec:results}, we present analysis results for the simulated data
-and NLA data. And in Section \ref{sec:discussion}, we end with a
+and real data. And in Section \ref{sec:discussion}, we end with a
 discussion and provide directions for future research.
@@
 select spatially balanced samples. Spatially balanced samples are useful
-because parameter estimates from these samples tend to vary less than
-parameter estimates from samples lacking spatial balance (Barabesi and
-Franceschi, 2011; Benedetti et al., 2017; Grafström and Lundström, 2013;
-Robertson et al., 2013; Stevens and Olsen, 2004; Wang et al., 2013). The
-first spatially balanced sampling algorithm to see widespread use was
-the generalized random tessellation stratified (GRTS) algorithm (Stevens
-and Olsen, 2004). After the GRTS algorithm was developed, several other
-spatially balanced sampling algorithms emerged, including stratified
-sampling with compact geographical strata (Walvoort et al., 2010), the
-local pivotal method (Grafström et al., 2012; Grafström and Matei,
-2018), spatially correlated Poisson sampling (Grafström, 2012), balanced
-acceptance sampling (Robertson et al., 2013), within-sample-distance
-sampling (Benedetti and Piersimoni, 2017), and Halton iterative
-partitioning sampling (Robertson et al., 2018). In this manuscript, we
-select spatially balanced samples using the GRTS algorithm because it is
+because parameter estimates from these samples tend to vary less (be
+more precise) than parameter estimates from samples lacking spatial
+balance (Barabesi and Franceschi, 2011; Benedetti et al., 2017;
+Grafström and Lundström, 2013; Robertson et al., 2013; Stevens and
+Olsen, 2004; Wang et al., 2013). The first spatially balanced sampling
+algorithm to see widespread use was the generalized random tessellation
+stratified (GRTS) algorithm (Stevens and Olsen, 2004). After the GRTS
+algorithm was developed, several other spatially balanced sampling
+algorithms emerged, including stratified sampling with compact
+geographical strata (Walvoort et al., 2010), the local pivotal method
+(Grafström et al., 2012; Grafström and Matei, 2018), spatially
+correlated Poisson sampling (Grafström, 2012), balanced acceptance
+sampling (Robertson et al., 2013), within-sample-distance sampling
+(Benedetti and Piersimoni, 2017), and Halton iterative partitioning
+sampling (Robertson et al., 2018). In this manuscript, we select
+spatially balanced samples using the GRTS algorithm because it is
 readily available in the \texttt{spsurvey} \textbf{\textsf{R}} package
@@
 (for the random location layout, locations were simulated separately for
-each trial). Next, SRS and GRTS samples were selected. Then,
-design-based and model-based inferences were used to estimate
-(design-based) or predict (model-based) the realized mean and construct
-95\% confidence (design-based) or 95\% prediction (model-based)
-intervals. With model-based inference, covariance parameters and the
-realized mean were estimated (using REML) separately for each trial.
-After all 2000 trials, we summarized the long-run performance of the
-sampling-inference combination in each scenario by calculating mean
-bias, root-mean-squared error, and interval coverage. Mean bias was
-taken as the average deviation between each trial's estimated (or
-predicted) mean (\(\hat{\mu}_i\)) and its realized mean (\(\mu_i\)):
-\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
-indexes the simulation trials. Because each trial had different realized
-populations, \(\mu_i\) changed with \(i\). Root-mean-squared error was
-taken as the square root of the average squared deviation between each
-trial's estimated (or predicted) mean and its realized mean:
+each trial). Next, a random SRS sample and a random GRTS sample were
+selected. Then, design-based and model-based inferences were used to
+estimate (design-based) or predict (model-based) the realized mean and
+construct 95\% confidence (design-based) or 95\% prediction
+(model-based) intervals. With model-based inference, covariance
+parameters and the realized mean were estimated (using REML) separately
+for each trial. After all 2000 trials, we summarized the long-run
+performance of the sampling-inference combination in each scenario by
+calculating mean bias, root-mean-squared error, and interval coverage.
+Mean bias was taken as the average deviation between each trial's
+estimated (or predicted) mean (\(\hat{\mu}_i\)) and its realized mean
+(\(\mu_i\)): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\),
+where \(i\) indexes the simulation trials. Because each trial had a
+different realized population, \(\mu_i\) changed with \(i\).
+Root-mean-squared error was taken as the square root of the average
+squared deviation between each trial's estimated (or predicted) mean and
+its realized mean:
 \(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}\).

### 2022-05-19 inst/manuscript/manuscript.Rmd
minor updates

@@
 The set of level-one cells is denoted by $\mathcal{A}_1$ and defined as $\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}$.
-Within each level-one cell, the inclusion probability for each population unit (which is pre-specified) is summed, and if any of these sums exceed one, a second level of cells is added.
+Within each level-one cell, the inclusion probability for each population unit (which is pre-specified) is summed, and if any of these sums are one or greater, a second level of cells is added.
 Then each level-one cell is split into four distinct, equally sized squares called level-two cells.
@@
 The set of level-two cells is denoted by $\mathcal{A}_2$ and defined as $\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}$.
-The inclusion probabilities within each level-two cell are summed, and if any of these sums exceed one, a third level of cells is added.
-This process continues for $k$ steps, until all level-$k$ cells have inclusion probability sums no larger than one.
+The inclusion probabilities within each level-two cell are summed, and if any of these sums are one or greater, a third level of cells is added.
+This process continues for $k$ steps, until all level-$k$ cells have inclusion probability sums less than one.
 Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$.
@@
 To address these two drawbacks simultaneously, @stevens2003variance proposed the local neighborhood variance estimator.
-The local neighborhood variance estimator does not rely on $\pi_{ij}$ and estimates the variance of $\hat{\tau}$ conditional on the random properties of the GRTS sample -- the idea being that this conditioning should yield a more precise estimate of $\hat{\tau}$.
+The local neighborhood variance estimator does not rely on $\pi_{ij}$ and estimates the variance of $\hat{\tau}$ conditional on the random properties of the GRTS sample -- the idea being that this conditioning should yield a more precise estimate of $\tau$.
 They show that the contribution from each sampled population unit to the overall variance is dominated by local variation.
@@
 The response was then exponentiated, yielding a skewed random variable whose total variance was 2.
-The skewed responses were used to evaluate performance of the sampling-inference approaches for data that were not normal but were still estimated using REML, which relies on a normal log-likelihood.
-Figure \ref{fig:sim_pops} shows an example of a realized population for the normal and skewed responses using the random layout and 50% DRE.
+The skewed responses were used to evaluate performance of the sampling-inference approaches for data that were not normally distributed but were still estimated using REML, which relies on a normal log-likelihood.
+Figure \ref{fig:sim_pops} shows an example of a realized population for the normal and skewed responses using the random location layout and 50% DRE.
 
@@
 The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
-Both ZMMI and Hg ppb seem to have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as the semivariance increases steadily with distance (meaning that observations nearby one another tend to be more similar than observations far apart from one another).
+Both ZMMI and Hg ppb seem to have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as the empirical semivariance increases steadily with distance (meaning that observations near one another tend to be more similar than observations far apart from one another).
 
@@
   \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
-  \caption{Semivariogram of the ZMMI population.}
+  \caption{Empirical semivariogram of the ZMMI population.}
   \label{fig:zmmi_sv_plot}
@@
   \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
-  \caption{Semivariogram of the Hg ppb population.}
+  \caption{Empirical semivariogram of the Hg ppb population.}
   \label{fig:mercury_sv_plot}
@@
   \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
-  \caption{NLA data relative RMSE for the four sampling-inference combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
-The solid, black lines separate the sample sizes.}
+  \caption{NLA data relative RMSE for the four sampling-inference combinations. The columns indicate the response type. The solid, black lines separate the sample sizes.}
   \label{fig:data_rmspe_eff}
@@
   \caption{NLA data interval coverage for the four sampling-inference combinations.
-The rows indicate the proportion of dependent error and the columns indicate the response type.
+The columns indicate the response type.
 The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
@@
 There are several benefits and drawbacks of the design-based and model-based approaches for finite population spatial sampling and inference.
-Some we have discussed, but others we have not, and they are worthy of consideration in future research.
+Some we have discussed, but others we have not, and they are worthy of discussion.
 First, we discuss advantages of the design-based approach.

### 2022-05-19 inst/manuscript/manuscript.tex
minor updates

@@
 each level-one cell, the inclusion probability for each population unit
-(which is pre-specified) is summed, and if any of these sums exceed one,
-a second level of cells is added. Then each level-one cell is split into
-four distinct, equally sized squares called level-two cells. Each
-level-two cell is randomly assigned a level-two address of 0, 1, 2, or
-3. The set of level-two cells is denoted by \(\mathcal{A}_2\) and
+(which is pre-specified) is summed, and if any of these sums are one or
+greater, a second level of cells is added. Then each level-one cell is
+split into four distinct, equally sized squares called level-two cells.
+Each level-two cell is randomly assigned a level-two address of 0, 1, 2,
+or 3. The set of level-two cells is denoted by \(\mathcal{A}_2\) and
 defined as
@@
 The inclusion probabilities within each level-two cell are summed, and
-if any of these sums exceed one, a third level of cells is added. This
-process continues for \(k\) steps, until all level-\(k\) cells have
-inclusion probability sums no larger than one. Then
+if any of these sums are one or greater, a third level of cells is
+added. This process continues for \(k\) steps, until all level-\(k\)
+cells have inclusion probability sums less than one. Then
 \(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).
@@
 GRTS sample -- the idea being that this conditioning should yield a more
-precise estimate of \(\hat{\tau}\). They show that the contribution from
-each sampled population unit to the overall variance is dominated by
-local variation. Thus the local neighborhood variance estimator is a
-weighted sum of variance estimates from each sampled population unit's
-local neighborhood. These local neighborhoods contain the sampled
-population unit itself and its three nearest neighbors (among all other
-sampled population units). For more details, see Stevens and Olsen
-(2003).
+precise estimate of \(\tau\). They show that the contribution from each
+sampled population unit to the overall variance is dominated by local
+variation. Thus the local neighborhood variance estimator is a weighted
+sum of variance estimates from each sampled population unit's local
+neighborhood. These local neighborhoods contain the sampled population
+unit itself and its three nearest neighbors (among all other sampled
+population units). For more details, see Stevens and Olsen (2003).
 
@@
 were used to evaluate performance of the sampling-inference approaches
-for data that were not normal but were still estimated using REML, which
-relies on a normal log-likelihood. Figure \ref{fig:sim_pops} shows an
-example of a realized population for the normal and skewed responses
-using the random layout and 50\% DRE.
+for data that were not normally distributed but were still estimated
+using REML, which relies on a normal log-likelihood. Figure
+\ref{fig:sim_pops} shows an example of a realized population for the
+normal and skewed responses using the random location layout and 50\%
+DRE.
 
@@
 spatial dependence in data. Both ZMMI and Hg ppb seem to have moderately
-strong spatial dependence (Figure \ref{fig:zmmi}), as the semivariance
-increases steadily with distance (meaning that observations nearby one
-another tend to be more similar than observations far apart from one
-another).
+strong spatial dependence (Figure \ref{fig:zmmi}), as the empirical
+semivariance increases steadily with distance (meaning that observations
+near one another tend to be more similar than observations far apart
+from one another).
 
@@
 inference. Some we have discussed, but others we have not, and they are
-worthy of consideration in future research. First, we discuss advantages
-of the design-based approach. Design-based inference is often
-computationally efficient, while model-based inference can be
-computationally burdensome, especially for likelihood-based estimation
-methods like REML that rely on the inverse of a covariance matrix.
-Design-based inference easily handles binary data through a
-straightforward application of the Horvitz-Thompson estimator. In
-contrast, analyzing binary data using model-based inference generally
-requires a logistic mixed regression model, the parameters of which can
-be difficult to estimate and interpret (Bolker et al., 2009). An
-advantage of design-based inference is that interval coverage is valid
-(has the proper coverage rate) as long as 1) the sample is sufficiently
-large to ensure the statistic's sampling distribution is approximately
-normal and 2) the variance estimator is consistent (Brus and De
-Gruijter, 1997; Särndal et al., 2003). This is because with the
-design-based approach, the sampling plan and inclusion probabilities are
-specified directly by the researcher. An advantage of SRS-DB not
-previously mentioned is that it is likely to be valid given the
-consistency of its variance estimator (Särndal et al., 2003). With the
-model-based approach, however, interval coverage is unlikely to be valid
-if the model assumptions made do not not accurately reflect reality.
-Whether model assumptions accurately reflect reality can be a
-challenging and sometimes impossible question to answer definitively.
+worthy of discussion. First, we discuss advantages of the design-based
+approach. Design-based inference is often computationally efficient,
+while model-based inference can be computationally burdensome,
+especially for likelihood-based estimation methods like REML that rely
+on the inverse of a covariance matrix. Design-based inference easily
+handles binary data through a straightforward application of the
+Horvitz-Thompson estimator. In contrast, analyzing binary data using
+model-based inference generally requires a logistic mixed regression
+model, the parameters of which can be difficult to estimate and
+interpret (Bolker et al., 2009). An advantage of design-based inference
+is that interval coverage is valid (has the proper coverage rate) as
+long as 1) the sample is sufficiently large to ensure the statistic's
+sampling distribution is approximately normal and 2) the variance
+estimator is consistent (Brus and De Gruijter, 1997; Särndal et al.,
+2003). This is because with the design-based approach, the sampling plan
+and inclusion probabilities are specified directly by the researcher. An
+advantage of SRS-DB not previously mentioned is that it is likely to be
+valid given the consistency of its variance estimator (Särndal et al.,
+2003). With the model-based approach, however, interval coverage is
+unlikely to be valid if the model assumptions made do not not accurately
+reflect reality. Whether model assumptions accurately reflect reality
+can be a challenging and sometimes impossible question to answer
+definitively.
 

### 2022-05-19 inst/manuscript/manuscript.tex
knitted version with correct citations.

@@
   \linenumbers % turns line numbering on
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
 \usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Pandoc header
 
@@
 \begin{document}
-
 \begin{frontmatter}
 
-  \title{A comparison of design-based and model-based approaches for finite
-population spatial sampling and inference.}
+  \title{A comparison of design-based and model-based approaches for
+finite population spatial sampling and inference.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{St.~Lawrence University Department of Mathematics, Computer Science, and
-Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{St.~Lawrence University Department of Mathematics,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+      \cortext[1]{Corresponding Author: Michael Dumelle
+(Dumelle.Michael@epa.gov)}
 
@@
   \item
-    The design-based and model-based approaches to frequentist statistical
-    inference rest on fundamentally different foundations. In the
-    design-based approach, inference relies on random sampling. In the
-    model-based approach, inference relies on distributional assumptions.
-    We compare the approaches in a finite population spatial context.
+    The design-based and model-based approaches to frequentist
+    statistical inference rest on fundamentally different foundations.
+    In the design-based approach, inference relies on random sampling.
+    In the model-based approach, inference relies on distributional
+    assumptions. We compare the approaches in a finite population
+    spatial context.
   \item
@@
   \item
-    When studying the simulated and real data, we found that regardless of
-    the strength of spatial dependence in the data, the generalized random
-    tessellation stratified (GRTS) algorithm, which explicitly
-    incorporates spatial locations into sampling, tends to outperform the
-    simple random sampling (SRS) algorithm, which does not explicitly
-    incorporate spatial locations into sampling. We also found that
-    model-based inference tends to outperform design-based inference, even
-    for skewed data where the model-based distributional assumptions are
-    violated. The performance gap between design-based inference and
-    model-based inference is small when GRTS samples are used but large
-    when SRS samples are used, suggesting that the sampling choice
-    (whether to use GRTS or SRS) is most important when performing
-    design-based inference.
+    When studying the simulated and real data, we found that regardless
+    of the strength of spatial dependence in the data, the generalized
+    random tessellation stratified (GRTS) algorithm, which explicitly
+    incorporates spatial locations into sampling, tends to outperform
+    the simple random sampling (SRS) algorithm, which does not
+    explicitly incorporate spatial locations into sampling. We also
+    found that model-based inference tends to outperform design-based
+    inference, even for skewed data where the model-based distributional
+    assumptions are violated. The performance gap between design-based
+    inference and model-based inference is small when GRTS samples are
+    used but large when SRS samples are used, suggesting that the
+    sampling choice (whether to use GRTS or SRS) is most important when
+    performing design-based inference.
   \item
@@
     model-based approaches for finite population spatial sampling and
-    inference that practitioners must consider when choosing between them.
-    We provide relevant background contextualizing each approach and study
-    their properties in a variety of scenarios, making recommendations for
-    use based on the practitioner's goals.
+    inference that practitioners must consider when choosing between
+    them. We provide relevant background contextualizing each approach
+    and study their properties in a variety of scenarios, making
+    recommendations for use based on the practitioner's goals.
   \end{enumerate}
@@
 design-based and model-based approaches for spatial data (Brus and De
-Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006) reviews
-the two approaches in an ecological context before introducing a
-``model-assisted'' variance estimator that combines aspects from each
-approach. In addition to Cooper (2006), there has been substantial
-research and development into estimators that use both design-based and
-model-based principles (see e.g., Sterba (2009) and Cicchitelli and
-Montanari (2012), and for Bayesian approaches, see Chan-Golston et al.
-(2020) and Hofman and Brus (2021)).
+Gruijter, 1997; Brus, 2021; Ver Hoef, 2002; Ver Hoef, 2008). Cooper
+(2006) reviews the two approaches in an ecological context before
+introducing a ``model-assisted'' variance estimator that combines
+aspects from each approach. In addition to Cooper (2006), there has been
+substantial research and development into estimators that use both
+design-based and model-based principles (see e.g., Sterba (2009) and
+Cicchitelli and Montanari (2012), and for Bayesian approaches, see
+Chan-Golston et al. (2020) and Hofman and Brus (2021)).
 
@@
 sample spatially balanced. By ``reflecting the layout of the population
-units'', we mean that if population units are concentrated in specific
+units,'' we mean that if population units are concentrated in specific
 areas, the units in the sample should be concentrated in the same areas.
@@
 stochastic process. We summarize some of the basic principles of FPBK
-next -- see Ver Hoef (2008) for technical details and see Higham, Ver
-Hoef, Madsen, et al. (2021) for an extension to cases of imperfect
-detection among population units. Let
+next -- see Ver Hoef (2008) for technical details and see Higham et al.
+(2021b) for an extension to cases of imperfect detection among
+population units. Let
 \({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
@@
 FPBK was implemented using the \texttt{sptotal} \textbf{\textsf{R}}
-package (Higham, Ver Hoef, Frank, et al., 2021).
+package (Higham et al., 2021a).
 
@@
 \leavevmode\hypertarget{ref-higham2021sptotal}{}%
-Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
+Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021a. Sptotal:
 Predicting totals and weighted sums from spatial data.
@@
 \leavevmode\hypertarget{ref-higham2021adjusting}{}%
-Higham, M., Ver Hoef, J., Madsen, L., Aderman, A., 2021. Adjusting a
+Higham, M., Ver Hoef, J., Madsen, L., Aderman, A., 2021b. Adjusting a
 finite population block kriging estimator for imperfect detection.

### 2022-06-09 inst/scripts/figures/data.R
minor updates

@@ -60,31 +60,6 @@ combo_data <- combo_data %>%
   ) %>% factor(levels = c("SRS-DB", "SRS-MB", "GRTS-DB", "GRTS-MB")))
 
 
-# combo_data <- combo_data %>%
-#   pivot_longer(cols = c(rel_efficiency, coverage), names_to = "stat", values_to = "value")
-# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-# resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
-# stat_labs <- c(rel_efficiency = "Relative rMS(P)E", coverage = "Interval Coverage")
-# rmspe_eff <- combo_data %>%
-#   ggplot(aes(x = n_factor_adj, y = value, colour = approach)) +
-#   facet_grid(
-#     stat ~ var,
-#     labeller = labeller(var = resptype_labs, stat = stat_labs)
-#   ) +
-#   scale_colour_manual(values = colour_scale) +
-#   geom_point(size = 2) +
-#   geom_vline(xintercept = c(75, 125), lty = "solid") +
-#   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
-#   theme_bw(base_size = base_size) +
-#   labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
-#   # lims(y = c(0.4, 1.1)) +
-#   theme(
-#     panel.grid.major = element_blank(),
-#     panel.grid.minor = element_blank(),
-#   )
-
-
-
 # figure 2
 # colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
 # colour_scale <- palette(hcl.colors(4, "viridis"))
@@ -110,26 +85,6 @@ rmspe_eff <- combo_data %>%
   )
 
 
-
-# figure 2
-# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-# resptype_labs <- c(MMI_ZOOP_NLA6 = "Response: Zooplankton MMI", TOTALHG_RESULT = "Response: Mercury")
-# rmspe_eff <- combo_data %>%
-#   ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
-#   facet_grid(
-#     rows = vars(var),
[…]

### 2022-06-09 inst/scripts/figures/simulation.R

@@ -85,25 +85,6 @@ rmspe_eff <- combo_data %>%
     panel.grid.minor = element_blank(),
   )
 
-# figure 2
-# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-# resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
-# psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
-# rmspe_eff <- combo_data %>%
-#   ggplot(aes(x = n_factor, y = rel_efficiency, colour = approach)) +
-#   facet_grid(
-#     psill_ratio ~ resptype,
-#     labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
-#   ) +
-#   geom_jitter(width = 0.24, size = 2) +
-#   scale_colour_manual(values = colour_scale) +
-#   theme_bw(base_size = base_size) +
-#   labs(x = "Sample Size", colour = "Approach", y = "Relative rMS(P)E") +
-#   lims(y = c(0.4, 1.1)) +
-#   theme(
-#     panel.grid.major = element_blank(),
-#     panel.grid.minor = element_blank(),
-#   )
 
 ### rmspe_eff
 if (write_out) {
@@ -115,60 +96,6 @@ if (write_out) {
     height = 4.39
   )
 }
-
-
-# figure 3
-# colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
-# resptype_labs <- c(normal = "Response: Normal", lognormal = "Response: Lognormal")
-# psill_ratio_labs <- c("0" = "Prop DE: 0", "0.5" = "Prop DE: 0.5", "0.9" = "Prop DE: 0.9")
-# mse_eff <- combo_data %>%
-#   ggplot(aes(x = n_factor_adj, y = mse_efficiency, colour = approach)) +
-#   facet_grid(
-#     psill_ratio ~ resptype,
-#     labeller = labeller(resptype = resptype_labs, psill_ratio = psill_ratio_labs)
-#   ) +
-#   geom_point(size = 2) +
-#   geom_vline(xintercept = c(75, 125), lty = "dashed") +
-#   scale_x_continuous(breaks = c(50, 100, 150), labels = c("50", "100", "200")) +
-#   scale_colour_manual(values = colour_scale) +
[…]

### 2022-06-09 inst/scripts/figures/simulation_exploratory.R

@@ -63,7 +63,7 @@ if (write_out) {
 }
 
 
-# # population simulations
+# # population simulations for single observation
 #
 # ## normal
 # df_norm <- data.frame(x = seq(-5, 5, length.out = 1000))

### 2022-06-09 inst/manuscript/manuscript.Rmd
minor updates

@@
 bibliography: mybibfile.bib
-linenumbers: true
+linenumbers: false
 numbersections: true
@@
 
-```{r, echo = FALSE, results = "asis"}
-# tab <- matrix(c("SRS-DB", "SRS-MB", "GRTS-DB", "GRTS-MB"),
-#        nrow = 2, byrow = TRUE) %>%
-#   as.table()
-# row.names(tab) <- c("IRS", "GRTS")
-# xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-inference combinations in the simulation study.
-The rows give the two types of sampling designs and the columns give the two types of analyses.")
-# align(xtab1) <- "r|ll"
-# names(xtab1) <- c("Design", "Model")
-# print(xtab1, comment = FALSE)
-# ## I really want "Analysis" to appear as a row header and "Sampling Design" 
-# ## to appear as a column header, but this isn't easy to figure out.
-# ## I couldn't figure this out with xtable either so it might need to be
-# ## made "by hand."
-```
-
 We evaluated performance of the four sampling-inference combinations in 36 different simulation scenarios.
@@
 \end{figure}
-
-```{r, echo = FALSE, results = "asis"}
-# n <- c("50", "100", "200")
-# site_locations <- c("Random", "Gridded", "")
-# psill_ratio <- c("0", "0.5", "0.9")
-# resp_type <- c("Normal", "Lognormal", "")
-# parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-# parm_tab[c(2, 4), 3] <- "-"
-# row.names(parm_tab) <- c("Sample Size (n)", "Location Layout", 
-#                          "Proportion of Dependent Error", "Response Type")
-# # parm_tab <- parm_tab[c(1, 2, 4, 3), ]
-# parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
-All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was 2.")
-# align(parm_xtab) <- "r|lll"
-# print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
-```
 
@@
 
-```{r apptab, results = "asis"}
-# res_df <- readr::read_csv(here("inst", "output", "application", "application.csv"))
-# # res_df$bias <- res_df$realized_mean - res_df$estimate
-# # res_df_drop <- res_df[ , c("approach", "estimate", "bias", "se", "lb", "ub")]
-# res_df_drop <- res_df[, ]
-# names(res_df_drop) <- c("Approach", "True Mean", "Est/Pred", "SE", "95% LB", "95% UB")
-# xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} For each sampling-inference combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\\% interval bounds (95\\% LB), and upper 95\\% interval bounds (95\\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.")
-# print(xtab_app, include.rownames = FALSE, comment = FALSE)
-```
-
 # Discussion {#sec:discussion}
@@
 
-# Conflict of Interest Statement {.unnumbered}
-
-There are no conflicts of interest for any of the authors.
-
-# Author Contribution Statement {.unnumbered}
-
-All authors conceived the ideas; All authors designed the methodology; MD and MH performed the simulations and analyzed the data; MD and MH led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
-
 # Data and Code Availability {.unnumbered}
@@
 
+# Conflict of Interest Statement {.unnumbered}
+
+There are no conflicts of interest for any of the authors.
+
+# Author Contribution Statement {.unnumbered}
+
+All authors conceived the ideas; All authors designed the methodology; Michael Dumelle and Matt Higham performed the simulations and analyzed the data; Michael Dumelle and Matt Higham led the writing of the manuscript; All authors contributed critically to the drafts and gave final approval for publication.
+
 # Supporting Information {.unnumbered}

### 2022-06-09 inst/manuscript/manuscript.tex
minor updates

@@
 \usepackage{lineno} % add
-  \linenumbers % turns line numbering on
 \providecommand{\tightlist}{%
@@
 
+\hypertarget{data-and-code-availability}{%
+\section*{Data and Code Availability}\label{data-and-code-availability}}
+\addcontentsline{toc}{section}{Data and Code Availability}
+
+This manuscript has a supplementary \textbf{\textsf{R}} package that
+contains all of the data and code used in its creation. The
+supplementary \textbf{\textsf{R}} package is hosted on GitHub.
+Instructions for download are available at
+
+\url{https://github.com/michaeldumelle/DvMsp}.
+
+If the manuscript is accepted, this repository will be archived in
+Zenodo.
+
 \hypertarget{conflict-of-interest-statement}{%
@@
 All authors conceived the ideas; All authors designed the methodology;
-MD and MH performed the simulations and analyzed the data; MD and MH led
-the writing of the manuscript; All authors contributed critically to the
-drafts and gave final approval for publication.
-
-\hypertarget{data-and-code-availability}{%
-\section*{Data and Code Availability}\label{data-and-code-availability}}
-\addcontentsline{toc}{section}{Data and Code Availability}
-
-This manuscript has a supplementary \textbf{\textsf{R}} package that
-contains all of the data and code used in its creation. The
-supplementary \textbf{\textsf{R}} package is hosted on GitHub.
-Instructions for download are available at
-
-\url{https://github.com/michaeldumelle/DvMsp}.
-
-If the manuscript is accepted, this repository will be archived in
-Zenodo.
+Michael Dumelle and Matt Higham performed the simulations and analyzed
+the data; Michael Dumelle and Matt Higham led the writing of the
+manuscript; All authors contributed critically to the drafts and gave
+final approval for publication.
 

### 2022-06-09 inst/manuscript/supporting.Rmd
minor updates

@@
 For the simulated data, we considered 36 parameter configurations -- the crossing of three sample sizes ($n = 50$, $n = 100$, $n = 200)$, two location layouts (random and gridded), three proportions of dependent random error (DRE) (0%, 50%, 90%), and two response types (normal and skewed).
-
-```{r, echo = FALSE, results = "asis"}
-# tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
-#        nrow = 2, byrow = TRUE) %>%
-#   as.table()
-# row.names(tab) <- c("IRS", "GRTS")
-# xtab1 <- xtable(tab, col.names = c("Design", "Model"), caption = "\\label{tab:designanalysis} Sampling-analysis combinations in the simulation study.
-The rows give the two types of sampling designs and the columns give the two types of analyses.")
-# align(xtab1) <- "r|ll"
-# names(xtab1) <- c("Design", "Model")
-# print(xtab1, comment = FALSE)
-# ## I really want "Analysis" to appear as a row header and "Sampling Design" 
-# ## to appear as a column header, but this isn't easy to figure out.
-# ## I couldn't figure this out with xtable either so it might need to be
-# ## made "by hand."
-```
-
-```{r, echo = FALSE, results = "asis"}
-# n <- c("50", "100", "200")
-# site_locations <- c("Random", "Gridded", "")
-# psill_ratio <- c("0", "0.5", "0.9")
-# resp_type <- c("Normal", "Lognormal", "")
-# parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
-# parm_tab[c(2, 4), 3] <- "-"
-# row.names(parm_tab) <- c("Sample Size (n)", "Layout", 
-#                          "Proportion of Dependent Error", "Response Type")
-# # library(knitr) kable with htlm has trouble with page breaks -- kable with latex harder for me than xtable
-# # kable(parm_tab, col.names = c("", "", ""),
-# #       caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
-# parm_xtab <- xtable(parm_tab, caption = "\\label{tab:parmtab} Simulation scenario options.
-All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios.
-In each simualtion scenario, the total variance was 2.")
-# align(parm_xtab) <- "r|lll"
-# print(parm_xtab, include.colnames = FALSE, hline.after = c(0, nrow(parm_xtab)), comment = FALSE)
-```
 

### 2022-06-10 inst/manuscript/manuscript.Rmd
minor updates

@@
 After the GRTS algorithm was developed, several other spatially balanced sampling algorithms emerged, including stratified sampling with compact geographical strata [@walvoort2010r], the local pivotal method [@grafstrom2012spatially; @grafstrom2018spatially], spatially correlated Poisson sampling [@grafstrom2012spatiallypoisson], balanced acceptance sampling [@robertson2013bas], within-sample-distance sampling [@benedetti2017spatially], and Halton iterative partitioning sampling [@robertson2018halton].
-In this manuscript, we select spatially balanced samples using the GRTS algorithm because it is readily available in the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units.
+In this manuscript, we select spatially balanced samples using the GRTS algorithm because it is readily available in the spsurvey \textbf{\textsf{R}} package [@dumelle2022spsurvey] and naturally accommodates finite and infinite sampling frames, unequal inclusion probabilities, and replacement units.
 Replacement units are additional population units that can be sampled when a population unit originally selected can no longer be sampled.
@@
 For SRS-MB and GRTS-MB, FPBK was used to estimate means and variances using restricted maximum likelihood.
-SRS, GRTS sampling, and design-based inference were implemented using the `spsurvey` \textbf{\textsf{R}} package [@dumelle2022spsurvey].
-FPBK was implemented using the `sptotal` \textbf{\textsf{R}} package [@higham2021sptotal].
+SRS, GRTS sampling, and design-based inference were implemented using the spsurvey \textbf{\textsf{R}} package [@dumelle2022spsurvey].
+FPBK was implemented using the sptotal \textbf{\textsf{R}} package [@higham2021sptotal].
 
@@
 The skewed responses were used to evaluate performance of the sampling-inference approaches for data that were not normally distributed but were still estimated using REML, which relies on a normal log-likelihood.
-Figure \ref{fig:sim_pops} shows an example of a realized population for the normal and skewed responses using the random location layout and 50% DRE.
+Fig.
+\ref{fig:sim_pops} shows an example of a realized population for the normal and skewed responses using the random location layout and 50% DRE.
 
@@
 This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data and code used in its creation.
-The supplementary \textbf{\textsf{R}} package is hosted on GitHub.
-Instructions for download are available at
-
-[https://github.com/michaeldumelle/DvMsp](https://github.com/michaeldumelle/DvMsp).
-
-If the manuscript is accepted, this repository will be archived in Zenodo.
+The supplementary \textbf{\textsf{R}} package is hosted on GitHub and archived on Zenodo.
+The Zenodo DOI is located here and in the references.
 

### 2022-06-10 inst/manuscript/manuscript.tex
minor updates

@@
 contains all of the data and code used in its creation. The
-supplementary \textbf{\textsf{R}} package is hosted on GitHub.
-Instructions for download are available at
-
-\url{https://github.com/michaeldumelle/DvMsp}.
-
-If the manuscript is accepted, this repository will be archived in
-Zenodo.
+supplementary \textbf{\textsf{R}} package is hosted on GitHub and
+archived on Zenodo. The Zenodo DOI is located here and in the
+references.
 

### 2022-06-11 inst/manuscript/manuscript.Rmd
minor updates pre zenodo

@@
 date: "`r Sys.Date()`"
-bibliography: mybibfile.bib
+bibliography: references.bib
 linenumbers: false
@@
 For Hg ppb, data were collected at 995 unique lakes (there were no replicates).
-The ZMMI and Hg ppb data are shown as spatial maps and as histograms in Figure \ref{fig:zmmi}.
+The ZMMI and Hg ppb data are shown as spatial maps and as histograms in Fig.
+\ref{fig:zmmi}.
 The ZMMI data tend to be highest near the coasts, lowest in the Central United States, are relatively symmetric, and have a mean of 55.05.
 The Hg ppb data tend to be highest in the Northeastern United States, lowest elsewhere, are skewed, and have a mean of 103.16 ppb.
-Also in Figure \ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
+Also in Fig.
+\ref{fig:zmmi} are separate spatial semivariograms for ZMMI and Hg ppb.
 The spatial semivariogram quantifies the halved average squared differences (semivariance) of responses whose separation (distance) falls within a separation class.
 The spatial semivariance is closely related to the spatial covariance, and spatial semivariograms are often used to gauge the strength of spatial dependence in data.
-Both ZMMI and Hg ppb seem to have moderately strong spatial dependence (Figure \ref{fig:zmmi}), as the empirical semivariance increases steadily with distance (meaning that observations near one another tend to be more similar than observations far apart from one another).
+Both ZMMI and Hg ppb seem to have moderately strong spatial dependence (Fig.
+\ref{fig:zmmi}), as the empirical semivariance increases steadily with distance (meaning that observations near one another tend to be more similar than observations far apart from one another).
 
@@
 As with the simulated data, coverages for the NLA data tend to increase with the sample sizes, coverages tend to be higher for symmetric responses than for skewed responses, and the local neighborhood variance was slightly too small for small $n$, yielding slightly lower interval coverages than the other sampling-inference combinations.
+
 Recall that model-based inference defines interval coverage properties across realized populations.
@@
 
-This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data and code used in its creation.
+This manuscript has a supplementary \textbf{\textsf{R}} package that contains all of the data, code, and figures used in its creation.
+All figures were created using the ggplot2 \textbf{\textsf{R}} package [@wickham2016ggplot2].
 The supplementary \textbf{\textsf{R}} package is hosted on GitHub and archived on Zenodo.

### 2022-06-11 inst/manuscript/manuscript.tex
minor updates pre zenodo

@@
 spatially balanced samples using the GRTS algorithm because it is
-readily available in the \texttt{spsurvey} \textbf{\textsf{R}} package
-(Dumelle et al., 2022) and naturally accommodates finite and infinite
-sampling frames, unequal inclusion probabilities, and replacement units.
+readily available in the spsurvey \textbf{\textsf{R}} package (Dumelle
+et al., 2022) and naturally accommodates finite and infinite sampling
+frames, unequal inclusion probabilities, and replacement units.
 Replacement units are additional population units that can be sampled
@@
 means and variances using restricted maximum likelihood. SRS, GRTS
-sampling, and design-based inference were implemented using the
-\texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle et al., 2022).
-FPBK was implemented using the \texttt{sptotal} \textbf{\textsf{R}}
-package (Higham et al., 2021a).
+sampling, and design-based inference were implemented using the spsurvey
+\textbf{\textsf{R}} package (Dumelle et al., 2022). FPBK was implemented
+using the sptotal \textbf{\textsf{R}} package (Higham et al., 2021a).
 
@@
 for data that were not normally distributed but were still estimated
-using REML, which relies on a normal log-likelihood. Figure
+using REML, which relies on a normal log-likelihood. Fig.
 \ref{fig:sim_pops} shows an example of a realized population for the
@@
 replicates). The ZMMI and Hg ppb data are shown as spatial maps and as
-histograms in Figure \ref{fig:zmmi}. The ZMMI data tend to be highest
-near the coasts, lowest in the Central United States, are relatively
+histograms in Fig. \ref{fig:zmmi}. The ZMMI data tend to be highest near
+the coasts, lowest in the Central United States, are relatively
 symmetric, and have a mean of 55.05. The Hg ppb data tend to be highest
 in the Northeastern United States, lowest elsewhere, are skewed, and
-have a mean of 103.16 ppb. Also in Figure \ref{fig:zmmi} are separate
+have a mean of 103.16 ppb. Also in Fig. \ref{fig:zmmi} are separate
 spatial semivariograms for ZMMI and Hg ppb. The spatial semivariogram
@@
 spatial dependence in data. Both ZMMI and Hg ppb seem to have moderately
-strong spatial dependence (Figure \ref{fig:zmmi}), as the empirical
+strong spatial dependence (Fig. \ref{fig:zmmi}), as the empirical
 semivariance increases steadily with distance (meaning that observations
@@
 for small \(n\), yielding slightly lower interval coverages than the
-other sampling-inference combinations. Recall that model-based inference
-defines interval coverage properties across realized populations. With
-the simulated data, we evaluated interval coverage across realized
-populations, but for the NLA data, we evaluated interval coverage within
-a single realized population. We did find that model-based coverages
-were similar to the design-based coverages, however, suggesting that for
-some realized populations it is reasonable to heuristically view data
-from separate random samples as being from approximately separate
-realized populations. But generally, if model-based intervals
-constructed from many random samples of a single realized population
-show improper coverage, this does not necessarily imply a deficiency in
-model-based inference. Tables for interval coverage in all six
-simulation scenarios are provided in the supporting information.
+other sampling-inference combinations.
+
+Recall that model-based inference defines interval coverage properties
+across realized populations. With the simulated data, we evaluated
+interval coverage across realized populations, but for the NLA data, we
+evaluated interval coverage within a single realized population. We did
+find that model-based coverages were similar to the design-based
+coverages, however, suggesting that for some realized populations it is
+reasonable to heuristically view data from separate random samples as
+being from approximately separate realized populations. But generally,
+if model-based intervals constructed from many random samples of a
+single realized population show improper coverage, this does not
+necessarily imply a deficiency in model-based inference. Tables for
+interval coverage in all six simulation scenarios are provided in the
+supporting information.
 
@@
 This manuscript has a supplementary \textbf{\textsf{R}} package that
-contains all of the data and code used in its creation. The
-supplementary \textbf{\textsf{R}} package is hosted on GitHub and
-archived on Zenodo. The Zenodo DOI is located here and in the
-references.
+contains all of the data, code, and figures used in its creation. All
+figures were created using the ggplot2 \textbf{\textsf{R}} package
+(Wickham, 2016). The supplementary \textbf{\textsf{R}} package is hosted
+on GitHub and archived on Zenodo. The Zenodo DOI is located here and in
+the references.
 
@@
 
+\leavevmode\hypertarget{ref-wickham2016ggplot2}{}%
+Wickham, H., 2016. ggplot2: Elegant graphics for data analysis.
+Springer.
+
 \leavevmode\hypertarget{ref-wolfinger1994computing}{}%

### 2022-06-11 inst/manuscript/supporting.Rmd
minor updates pre zenodo

@@
                   n = n,
-                  MB = mpbias,
+                  MBias = mpbias,
                   RMSE = rmspe,
@@
   digits = 4,
-  caption = paste0("Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
+  caption = paste0("Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in simulation scenario ", table_value, "."),
    type = "latex",
@@
   digits = 4,
-  caption = paste0("Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in a data application scenario ", table_value, "."),
+  caption = paste0("Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\\% interval coverage (Coverage) in a data application scenario ", table_value, "."),
    type = "latex",

### 2022-06-11 inst/manuscript/supporting.tex
minor review updates
minor updates
minor updates pre zenodo

@@
   \linenumbers % turns line numbering on
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% Pandoc citation processing
+
+% Pandoc header
 
@@
 \begin{document}
-
 \begin{frontmatter}
 
-  \title{A comparison of design-based and model-based approaches for finite
-population spatial data -- Supporting information.}
+  \title{A comparison of design-based and model-based approaches for
+finite population spatial data -- Supporting information.}
     \author[USEPA]{Michael Dumelle\corref{1}}
@@
 
-      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
-Corvallis, Oregon, 97333}
-    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
-and Statistics, 23 Romoda Drive, Canton, New York, 13617}
-    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
-Corvallis, Oregon, 97331}
-    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
-Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
-      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}
+      \address[USEPA]{United States Environmental Protection Agency, 200
+SW 35th St, Corvallis, Oregon, 97333}
+    \address[STLAW]{St.~Lawrence University Department of Mathematics,
+Computer Science, and Statistics, 23 Romoda Drive, Canton, New York,
+13617}
+    \address[OSU]{Oregon State University Department of Statistics, 239
+Weniger Hall, Corvallis, Oregon, 97331}
+    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science
+Center, National Oceanic and Atmospheric Administration, Seattle,
+Washington, 98115}
+      \cortext[1]{Corresponding Author: Michael Dumelle
+(Dumelle.Michael@epa.gov)}
 
@@
   \hline
-Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
+Approach & Layout & Response & DRE\% & n & MBias & RMSE & Coverage \\ 
   \hline
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 1.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 1.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 2.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 2.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 3.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 3.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 4.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 4.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 5.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 5.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 6.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 6.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 7.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 7.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 8.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 8.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 9.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 9.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 10.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 10.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 11.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 11.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 12.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 12.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 13.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 13.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 14.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 14.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 15.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 15.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 16.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 16.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 17.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 17.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 18.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 18.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 19.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 19.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 20.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 20.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 21.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 21.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 22.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 22.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 23.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 23.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 24.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 24.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 25.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 25.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 26.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 26.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 27.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 27.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 28.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 28.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 29.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 29.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 30.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 30.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 31.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 31.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 32.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 32.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 33.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 33.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 34.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 34.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 35.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 35.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 36.}
+\caption{Sampling-inference combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in simulation scenario 36.}
 \end{table}
@@
   \hline
-Approach & Response & n & MB & RMSE & Coverage \\ 
+Approach & Response & n & MBias & RMSE & Coverage \\ 
   \hline
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 1.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 1.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 2.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 2.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 3.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 3.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 4.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 4.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 5.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 5.}
 \end{table}
@@
 }
-\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MB), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 6.}
+\caption{Sampling-inference combination (Approach), response type (Response), sample size (n), mean bias (MBias), root-mean-squared error (RMSE), and 95\% interval coverage (Coverage) in a data application scenario 6.}
 \end{table}

### 2022-06-11 inst/manuscript/manuscript.tex
added zenodo reference

@@
 readily available in the spsurvey \textbf{\textsf{R}} package (Dumelle
-et al., 2022) and naturally accommodates finite and infinite sampling
+et al., 2022b) and naturally accommodates finite and infinite sampling
 frames, unequal inclusion probabilities, and replacement units.
@@
 sampling, and design-based inference were implemented using the spsurvey
-\textbf{\textsf{R}} package (Dumelle et al., 2022). FPBK was implemented
-using the sptotal \textbf{\textsf{R}} package (Higham et al., 2021a).
+\textbf{\textsf{R}} package (Dumelle et al., 2022b). FPBK was
+implemented using the sptotal \textbf{\textsf{R}} package (Higham et
+al., 2021a).
 
@@
 (Wickham, 2016). The supplementary \textbf{\textsf{R}} package is hosted
-on GitHub and archived on Zenodo. The Zenodo DOI is located here and in
-the references.
+on GitHub and archived on Zenodo. The Zenodo DOI link is
+https://doi.org/10.5281/zenodo.6633805 (Dumelle et al., 2022a).
 
@@
 
+\leavevmode\hypertarget{ref-dumelle2022DvMsp}{}%
+Dumelle, M., Higham, M., Hoef, J.M.V., Olsen, A.R., Madsen, L., 2022a.
+USEPA/DvMsp: Publication release. Zenodo.
+doi:\href{https://doi.org/10.5281/zenodo.6633805}{10.5281/zenodo.6633805}
+
 \leavevmode\hypertarget{ref-dumelle2022spsurvey}{}%
-Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2022. Spsurvey:
+Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2022b. Spsurvey:
 Spatial sampling design and analysis.

### 2022-06-11 inst/manuscript/manuscript.Rmd
added zenodo reference
minor updates

@@
 Then $\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}$.
-Figure$~$\ref{fig:grts} provides some intuition regarding the assignment of level-one and level-two cells.
+Fig.$~$\ref{fig:grts} provides some intuition regarding the assignment of level-one and level-two cells.
 
@@
 \end{equation}
-where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, $\bm{\beta}$ is the parameter vector of fixed effects, and $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
+where $\mathbf{X}_s$ and $\mathbf{X}_u$ are the design matrices for the sampled and unsampled population units, respectively, $\bm{\beta}$ is the parameter vector of fixed effects, and $\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]^T$, where $\bm{\delta}_s$ and $\bm{\delta}_u$ are random errors for the sampled and unsampled population units, respectively.
 
@@
 \end{subfigure}
-\caption{Histograms of single realized populations simulated for the normal and skewed respones using the random layout and 50\% DRE.}
+\caption{Histograms of single realized populations simulated for the normal and skewed responses using the random layout and 50\% DRE.}
 \label{fig:sim_pops}
@@
 The supplementary \textbf{\textsf{R}} package is hosted on GitHub and archived on Zenodo.
-The Zenodo DOI is located here and in the references.
+The Zenodo DOI link is https://doi.org/10.5281/zenodo.6633805 [@dumelle2022DvMsp].
 

### 2022-06-11 inst/manuscript/manuscript.tex
minor updates

@@
 \(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).
-Figure\(~\)\ref{fig:grts} provides some intuition regarding the
-assignment of level-one and level-two cells.
+Fig.\(~\)\ref{fig:grts} provides some intuition regarding the assignment
+of level-one and level-two cells.
 
@@
 respectively, \(\bm{\beta}\) is the parameter vector of fixed effects,
-and \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
+and \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]^T\), where
 \(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the

### 2022-06-11 inst/manuscript/supporting.Rmd
minor updates

@@
 date: "`r Sys.Date()`"
-linenumbers: true
+linenumbers: false
 numbersections: true
